---
title: "HSA clinical records coded exploration"
format: html
editor: visual
---

```{r}
library(tidyverse)
folder_path <- here::here("Data")

```

Checking number of HSA dogs coded by Josie etc and how many had histopath vs not

```{r}
HSA_manually_coded<-read.csv(here::here("Data/HSA_cases_clinical_records_coded.csv"))

analysed_manuscript_dogs <-read.csv(here::here("Output/HSA_cohort_all_RFs_data.csv"))
#keep only HSA cases in manuscript analyses
HSA_analysed_manuscript_dogs <-analysed_manuscript_dogs %>%
  dplyr::filter(had_HSA==1)
#keep only HSA manually coded cases that have the records read and coded
HSA_manually_coded_coded <- HSA_manually_coded %>%
  dplyr:: filter(Male.Female != "")
#there is 318 / 490  ~61% coded

#keep only 
```

Check how many of these match up to the ones used in paper ie. so which ones have been recorded so that we can confirm there was an "intent to treat"

```{r}
#rename PatientID to subject_id
HSA_manually_coded_coded <- HSA_manually_coded_coded %>%
  rename(subject_id =PatientID)

unique_rows <- unique(HSA_manually_coded_coded$subject_id)
HSA_in_paper_and_manually_coded <- HSA_analysed_manuscript_dogs %>% left_join(HSA_manually_coded_coded,by="subject_id")

#there aresomehow ~30 more rows made , find duplicates and determine why
duplicates <-HSA_in_paper_and_manually_coded %>%
group_by(subject_id) %>%
  filter(n() > 1) %>%
  arrange(subject_id)

#are these dupes from RHS dataset? or LHS?
original_df_dupes <-HSA_manually_coded_coded %>%
    group_by(subject_id) %>%
  filter(n() > 1) %>%
  arrange(subject_id)
```

There are dupes in the manual coding df- why?

write to csv to read these better

```{r}
write.csv(original_df_dupes,here::here("Output/HSA_manual_coding_dupes.csv"))
```

There is 54 duplicates = 26 dogs duplicated

Josie and student did lots of duplicates, Josie has QCed using some of the duplicates and colour coded in the original google sheet (duplicates all yellow colour in PatientID column, white row is use this row of the two not the green one

<https://docs.google.com/spreadsheets/d/1WyjMUCg7ToER7m2Ptt5wr1aEPtgdjvkc6148D5br5mo/edit?gid=0#gid=0>

There are about 6 dogs that have duplicates that have not been QCed:

020679,025385,026264, 027803,028584,034118

Since CSV doesnt carry this colour info over make a new column in it (QC_duplicate_use YN for dupes else NA)

Check how many records of the tier 1 patients are already coded up manually

```{r}
tiered_HSA <- read.csv(here::here("Data/study_endpoints.csv"))

tiered_HSA<- tiered_HSA %>%
  mutate(across(where(is.character), str_to_lower))


```

Tier 1 HSA dogs

```{r}
HSA_tier1 <- tiered_HSA %>%
  dplyr::filter(tier_of_confidence==1)%>%
  dplyr::filter(str_detect(tracked_condition, "hemangios"))
```

How many of these 371 are already coded:

```{r}
HSA_manually_coded_coded_unique_only <- HSA_manually_coded_coded %>%
  
already_coded <- HSA_man
```
