---
title: "GRLS clinic postcode deprivation linkup"
format: html
editor: visual
---

## GRLS clinic postcode to deprivation linkup

Further info on MDI = https://www.census.gov/topics/income-poverty/poverty/about/related-sites/rates.html

US equiv of UK IMD

Using vet clinic zip codes/postal code to get a FIPS code then link this to county level MDI

```{r}
library(tidyverse)
#GRLS clinics
clinics <- read.csv("C:/Users/ctaylor18/OneDrive - Royal Veterinary College/post doc/GRLS/data updates June 2024/vet_address.csv")

#strip to only have first 5 digits of postcode as some have a second part for full zipcode
clinic <- clinics %>%
  mutate(postal_code = substr(postal_code, 1, 5))


clinics$postal_code <- as.numeric(clinics$postal_code)
#FIPS database from https://simplemaps.com/data/us-zips
FIPS <- read.csv("C:/Users/ctaylor18/OneDrive - Royal Veterinary College/post doc/GRLS/data updates June 2024/simplemaps_uszips_basicv1.86/uszips.csv")
FIPS <- FIPS %>%
  rename(postal_code = zip)

#join clinics up to the FIPS data
clinics_and_fips <- clinics %>% left_join(FIPS, by = "postal_code")

#delete redundant columns
clinics_and_fips2 <- clinics_and_fips %>%
  dplyr::select(-c(address_line2))

```

MDI data from https://www.census.gov/topics/income-poverty/poverty/about/related-sites/rates.html - just going to use most recent values available of 2019 MDI

```{r}
MDI_2019 <- read.csv("C:/Users/ctaylor18/OneDrive - Royal Veterinary College/post doc/GRLS/data updates June 2024/MDI_2019.csv")
MDI_2019 <- MDI_2019 %>%
  rename(county_fips=County)



```

Need to calculate quintiles for MDI 2019 rate across all counties and then apply this quintile value to our clinics

```{r}
MDI_quintiles <- quantile(MDI_2019$MDI.rate, probs = seq(0, 1, 0.2))


MDI_2019$MDI_quintile <- cut(MDI_2019$MDI.rate, breaks = MDI_quintiles, include.lowest = TRUE, labels = 1:5)

```

Join MDI data to clinics

```{r}
clinics_and_fips2_mdi <- clinics_and_fips2 %>% left_join (MDI_2019,by="county_fips" )

write.csv(clinics_and_fips2_mdi,"C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/GRLS_dogs_MDI_2019.csv")
```
