---
title: "Cox HSA time to diagnosis"
format: html
editor: visual
---

# Time to HSA diagnosis in GRLS cohort

-   Examining all 3044 dogs in cohort through Cox PH model to determine time to HSA diagnosis

-   Endpoint - date of HSA diagnosis or censor date December 1 2024

-   Censor status - 1= had HSA diagnosed before Dec 1 2024 and 0= no HSA diagnosis by Dec 1 2024

## Import data

```{r, echo=FALSE}
library(tidyverse)
library(dplyr)
library(survival)
library(ggplot2)
library(tibble)
library(kableExtra)
#devtools::install_github("zabore/ezfun")
ezfun::set_ccf_palette("contrast")
library(lubridate)
library(ggsurvfit)
library(gtsummary)
library(tidycmprsk)
library(gtable)
#devtools::install_github("zabore/condsurv")
library(condsurv)
library(survminer)
library(cardx)
library(finalfit)

#load in dataset
cohort <- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_cohort_all_RFs_data.csv")
colnames(cohort)
cohort <- cohort %>%
  dplyr::select(-c(X))
cohort <- cohort %>%
  rename(had_HSA = had_HSA.x.x)

HSA <- cohort %>%
  dplyr::filter(had_HSA==1)
non_HSA <- cohort %>%
   dplyr::filter(had_HSA==0)


```

## Descriptives of cohort

Age at endpoint

```{r}
cohort$age_at_final_date <- as.numeric(cohort$age_at_final_date)

summary(cohort$age_at_final_date)

summary(HSA$age_at_final_date)
summary(non_HSA$age_at_final_date)
```

Also split age into quartiles

```{r}
age_summary <- cohort %>%
  summarise(
      Mean  = mean(age_at_final_date),
      SD = sd(age_at_final_date),
      Median = median(age_at_final_date),
      LQ = quantile(age_at_final_date, 0.25),
      UQ = quantile(age_at_final_date, 0.75),
      Min = min(age_at_final_date),
      Max = max(age_at_final_date)
    )
age_summary
cohort <- cohort %>%
  mutate(
    age_at_final_date_quartiles = case_when(
      (age_at_final_date < age_summary$LQ) ~ "0.74-<8.47 years",
      (age_at_final_date >= age_summary$LQ & age_at_final_date < age_summary$Median) ~ ">=8.47-<10.68 years",
      (age_at_final_date >= age_summary$Median & age_at_final_date < age_summary$UQ) ~ ">=10.68-<11.64 years",
      (age_at_final_date >= age_summary$UQ) ~ ">=11.64-14.51 years",
    )
  )
```

```{r}
cross_table <- table(cohort$had_HSA,cohort$age_at_final_date_quartiles)
cross_table_percent <- prop.table(cross_table) * 100
print(cross_table)
print(cross_table_percent)

```

Neuter status endpoint

```{r}
cross_table <- table(cohort$had_HSA,cohort$neuter)
cross_table_percent <- prop.table(cross_table) * 100
print(cross_table)
print(cross_table_percent)

```

Sex

```{r}
cross_table <- table(cohort$had_HSA,cohort$sex_status)
# Get column-wise percentages
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

```{r}
cross_table <- table(cohort$had_HSA,cohort$sex)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

```{r}
#tidy up redundant cols - not wanted for univar and those that are essentially duplicats
cols_to_keep <- c("subject_id", "MDI_quintile",  "owner_MDI_quintile",
  "area_type_mode_5yrs_prior", "house_type_mode_5yrs_prior",
  "water_source_mode_5yrs_prior", "pipes_metal_any_mode_5yrs_prior",
  "pipes_plastic_any_mode_5yrs_prior", "heating_fuel_primary_mode_5yrs_prior",
  "cooking_fuel_primary_mode_5yrs_prior", "heating_fuel_secondary_mode_5yrs_prior",
  "cooking_fuel_secondary_mode_5yrs_prior", "year_in_study_diagnosis_or_final_record_year.y",
  "region_name_mode_5yrs_prior", "state_mode_5yrs_prior", 
  "final_date.y", 
  "chronic_inflam_present", "study_year_first_chronic_inflam",
  "first_chronic_inflam_condition", "study_year_latest_chronic_inflam",
  "num_lifetime_chronic_inflam", "all_comorbidities_ever", "diagnosis_after_comorb",
  "time_between_first_chronic_inflam_endpoint_diagnosis", "comorb_within_5_years_diagnosis_TF",
  "years_between_diagnosis_and_last_chronic_inflam", "number_comorbs_present",
  "comorbidities", "intestinal_parasitism", "tick_borne_parasitism",
  "other_parasitism", "all_parasitism", "infectious", "orthopaedic",
  "immune_mediated", "cardiovascular", "gastrointestinal",
  "inflammatory_other", "chronic_inflammatory", "lifelong_comorb_present.x",
  "lifelong_comorbidities", "lifelong_orthopaedic", "lifelong_immune_mediated",
  "lifelong_cardiovascular", "lifelong_gastrointestinal",
  "lifelong_inflammatory_other", "lifelong_chronic_inflammatory",
  "number_lifelong_comorbs_present", "had_HSA",
  "mean_activity_level_overview", "USDA_30min_sy1_2", "activity_1h_min_sy1_2",
  "activity_KC_2h_min_sy1_2", "avg_activity_duration_sy1_2",
  "avg_activity_intensity_sy1_2", "avg_activity_freq_sy1_2",
  "avg_activity_freq_sy3", "avg_activity_intensity_sy3",
  "avg_intensity_5y_prior_endpoint", "avg_frequency_5y_prior_endpoint",
  "use_aerosol_study_years_early_life", "use_air_cleaner_study_years_early_life",
  "use_hepa_filter_study_years_early_life", "use_moth_balls_study_years_early_life",
  "use_incense_or_candles_study_years_early_life", "smoke_exposure_study_years_early_life",
  "any_treated_weeds_study_years_early_life", "any_treated_insects_study_years_early_life",
  "any_treated_fertilizer_study_years_early_life", "use_aerosol_sum_early_life",
  "use_air_cleaner_sum_early_life", "use_hepa_filter_sum_early_life",
  "use_moth_balls_sum_early_life", "use_incense_or_candles_sum_early_life",
  "smoke_exposure_sum_early_life", "any_treated_weeds_sum_early_life",
  "any_treated_insects_sum_early_life", "any_treated_fertilizer_sum_early_life",
  "use_aerosol_study_years_rest_of_life", "use_air_cleaner_study_years_rest_of_life",
  "use_hepa_filter_study_years_rest_of_life", "use_moth_balls_study_years_rest_of_life",
  "use_incense_or_candles_study_years_rest_of_life", "smoke_exposure_study_years_rest_of_life",
  "any_treated_weeds_study_years_rest_of_life", "any_treated_insects_study_years_rest_of_life",
  "any_treated_fertilizer_study_years_rest_of_life", "use_aerosol_sum_rest_of_life",
  "use_air_cleaner_sum_rest_of_life", "use_hepa_filter_sum_rest_of_life",
  "use_moth_balls_sum_rest_of_life", "use_incense_or_candles_sum_rest_of_life",
  "smoke_exposure_sum_rest_of_life", "any_treated_weeds_sum_rest_of_life",
  "any_treated_insects_sum_rest_of_life", "any_treated_fertilizer_sum_rest_of_life",
  "use_aerosol_study_years_whole_life", "use_air_cleaner_study_years_whole_life",
  "use_hepa_filter_study_years_whole_life", "use_moth_balls_study_years_whole_life",
  "use_incense_or_candles_study_years_whole_life", "smoke_exposure_study_years_whole_life",
  "any_treated_weeds_study_years_whole_life", "any_treated_insects_study_years_whole_life",
  "any_treated_fertilizer_study_years_whole_life", "use_aerosol_sum_whole_life",
  "use_air_cleaner_sum_whole_life", "use_hepa_filter_sum_whole_life",
  "use_moth_balls_sum_whole_life", "use_incense_or_candles_sum_whole_life",
  "smoke_exposure_sum_whole_life", "any_treated_weeds_sum_whole_life",
  "any_treated_insects_sum_whole_life", "any_treated_fertilizer_sum_whole_life",
  "hours_of_smoke_early_life_total_dosage", "hours_of_smoke_rest_of_life_total_dosage",
  "hours_of_smoke_whole_life_total_dosage", "in.the.garage_YN_study_years_early_life",
  "in.the.house_YN_study_years_early_life", "outside_YN_study_years_early_life",
  "in.the.garage_YN_sum_early_life", "in.the.house_YN_sum_early_life",
  "outside_YN_sum_early_life", "in.the.garage_YN_study_years_rest_of_life",
  "in.the.house_YN_study_years_rest_of_life", "outside_YN_study_years_rest_of_life",
  "in.the.garage_YN_sum_rest_of_life", "in.the.house_YN_sum_rest_of_life",
  "outside_YN_sum_rest_of_life", "in.the.garage_YN_study_years_whole_life",
  "in.the.house_YN_study_years_whole_life", "outside_YN_study_years_whole_life",
  "in.the.garage_YN_sum_whole_life", "in.the.house_YN_sum_whole_life",
  "outside_YN_sum_whole_life", "majority_sleep_location_lifetime",
  "lifestyle", "main_lifestyle", "main_lifestyle_category",
  "year_in_study_diagnosis_or_final_record_year",
  "main_lifestyle_category_mode_5yrs_prior", "sleep_location_mode_5yrs_prior",  "use_aerosol_exposed_5yrs_prior", "use_air_cleaner_exposed_5yrs_prior",
  "use_hepa_filter_exposed_5yrs_prior", "use_moth_balls_exposed_5yrs_prior",
  "use_incense_or_candles_exposed_5yrs_prior", "smoke_exposure_exposed_5yrs_prior",
  "any_treated_weeds_exposed_5yrs_prior", "any_treated_insects_exposed_5yrs_prior",
  "any_treated_fertilizer_exposed_5yrs_prior", "hours_of_smoke_exposure_hours_5yrs_prior",
  "sex", "neuter", "mismating_ever", "neutered_on_date", "neutered_reason",
  "neutered_ever_bred", "spayed_while_in_heat", "spayed_while_pregnant",
  "prevent_heat_ever", "no_pregnancy_ever", "intact_sired_litters_natural_ever",
  "neutered_females_bred_with_count_ever", "neutered_sired_litters_natural_ever",
  "age_at_neuter_months", "age_at_neuter_years", "heats_before_neutering",
  "preg_pre_neuter", "mated_pre_neuter", "num_bred_before_neutering",
  "num_breeding_before_neutering", "ever_bred_from", "birth_date", "sex_status",
  "date_of_birth", "final_date2", "age_at_final_date", "avg_weight_5y_prior_endpoint",
  "avg_height_5y_prior_endpoint", "avg_purina_BCS_5y_prior_endpoint",
  "mode_purina_BCS_5y_prior_endpoint", "half_anti_inflam_lifetime",
  "half_anti_microbials_lifetime", "half_dietary_supplements_lifetime",
  "half_enteric_lifetime", "half_fluid_metabolites_lifetime",
  "half_hormones_and_related_lifetime", "half_immunologicals_lifetime",
  "half_misc_lifetime", "half_neurological_lifetime", "half_anti_parasites_lifetime",
  "half_anti_septics_lifetime", "half_cardio_respiratory_lifetime",
  "half_diuretics_lifetime", "half_anti_hist_lifetime", "half_anti_viral_lifetime",
  "quartile_anti_inflam_lifetime", "quartile_anti_microbials_lifetime",
  "quartile_dietary_supplements_lifetime", "quartile_enteric_lifetime",
  "quartile_fluid_metabolites_lifetime", "quartile_hormones_and_related_lifetime",
  "quartile_immunologicals_lifetime", "quartile_misc_lifetime",
  "quartile_neurological_lifetime", "quartile_anti_parasites_lifetime",
  "quartile_anti_septics_lifetime", "quartile_cardio_respiratory_lifetime",
  "quartile_diuretics_lifetime", "quartile_anti_hist_lifetime",
  "quartile_anti_viral_lifetime", "half_anti_inflam_5y_prior_endpoint",
  "half_anti_microbials_5y_prior_endpoint", "half_enteric_5y_prior_endpoint",
  "half_immunologicals_5y_prior_endpoint", "half_misc_5y_prior_endpoint",
  "half_neurological_5y_prior_endpoint", "half_anti_parasites_5y_prior_endpoint",
  "half_dietary_supplements_5y_prior_endpoint", "half_anti_septics_5y_prior_endpoint",
  "half_hormones_and_related_5y_prior_endpoint", "half_cardio_respiratory_5y_prior_endpoint",
  "half_fluid_metabolites_5y_prior_endpoint", "half_diuretics_5y_prior_endpoint",
  "half_anti_hist_5y_prior_endpoint", "half_anti_viral_5y_prior_endpoint",
  "quartile_anti_inflam_5y_prior_endpoint", "quartile_anti_microbials_5y_prior_endpoint",
  "quartile_enteric_5y_prior_endpoint", "quartile_immunologicals_5y_prior_endpoint",
  "quartile_misc_5y_prior_endpoint", "quartile_neurological_5y_prior_endpoint",
  "quartile_anti_parasites_5y_prior_endpoint", "quartile_dietary_supplements_5y_prior_endpoint",
  "quartile_anti_septics_5y_prior_endpoint", "quartile_hormones_and_related_5y_prior_endpoint",
  "quartile_cardio_respiratory_5y_prior_endpoint", "quartile_fluid_metabolites_5y_prior_endpoint",
  "quartile_diuretics_5y_prior_endpoint", "quartile_anti_hist_5y_prior_endpoint",
  "quartile_anti_viral_5y_prior_endpoint", "age_at_final_date_quartiles"
)

cohort_abr <- cohort %>%
  dplyr::select(all_of(cols_to_keep))

#also need to add in a survival time variable (days/weeks/months/years)
enrol_date <-read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Data/dog_profile.csv")
enrol_date <- enrol_date %>%
  dplyr::select(c(subject_id, enrolled_date))

cohort_abr<- cohort_abr%>%
  left_join(enrol_date,by="subject_id")
cohort_abr <- cohort_abr %>%
  mutate(final_date =as.Date(final_date.y,format="%Y-%m-%d"),
         enrolled_date = as.Date(enrolled_date,format="%Y-%m-%d") )
  
  
#survival time days
#survival time weeks
#survival time months
#survival time years
cohort_abr <- cohort_abr %>%
  mutate(
    survival_time_days2 = round(as.numeric(difftime(final_date, enrolled_date, units = "days")), 2),
    survival_time_weeks = round(as.numeric(difftime(final_date, enrolled_date, units = "weeks")), 2),
    survival_time_months = round(time_length(interval(enrolled_date, final_date), "months"), 2),
    survival_time_year = round(time_length(interval(enrolled_date, final_date), "years"), 2)
  )
summary(cohort_abr$survival_time_days2)
summary(cohort_abr$survival_time_weeks)
summary(cohort_abr$survival_time_months)
summary(cohort_abr$survival_time_year)


# set HSA status as numeric for KM curves
cohort_abr$had_HSA <-as.numeric(cohort_abr$had_HSA)
```

```{r}
cohort_abr<- cohort_abr %>%
  mutate(avg_weight_5y_prior_endpoint_kg = avg_weight_5y_prior_endpoint * 0.453)


#how does data look now
cohort_abr %>%
    summarize(
    mean_weight = mean(avg_weight_5y_prior_endpoint_kg, na.rm = TRUE),
    median_weight = median(avg_weight_5y_prior_endpoint_kg, na.rm = TRUE),
    sd_weight = sd(avg_weight_5y_prior_endpoint_kg, na.rm = TRUE),
    min_weight = min(avg_weight_5y_prior_endpoint_kg, na.rm = TRUE),
    max_weight= max(avg_weight_5y_prior_endpoint_kg, na.rm = TRUE),
    n = n(),
    n_missing = sum(is.na(avg_weight_5y_prior_endpoint_kg))
  )
#remove outliers >1.5X above IQR

Q1 <- quantile(cohort_abr$avg_weight_5y_prior_endpoint_kg, 0.25, na.rm = TRUE)
Q3 <- quantile(cohort_abr$avg_weight_5y_prior_endpoint_kg, 0.75, na.rm = TRUE)
IQR <- Q3 - Q1

cohort_abr <- cohort_abr %>%
  mutate(avg_weight_5y_prior_endpoint_kg_outliers_removed= ifelse(avg_weight_5y_prior_endpoint_kg < (Q1 - 1.5 * IQR) | avg_weight_5y_prior_endpoint_kg > (Q3 + 1.5 * IQR), NA, avg_weight_5y_prior_endpoint_kg))


#height
cohort_abr<- cohort_abr%>%
  mutate(avg_height_5y_prior_endpoint_cm = avg_height_5y_prior_endpoint * 2.54)


#how does data look now
cohort_abr %>%
    summarize(
    mean_height = mean(avg_height_5y_prior_endpoint_cm, na.rm = TRUE),
    median_height = median(avg_height_5y_prior_endpoint_cm, na.rm = TRUE),
    sd_height = sd(avg_height_5y_prior_endpoint_cm, na.rm = TRUE),
    min_height = min(avg_height_5y_prior_endpoint_cm, na.rm = TRUE),
    max_height= max(avg_height_5y_prior_endpoint_cm, na.rm = TRUE),
    n = n(),
    n_missing = sum(is.na(avg_height_5y_prior_endpoint_cm))
  )
#remove outliers >1.5X above IQR

Q1 <- quantile(cohort_abr$avg_height_5y_prior_endpoint_cm, 0.25, na.rm = TRUE)
Q3 <- quantile(cohort_abr$avg_height_5y_prior_endpoint_cm, 0.75, na.rm = TRUE)
IQR <- Q3 - Q1

cohort_abr <- cohort_abr %>%
  mutate(avg_height_5y_prior_endpoint_cm_outliers_removed= ifelse(avg_height_5y_prior_endpoint_cm < (Q1 - 1.5 * IQR) | avg_height_5y_prior_endpoint_cm > (Q3 + 1.5 * IQR), NA, avg_height_5y_prior_endpoint_cm))
```

I think months maybe is best option of the 4 (days,weeks, months, years)

## More final bits of data tidying

```{r}
#purina BCS average
cohort_abr$avg_purina_BCS_5y_prior_endpoint

bcs_summary <- cohort_abr %>%
  summarise(
      Mean  = mean(avg_purina_BCS_5y_prior_endpoint,na.rm=T),
      SD = sd(avg_purina_BCS_5y_prior_endpoint,na.rm=T),
      Median = median(avg_purina_BCS_5y_prior_endpoint,na.rm=T),
      LQ = quantile(avg_purina_BCS_5y_prior_endpoint, 0.25,na.rm=T),
      UQ = quantile(avg_purina_BCS_5y_prior_endpoint, 0.75,na.rm=T),
      Min = min(avg_purina_BCS_5y_prior_endpoint,na.rm=T),
      Max = max(avg_purina_BCS_5y_prior_endpoint,na.rm=T)
    )
bcs_summary
cohort_abr <- cohort_abr %>%
  mutate(
    avg_purina_BCS_5y_prior_endpoint_quartiles = case_when(
      (avg_purina_BCS_5y_prior_endpoint < bcs_summary$LQ) ~ "3.3-<5",
      (avg_purina_BCS_5y_prior_endpoint >= bcs_summary$LQ & avg_purina_BCS_5y_prior_endpoint < bcs_summary$Median) ~ ">=5-<5.28",
      (avg_purina_BCS_5y_prior_endpoint >= bcs_summary$Median & avg_purina_BCS_5y_prior_endpoint < bcs_summary$UQ) ~ ">=5.28-<5.86",
      (avg_purina_BCS_5y_prior_endpoint>= bcs_summary$UQ) ~ ">=5.86 - 9",
    )
  )
#weight quartiles
weight_summary <- cohort_abr %>%
  summarise(
      Mean  = mean(avg_weight_5y_prior_endpoint_kg_outliers_removed,na.rm=T),
      SD = sd(avg_weight_5y_prior_endpoint_kg_outliers_removed,na.rm=T),
      Median = median(avg_weight_5y_prior_endpoint_kg_outliers_removed,na.rm=T),
      LQ = quantile(avg_weight_5y_prior_endpoint_kg_outliers_removed, 0.25,na.rm=T),
      UQ = quantile(avg_weight_5y_prior_endpoint_kg_outliers_removed, 0.75,na.rm=T),
      Min = min(avg_weight_5y_prior_endpoint_kg_outliers_removed,na.rm=T),
      Max = max(avg_weight_5y_prior_endpoint_kg_outliers_removed,na.rm=T)
    )
weight_summary
cohort_abr <- cohort_abr %>%
  mutate(
    avg_weight_5y_prior_endpoint_quartiles = case_when(
      (avg_weight_5y_prior_endpoint_kg_outliers_removed < weight_summary$LQ) ~ "9.75-<21.31kg",
      (avg_weight_5y_prior_endpoint_kg_outliers_removed >= weight_summary$LQ & avg_weight_5y_prior_endpoint_kg_outliers_removed < weight_summary$Median) ~ ">=21.31-<24.66kg",
      (avg_weight_5y_prior_endpoint_kg_outliers_removed >= weight_summary$Median & avg_weight_5y_prior_endpoint_kg_outliers_removed < weight_summary$UQ) ~ ">=24.66-<29.37kg",
      (avg_weight_5y_prior_endpoint_kg_outliers_removed>= weight_summary$UQ) ~ ">=29.37-41.52kg",
    )
  )
#hieght quartiles
height_summary <- cohort_abr %>%
  summarise(
      Mean  = mean(avg_height_5y_prior_endpoint_cm_outliers_removed,na.rm=T),
      SD = sd(avg_height_5y_prior_endpoint_cm_outliers_removed,na.rm=T),
      Median = median(avg_height_5y_prior_endpoint_cm_outliers_removed,na.rm=T),
      LQ = quantile(avg_height_5y_prior_endpoint_cm_outliers_removed, 0.25,na.rm=T),
      UQ = quantile(avg_height_5y_prior_endpoint_cm_outliers_removed, 0.75,na.rm=T),
      Min = min(avg_height_5y_prior_endpoint_cm_outliers_removed,na.rm=T),
      Max = max(avg_height_5y_prior_endpoint_cm_outliers_removed,na.rm=T)
    )
height_summary
cohort_abr <- cohort_abr %>%
  mutate(
    avg_height_5y_prior_endpoint_quartiles = case_when(
      (avg_height_5y_prior_endpoint_cm_outliers_removed< height_summary$LQ) ~ "49.13-<56.07cm",
      (avg_height_5y_prior_endpoint_cm_outliers_removed >= height_summary$LQ & avg_height_5y_prior_endpoint_cm_outliers_removed < height_summary$Median) ~ ">=56.07-<58.42cm",
      (avg_height_5y_prior_endpoint_cm_outliers_removed >= height_summary$Median & avg_height_5y_prior_endpoint_cm_outliers_removed< height_summary$UQ) ~ ">=58.42-<60.32cm",
      (avg_height_5y_prior_endpoint_cm_outliers_removed>= height_summary$UQ) ~ ">=60.32cm-67.73cm",
    )
  )
#age at neuter quartiles
neut_summary <- cohort_abr %>%
  summarise(
      Mean  = mean(age_at_neuter_years,na.rm=T),
      SD = sd(age_at_neuter_years,na.rm=T),
      Median = median(age_at_neuter_years,na.rm=T),
      LQ = quantile(age_at_neuter_years, 0.25,na.rm=T),
      UQ = quantile(age_at_neuter_years, 0.75,na.rm=T),
      Min = min(age_at_neuter_years,na.rm=T),
      Max = max(age_at_neuter_years,na.rm=T)
    )
neut_summary

cohort_abr <- cohort_abr %>%
  mutate(
    age_at_neuter_years_quartiles = case_when(
      #so that entire dogs dont end up coded as Not_recorded
      neuter == "Entire" ~ "Entire",
      (age_at_neuter_years< neut_summary$LQ) ~ "0.15-<0.52y",
      (age_at_neuter_years >= neut_summary$LQ & age_at_neuter_years < neut_summary$Median) ~ ">=0.52-<0.83y",
      (age_at_neuter_years >= neut_summary$Median & age_at_neuter_years < neut_summary$UQ) ~ ">=0.83-2.44y",
      (age_at_neuter_years>= neut_summary$UQ) ~ ">=2.44-9.59y",
    )
  )

#hours of smoke - not worth quartiling as so many are 0
smoke_5y_summary <- cohort_abr %>%
  filter(hours_of_smoke_exposure_hours_5yrs_prior > 0) %>%
  summarise(
      Mean  = mean(hours_of_smoke_exposure_hours_5yrs_prior,na.rm=T),
      SD = sd(hours_of_smoke_exposure_hours_5yrs_prior,na.rm=T),
      Median = median(hours_of_smoke_exposure_hours_5yrs_prior,na.rm=T),
      LQ = quantile(hours_of_smoke_exposure_hours_5yrs_prior, 0.25,na.rm=T),
      UQ = quantile(hours_of_smoke_exposure_hours_5yrs_prior, 0.75,na.rm=T),
      Min = min(hours_of_smoke_exposure_hours_5yrs_prior,na.rm=T),
      Max = max(hours_of_smoke_exposure_hours_5yrs_prior,na.rm=T),
         T1 = quantile(hours_of_smoke_exposure_hours_5yrs_prior, 1/3, na.rm = TRUE),  # First tercile (33.3%)
    T2 = quantile(hours_of_smoke_exposure_hours_5yrs_prior, 2/3, na.rm = TRUE)   # Second
    )
smoke_5y_summary

cohort_abr <- cohort_abr %>%
  mutate(
    hours_of_smoke_exposure_hours_5yrs_prior_terciles = case_when(
      hours_of_smoke_exposure_hours_5yrs_prior == 0 ~ "no_smoke",
      hours_of_smoke_exposure_hours_5yrs_prior >= smoke_5y_summary$Min & hours_of_smoke_exposure_hours_5yrs_prior < smoke_5y_summary$T1 ~ ">0-<4h",
      hours_of_smoke_exposure_hours_5yrs_prior >= smoke_5y_summary$T1 & hours_of_smoke_exposure_hours_5yrs_prior < smoke_5y_summary$T2 ~ ">=4-<14h",
      hours_of_smoke_exposure_hours_5yrs_prior >= smoke_5y_summary$T2 ~ ">=14-368h",
      TRUE ~ NA_character_  # Handles any unexpected values
    )
  )

#smoke early years
cohort_abr$hours_of_smoke_early_life_total_dosage <- as.numeric(cohort_abr$hours_of_smoke_early_life_total_dosage)
smoke_early_summary <- cohort_abr %>%
  filter(hours_of_smoke_early_life_total_dosage > 0) %>%
  summarise(
      Mean  = mean(hours_of_smoke_early_life_total_dosage,na.rm=T),
      SD = sd(hours_of_smoke_early_life_total_dosage,na.rm=T),
      Median = median(hours_of_smoke_early_life_total_dosage,na.rm=T),
      LQ = quantile(hours_of_smoke_early_life_total_dosage, 0.25,na.rm=T),
      UQ = quantile(hours_of_smoke_early_life_total_dosage, 0.75,na.rm=T),
      Min = min(hours_of_smoke_early_life_total_dosage,na.rm=T),
      Max = max(hours_of_smoke_early_life_total_dosage,na.rm=T),
         T1 = quantile(hours_of_smoke_early_life_total_dosage, 1/3, na.rm = TRUE),  # First tercile (33.3%)
    T2 = quantile(hours_of_smoke_early_life_total_dosage, 2/3, na.rm = TRUE)   # Second
    )
smoke_early_summary

cohort_abr <- cohort_abr %>%
  mutate(
    hours_of_smoke_early_life_total_dosage_terciles = case_when(
      hours_of_smoke_early_life_total_dosage == 0 ~ "no_smoke",
      hours_of_smoke_early_life_total_dosage >= smoke_early_summary$Min & hours_of_smoke_early_life_total_dosage < smoke_early_summary$T1 ~ ">0-<2h",
      hours_of_smoke_early_life_total_dosage>= smoke_early_summary$T1 & hours_of_smoke_early_life_total_dosage < smoke_early_summary$T2 ~ ">=2-<7h",
      hours_of_smoke_early_life_total_dosage >= smoke_early_summary$T2 ~ ">=7-72h",
      TRUE ~ NA_character_  # Handles any unexpected values
    )
  )
#smoke rest of life
cohort_abr$hours_of_smoke_rest_of_life_total_dosage <- as.numeric(cohort_abr$hours_of_smoke_rest_of_life_total_dosage)
smoke_rest_summary <- cohort_abr %>%
  filter(hours_of_smoke_rest_of_life_total_dosage > 0) %>%
  summarise(
      Mean  = mean(hours_of_smoke_rest_of_life_total_dosage,na.rm=T),
      SD = sd(hours_of_smoke_rest_of_life_total_dosage,na.rm=T),
      Median = median(hours_of_smoke_rest_of_life_total_dosage,na.rm=T),
      LQ = quantile(hours_of_smoke_rest_of_life_total_dosage, 0.25,na.rm=T),
      UQ = quantile(hours_of_smoke_rest_of_life_total_dosage, 0.75,na.rm=T),
      Min = min(hours_of_smoke_rest_of_life_total_dosage,na.rm=T),
      Max = max(hours_of_smoke_rest_of_life_total_dosage,na.rm=T),
         T1 = quantile(hours_of_smoke_rest_of_life_total_dosage, 1/3, na.rm = TRUE),  # First tercile (33.3%)
    T2 = quantile(hours_of_smoke_rest_of_life_total_dosage, 2/3, na.rm = TRUE)   # Second
    )
smoke_rest_summary

cohort_abr <- cohort_abr %>%
  mutate(
    hours_of_smoke_rest_of_life_total_dosage_terciles = case_when(
      hours_of_smoke_rest_of_life_total_dosage == 0 ~ "no_smoke",
      hours_of_smoke_rest_of_life_total_dosage >= smoke_rest_summary$Min & hours_of_smoke_rest_of_life_total_dosage < smoke_rest_summary$T1 ~ ">0-<2h",
      hours_of_smoke_rest_of_life_total_dosage >= smoke_rest_summary$T1 & hours_of_smoke_rest_of_life_total_dosage < smoke_rest_summary$T2 ~ ">=2-<9h",
      hours_of_smoke_rest_of_life_total_dosage >= smoke_rest_summary$T2 ~ ">=9-378h",
      TRUE ~ NA_character_  # Handles any unexpected values
    )
  )
#smoke whole life
cohort_abr$hours_of_smoke_whole_life_total_dosage <- as.numeric(cohort_abr$hours_of_smoke_whole_life_total_dosage)
smoke_whole_summary <- cohort_abr %>%
  filter(hours_of_smoke_whole_life_total_dosage > 0) %>%
  summarise(
      Mean  = mean(hours_of_smoke_whole_life_total_dosage,na.rm=T),
      SD = sd(hours_of_smoke_whole_life_total_dosage,na.rm=T),
      Median = median(hours_of_smoke_whole_life_total_dosage,na.rm=T),
      LQ = quantile(hours_of_smoke_whole_life_total_dosage, 0.25,na.rm=T),
      UQ = quantile(hours_of_smoke_whole_life_total_dosage, 0.75,na.rm=T),
      Min = min(hours_of_smoke_whole_life_total_dosage,na.rm=T),
      Max = max(hours_of_smoke_whole_life_total_dosage,na.rm=T),
         T1 = quantile(hours_of_smoke_whole_life_total_dosage, 1/3, na.rm = TRUE),  # First tercile (33.3%)
    T2 = quantile(hours_of_smoke_whole_life_total_dosage, 2/3, na.rm = TRUE)   # Second
    )
smoke_whole_summary

cohort_abr <- cohort_abr %>%
  mutate(
    hours_of_smoke_whole_life_total_dosage_terciles = case_when(
      hours_of_smoke_whole_life_total_dosage == 0 ~ "no_smoke",
      hours_of_smoke_whole_life_total_dosage>= smoke_whole_summary$Min & hours_of_smoke_whole_life_total_dosage< smoke_whole_summary$T1 ~ ">0-<2h",
      hours_of_smoke_whole_life_total_dosage >= smoke_whole_summary$T1 & hours_of_smoke_whole_life_total_dosage < smoke_whole_summary$T2 ~ ">=2-<10h",
     hours_of_smoke_whole_life_total_dosage >= smoke_whole_summary$T2 ~ ">=10-388h",
      TRUE ~ NA_character_  # Handles any unexpected values
    )
  )

# num bred before neutering
num_breed_summary <- cohort_abr %>%
  filter(num_breeding_before_neutering > 0) %>%  # Exclude 0 values
  summarise(
    Mean  = mean(num_breeding_before_neutering, na.rm = TRUE),
    SD = sd(num_breeding_before_neutering, na.rm = TRUE),
    Median = median(num_breeding_before_neutering, na.rm = TRUE),
    LQ = quantile(num_breeding_before_neutering, 0.25, na.rm = TRUE),
    UQ = quantile(num_breeding_before_neutering, 0.75, na.rm = TRUE),
    Min = min(num_breeding_before_neutering, na.rm = TRUE),
    Max = max(num_breeding_before_neutering, na.rm = TRUE),
    T1 = quantile(num_breeding_before_neutering, 1/3, na.rm = TRUE),  # First tercile (33.3%)
    T2 = quantile(num_breeding_before_neutering, 2/3, na.rm = TRUE)   # Second tercile (66.6%)
  )

num_breed_summary
cohort_abr <- cohort_abr %>%
  mutate(
    num_breeding_before_neutering_terciles = case_when(
      num_breeding_before_neutering == 0 ~ "not_bred",
      num_breeding_before_neutering >= num_breed_summary$Min & num_breeding_before_neutering < num_breed_summary$T1 ~ ">0-<3h",
      num_breeding_before_neutering >= num_breed_summary$T1 & num_breeding_before_neutering < num_breed_summary$T2 ~ ">=3-<9h",
      num_breeding_before_neutering >= num_breed_summary$T2 ~ ">=9-142h",
      TRUE ~ NA_character_  # Handles any unexpected values
    )
  )
#group US states
table(cohort_abr$state_mode_5yrs_prior)
# Define state abbreviations grouped by region (all lowercase)
northeast <- c("ct", "me", "ma", "nh", "nj", "ny", "pa", "ri", "vt")
midwest <- c("il", "in", "ia", "ks", "mi", "mn", "mo", "ne", "nd", "oh", "sd", "wi")
south <- c("al", "ar", "de", "fl", "ga", "ky", "la", "md", "ms", "nc", "ok", "sc", "tn", "tx", "va", "wv")
west <- c("az", "ca", "co", "id", "mt", "nv", "nm", "or", "ut", "wa", "wy")

# Assuming your dataframe is called df and the state column is `state`
cohort_abr <- cohort_abr %>%
  mutate(region_mode_5y_prior = case_when(
    state_mode_5yrs_prior %in% northeast ~ "northeast",
    state_mode_5yrs_prior %in% midwest ~ "midwest",
    state_mode_5yrs_prior %in% south ~ "south",
    state_mode_5yrs_prior %in% west ~ "west",
    TRUE ~ "unknown"  # Handle any unexpected values
  ))


# exercise intensity
#5y prior to endpoint
cohort_abr$avg_intensity_5y_prior_endpoint <-as.numeric(cohort_abr$avg_intensity_5y_prior_endpoint)
int_summary <- cohort_abr %>%
  summarise(
      Mean  = mean(avg_intensity_5y_prior_endpoint,na.rm=T),
      SD = sd(avg_intensity_5y_prior_endpoint,na.rm=T),
      Median = median(avg_intensity_5y_prior_endpoint,na.rm=T),
      LQ = quantile(avg_intensity_5y_prior_endpoint, 0.25,na.rm=T),
      UQ = quantile(avg_intensity_5y_prior_endpoint, 0.75,na.rm=T),
      Min = min(avg_intensity_5y_prior_endpoint,na.rm=T),
      Max = max(avg_intensity_5y_prior_endpoint,na.rm=T)
    )
int_summary


cohort_abr <- cohort_abr %>%
  mutate(
    avg_intensity_5y_prior_endpoint_quartiles = case_when(
      (avg_intensity_5y_prior_endpoint< int_summary$LQ) ~ "0-<50",
      (avg_intensity_5y_prior_endpoint >= int_summary$LQ & avg_intensity_5y_prior_endpoint < int_summary$Median) ~ ">=50-<61",
      (avg_intensity_5y_prior_endpoint>= int_summary$Median & avg_intensity_5y_prior_endpoint < int_summary$UQ) ~ ">=61-<68.9",
      (avg_intensity_5y_prior_endpoint>= int_summary$UQ) ~ ">=68.9-100",
    )
  )

# activity frequency
#early
#attempting dynamic labelling for these quartiles
# Compute quartiles dynamically


# Function to calculate quartiles or terciles dynamically
get_breaks <- function(x, probs) {
  breaks <- quantile(x, probs = probs, na.rm = TRUE)
  
  # If breaks are not unique, switch to terciles
  if (length(unique(breaks)) < length(probs)) {
    message("Quartiles not unique, switching to terciles.")
    breaks <- quantile(x, probs = c(0, 1/3, 2/3, 1), na.rm = TRUE)
  }
  
  return(breaks)
}

# Calculate quartiles first, switch to terciles if needed
breaks <- get_breaks(cohort_abr$avg_activity_freq_sy1_2, c(0, 0.25, 0.5, 0.75, 1))

# Assign labels dynamically
labels <- paste0("≥", round(head(breaks, -1), 2), " - <", round(tail(breaks, -1), 2))

# Apply the breaks using `cut()`
cohort_abr <- cohort_abr %>%
  mutate(avg_activity_freq_sy1_2_terciles = cut(avg_activity_freq_sy1_2, 
                                                breaks = breaks, 
                                                include.lowest = TRUE, 
                                                labels = labels))

# Check output
table(cohort_abr$avg_activity_freq_sy1_2_terciles)

#rest
# Calculate quartiles first, switch to terciles if needed
cohort_abr$avg_activity_freq_sy3 <- as.numeric(cohort_abr$avg_activity_freq_sy3)
breaks <- get_breaks(cohort_abr$avg_activity_freq_sy3, c(0, 0.25, 0.5, 0.75, 1))

# Assign labels dynamically
labels <- paste0("≥", round(head(breaks, -1), 2), " - <", round(tail(breaks, -1), 2))

# Apply the breaks using `cut()`
cohort_abr <- cohort_abr %>%
  mutate(avg_activity_freq_sy3_terciles = cut(avg_activity_freq_sy3, 
                                                breaks = breaks, 
                                                include.lowest = TRUE, 
                                                labels = labels))

# Check output
table(cohort_abr$avg_activity_freq_sy3_terciles)


## intensity
#sy1_2
# Calculate quartiles first, switch to terciles if needed - not enough breaks for this one
#cohort_abr$avg_activity_intensity_sy1_2 <- as.numeric(cohort_abr$avg_activity_intensity_sy1_2)
#breaks <- get_breaks(cohort_abr$avg_activity_intensity_sy1_2, c(0, 0.25, 0.5, 0.75, 1))

# Assign labels dynamically
#labels <- paste0("≥", round(head(breaks, -1), 2), " - <", round(tail(breaks, -1), 2))

# Apply the breaks using `cut()`
#cohort_abr <- cohort_abr %>%
#  mutate(avg_activity_intensity_sy1_2_terciles = cut(avg_activity_intensity_sy1_2, 
                               #                 breaks = breaks, 
                               #                 include.lowest = TRUE, 
                                #                labels = labels))

# Check output
#table(cohort_abr$avg_activity_intensity_sy1_2_terciles)
#sy3+
cohort_abr$avg_activity_intensity_sy3 <- as.numeric(cohort_abr$avg_activity_intensity_sy3)
breaks <- get_breaks(cohort_abr$avg_activity_intensity_sy3, c(0, 0.25, 0.5, 0.75, 1))

# Assign labels dynamically
labels <- paste0("≥", round(head(breaks, -1), 2), " - <", round(tail(breaks, -1), 2))

# Apply the breaks using `cut()`
cohort_abr <- cohort_abr %>%
  mutate(avg_activity_intensity_sy3_quartiles = cut(avg_activity_freq_sy3, 
                                                breaks = breaks, 
                                                include.lowest = TRUE, 
                                                labels = labels))

#frequency
#SY1-2
cohort_abr$avg_activity_freq_sy1_2 <- as.numeric(cohort_abr$avg_activity_freq_sy1_2)
breaks <- get_breaks(cohort_abr$avg_activity_freq_sy1_2, c(0, 0.25, 0.5, 0.75, 1))


# Assign labels dynamically
labels <- paste0("≥", round(head(breaks, -1), 2), " - <", round(tail(breaks, -1), 2))

# Apply the breaks using `cut()`
cohort_abr <- cohort_abr %>%
  mutate(avg_activity_freq_sy1_2_quartiles = cut(avg_activity_freq_sy1_2, 
                                                breaks = breaks, 
                                                include.lowest = TRUE, 
                                                labels = labels))

cohort_abr$avg_frequency_5y_prior_endpoint <- as.numeric(cohort_abr$avg_frequency_5y_prior_endpoint)
breaks <- get_breaks(cohort_abr$avg_frequency_5y_prior_endpoint, c(0, 0.25, 0.5, 0.75, 1))


# Assign labels dynamically
labels <- paste0("≥", round(head(breaks, -1), 2), " - <", round(tail(breaks, -1), 2))

# Apply the breaks using `cut()`
cohort_abr <- cohort_abr %>%
  mutate(avg_frequency_5y_prior_endpoint_quartiles = cut(avg_frequency_5y_prior_endpoint, 
                                                breaks = breaks, 
                                                include.lowest = TRUE, 
                                                labels = labels))


#duration
cohort_abr$avg_activity_duration_sy1_2 <- as.numeric(cohort_abr$avg_activity_duration_sy1_2)
breaks <- get_breaks(cohort_abr$avg_activity_duration_sy1_2, c(0, 0.25, 0.5, 0.75, 1))


# Assign labels dynamically
labels <- paste0("≥", round(head(breaks, -1), 2), " - <", round(tail(breaks, -1), 2))

# Apply the breaks using `cut()`
cohort_abr <- cohort_abr %>%
  mutate(avg_activity_duration_sy1_2_quartiles = cut(avg_activity_duration_sy1_2, 
                                                breaks = breaks, 
                                                include.lowest = TRUE, 
                                                labels = labels))


#incense - early life and rest of life not enough breaks to do terciles
#whole of life

cohort_abr$use_incense_or_candles_sum_whole_life <- as.numeric(cohort_abr$use_incense_or_candles_sum_whole_life)
breaks <- get_breaks(cohort_abr$use_incense_or_candles_sum_whole_life, c(0, 0.25, 0.5, 0.75, 1))


# Assign labels dynamically
labels <- paste0("≥", round(head(breaks, -1), 2), " - <", round(tail(breaks, -1), 2))

# Apply the breaks using `cut()`
cohort_abr <- cohort_abr %>%
  mutate(use_incense_or_candles_sum_whole_life_terciles = cut(use_incense_or_candles_sum_whole_life, 
                                                breaks = breaks, 
                                                include.lowest = TRUE, 
                                                labels = labels))


#weeds - all 0s - not quartiles/terciles

#insect - not enough variation for early life, rest of life, whole life


#aerosol - not enough var 
#hepa - not enough var


###household
#whole
cohort_abr$in.the.house_YN_sum_whole_life <- as.numeric(cohort_abr$in.the.house_YN_sum_whole_life)
breaks <- get_breaks(cohort_abr$in.the.house_YN_sum_whole_life, c(0, 0.25, 0.5, 0.75, 1))


# Assign labels dynamically
labels <- paste0("≥", round(head(breaks, -1), 2), " - <", round(tail(breaks, -1), 2))

# Apply the breaks using `cut()`
cohort_abr <- cohort_abr %>%
  mutate(in.the.house_YN_sum_whole_life_quartiles = cut(in.the.house_YN_sum_whole_life, 
                                                breaks = breaks, 
                                                include.lowest = TRUE, 
                                                labels = labels))
#rest
cohort_abr$in.the.house_YN_sum_rest_of_life <- as.numeric(cohort_abr$in.the.house_YN_sum_rest_of_life)
breaks <- get_breaks(cohort_abr$in.the.house_YN_sum_rest_of_life, c(0, 0.25, 0.5, 0.75, 1))


# Assign labels dynamically
labels <- paste0("≥", round(head(breaks, -1), 2), " - <", round(tail(breaks, -1), 2))

# Apply the breaks using `cut()`
cohort_abr <- cohort_abr %>%
  mutate(in.the.house_YN_sum_rest_of_life_quartiles = cut(in.the.house_YN_sum_rest_of_life, 
                                                breaks = breaks, 
                                                include.lowest = TRUE, 
                                                labels = labels))
#early
cohort_abr$in.the.house_YN_sum_early_life <- as.numeric(cohort_abr$in.the.house_YN_sum_early_life)
breaks <- get_breaks(cohort_abr$in.the.house_YN_sum_early_life, c(0, 0.25, 0.5, 0.75, 1))


# Assign labels dynamically
labels <- paste0("≥", round(head(breaks, -1), 2), " - <", round(tail(breaks, -1), 2))

# Apply the breaks using `cut()`
cohort_abr <- cohort_abr %>%
  mutate(in.the.house_YN_sum_early_life_terciles = cut(in.the.house_YN_sum_early_life, 
                                                breaks = breaks, 
                                                include.lowest = TRUE, 
                                                labels = labels))



#outside - not enough var

#garage - not enough var


```

```{r}
#last tidying up of variables
cohort_abr <- cohort_abr %>%
  dplyr::mutate(heating_fuel_primary_mode_5yrs_prior = case_when(
    heating_fuel_primary_mode_5yrs_prior == "natural gas" ~ "natural_gas",
    TRUE ~ heating_fuel_primary_mode_5yrs_prior
  ))



#recode 0 in neutered_reason to not_neutered and unknown as not_recorded

cohort_abr <- cohort_abr %>%
  mutate(neutered_reason = recode(neutered_reason, "0" = "Entire"))


cohort_abr <- cohort_abr %>%
  mutate(
    heating_fuel_primary_mode_5yrs_prior = recode(heating_fuel_primary_mode_5yrs_prior, "unknown" = "not_recorded"),
    cooking_fuel_primary_mode_5yrs_prior = recode(cooking_fuel_primary_mode_5yrs_prior, "unknown" = "not_recorded"),
    heating_fuel_secondary_mode_5yrs_prior = recode(heating_fuel_secondary_mode_5yrs_prior, "unknown" = "not_recorded"),
    cooking_fuel_secondary_mode_5yrs_prior = recode(cooking_fuel_secondary_mode_5yrs_prior, "unknown" = "not_recorded"),
    region_mode_5y_prior = recode(region_mode_5y_prior, "unknown" = "not_recorded")
  )
```

```{r}
#replace all missing with NA (so "" and NA)
cohort_abr <- cohort_abr %>%
  mutate(across(everything(), as.character)) %>%  # Ensure all columns are character
  mutate(across(everything(), ~na_if(.x, "n/a"))) %>%
  mutate(across(everything(), ~na_if(.x, "NA"))) %>%
  mutate(across(everything(), ~na_if(.x, "nr"))) %>%
  mutate(across(everything(), ~na_if(.x, "not recorded"))) %>%
  mutate(across(everything(), ~na_if(.x, ""))) %>%
  mutate(across(everything(), ~replace_na(.x, "not_recorded")))
cohort_abr$survival_time_months <- as.numeric(cohort_abr$survival_time_months)


cohort_abr$status_factor <- as.factor(as.character(cohort_abr$had_HSA))
cohort_abr$status <-as.numeric(cohort_abr$had_HSA) 
cohort_abr$survival_time_months <- as.numeric(cohort_abr$survival_time_months)


#manually assign height, weight, BCS factor levles and age at neuter and age at diagnosis 
#height
cohort_abr <- cohort_abr%>%
  mutate(avg_height_5y_prior_endpoint_quartiles = factor(avg_height_5y_prior_endpoint_quartiles, levels = c("49.13-<56.07cm", ">=56.07-<58.42cm", ">=58.42-<60.32cm", ">=60.32cm-67.73cm","not_recorded")))
#weight
cohort_abr <- cohort_abr%>%
  mutate(avg_weight_5y_prior_endpoint_quartiles = factor(avg_weight_5y_prior_endpoint_quartiles, levels = c("9.75-<21.31kg", ">=21.31-<24.66kg", ">=24.66-<29.37kg", ">=29.37-41.52kg","not_recorded")))
#bcs
cohort_abr <- cohort_abr%>%
  mutate(avg_purina_BCS_5y_prior_endpoint_quartiles = factor(avg_purina_BCS_5y_prior_endpoint_quartiles, levels = c("3.3-<5", ">=5-<5.28", ">=5.28-<5.86", ">=5.86 - 9","not_recorded")))
#neut age

cohort_abr <- cohort_abr%>%
  mutate(age_at_neuter_years_quartiles = factor(age_at_neuter_years_quartiles, levels = c("0.15-<0.52y", ">=0.52-<0.83y", ">=0.83-2.44y", ">=2.44-9.59y","Entire","not_recorded")))
#diagnosis age
cohort_abr <- cohort_abr%>%
  mutate(age_at_final_date_quartiles = factor(age_at_final_date_quartiles, levels = c("0.74-<8.47 years", ">=8.47-<10.68 years", ">=10.68-<11.64 years", ">=11.64-14.51 years","not_recorded")))

#change spayed while in heat to account for males and entire females
cohort_abr <- cohort_abr %>%
  mutate(spayed_while_in_heat = case_when(
    sex== "M" ~ "Male",
    sex_status == "Intact Female" ~ "Intact Female",
    spayed_while_in_heat == "0" ~ "0",
    spayed_while_in_heat == "1" ~"1"
  ))
#same as above account for unnetered and males 
heats <- cohort_abr %>%
# Convert to numeric
  filter(heats_before_neutering != 0) %>% 
# Remove rows where the column is 0
  dplyr::select(heats_before_neutering) %>%
  filter(heats_before_neutering != "not_recorded") # Keep only this column
heats <- as.numeric(heats$heats_before_neutering)

breaks <- get_breaks(heats, c(0, 0.25, 0.5, 0.75, 1))


# Assign labels dynamically
labels <- paste0("≥", round(head(breaks, -1), 2), " - <", round(tail(breaks, -1), 2))
cohort_abr <- cohort_abr %>%
  mutate(heats_before_neutering_terciles = case_when(
    sex== "M" ~ "Male",
    sex_status == "Intact Female" ~ "Intact Female",
    heats_before_neutering == "0" ~ "0",
    heats_before_neutering == "1" ~"1-<=2",
    heats_before_neutering == "2" ~"1-<=2",
    heats_before_neutering == "3" ~">2-<=4",
    heats_before_neutering == "4" ~">2-<=4",
    heats_before_neutering == "5" ~">4-12",
    heats_before_neutering == "6" ~">4-12",
    heats_before_neutering == "7" ~">4-12",
    heats_before_neutering == "8" ~">4-12",
    heats_before_neutering == "9" ~">4-12",
    heats_before_neutering == "10" ~">4-12",
    heats_before_neutering == "11" ~">4-12",
    heats_before_neutering == "12" ~">4-12"
  ))


#mode BCS classified
cohort_abr <- cohort_abr %>%
  mutate(mode_BCS_classify = case_when(
    mode_purina_BCS_5y_prior_endpoint == "1" ~"underweight",
    mode_purina_BCS_5y_prior_endpoint == "2" ~"underweight",
    mode_purina_BCS_5y_prior_endpoint == "3" ~"underweight",
    mode_purina_BCS_5y_prior_endpoint == "4" ~"ideal",
    mode_purina_BCS_5y_prior_endpoint == "5" ~"ideal",
    mode_purina_BCS_5y_prior_endpoint == "6" ~"overweight",
    mode_purina_BCS_5y_prior_endpoint == "7" ~"overweight",
    mode_purina_BCS_5y_prior_endpoint == "8" ~"overweight",
    mode_purina_BCS_5y_prior_endpoint == "9" ~"overweight",
    mode_purina_BCS_5y_prior_endpoint == "not_recorded"~"not_recorded"
  ))

#
cohort_abr <- cohort_abr%>%
  mutate(heats_before_neutering_terciles = factor(heats_before_neutering_terciles, levels = c("0", "1-<=2", ">2-<=4", ">4-12","Intact Female","Male","not_recorded")))



```

Combine the lifelong comorbidity and 5y prior t endpoint comorbidities for cardiovascular, orthopaedic, immune med, GI, other chronic inflam

```{r}
#replace Y and N with 0 and 1
cohort_abr <- cohort_abr %>%
  mutate(across(c(cardiovascular, orthopaedic,immune_mediated,gastrointestinal,chronic_inflammatory), ~ ifelse(. == "Y", 1, 0)))

cohort_abr <- cohort_abr %>%
  mutate(across(c(lifelong_cardiovascular,lifelong_orthopaedic,lifelong_immune_mediated,lifelong_gastrointestinal,lifelong_chronic_inflammatory), as.numeric))
cohort_abr <- cohort_abr %>%
  mutate(comorb_cardiovascular_all = cardiovascular + lifelong_cardiovascular,
         comorb_orthopaedic_all = orthopaedic + lifelong_orthopaedic,
         comorb_immune_med_all = immune_mediated + lifelong_immune_mediated,
         comorb_gastrointestinal_all = gastrointestinal+lifelong_gastrointestinal,
         comorb_chronic_inflam_all = chronic_inflammatory + lifelong_chronic_inflammatory
    
  )
#in these new columns now replace the 2 with 1 so have an UTD count of all possible comorbs 
cohort_abr <- cohort_abr %>%
  mutate(across(c(comorb_cardiovascular_all,comorb_orthopaedic_all,comorb_immune_med_all,comorb_gastrointestinal_all,comorb_chronic_inflam_all), ~ ifelse(. == 2, 1, .)))
```

```{r}
write.csv(cohort_abr,"C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_cohort_all_RFs_data_tidied_from_cox.csv")
```

## Setting factors with baselines correct

```{r}
#ensure not_recorded is default not baseline
convert_and_relevel <- function(df) {
  df <- df %>%
    mutate(across(where(is.character), ~ factor(.))) %>%  # Convert all character columns to factors
    mutate(across(where(is.factor), ~ fct_relevel(., "not_recorded", after = Inf)))  # Move "not_recorded" to last place
  return(df)
}

# Apply to your dataset
cohort_abr <- convert_and_relevel(cohort_abr)


```

## Kaplan Meier curves

```{r}
survfit(Surv(cohort_abr$survival_time_months, cohort_abr$status) ~ 1, data = cohort_abr) %>% 
  ggsurvfit() +
  labs(
    x = "Months",
    y = "Overall survival probability of cohort"
  ) +  add_confidence_interval() +
  add_risktable()
```

```{r}
survdiff(Surv(survival_time_months, status) ~ neuter, data = cohort_abr)

#KM for neuter
dependent_os = "Surv(survival_time_months, status)"
explanatory = c("neuter")

cohort_abr %>% 
    surv_plot(dependent_os, explanatory, pval = TRUE)
```

## Univariable Cox PH of variables

-   add in full list of variables

-   make sure factors make sense and baseline isnt things like not recorded etc.

-   there is a variable with \<2 levels in it

-   NB way function is written is surv(survival_time_days) could set to months to calc for more appropriate month of survival

```{r}


check_levels <- function(df) {
  sapply(df, function(x) length(unique(na.omit(x))) >= 2)
}

# Example usage
result <- check_levels(cohort_abr)

# Print columns that do not meet the condition
names(result[result == FALSE])


source('C:/Users/ctaylor18/GitHub/HSA-VC-survival-analyses/HSA-VC-survival-analyses/code/functions/univar_cox_ph_table_funct.R')

cohort_abr$status_factor <- as.factor(as.character(cohort_abr$had_HSA))
cohort_abr$status <-as.numeric(cohort_abr$had_HSA) 
cohort_abr$survival_time_months <- as.numeric(cohort_abr$survival_time_months)

cohort_abr$survival_time_days <- as.numeric(cohort_abr$survival_time_months)


#keep a second dataset of this for NA removed from univar
cohort_no_not_rec <- cohort_abr


# Usage example
perform_univariable_analysis(
  data = cohort_abr,
  dependent_var = "status",
  dependent_var_factor="status_factor",
  explanatory_vars = c("MDI_quintile","owner_MDI_quintile","area_type_mode_5yrs_prior","house_type_mode_5yrs_prior", 
             "water_source_mode_5yrs_prior", "pipes_metal_any_mode_5yrs_prior", "pipes_plastic_any_mode_5yrs_prior", 
             "heating_fuel_primary_mode_5yrs_prior", "cooking_fuel_primary_mode_5yrs_prior", "heating_fuel_secondary_mode_5yrs_prior", 
             "cooking_fuel_secondary_mode_5yrs_prior", "region_mode_5y_prior","chronic_inflam_present","intestinal_parasitism","tick_borne_parasitism","other_parasitism","all_parasitism","infectious","orthopaedic","immune_mediated","cardiovascular","gastrointestinal","inflammatory_other","lifelong_comorb_present.x","lifelong_orthopaedic","lifelong_immune_mediated","lifelong_cardiovascular","lifelong_gastrointestinal","lifelong_inflammatory_other","lifelong_chronic_inflammatory","number_lifelong_comorbs_present",  #"number_comorbs_present_lifetime", 
            # "study_year_first_chronic_inflam", #"chronic_inflam_present_5y", #"number_comorbs_present_5y", 
            # "chronic_inflam_present_no_min_year", "number_comorbs_present_no_min_year",  "diagnosis_after_comorb",  "time_between_first_comorb_and_endpoint",  
              "num_lifetime_chronic_inflam", "intestinal_parasitism","tick_borne_parasitism","other_parasitism","all_parasitism","infectious","orthopaedic","immune_mediated","cardiovascular","gastrointestinal","inflammatory_other","chronic_inflammatory","comorb_cardiovascular_all","comorb_orthopaedic_all","comorb_immune_med_all","comorb_gastrointestinal_all","comorb_chronic_inflam_all", "USDA_30min_sy1_2", "activity_1h_min_sy1_2", 
             "activity_KC_2h_min_sy1_2", "avg_activity_duration_sy1_2_quartiles",#"avg_activity_intensity_sy1_2_terciles",
 "avg_activity_freq_sy1_2_terciles", 
             "avg_activity_freq_sy3_terciles", "avg_activity_intensity_sy3_quartiles", "avg_intensity_5y_prior_endpoint_quartiles", "avg_frequency_5y_prior_endpoint_quartiles", 
             "use_aerosol_study_years_early_life", "use_air_cleaner_study_years_early_life", "use_hepa_filter_study_years_early_life", 
             "use_moth_balls_study_years_early_life", "use_incense_or_candles_study_years_early_life", "smoke_exposure_study_years_early_life", 
       #      "any_treated_weeds_study_years_early_life", "any_treated_insects_study_years_early_life", #"any_treated_fertilizer_study_years_early_life", 
         #    "use_aerosol_sum_early_life", "use_air_cleaner_sum_early_life", "use_hepa_filter_sum_early_life", "use_moth_balls_sum_early_life", 
          #   "use_incense_or_candles_sum_early_life", 
 #"smoke_exposure_sum_early_life",
 #"any_treated_weeds_sum_early_life", 
    #         "any_treated_insects_sum_early_life", #"any_treated_fertilizer_sum_early_life", 
 "use_aerosol_study_years_rest_of_life", 
             "use_air_cleaner_study_years_rest_of_life", "use_hepa_filter_study_years_rest_of_life", "use_moth_balls_study_years_rest_of_life", 
             "use_incense_or_candles_study_years_rest_of_life", "smoke_exposure_study_years_rest_of_life", "any_treated_weeds_study_years_rest_of_life", 
             "any_treated_insects_study_years_rest_of_life", "any_treated_fertilizer_study_years_rest_of_life", #"use_aerosol_sum_rest_of_life", 
 #            "use_air_cleaner_sum_rest_of_life", "use_hepa_filter_sum_rest_of_life", "use_moth_balls_sum_rest_of_life", 
          #   "use_incense_or_candles_sum_rest_of_life", 
# "smoke_exposure_sum_rest_of_life", 
# "any_treated_weeds_sum_rest_of_life", 
            # "any_treated_insects_sum_rest_of_life", "any_treated_fertilizer_sum_rest_of_life", 
"use_aerosol_study_years_whole_life", 
             "use_air_cleaner_study_years_whole_life", "use_hepa_filter_study_years_whole_life", "use_moth_balls_study_years_whole_life", 
             "use_incense_or_candles_study_years_whole_life", "smoke_exposure_study_years_whole_life", "any_treated_weeds_study_years_whole_life", 
             "any_treated_insects_study_years_whole_life", "any_treated_fertilizer_study_years_whole_life",# "use_aerosol_sum_whole_life", 
           #  "use_air_cleaner_sum_whole_life", "use_hepa_filter_sum_whole_life", "use_moth_balls_sum_whole_life",# "use_incense_or_candles_sum_whole_life", 
            # "smoke_exposure_sum_whole_life", 
 #"any_treated_weeds_sum_whole_life", "any_treated_insects_sum_whole_life", 
            # "any_treated_fertilizer_sum_whole_life",
 "hours_of_smoke_early_life_total_dosage_terciles", "hours_of_smoke_rest_of_life_total_dosage_terciles", 
             "hours_of_smoke_whole_life_total_dosage_terciles", "in.the.garage_YN_study_years_early_life", "in.the.house_YN_study_years_early_life", 
             "outside_YN_study_years_early_life",  "in.the.house_YN_sum_early_life_terciles", 
             "in.the.garage_YN_study_years_rest_of_life", "in.the.house_YN_study_years_rest_of_life", "outside_YN_study_years_rest_of_life", 
              "in.the.house_YN_sum_rest_of_life_quartiles",  
             "in.the.garage_YN_study_years_whole_life", "in.the.house_YN_study_years_whole_life", "outside_YN_study_years_whole_life", 
             "in.the.house_YN_sum_whole_life_quartiles",  
             "majority_sleep_location_lifetime",  "main_lifestyle_category", 
 "main_lifestyle_category_mode_5yrs_prior", "sleep_location_mode_5yrs_prior", 
             "use_aerosol_exposed_5yrs_prior", "use_air_cleaner_exposed_5yrs_prior", "use_hepa_filter_exposed_5yrs_prior", 
             "use_moth_balls_exposed_5yrs_prior", "use_incense_or_candles_exposed_5yrs_prior", "smoke_exposure_exposed_5yrs_prior", 
             "any_treated_weeds_exposed_5yrs_prior", "any_treated_insects_exposed_5yrs_prior", "any_treated_fertilizer_exposed_5yrs_prior", 
             "hours_of_smoke_exposure_hours_5yrs_prior_terciles", "sex", "neuter", "mismating_ever",
             "neutered_reason", "neutered_ever_bred", "spayed_while_in_heat", "spayed_while_pregnant", "prevent_heat_ever", 
             "no_pregnancy_ever", "intact_sired_litters_natural_ever", "neutered_females_bred_with_count_ever", 
             "neutered_sired_litters_natural_ever", "heats_before_neutering", "heats_before_neutering_terciles",
             "preg_pre_neuter",# "mated_pre_neuter",
 "num_breeding_before_neutering_terciles", "ever_bred_from", 
              "avg_weight_5y_prior_endpoint_quartiles", "avg_height_5y_prior_endpoint_quartiles", 
             "avg_purina_BCS_5y_prior_endpoint_quartiles", "mode_purina_BCS_5y_prior_endpoint", "mode_BCS_classify", "half_anti_microbials_lifetime", "half_dietary_supplements_lifetime", 
             "half_enteric_lifetime", "half_fluid_metabolites_lifetime", "half_hormones_and_related_lifetime", "half_immunologicals_lifetime", 
             "half_misc_lifetime", "half_neurological_lifetime", "half_anti_parasites_lifetime", "half_anti_septics_lifetime", 
             "half_cardio_respiratory_lifetime", "half_diuretics_lifetime", "half_anti_hist_lifetime", 
             "quartile_anti_inflam_lifetime", "quartile_anti_microbials_lifetime", "quartile_dietary_supplements_lifetime", 
            #not enough data for quartiles "quartile_enteric_lifetime", "quartile_fluid_metabolites_lifetime", 
 "quartile_hormones_and_related_lifetime", 
             "quartile_immunologicals_lifetime", "quartile_misc_lifetime", "quartile_neurological_lifetime", 
             "quartile_anti_parasites_lifetime", 
 #not enough data for quartiles "quartile_anti_septics_lifetime", "quartile_cardio_respiratory_lifetime",   "quartile_diuretics_lifetime", "quartile_anti_hist_lifetime", 
             "half_anti_inflam_5y_prior_endpoint", "half_anti_microbials_5y_prior_endpoint", "half_enteric_5y_prior_endpoint", 
             "half_immunologicals_5y_prior_endpoint", "half_misc_5y_prior_endpoint", "half_neurological_5y_prior_endpoint", 
             "half_anti_parasites_5y_prior_endpoint", "half_dietary_supplements_5y_prior_endpoint", 
             "half_anti_septics_5y_prior_endpoint", "half_hormones_and_related_5y_prior_endpoint", 
             "half_cardio_respiratory_5y_prior_endpoint", "half_fluid_metabolites_5y_prior_endpoint", 
             "half_diuretics_5y_prior_endpoint", "half_anti_hist_5y_prior_endpoint", 
             "quartile_anti_inflam_5y_prior_endpoint", "quartile_anti_microbials_5y_prior_endpoint", 
            # "quartile_enteric_5y_prior_endpoint",
 "quartile_immunologicals_5y_prior_endpoint", 
             "quartile_misc_5y_prior_endpoint", "quartile_neurological_5y_prior_endpoint", 
             "quartile_anti_parasites_5y_prior_endpoint", "quartile_dietary_supplements_5y_prior_endpoint", 
       #      "quartile_anti_septics_5y_prior_endpoint", 
 "quartile_hormones_and_related_5y_prior_endpoint", 
          #   "quartile_cardio_respiratory_5y_prior_endpoint", #"quartile_fluid_metabolites_5y_prior_endpoint", 
         #    "quartile_diuretics_5y_prior_endpoint", "quartile_anti_hist_5y_prior_endpoint", 
             #"quartile_anti_viral_5y_prior_endpoint",
 "age_at_final_date_quartiles","age_at_neuter_years_quartiles"),
  output_csv = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_time_to_HSA_cox_ph_univar.csv",
  output_doc = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_time_to_HSA_cox_ph_univar.docx"
)


```

Create a counts cross table for each variable:

```{r}
library(finalfit)
explanatory = c("MDI_quintile","owner_MDI_quintile","area_type_mode_5yrs_prior","house_type_mode_5yrs_prior", 
             "water_source_mode_5yrs_prior", "pipes_metal_any_mode_5yrs_prior", "pipes_plastic_any_mode_5yrs_prior", 
             "heating_fuel_primary_mode_5yrs_prior", "cooking_fuel_primary_mode_5yrs_prior", "heating_fuel_secondary_mode_5yrs_prior", 
             "cooking_fuel_secondary_mode_5yrs_prior", "region_mode_5y_prior","chronic_inflam_present","intestinal_parasitism","tick_borne_parasitism","other_parasitism","all_parasitism","infectious","orthopaedic","immune_mediated","cardiovascular","gastrointestinal","inflammatory_other","lifelong_comorb_present.x","lifelong_orthopaedic","lifelong_immune_mediated","lifelong_cardiovascular","lifelong_gastrointestinal","lifelong_inflammatory_other","lifelong_chronic_inflammatory","number_lifelong_comorbs_present",  #"number_comorbs_present_lifetime", 
            # "study_year_first_chronic_inflam", #"chronic_inflam_present_5y", #"number_comorbs_present_5y", 
            # "chronic_inflam_present_no_min_year", "number_comorbs_present_no_min_year",  "diagnosis_after_comorb",  "time_between_first_comorb_and_endpoint",  
              "num_lifetime_chronic_inflam", "intestinal_parasitism","tick_borne_parasitism","other_parasitism","all_parasitism","infectious","orthopaedic","immune_mediated","cardiovascular","gastrointestinal","inflammatory_other","chronic_inflammatory","comorb_cardiovascular_all","comorb_orthopaedic_all","comorb_immune_med_all","comorb_gastrointestinal_all","comorb_chronic_inflam_all", 
 "USDA_30min_sy1_2", "activity_1h_min_sy1_2", 
             "activity_KC_2h_min_sy1_2", "avg_activity_duration_sy1_2_quartiles",#"avg_activity_intensity_sy1_2_terciles",
 "avg_activity_freq_sy1_2_terciles", 
             "avg_activity_freq_sy3_terciles", "avg_activity_intensity_sy3_quartiles", "avg_intensity_5y_prior_endpoint_quartiles", "avg_frequency_5y_prior_endpoint_quartiles", 
             "use_aerosol_study_years_early_life", "use_air_cleaner_study_years_early_life", "use_hepa_filter_study_years_early_life", 
             "use_moth_balls_study_years_early_life", "use_incense_or_candles_study_years_early_life", "smoke_exposure_study_years_early_life", 
       #      "any_treated_weeds_study_years_early_life", "any_treated_insects_study_years_early_life", #"any_treated_fertilizer_study_years_early_life", 
         #    "use_aerosol_sum_early_life", "use_air_cleaner_sum_early_life", "use_hepa_filter_sum_early_life", "use_moth_balls_sum_early_life", 
          #   "use_incense_or_candles_sum_early_life", 
 #"smoke_exposure_sum_early_life",
 #"any_treated_weeds_sum_early_life", 
    #         "any_treated_insects_sum_early_life", #"any_treated_fertilizer_sum_early_life", 
 "use_aerosol_study_years_rest_of_life", 
             "use_air_cleaner_study_years_rest_of_life", "use_hepa_filter_study_years_rest_of_life", "use_moth_balls_study_years_rest_of_life", 
             "use_incense_or_candles_study_years_rest_of_life", "smoke_exposure_study_years_rest_of_life", "any_treated_weeds_study_years_rest_of_life", 
             "any_treated_insects_study_years_rest_of_life", "any_treated_fertilizer_study_years_rest_of_life", #"use_aerosol_sum_rest_of_life", 
 #            "use_air_cleaner_sum_rest_of_life", "use_hepa_filter_sum_rest_of_life", "use_moth_balls_sum_rest_of_life", 
          #   "use_incense_or_candles_sum_rest_of_life", 
# "smoke_exposure_sum_rest_of_life", 
# "any_treated_weeds_sum_rest_of_life", 
            # "any_treated_insects_sum_rest_of_life", "any_treated_fertilizer_sum_rest_of_life", 
"use_aerosol_study_years_whole_life", 
             "use_air_cleaner_study_years_whole_life", "use_hepa_filter_study_years_whole_life", "use_moth_balls_study_years_whole_life", 
             "use_incense_or_candles_study_years_whole_life", "smoke_exposure_study_years_whole_life", "any_treated_weeds_study_years_whole_life", 
             "any_treated_insects_study_years_whole_life", "any_treated_fertilizer_study_years_whole_life",# "use_aerosol_sum_whole_life", 
           #  "use_air_cleaner_sum_whole_life", "use_hepa_filter_sum_whole_life", "use_moth_balls_sum_whole_life",# "use_incense_or_candles_sum_whole_life", 
            # "smoke_exposure_sum_whole_life", 
 #"any_treated_weeds_sum_whole_life", "any_treated_insects_sum_whole_life", 
            # "any_treated_fertilizer_sum_whole_life",
 "hours_of_smoke_early_life_total_dosage_terciles", "hours_of_smoke_rest_of_life_total_dosage_terciles", 
             "hours_of_smoke_whole_life_total_dosage_terciles", "in.the.garage_YN_study_years_early_life", "in.the.house_YN_study_years_early_life", 
             "outside_YN_study_years_early_life",  "in.the.house_YN_sum_early_life_terciles", 
             "in.the.garage_YN_study_years_rest_of_life", "in.the.house_YN_study_years_rest_of_life", "outside_YN_study_years_rest_of_life", 
              "in.the.house_YN_sum_rest_of_life_quartiles",  
             "in.the.garage_YN_study_years_whole_life", "in.the.house_YN_study_years_whole_life", "outside_YN_study_years_whole_life", 
             "in.the.house_YN_sum_whole_life_quartiles",  
             "majority_sleep_location_lifetime",  "main_lifestyle_category", 
 "main_lifestyle_category_mode_5yrs_prior", "sleep_location_mode_5yrs_prior", 
             "use_aerosol_exposed_5yrs_prior", "use_air_cleaner_exposed_5yrs_prior", "use_hepa_filter_exposed_5yrs_prior", 
             "use_moth_balls_exposed_5yrs_prior", "use_incense_or_candles_exposed_5yrs_prior", "smoke_exposure_exposed_5yrs_prior", 
             "any_treated_weeds_exposed_5yrs_prior", "any_treated_insects_exposed_5yrs_prior", "any_treated_fertilizer_exposed_5yrs_prior", 
             "hours_of_smoke_exposure_hours_5yrs_prior_terciles", "sex", "neuter", "mismating_ever",
             "neutered_reason", "neutered_ever_bred", "spayed_while_in_heat", "spayed_while_pregnant", "prevent_heat_ever", 
             "no_pregnancy_ever", "intact_sired_litters_natural_ever", "neutered_females_bred_with_count_ever", 
             "neutered_sired_litters_natural_ever", "heats_before_neutering", "heats_before_neutering_terciles",
             "preg_pre_neuter",# "mated_pre_neuter",
 "num_breeding_before_neutering_terciles", "ever_bred_from", 
              "avg_weight_5y_prior_endpoint_quartiles", "avg_height_5y_prior_endpoint_quartiles", 
             "avg_purina_BCS_5y_prior_endpoint_quartiles", "mode_purina_BCS_5y_prior_endpoint", "mode_BCS_classify", "half_anti_microbials_lifetime", "half_dietary_supplements_lifetime", 
             "half_enteric_lifetime", "half_fluid_metabolites_lifetime", "half_hormones_and_related_lifetime", "half_immunologicals_lifetime", 
             "half_misc_lifetime", "half_neurological_lifetime", "half_anti_parasites_lifetime", "half_anti_septics_lifetime", 
             "half_cardio_respiratory_lifetime", "half_diuretics_lifetime", "half_anti_hist_lifetime", 
             "quartile_anti_inflam_lifetime", "quartile_anti_microbials_lifetime", "quartile_dietary_supplements_lifetime", 
            #not enough data for quartiles "quartile_enteric_lifetime", "quartile_fluid_metabolites_lifetime", 
 "quartile_hormones_and_related_lifetime", 
             "quartile_immunologicals_lifetime", "quartile_misc_lifetime", "quartile_neurological_lifetime", 
             "quartile_anti_parasites_lifetime", 
 #not enough data for quartiles "quartile_anti_septics_lifetime", "quartile_cardio_respiratory_lifetime",   "quartile_diuretics_lifetime", "quartile_anti_hist_lifetime", 
             "half_anti_inflam_5y_prior_endpoint", "half_anti_microbials_5y_prior_endpoint", "half_enteric_5y_prior_endpoint", 
             "half_immunologicals_5y_prior_endpoint", "half_misc_5y_prior_endpoint", "half_neurological_5y_prior_endpoint", 
             "half_anti_parasites_5y_prior_endpoint", "half_dietary_supplements_5y_prior_endpoint", 
             "half_anti_septics_5y_prior_endpoint", "half_hormones_and_related_5y_prior_endpoint", 
             "half_cardio_respiratory_5y_prior_endpoint", "half_fluid_metabolites_5y_prior_endpoint", 
             "half_diuretics_5y_prior_endpoint", "half_anti_hist_5y_prior_endpoint", 
             "quartile_anti_inflam_5y_prior_endpoint", "quartile_anti_microbials_5y_prior_endpoint", 
            # "quartile_enteric_5y_prior_endpoint",
 "quartile_immunologicals_5y_prior_endpoint", 
             "quartile_misc_5y_prior_endpoint", "quartile_neurological_5y_prior_endpoint", 
             "quartile_anti_parasites_5y_prior_endpoint", "quartile_dietary_supplements_5y_prior_endpoint", 
       #      "quartile_anti_septics_5y_prior_endpoint", 
 "quartile_hormones_and_related_5y_prior_endpoint", 
          #   "quartile_cardio_respiratory_5y_prior_endpoint", #"quartile_fluid_metabolites_5y_prior_endpoint", 
         #    "quartile_diuretics_5y_prior_endpoint", "quartile_anti_hist_5y_prior_endpoint", 
             #"quartile_anti_viral_5y_prior_endpoint",
 "age_at_final_date_quartiles","age_at_neuter_years_quartiles")
dependent = "status_factor" # had HSA or no
results <- cohort_abr %>%
  summary_factorlist(dependent, explanatory,
  p=TRUE, add_dependent_label=TRUE) -> t1
knitr::kable(t1, row.names=FALSE, align=c("l", "l", "r", "r", "r"))

results_df <- as.data.frame(results)

# Write the results to CSV
write.csv(results_df, "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_cross_table_case_control_output.csv", row.names = FALSE)

#save as worddoc
# Install and load the required packages
library(officer)
library(flextable)

# Create a Word document
doc <- read_docx()

# Convert finalfit output to a flextable
results_flextable <- flextable(results_df)

# Add the flextable to the Word document
doc <- doc %>%
  body_add_flextable(results_flextable) %>%
  body_add_par("Finalfit Regression Results", style = "heading 1")

# Save the Word document
print(doc, target = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_cross_table_case_control_output.docx")
```

## Multivariable Cox PH

```{r}
lrt_pvalues <-read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_time_to_HSA_cox_ph_univar.csv")
multivar_variables <- lrt_pvalues %>%
  filter(lrt_pvalue <= 0.2) %>%
  distinct(variable)
cols_to_keep = multivar_variables$variable
cols_to_keep <- append(cols_to_keep, "status")
cols_to_keep <- append(cols_to_keep, "survival_time_days")
cols_to_keep <- append(cols_to_keep, "status_factor")
cohort_abr_sig <- cohort_abr[,cols_to_keep]

lrt_pvalues$pval <- sprintf("%.20f", lrt_pvalues$lrt_pvalue)
```

Variables to take forward (p\<0.2):

```{r}
print(multivar_variables)
```

Assess which variables are correlated:

```{r,error=TRUE}
source("C:/Users/ctaylor18/GitHub/HSA-VC-enviro-analyses/functions/corr_matrix_creation.R")
#use sig variables list generated above

create_corr_matrix(cohort_abr_sig,c(1:70), "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_survival.csv", "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_survival2.csv","C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_survival_corrvalue.csv","C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_survival.png")
```

### Correlated variables:

```{r}
corr_values <- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_survival_corrvalue.csv")

#ditch rows where variable 1 and 2 values are same as in another row :
corr_values2 <- corr_values %>%
  dplyr::select(-c(X))

corr_values3 <- corr_values2 %>%
  mutate(Var1_sorted = pmin(Variable1, Variable2),  
         Var2_sorted = pmax(Variable1, Variable2)) %>%  
  distinct(Var1_sorted, Var2_sorted, Correlation, .keep_all = TRUE) %>%  
  dplyr::select(-Var1_sorted, -Var2_sorted)  # Remove helper columns


print(corr_values3)
```

-   Home enviro

    -   Sleep location and primary heating source

    -   early vs rest vs whole life for variables like incense, candles, moth balls etc

-   Dog health

    -   comorbs lifetime vs 5y

    -   half vs quartiles for drug admin

    -   lifetime vs 5y prior for drug admin

    -   anti viral and fluid halves 5y prior endpoint

-   Activity/lifestyle

    -   activity intensity vs frequency

    -   lifestyle category and sleep location

-   Repro

    -   sex and no preg, heats before neutering, age at neutering

-   Comorbs

    -   Ortho 5y prior and all orthopaedic

    -   GI 5y and all GI

Multivar model building:

-   Stepwise with most sig initially

```{r}
source("C:/Users/ctaylor18/GitHub/HSA-VC-survival-analyses/HSA-VC-survival-analyses/code/functions/multivar_cox_model_analyse_AIC_funct.R")

#null model
cox_model_null <- coxph(Surv(survival_time_days, status) ~1
, data = cohort_abr_sig)
summary(cox_model_null)
```

### Weight quartiles 5y prior

```{r}
cox_model_2<- coxph(Surv(survival_time_days, status) ~  
                      #age_at_final_date_quartiles
                       avg_weight_5y_prior_endpoint_quartiles
, data = cohort_abr_sig)



analyse_model(cox_model_2,cohort_abr_sig,cox_model_null)

```

### Purina BCS 5y prior

```{r}
cox_model_3<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles
, data = cohort_abr_sig)

vif(cox_model_3)


analyse_model(cox_model_3,cohort_abr_sig,cox_model_2)

explanatory <- c("avg_weight_5y_prior_endpoint_quartiles", "avg_purina_BCS_5y_prior_endpoint_quartiles")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

### Activity

-   rest of life terciles

    -   frequency or intensity

-   frequency best fit via AIC

-   rest of life(SY3+) vs 5y prior

    -   SY3+ improved model parsimonious-ness with BIC \~100 less than 5y prior mdoels

```{r}
cox_model_4a<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles
, data = cohort_abr_sig)

vif(cox_model_4a)


analyse_model(cox_model_4a,cohort_abr_sig,cox_model_3)

cox_model_4b<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_intensity_sy3_quartiles
, data = cohort_abr_sig)

vif(cox_model_4b)


analyse_model(cox_model_4b,cohort_abr_sig,cox_model_3)

#5y prior rather than SY3+
cox_model_4c<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles
, data = cohort_abr_sig)
analyse_model(cox_model_4c,cohort_abr_sig,cox_model_3)

cox_model_4d<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_intensity_5y_prior_endpoint_quartiles
, data = cohort_abr_sig)
analyse_model(cox_model_4d,cohort_abr_sig,cox_model_3)
```

### Height

-   doesnt improve model

```{r}
cox_model_5<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles + avg_height_5y_prior_endpoint_quartiles
, data = cohort_abr_sig)

vif(cox_model_5)


analyse_model(cox_model_5,cohort_abr_sig,cox_model_4a)
```

### Parasite 5y prior quartiles

```{r}
cox_model_6<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles + quartile_anti_parasites_5y_prior_endpoint
, data = cohort_abr_sig)

vif(cox_model_6)


analyse_model(cox_model_6,cohort_abr_sig,cox_model_4a)

explanatory <- c("avg_weight_5y_prior_endpoint_quartiles", "avg_purina_BCS_5y_prior_endpoint_quartiles","avg_activity_freq_sy3_terciles","quartile_anti_parasites_5y_prior_endpoint")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

### In the house rest of life main sleep location

-   not include

```{r}
cox_model_7<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles + quartile_anti_parasites_5y_prior_endpoint+in.the.house_YN_study_years_rest_of_life
, data = cohort_abr_sig)

vif(cox_model_7)


analyse_model(cox_model_7,cohort_abr_sig,cox_model_6)


```

### Main lifestyle category 5y prior

-   approaching VIF of 5 so assessed if activity and lifestyle which one was better for model - activity

```{r}
cox_model_8<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles + quartile_anti_parasites_5y_prior_endpoint+main_lifestyle_category_mode_5yrs_prior
, data = cohort_abr_sig)

vif(cox_model_8)


analyse_model(cox_model_8,cohort_abr_sig,cox_model_6)

explanatory <- c("avg_weight_5y_prior_endpoint_quartiles", "avg_purina_BCS_5y_prior_endpoint_quartiles","avg_activity_freq_sy3_terciles","quartile_anti_parasites_5y_prior_endpoint", "main_lifestyle_category_mode_5yrs_prior")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod

#assess if activity freq SY3 tercile or main lifestyle better model (as VIF is 4.5 and 3.5 for these)
cox_model_8a<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles +  quartile_anti_parasites_5y_prior_endpoint+main_lifestyle_category_mode_5yrs_prior
, data = cohort_abr_sig)

vif(cox_model_8a)

cox_model_8b<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles +  quartile_anti_parasites_5y_prior_endpoint+ avg_activity_freq_sy3_terciles
, data = cohort_abr_sig)

vif(cox_model_8b)
analyse_model(cox_model_8a,cohort_abr_sig,cox_model_8b)

```

### Pipes plastic 5y prior

-   does not improve model

```{r}
cox_model_9<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles + quartile_anti_parasites_5y_prior_endpoint+pipes_plastic_any_mode_5yrs_prior
, data = cohort_abr_sig)

vif(cox_model_9)


analyse_model(cox_model_9,cohort_abr_sig,cox_model_6)


```

### House type 5y prior

-   does not improve model

```{r}
cox_model_10<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles + quartile_anti_parasites_5y_prior_endpoint+house_type_mode_5yrs_prior
, data = cohort_abr_sig)

vif(cox_model_10)


analyse_model(cox_model_10,cohort_abr_sig,cox_model_6)
```

### Sleep location 5y prior

-   does not improve model

```{r}
cox_model_11<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles + quartile_anti_parasites_5y_prior_endpoint+sleep_location_mode_5yrs_prior
, data = cohort_abr_sig)

vif(cox_model_11)


analyse_model(cox_model_11,cohort_abr_sig,cox_model_6)
```

### Primary cooking fuel

-   not significant

```{r}
cox_model_12<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles + quartile_anti_parasites_5y_prior_endpoint+cooking_fuel_primary_mode_5yrs_prior
, data = cohort_abr_sig)

vif(cox_model_12)


analyse_model(cox_model_12,cohort_abr_sig,cox_model_6)
```

### Water source

```{r}
cox_model_13<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles + quartile_anti_parasites_5y_prior_endpoint+water_source_mode_5yrs_prior
, data = cohort_abr_sig)

vif(cox_model_13)


analyse_model(cox_model_13,cohort_abr_sig,cox_model_6)
```

### Sex

-   does improve model

```{r}
cox_model_14<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles + quartile_anti_parasites_5y_prior_endpoint+sex
, data = cohort_abr_sig)

vif(cox_model_14)


analyse_model(cox_model_14,cohort_abr_sig,cox_model_6)

explanatory <- c("avg_weight_5y_prior_endpoint_quartiles", "avg_purina_BCS_5y_prior_endpoint_quartiles","avg_activity_freq_sy3_terciles","quartile_anti_parasites_5y_prior_endpoint", "sex")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

### No pregnancy ever

-   sex and no_preg ever are collinear

-   assess which is better - equivocal so keep sex in model

```{r}
cox_model_15<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles + quartile_anti_parasites_5y_prior_endpoint+sex+no_pregnancy_ever
, data = cohort_abr_sig)

vif(cox_model_15)


analyse_model(cox_model_15,cohort_abr_sig,cox_model_14)

cox_model_15a<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles + quartile_anti_parasites_5y_prior_endpoint+no_pregnancy_ever
, data = cohort_abr_sig)




vif(cox_model_15a)
analyse_model(cox_model_15a,cohort_abr_sig,cox_model_14)

```

### Area type

-   does not improve model

```{r}
cox_model_16<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles + quartile_anti_parasites_5y_prior_endpoint+sex
+ area_type_mode_5yrs_prior, data = cohort_abr_sig)

vif(cox_model_16)


analyse_model(cox_model_16,cohort_abr_sig,cox_model_14)
```

### Primary heating fuel

-   does not improve model

```{r}
cox_model_17<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles + quartile_anti_parasites_5y_prior_endpoint+sex
+ heating_fuel_primary_mode_5yrs_prior, data = cohort_abr_sig)

vif(cox_model_17)


analyse_model(cox_model_17,cohort_abr_sig,cox_model_14)
```

### Heats before neutering terciles

-   cant have sex and heats before neutering in same model as collinear

-   keep sex

```{r,error=TRUE}
cox_model_18<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles + quartile_anti_parasites_5y_prior_endpoint+sex
+ heats_before_neutering_terciles, data = cohort_abr_sig)

#vif(cox_model_18)
cox_model_18a<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles + quartile_anti_parasites_5y_prior_endpoint
+ heats_before_neutering_terciles, data = cohort_abr_sig)

#vif(cox_model_18a)

cox_model_18b<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles + quartile_anti_parasites_5y_prior_endpoint
, data = cohort_abr_sig)


analyse_model(cox_model_18a,cohort_abr_sig,cox_model_18b)
```

### Use of incense whole life

```{r}
cox_model_19<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles + quartile_anti_parasites_5y_prior_endpoint+sex
+ use_incense_or_candles_study_years_whole_life, data = cohort_abr_sig)

vif(cox_model_19)


analyse_model(cox_model_19,cohort_abr_sig,cox_model_14)
```

### Primarily outside whole life/rest of life

-   not improving

```{r}
cox_model_20<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles + quartile_anti_parasites_5y_prior_endpoint+sex
+ use_incense_or_candles_study_years_whole_life+outside_YN_study_years_rest_of_life, data = cohort_abr_sig)

vif(cox_model_20)


analyse_model(cox_model_20,cohort_abr_sig,cox_model_19)

cox_model_20a<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles + quartile_anti_parasites_5y_prior_endpoint+sex
+ use_incense_or_candles_study_years_whole_life+outside_YN_study_years_whole_life, data = cohort_abr_sig)

vif(cox_model_20a)


analyse_model(cox_model_20a,cohort_abr_sig,cox_model_19)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles", "avg_purina_BCS_5y_prior_endpoint_quartiles","avg_activity_freq_sy3_terciles","quartile_anti_parasites_5y_prior_endpoint", "sex", "use_incense_or_candles_study_years_whole_life","outside_YN_study_years_whole_life")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

### Hepa filter use rest of life

-   does not improve model

```{r}
cox_model_21<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles + quartile_anti_parasites_5y_prior_endpoint+sex
+ use_incense_or_candles_study_years_whole_life +use_hepa_filter_study_years_whole_life, data = cohort_abr_sig)

vif(cox_model_21)


analyse_model(cox_model_21,cohort_abr_sig,cox_model_19)
```

### Age at neuter quartiles

-   does not improve model

```{r}
cox_model_22<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles + quartile_anti_parasites_5y_prior_endpoint+sex
+ use_incense_or_candles_study_years_whole_life +age_at_neuter_years_quartiles, data = cohort_abr_sig)

vif(cox_model_22)


analyse_model(cox_model_22,cohort_abr_sig,cox_model_19)
```

### Early life activity duration quartiles

```{r}
cox_model_23<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles + quartile_anti_parasites_5y_prior_endpoint+sex
+ use_incense_or_candles_study_years_whole_life +avg_activity_duration_sy1_2_quartiles, data = cohort_abr_sig)

vif(cox_model_23)


analyse_model(cox_model_23,cohort_abr_sig,cox_model_19)

explanatory <- c("avg_weight_5y_prior_endpoint_quartiles", "avg_purina_BCS_5y_prior_endpoint_quartiles","avg_activity_freq_sy3_terciles","quartile_anti_parasites_5y_prior_endpoint", "sex", "use_incense_or_candles_study_years_whole_life","avg_activity_duration_sy1_2_quartiles")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

### Outside early life

```{r}

cox_model_24<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles + quartile_anti_parasites_5y_prior_endpoint+sex
+ use_incense_or_candles_study_years_whole_life +avg_activity_duration_sy1_2_quartiles +outside_YN_study_years_early_life, data = cohort_abr_sig)

vif(cox_model_24)


analyse_model(cox_model_24,cohort_abr_sig,cox_model_23)

explanatory <- c("avg_weight_5y_prior_endpoint_quartiles", "avg_purina_BCS_5y_prior_endpoint_quartiles","avg_activity_freq_sy3_terciles","quartile_anti_parasites_5y_prior_endpoint", "sex", "use_incense_or_candles_study_years_whole_life","avg_activity_duration_sy1_2_quartiles","outside_YN_study_years_early_life")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

### In house early and whole life terciles

-   non sig

```{r}
#early life
cox_model_25a<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex +avg_activity_duration_sy1_2_quartiles +in.the.house_YN_sum_early_life_terciles, data = cohort_abr_sig)

vif(cox_model_25a)

analyse_model(cox_model_25a,cohort_abr_sig,cox_model_23)
#whole life
cox_model_25b<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +in.the.house_YN_sum_whole_life_quartiles, data = cohort_abr_sig)

vif(cox_model_25b)
analyse_model(cox_model_25b,cohort_abr_sig,cox_model_23)

```

### Neutering reason

```{r}
cox_model_26<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles+ neutered_reason, data = cohort_abr_sig)

vif(cox_model_26)
cox_model_23<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles, data = cohort_abr_sig)

#analyse_model(cox_model_26,cohort_abr_sig,cox_model_23)

explanatory <- c("avg_weight_5y_prior_endpoint_quartiles", "avg_purina_BCS_5y_prior_endpoint_quartiles","avg_frequency_5y_prior_endpoint_quartiles","quartile_anti_parasites_5y_prior_endpoint", "sex", "avg_activity_duration_sy1_2_quartiles","neutered_reason")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

### Secondary cooking fuel

```{r}
cox_model_27<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason + cooking_fuel_secondary_mode_5yrs_prior, data = cohort_abr_sig)

vif(cox_model_27)
analyse_model(cox_model_27,cohort_abr_sig,cox_model_26)
explanatory <- c("avg_weight_5y_prior_endpoint_quartiles", "avg_purina_BCS_5y_prior_endpoint_quartiles","avg_frequency_5y_prior_endpoint_quartiles","quartile_anti_parasites_5y_prior_endpoint", "sex", "use_incense_or_candles_study_years_whole_life","avg_activity_duration_sy1_2_quartiles","neutered_reason","cooking_fuel_secondary_mode_5yrs_prior")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

### Neurological drugs lifetime half

```{r}
cox_model_28<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason +half_neurological_lifetime, data = cohort_abr_sig)

#vif(cox_model_28)


analyse_model(cox_model_28,cohort_abr_sig,cox_model_26)

explanatory <- c("avg_weight_5y_prior_endpoint_quartiles", "avg_purina_BCS_5y_prior_endpoint_quartiles","avg_frequency_5y_prior_endpoint_quartiles","quartile_anti_parasites_5y_prior_endpoint", "sex", "avg_activity_duration_sy1_2_quartiles","neutered_reason","half_neurological_lifetime")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

### Enteric meds lifetime half or 5y prior half

```{r}
cox_model_29<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason+half_enteric_lifetime, data = cohort_abr_sig)

#vif(cox_model_29)


analyse_model(cox_model_29,cohort_abr_sig,cox_model_26)

cox_model_29a<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason +half_enteric_5y_prior_endpoint, data = cohort_abr_sig)

#vif(cox_model_29a)


analyse_model(cox_model_29a,cohort_abr_sig,cox_model_26)
```

### Immune mediated

```{r}
cox_model_30 <- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason + immune_mediated, data = cohort_abr_sig)

vif(cox_model_30)


analyse_model(cox_model_30,cohort_abr_sig,cox_model_26)
```

### lived in garage early life or whole life

```{r}
cox_model_31<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason+
in.the.garage_YN_study_years_early_life, data = cohort_abr_sig)

analyse_model(cox_model_31,cohort_abr_sig,cox_model_26)
cox_model_31a<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason+
in.the.garage_YN_study_years_whole_life, data = cohort_abr_sig)

analyse_model(cox_model_31a,cohort_abr_sig,cox_model_26)

```

### anti septic half lifetime

```{r}
cox_model_32<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex +avg_activity_duration_sy1_2_quartiles +neutered_reason+
half_anti_septics_lifetime, data = cohort_abr_sig)

#vif(cox_model_32)


analyse_model(cox_model_32,cohort_abr_sig,cox_model_26)
```

### half anti inflams 5y prior to endpoint

```{r}
cox_model_33<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason +
half_anti_inflam_5y_prior_endpoint, data = cohort_abr_sig)

#vif(cox_model_33)


analyse_model(cox_model_33,cohort_abr_sig,cox_model_26)

explanatory <- c("avg_weight_5y_prior_endpoint_quartiles", "avg_purina_BCS_5y_prior_endpoint_quartiles","avg_frequency_5y_prior_endpoint_quartiles","quartile_anti_parasites_5y_prior_endpoint", "sex", "use_incense_or_candles_study_years_whole_life","avg_activity_duration_sy1_2_quartiles","neutered_reason","half_anti_inflam_5y_prior_endpoint")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

### incense rest of life

```{r}
cox_model_34<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason+
use_incense_or_candles_study_years_rest_of_life, data = cohort_abr_sig)

vif(cox_model_34)
analyse_model(cox_model_34,cohort_abr_sig,cox_model_26)

cox_model_34<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason+
use_incense_or_candles_study_years_whole_life, data = cohort_abr_sig)

vif(cox_model_34)

analyse_model(cox_model_34,cohort_abr_sig,cox_model_26)

```

### aerosol whole life

```{r}
cox_model_35<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason+
use_incense_or_candles_study_years_whole_life+
use_aerosol_study_years_whole_life, data = cohort_abr_sig)

vif(cox_model_35)
analyse_model(cox_model_35,cohort_abr_sig,cox_model_34)

```

### smoke exposure early life

```{r}
cox_model_36<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason+
use_incense_or_candles_study_years_whole_life+
smoke_exposure_study_years_early_life, data = cohort_abr_sig)

vif(cox_model_36)
analyse_model(cox_model_36,cohort_abr_sig,cox_model_34)

```

### moth balls rest of life

```{r}
cox_model_37<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason+
use_incense_or_candles_study_years_whole_life +
use_moth_balls_study_years_rest_of_life, data = cohort_abr_sig)

vif(cox_model_37)
analyse_model(cox_model_37,cohort_abr_sig,cox_model_34)

```

### half cardio resp lifetime

```{r}
cox_model_38<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason+
use_incense_or_candles_study_years_whole_life +
half_cardio_respiratory_lifetime, data = cohort_abr_sig)

#vif(cox_model_38)


analyse_model(cox_model_38,cohort_abr_sig,cox_model_34)
```

### half fluid metabolites lifetime

```{r}
cox_model_39<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason+
use_incense_or_candles_study_years_whole_life+
half_fluid_metabolites_5y_prior_endpoint
, data = cohort_abr_sig)

#vif(cox_model_39)


analyse_model(cox_model_39,cohort_abr_sig,cox_model_34)
```

### Gastrointestinal comorbidities- lifelong and 5y and all cause

-   all sig in model and all equivocal AIC/BIC

```{r}
#lifelong
cox_model_40a<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason+
use_incense_or_candles_study_years_whole_life+
half_fluid_metabolites_5y_prior_endpoint + lifelong_gastrointestinal
, data = cohort_abr_sig)

#vif(cox_model_40a)


analyse_model(cox_model_40a,cohort_abr_sig,cox_model_34)
#5y prior
cox_model_40b<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason+
use_incense_or_candles_study_years_whole_life+
half_fluid_metabolites_5y_prior_endpoint + gastrointestinal
, data = cohort_abr_sig)

#vif(cox_model_40b)


analyse_model(cox_model_40b,cohort_abr_sig,cox_model_34)

#all cause
cox_model_40c<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason+
use_incense_or_candles_study_years_whole_life+
half_fluid_metabolites_5y_prior_endpoint + comorb_gastrointestinal_all

, data = cohort_abr_sig)

#vif(cox_model_40b)


analyse_model(cox_model_40c,cohort_abr_sig,cox_model_34)
```

### Inflammatory other comorbdities 5y

-   non sig

```{r}
cox_model_41<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason+
use_incense_or_candles_study_years_whole_life+
half_fluid_metabolites_5y_prior_endpoint + lifelong_gastrointestinal + inflammatory_other
, data = cohort_abr_sig)

#vif(cox_model_40a)


analyse_model(cox_model_41,cohort_abr_sig,cox_model_40a)
```

### Orthopaedic comorbidities - 5y

-   non sig

```{r}
#5y
cox_model_42a<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason+
use_incense_or_candles_study_years_whole_life+
half_fluid_metabolites_5y_prior_endpoint + lifelong_gastrointestinal + orthopaedic
, data = cohort_abr_sig)

#vif(cox_model_40a)


analyse_model(cox_model_42a,cohort_abr_sig,cox_model_40a)


```

### Immune mediated 5y

-   non sig

```{r}
cox_model_43<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason+
use_incense_or_candles_study_years_whole_life+
half_fluid_metabolites_5y_prior_endpoint + lifelong_gastrointestinal + immune_mediated
, data = cohort_abr_sig)

#vif(cox_model_40a)


analyse_model(cox_model_43,cohort_abr_sig,cox_model_40a)
```

### Chronic inflam present 5y

-   non sig

```{r}
cox_model_44<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason+
use_incense_or_candles_study_years_whole_life+
half_fluid_metabolites_5y_prior_endpoint + lifelong_gastrointestinal + chronic_inflam_present

, data = cohort_abr_sig)

#vif(cox_model_40a)


analyse_model(cox_model_44,cohort_abr_sig,cox_model_40a)
```

```{r}
explanatory <- c("avg_weight_5y_prior_endpoint_quartiles", "avg_purina_BCS_5y_prior_endpoint_quartiles","avg_frequency_5y_prior_endpoint_quartiles","quartile_anti_parasites_5y_prior_endpoint", "sex", "use_incense_or_candles_study_years_whole_life","avg_activity_duration_sy1_2_quartiles","neutered_reason","lifelong_gastrointestinal")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Remove neuter and activity duration and BCS and incense and lifelong gastrointestinal as nonsig in final model

```{r}
explanatory <- c("avg_weight_5y_prior_endpoint_quartiles", "avg_frequency_5y_prior_endpoint_quartiles","quartile_anti_parasites_5y_prior_endpoint", "sex")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Should weight and BCS be in same model

```{r}
cox_model_no_BCS<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles +  avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason+
use_incense_or_candles_study_years_whole_life
, data = cohort_abr_sig)

vif(cox_model_no_BCS)
cox_model_weight_and_BCS<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason+
use_incense_or_candles_study_years_whole_life
, data = cohort_abr_sig)

vif(cox_model_weight_and_BCS)

analyse_model(cox_model_weight_and_BCS,cohort_abr_sig,cox_model_no_BCS)


cox_model_no_weight<- coxph(Surv(survival_time_days, status) ~    avg_purina_BCS_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex
 +avg_activity_duration_sy1_2_quartiles +neutered_reason+
use_incense_or_candles_study_years_whole_life
, data = cohort_abr_sig)

vif(cox_model_no_weight)


analyse_model(cox_model_weight_and_BCS,cohort_abr_sig,cox_model_no_weight)

```

Final model output:

```{r}
# Exclude frailty from the explanatory variables for finalfit output
dependent = 'Surv(survival_time_days,status)'

explanatory_fixed = c("avg_weight_5y_prior_endpoint_quartiles", "avg_frequency_5y_prior_endpoint_quartiles","quartile_anti_parasites_5y_prior_endpoint", "sex")

# Generate the finalfit table for fixed effects only
HSA_multivar_cox_ph_table <- cohort_abr_sig %>% 
  finalfit(dependent, explanatory_fixed)

# Print the table
print(knitr::kable(HSA_multivar_cox_ph_table, row.names=FALSE, align=c("l", "l", "r", "r", "r")))
```

```{r}
cox_model_f<- coxph(Surv(survival_time_days, status) ~  avg_weight_5y_prior_endpoint_quartiles + avg_frequency_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint+sex


, data = cohort_abr_sig)

library(gtsummary)
tbl_regression(cox_model_f, exponentiate = TRUE)
```

```{r}
ggforest(cox_model_f, data = cohort_abr_sig, fontsize = 1) + ggtitle("Factors associated with time to HSA diagnosis")

```

### Post hoc assessment

-   majority of variables violate PH assumption

-   those that do - weight, activity, parasites, sex, neutering reason

-   those that do not - BCS, use of incense, activity duration avg, whether outside or not for sleeping in early years

```{r}
dependent = "Surv(survival_time_days, status)"
explanatory = c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_frequency_5y_prior_endpoint_quartiles","quartile_anti_parasites_5y_prior_endpoint", "sex"
                )

ph_schoen <- cohort_abr_sig %>% 
    coxphmulti(dependent, explanatory) %>% 
    cox.zph() %>% 
    {zph_result <<- .} %>% 
    plot(var=4)

zph_result

```

Manually assessing with KM plots

Weight

```{r}

dependent_os = "Surv(survival_time_months, status)"
explanatory = c("avg_weight_5y_prior_endpoint_quartiles")

cohort_abr %>% 
    surv_plot(dependent_os, explanatory, pval = TRUE)
```

Frequency exercise

```{r}


dependent_os = "Surv(survival_time_months, status)"
explanatory = c("avg_frequency_5y_prior_endpoint_quartiles")

cohort_abr %>% 
    surv_plot(dependent_os, explanatory, pval = TRUE)
```

Parasite tx

```{r}


dependent_os = "Surv(survival_time_months, status)"
explanatory = c("quartile_anti_parasites_5y_prior_endpoint")

cohort_abr %>% 
    surv_plot(dependent_os, explanatory, pval = TRUE)
```

Sex

```{r}


dependent_os = "Surv(survival_time_months, status)"
explanatory = c("sex")

cohort_abr %>% 
    surv_plot(dependent_os, explanatory, pval = TRUE)
```

## Second version analyses

-   only taking forward variables where they are p\<0.2 when "not_recorded is excluded

```{r}
#swap not_recorded for NA and exclude from univar this way
cohort_no_not_rec <- cohort_no_not_rec%>%
  mutate(across(where(~ is.character(.) | is.factor(.)), ~ na_if(as.character(.), "not_recorded")))


```

Create just chi-sq table without NA included as a not_recorded variable

```{r}

check_levels <- function(df) {
  sapply(df, function(x) length(unique(na.omit(x))) >= 2)
}

# Example usage
result <- check_levels(cohort_no_not_rec)

# Print columns that do not meet the condition
names(result[result == FALSE])


source('C:/Users/ctaylor18/GitHub/HSA-VC-survival-analyses/HSA-VC-survival-analyses/code/functions/univar_cox_ph_table_funct.R')

cohort_no_not_rec$status_factor <- as.factor(as.character(cohort_no_not_rec$had_HSA))
cohort_no_not_rec$status <-as.numeric(cohort_no_not_rec$had_HSA) 
cohort_no_not_rec$survival_time_months <- as.numeric(cohort_no_not_rec$survival_time_months)

cohort_no_not_rec$survival_time_days <- as.numeric(cohort_no_not_rec$survival_time_months)




explanatory = c("MDI_quintile","owner_MDI_quintile","area_type_mode_5yrs_prior","house_type_mode_5yrs_prior", 
             "water_source_mode_5yrs_prior", "pipes_metal_any_mode_5yrs_prior", "pipes_plastic_any_mode_5yrs_prior", 
             "heating_fuel_primary_mode_5yrs_prior", "cooking_fuel_primary_mode_5yrs_prior", "heating_fuel_secondary_mode_5yrs_prior", 
             "cooking_fuel_secondary_mode_5yrs_prior", "region_mode_5y_prior",  #"number_comorbs_present_lifetime", 
            # "study_year_first_chronic_inflam", #"chronic_inflam_present_5y", #"number_comorbs_present_5y", 
            # "chronic_inflam_present_no_min_year", "number_comorbs_present_no_min_year",  "diagnosis_after_comorb",  "time_between_first_comorb_and_endpoint",  
              "num_lifetime_chronic_inflam", "intestinal_parasitism","tick_borne_parasitism","other_parasitism","all_parasitism","infectious","orthopaedic","immune_mediated","cardiovascular","gastrointestinal","inflammatory_other","chronic_inflammatory",
 "USDA_30min_sy1_2", "activity_1h_min_sy1_2", 
             "activity_KC_2h_min_sy1_2", "avg_activity_duration_sy1_2_quartiles",#"avg_activity_intensity_sy1_2_terciles",
 "avg_activity_freq_sy1_2_terciles", 
             "avg_activity_freq_sy3_terciles", "avg_activity_intensity_sy3_quartiles", "avg_intensity_5y_prior_endpoint_quartiles", "avg_frequency_5y_prior_endpoint_quartiles", 
             "use_aerosol_study_years_early_life", "use_air_cleaner_study_years_early_life", "use_hepa_filter_study_years_early_life", 
             "use_moth_balls_study_years_early_life", "use_incense_or_candles_study_years_early_life", "smoke_exposure_study_years_early_life", 
       #      "any_treated_weeds_study_years_early_life", "any_treated_insects_study_years_early_life", #"any_treated_fertilizer_study_years_early_life", 
         #    "use_aerosol_sum_early_life", "use_air_cleaner_sum_early_life", "use_hepa_filter_sum_early_life", "use_moth_balls_sum_early_life", 
          #   "use_incense_or_candles_sum_early_life", 
 #"smoke_exposure_sum_early_life",
 #"any_treated_weeds_sum_early_life", 
    #         "any_treated_insects_sum_early_life", #"any_treated_fertilizer_sum_early_life", 
 "use_aerosol_study_years_rest_of_life", 
             "use_air_cleaner_study_years_rest_of_life", "use_hepa_filter_study_years_rest_of_life", "use_moth_balls_study_years_rest_of_life", 
             "use_incense_or_candles_study_years_rest_of_life", "smoke_exposure_study_years_rest_of_life", "any_treated_weeds_study_years_rest_of_life", 
             "any_treated_insects_study_years_rest_of_life", "any_treated_fertilizer_study_years_rest_of_life", #"use_aerosol_sum_rest_of_life", 
 #            "use_air_cleaner_sum_rest_of_life", "use_hepa_filter_sum_rest_of_life", "use_moth_balls_sum_rest_of_life", 
          #   "use_incense_or_candles_sum_rest_of_life", 
# "smoke_exposure_sum_rest_of_life", 
# "any_treated_weeds_sum_rest_of_life", 
            # "any_treated_insects_sum_rest_of_life", "any_treated_fertilizer_sum_rest_of_life", 
"use_aerosol_study_years_whole_life", 
             "use_air_cleaner_study_years_whole_life", "use_hepa_filter_study_years_whole_life", "use_moth_balls_study_years_whole_life", 
             "use_incense_or_candles_study_years_whole_life", "smoke_exposure_study_years_whole_life", "any_treated_weeds_study_years_whole_life", 
             "any_treated_insects_study_years_whole_life", "any_treated_fertilizer_study_years_whole_life",# "use_aerosol_sum_whole_life", 
           #  "use_air_cleaner_sum_whole_life", "use_hepa_filter_sum_whole_life", "use_moth_balls_sum_whole_life",# "use_incense_or_candles_sum_whole_life", 
            # "smoke_exposure_sum_whole_life", 
 #"any_treated_weeds_sum_whole_life", "any_treated_insects_sum_whole_life", 
            # "any_treated_fertilizer_sum_whole_life",
 "hours_of_smoke_early_life_total_dosage_terciles", "hours_of_smoke_rest_of_life_total_dosage_terciles", 
             "hours_of_smoke_whole_life_total_dosage_terciles", "in.the.garage_YN_study_years_early_life", "in.the.house_YN_study_years_early_life", 
             "outside_YN_study_years_early_life",  "in.the.house_YN_sum_early_life_terciles", 
             "in.the.garage_YN_study_years_rest_of_life", "in.the.house_YN_study_years_rest_of_life", "outside_YN_study_years_rest_of_life", 
              "in.the.house_YN_sum_rest_of_life_quartiles",  
             "in.the.garage_YN_study_years_whole_life", "in.the.house_YN_study_years_whole_life", "outside_YN_study_years_whole_life", 
             "in.the.house_YN_sum_whole_life_quartiles",  
             "majority_sleep_location_lifetime",  "main_lifestyle_category", 
 "main_lifestyle_category_mode_5yrs_prior", "sleep_location_mode_5yrs_prior", 
             "use_aerosol_exposed_5yrs_prior", "use_air_cleaner_exposed_5yrs_prior", "use_hepa_filter_exposed_5yrs_prior", 
             "use_moth_balls_exposed_5yrs_prior", "use_incense_or_candles_exposed_5yrs_prior", "smoke_exposure_exposed_5yrs_prior", 
             "any_treated_weeds_exposed_5yrs_prior", "any_treated_insects_exposed_5yrs_prior", "any_treated_fertilizer_exposed_5yrs_prior", 
             "hours_of_smoke_exposure_hours_5yrs_prior_terciles", "sex", "neuter", "mismating_ever",
             "neutered_reason", "neutered_ever_bred", "spayed_while_in_heat", "spayed_while_pregnant", "prevent_heat_ever", 
             "no_pregnancy_ever", "intact_sired_litters_natural_ever", "neutered_females_bred_with_count_ever", 
             "neutered_sired_litters_natural_ever", "heats_before_neutering", "heats_before_neutering_terciles",
             "preg_pre_neuter",# "mated_pre_neuter",
 "num_breeding_before_neutering_terciles", "ever_bred_from", 
              "avg_weight_5y_prior_endpoint_quartiles", "avg_height_5y_prior_endpoint_quartiles", 
             "avg_purina_BCS_5y_prior_endpoint_quartiles", "mode_purina_BCS_5y_prior_endpoint", "mode_BCS_classify", "half_anti_microbials_lifetime", "half_dietary_supplements_lifetime", 
             "half_enteric_lifetime", "half_fluid_metabolites_lifetime", "half_hormones_and_related_lifetime", "half_immunologicals_lifetime", 
             "half_misc_lifetime", "half_neurological_lifetime", "half_anti_parasites_lifetime", "half_anti_septics_lifetime", 
             "half_cardio_respiratory_lifetime", "half_diuretics_lifetime", "half_anti_hist_lifetime", 
             "quartile_anti_inflam_lifetime", "quartile_anti_microbials_lifetime", "quartile_dietary_supplements_lifetime", 
            #not enough data for quartiles "quartile_enteric_lifetime", "quartile_fluid_metabolites_lifetime", 
 "quartile_hormones_and_related_lifetime", 
             "quartile_immunologicals_lifetime", "quartile_misc_lifetime", "quartile_neurological_lifetime", 
             "quartile_anti_parasites_lifetime", 
 #not enough data for quartiles "quartile_anti_septics_lifetime", "quartile_cardio_respiratory_lifetime",   "quartile_diuretics_lifetime", "quartile_anti_hist_lifetime", 
             "half_anti_inflam_5y_prior_endpoint", "half_anti_microbials_5y_prior_endpoint", "half_enteric_5y_prior_endpoint", 
             "half_immunologicals_5y_prior_endpoint", "half_misc_5y_prior_endpoint", "half_neurological_5y_prior_endpoint", 
             "half_anti_parasites_5y_prior_endpoint", "half_dietary_supplements_5y_prior_endpoint", 
             "half_anti_septics_5y_prior_endpoint", "half_hormones_and_related_5y_prior_endpoint", 
             "half_cardio_respiratory_5y_prior_endpoint", "half_fluid_metabolites_5y_prior_endpoint", 
             "half_diuretics_5y_prior_endpoint", "half_anti_hist_5y_prior_endpoint", 
             "quartile_anti_inflam_5y_prior_endpoint", "quartile_anti_microbials_5y_prior_endpoint", 
            # "quartile_enteric_5y_prior_endpoint",
 "quartile_immunologicals_5y_prior_endpoint", 
             "quartile_misc_5y_prior_endpoint", "quartile_neurological_5y_prior_endpoint", 
             "quartile_anti_parasites_5y_prior_endpoint", "quartile_dietary_supplements_5y_prior_endpoint", 
       #      "quartile_anti_septics_5y_prior_endpoint", 
 "quartile_hormones_and_related_5y_prior_endpoint", 
          #   "quartile_cardio_respiratory_5y_prior_endpoint", #"quartile_fluid_metabolites_5y_prior_endpoint", 
         #    "quartile_diuretics_5y_prior_endpoint", "quartile_anti_hist_5y_prior_endpoint", 
             #"quartile_anti_viral_5y_prior_endpoint",
 "age_at_final_date_quartiles","age_at_neuter_years_quartiles")
dependent = "status_factor" 
results <- cohort_no_not_rec %>%
  summary_factorlist(dependent, explanatory,
  p=TRUE, add_dependent_label=TRUE,na_include = FALSE) -> t1
knitr::kable(t1, row.names=FALSE, align=c("l", "l", "r", "r", "r"))

results_df <- as.data.frame(results)

# Write the results to CSV
write.csv(results_df, "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_cross_table_case_control_output_no_missing.csv", row.names = FALSE)

#save as worddoc
# Install and load the required packages
library(officer)
library(flextable)

# Create a Word document
doc <- read_docx()

# Convert finalfit output to a flextable
results_flextable <- flextable(results_df)

# Add the flextable to the Word document
doc <- doc %>%
  body_add_flextable(results_flextable) %>%
  body_add_par("Finalfit Regression Results", style = "heading 1")

# Save the Word document
print(doc, target = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_cross_table_case_control_output_no_missing.docx")




perform_univariable_analysis(
  data = cohort_no_not_rec,
  dependent_var = "status",
  dependent_var_factor="status_factor",
  explanatory_vars = c("MDI_quintile","owner_MDI_quintile","area_type_mode_5yrs_prior","house_type_mode_5yrs_prior", 
             "water_source_mode_5yrs_prior", "pipes_metal_any_mode_5yrs_prior", "pipes_plastic_any_mode_5yrs_prior", 
             "heating_fuel_primary_mode_5yrs_prior", "cooking_fuel_primary_mode_5yrs_prior", "heating_fuel_secondary_mode_5yrs_prior", 
             "cooking_fuel_secondary_mode_5yrs_prior", "region_mode_5y_prior",  #"number_comorbs_present_lifetime", 
            # "study_year_first_chronic_inflam", #"chronic_inflam_present_5y", #"number_comorbs_present_5y", 
            # "chronic_inflam_present_no_min_year", "number_comorbs_present_no_min_year",  "diagnosis_after_comorb",  "time_between_first_comorb_and_endpoint",  
              "num_lifetime_chronic_inflam", "intestinal_parasitism","tick_borne_parasitism","other_parasitism","all_parasitism","infectious","orthopaedic","immune_mediated","cardiovascular","gastrointestinal","inflammatory_other","chronic_inflammatory",
 "USDA_30min_sy1_2", "activity_1h_min_sy1_2", 
             "activity_KC_2h_min_sy1_2", "avg_activity_duration_sy1_2_quartiles",#"avg_activity_intensity_sy1_2_terciles",
 "avg_activity_freq_sy1_2_terciles", 
             "avg_activity_freq_sy3_terciles", "avg_activity_intensity_sy3_quartiles", "avg_intensity_5y_prior_endpoint_quartiles", "avg_frequency_5y_prior_endpoint_quartiles", 
             "use_aerosol_study_years_early_life", "use_air_cleaner_study_years_early_life", "use_hepa_filter_study_years_early_life", 
             "use_moth_balls_study_years_early_life", "use_incense_or_candles_study_years_early_life", "smoke_exposure_study_years_early_life", 
       #      "any_treated_weeds_study_years_early_life", "any_treated_insects_study_years_early_life", #"any_treated_fertilizer_study_years_early_life", 
         #    "use_aerosol_sum_early_life", "use_air_cleaner_sum_early_life", "use_hepa_filter_sum_early_life", "use_moth_balls_sum_early_life", 
          #   "use_incense_or_candles_sum_early_life", 
 #"smoke_exposure_sum_early_life",
 #"any_treated_weeds_sum_early_life", 
    #         "any_treated_insects_sum_early_life", #"any_treated_fertilizer_sum_early_life", 
 "use_aerosol_study_years_rest_of_life", 
             "use_air_cleaner_study_years_rest_of_life", "use_hepa_filter_study_years_rest_of_life", "use_moth_balls_study_years_rest_of_life", 
             "use_incense_or_candles_study_years_rest_of_life", "smoke_exposure_study_years_rest_of_life", "any_treated_weeds_study_years_rest_of_life", 
             "any_treated_insects_study_years_rest_of_life", "any_treated_fertilizer_study_years_rest_of_life", #"use_aerosol_sum_rest_of_life", 
 #            "use_air_cleaner_sum_rest_of_life", "use_hepa_filter_sum_rest_of_life", "use_moth_balls_sum_rest_of_life", 
          #   "use_incense_or_candles_sum_rest_of_life", 
# "smoke_exposure_sum_rest_of_life", 
# "any_treated_weeds_sum_rest_of_life", 
            # "any_treated_insects_sum_rest_of_life", "any_treated_fertilizer_sum_rest_of_life", 
"use_aerosol_study_years_whole_life", 
             "use_air_cleaner_study_years_whole_life", "use_hepa_filter_study_years_whole_life", "use_moth_balls_study_years_whole_life", 
             "use_incense_or_candles_study_years_whole_life", "smoke_exposure_study_years_whole_life", "any_treated_weeds_study_years_whole_life", 
             "any_treated_insects_study_years_whole_life", "any_treated_fertilizer_study_years_whole_life",# "use_aerosol_sum_whole_life", 
           #  "use_air_cleaner_sum_whole_life", "use_hepa_filter_sum_whole_life", "use_moth_balls_sum_whole_life",# "use_incense_or_candles_sum_whole_life", 
            # "smoke_exposure_sum_whole_life", 
 #"any_treated_weeds_sum_whole_life", "any_treated_insects_sum_whole_life", 
            # "any_treated_fertilizer_sum_whole_life",
 "hours_of_smoke_early_life_total_dosage_terciles", "hours_of_smoke_rest_of_life_total_dosage_terciles", 
             "hours_of_smoke_whole_life_total_dosage_terciles", "in.the.garage_YN_study_years_early_life", "in.the.house_YN_study_years_early_life", 
             "outside_YN_study_years_early_life",  "in.the.house_YN_sum_early_life_terciles", 
             "in.the.garage_YN_study_years_rest_of_life", "in.the.house_YN_study_years_rest_of_life", "outside_YN_study_years_rest_of_life", 
              "in.the.house_YN_sum_rest_of_life_quartiles",  
             "in.the.garage_YN_study_years_whole_life", "in.the.house_YN_study_years_whole_life", "outside_YN_study_years_whole_life", 
             "in.the.house_YN_sum_whole_life_quartiles",  
             "majority_sleep_location_lifetime",  "main_lifestyle_category", 
 "main_lifestyle_category_mode_5yrs_prior", "sleep_location_mode_5yrs_prior", 
             "use_aerosol_exposed_5yrs_prior", "use_air_cleaner_exposed_5yrs_prior", "use_hepa_filter_exposed_5yrs_prior", 
             "use_moth_balls_exposed_5yrs_prior", "use_incense_or_candles_exposed_5yrs_prior", "smoke_exposure_exposed_5yrs_prior", 
             "any_treated_weeds_exposed_5yrs_prior", "any_treated_insects_exposed_5yrs_prior", "any_treated_fertilizer_exposed_5yrs_prior", 
             "hours_of_smoke_exposure_hours_5yrs_prior_terciles", "sex", "neuter", "mismating_ever",
             "neutered_reason", "neutered_ever_bred", "spayed_while_in_heat", "spayed_while_pregnant", "prevent_heat_ever", 
             "no_pregnancy_ever", "intact_sired_litters_natural_ever", "neutered_females_bred_with_count_ever", 
             "neutered_sired_litters_natural_ever", "heats_before_neutering", "heats_before_neutering_terciles",
             "preg_pre_neuter",# "mated_pre_neuter",
 "num_breeding_before_neutering_terciles", "ever_bred_from", 
              "avg_weight_5y_prior_endpoint_quartiles", "avg_height_5y_prior_endpoint_quartiles", 
             "avg_purina_BCS_5y_prior_endpoint_quartiles", "mode_purina_BCS_5y_prior_endpoint", "mode_BCS_classify", "half_anti_microbials_lifetime", "half_dietary_supplements_lifetime", 
             "half_enteric_lifetime", "half_fluid_metabolites_lifetime", "half_hormones_and_related_lifetime", "half_immunologicals_lifetime", 
             "half_misc_lifetime", "half_neurological_lifetime", "half_anti_parasites_lifetime", "half_anti_septics_lifetime", 
             "half_cardio_respiratory_lifetime", "half_diuretics_lifetime", "half_anti_hist_lifetime", 
             "quartile_anti_inflam_lifetime", "quartile_anti_microbials_lifetime", "quartile_dietary_supplements_lifetime", 
            #not enough data for quartiles "quartile_enteric_lifetime", "quartile_fluid_metabolites_lifetime", 
 "quartile_hormones_and_related_lifetime", 
             "quartile_immunologicals_lifetime", "quartile_misc_lifetime", "quartile_neurological_lifetime", 
             "quartile_anti_parasites_lifetime", 
 #not enough data for quartiles "quartile_anti_septics_lifetime", "quartile_cardio_respiratory_lifetime",   "quartile_diuretics_lifetime", "quartile_anti_hist_lifetime", 
             "half_anti_inflam_5y_prior_endpoint", "half_anti_microbials_5y_prior_endpoint", "half_enteric_5y_prior_endpoint", 
             "half_immunologicals_5y_prior_endpoint", "half_misc_5y_prior_endpoint", "half_neurological_5y_prior_endpoint", 
             "half_anti_parasites_5y_prior_endpoint", "half_dietary_supplements_5y_prior_endpoint", 
             "half_anti_septics_5y_prior_endpoint", "half_hormones_and_related_5y_prior_endpoint", 
             "half_cardio_respiratory_5y_prior_endpoint", "half_fluid_metabolites_5y_prior_endpoint", 
             "half_diuretics_5y_prior_endpoint", "half_anti_hist_5y_prior_endpoint", 
             "quartile_anti_inflam_5y_prior_endpoint", "quartile_anti_microbials_5y_prior_endpoint", 
            # "quartile_enteric_5y_prior_endpoint",
 "quartile_immunologicals_5y_prior_endpoint", 
             "quartile_misc_5y_prior_endpoint", "quartile_neurological_5y_prior_endpoint", 
             "quartile_anti_parasites_5y_prior_endpoint", "quartile_dietary_supplements_5y_prior_endpoint", 
       #      "quartile_anti_septics_5y_prior_endpoint", 
 "quartile_hormones_and_related_5y_prior_endpoint", 
          #   "quartile_cardio_respiratory_5y_prior_endpoint", #"quartile_fluid_metabolites_5y_prior_endpoint", 
         #    "quartile_diuretics_5y_prior_endpoint", "quartile_anti_hist_5y_prior_endpoint", 
             #"quartile_anti_viral_5y_prior_endpoint",
 "age_at_final_date_quartiles","age_at_neuter_years_quartiles"),
  output_csv = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_time_to_HSA_cox_ph_univar_no_missing.csv",
  output_doc = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_time_to_HSA_cox_ph_univar_no_missing.docx"
)
```

## Multivariable Cox PH

```{r}
lrt_pvalues <-read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_time_to_HSA_cox_ph_univar_no_missing.csv")
multivar_variables <- lrt_pvalues %>%
  filter(lrt_pvalue <= 0.2) %>%
  distinct(variable)
cols_to_keep = multivar_variables$variable
cols_to_keep <- append(cols_to_keep, "status")
cols_to_keep <- append(cols_to_keep, "survival_time_days")
cols_to_keep <- append(cols_to_keep, "status_factor")
cohort_no_not_rec_sig <- cohort_no_not_rec[,cols_to_keep]

lrt_pvalues$pval <- sprintf("%.20f", lrt_pvalues$lrt_pvalue)
```

Variables to take forward (p\<0.2):

```{r}
print(multivar_variables)
```

Assess which variables are correlated:

```{r,error=TRUE}
source("C:/Users/ctaylor18/GitHub/HSA-VC-enviro-analyses/functions/corr_matrix_creation.R")
#use sig variables list generated above

create_corr_matrix(cohort_no_not_rec_sig,c(1:60), "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_survival_no_missing.csv", "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_survival2_no_missing.csv","C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_survival_corrvalue_no_missing.csv","C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_survival_no_missing.png")
```

### Correlated variables:

```{r,error=TRUE}
corr_values <- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_survival_corrvalue.csv")

#ditch rows where variable 1 and 2 values are same as in another row :
corr_values2 <- corr_values %>%
  dplyr::select(-c(X))

corr_values3 <- corr_values2 %>%
  mutate(Var1_sorted = pmin(Variable1, Variable2),  
         Var2_sorted = pmax(Variable1, Variable2)) %>%  
  distinct(Var1_sorted, Var2_sorted, Correlation, .keep_all = TRUE) %>%  
  dplyr::select(-Var1_sorted, -Var2_sorted)  # Remove helper columns


print(corr_values3)
```

Correlated variables:

-   Similar as above

#### Multivariable stepwise

-   starting with smallest p value variables

-   also not retaining variables where only the 'not_recorded' is the only sig category

Weight

```{r}
source("C:/Users/ctaylor18/GitHub/HSA-VC-survival-analyses/HSA-VC-survival-analyses/code/functions/multivar_cox_model_analyse_AIC_funct.R")

#null model
cox_model_null <- coxph(Surv(survival_time_days, status) ~1
, data = cohort_abr_sig)
summary(cox_model_null)

#weight model
cox_model_1 <- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles
, data = cohort_abr_sig)
summary(cox_model_1)

analyse_model(cox_model_1,cohort_abr_sig,cox_model_null)
```

Activity freq SY3

```{r}
cox_model_2 <- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles
, data = cohort_abr_sig)
summary(cox_model_2)

analyse_model(cox_model_2,cohort_abr_sig,cox_model_1)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Antiparasites quartiles/halves

```{r}
cox_model_3 <- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint
, data = cohort_abr_sig)
summary(cox_model_3)

analyse_model(cox_model_3,cohort_abr_sig,cox_model_2)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

In the house rest of life and quartiles

-   neither improve model

```{r}
#rest of life
cox_model_4 <- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + in.the.house_YN_study_years_rest_of_life
, data = cohort_abr_sig)
summary(cox_model_4)

analyse_model(cox_model_4,cohort_abr_sig,cox_model_3)
#quartiles
cox_model_4a <- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + in.the.house_YN_sum_whole_life_quartiles
, data = cohort_abr_sig)
summary(cox_model_4a)

analyse_model(cox_model_4a,cohort_abr_sig,cox_model_3)
```

Spayed when in heat

-   exclude as only male cateogry sig

```{r}
cox_model_5 <- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + spayed_while_in_heat
, data = cohort_abr_sig)
summary(cox_model_5)

analyse_model(cox_model_5,cohort_abr_sig,cox_model_3)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","spayed_while_in_heat")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Sex

```{r}
cox_model_6 <- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex
, data = cohort_abr_sig)
summary(cox_model_6)

analyse_model(cox_model_6,cohort_abr_sig,cox_model_3)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Outside whole life/rest of life

```{r}
cox_model_7a <- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex +outside_YN_study_years_whole_life
, data = cohort_abr_sig)
summary(cox_model_7a)

analyse_model(cox_model_7a,cohort_abr_sig,cox_model_6)

cox_model_7b<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex +outside_YN_study_years_rest_of_life
, data = cohort_abr_sig)
summary(cox_model_7b)

analyse_model(cox_model_7b,cohort_abr_sig,cox_model_6)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","outside_YN_study_years_whole_life")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

No pregnancy ever

-   sex and no pregnancy correlated so cannot keep both in

-   no pregnancy ever doesnt improve model

```{r}
cox_model_8<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex +outside_YN_study_years_rest_of_life +no_pregnancy_ever
, data = cohort_abr_sig)
summary(cox_model_8)

analyse_model(cox_model_8,cohort_abr_sig,cox_model_6)
vif(cox_model_8)

```

Main lifestyle category 5y prior and ever

-says improve model LRT but no levels are significant = not retained

```{r}
cox_model_9a<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex +outside_YN_study_years_rest_of_life +main_lifestyle_category
, data = cohort_abr_sig)
summary(cox_model_9a)

analyse_model(cox_model_9a,cohort_abr_sig,cox_model_6)
vif(cox_model_9a)

cox_model_9b<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex +outside_YN_study_years_rest_of_life +main_lifestyle_category_mode_5yrs_prior
, data = cohort_abr_sig)
summary(cox_model_9b)

analyse_model(cox_model_9b,cohort_abr_sig,cox_model_6)
vif(cox_model_9b)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","main_lifestyle_category")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Heats before neutering terciles

-   cant have sex and heats

```{r,error=TRUE}
cox_model_10<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex +outside_YN_study_years_rest_of_life +heats_before_neutering_terciles
, data = cohort_abr_sig)
summary(cox_model_10)

analyse_model(cox_model_10,cohort_abr_sig,cox_model_6)
vif(cox_model_10)
```

Incense whole life

-   not significant

```{r}
cox_model_11<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex +use_incense_or_candles_study_years_whole_life
, data = cohort_abr_sig)
summary(cox_model_11)

analyse_model(cox_model_11,cohort_abr_sig,cox_model_6)
vif(cox_model_11)

```

Mismating ever

-   doesnt improve model

```{r}
cox_model_12<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex +mismating_ever
, data = cohort_abr_sig)
summary(cox_model_12)

analyse_model(cox_model_12,cohort_abr_sig,cox_model_6)
vif(cox_model_12)
```

Use hepa filters rest of life

```{r}
cox_model_13<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life
, data = cohort_abr_sig)
summary(cox_model_13)

analyse_model(cox_model_13,cohort_abr_sig,cox_model_6)
vif(cox_model_13)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Avg height

-   improves model but no categories significant = not retain

```{r}
cox_model_14<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + avg_height_5y_prior_endpoint_quartiles
, data = cohort_abr_sig)
summary(cox_model_14)

analyse_model(cox_model_14,cohort_abr_sig,cox_model_13)
vif(cox_model_14)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","avg_height_5y_prior_endpoint_quartiles")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Use air cleaner - rest, whole,early

-   none improve model

```{r}
#5y prior
cox_model_16a<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + use_air_cleaner_exposed_5yrs_prior
, data = cohort_abr_sig)
summary(cox_model_16a)

analyse_model(cox_model_16a,cohort_abr_sig,cox_model_13)
vif(cox_model_16a)
#whole life
cox_model_16b<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + use_air_cleaner_study_years_whole_life
, data = cohort_abr_sig)
summary(cox_model_16b)

analyse_model(cox_model_16b,cohort_abr_sig,cox_model_13)
vif(cox_model_16b)
#rest of life

cox_model_16c<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + use_air_cleaner_study_years_rest_of_life
, data = cohort_abr_sig)
summary(cox_model_16c)

analyse_model(cox_model_16c,cohort_abr_sig,cox_model_13)
vif(cox_model_16c)
```

Age at neuter

-   corr with sex so need to choose which to include

-   also create an interaction variable between these two

```{r}
cox_model_17<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + age_at_neuter_years_quartiles
, data = cohort_abr_sig)
summary(cox_model_17)

analyse_model(cox_model_17,cohort_abr_sig,cox_model_13)
vif(cox_model_17)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","age_at_neuter_years_quartiles")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

In the garage

-   early years sig

```{r}
#early years
cox_model_18<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life
, data = cohort_abr_sig)
summary(cox_model_18)

analyse_model(cox_model_18,cohort_abr_sig,cox_model_13)
vif(cox_model_18)

explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","in.the.garage_YN_study_years_early_life")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod

#whole life
cox_model_18a<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_whole_life
, data = cohort_abr_sig)
summary(cox_model_18a)

analyse_model(cox_model_18a,cohort_abr_sig,cox_model_13)
vif(cox_model_18a)


```

Avg purina BCS

-   only sig for not recorded category once in model, elect not to retain

```{r}
cox_model_19<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +avg_purina_BCS_5y_prior_endpoint_quartiles
, data = cohort_abr_sig)
summary(cox_model_19)

analyse_model(cox_model_19,cohort_abr_sig,cox_model_18)
vif(cox_model_19)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","avg_purina_BCS_5y_prior_endpoint_quartiles")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Avg activity duration early years

-   although sig it is only due to the not recorded variable, not retained

```{r}
cox_model_20<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +avg_activity_duration_sy1_2_quartiles
, data = cohort_abr_sig)
summary(cox_model_20)

analyse_model(cox_model_20,cohort_abr_sig,cox_model_18)
vif(cox_model_20)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","avg_activity_duration_sy1_2_quartiles")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Outside early years

-   sig

```{r}
cox_model_21<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
, data = cohort_abr_sig)
summary(cox_model_21)

analyse_model(cox_model_21,cohort_abr_sig,cox_model_18)
vif(cox_model_21)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","outside_YN_study_years_early_life")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Neutering reason

-   only neuter category significant is 'Entire' = not retain

```{r}
cox_model_22<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
+ neutered_reason, data = cohort_abr_sig)
summary(cox_model_22)

analyse_model(cox_model_22,cohort_abr_sig,cox_model_21)
vif(cox_model_22)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","outside_YN_study_years_early_life","neutered_reason")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Activity freq 5y prior endpoint quartiles

-   sig

```{r}
cox_model_23<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
+ avg_frequency_5y_prior_endpoint_quartiles
, data = cohort_abr_sig)
summary(cox_model_23)

analyse_model(cox_model_23,cohort_abr_sig,cox_model_21)
vif(cox_model_23)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","outside_YN_study_years_early_life","avg_frequency_5y_prior_endpoint_quartiles
")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Half neuro lifetime

-   not sig

```{r,error=TRUE}
cox_model_24<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
+ avg_frequency_5y_prior_endpoint_quartiles + half_neurological_lifetime
, data = cohort_abr_sig)
summary(cox_model_24)

analyse_model(cox_model_24,cohort_abr_sig,cox_model_23)
vif(cox_model_24)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","outside_YN_study_years_early_life","avg_frequency_5y_prior_endpoint_quartiles
", "half_neurological_lifetime")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Half enteric lifetime

-   sig but collinear with parasites, keep parasites

```{r,error=TRUE}
cox_model_25<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
+ avg_frequency_5y_prior_endpoint_quartiles + half_enteric_lifetime
, data = cohort_abr_sig)
summary(cox_model_25)

analyse_model(cox_model_25,cohort_abr_sig,cox_model_23)
vif(cox_model_25)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","outside_YN_study_years_early_life","avg_frequency_5y_prior_endpoint_quartiles", "half_enteric_lifetime")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Use air cleaner rest of life/5y

```{r}
cox_model_26<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
+ avg_frequency_5y_prior_endpoint_quartiles + use_air_cleaner_exposed_5yrs_prior
, data = cohort_abr_sig)
summary(cox_model_26)

analyse_model(cox_model_26,cohort_abr_sig,cox_model_23)
vif(cox_model_26)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","outside_YN_study_years_early_life","avg_frequency_5y_prior_endpoint_quartiles", "use_air_cleaner_exposed_5yrs_prior")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Half enteric 5y

-   not sig

```{r}
cox_model_27<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
+ avg_frequency_5y_prior_endpoint_quartiles + half_enteric_5y_prior_endpoint
, data = cohort_abr_sig)
summary(cox_model_27)

analyse_model(cox_model_27,cohort_abr_sig,cox_model_23)
#vif(cox_model_27)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","outside_YN_study_years_early_life","avg_frequency_5y_prior_endpoint_quartiles", "half_enteric_5y_prior_endpoint")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Immune med comorbs

-   not significant

```{r}
cox_model_28<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
+ avg_frequency_5y_prior_endpoint_quartiles + immune_mediated
, data = cohort_abr_sig)
summary(cox_model_28)

analyse_model(cox_model_28,cohort_abr_sig,cox_model_23)
#vif(cox_model_28)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","outside_YN_study_years_early_life","avg_frequency_5y_prior_endpoint_quartiles", "immune_mediated")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Cooking fuel secondary

-   sig but only not recorded category, not retained

```{r}
cox_model_29<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
+ avg_frequency_5y_prior_endpoint_quartiles + cooking_fuel_secondary_mode_5yrs_prior
, data = cohort_abr_sig)
summary(cox_model_29)

analyse_model(cox_model_29,cohort_abr_sig,cox_model_23)
vif(cox_model_29)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","outside_YN_study_years_early_life","avg_frequency_5y_prior_endpoint_quartiles", "cooking_fuel_secondary_mode_5yrs_prior")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Prevent heat ever

-   not sig

```{r}
cox_model_30<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
+ avg_frequency_5y_prior_endpoint_quartiles + prevent_heat_ever
, data = cohort_abr_sig)
summary(cox_model_30)

analyse_model(cox_model_30,cohort_abr_sig,cox_model_23)
vif(cox_model_30)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","outside_YN_study_years_early_life","avg_frequency_5y_prior_endpoint_quartiles", "prevent_heat_ever")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Half anti septics lifetime

```{r,error=TRUE}
cox_model_31<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
+ avg_frequency_5y_prior_endpoint_quartiles + half_anti_septics_lifetime
, data = cohort_abr_sig)
summary(cox_model_31)

analyse_model(cox_model_31,cohort_abr_sig,cox_model_23)
vif(cox_model_31)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","outside_YN_study_years_early_life","avg_frequency_5y_prior_endpoint_quartiles", "half_anti_septics_lifetime")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Aerosol whole life

-   not sig

```{r}
cox_model_32<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
+ avg_frequency_5y_prior_endpoint_quartiles + use_aerosol_study_years_whole_life
, data = cohort_abr_sig)
summary(cox_model_32)

analyse_model(cox_model_32,cohort_abr_sig,cox_model_23)
vif(cox_model_32)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","outside_YN_study_years_early_life","avg_frequency_5y_prior_endpoint_quartiles", "use_aerosol_study_years_whole_life")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Smoke exposure early years

-   not sig

```{r}
cox_model_33<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
+ avg_frequency_5y_prior_endpoint_quartiles + smoke_exposure_study_years_early_life
, data = cohort_abr_sig)
summary(cox_model_33)

analyse_model(cox_model_33,cohort_abr_sig,cox_model_23)
vif(cox_model_33)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","outside_YN_study_years_early_life","avg_frequency_5y_prior_endpoint_quartiles", "smoke_exposure_study_years_early_life")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Use mothballs rest of life

-   not significant

```{r}
cox_model_34<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
+ avg_frequency_5y_prior_endpoint_quartiles + use_moth_balls_study_years_rest_of_life
, data = cohort_abr_sig)
summary(cox_model_34)

analyse_model(cox_model_34,cohort_abr_sig,cox_model_23)
vif(cox_model_34)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","outside_YN_study_years_early_life","avg_frequency_5y_prior_endpoint_quartiles", "use_moth_balls_study_years_rest_of_life")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Half cardioresp lifetime

-   not sig

```{r,error=TRUE}
cox_model_35<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
+ avg_frequency_5y_prior_endpoint_quartiles + half_cardio_respiratory_lifetime
, data = cohort_abr_sig)
summary(cox_model_35)

analyse_model(cox_model_35,cohort_abr_sig,cox_model_23)
vif(cox_model_35)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","outside_YN_study_years_early_life","avg_frequency_5y_prior_endpoint_quartiles", "half_cardio_respiratory_lifetime")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Mode Purina BCS

-   although sig, output is nonsensical

```{r}
cox_model_36<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
+ avg_frequency_5y_prior_endpoint_quartiles + mode_purina_BCS_5y_prior_endpoint
, data = cohort_abr_sig)
summary(cox_model_36)

analyse_model(cox_model_36,cohort_abr_sig,cox_model_23)
vif(cox_model_36)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","outside_YN_study_years_early_life","avg_frequency_5y_prior_endpoint_quartiles", "mode_purina_BCS_5y_prior_endpoint")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Pipes plastic 5y

-   only sig due to not recorded category

```{r}
cox_model_37<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
+ avg_frequency_5y_prior_endpoint_quartiles + pipes_plastic_any_mode_5yrs_prior
, data = cohort_abr_sig)
summary(cox_model_37)

analyse_model(cox_model_37,cohort_abr_sig,cox_model_23)
vif(cox_model_37)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","outside_YN_study_years_early_life","avg_frequency_5y_prior_endpoint_quartiles", "pipes_plastic_any_mode_5yrs_prior")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Exercise intensity 5y prior

-   cant have freq and intensity in model, freq is more informative but surely overlapping with freq sy3 terciels, keep in 5y prior one

```{r}
cox_model_38<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + avg_activity_freq_sy3_terciles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
+ avg_frequency_5y_prior_endpoint_quartiles + avg_intensity_5y_prior_endpoint_quartiles
, data = cohort_abr_sig)
summary(cox_model_38)

analyse_model(cox_model_38,cohort_abr_sig,cox_model_23)
#vif(cox_model_38)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,"avg_activity_freq_sy3_terciles",
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","outside_YN_study_years_early_life","avg_frequency_5y_prior_endpoint_quartiles", "avg_intensity_5y_prior_endpoint_quartiles")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Heating fuel primary 5y prior

-   sig but only due to not recorded category

```{r}
cox_model_39<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
+ avg_frequency_5y_prior_endpoint_quartiles + heating_fuel_primary_mode_5yrs_prior
, data = cohort_abr_sig)
summary(cox_model_39)

analyse_model(cox_model_39,cohort_abr_sig,cox_model_23)
#vif(cox_model_38)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","outside_YN_study_years_early_life","avg_frequency_5y_prior_endpoint_quartiles", "heating_fuel_primary_mode_5yrs_prior")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Chronic inflam present no min year

-   not sig

```{r}
cox_model_40<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
+ avg_frequency_5y_prior_endpoint_quartiles 
, data = cohort_abr_sig)
summary(cox_model_40)

analyse_model(cox_model_40,cohort_abr_sig,cox_model_23)
vif(cox_model_40)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","outside_YN_study_years_early_life","avg_frequency_5y_prior_endpoint_quartiles")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

House type

-   not sig

```{r}
cox_model_41<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
+ avg_frequency_5y_prior_endpoint_quartiles + house_type_mode_5yrs_prior
, data = cohort_abr_sig)
summary(cox_model_41)

analyse_model(cox_model_41,cohort_abr_sig,cox_model_23)
vif(cox_model_41)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","outside_YN_study_years_early_life","avg_frequency_5y_prior_endpoint_quartiles", "house_type_mode_5yrs_prior")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Half fluid metabolites

-   not sig

```{r,error=TRUE}
cox_model_42<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles+quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
+ avg_frequency_5y_prior_endpoint_quartiles + half_fluid_metabolites_5y_prior_endpoint
, data = cohort_abr_sig)
summary(cox_model_42)

analyse_model(cox_model_42,cohort_abr_sig,cox_model_23)
vif(cox_model_42)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","outside_YN_study_years_early_life","avg_frequency_5y_prior_endpoint_quartiles", "house_type_mode_5yrs_prior")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Sleep location

-   only not recorded is sig

```{r}
cox_model_43<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
+ avg_frequency_5y_prior_endpoint_quartiles + sleep_location_mode_5yrs_prior
, data = cohort_abr_sig)
summary(cox_model_43)

analyse_model(cox_model_43,cohort_abr_sig,cox_model_23)
vif(cox_model_43)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","outside_YN_study_years_early_life","avg_frequency_5y_prior_endpoint_quartiles", "sleep_location_mode_5yrs_prior")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

Assess final model: -Same as above method

```{r}
cox_model_final<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint + sex + use_hepa_filter_study_years_rest_of_life + in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
+ avg_frequency_5y_prior_endpoint_quartiles 
, data = cohort_abr_sig)
summary(cox_model_final)

vif(cox_model_final)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,
                 "quartile_anti_parasites_5y_prior_endpoint","sex","use_hepa_filter_study_years_rest_of_life","outside_YN_study_years_early_life","avg_frequency_5y_prior_endpoint_quartiles")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod

#remove hepa as not sig 

cox_model_final<- coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint + sex+ in.the.garage_YN_study_years_early_life +outside_YN_study_years_early_life
+ avg_frequency_5y_prior_endpoint_quartiles 
, data = cohort_abr_sig)
summary(cox_model_final)

vif(cox_model_final)


explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,
                 "quartile_anti_parasites_5y_prior_endpoint","sex",
                 "avg_frequency_5y_prior_endpoint_quartiles")
dependent = 'Surv(survival_time_days,status)'
# Summarize the Cox model output in a table
mod <- cohort_abr_sig %>%
  coxphmulti(dependent,explanatory)%>%
  fit2df()

mod
```

#### Post-hoc assessment:

##### assess for confounding - OK

```{r}
explanatory <- c("avg_weight_5y_prior_endpoint_quartiles" ,
                 "quartile_anti_parasites_5y_prior_endpoint","sex","avg_frequency_5y_prior_endpoint_quartiles")
dependent = 'Surv(survival_time_days,status)'
table<- cohort_abr_sig%>% 
    finalfit(dependent, explanatory, add_dependent_label = FALSE) %>% 
    rename("Overall survival" = label) %>% 
    rename(" " = levels) %>% 
    rename("  " = all)
table
```

##### Assess for interactions

-   weight and parasite quartiles significant , all else not in pairwise

```{r}
# Load the survival package
library(survival)


# Function to assess pairwise interactions using LRT
assess_pairwise_interactions <- function(data, time_var, status_var, variables) {
  
  # Initialize a dataframe to store the results
  results <- data.frame(Interaction = character(), AIC = numeric(), LRT_pvalue = numeric(), stringsAsFactors = FALSE)
  
  # Loop through each pair of specified variables
  for (i in 1:(length(variables) - 1)) {
    for (j in (i + 1):length(variables)) {
      
      # Get the current pair of variables
      var1 <- variables[i]
      var2 <- variables[j]
      
      # Create the formula for the base model (without interaction)
      base_formula <- as.formula(paste("Surv(", time_var, ",", status_var, ") ~", var1, "+", var2))
      
      # Create the formula for the interaction model (with interaction term)
      interaction_formula <- as.formula(paste("Surv(", time_var, ",", status_var, ") ~", 
                                              var1, "+", var2, "+", paste(var1, var2, sep = "*")))
      
      # Fit the base model
      base_model <- coxph(base_formula, data = data)
      
      # Fit the interaction model
      interaction_model <- coxph(interaction_formula, data = data)
      
      # Calculate the AIC of the interaction model
      interaction_aic <- AIC(interaction_model)
      
      # Perform Likelihood Ratio Test (LRT) to compare models
      lrt <- anova(base_model, interaction_model, test = "Chisq")
      lrt_pvalue <- lrt$`Pr(>|Chi|)`[2]  # Extract p-value for interaction
      
      # Store the results
      results <- rbind(results, data.frame(Interaction = paste(var1, var2, sep = "*"), 
                                           AIC = interaction_aic, 
                                           LRT_pvalue = lrt_pvalue))
      
      # Display results for each interaction model
      print(paste("Model with interaction between:", var1, "and", var2))
      print(summary(interaction_model))
      print(paste("AIC:", interaction_aic))
      print(paste("LRT p-value:", lrt_pvalue))
      cat("\n-----------------------------------------------------\n")
    }
  }
  
  # Order results by AIC
  results <- results[order(results$AIC), ]
  
  # Display the results
  print("Pairwise Interaction Terms Ranked by AIC:")
  print(results)
  
  return(results)
}
# Specify the time and status variables along with the variables you want to assess for interactions
time_variable <- "survival_time_days"
status_variable <- "status"
specified_variables <- c("avg_weight_5y_prior_endpoint_quartiles" ,
                 "quartile_anti_parasites_5y_prior_endpoint","sex","avg_frequency_5y_prior_endpoint_quartiles")

# Assess pairwise interactions using only the two terms in each model
interaction_results <- assess_pairwise_interactions(cohort_abr_sig, time_variable, status_variable, specified_variables)
```

-   Assess weight \* parasite in multivar model

```{r}
int <-coxph(Surv(survival_time_days, status) ~ avg_weight_5y_prior_endpoint_quartiles + quartile_anti_parasites_5y_prior_endpoint + sex
+ avg_frequency_5y_prior_endpoint_quartiles + avg_weight_5y_prior_endpoint_quartiles* quartile_anti_parasites_5y_prior_endpoint  , data = cohort_abr_sig) 
AIC(int)
summary(int)
# Perform the likelihood ratio test
lr_test <- anova(int, cox_model_final)
#
print(lr_test)
analyse_model(int,HSA_multi_abr_day1,cox_model_final)
```

```{r}
library(broom)

cox_summary <- broom::tidy(int, exponentiate = TRUE, conf.int = TRUE) 
cox_summary <- cox_summary %>%
 mutate(
    estimate = sprintf("%.4f", estimate),  # HR to 4 decimal places
    conf.low = sprintf("%.4f", conf.low),  # Lower 95% CI
    conf.high = sprintf("%.4f", conf.high), # Upper 95% CI
    p.value = format(p.value, scientific = FALSE, digits = 4)  # Format p-value
  )

print(cox_summary)
```

##### Proportional hazards assumption

-   almost all variables significant in Schoenfeld residuals

```{r}
dependent = "Surv(survival_time_days, status)"
explanatory = c("avg_weight_5y_prior_endpoint_quartiles" ,
                 "quartile_anti_parasites_5y_prior_endpoint","sex","avg_frequency_5y_prior_endpoint_quartiles"
                )

ph_schoen <- cohort_abr_sig %>% 
    coxphmulti(dependent, explanatory) %>% 
    cox.zph() %>% 
    {zph_result <<- .} %>% 
    plot(var=4)

zph_result
```

Visualise proportional hazards with KM for each variable

-   Weight

```{r}

dependent_os = "Surv(survival_time_months, status)"
explanatory = c("avg_weight_5y_prior_endpoint_quartiles")

cohort_abr %>% 
    surv_plot(dependent_os, explanatory, pval = TRUE)
```

-   Parasite treatment

```{r}
survdiff(Surv(survival_time_months, status) ~ quartile_anti_parasites_5y_prior_endpoint, data = cohort_abr)

dependent_os = "Surv(survival_time_months, status)"
explanatory = c("quartile_anti_parasites_5y_prior_endpoint")

cohort_abr %>% 
    surv_plot(dependent_os, explanatory, pval = TRUE)
```

-   Sex

```{r}
survdiff(Surv(survival_time_months, status) ~ sex, data = cohort_abr)

dependent_os = "Surv(survival_time_months, status)"
explanatory = c("sex")

cohort_abr %>% 
    surv_plot(dependent_os, explanatory, pval = TRUE)
```

-   Exercise frequency

```{r}
survdiff(Surv(survival_time_months, status) ~ avg_frequency_5y_prior_endpoint_quartiles, data = cohort_abr)

dependent_os = "Surv(survival_time_months, status)"
explanatory = c("avg_frequency_5y_prior_endpoint_quartiles")

cohort_abr %>% 
    surv_plot(dependent_os, explanatory, pval = TRUE)
```
