---
title: "GRLS log reg HSA revised analysis approach"
format: html
editor: visual
---

# Variables associated with HSA in GRLS cohort

Load in dataset and packages

```{r}
library(tidyverse)
library(tidyverse)
library(dplyr)
library(survival)
library(ggplot2)
library(tibble)
library(kableExtra)
library(epiDisplay)
library(lme4)
#devtools::install_github("zabore/ezfun")
ezfun::set_ccf_palette("contrast")
library(lubridate)
library(ggsurvfit)
library(gtsummary)
library(tidycmprsk)
library(gtable)
#devtools::install_github("zabore/condsurv")
library(condsurv)
library(survminer)
library(cardx)
library(finalfit)
library(crosstable)
library(kableExtra)
library(officer)
library(flextable)
# data was preliminarily tidied in GRLS cox HSA time to diagnosis script before changing to log reg analyses
cohort_abr <-read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_cohort_all_RFs_data_tidied_from_cox.csv")
```

## Updated weight data

Incorporate in updated weight from Zoe's work (she has modified the erroneus lb kg issues also now all only adult weight so no \<18m weights in here)

```{r}
updated_weight <- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Data/weight_cleaned_grls.csv")

#existing weight columns are average and median 5y prior to endpoint and either split into quartiles or biological

#use the updated_weight df to create these same variables again + whole life weight

weight_df <- updated_weight %>%
  mutate(
    weight_kg = case_when(
      str_to_lower(weight_units) %in% c("lb", "lbs") ~ weight * 0.45359237,
      str_to_lower(weight_units) == "kg" ~ weight,
      TRUE ~ NA_real_
    )
  )

median_weight <- weight_df %>%
  filter(!is.na(weight_kg)) %>%
  group_by(subject_id) %>%
  arrange(year_in_study, .by_group = TRUE) %>%
  summarise(
    median_weight_lifetime_updated = median(weight_kg, na.rm = TRUE),

    median_weight_5y_prior_endpoint_updated = median(
      weight_kg[year_in_study %in% tail(sort(unique(year_in_study)), 5)],
      na.rm = TRUE
    ),

    .groups = "drop"
  )

#add into main df and see how different values are
cohort_abr_updated_weight2<- cohort_abr %>%
  left_join(median_weight, by="subject_id")


#check how diff
  
  
summarise(
  cohort_abr_updated_weight2,
  old_mean = mean(as.numeric(na_if(median_weight_5y_prior_endpoint_kg, "not_recorded")), na.rm = TRUE),
  old_median = median(as.numeric(na_if(median_weight_5y_prior_endpoint_kg, "not_recorded")), na.rm = TRUE),
  old_q1 = quantile(as.numeric(na_if(median_weight_5y_prior_endpoint_kg, "not_recorded")), 0.25, na.rm = TRUE),
  old_q3 = quantile(as.numeric(na_if(median_weight_5y_prior_endpoint_kg, "not_recorded")), 0.75, na.rm = TRUE),

  new_mean = mean(median_weight_5y_prior_endpoint_updated, na.rm = TRUE),
  new_median = median(median_weight_5y_prior_endpoint_updated, na.rm = TRUE),
  new_q1 = quantile(median_weight_5y_prior_endpoint_updated, 0.25, na.rm = TRUE),
  new_q3 = quantile(median_weight_5y_prior_endpoint_updated, 0.75, na.rm = TRUE)
)  


#remove old weight rows from df
cohort_abr_updated_weight2<- cohort_abr_updated_weight2 %>%
  dplyr::select(-c("median_weight_5y_prior_endpoint_kg","avg_weight_5y_prior_endpoint_kg","avg_weight_5y_prior_endpoint_kg_outliers_removed","median_weight_5y_prior_endpoint_any_records_kg","median_weight_5y_prior_endpoint_kg_outliers_removed","median_weight_5y_prior_endpoint_any_records_kg","median_weight_5y_prior_endpoint_any_records_kg_outliers_removed",
                  "avg_weight_5y_prior_endpoint_quartiles",
                  "median_weight_5y_prior_endpoint_kg_outliers_removed_quartiles","avg_weight_5y_prior_endpoint_kg_biological","median_weight_5y_prior_endpoint_any_records_kg_outliers_removed_biological",
                  "median_weight_5y_prior_endpoint_kg_outliers_removed_biological"
                  ))

#now recode the new weight data into quartiles
cuts <- quantile(
  cohort_abr_updated_weight2$median_weight_5y_prior_endpoint_updated,
  probs = c(0, 0.25, 0.5, 0.75, 1),
  na.rm = TRUE
)

labels <- c(
  paste0(round(cuts[1], 1), "–", round(cuts[2], 1), " kg"),
  paste0(round(cuts[2], 1), "–", round(cuts[3], 1), " kg"),
  paste0(round(cuts[3], 1), "–", round(cuts[4], 1), " kg"),
  paste0(round(cuts[4], 1), "–", round(cuts[5], 1), " kg")
)

cohort_abr_updated_weight2 <- cohort_abr_updated_weight2%>%
  mutate(
    weight_quartile = case_when(
      is.na(median_weight_5y_prior_endpoint_updated) ~ "not_recorded",
      TRUE ~ as.character(cut(
        median_weight_5y_prior_endpoint_updated,
        breaks = cuts,
        include.lowest = TRUE,
        labels = labels
      ))
    ),
    weight_quartile = factor(
      weight_quartile,
      levels = c(labels, "not_recorded")
    )
  )


#also creating a 'biological' weight variable
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    weight_biological = case_when(
      is.na(median_weight_5y_prior_endpoint_updated) ~ "not_recorded",
      median_weight_5y_prior_endpoint_updated <= 25 ~ "≤25 kg",
      median_weight_5y_prior_endpoint_updated <= 35 ~ ">25–≤35 kg",
      median_weight_5y_prior_endpoint_updated > 35 ~ ">35 kg"
    ),
    weight_biological = factor(
      weight_biological,
      levels = c(
        "≤25 kg",
        ">25–≤35 kg",
        ">35 kg",
        "not_recorded"
      )
    )
  )
```

Converting to factors for analysis:

```{r}

cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    across(
      everything(),
      as.character
    )
  ) %>%
  mutate(
    across(
      everything(),
      ~ na_if(.x, "n/a")
    )
  ) %>%
  mutate(
    across(
      everything(),
      ~ na_if(.x, "NA")
    )
  ) %>%
  mutate(
    across(
      everything(),
      ~ na_if(.x, "nr")
    )
  ) %>%
  mutate(
    across(
      everything(),
      ~ na_if(.x, "not recorded")
    )
  ) %>%
  mutate(
    across(
      everything(),
      ~ na_if(.x, "")
    )  ) %>%
  mutate(
    across(
      everything(),
      ~ replace_na(.x, "not_recorded")
    )
  )

# Factor version of outcome for descriptive tables
cohort_abr_updated_weight2$status_factor <- factor(
  cohort_abr_updated_weight2$had_HSA
)

# Numeric binary outcome for logistic regression
cohort_abr_updated_weight2$status <- as.numeric(
  as.character(cohort_abr_updated_weight2$had_HSA)
)

cohort_abr_updated_weight2$had_HSA <- as.numeric(
  as.character(cohort_abr_updated_weight2$had_HSA)
)

# Convert survival time back to numeric
cohort_abr_updated_weight2$survival_time_months <- as.numeric(
  as.character(cohort_abr_updated_weight2$survival_time_months)
)


#change spayed while in heat to account for males and entire females
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    spayed_while_in_heat = case_when(
      sex == "M" ~ "Male",
      sex_status == "Intact Female" ~ "Intact Female",
      spayed_while_in_heat == "0" ~ "0",
      spayed_while_in_heat == "1" ~ "1",
      TRUE ~ "not_recorded"
    )
  )


# Convert heats before neutering to numeric temporarily
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    heats_before_neutering_numeric = case_when(
      heats_before_neutering == "not_recorded" ~ NA_real_,
      TRUE ~ suppressWarnings(
        as.numeric(as.character(heats_before_neutering))
      )
    )
  )

# Create a temporary vector for inspecting non-zero recorded values
heats <- cohort_abr_updated_weight2 %>%
  filter(
    !is.na(heats_before_neutering_numeric),
    heats_before_neutering_numeric != 0
  ) %>%
  pull(heats_before_neutering_numeric)


# Function to calculate quartile or tercile breaks dynamically
get_breaks <- function(x, probs) {

  breaks <- quantile(
    x,
    probs = probs,
    na.rm = TRUE
  )

  if (length(unique(breaks)) < length(probs)) {

    message(
      "Quartile breaks are not unique; switching to terciles."
    )

    breaks <- quantile(
      x,
      probs = c(0, 1 / 3, 2 / 3, 1),
      na.rm = TRUE
    )
  }

  breaks
}


breaks <- get_breaks(
  heats,
  c(0, 0.25, 0.5, 0.75, 1)
)


labels <- paste0(
  "≥",
  round(head(breaks, -1), 2),
  " - <",
  round(tail(breaks, -1), 2)
)


# Create manually grouped heats-before-neutering variable
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    heats_before_neutering_tercile = case_when(
      sex == "M" ~ "Male",
      sex_status == "Intact Female" ~ "Intact Female",

      heats_before_neutering_numeric == 0 ~ "0",

      heats_before_neutering_numeric %in% c(
        1, 2
      ) ~ "1-<=2",

      heats_before_neutering_numeric %in% c(
        3, 4
      ) ~ ">2-<=4",

      heats_before_neutering_numeric >= 5 &
        heats_before_neutering_numeric <= 12 ~ ">4-12",

      TRUE ~ "not_recorded"
    )
  )



# Create combined sex-neuter variable
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    sex_neuter = paste(
      sex,
      neuter,
      sep = " "
    )
  )


#------------------------------------------------------------
# CONVERT CHARACTER VARIABLES TO FACTORS
#------------------------------------------------------------

# Convert all remaining character columns to factors.
# Move "not_recorded" to the final level so it is not used
# as the default reference category.

convert_and_relevel <- function(df) {
df %>%
    mutate(
      across(
        where(is.character),
        factor
      )
    ) %>%
    mutate(
      across(
        where(is.factor),
        ~ {
          if ("not_recorded" %in% levels(.x)) {
            forcats::fct_relevel(
              .x,
              "not_recorded",
              after = Inf
            )
          } else {
            .x
          }
        }
      )
    )
}


cohort_abr_updated_weight2 <- convert_and_relevel(
  cohort_abr_updated_weight2
)
#------------------------------------------------------------
# RESTORE CONTINUOUS VARIABLES TO NUMERIC
#------------------------------------------------------------

# These variables must be numeric so that logistic regression fits
# one coefficient per one-unit increase rather than one coefficient
# for every observed value.

continuous_variables <- c(
  "mean_activity_level_overview",

  "whole_activity_mean",
  "early_activity_mean",
  "rest_activity_mean",
  "prior5y_activity_mean",

  "avg_activity_duration_sy1_2",
  "avg_activity_intensity_sy1_2",
  "avg_activity_freq_sy1_2",

  "avg_activity_freq_sy3",
  "avg_activity_intensity_sy3",

  "avg_intensity_5y_prior_endpoint",
  "avg_frequency_5y_prior_endpoint",

  "avg_height_5y_prior_endpoint",

  "median_weight_lifetime_updated",
  "median_weight_5y_prior_endpoint_updated"
)

# Check that all named variables exist
missing_continuous_variables <- setdiff(
  continuous_variables,
  names(cohort_abr_updated_weight2)
)

if (length(missing_continuous_variables) > 0) {

  warning(
    paste(
      "These continuous variables are missing from the dataframe:",
      paste(
        missing_continuous_variables,
        collapse = ", "
      )
    )
  )
}


# Retain only continuous variables that actually exist
continuous_variables_present <- intersect(
  continuous_variables,
  names(cohort_abr_updated_weight2)
)

# Convert factor -> character -> numeric
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    across(
      all_of(continuous_variables_present),
      ~ suppressWarnings(
        as.numeric(as.character(.x))
      )
    )
  )


#------------------------------------------------------------
# CHECK NUMERIC CONVERSION
#------------------------------------------------------------
continuous_variable_classes <- tibble(
  variable = continuous_variables_present,
  variable_class = vapply(
    cohort_abr_updated_weight2[
      continuous_variables_present
    ],
    function(x) {
      paste(
        class(x),
        collapse = "/"
      )
    },
    character(1)
  ),
  number_missing = vapply(
    cohort_abr_updated_weight2[
      continuous_variables_present
    ],
    function(x) {
      sum(is.na(x))
    },
    integer(1)
  ),
  number_unique = vapply(
    cohort_abr_updated_weight2[
      continuous_variables_present
    ],
    function(x) {
      dplyr::n_distinct(
        x,
        na.rm = TRUE
      )
    },
    integer(1)
  )
)
continuous_variable_classes


# Stop the script if any continuous variable is still not numeric
non_numeric_continuous_variables <- continuous_variable_classes %>%
  filter(
    variable_class != "numeric"
  )

if (nrow(non_numeric_continuous_variables) > 0) {

  stop(
    paste(
      "The following intended continuous variables are not numeric:",
      paste(
        non_numeric_continuous_variables$variable,
        collapse = ", "
      )
    )
  )
}


#------------------------------------------------------------
# SET MANUAL FACTOR LEVELS
#------------------------------------------------------------
# Heats before neutering
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    heats_before_neutering_tercile = factor(
      heats_before_neutering_tercile,
      levels = c(
        "0",
        "1-<=2",
        ">2-<=4",
        ">4-12",
        "Intact Female",
        "Male",
        "not_recorded"
      )
    )
  )


# Neutering reason
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    neutered_reason = factor(
      neutered_reason,
      levels = c(
        "Elective",
        "Medical Reason",
        "Entire",
        "Unknown",
        "not_recorded"
      )
    )
  )
# Mode BCS
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    mode_purina_BCS_5y_prior_endpoint = factor(
      mode_purina_BCS_5y_prior_endpoint,
      levels = c(
        "5",
        "3",
        "4",
        "6",
        "7",
        "8",
        "9",
        "1",
        "2",
        "not_recorded"
      )
    )
  )
# Reclassify mode BCS
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    mode_BCS_classify = case_when(
      as.character(
        mode_purina_BCS_5y_prior_endpoint
      ) %in% c(
        "1", "2", "3"
      ) ~ "underweight",

      as.character(
        mode_purina_BCS_5y_prior_endpoint
      ) %in% c(
        "4", "5"
      ) ~ "ideal",

      as.character(
        mode_purina_BCS_5y_prior_endpoint
      ) %in% c(
        "6", "7", "8", "9"
      ) ~ "overweight",

      TRUE ~ "not_recorded"
    ),
    mode_BCS_classify = factor(
      mode_BCS_classify,
      levels = c(
        "ideal",
        "underweight",
        "overweight",
        "not_recorded"
      )
    )
  )
# Age quartiles
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    age_at_final_date_quartiles = factor(
      age_at_final_date_quartiles,
      levels = c(
        "0.74-<8.47 years",
        ">=8.47-<10.68 years",
        ">=10.68-<11.64 years",
        ">=11.64-14.51 years",
        "not_recorded"
      )
    )
  )


# Biological age groups
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    age_at_final_date_bio = factor(
      age_at_final_date_bio,
      levels = c(
        ">5-<8y",
        "<5y",
        ">8-<11y",
        ">11-<13y",
        ">13y",
        "not_recorded"
      )
    )
  )


# Golden Retriever Lifetime Study age groups
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    age_at_final_date_GB = factor(
      age_at_final_date_GB,
      levels = c(
        ">7-<9y",
        "<5y",
        ">5-<7y",
        ">9-<11y",
        ">11-<13y",
        ">13y",
        "not_recorded"
      )
    )
  )


# Height quartiles
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    avg_height_5y_prior_endpoint_quartiles = factor(
      avg_height_5y_prior_endpoint_quartiles,
      levels = c(
        "49.80-<56.30cm",
        ">=56.30-<58.29cm",
        ">=58.29-<60.32cm",
        ">=60.32cm-67.73cm",
        "not_recorded"
      )
    )
  )
# Average BCS quartiles
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    avg_purina_BCS_5y_prior_endpoint_quartiles = factor(
      avg_purina_BCS_5y_prior_endpoint_quartiles,
      levels = c(
        "3.3-<5",
        ">=5-<5.28",
        ">=5.28-<5.86",
        ">=5.86 - 9",
        "not_recorded"
      )
    )
  )


# Age at neutering quartiles
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    age_at_neuter_years_quartiles = factor(
      age_at_neuter_years_quartiles,
      levels = c(
        "0.15-<0.52y",
        ">=0.52-<0.83y",
        ">=0.83-2.44y",
        ">=2.44-9.59y",
        "Entire",
        "not_recorded"
      )
    )
  )
# Number bred before neutering
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    num_breeding_before_neutering_terciles = factor(
      num_breeding_before_neutering_terciles,
      levels = c(
        ">0-<3",
        ">=3-<9",
        ">=9-142",
        "not_bred",
        "not_recorded"
      )
    )
  )


# NSAID exposure
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    NSAID_5y_quantiles = factor(
      NSAID_5y_quantiles,
      levels = c(
        "no_NSAID_admin",
        "Q1(1)",
        "Q2(1-2)",
        "Q3(2-3)",
        "Q4(3-11)",
        "no_med_use_recorded",
        "not_recorded"
      )
    )
  )
# Endoparasiticide exposure
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    endo_5y_quantiles = factor(
      endo_5y_quantiles,
      levels = c(
        "no_endo_admin",
        "Q1(1-2)",
        "Q2(2-3)",
        "Q3(3-4)",
        "Q4(4-17)",
        "no_med_use_recorded",
        "not_recorded"
      )
    )
  )


# Ectoparasiticide exposure
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    ecto_5y_quantiles = factor(
      ecto_5y_quantiles,
      levels = c(
        "no_ecto_admin",
        "Q1(1)",
        "Q2(1-2)",
        "Q3(2-4)",
        "Q4(4-17)",
        "no_med_use_recorded",
        "not_recorded"
      )
    )
  )


# Biological weight categories
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    weight_biological = factor(
      weight_biological,
      levels = c(
        "≤25 kg",
        ">25–≤35 kg",
        ">35 kg",
        "not_recorded"
      )
    )
  )


# Weight quartiles
if ("weight_quartile" %in% names(cohort_abr_updated_weight2)) {

  current_weight_quartile_levels <- levels(
    factor(
      cohort_abr_updated_weight2$weight_quartile
    )
  )

  recorded_weight_quartile_levels <- setdiff(
    current_weight_quartile_levels,
    "not_recorded"
  )

  cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
    mutate(
      weight_quartile = factor(
        weight_quartile,
        levels = c(
          recorded_weight_quartile_levels,
          "not_recorded"
        )
      )
    )
}


#------------------------------------------------------------
# MOVE MEDICATION MISSING CATEGORY TO FINAL LEVEL
#------------------------------------------------------------

# Prefer named variables rather than fixed column positions because column
# positions may change when variables are added or removed.

medication_variables <- unique(
  unlist(
    list(
      grep(
        "^half_",
        names(cohort_abr_updated_weight2),
        value = TRUE
      ),
      grep(
        "^quartile_",
        names(cohort_abr_updated_weight2),
        value = TRUE
      ),
      c(
        "NSAID_5y_quantiles",
        "endo_5y_quantiles",
        "ecto_5y_quantiles"
      )
    )
  )
)


medication_variables <- intersect(
  medication_variables,
  names(cohort_abr_updated_weight2)
)


for (variable in medication_variables) {

  if (is.factor(cohort_abr_updated_weight2[[variable]])) {

    current_levels <- levels(
      cohort_abr_updated_weight2[[variable]]
    )

    if (
      "no_medication_records_for_this_time_period" %in%
        current_levels
    ) {

      cohort_abr_updated_weight2[[variable]] <- factor(
        cohort_abr_updated_weight2[[variable]],
        levels = c(
          setdiff(
            current_levels,
            "no_medication_records_for_this_time_period"
          ),
          "no_medication_records_for_this_time_period"
        )
      )
    }
  }
}


#------------------------------------------------------------
# FINAL VALIDATION OF CONTINUOUS VARIABLES
#------------------------------------------------------------

# Check again after all factor-level changes to ensure no continuous
# variables have accidentally been converted back to factors.

final_continuous_check <- tibble(
  variable = continuous_variables_present,
  variable_class = vapply(
    cohort_abr_updated_weight2[
      continuous_variables_present
    ],
    function(x) {
      paste(
        class(x),
        collapse = "/"
      )
    },
    character(1)
  ),
  number_unique = vapply(
    cohort_abr_updated_weight2[
      continuous_variables_present
    ],
    function(x) {
      dplyr::n_distinct(
        x,
        na.rm = TRUE
      )
    },
    integer(1)
  ),
  number_missing = vapply(
    cohort_abr_updated_weight2[
      continuous_variables_present
    ],
    function(x) {
      sum(is.na(x))
    },
    integer(1)
  )
)

final_continuous_check


if (
  any(
    final_continuous_check$variable_class != "numeric"
  )
) {

  stop(
    paste(
      "Continuous-variable conversion failed for:",
      paste(
        final_continuous_check$variable[
          final_continuous_check$variable_class !=
            "numeric"
        ],
        collapse = ", "
      )
    )
  )
}


# Specific check for the activity variable that previously caused problems
class(
  cohort_abr_updated_weight2$avg_activity_freq_sy3
)

is.numeric(
  cohort_abr_updated_weight2$avg_activity_freq_sy3
)

is.factor(
  cohort_abr_updated_weight2$avg_activity_freq_sy3
)


# Expected output:
#
# [1] "numeric"
# [1] TRUE
# [1] FALSE


#------------------------------------------------------------
# OPTIONAL DATAFRAME WITH "not_recorded" REPLACED BY NA
#------------------------------------------------------------

# Use this version if you want not_recorded categories excluded
# automatically by complete-case regression models.

cohort_no_not_rec <- cohort_abr_updated_weight2 %>%
 mutate(
   across(
    where(
       ~ is.character(.x) || is.factor(.x)
      ),
     ~ factor(
        na_if(
         as.character(.x),
          "not_recorded"
      )
     )
    )
  )


# Recreate BCS classification in the no-not-recorded dataframe
cohort_no_not_rec <- cohort_no_not_rec %>%
  mutate(
    mode_BCS_classify = case_when(
      as.character(
        mode_purina_BCS_5y_prior_endpoint
      ) %in% c(
        "1", "2", "3"
      ) ~ "underweight",

      as.character(
        mode_purina_BCS_5y_prior_endpoint
      ) %in% c(
        "4", "5"
      ) ~ "ideal",

      as.character(
        mode_purina_BCS_5y_prior_endpoint
      ) %in% c(
        "6", "7", "8", "9"
      ) ~ "overweight",

      TRUE ~ NA_character_
    ),

    mode_BCS_classify = factor(
      mode_BCS_classify,
      levels = c(
        "ideal",
        "underweight",
        "overweight"
      )
    )
  )


#------------------------------------------------------------
# WRITE COLUMN NAMES
#------------------------------------------------------------

colnames_df <- tibble(
  variable = names(cohort_no_not_rec)
)



colnames_df <- colnames(cohort_no_not_rec)
write.csv(colnames_df,"C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_col_names.csv")
```

## Regression analyses

**Fixed/forced covariates ie. in ALL models**

```{r}
exposome_data <- cohort_abr_updated_weight2 %>%
  mutate(
    had_HSA = as.integer(as.character(had_HSA))
  )

#forced covariates ie. the model will keep these in always and the exposures will be added in and out of different models
forced_covariates <- c(
  "age_at_final_date_bio",
  "sex",
  "neuter"
  #or possibly sex_status which is sex and neuter status combined
)
```

### **Exposome groups**

#### Built environment and socioeconomic

#### Household chemical exposure

#### Lifestyle/sleep location

#### Dog activity

#### Body condition/weight

#### Reproductive exposures

#### Medication types

##### NSAIDs(anti-inflams)

##### Antimicrobials

##### Dietary supplements

##### Enteric medications

##### Hormones

##### Immunologicals

##### Neurologicals

##### Anti-parasites

##### Anti-septics

##### Cardiorespiratory

##### Diuretics

##### Antihistamines

##### Antivirals

#### Comorbidities

##### Cardiovascular

##### Orthopaedic

##### Immune mediated

##### Chronic inflammatory

##### Parasitism

```{r}
exposure_families <- list(

  #------------------------------------------------------------
  # BUILT ENVIRONMENT AND SOCIOECONOMIC
  #------------------------------------------------------------

  built_environment_socioeconomic = c(
    "MDI_quintile",
    "owner_MDI_quintile",
    "area_type_mode_5yrs_prior",
    "house_type_mode_5yrs_prior",
    "water_source_mode_5yrs_prior",
    "pipes_metal_any_mode_5yrs_prior",
    "pipes_plastic_any_mode_5yrs_prior",
    "heating_fuel_primary_mode_5yrs_prior",
    "cooking_fuel_primary_mode_5yrs_prior",
    "heating_fuel_secondary_mode_5yrs_prior",
    "cooking_fuel_secondary_mode_5yrs_prior",
    "region_mode_5y_prior"
  ),

  #------------------------------------------------------------
  # HOUSEHOLD CHEMICAL EXPOSURE
  #------------------------------------------------------------

  household_chemical_exposure = c(
    "use_aerosol_study_years_early_life",
    "use_aerosol_study_years_rest_of_life",
    "use_aerosol_study_years_whole_life",
    "use_aerosol_exposed_5yrs_prior",

    "use_air_cleaner_study_years_early_life",
    "use_air_cleaner_study_years_rest_of_life",
    "use_air_cleaner_study_years_whole_life",
    "use_air_cleaner_exposed_5yrs_prior",

    "use_hepa_filter_study_years_early_life",
    "use_hepa_filter_study_years_rest_of_life",
    "use_hepa_filter_study_years_whole_life",
    "use_hepa_filter_exposed_5yrs_prior",

    "use_moth_balls_study_years_early_life",
    "use_moth_balls_study_years_rest_of_life",
    "use_moth_balls_study_years_whole_life",
    "use_moth_balls_exposed_5yrs_prior",

    "use_incense_or_candles_study_years_early_life",
    "use_incense_or_candles_study_years_rest_of_life",
    "use_incense_or_candles_study_years_whole_life",
    "use_incense_or_candles_exposed_5yrs_prior",

    "smoke_exposure_study_years_early_life",
    "smoke_exposure_study_years_rest_of_life",
    "smoke_exposure_study_years_whole_life",
    "smoke_exposure_exposed_5yrs_prior",

    "hours_of_smoke_early_life_total_dosage_terciles",
    "hours_of_smoke_rest_of_life_total_dosage_terciles",
    "hours_of_smoke_whole_life_total_dosage_terciles",
    "hours_of_smoke_exposure_hours_5yrs_prior_terciles",

    "any_treated_weeds_study_years_early_life",
    "any_treated_weeds_study_years_rest_of_life",
    "any_treated_weeds_study_years_whole_life",
    "any_treated_weeds_exposed_5yrs_prior",

    "any_treated_insects_study_years_early_life",
    "any_treated_insects_study_years_rest_of_life",
    "any_treated_insects_study_years_whole_life",
    "any_treated_insects_exposed_5yrs_prior",

    "any_treated_fertilizer_study_years_early_life",
    "any_treated_fertilizer_study_years_rest_of_life",
    "any_treated_fertilizer_study_years_whole_life",
    "any_treated_fertilizer_exposed_5yrs_prior"
  ),

  #------------------------------------------------------------
  # LIFESTYLE / SLEEP LOCATION
  #------------------------------------------------------------

  lifestyle_sleep_location = c(
    "majority_sleep_location_lifetime",
    "main_lifestyle_category",
    "main_lifestyle_category_mode_5yrs_prior",
    "sleep_location_mode_5yrs_prior",

    "in.the.garage_YN_study_years_early_life",
    "in.the.garage_YN_study_years_rest_of_life",
    "in.the.garage_YN_study_years_whole_life",

    "in.the.house_YN_study_years_early_life",
    "in.the.house_YN_study_years_rest_of_life",
    "in.the.house_YN_study_years_whole_life",

    "outside_YN_study_years_early_life",
    "outside_YN_study_years_rest_of_life",
    "outside_YN_study_years_whole_life",

    "in.the.house_YN_sum_early_life_quartile",
    "in.the.house_YN_sum_rest_of_life_quartile",
    "in.the.house_YN_sum_whole_life_quartile"
  ),

  #------------------------------------------------------------
  # DOG ACTIVITY
  #------------------------------------------------------------

  dog_activity = c(

  # Overall activity
  "mean_activity_level_overview",
  "mode_activity_level_overview",

  # Time spent exercising
  "USDA_30min",
  "activity_1h_min",
  "activity_KC_2h_min",
  "avg_activity_duration_sy1_2",
  "avg_activity_duration_sy1_2_quartile",

  # Intensity
  "avg_activity_intensity_sy1_2",
  "avg_activity_intensity_sy3",
  "avg_intensity_5y_prior_endpoint",
  "avg_activity_intensity_sy3_quartile",
  "avg_intensity_5y_prior_endpoint_quartile",
  "mode_intensity_5y_prior_endpoint",

  # Frequency
  "avg_activity_freq_sy1_2",
  "avg_activity_freq_sy3",
  "avg_frequency_5y_prior_endpoint",
  "avg_activity_freq_sy1_2_quartile",
  "avg_activity_freq_sy3_quartile",
  "avg_frequency_5y_prior_endpoint_quartile",
  "mode_frequency_5y_prior_endpoint",

  # Whole-life summaries
  "whole_activity_mode",
  "whole_activity_mean",
  "whole_activity_median",
  "whole_activity_min",
  "whole_activity_max",

  # Early-life summaries
  "early_activity_mode",
  "early_activity_mean",
  "early_activity_median",
  "early_activity_min",
  "early_activity_max",

  # Rest-of-life summaries
  "rest_activity_mode",
  "rest_activity_mean",
  "rest_activity_median",
  "rest_activity_min",
  "rest_activity_max",

  # Five-year-prior summaries
  "prior5y_activity_mode",
  "prior5y_activity_mean",
  "prior5y_activity_median",
  "prior5y_activity_min",
  "prior5y_activity_max"

  ),

  #------------------------------------------------------------
  # BODY CONDITION / WEIGHT
  #------------------------------------------------------------

  body_condition_weight = c(
    "weight_biological",
    "weight_quartile",
    "avg_height_5y_prior_endpoint_quartiles",
   # "avg_purina_BCS_5y_prior_endpoint_quartiles", <- doesnt really make sense to do avg when its a whole number value 
    "mode_BCS_classify"
  ),

  #------------------------------------------------------------
  # REPRODUCTIVE EXPOSURES
  #------------------------------------------------------------

  reproductive_exposures = c(
    "age_at_neuter_years_quartiles",
    "mismating_ever",
    "neutered_reason",
    "neutered_ever_bred",
    "spayed_while_in_heat",
    "spayed_while_pregnant",
    "prevent_heat_ever",
    "no_pregnancy_ever",
    "heats_before_neutering_terciles",
    "preg_pre_neuter",
    "mated_pre_neuter",
    "num_breeding_before_neutering_terciles",
    "ever_bred_from"
  ),

  #------------------------------------------------------------
  # INDIVIDUAL MEDICATION FAMILIES
  #------------------------------------------------------------

  nsaids_anti_inflammatories = c(
    "NSAID_ever",
    "NSAID_5y",
    "NSAID_5y_quantiles",
    "half_anti_inflam_lifetime",
    "quartile_anti_inflam_lifetime",
    "half_anti_inflam_5y_prior_endpoint",
    "quartile_anti_inflam_5y_prior_endpoint"
  ),

  antimicrobials = c(
    "half_anti_microbials_lifetime",
    "quartile_anti_microbials_lifetime",
    "half_anti_microbials_5y_prior_endpoint",
    "quartile_anti_microbials_5y_prior_endpoint"
  ),

  dietary_supplements = c(
    "half_dietary_supplements_lifetime",
    "quartile_dietary_supplements_lifetime",
    "half_dietary_supplements_5y_prior_endpoint",
    "quartile_dietary_supplements_5y_prior_endpoint"
  ),

  enteric_medications = c(
    "half_enteric_lifetime",
    "quartile_enteric_lifetime",
    "half_enteric_5y_prior_endpoint",
    "quartile_enteric_5y_prior_endpoint"
  ),

  hormones = c(
    "half_hormones_and_related_lifetime",
    "quartile_hormones_and_related_lifetime",
    "half_hormones_and_related_5y_prior_endpoint",
    "quartile_hormones_and_related_5y_prior_endpoint",
    "steroid_ever",
    "steroid_5y"
  ),

  immunologicals = c(
    "half_immunologicals_lifetime",
    "quartile_immunologicals_lifetime",
    "half_immunologicals_5y_prior_endpoint",
    "quartile_immunologicals_5y_prior_endpoint"
  ),

  neurologicals = c(
    "half_neurological_lifetime",
    "quartile_neurological_lifetime",
    "half_neurological_5y_prior_endpoint",
    "quartile_neurological_5y_prior_endpoint"
  ),

  anti_parasites = c(
    "half_anti_parasites_lifetime",
    "quartile_anti_parasites_lifetime",
    "half_anti_parasites_5y_prior_endpoint",
    "quartile_anti_parasites_5y_prior_endpoint",
    "ecto_5y_quantiles",
    "endo_5y_quantiles",
    "fenbendazole_ever",
    "fenbendazole_5y"
  ),

  anti_septics = c(
    "half_anti_septics_lifetime",
    "quartile_anti_septics_lifetime",
    "half_anti_septics_5y_prior_endpoint",
    "quartile_anti_septics_5y_prior_endpoint"
  ),

  cardiorespiratory_medications = c(
    "half_cardio_respiratory_lifetime",
    "quartile_cardio_respiratory_lifetime",
    "half_cardio_respiratory_5y_prior_endpoint",
    "quartile_cardio_respiratory_5y_prior_endpoint"
  ),

  diuretics = c(
    "half_diuretics_lifetime",
    "quartile_diuretics_lifetime",
    "half_diuretics_5y_prior_endpoint",
    "quartile_diuretics_5y_prior_endpoint"
  ),

  antihistamines = c(
    "half_anti_hist_lifetime",
    "quartile_anti_hist_lifetime",
    "half_anti_hist_5y_prior_endpoint",
    "quartile_anti_hist_5y_prior_endpoint"
  ),

  antivirals = c(
    "half_anti_viral_lifetime",
    "quartile_anti_viral_lifetime",
    "half_anti_viral_5y_prior_endpoint",
    "quartile_anti_viral_5y_prior_endpoint"
  ),

  #------------------------------------------------------------
  # COMORBIDITIES: OVERALL
  #------------------------------------------------------------

  comorbidities = c(
    "chronic_inflam_present_lifetime",
    "number_comorbs_present_lifetime",
    "chronic_inflam_present_5y",
    "number_comorbs_present_5y",
    "chronic_inflam_present_no_min_year",
    "number_comorbs_present_no_min_year",
    "lifelong_comorb_present.x",
    "number_lifelong_comorbs_present"
  ),

  #------------------------------------------------------------
  # INDIVIDUAL COMORBIDITY FAMILIES
  #------------------------------------------------------------

  cardiovascular_comorbidity = c(
    "cardiovascular",
    "lifelong_cardiovascular",
    "comorb_cardiovascular_all"
  ),

  orthopaedic_comorbidity = c(
    "orthopaedic",
    "lifelong_orthopaedic",
    "comorb_orthopaedic_all"
  ),

  immune_mediated_comorbidity = c(
    "immune_mediated",
    "lifelong_immune_mediated",
    "comorb_immune_med_all"
  ),

  chronic_inflammatory_comorbidity = c(
    "chronic_inflammatory",
    "lifelong_chronic_inflammatory",
    "lifelong_inflammatory_other",
    "comorb_chronic_inflam_all"
  ),

  #------------------------------------------------------------
  # PARASITISM
  #------------------------------------------------------------

  parasitism = c(
    "intestinal_parasitism",
    "tick_borne_parasitism",
    "other_parasitism",
    "all_parasitism"
  )
)
```

```{r}
exposure_families_requested <- exposure_families

missing_family_variables <- imap_dfr(
  exposure_families_requested,
  function(variables, family) {

    missing <- setdiff(
      variables,
      names(exposome_data)
    )

    tibble(
      exposure_family = family,
      missing_variable = missing
    )
  }
)

missing_family_variables
```

```{r}
analysis_df <- cohort_abr_updated_weight2
outcome_var <- "had_HSA"
requested_exposure_families <- exposure_families
missing_exposure_variables <- imap_dfr(
  requested_exposure_families,
  function(variables, family) {

    missing_variables <- setdiff(
      variables,
      names(analysis_df)
    )

    tibble(
      exposure_family = family,
      variable = missing_variables
    )
  }
)
exposure_families <- map(
  requested_exposure_families,
  ~ intersect(.x, names(analysis_df))
)
exposure_families <- exposure_families[
  lengths(exposure_families) > 0
]
tibble(
  exposure_family = names(exposure_families),
  number_variables = lengths(exposure_families)
)
```

```{r}
screen_exposure <- function(
    data,
    exposure,
    outcome = "had_HSA"
) {

  x <- data[[exposure]]
  y <- data[[outcome]]

  complete <- !is.na(x) & !is.na(y)

  x_complete <- x[complete]
  y_complete <- y[complete]

  tibble(
    exposure = exposure,
    variable_class = paste(class(x), collapse = "/"),
    n_total = nrow(data),
    n_observed = sum(!is.na(x)),
    n_missing = sum(is.na(x)),
    percent_missing = round(
      100 * mean(is.na(x)),
      1
    ),
    n_complete_outcome_exposure = sum(complete),
    cases_complete = sum(y_complete == 1),
    controls_complete = sum(y_complete == 0),
    n_unique_observed = n_distinct(x_complete),
    suitable_for_model = (
      sum(complete) > 0 &&
      n_distinct(x_complete) >= 2 &&
      n_distinct(y_complete) == 2
    )
  )
}
```

```{r}
exposure_screening <- imap_dfr(
  exposure_families,
  function(variables, family) {

    map_dfr(
      variables,
      ~ screen_exposure(
        data = analysis_df,
        exposure = .x,
        outcome = outcome_var
      )
    ) %>%
      mutate(
        exposure_family = family,
        .before = 1
      )
  }
)

exposure_screening %>%
  filter(!suitable_for_model)
```

```{r}
usable_exposure_families <- split(
  exposure_screening$exposure[
    exposure_screening$suitable_for_model
  ],
  exposure_screening$exposure_family[
    exposure_screening$suitable_for_model
  ]
)
```

```{r}
#ensure the numeric activity columns are not being turned back into factors again!
# Identify every column containing "mean" or "avg" in its name
mean_avg_variables <- names(analysis_df)[
  stringr::str_detect(
    names(analysis_df),
    stringr::regex(
      "mean|avg",
      ignore_case = TRUE
    )
  )
]

# Review which columns have been selected
mean_avg_variables


# Convert selected columns from factor/character to numeric
#
# "not_recorded" and other non-numeric values become NA.
# factor -> character -> numeric prevents factor level numbers
# being used accidentally.

analysis_df <- analysis_df %>%
  dplyr::mutate(
    dplyr::across(
      dplyr::all_of(mean_avg_variables),
      ~ suppressWarnings(
        as.numeric(
          dplyr::na_if(
            as.character(.x),
            "not_recorded"
          )
        )
      )
    )
  )


#------------------------------------------------------------
# CHECK THAT CONVERSION WORKED
#------------------------------------------------------------

mean_avg_variable_check <- tibble::tibble(
  variable = mean_avg_variables,

  variable_class = vapply(
    analysis_df[mean_avg_variables],
    function(x) {
      paste(
        class(x),
        collapse = "/"
      )
    },
    character(1)
  ),

  number_observed = vapply(
    analysis_df[mean_avg_variables],
    function(x) {
      sum(!is.na(x))
    },
    integer(1)
  ),

  number_missing = vapply(
    analysis_df[mean_avg_variables],
    function(x) {
      sum(is.na(x))
    },
    integer(1)
  ),

  number_unique = vapply(
    analysis_df[mean_avg_variables],
    function(x) {
      dplyr::n_distinct(
        x,
        na.rm = TRUE
      )
    },
    integer(1)
  )
)

mean_avg_variable_check


# Specifically confirm the problem variable
class(
  analysis_df$avg_activity_freq_sy3
)

is.numeric(
  analysis_df$avg_activity_freq_sy3
)

is.factor(
  analysis_df$avg_activity_freq_sy3
)

```

\

```{r}
make_exposure_counts <- function(
    data,
    exposure,
    outcome = "had_HSA"
) {

  data %>%
    transmute(
      exposure_level = as.character(.data[[exposure]]),
      outcome_value = .data[[outcome]]
    ) %>%
    mutate(
      exposure_level = replace_na(
        exposure_level,
        "missing"
      ),
      outcome_value = as.character(outcome_value)
    ) %>%
    count(
      exposure_level,
      outcome_value,
      name = "n",
      .drop = FALSE
    ) %>%
    pivot_wider(
      names_from = outcome_value,
      values_from = n,
      values_fill = 0,
      names_prefix = "outcome_"
    ) %>%
    mutate(
      exposure = exposure,
      .before = 1
    )
}
```

```{r}
exposure_level_counts <- imap_dfr(
  usable_exposure_families,
  function(variables, family) {

    map_dfr(
      variables,
      ~ make_exposure_counts(
        data = analysis_df,
        exposure = .x,
        outcome = outcome_var
      )
    ) %>%
      mutate(
        exposure_family = family,
        .before = 1
      )
  }
)
```

```{r}
make_exposure_counts <- function(
    data,
    exposure,
    outcome = "had_HSA"
) {

  # Treat genuinely continuous variables as one exposure,
  # rather than producing one count row per numeric value
  if (is.numeric(data[[exposure]])) {

    data %>%
      dplyr::summarise(
        exposure = exposure,
        exposure_level = "continuous",
        outcome_0 = sum(
          .data[[outcome]] == 0 &
            !is.na(.data[[exposure]]),
          na.rm = TRUE
        ),
        outcome_1 = sum(
          .data[[outcome]] == 1 &
            !is.na(.data[[exposure]]),
          na.rm = TRUE
        ),
        missing = sum(
          is.na(.data[[exposure]])
        )
      )

  } else {

    # Categorical variables retain one row per exposure level
    data %>%
      dplyr::transmute(
        exposure_level = as.character(
          .data[[exposure]]
        ),
        outcome_value = .data[[outcome]]
      ) %>%
      dplyr::mutate(
        exposure_level = tidyr::replace_na(
          exposure_level,
          "missing"
        ),
        outcome_value = as.character(
          outcome_value
        )
      ) %>%
      dplyr::count(
        exposure_level,
        outcome_value,
        name = "n",
        .drop = FALSE
      ) %>%
      tidyr::pivot_wider(
        names_from = outcome_value,
        values_from = n,
        values_fill = 0,
        names_prefix = "outcome_"
      ) %>%
      dplyr::mutate(
        exposure = exposure,
        missing = NA_integer_,
        .before = 1
      )
  }
}


#------------------------------------------------------------
# CREATE EXPOSURE LEVEL COUNTS
#------------------------------------------------------------

exposure_level_counts <- purrr::imap_dfr(
  usable_exposure_families,
  function(variables, family) {

    purrr::map_dfr(
      variables,
      ~ make_exposure_counts(
        data = analysis_df,
        exposure = .x,
        outcome = outcome_var
      )
    ) %>%
      dplyr::mutate(
        exposure_family = family,
        .before = 1
      )
  }
)


exposure_level_counts
```

```{r}
exposure_level_counts

```

```{r}
write_csv(
  exposure_level_counts,
  "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/exposome_exposure_level_counts.csv"
)
```

### Univariable

```{r}
fit_crude_exposure_model <- function(
    data,
    exposure,
    outcome = "had_HSA"
) {

  required_variables <- c(
    outcome,
    exposure
  )

  missing_variables <- setdiff(
    required_variables,
    names(data)
  )

  if (length(missing_variables) > 0) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = NA_character_,
        n = NA_integer_,
        cases = NA_integer_,
        controls = NA_integer_,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = NA_real_,
        overall_LRT_df = NA_real_,
        overall_LRT_p = NA_real_,
        AIC = NA_real_,
        converged = FALSE,
        warning = paste(
          "Missing variable(s):",
          paste(
            missing_variables,
            collapse = ", "
          )
        )
      )
    )
  }

  model_data <- data %>%
    dplyr::select(
      dplyr::all_of(
        required_variables
      )
    ) %>%
    tidyr::drop_na()

  n_model <- nrow(model_data)

  n_cases <- sum(
    model_data[[outcome]] == 1,
    na.rm = TRUE
  )

  n_controls <- sum(
    model_data[[outcome]] == 0,
    na.rm = TRUE
  )

  outcome_variation <- dplyr::n_distinct(
    model_data[[outcome]]
  )

  exposure_variation <- dplyr::n_distinct(
    model_data[[exposure]]
  )

  if (
    n_model == 0 ||
      outcome_variation < 2 ||
      exposure_variation < 2
  ) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = NA_character_,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = NA_real_,
        overall_LRT_df = NA_real_,
        overall_LRT_p = NA_real_,
        AIC = NA_real_,
        converged = FALSE,
        warning = "Insufficient variation in outcome or exposure"
      )
    )
  }

  full_formula <- stats::reformulate(
    termlabels = exposure,
    response = outcome
  )

  null_formula <- stats::as.formula(
    paste0(outcome, " ~ 1")
  )

  warning_messages <- character()

  full_model <- tryCatch(

    withCallingHandlers(

      stats::glm(
        formula = full_formula,
        data = model_data,
        family = stats::binomial()
      ),

      warning = function(w) {

        warning_messages <<- c(
          warning_messages,
          conditionMessage(w)
        )

        invokeRestart("muffleWarning")
      }
    ),

    error = function(e) {
      e
    }
  )

  if (inherits(full_model, "error")) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = NA_character_,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = NA_real_,
        overall_LRT_df = NA_real_,
        overall_LRT_p = NA_real_,
        AIC = NA_real_,
        converged = FALSE,
        warning = conditionMessage(full_model)
      )
    )
  }

  null_model <- tryCatch(

    stats::glm(
      formula = null_formula,
      data = model_data,
      family = stats::binomial()
    ),

    error = function(e) {
      e
    }
  )

  if (inherits(null_model, "error")) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = NA_character_,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = NA_real_,
        overall_LRT_df = NA_real_,
        overall_LRT_p = NA_real_,
        AIC = stats::AIC(full_model),
        converged = full_model$converged,
        warning = conditionMessage(null_model)
      )
    )
  }

  LRT_table <- tryCatch(

    stats::anova(
      null_model,
      full_model,
      test = "LRT"
    ),

    error = function(e) {
      NULL
    }
  )

  if (is.null(LRT_table)) {

    overall_LRT_chisq <- NA_real_
    overall_LRT_df <- NA_real_
    overall_LRT_p <- NA_real_

  } else {

    overall_LRT_chisq <- LRT_table$Deviance[2]
    overall_LRT_df <- LRT_table$Df[2]
    overall_LRT_p <- LRT_table$`Pr(>Chi)`[2]
  }

  exposure_object <- model_data[[exposure]]

  if (
    is.factor(exposure_object) ||
      is.character(exposure_object)
  ) {

    factor_exposure <- factor(exposure_object)

    reference_level <- levels(
      factor_exposure
    )[1]

  } else {

    reference_level <- "Per one-unit increase"
  }

  tidy_results <- tryCatch(

    broom::tidy(
      full_model,
      conf.int = TRUE,
      exponentiate = FALSE
    ),

    error = function(e) {
      NULL
    }
  )

  if (is.null(tidy_results)) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = reference_level,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = overall_LRT_chisq,
        overall_LRT_df = overall_LRT_df,
        overall_LRT_p = overall_LRT_p,
        AIC = stats::AIC(full_model),
        converged = full_model$converged,
        warning = "Could not tidy model coefficients"
      )
    )
  }

  tidy_results <- tidy_results %>%
    dplyr::filter(
      term != "(Intercept)"
    )

  if (nrow(tidy_results) == 0) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = reference_level,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = overall_LRT_chisq,
        overall_LRT_df = overall_LRT_df,
        overall_LRT_p = overall_LRT_p,
        AIC = stats::AIC(full_model),
        converged = full_model$converged,
        warning = "No exposure coefficient returned"
      )
    )
  }

  final_warning <- if (
    length(warning_messages) == 0
  ) {

    NA_character_

  } else {

    paste(
      unique(warning_messages),
      collapse = "; "
    )
  }

  tidy_results %>%
    dplyr::transmute(
      exposure = exposure,
      term = term,
      reference_level = reference_level,
      n = n_model,
      cases = n_cases,
      controls = n_controls,
      OR = exp(estimate),
      CI_lower = exp(conf.low),
      CI_upper = exp(conf.high),
      wald_z = statistic,
      wald_p = p.value,
      overall_LRT_chisq = overall_LRT_chisq,
      overall_LRT_df = overall_LRT_df,
      overall_LRT_p = overall_LRT_p,
      AIC = stats::AIC(full_model),
      converged = full_model$converged,
      warning = final_warning
    )
}


```

```{r}
crude_results <- imap_dfr(
  usable_exposure_families,
  function(variables, family) {

    map_dfr(
      variables,
      ~ fit_crude_exposure_model(
        data = analysis_df,
        exposure = .x,
        outcome = outcome_var
      )
    ) %>%
      mutate(
        exposure_family = family,
        model_type = "Crude",
        .before = 1
      )
  }
)
```

```{r}
test_exposure <- usable_exposure_families[[1]][1]

test_exposure
```

```{r}
test_result <- fit_crude_exposure_model(
  data = analysis_df,
  exposure = test_exposure,
  outcome = outcome_var
)

test_result
```

```{r}
crude_results <- purrr::imap_dfr(
  usable_exposure_families,
  function(variables, family) {

    family_results <- purrr::map_dfr(
      variables,
      function(variable_name) {

        fit_crude_exposure_model(
          data = analysis_df,
          exposure = variable_name,
          outcome = outcome_var
        )
      }
    )

    family_results %>%
      dplyr::mutate(
        exposure_family = family,
        model_type = "Crude",
        .before = 1
      )
  }
)
```

```{r}
crude_results %>%
  dplyr::distinct(
    exposure_family,
    exposure
  ) %>%
  nrow()
```

Checking model problems

```{r}
crude_model_problems <- crude_results %>%
  dplyr::filter(
    !converged |
      !is.na(warning)
  ) %>%
  dplyr::distinct(
    exposure_family,
    exposure,
    n,
    cases,
    controls,
    converged,
    warning
  ) %>%
  dplyr::arrange(
    exposure_family,
    exposure
  )
```

One row per exposure variable table

```{r}
crude_overall_tests <- crude_results %>%
  dplyr::distinct(
    exposure_family,
    exposure,
    n,
    cases,
    controls,
    overall_LRT_chisq,
    overall_LRT_df,
    overall_LRT_p,
    AIC,
    converged,
    warning
  )
crude_overall_tests <- crude_overall_tests %>%
  dplyr::group_by(
    exposure_family
  ) %>%
  dplyr::mutate(
    LRT_p_FDR_within_family = stats::p.adjust(
      overall_LRT_p,
      method = "BH"
    )
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    LRT_p_FDR_global = stats::p.adjust(
      overall_LRT_p,
      method = "BH"
    )
  ) %>%
  dplyr::arrange(
    exposure_family,
    overall_LRT_p
  )
```

```{r}
crude_results <- crude_results %>%
  dplyr::left_join(
    crude_overall_tests %>%
      dplyr::select(
        exposure_family,
        exposure,
        LRT_p_FDR_within_family,
        LRT_p_FDR_global
      ),
    by = c(
      "exposure_family",
      "exposure"
    )
  )
```

```{r}
format_p_value <- function(p) {

  dplyr::case_when(
    is.na(p) ~ NA_character_,
    p < 0.001 ~ "<0.001",
    TRUE ~ sprintf("%.3f", p)
  )
}
```

```{r}
crude_results_formatted <- crude_results %>%
  dplyr::mutate(

    OR_95CI = dplyr::case_when(
      is.na(OR) ~ NA_character_,
      TRUE ~ sprintf(
        "%.2f (%.2f–%.2f)",
        OR,
        CI_lower,
        CI_upper
      )
    ),

    Wald_p_display = format_p_value(
      wald_p
    ),

    LRT_p_display = format_p_value(
      overall_LRT_p
    ),

    family_FDR_display = format_p_value(
      LRT_p_FDR_within_family
    ),

    global_FDR_display = format_p_value(
      LRT_p_FDR_global
    )
  ) %>%
  dplyr::select(
    exposure_family,
    exposure,
    term,
    reference_level,
    n,
    cases,
    controls,
    OR_95CI,
    OR,
    CI_lower,
    CI_upper,
    wald_z,
    wald_p,
    Wald_p_display,
    overall_LRT_chisq,
    overall_LRT_df,
    overall_LRT_p,
    LRT_p_display,
    LRT_p_FDR_within_family,
    family_FDR_display,
    LRT_p_FDR_global,
    global_FDR_display,
    AIC,
    converged,
    warning
  ) %>%
  dplyr::arrange(
    exposure_family,
    overall_LRT_p,
    exposure,
    term
  )


#write to csv
write.csv(crude_results_formatted,"C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_log_reg_family_grouped.csv")


```

Table per family

Need to produce prettier tables

-   p\<0.001, values to 2 dp

```{r}
format_p <- function(x, digits = 3) {
  case_when(
    is.na(x) ~ NA_character_,
    x < 0.001 ~ "<0.001",
    TRUE ~ formatC(x, format = "f", digits = digits)
  )
}

crude_results_pretty <- crude_results %>%
  mutate(
    # OR and confidence limits to 3 decimal places
    OR_display = sprintf("%.3f", OR),
    CI_lower_display = sprintf("%.3f", CI_lower),
    CI_upper_display = sprintf("%.3f", CI_upper),

    # Combined OR (95% CI)
    OR_95CI = sprintf(
      "%.3f (%.3f–%.3f)",
      OR,
      CI_lower,
      CI_upper
    ),

    # Other numerical results
    wald_z_display = sprintf("%.3f", wald_z),
    overall_LRT_chisq_display = sprintf("%.3f", overall_LRT_chisq),
    AIC_display = sprintf("%.3f", AIC),

    # P-values
    Wald_p_display = format_p(wald_p),
    LRT_p_display = format_p(overall_LRT_p),
    family_FDR_display = format_p(LRT_p_FDR_within_family),
    global_FDR_display = format_p(LRT_p_FDR_global)
  )


crude_results_pretty 
```

Format further differences:

-   reference level in table not as column and also family grouping

```{r}
library(dplyr)
library(stringr)
library(purrr)
library(tibble)


#===============================================================================
# Formatting functions
#===============================================================================

format_p <- function(x, digits = 3) {
  dplyr::case_when(
    is.na(x)  ~ "",
    x < 0.001 ~ "<0.001",
    TRUE ~ formatC(
      x,
      format = "f",
      digits = digits
    )
  )
}


format_count_percent <- function(count, denominator, digits = 1) {
  dplyr::case_when(
    is.na(count) |
      is.na(denominator) |
      denominator == 0 ~ "",

    TRUE ~ sprintf(
      paste0("%d (%.", digits, "f%%)"),
      as.integer(count),
      100 * count / denominator
    )
  )
}


# Apply identical cleaning to:
# 1. coefficient categories
# 2. count-table categories
# 3. reference categories

pretty_category <- function(x) {
  x <- as.character(x)

  dplyr::case_when(
    is.na(x) ~ NA_character_,

    x == "" |
      x == "continuous" ~
      "Continuous",

    x == "no_med_use_recorded" ~
      "No medication use recorded",

    x == "no_medication_records_for_this_time_period" ~
      "No medication records for this time period",

    TRUE ~ x
  )
}



crude_model_display <- crude_results %>%
  dplyr::mutate(
    exposure = as.character(exposure),
    term = as.character(term),

    # For continuous variables, term usually equals exposure.
    # For factors, remove the exposure name from the start of term.
    Category_raw = dplyr::case_when(
      term == exposure ~ "continuous",

      stringr::str_starts(term, exposure) ~
        stringr::str_sub(
          term,
          start = stringr::str_length(exposure) + 1L
        ),

      TRUE ~ term
    ),

    Category = pretty_category(Category_raw),

    `Crude OR (95% CI)` = dplyr::case_when(
      is.na(OR) |
        is.na(CI_lower) |
        is.na(CI_upper) ~ "",

      TRUE ~ sprintf(
        "%.3f (%.3f–%.3f)",
        OR,
        CI_lower,
        CI_upper
      )
    ),

    `Wald p-value` = format_p(wald_p)
  ) %>%
  dplyr::select(
    exposure_family,
    exposure,
    Category,
    `Crude OR (95% CI)`,
    `Wald p-value`
  )



reference_lookup <- crude_results %>%
  dplyr::filter(
    !is.na(reference_level),
    reference_level != ""
  ) %>%
  dplyr::transmute(
    exposure_family,
    exposure,
    ReferenceCategory = pretty_category(reference_level)
  ) %>%
  dplyr::distinct()




reference_lookup %>%
  dplyr::count(
    exposure_family,
    exposure
  ) %>%
  dplyr::filter(n > 1)


crude_overall_p <- crude_results %>%
  dplyr::group_by(
    exposure_family,
    exposure
  ) %>%
  dplyr::summarise(
    `LRT p-value` = format_p(
      dplyr::first(
        overall_LRT_p[!is.na(overall_LRT_p)],
        default = NA_real_
      )
    ),

    `Family FDR p-value` = format_p(
      dplyr::first(
        LRT_p_FDR_within_family[
          !is.na(LRT_p_FDR_within_family)
        ],
        default = NA_real_
      )
    ),

    `Global FDR p-value` = format_p(
      dplyr::first(
        LRT_p_FDR_global[
          !is.na(LRT_p_FDR_global)
        ],
        default = NA_real_
      )
    ),

    .groups = "drop"
  )


crude_counts_display <- exposure_level_counts %>%
  dplyr::group_by(
    exposure_family,
    exposure
  ) %>%
  dplyr::mutate(
    category_original_order = dplyr::row_number(),

    total_cases = sum(
      outcome_1,
      na.rm = TRUE
    ),

    total_non_cases = sum(
      outcome_0,
      na.rm = TRUE
    ),

    `Cases (%)` = format_count_percent(
      outcome_1,
      total_cases
    ),

    `Non-cases (%)` = format_count_percent(
      outcome_0,
      total_non_cases
    ),

    Category = pretty_category(exposure_level)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::select(
    exposure_family,
    exposure,
    Category,
    category_original_order,
    `Cases (%)`,
    `Non-cases (%)`
  )


family_order_lookup <- tibble::tibble(
  exposure_family = names(exposure_families),
  family_order = seq_along(exposure_families)
)


exposure_order_lookup <- purrr::imap_dfr(
  exposure_families,
  function(variables, family_name) {
    tibble::tibble(
      exposure_family = family_name,
      exposure = variables,
      exposure_order = seq_along(variables)
    )
  }
) %>%
  dplyr::distinct(
    exposure_family,
    exposure,
    .keep_all = TRUE
  )


crude_results_table <- crude_counts_display %>%

  # Add the reference category to every row for each exposure
  dplyr::left_join(
    reference_lookup,
    by = c(
      "exposure_family",
      "exposure"
    )
  ) %>%

  # Add OR and Wald p-value to the relevant non-reference rows
  dplyr::left_join(
    crude_model_display,
    by = c(
      "exposure_family",
      "exposure",
      "Category"
    )
  ) %>%

  # Add the overall LRT and FDR p-values
  dplyr::left_join(
    crude_overall_p,
    by = c(
      "exposure_family",
      "exposure"
    )
  ) %>%

  # Add the desired family and exposure order
  dplyr::left_join(
    family_order_lookup,
    by = "exposure_family"
  ) %>%

  dplyr::left_join(
    exposure_order_lookup,
    by = c(
      "exposure_family",
      "exposure"
    )
  ) %>%

  dplyr::mutate(
    is_reference =
      !is.na(ReferenceCategory) &
      Category == ReferenceCategory,

    # Put the reference category first
    reference_order = dplyr::if_else(
      is_reference,
      0L,
      1L
    ),

    # Add the reference OR manually
    `Crude OR (95% CI)` = dplyr::case_when(
      is_reference ~ "1.000 (reference)",
      TRUE ~ dplyr::coalesce(
        `Crude OR (95% CI)`,
        ""
      )
    ),

    # Reference categories have no coefficient-level Wald p-value
    `Wald p-value` = dplyr::case_when(
      is_reference ~ "",
      TRUE ~ dplyr::coalesce(
        `Wald p-value`,
        ""
      )
    ),

    # Put unrecognised families/exposures at the end
    family_order = dplyr::coalesce(
      family_order,
      length(exposure_families) + 1L
    ),

    exposure_order = dplyr::coalesce(
      exposure_order,
      999999L
    )
  ) %>%

  dplyr::arrange(
    family_order,
    exposure_order,
    reference_order,
    category_original_order
  )



crude_results_table <- crude_results_table %>%
  dplyr::group_by(
    exposure_family,
    exposure
  ) %>%
  dplyr::mutate(
    `LRT p-value` = dplyr::if_else(
      dplyr::row_number() == 1L,
      `LRT p-value`,
      ""
    ),

    `Family FDR p-value` = dplyr::if_else(
      dplyr::row_number() == 1L,
      `Family FDR p-value`,
      ""
    ),

    `Global FDR p-value` = dplyr::if_else(
      dplyr::row_number() == 1L,
      `Global FDR p-value`,
      ""
    )
  ) %>%
  dplyr::ungroup()




#tidier ready for manuscript table
crude_results_table2 <- crude_results_table %>%
  dplyr::transmute(
    Group = exposure_family,
    Variable = exposure,
    Category,
    `Cases (%)`,
    `Non-cases (%)`,
    `Crude OR (95% CI)`,
    `Wald p-value`,
    `LRT p-value`,
    `Family FDR p-value`,
    `Global FDR p-value`
  )



crude_results_table3 <- crude_results_table2 %>%
  dplyr::mutate(
    Group_original = Group,
    Variable_original = Variable
  ) %>%

  # Display each variable name only once
  dplyr::group_by(
    Group_original,
    Variable_original
  ) %>%
  dplyr::mutate(
    Variable = dplyr::if_else(
      dplyr::row_number() == 1L,
      Variable,
      ""
    )
  ) %>%
  dplyr::ungroup() %>%

  # Display each exposure family only once
  dplyr::group_by(Group_original) %>%
  dplyr::mutate(
    Group = dplyr::if_else(
      dplyr::row_number() == 1L,
      Group,
      ""
    )
  ) %>%
  dplyr::ungroup() %>%

  dplyr::select(
    -Group_original,
    -Variable_original
  )
```

Need to ensure correct reference level (need to re

```{r}
install.packages(c("writexl", "readr", "officer", "flextable"))
```

factorise at start before this analysis
