---
title: "GRLS log reg HSA revised analysis approach"
format: html
editor: visual
---

# Variables associated with HSA in GRLS cohort

Load in dataset and packages

```{r}
library(tidyverse)
library(tidyverse)
library(dplyr)
library(survival)
library(ggplot2)
library(tibble)
library(kableExtra)
library(epiDisplay)
library(lme4)
#devtools::install_github("zabore/ezfun")
ezfun::set_ccf_palette("contrast")
library(lubridate)
library(ggsurvfit)
library(gtsummary)
library(tidycmprsk)
library(gtable)
#devtools::install_github("zabore/condsurv")
library(condsurv)
library(survminer)
library(cardx)
library(finalfit)
library(crosstable)
library(kableExtra)
library(officer)
library(flextable)
# data was preliminarily tidied in GRLS cox HSA time to diagnosis script before changing to log reg analyses
cohort_abr_updated_weight2 <-read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_cohort_all_RFs_data_tidied_from_cox.csv")
```

## Updated weight data

Incorporate in updated weight from Zoe's work (she has modified the erroneus lb kg issues also now all only adult weight so no \<18m weights in here)

```{r}
updated_weight <- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Data/weight_cleaned_grls.csv")

#existing weight columns are average and median 5y prior to endpoint and either split into quartiles or biological

#use the updated_weight df to create these same variables again + whole life weight

weight_df <- updated_weight %>%
  mutate(
    weight_kg = case_when(
      str_to_lower(weight_units) %in% c("lb", "lbs") ~ weight * 0.45359237,
      str_to_lower(weight_units) == "kg" ~ weight,
      TRUE ~ NA_real_
    )
  )

median_weight <- weight_df %>%
  filter(!is.na(weight_kg)) %>%
  group_by(subject_id) %>%
  arrange(year_in_study, .by_group = TRUE) %>%
  summarise(
    median_weight_lifetime_updated = median(weight_kg, na.rm = TRUE),

    median_weight_5y_prior_endpoint_updated = median(
      weight_kg[year_in_study %in% tail(sort(unique(year_in_study)), 5)],
      na.rm = TRUE
    ),

    .groups = "drop"
  )

#add into main df and see how different values are
cohort_abr_updated_weight2_updated_weight <- cohort_abr_updated_weight2 %>%
  left_join(median_weight, by="subject_id")


#check how diff
  
  
summarise(
  cohort_abr_updated_weight2_updated_weight,
  old_mean = mean(as.numeric(na_if(median_weight_5y_prior_endpoint_kg, "not_recorded")), na.rm = TRUE),
  old_median = median(as.numeric(na_if(median_weight_5y_prior_endpoint_kg, "not_recorded")), na.rm = TRUE),
  old_q1 = quantile(as.numeric(na_if(median_weight_5y_prior_endpoint_kg, "not_recorded")), 0.25, na.rm = TRUE),
  old_q3 = quantile(as.numeric(na_if(median_weight_5y_prior_endpoint_kg, "not_recorded")), 0.75, na.rm = TRUE),

  new_mean = mean(median_weight_5y_prior_endpoint_updated, na.rm = TRUE),
  new_median = median(median_weight_5y_prior_endpoint_updated, na.rm = TRUE),
  new_q1 = quantile(median_weight_5y_prior_endpoint_updated, 0.25, na.rm = TRUE),
  new_q3 = quantile(median_weight_5y_prior_endpoint_updated, 0.75, na.rm = TRUE)
)  


#remove old weight rows from df
cohort_abr_updated_weight2_updated_weight2 <- cohort_abr_updated_weight2_updated_weight %>%
  dplyr::select(-c("median_weight_5y_prior_endpoint_kg","avg_weight_5y_prior_endpoint_kg","avg_weight_5y_prior_endpoint_kg_outliers_removed","median_weight_5y_prior_endpoint_any_records_kg","median_weight_5y_prior_endpoint_kg_outliers_removed","median_weight_5y_prior_endpoint_any_records_kg","median_weight_5y_prior_endpoint_any_records_kg_outliers_removed",
                  "avg_weight_5y_prior_endpoint_quartiles",
                  "median_weight_5y_prior_endpoint_kg_outliers_removed_quartiles","avg_weight_5y_prior_endpoint_kg_biological","median_weight_5y_prior_endpoint_any_records_kg_outliers_removed_biological",
                  "median_weight_5y_prior_endpoint_kg_outliers_removed_biological"
                  ))

#now recode the new weight data into quartiles
cuts <- quantile(
  cohort_abr_updated_weight2_updated_weight2$median_weight_5y_prior_endpoint_updated,
  probs = c(0, 0.25, 0.5, 0.75, 1),
  na.rm = TRUE
)

labels <- c(
  paste0(round(cuts[1], 1), "–", round(cuts[2], 1), " kg"),
  paste0(round(cuts[2], 1), "–", round(cuts[3], 1), " kg"),
  paste0(round(cuts[3], 1), "–", round(cuts[4], 1), " kg"),
  paste0(round(cuts[4], 1), "–", round(cuts[5], 1), " kg")
)

cohort_abr_updated_weight2_updated_weight2 <- cohort_abr_updated_weight2_updated_weight2 %>%
  mutate(
    weight_quartile = case_when(
      is.na(median_weight_5y_prior_endpoint_updated) ~ "not_recorded",
      TRUE ~ as.character(cut(
        median_weight_5y_prior_endpoint_updated,
        breaks = cuts,
        include.lowest = TRUE,
        labels = labels
      ))
    ),
    weight_quartile = factor(
      weight_quartile,
      levels = c(labels, "not_recorded")
    )
  )


#also creating a 'biological' weight variable
cohort_abr_updated_weight2_updated_weight2 <- cohort_abr_updated_weight2_updated_weight2 %>%
  mutate(
    weight_biological = case_when(
      is.na(median_weight_5y_prior_endpoint_updated) ~ "not_recorded",
      median_weight_5y_prior_endpoint_updated <= 25 ~ "≤25 kg",
      median_weight_5y_prior_endpoint_updated <= 35 ~ ">25–≤35 kg",
      median_weight_5y_prior_endpoint_updated > 35 ~ ">35 kg"
    ),
    weight_biological = factor(
      weight_biological,
      levels = c(
        "≤25 kg",
        ">25–≤35 kg",
        ">35 kg",
        "not_recorded"
      )
    )
  )
```

Converting to factors for analysis:

```{r}

#replace all missing with NA (so "" and NA)
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(across(everything(), as.character)) %>%  # Ensure all columns are character
  mutate(across(everything(), ~na_if(.x, "n/a"))) %>%
  mutate(across(everything(), ~na_if(.x, "NA"))) %>%
  mutate(across(everything(), ~na_if(.x, "nr"))) %>%
  mutate(across(everything(), ~na_if(.x, "not recorded"))) %>%
  mutate(across(everything(), ~na_if(.x, ""))) %>%
  mutate(across(everything(), ~replace_na(.x, "not_recorded")))
cohort_abr_updated_weight2$survival_time_months <- as.numeric(cohort_abr_updated_weight2$survival_time_months)


cohort_abr_updated_weight2$status_factor <- as.factor(as.character(cohort_abr_updated_weight2$had_HSA))
cohort_abr_updated_weight2$status <-as.numeric(cohort_abr_updated_weight2$had_HSA) 
cohort_abr_updated_weight2$survival_time_months <- as.numeric(cohort_abr_updated_weight2$survival_time_months)


#change spayed while in heat to account for males and entire females
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(spayed_while_in_heat = case_when(
    sex== "M" ~ "Male",
    sex_status == "Intact Female" ~ "Intact Female",
    spayed_while_in_heat == "0" ~ "0",
    spayed_while_in_heat == "1" ~"1"
  ))
#same as above account for unnetered and males 
cohort_abr_updated_weight2$heats_before_neutering <- as.numeric(cohort_abr_updated_weight2$heats_before_neutering)
heats <- cohort_abr_updated_weight2 %>%
# Convert to numeric
  filter(heats_before_neutering != 0) %>% 
# Remove rows where the column is 0
  filter(heats_before_neutering != "not_recorded")%>%
  dplyr::select(heats_before_neutering)# Keep only this column
heats <- as.numeric(heats$heats_before_neutering)

# Function to calculate quartile or tercile dynamically
get_breaks <- function(x, probs) {
  breaks <- quantile(x, probs = probs, na.rm = TRUE)
  
  # If breaks are not unique, switch to tercile
  if (length(unique(breaks)) < length(probs)) {
    message("quartile not unique, switching to tercile.")
    breaks <- quantile(x, probs = c(0, 1/3, 2/3, 1), na.rm = TRUE)
  }
  
  return(breaks)
}

breaks <- get_breaks(heats, c(0, 0.25, 0.5, 0.75, 1))


# Assign labels dynamically
labels <- paste0("≥", round(head(breaks, -1), 2), " - <", round(tail(breaks, -1), 2))


cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(heats_before_neutering_tercile = case_when(
    sex== "M" ~ "Male",
    sex_status == "Intact Female" ~ "Intact Female",
    heats_before_neutering == "0" ~ "0",
    heats_before_neutering == "1" ~"1-<=2",
    heats_before_neutering == "2" ~"1-<=2",
    heats_before_neutering == "3" ~">2-<=4",
    heats_before_neutering == "4" ~">2-<=4",
    heats_before_neutering == "5" ~">4-12",
    heats_before_neutering == "6" ~">4-12",
    heats_before_neutering == "7" ~">4-12",
    heats_before_neutering == "8" ~">4-12",
    heats_before_neutering == "9" ~">4-12",
    heats_before_neutering == "10" ~">4-12",
    heats_before_neutering == "11" ~">4-12",
    heats_before_neutering == "12" ~">4-12"
  ))


#create combined sex neuter variable
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(sex_neuter = paste(sex, neuter, sep = " "))


#ensure not_recorded is default not baseline for regression analyses as isnt a good comparison option
convert_and_relevel <- function(df) {
  df <- df %>%
    mutate(across(where(is.character), ~ factor(.))) %>%  # Convert all character columns to factors
    mutate(across(where(is.factor), ~ fct_relevel(., "not_recorded", after = Inf)))  # Move "not_recorded" to last place
  return(df)
}

# Apply to your dataset
cohort_abr_updated_weight2 <- convert_and_relevel(cohort_abr_updated_weight2)

#set the levels i would like for comparison

cohort_abr_updated_weight2 <- cohort_abr_updated_weight2%>%
  mutate(heats_before_neutering_tercile = factor(heats_before_neutering_tercile, levels = c("0", "1-<=2", ">2-<=4", ">4-12","Intact Female","Male","not_recorded")))


#neuter reason levels
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2%>%
  mutate(neutered_reason = factor(neutered_reason, levels = c("Elective",  "Medical Reason", "Entire","Unknown")))


cohort_abr_updated_weight2 <- cohort_abr_updated_weight2%>%
  mutate(mode_purina_BCS_5y_prior_endpoint = factor(mode_purina_BCS_5y_prior_endpoint, levels = c("5", "3", "4", "6","7","8","9","not_recorded")))

#change BCS to be reclassified as over underweight or ideal
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(mode_BCS_classify = case_when(
    mode_purina_BCS_5y_prior_endpoint == "1" ~"underweight",
    mode_purina_BCS_5y_prior_endpoint == "2" ~"underweight",
    mode_purina_BCS_5y_prior_endpoint == "3" ~"underweight",
    mode_purina_BCS_5y_prior_endpoint == "4" ~"ideal",
    mode_purina_BCS_5y_prior_endpoint == "5" ~"ideal",
    mode_purina_BCS_5y_prior_endpoint == "6" ~"overweight",
    mode_purina_BCS_5y_prior_endpoint == "7" ~"overweight",
    mode_purina_BCS_5y_prior_endpoint == "8" ~"overweight",
    mode_purina_BCS_5y_prior_endpoint == "9" ~"overweight",
  ))

#set age levels
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2%>%
  mutate(age_at_final_date_quartiles = factor(age_at_final_date_quartiles, levels = c("0.74-<8.47 years", ">=8.47-<10.68 years", ">=10.68-<11.64 years", ">=11.64-14.51 years")))

cohort_abr_updated_weight2 <- cohort_abr_updated_weight2%>%
  mutate(age_at_final_date_bio = factor(age_at_final_date_bio, levels = c(">5-<8y","<5y",  ">8-<11y", ">11-<13y",">13y")))

cohort_abr_updated_weight2 <- cohort_abr_updated_weight2%>%
  mutate(age_at_final_date_GB = factor(age_at_final_date_GB, levels = c(">7-<9y","<5y", ">5-<7y",  ">9-<11y",">11-<13y",">13y")))


#manually assign height, weight, BCS factor levles and age at neuter and age at diagnosis 
#height
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2%>%
  mutate(avg_height_5y_prior_endpoint_quartiles = factor(avg_height_5y_prior_endpoint_quartiles, levels = c("49.80-<56.30cm", ">=56.30-<58.29cm", ">=58.29-<60.32cm", ">=60.32cm-67.73cm","not_recorded")))

cohort_abr_updated_weight2 <- cohort_abr_updated_weight2%>%
  mutate(avg_purina_BCS_5y_prior_endpoint_quartiles = factor(avg_purina_BCS_5y_prior_endpoint_quartiles, levels = c("3.3-<5", ">=5-<5.28", ">=5.28-<5.86", ">=5.86 - 9","not_recorded")))
#neut age

cohort_abr_updated_weight2 <- cohort_abr_updated_weight2%>%
  mutate(age_at_neuter_years_quartiles = factor(age_at_neuter_years_quartiles, levels = c("0.15-<0.52y", ">=0.52-<0.83y", ">=0.83-2.44y", ">=2.44-9.59y","Entire","not_recorded")))
#diagnosis age
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2%>%
  mutate(age_at_final_date_quartiles = factor(age_at_final_date_quartiles, levels = c("0.74-<8.47 years", ">=8.47-<10.68 years", ">=10.68-<11.64 years", ">=11.64-14.51 years","not_recorded")))

#breed tercile levels
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2%>%
  mutate(num_breeding_before_neutering_terciles = factor(num_breeding_before_neutering_terciles, levels = c(">0-<3",">=3-<9",">=9-142","not_bred","not_recorded")))


#recode NSAID quartiles

cohort_abr_updated_weight2 <- cohort_abr_updated_weight2%>%
  mutate(NSAID_5y_quantiles = factor(NSAID_5y_quantiles, levels = c("no_NSAID_admin", "Q1(1)", "Q2(1-2)", "Q3(2-3)","Q4(3-11)","no_med_use_recorded")))

cohort_abr_updated_weight2 <- cohort_abr_updated_weight2%>%
  mutate(endo_5y_quantiles = factor(endo_5y_quantiles, levels = c("no_endo_admin", "Q1(1-2)", "Q2(2-3)", "Q3(3-4)","Q4(4-17)","no_med_use_recorded")))

cohort_abr_updated_weight2 <- cohort_abr_updated_weight2%>%
  mutate(ecto_5y_quantiles = factor(ecto_5y_quantiles, levels = c("no_ecto_admin", "Q1(1)", "Q2(1-2)", "Q3(2-4)","Q4(4-17)","no_med_use_recorded")))
###setting "no_medication_records_for_this_time_periodistered" as bottom factor level for drug variables as think it actually means drug use not recorded so doesnt make a good baseline
# Vector of the column names to modify

cols_to_modify <- 182:249 

for (i in cols_to_modify) {
  current_levels <- levels(cohort_abr_updated_weight2[[i]])
  if ("no_medication_records_for_this_time_period" %in% current_levels) {
    new_levels <- c(setdiff(current_levels, "no_medication_records_for_this_time_period"), "no_medication_records_for_this_time_period")
    cohort_abr_updated_weight2[[i]] <- factor(cohort_abr_updated_weight2[[i]], levels = new_levels)
  }
}
```

Activity dataset edits:

```{r}
#SY1 and 2
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(avg_activity_duration_sy1_2 = as.numeric(as.character(avg_activity_duration_sy1_2))) %>%
  mutate(avg_activity_duration_sy1_2_splits = case_when(
    avg_activity_duration_sy1_2 <= 20 ~ "less_than_30min",
    avg_activity_duration_sy1_2 >20 & avg_activity_duration_sy1_2 <= 40 ~ "between_30_60_min",
    avg_activity_duration_sy1_2 >40 & avg_activity_duration_sy1_2<= 60 ~ "between_60_90_min",
    avg_activity_duration_sy1_2 >60 & avg_activity_duration_sy1_2 <= 80 ~ "between_90_120_min",
    avg_activity_duration_sy1_2 >80 & avg_activity_duration_sy1_2 <= 100 ~ "more_than_120_min"
  ))

cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(avg_activity_freq_sy1_2 = as.numeric(as.character(avg_activity_freq_sy1_2))) %>%
  mutate(avg_activity_freq_sy1_2_splits = case_when(
    avg_activity_freq_sy1_2 <= 20 ~ "never",
    avg_activity_freq_sy1_2 >20 & avg_activity_freq_sy1_2<= 40 ~ "less_than_weekly",
    avg_activity_freq_sy1_2 >40 & avg_activity_freq_sy1_2<= 60 ~ "min_weekly",
    avg_activity_freq_sy1_2 >60 & avg_activity_freq_sy1_2 <= 80 ~ "daily",
    avg_activity_freq_sy1_2 >80 & avg_activity_freq_sy1_2<= 100 ~ "multi_daily"
  ))


cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(avg_activity_intensity_sy1_2 = as.numeric(as.character(avg_activity_intensity_sy1_2))) %>%
  mutate(avg_activity_intensity_sy1_2_splits = case_when(
    avg_activity_intensity_sy1_2 <= 33.3 ~ "low",
    avg_activity_intensity_sy1_2 >33.3 & avg_activity_intensity_sy1_2<= 66.7 ~ "moderate",
    avg_activity_intensity_sy1_2 >66.7 & avg_activity_intensity_sy1_2<= 100 ~ "high",

  ))

#SY3 + intensity and freq


cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(avg_activity_freq_sy3 = as.numeric(as.character(avg_activity_freq_sy3))) %>%
  mutate(avg_activity_freq_sy3_splits = case_when(
    avg_activity_freq_sy3 <= 20 ~ "rarely",
    avg_activity_freq_sy3 >20 & avg_activity_freq_sy3<= 40 ~ "monthly",
    avg_activity_freq_sy3 >40 & avg_activity_freq_sy3<= 60 ~ "weekly",
    avg_activity_freq_sy3 >60 & avg_activity_freq_sy3 <= 80 ~ "daily",
    avg_activity_freq_sy3 >80 & avg_activity_freq_sy3<= 100 ~ "multi_daily"
  ))


cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(avg_activity_intensity_sy3 = as.numeric(as.character(avg_activity_intensity_sy3))) %>%
  mutate(avg_activity_intensity_sy3_splits = case_when(
    avg_activity_intensity_sy3 <= 20 ~ "slow_walks",
    avg_activity_intensity_sy3 >20 & avg_activity_intensity_sy3<= 40 ~ "average_walks",
    avg_activity_intensity_sy3 >40 & avg_activity_intensity_sy3<= 60 ~ "brisk_walks",
    avg_activity_intensity_sy3 >60 & avg_activity_intensity_sy3 <= 80 ~ "jogs",
    avg_activity_intensity_sy3 >80 & avg_activity_intensity_sy3<= 100 ~ "runs"
  ))



#5y prior endpoint  freq


cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(avg_frequency_5y_prior_endpoint = as.numeric(as.character(avg_activity_freq_sy3))) %>%
  mutate(avg_frequency_5y_prior_endpoint_splits = case_when(
    avg_frequency_5y_prior_endpoint <= 20 ~ "rarely",
    avg_frequency_5y_prior_endpoint >20 & avg_frequency_5y_prior_endpoint<= 40 ~ "monthly",
    avg_frequency_5y_prior_endpoint >40 & avg_frequency_5y_prior_endpoint<= 60 ~ "weekly",
    avg_frequency_5y_prior_endpoint >60 & avg_frequency_5y_prior_endpoint <= 80 ~ "daily",
    avg_frequency_5y_prior_endpoint >80 &avg_frequency_5y_prior_endpoint<= 100 ~ "multi_daily"
  ))
#5y prior endpoint intensity
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(avg_intensity_5y_prior_endpoint = as.numeric(as.character(avg_intensity_5y_prior_endpoint))) %>%
  mutate(avg_intensity_5y_prior_endpoint_splits = case_when(
    avg_intensity_5y_prior_endpoint <= 20 ~ "slow_walks",
    avg_intensity_5y_prior_endpoint >20 & avg_intensity_5y_prior_endpoint<= 40 ~ "average_walks",
    avg_intensity_5y_prior_endpoint >40 & avg_intensity_5y_prior_endpoint<= 60 ~ "brisk_walks",
    avg_intensity_5y_prior_endpoint >60 & avg_intensity_5y_prior_endpoint <= 80 ~ "jogs",
    avg_intensity_5y_prior_endpoint >80 & avg_intensity_5y_prior_endpoint<= 100 ~ "runs"
  ))
#recode into larger groupings e.g grouping together rarely + monthly for frequency
#freq
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(avg_activity_freq_sy3_splits_grouped = case_when(
    avg_activity_freq_sy3_splits == "monthly" ~ "rarely",
    avg_activity_freq_sy3_splits == "multi_daily"~"daily_or_more" ,
    avg_activity_freq_sy3_splits == "daily" ~"daily_or_more" ,
    avg_activity_freq_sy3_splits == "weekly" ~"weekly" ,
    avg_activity_freq_sy3_splits == "rarely" ~ "rarely",
  ))

#intensity
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(avg_activity_intensity_sy3_splits_grouped = case_when(
    avg_activity_intensity_sy3_splits == "jogs" ~ "high_intensity",
    avg_activity_intensity_sy3_splits == "runs" ~"high_intensity" ,
    avg_activity_intensity_sy3_splits == "brisk_walks" ~"moderate_intensity",
    avg_activity_intensity_sy3_splits =="average_walks" ~ "low_intensity",
     avg_activity_intensity_sy3_splits =="slow_walks" ~ "low_intensity"
  ))
#5y prior freq
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(avg_frequency_5y_prior_endpoint_splits_grouped = case_when(
    avg_frequency_5y_prior_endpoint_splits == "monthly" ~ "rarely",
    avg_frequency_5y_prior_endpoint_splits == "multi_daily"~"daily_or_more" ,
    avg_frequency_5y_prior_endpoint_splits == "daily" ~"daily_or_more" ,
    avg_frequency_5y_prior_endpoint_splits == "weekly" ~"weekly" ,
    avg_frequency_5y_prior_endpoint_splits == "rarely" ~"rarely" ,
  ))

#5y prior intensity
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(avg_intensity_5y_prior_endpoint_splits_grouped = case_when(
    avg_intensity_5y_prior_endpoint_splits == "jogs" ~ "high_intensity",
    avg_intensity_5y_prior_endpoint_splits == "runs" ~"high_intensity" ,
    avg_intensity_5y_prior_endpoint_splits == "brisk_walks" ~"moderate_intensity",
    avg_intensity_5y_prior_endpoint_splits =="average_walks" ~ "low_intensity",
    avg_intensity_5y_prior_endpoint_splits =="slow_walks" ~ "low_intensity"
  ))

cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(across(c(early_activity_mean, rest_activity_mean, whole_activity_mean,prior5y_activity_mean), as.numeric))

#swap not_recorded for NA and exclude from univar this way
cohort_no_not_rec <- cohort_abr_updated_weight2 %>%
  mutate(across(
    where(~ is.character(.) | is.factor(.)),
    ~ factor(na_if(as.character(.), "not_recorded"))
  ))

#change BCS to be over underweight
cohort_no_not_rec <- cohort_no_not_rec %>%
  mutate(mode_BCS_classify = case_when(
    mode_purina_BCS_5y_prior_endpoint == "1" ~"underweight",
    mode_purina_BCS_5y_prior_endpoint == "2" ~"underweight",
    mode_purina_BCS_5y_prior_endpoint == "3" ~"underweight",
    mode_purina_BCS_5y_prior_endpoint == "4" ~"ideal",
    mode_purina_BCS_5y_prior_endpoint == "5" ~"ideal",
    mode_purina_BCS_5y_prior_endpoint == "6" ~"overweight",
    mode_purina_BCS_5y_prior_endpoint == "7" ~"overweight",
    mode_purina_BCS_5y_prior_endpoint == "8" ~"overweight",
    mode_purina_BCS_5y_prior_endpoint == "9" ~"overweight",
    mode_purina_BCS_5y_prior_endpoint == "not_recorded" ~"not_recorded",
  ))

colnames_df <- colnames(cohort_no_not_rec)
write.csv(colnames_df,"C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_col_names.csv")
```

## Regression analyses

**Fixed/forced covariates ie. in ALL models**

```{r}
exposome_data <- cohort_abr_updated_weight2 %>%
  mutate(
    had_HSA = as.integer(as.character(had_HSA))
  )

#forced covariates ie. the model will keep these in always and the exposures will be added in and out of different models
forced_covariates <- c(
  "age_at_final_date_bio",
  "sex",
  "neuter"
  #or possibly sex_status which is sex and neuter status combined
)
```

### **Exposome groups**

#### Built environment and socioeconomic

#### Household chemical exposure

#### Lifestyle/sleep location

#### Dog activity

#### Body condition/weight

#### Reproductive exposures

#### Medication types

##### NSAIDs(anti-inflams)

##### Antimicrobials

##### Dietary supplements

##### Enteric medications

##### Hormones

##### Immunologicals

##### Neurologicals

##### Anti-parasites

##### Anti-septics

##### Cardiorespiratory 

##### Diuretics

##### Antihistamines

##### Antivirals

#### Comorbidities

##### Cardiovascular

##### Orthopaedic

##### Immune mediated

##### Chronic inflammatory

##### Parasitism

```{r}
exposure_families <- list(

  #------------------------------------------------------------
  # BUILT ENVIRONMENT AND SOCIOECONOMIC
  #------------------------------------------------------------

  built_environment_socioeconomic = c(
    "MDI_quintile",
    "owner_MDI_quintile",
    "area_type_mode_5yrs_prior",
    "house_type_mode_5yrs_prior",
    "water_source_mode_5yrs_prior",
    "pipes_metal_any_mode_5yrs_prior",
    "pipes_plastic_any_mode_5yrs_prior",
    "heating_fuel_primary_mode_5yrs_prior",
    "cooking_fuel_primary_mode_5yrs_prior",
    "heating_fuel_secondary_mode_5yrs_prior",
    "cooking_fuel_secondary_mode_5yrs_prior",
    "region_mode_5y_prior"
  ),

  #------------------------------------------------------------
  # HOUSEHOLD CHEMICAL EXPOSURE
  #------------------------------------------------------------

  household_chemical_exposure = c(
    "use_aerosol_study_years_early_life",
    "use_aerosol_study_years_rest_of_life",
    "use_aerosol_study_years_whole_life",
    "use_aerosol_exposed_5yrs_prior",

    "use_air_cleaner_study_years_early_life",
    "use_air_cleaner_study_years_rest_of_life",
    "use_air_cleaner_study_years_whole_life",
    "use_air_cleaner_exposed_5yrs_prior",

    "use_hepa_filter_study_years_early_life",
    "use_hepa_filter_study_years_rest_of_life",
    "use_hepa_filter_study_years_whole_life",
    "use_hepa_filter_exposed_5yrs_prior",

    "use_moth_balls_study_years_early_life",
    "use_moth_balls_study_years_rest_of_life",
    "use_moth_balls_study_years_whole_life",
    "use_moth_balls_exposed_5yrs_prior",

    "use_incense_or_candles_study_years_early_life",
    "use_incense_or_candles_study_years_rest_of_life",
    "use_incense_or_candles_study_years_whole_life",
    "use_incense_or_candles_exposed_5yrs_prior",

    "smoke_exposure_study_years_early_life",
    "smoke_exposure_study_years_rest_of_life",
    "smoke_exposure_study_years_whole_life",
    "smoke_exposure_exposed_5yrs_prior",

    "hours_of_smoke_early_life_total_dosage_terciles",
    "hours_of_smoke_rest_of_life_total_dosage_terciles",
    "hours_of_smoke_whole_life_total_dosage_terciles",
    "hours_of_smoke_exposure_hours_5yrs_prior_terciles",

    "any_treated_weeds_study_years_early_life",
    "any_treated_weeds_study_years_rest_of_life",
    "any_treated_weeds_study_years_whole_life",
    "any_treated_weeds_exposed_5yrs_prior",

    "any_treated_insects_study_years_early_life",
    "any_treated_insects_study_years_rest_of_life",
    "any_treated_insects_study_years_whole_life",
    "any_treated_insects_exposed_5yrs_prior",

    "any_treated_fertilizer_study_years_early_life",
    "any_treated_fertilizer_study_years_rest_of_life",
    "any_treated_fertilizer_study_years_whole_life",
    "any_treated_fertilizer_exposed_5yrs_prior"
  ),

  #------------------------------------------------------------
  # LIFESTYLE / SLEEP LOCATION
  #------------------------------------------------------------

  lifestyle_sleep_location = c(
    "majority_sleep_location_lifetime",
    "main_lifestyle_category",
    "main_lifestyle_category_mode_5yrs_prior",
    "sleep_location_mode_5yrs_prior",

    "in.the.garage_YN_study_years_early_life",
    "in.the.garage_YN_study_years_rest_of_life",
    "in.the.garage_YN_study_years_whole_life",

    "in.the.house_YN_study_years_early_life",
    "in.the.house_YN_study_years_rest_of_life",
    "in.the.house_YN_study_years_whole_life",

    "outside_YN_study_years_early_life",
    "outside_YN_study_years_rest_of_life",
    "outside_YN_study_years_whole_life",

    "in.the.house_YN_sum_early_life_quartile",
    "in.the.house_YN_sum_rest_of_life_quartile",
    "in.the.house_YN_sum_whole_life_quartile"
  ),

  #------------------------------------------------------------
  # DOG ACTIVITY
  #------------------------------------------------------------

  dog_activity = c(
    "mean_activity_level_overview",

    "avg_activity_freq_sy1_2_quartile",
    "avg_activity_duration_sy1_2_quartile",
    "avg_activity_intensity_sy3_quartile",
    "avg_activity_freq_sy3_quartile",

    "avg_frequency_5y_prior_endpoint_quartile",
    "avg_intensity_5y_prior_endpoint_quartile",

    "early_activity_mode",
    "early_activity_mean",
    "early_activity_median",

    "rest_activity_mode",
    "rest_activity_mean",
    "rest_activity_median",

    "whole_activity_mode",
    "whole_activity_mean",
    "whole_activity_median",

    "prior5y_activity_mode",
    "prior5y_activity_mean",
    "prior5y_activity_median"
  ),

  #------------------------------------------------------------
  # BODY CONDITION / WEIGHT
  #------------------------------------------------------------

  body_condition_weight = c(
    "weight_biological",
    "weight_quartile",
    "avg_height_5y_prior_endpoint_quartiles",
    "avg_purina_BCS_5y_prior_endpoint_quartiles",
    "mode_BCS_classify"
  ),

  #------------------------------------------------------------
  # REPRODUCTIVE EXPOSURES
  #------------------------------------------------------------

  reproductive_exposures = c(
    "age_at_neuter_years_quartiles",
    "mismating_ever",
    "neutered_reason",
    "neutered_ever_bred",
    "spayed_while_in_heat",
    "spayed_while_pregnant",
    "prevent_heat_ever",
    "no_pregnancy_ever",
    "heats_before_neutering_terciles",
    "preg_pre_neuter",
    "mated_pre_neuter",
    "num_breeding_before_neutering_terciles",
    "ever_bred_from"
  ),

  #------------------------------------------------------------
  # MEDICATION TYPES: BROAD OVERVIEW
  #------------------------------------------------------------

  medication_types = c(
    "half_anti_inflam_lifetime",
    "half_anti_microbials_lifetime",
    "half_dietary_supplements_lifetime",
    "half_enteric_lifetime",
    "half_hormones_and_related_lifetime",
    "half_immunologicals_lifetime",
    "half_neurological_lifetime",
    "half_anti_parasites_lifetime",
    "half_anti_septics_lifetime",
    "half_cardio_respiratory_lifetime",
    "half_diuretics_lifetime",
    "half_anti_hist_lifetime",
    "half_anti_viral_lifetime"
  ),

  #------------------------------------------------------------
  # INDIVIDUAL MEDICATION FAMILIES
  #------------------------------------------------------------

  nsaids_anti_inflammatories = c(
    "NSAID_ever",
    "NSAID_5y",
    "NSAID_5y_quantiles",
    "half_anti_inflam_lifetime",
    "quartile_anti_inflam_lifetime",
    "half_anti_inflam_5y_prior_endpoint",
    "quartile_anti_inflam_5y_prior_endpoint"
  ),

  antimicrobials = c(
    "half_anti_microbials_lifetime",
    "quartile_anti_microbials_lifetime",
    "half_anti_microbials_5y_prior_endpoint",
    "quartile_anti_microbials_5y_prior_endpoint"
  ),

  dietary_supplements = c(
    "half_dietary_supplements_lifetime",
    "quartile_dietary_supplements_lifetime",
    "half_dietary_supplements_5y_prior_endpoint",
    "quartile_dietary_supplements_5y_prior_endpoint"
  ),

  enteric_medications = c(
    "half_enteric_lifetime",
    "quartile_enteric_lifetime",
    "half_enteric_5y_prior_endpoint",
    "quartile_enteric_5y_prior_endpoint"
  ),

  hormones = c(
    "half_hormones_and_related_lifetime",
    "quartile_hormones_and_related_lifetime",
    "half_hormones_and_related_5y_prior_endpoint",
    "quartile_hormones_and_related_5y_prior_endpoint",
    "steroid_ever",
    "steroid_5y"
  ),

  immunologicals = c(
    "half_immunologicals_lifetime",
    "quartile_immunologicals_lifetime",
    "half_immunologicals_5y_prior_endpoint",
    "quartile_immunologicals_5y_prior_endpoint"
  ),

  neurologicals = c(
    "half_neurological_lifetime",
    "quartile_neurological_lifetime",
    "half_neurological_5y_prior_endpoint",
    "quartile_neurological_5y_prior_endpoint"
  ),

  anti_parasites = c(
    "half_anti_parasites_lifetime",
    "quartile_anti_parasites_lifetime",
    "half_anti_parasites_5y_prior_endpoint",
    "quartile_anti_parasites_5y_prior_endpoint",
    "ecto_5y_quantiles",
    "endo_5y_quantiles",
    "fenbendazole_ever",
    "fenbendazole_5y"
  ),

  anti_septics = c(
    "half_anti_septics_lifetime",
    "quartile_anti_septics_lifetime",
    "half_anti_septics_5y_prior_endpoint",
    "quartile_anti_septics_5y_prior_endpoint"
  ),

  cardiorespiratory_medications = c(
    "half_cardio_respiratory_lifetime",
    "quartile_cardio_respiratory_lifetime",
    "half_cardio_respiratory_5y_prior_endpoint",
    "quartile_cardio_respiratory_5y_prior_endpoint"
  ),

  diuretics = c(
    "half_diuretics_lifetime",
    "quartile_diuretics_lifetime",
    "half_diuretics_5y_prior_endpoint",
    "quartile_diuretics_5y_prior_endpoint"
  ),

  antihistamines = c(
    "half_anti_hist_lifetime",
    "quartile_anti_hist_lifetime",
    "half_anti_hist_5y_prior_endpoint",
    "quartile_anti_hist_5y_prior_endpoint"
  ),

  antivirals = c(
    "half_anti_viral_lifetime",
    "quartile_anti_viral_lifetime",
    "half_anti_viral_5y_prior_endpoint",
    "quartile_anti_viral_5y_prior_endpoint"
  ),

  #------------------------------------------------------------
  # COMORBIDITIES: OVERALL
  #------------------------------------------------------------

  comorbidities = c(
    "chronic_inflam_present_lifetime",
    "number_comorbs_present_lifetime",
    "chronic_inflam_present_5y",
    "number_comorbs_present_5y",
    "chronic_inflam_present_no_min_year",
    "number_comorbs_present_no_min_year",
    "lifelong_comorb_present.x",
    "number_lifelong_comorbs_present"
  ),

  #------------------------------------------------------------
  # INDIVIDUAL COMORBIDITY FAMILIES
  #------------------------------------------------------------

  cardiovascular_comorbidity = c(
    "cardiovascular",
    "lifelong_cardiovascular",
    "comorb_cardiovascular_all"
  ),

  orthopaedic_comorbidity = c(
    "orthopaedic",
    "lifelong_orthopaedic",
    "comorb_orthopaedic_all"
  ),

  immune_mediated_comorbidity = c(
    "immune_mediated",
    "lifelong_immune_mediated",
    "comorb_immune_med_all"
  ),

  chronic_inflammatory_comorbidity = c(
    "chronic_inflammatory",
    "lifelong_chronic_inflammatory",
    "lifelong_inflammatory_other",
    "comorb_chronic_inflam_all"
  ),

  #------------------------------------------------------------
  # PARASITISM
  #------------------------------------------------------------

  parasitism = c(
    "intestinal_parasitism",
    "tick_borne_parasitism",
    "other_parasitism",
    "all_parasitism"
  )
)
```

```{r}
exposure_families_requested <- exposure_families

missing_family_variables <- imap_dfr(
  exposure_families_requested,
  function(variables, family) {

    missing <- setdiff(
      variables,
      names(exposome_data)
    )

    tibble(
      exposure_family = family,
      missing_variable = missing
    )
  }
)

missing_family_variables
```

```{r}
analysis_df <- cohort_abr_updated_weight2
outcome_var <- "had_HSA"
requested_exposure_families <- exposure_families
missing_exposure_variables <- imap_dfr(
  requested_exposure_families,
  function(variables, family) {

    missing_variables <- setdiff(
      variables,
      names(analysis_df)
    )

    tibble(
      exposure_family = family,
      variable = missing_variables
    )
  }
)
exposure_families <- map(
  requested_exposure_families,
  ~ intersect(.x, names(analysis_df))
)
exposure_families <- exposure_families[
  lengths(exposure_families) > 0
]
tibble(
  exposure_family = names(exposure_families),
  number_variables = lengths(exposure_families)
)
```

```{r}
screen_exposure <- function(
    data,
    exposure,
    outcome = "had_HSA"
) {

  x <- data[[exposure]]
  y <- data[[outcome]]

  complete <- !is.na(x) & !is.na(y)

  x_complete <- x[complete]
  y_complete <- y[complete]

  tibble(
    exposure = exposure,
    variable_class = paste(class(x), collapse = "/"),
    n_total = nrow(data),
    n_observed = sum(!is.na(x)),
    n_missing = sum(is.na(x)),
    percent_missing = round(
      100 * mean(is.na(x)),
      1
    ),
    n_complete_outcome_exposure = sum(complete),
    cases_complete = sum(y_complete == 1),
    controls_complete = sum(y_complete == 0),
    n_unique_observed = n_distinct(x_complete),
    suitable_for_model = (
      sum(complete) > 0 &&
      n_distinct(x_complete) >= 2 &&
      n_distinct(y_complete) == 2
    )
  )
}
```

```{r}
exposure_screening <- imap_dfr(
  exposure_families,
  function(variables, family) {

    map_dfr(
      variables,
      ~ screen_exposure(
        data = analysis_df,
        exposure = .x,
        outcome = outcome_var
      )
    ) %>%
      mutate(
        exposure_family = family,
        .before = 1
      )
  }
)

exposure_screening %>%
  filter(!suitable_for_model)
```

```{r}
usable_exposure_families <- split(
  exposure_screening$exposure[
    exposure_screening$suitable_for_model
  ],
  exposure_screening$exposure_family[
    exposure_screening$suitable_for_model
  ]
)
```

\

```{r}
make_exposure_counts <- function(
    data,
    exposure,
    outcome = "had_HSA"
) {

  data %>%
    transmute(
      exposure_level = as.character(.data[[exposure]]),
      outcome_value = .data[[outcome]]
    ) %>%
    mutate(
      exposure_level = replace_na(
        exposure_level,
        "missing"
      ),
      outcome_value = as.character(outcome_value)
    ) %>%
    count(
      exposure_level,
      outcome_value,
      name = "n",
      .drop = FALSE
    ) %>%
    pivot_wider(
      names_from = outcome_value,
      values_from = n,
      values_fill = 0,
      names_prefix = "outcome_"
    ) %>%
    mutate(
      exposure = exposure,
      .before = 1
    )
}
```

```{r}
exposure_level_counts <- imap_dfr(
  usable_exposure_families,
  function(variables, family) {

    map_dfr(
      variables,
      ~ make_exposure_counts(
        data = analysis_df,
        exposure = .x,
        outcome = outcome_var
      )
    ) %>%
      mutate(
        exposure_family = family,
        .before = 1
      )
  }
)
```

```{r}
exposure_level_counts
```

```{r}
write_csv(
  exposure_level_counts,
  "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/exposome_exposure_level_counts.csv"
)
```

### Univariable

```{r}
# ============================================================
# CRUDE LOGISTIC REGRESSION FUNCTION
# ============================================================

fit_crude_exposure_model <- function(
    data,
    exposure,
    outcome = "had_HSA"
) {

  required_variables <- c(
    outcome,
    exposure
  )

  missing_variables <- setdiff(
    required_variables,
    names(data)
  )

  if (length(missing_variables) > 0) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = NA_character_,
        n = NA_integer_,
        cases = NA_integer_,
        controls = NA_integer_,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = NA_real_,
        overall_LRT_df = NA_real_,
        overall_LRT_p = NA_real_,
        AIC = NA_real_,
        converged = FALSE,
        warning = paste(
          "Missing variable(s):",
          paste(
            missing_variables,
            collapse = ", "
          )
        )
      )
    )
  }

  model_data <- data %>%
    dplyr::select(
      dplyr::all_of(
        required_variables
      )
    ) %>%
    tidyr::drop_na()

  n_model <- nrow(model_data)

  n_cases <- sum(
    model_data[[outcome]] == 1,
    na.rm = TRUE
  )

  n_controls <- sum(
    model_data[[outcome]] == 0,
    na.rm = TRUE
  )

  outcome_variation <- dplyr::n_distinct(
    model_data[[outcome]]
  )

  exposure_variation <- dplyr::n_distinct(
    model_data[[exposure]]
  )

  if (
    n_model == 0 ||
      outcome_variation < 2 ||
      exposure_variation < 2
  ) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = NA_character_,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = NA_real_,
        overall_LRT_df = NA_real_,
        overall_LRT_p = NA_real_,
        AIC = NA_real_,
        converged = FALSE,
        warning = "Insufficient variation in outcome or exposure"
      )
    )
  }

  full_formula <- stats::reformulate(
    termlabels = exposure,
    response = outcome
  )

  null_formula <- stats::reformulate(
    termlabels = NULL,
    response = outcome
  )

  warning_messages <- character()

  full_model <- tryCatch(

    withCallingHandlers(

      stats::glm(
        formula = full_formula,
        data = model_data,
        family = stats::binomial()
      ),

      warning = function(w) {

        warning_messages <<- c(
          warning_messages,
          conditionMessage(w)
        )

        invokeRestart("muffleWarning")
      }
    ),

    error = function(e) {
      e
    }
  )

  if (inherits(full_model, "error")) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = NA_character_,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = NA_real_,
        overall_LRT_df = NA_real_,
        overall_LRT_p = NA_real_,
        AIC = NA_real_,
        converged = FALSE,
        warning = conditionMessage(full_model)
      )
    )
  }

  null_model <- tryCatch(

    stats::glm(
      formula = null_formula,
      data = model_data,
      family = stats::binomial()
    ),

    error = function(e) {
      e
    }
  )

  if (inherits(null_model, "error")) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = NA_character_,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = NA_real_,
        overall_LRT_df = NA_real_,
        overall_LRT_p = NA_real_,
        AIC = stats::AIC(full_model),
        converged = full_model$converged,
        warning = conditionMessage(null_model)
      )
    )
  }

  LRT_table <- tryCatch(

    stats::anova(
      null_model,
      full_model,
      test = "LRT"
    ),

    error = function(e) {
      NULL
    }
  )

  if (is.null(LRT_table)) {

    overall_LRT_chisq <- NA_real_
    overall_LRT_df <- NA_real_
    overall_LRT_p <- NA_real_

  } else {

    overall_LRT_chisq <- LRT_table$Deviance[2]
    overall_LRT_df <- LRT_table$Df[2]
    overall_LRT_p <- LRT_table$`Pr(>Chi)`[2]
  }

  exposure_object <- model_data[[exposure]]

  if (
    is.factor(exposure_object) ||
      is.character(exposure_object)
  ) {

    factor_exposure <- factor(exposure_object)

    reference_level <- levels(
      factor_exposure
    )[1]

  } else {

    reference_level <- "Per one-unit increase"
  }

  tidy_results <- tryCatch(

    broom::tidy(
      full_model,
      conf.int = TRUE,
      exponentiate = FALSE
    ),

    error = function(e) {
      NULL
    }
  )

  if (is.null(tidy_results)) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = reference_level,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = overall_LRT_chisq,
        overall_LRT_df = overall_LRT_df,
        overall_LRT_p = overall_LRT_p,
        AIC = stats::AIC(full_model),
        converged = full_model$converged,
        warning = "Could not tidy model coefficients"
      )
    )
  }

  tidy_results <- tidy_results %>%
    dplyr::filter(
      term != "(Intercept)"
    )

  if (nrow(tidy_results) == 0) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = reference_level,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = overall_LRT_chisq,
        overall_LRT_df = overall_LRT_df,
        overall_LRT_p = overall_LRT_p,
        AIC = stats::AIC(full_model),
        converged = full_model$converged,
        warning = "No exposure coefficient returned"
      )
    )
  }

  final_warning <- if (
    length(warning_messages) == 0
  ) {

    NA_character_

  } else {

    paste(
      unique(warning_messages),
      collapse = "; "
    )
  }

  tidy_results %>%
    dplyr::transmute(
      exposure = exposure,
      term = term,
      reference_level = reference_level,
      n = n_model,
      cases = n_cases,
      controls = n_controls,
      OR = exp(estimate),
      CI_lower = exp(conf.low),
      CI_upper = exp(conf.high),
      wald_z = statistic,
      wald_p = p.value,
      overall_LRT_chisq = overall_LRT_chisq,
      overall_LRT_df = overall_LRT_df,
      overall_LRT_p = overall_LRT_p,
      AIC = stats::AIC(full_model),
      converged = full_model$converged,
      warning = final_warning
    )
}
```

```{r}

```

```{r}
fit_crude_exposure_model <- function(
    data,
    exposure,
    outcome = "had_HSA"
) {

  required_variables <- c(
    outcome,
    exposure
  )

  missing_variables <- setdiff(
    required_variables,
    names(data)
  )

  if (length(missing_variables) > 0) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = NA_character_,
        n = NA_integer_,
        cases = NA_integer_,
        controls = NA_integer_,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = NA_real_,
        overall_LRT_df = NA_real_,
        overall_LRT_p = NA_real_,
        AIC = NA_real_,
        converged = FALSE,
        warning = paste(
          "Missing variable(s):",
          paste(
            missing_variables,
            collapse = ", "
          )
        )
      )
    )
  }

  model_data <- data %>%
    dplyr::select(
      dplyr::all_of(
        required_variables
      )
    ) %>%
    tidyr::drop_na()

  n_model <- nrow(model_data)

  n_cases <- sum(
    model_data[[outcome]] == 1,
    na.rm = TRUE
  )

  n_controls <- sum(
    model_data[[outcome]] == 0,
    na.rm = TRUE
  )

  outcome_variation <- dplyr::n_distinct(
    model_data[[outcome]]
  )

  exposure_variation <- dplyr::n_distinct(
    model_data[[exposure]]
  )

  if (
    n_model == 0 ||
      outcome_variation < 2 ||
      exposure_variation < 2
  ) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = NA_character_,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = NA_real_,
        overall_LRT_df = NA_real_,
        overall_LRT_p = NA_real_,
        AIC = NA_real_,
        converged = FALSE,
        warning = "Insufficient variation in outcome or exposure"
      )
    )
  }

  full_formula <- stats::reformulate(
    termlabels = exposure,
    response = outcome
  )

  null_formula <- stats::as.formula(
    paste0(outcome, " ~ 1")
  )

  warning_messages <- character()

  full_model <- tryCatch(

    withCallingHandlers(

      stats::glm(
        formula = full_formula,
        data = model_data,
        family = stats::binomial()
      ),

      warning = function(w) {

        warning_messages <<- c(
          warning_messages,
          conditionMessage(w)
        )

        invokeRestart("muffleWarning")
      }
    ),

    error = function(e) {
      e
    }
  )

  if (inherits(full_model, "error")) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = NA_character_,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = NA_real_,
        overall_LRT_df = NA_real_,
        overall_LRT_p = NA_real_,
        AIC = NA_real_,
        converged = FALSE,
        warning = conditionMessage(full_model)
      )
    )
  }

  null_model <- tryCatch(

    stats::glm(
      formula = null_formula,
      data = model_data,
      family = stats::binomial()
    ),

    error = function(e) {
      e
    }
  )

  if (inherits(null_model, "error")) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = NA_character_,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = NA_real_,
        overall_LRT_df = NA_real_,
        overall_LRT_p = NA_real_,
        AIC = stats::AIC(full_model),
        converged = full_model$converged,
        warning = conditionMessage(null_model)
      )
    )
  }

  LRT_table <- tryCatch(

    stats::anova(
      null_model,
      full_model,
      test = "LRT"
    ),

    error = function(e) {
      NULL
    }
  )

  if (is.null(LRT_table)) {

    overall_LRT_chisq <- NA_real_
    overall_LRT_df <- NA_real_
    overall_LRT_p <- NA_real_

  } else {

    overall_LRT_chisq <- LRT_table$Deviance[2]
    overall_LRT_df <- LRT_table$Df[2]
    overall_LRT_p <- LRT_table$`Pr(>Chi)`[2]
  }

  exposure_object <- model_data[[exposure]]

  if (
    is.factor(exposure_object) ||
      is.character(exposure_object)
  ) {

    factor_exposure <- factor(exposure_object)

    reference_level <- levels(
      factor_exposure
    )[1]

  } else {

    reference_level <- "Per one-unit increase"
  }

  tidy_results <- tryCatch(

    broom::tidy(
      full_model,
      conf.int = TRUE,
      exponentiate = FALSE
    ),

    error = function(e) {
      NULL
    }
  )

  if (is.null(tidy_results)) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = reference_level,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = overall_LRT_chisq,
        overall_LRT_df = overall_LRT_df,
        overall_LRT_p = overall_LRT_p,
        AIC = stats::AIC(full_model),
        converged = full_model$converged,
        warning = "Could not tidy model coefficients"
      )
    )
  }

  tidy_results <- tidy_results %>%
    dplyr::filter(
      term != "(Intercept)"
    )

  if (nrow(tidy_results) == 0) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = reference_level,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = overall_LRT_chisq,
        overall_LRT_df = overall_LRT_df,
        overall_LRT_p = overall_LRT_p,
        AIC = stats::AIC(full_model),
        converged = full_model$converged,
        warning = "No exposure coefficient returned"
      )
    )
  }

  final_warning <- if (
    length(warning_messages) == 0
  ) {

    NA_character_

  } else {

    paste(
      unique(warning_messages),
      collapse = "; "
    )
  }

  tidy_results %>%
    dplyr::transmute(
      exposure = exposure,
      term = term,
      reference_level = reference_level,
      n = n_model,
      cases = n_cases,
      controls = n_controls,
      OR = exp(estimate),
      CI_lower = exp(conf.low),
      CI_upper = exp(conf.high),
      wald_z = statistic,
      wald_p = p.value,
      overall_LRT_chisq = overall_LRT_chisq,
      overall_LRT_df = overall_LRT_df,
      overall_LRT_p = overall_LRT_p,
      AIC = stats::AIC(full_model),
      converged = full_model$converged,
      warning = final_warning
    )
}


```

```{r}
crude_results <- imap_dfr(
  usable_exposure_families,
  function(variables, family) {

    map_dfr(
      variables,
      ~ fit_crude_exposure_model(
        data = analysis_df,
        exposure = .x,
        outcome = outcome_var
      )
    ) %>%
      mutate(
        exposure_family = family,
        model_type = "Crude",
        .before = 1
      )
  }
)
```

```{r}
test_exposure <- usable_exposure_families[[1]][1]

test_exposure
```

```{r}
test_result <- fit_crude_exposure_model(
  data = analysis_df,
  exposure = test_exposure,
  outcome = outcome_var
)

test_result
```

```{r}
crude_results <- purrr::imap_dfr(
  usable_exposure_families,
  function(variables, family) {

    family_results <- purrr::map_dfr(
      variables,
      function(variable_name) {

        fit_crude_exposure_model(
          data = analysis_df,
          exposure = variable_name,
          outcome = outcome_var
        )
      }
    )

    family_results %>%
      dplyr::mutate(
        exposure_family = family,
        model_type = "Crude",
        .before = 1
      )
  }
)
```

```{r}
crude_results %>%
  dplyr::distinct(
    exposure_family,
    exposure
  ) %>%
  nrow()
```

Checking model problems

```{r}
crude_model_problems <- crude_results %>%
  dplyr::filter(
    !converged |
      !is.na(warning)
  ) %>%
  dplyr::distinct(
    exposure_family,
    exposure,
    n,
    cases,
    controls,
    converged,
    warning
  ) %>%
  dplyr::arrange(
    exposure_family,
    exposure
  )
```

One row per exposure variable table

```{r}
crude_overall_tests <- crude_results %>%
  dplyr::distinct(
    exposure_family,
    exposure,
    n,
    cases,
    controls,
    overall_LRT_chisq,
    overall_LRT_df,
    overall_LRT_p,
    AIC,
    converged,
    warning
  )
crude_overall_tests <- crude_overall_tests %>%
  dplyr::group_by(
    exposure_family
  ) %>%
  dplyr::mutate(
    LRT_p_FDR_within_family = stats::p.adjust(
      overall_LRT_p,
      method = "BH"
    )
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    LRT_p_FDR_global = stats::p.adjust(
      overall_LRT_p,
      method = "BH"
    )
  ) %>%
  dplyr::arrange(
    exposure_family,
    overall_LRT_p
  )
```

```{r}
crude_results <- crude_results %>%
  dplyr::left_join(
    crude_overall_tests %>%
      dplyr::select(
        exposure_family,
        exposure,
        LRT_p_FDR_within_family,
        LRT_p_FDR_global
      ),
    by = c(
      "exposure_family",
      "exposure"
    )
  )
```

```{r}
format_p_value <- function(p) {

  dplyr::case_when(
    is.na(p) ~ NA_character_,
    p < 0.001 ~ "<0.001",
    TRUE ~ sprintf("%.3f", p)
  )
}
```

```{r}
crude_results_formatted <- crude_results %>%
  dplyr::mutate(

    OR_95CI = dplyr::case_when(
      is.na(OR) ~ NA_character_,
      TRUE ~ sprintf(
        "%.2f (%.2f–%.2f)",
        OR,
        CI_lower,
        CI_upper
      )
    ),

    Wald_p_display = format_p_value(
      wald_p
    ),

    LRT_p_display = format_p_value(
      overall_LRT_p
    ),

    family_FDR_display = format_p_value(
      LRT_p_FDR_within_family
    ),

    global_FDR_display = format_p_value(
      LRT_p_FDR_global
    )
  ) %>%
  dplyr::select(
    exposure_family,
    exposure,
    term,
    reference_level,
    n,
    cases,
    controls,
    OR_95CI,
    OR,
    CI_lower,
    CI_upper,
    wald_z,
    wald_p,
    Wald_p_display,
    overall_LRT_chisq,
    overall_LRT_df,
    overall_LRT_p,
    LRT_p_display,
    LRT_p_FDR_within_family,
    family_FDR_display,
    LRT_p_FDR_global,
    global_FDR_display,
    AIC,
    converged,
    warning
  ) %>%
  dplyr::arrange(
    exposure_family,
    overall_LRT_p,
    exposure,
    term
  )
```

Table per family

Need to produce prettier tables

Need to ensure correct reference level (need to refactorise at start before this analysis
