---
title: "Haemangiosarcoma dog descriptive stats"
format: html
editor: visual
---

## Haemangiosarcoma GRLS cohort descriptives and MST

## NB. this is all (presumptive and confirmatory diagnosed) cases

```{r}
library(tidyverse)
library(dplyr)
library(survival)
library(ggplot2)
library(tibble)
library(kableExtra)
#devtools::install_github("zabore/ezfun")
ezfun::set_ccf_palette("contrast")
library(lubridate)
library(ggsurvfit)
library(gtsummary)
library(tidycmprsk)
library(gtable)
#devtools::install_github("zabore/condsurv")
library(condsurv)
library(survminer)
library(cardx)
library(finalfit)
#whole grls cohort
cohort <- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_cohort_all_RFs_data_tidied_from_cox.csv")
HSA_dogs <- cohort %>%
  filter(had_HSA==1)

endpoints <- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Data/study_endpoints.csv")
#
#keeping only relevant columns from each df
HSA_dogs <- HSA_dogs %>%
  dplyr::select(c(subject_id,age_at_final_date))
endpoints <- endpoints %>%
  dplyr::select(c(subject_id,year_in_study,diagnosis_date,tracked_condition,location,phenotype,status,tier_of_confidence,is_cause_of_death))
HSA_dogs <- HSA_dogs %>%
  left_join(endpoints,by="subject_id")
#NB some HSA dogs have multi cancers - 606 rows but 490 unique IDs
#only keep haemangiosarcoma relevant rows - 502 dogs
HSA_dogs2<- HSA_dogs %>%
  filter(grepl("Hemangiosarc", tracked_condition))
#check the reason for 10 extra rows is that multi rows same dog - it is
multi_HSA <- HSA_dogs2 %>%
  add_count(subject_id, name = "count") %>%  # adds a count column while keeping all other cols
  filter(count > 1) %>%
  arrange(desc(count))


#incorp in survival time days - get date of diagnosis from endpoints and date of death from profile 
profile <- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Data/dog_profile.csv")
profile <- profile %>%
  dplyr::select(c(subject_id,enrolled_date,death_date,is_euthanized,birth_date))

HSA_dogs2 <- HSA_dogs2 %>%
  left_join(profile,by="subject_id")

```

## Time enrolled in study, median age at enrollment

```{r}
study_times <- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_study_pop_with_final_date.csv")

#link to dog profile to calc age at enrollment and time enrolled in study]

study_times_HSA_dogs2 <- profile %>% left_join(study_times, by="subject_id")


#calc difference between enrollment date and final date

#all dogs
study_times_HSA_dogs2 <- study_times_HSA_dogs2 %>%
mutate(
  start_date = as.Date(enrolled_date),
  end_date = as.Date(final_date),
  days_between = as.numeric(end_date -start_date)
)

study_times_HSA_dogs2 <- study_times_HSA_dogs2 %>%
  mutate(
    years_between = days_between / 365.25
  )

all_dogs <- study_times_HSA_dogs2 %>%
  summarise(
    median_years = median(years_between, na.rm = TRUE),
    min_years = min(years_between, na.rm = TRUE),
    max_years = max(years_between, na.rm = TRUE),
    IQR_years = IQR(years_between, na.rm = TRUE),
    q1_years = quantile(years_between, 0.25, na.rm = TRUE),
    q3_years = quantile(years_between, 0.75, na.rm = TRUE)
  )

#HSA dogs
HSA_dogs_summary <- study_times_HSA_dogs2 %>%
  filter(had_HSA==1)%>%
  summarise(
    median_years = median(years_between, na.rm = TRUE),
    min_years = min(years_between, na.rm = TRUE),
    max_years = max(years_between, na.rm = TRUE),
    IQR_years = IQR(years_between, na.rm = TRUE),
    q1_years = quantile(years_between, 0.25, na.rm = TRUE),
    q3_years = quantile(years_between, 0.75, na.rm = TRUE)
  )

#non-HSA dogs
non_HSA_dogs_summary <- study_times_HSA_dogs2 %>%
  filter(had_HSA==0)%>%
  summarise(
    median_years = median(years_between, na.rm = TRUE),
    min_years = min(years_between, na.rm = TRUE),
    max_years = max(years_between, na.rm = TRUE),
    IQR_years = IQR(years_between, na.rm = TRUE),
    q1_years = quantile(years_between, 0.25, na.rm = TRUE),
    q3_years = quantile(years_between, 0.75, na.rm = TRUE)
  )
```

Proportion of dogs still enrolled by Dec 31 2024

```{r}
#report number of dogs with Dec 31 2024 date
still_enrolled <- study_times_HSA_dogs2 %>%
  filter(final_date == "2024-12-31")
```

## Age

```{r}
non_HSA_dogs_summary_age <- cohort %>%
  filter(had_HSA==0)%>%
  summarise(
    median_years = median(age_at_final_date, na.rm = TRUE),
    min_years = min(age_at_final_date, na.rm = TRUE),
    max_years = max(age_at_final_date, na.rm = TRUE),
    IQR_years = IQR(age_at_final_date, na.rm = TRUE),
    q1_years = quantile(age_at_final_date, 0.25, na.rm = TRUE),
    q3_years = quantile(age_at_final_date, 0.75, na.rm = TRUE)
  )
non_HSA_dogs_summary_age

HSA_dogs_summary_age <- cohort %>%
  filter(had_HSA==1)%>%
  summarise(
    median_years = median(age_at_final_date, na.rm = TRUE),
    min_years = min(age_at_final_date, na.rm = TRUE),
    max_years = max(age_at_final_date, na.rm = TRUE),
    IQR_years = IQR(age_at_final_date, na.rm = TRUE),
    q1_years = quantile(age_at_final_date, 0.25, na.rm = TRUE),
    q3_years = quantile(age_at_final_date, 0.75, na.rm = TRUE)
  )

HSA_dogs_summary_age

all_HSA_dogs_summary_age <- cohort %>%
  summarise(
    median_years = median(age_at_final_date, na.rm = TRUE),
    min_years = min(age_at_final_date, na.rm = TRUE),
    max_years = max(age_at_final_date, na.rm = TRUE),
    IQR_years = IQR(age_at_final_date, na.rm = TRUE),
    q1_years = quantile(age_at_final_date, 0.25, na.rm = TRUE),
    q3_years = quantile(age_at_final_date, 0.75, na.rm = TRUE)
  )

all_HSA_dogs_summary_age
```

## Haemangiosarcoma locations

```{r}
#these are the broader locations -cardiac, splenic, visceral, cutaneous
locations_main <- HSA_dogs2 %>%
  count(tracked_condition, name = "count") %>%
  mutate(percent = round(100 * count / sum(count), 1)) %>%
  arrange(desc(count))

#create table count of ALL locations from locations column NB. more than one location per dog
locations <- HSA_dogs2$location

# Step 1: Replace " and " with commas to standardize delimiters
locations_clean <- gsub(" and ", ", ", locations, ignore.case = TRUE)

# Step 2: Remove unwanted strings (e.g., "suspect", "presum", "likely", directional terms)
# Use regex and ignore case for better coverage
locations_clean <- tolower(locations_clean)  # convert early for consistent matching
locations_clean <- gsub("suspect[^,]*", "", locations_clean)
locations_clean <- gsub("presum[^,]*", "", locations_clean)
locations_clean <- gsub("likel[^,]*", "", locations_clean)
locations_clean <- gsub("probable[^,]*", "", locations_clean)
locations_clean <- gsub("\\bleft\\b", "", locations_clean)
locations_clean <- gsub("\\bright\\b", "", locations_clean)

# Optional: remove "mass on", "in the", etc. if those phrases occur
locations_clean <- gsub("mass on ", "", locations_clean)
locations_clean <- gsub("in the ", "", locations_clean)
locations_clean <- gsub("area of ", "", locations_clean)
locations_clean <- gsub("\\s+", " ", locations_clean)  # condense multiple spaces

# Step 3: Split into individual locations by comma
location_split <- strsplit(locations_clean, ",")

# Step 4: Trim and flatten into a vector
all_locations <- trimws(unlist(location_split))

# Step 5: Standardize common synonyms (e.g., "cardiac", "ventricle" → "heart")
all_locations[all_locations %in% c("heart", "cardiac", "right atrial cavity", "ventricle", "right atrium","atrium","atrium/auricle","atrial","heart (atrium)","heart (ventricular wall)","heart (auricle)","subtle infiltration in section of atrium)","heart (mass","heart ( auricle)","ventricular free wall)","heart ( atrium)","heart ( auricle)", "near heart base","heart (atrium","heart (atrial tissue)" ,"heart (","heart base","heart; atrium","heart (interventricular septum just under the annulus of the posterior tricuspid valve leaflet)","myocardium/heart","myocardium","atrial cavity","atrio-ventricular groove", "atrium (","atrium of the heart","heart ( aricular)","heart; atrium","heart base","myocardium)","ventricle","heart ( auricle","ventricle)","auricle")] <- "heart"
all_locations[all_locations %in% c("skin (lower lip", "skin", "skin (abdomen)", "scrotum","dermal","haired skin")] <- "skin"
all_locations[all_locations %in% c("lungs", "lung", "thoracic cavity", "thorax","all lung lobes","lung.","lungs (","lungs (")] <- "lung"
all_locations[all_locations %in% c("liver", "hepatic","liver (","liver: disseminated hemangiosarcoma")] <- "liver"
all_locations[all_locations %in% c("spleen", "splenic","spleen (")] <- "spleen"
all_locations[all_locations %in% c("abdominal cavity", "abdomen")] <- "abdomen"
all_locations[all_locations %in% c("jejunum", "small intestine","intestinal serosa","intestines","small intestine (muscular wall)","small intestine (muscularis mucosa)")] <- "intestine"
all_locations[all_locations %in% c("kidneys", "kidney","kidney (possible)","bilateral kidneys"  )] <- "kidney"
all_locations[all_locations %in% c("retroperitoneal mass", "retroperitoneum ("  ,"retroperitoneal cavity","retroperitoneal","retroperitoneal space")] <- "retroperitoneum"
all_locations[all_locations %in% c("cerebellum","cerebreum","cerebrum","cerebellum)","brain (cerebrum","brain (cerebrum)")] <- "brain"
all_locations[all_locations %in% c("mesenteric adipose tissue","peripancreatic adipose","adipose tissue; cranial thigh","fibroadipose tissue:","fibroadipose tissue (","fat (mesentery?)","fat (tibial area)","adipose")] <- "adipose tissue"
all_locations[all_locations %in% c("gastric serosa","stomach (muscularis mucosa)")] <- "stomach"

all_locations[all_locations %in% c("to the urinary bladder caudal aspect of the abdomen","bladder","urinary bladder")] <- "bladder"
all_locations[all_locations %in% c("tricept","skeletal muscle/adipose tissue ","epaxial muscle","muscle","skeletal muscle/adipose tissue","muscle ( flank)","skeletal muscle")] <- "muscle"
all_locations[all_locations %in% c("adrenal")] <- "adrenal gland"
all_locations[all_locations %in% c("rib"," bone (proximal humerus)","bone (proximal humerus)","carpus","15 vertebral body","l5 vertebral body","tibial)","bone (proximal humerus)")] <- "bone"
all_locations[all_locations %in% c("muscle (diaphragm")] <- "diaphragm"
#NB skin techincally cutis
all_locations[all_locations %in% c("subcutaneous","subcutaneous mass ( flank)","lateral mid thorax","subcutis")] <- "skin"
all_locations[all_locations %in% c("adrenal glands")] <- "adrenal gland"
all_locations[all_locations %in% c("mesentery/omentum","mesenetery")] <- "mesentery"
#combine all the weird misc ones into "other" and "unknown"
all_locations[all_locations %in% c("abdomen)","abdominal fluid","abdominal mass","body wall","caudal abdominal mass","hindlimb amputation site","leg mass","mid abdomen -","mouth","mucosal lined tissue","side","shoulder mass", "shoulder/scapula","shoulder","front leg","thoaracic inlet","thoracic deep to scapula","chest","digit","abdomen")] <- "other"
all_locations[all_locations %in% c("(","masses","possibly spleen or liver","previously found in spleen","unspecified sites: tisse
","spleen previously removed","unspecified sites: tisse")] <- "unknown"


# Step 6: Remove empty strings (resulting from deleted words)
all_locations <- all_locations[all_locations != ""]

# Step 7: Create and print counts
location_counts <- sort(table(all_locations), decreasing = TRUE)
print(location_counts)



# Step 7: Create table of counts and percentages
location_counts <- sort(table(all_locations), decreasing = TRUE)

# Convert to data.frame
location_df <- as.data.frame(location_counts)
colnames(location_df) <- c("location", "count")

# Add percentage column
location_df$percent <- round(100 * location_df$count / sum(location_df$count), 1)

# View result
print(location_df)

# Combine count and percent into one column like "420 (80)"
location_df$count_percent <- paste0(location_df$count, " (", location_df$percent, ")")

# Optional: keep just location and the combined column
location_summary <- location_df[, c("location","count","count_percent", "count_percent")]

# View result
print(location_summary)
#write to csv for manuscript
write.csv(location_summary,"C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/GRLS_HSA_location_counts.csv")
```

Location barplot

```{r,eval=FALSE,echo=FALSE}

# Create the data
location_data <- tribble(
  ~Location, ~Number_percent,
  "spleen", "252 (23.9)",
  "cardiac", "240 (22.8)",
  "lung", "170 (16.1)",
  "liver", "155 (14.7)",
  "kidney", "59 (5.6)",
  "other", "31 (2.9)",
  "skin", "17 (1.6)",
  "omentum", "15 (1.4)",
  "adrenal gland", "14 (1.3)",
  "adipose tissue", "12 (1.1)",
  "unknown", "12 (1.1)",
  "muscle", "11 (1)",
  "brain", "9 (0.9)",
  "intestine", "9 (0.9)",
  "pancreas", "7 (0.7)",
  "bone", "6 (0.6)",
  "diaphragm", "6 (0.6)",
  "retroperitoneum", "6 (0.6)",
  "bladder", "4 (0.4)",
  "spine", "3 (0.3)",
  "stomach", "3 (0.3)",
  "thyroid", "3 (0.3)",
  "esophagus", "2 (0.2)",
  "lymph node", "2 (0.2)",
  "mesentery", "2 (0.2)",
  "spinal cord", "2 (0.2)",
  "prostate area", "1 (0.1)"
)

location_data2 <- location_data

# Extract count and percentage
location_data <- location_data %>%
  separate(Number_percent, into = c("Number", "Percent"), sep = " ", convert = TRUE) %>%
  mutate(Percent = parse_number(Percent))

# Plot
ggplot(location_data, aes(x = fct_reorder(Location, Number), y = Number)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Distribution of Tumor Locations",
    x = NULL,
    y = "Number of Cases"
  ) +
  theme_minimal(base_size = 14)
```

```{r}
library(dplyr)
library(ggplot2)
library(forcats)

plot2<- location_df %>%
  ggplot(aes(x = fct_reorder(location, count), y = count)) +
  geom_col(fill = "grey30", colour = "black", width = 0.8) +
  geom_text(aes(label = count_percent),
            hjust = -0.15,
            size = 4) +
  coord_flip() +
  expand_limits(y = max(location_df$count) * 1.15) +
  labs(
    title = "Distribution of Haemangiosarcoma Locations",
    x = NULL,
    y = "Number of cases"
  ) +
  theme_classic(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

ggsave("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/GRLS_HSA_locations_barplot.tiff", plot = plot2, width = 8, height = 6, dpi = 600)

# Plot
ggplot(location_df, aes(x = fct_reorder(location, count), y = count)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = paste0(percent, "%")), 
            hjust = -0.1, size = 4) +
  coord_flip() +
  labs(
    title = "Haemangiosarcoma Locations by Frequency",
    x = NULL,
    y = "Number of Cases"
  ) +
  theme_minimal(base_size = 14) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))  # extra space for labels



plot <- ggplot(location_df, aes(x = fct_reorder(location, count), y = count)) +
  geom_col(fill = "grey30", width = 0.7) +
  geom_text(aes(label = paste0(percent, "%")), 
            hjust = -0.1, size = 4.2) +
  coord_flip() +
  labs(
    title = NULL,
    x = NULL,
    y = "Frequency of haemangiosarcoma locations"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text = element_text(size = 12),
    axis.title.y = element_text(size = 13),
    axis.title.x = element_text(size = 13),
    plot.margin = margin(10, 30, 10, 10),
    axis.line = element_line(colour = "black")
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2)))


ggsave("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/GRLS_HSA_locations_barplot.tiff", plot = plot, width = 8, height = 6, dpi = 600)

```

## Haemangiosarcoma location MSTs

Survival time days calculation

```{r}
#funct to calc time between events
source("C:/Users/ctaylor18/GitHub/GRLS_analyses/Code/functions/age_calc_funct.R")

#convert char dates to date format
#HSA_dogs2$diagnosis_date2 <- as.Date(HSA_dogs2$diagnosis_date)
#HSA_dogs2$death_date2 <- as.Date(HSA_dogs2$death_date)


HSA_dogs3 <- age_calc(HSA_dogs2,"subject_id","diagnosis_date","death_date")

#across all these rows replace any negative values with 0
HSA_dogs3 <- HSA_dogs3 %>%
  mutate(across(13:16, ~ as.numeric(.))) %>%   # convert to numeric
  mutate(across(13:16, ~ ifelse(. < 0, 0, .))) # replace negatives with 0


#need to create a status variable for calc MST - status by 31/12/24. 0 = censored, 1=dead
#outcome where if do not have a death date = 0 and if do =1 
HSA_dogs3 <- HSA_dogs3 %>%
  mutate(outcome = if_else(time_between_diagnosis_date_death_date_days == "not recorded", 0, 1))



column_name <- "outcome"

HSA_dogs4 <- HSA_dogs3 %>%
  mutate(status = ifelse(is.na(.data[[column_name]]), 0, 1))

```

```{r}
splenic<- HSA_dogs4 %>%
  filter(grepl("splen",tracked_condition))
cardiac <- HSA_dogs4%>%
  filter(grepl("card",tracked_condition))
visceral<-HSA_dogs4%>%
  filter(grepl("visc",tracked_condition))
cutaneous<- HSA_dogs4 %>%
  filter(grepl("cut",tracked_condition))
other<- HSA_dogs4 %>%
  filter(grepl("other",tracked_condition))

```

### All HSA

```{r}
survfit(Surv(HSA_dogs4$time_between_diagnosis_date_death_date_days, HSA_dogs4$status) ~ 1, data = HSA_dogs4) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) +  add_confidence_interval() +
  add_risktable()

summary(HSA_dogs4$time_between_diagnosis_date_death_date_days)

survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = HSA_dogs4)
#probability of surviving to one year
summary(survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = HSA_dogs4), times = 365.25)

#1 year survival rate
survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = HSA_dogs4) %>% 
  tbl_survfit(
    times = 365.25,
    label_header = "**1-year survival (95% CI)**"
  )

quantile(HSA_dogs4$time_between_diagnosis_date_death_date_days, probs = c(0.25, 0.75), na.rm = TRUE)

```

### Splenic

```{r}
survfit(Surv(splenic$time_between_diagnosis_date_death_date_days, splenic$status) ~ 1, data = splenic) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) +  add_confidence_interval() +
  add_risktable()

summary(splenic$time_between_diagnosis_date_death_date_days)
#median survival time is 5 days 

survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = splenic)
#probability of surviving to one year
summary(survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = splenic), times = 365.25)

#1 year survival rate
survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = splenic) %>% 
  tbl_survfit(
    times = 365.25,
    label_header = "**1-year survival (95% CI)**"
  )

quantile(splenic$time_between_diagnosis_date_death_date_days, probs = c(0.25, 0.75), na.rm = TRUE)
range(splenic$time_between_diagnosis_date_death_date_days,na.rm = TRUE)
```

### Cardiac

```{r}
survfit(Surv(cardiac$time_between_diagnosis_date_death_date_days, cardiac$status) ~ 1, data = cardiac) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) +  add_confidence_interval() +
  add_risktable()

summary(cardiac$time_between_diagnosis_date_death_date_days)
#median survival time is 5 days 

survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = cardiac)
#probability of surviving to one year
summary(survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = cardiac), times = 365.25)

#1 year survival rate
survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = cardiac) %>% 
  tbl_survfit(
    times = 365.25,
    label_header = "**1-year survival (95% CI)**"
  )

quantile(cardiac$time_between_diagnosis_date_death_date_days, probs = c(0.25, 0.75), na.rm = TRUE)
range(cardiac$time_between_diagnosis_date_death_date_days,na.rm = TRUE)
```

### Visceral

```{r}
survfit(Surv(visceral$time_between_diagnosis_date_death_date_days, visceral$status) ~ 1, data = visceral) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) +  add_confidence_interval() +
  add_risktable()

summary(visceral$time_between_diagnosis_date_death_date_days)

survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = visceral)
#probability of surviving to one year
#NB needs to be set at 364 here as the single 365d case makes this not work as everyone has had the case by the 365d point
summary(survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = visceral), times = 365.25)
 
#1 year survival rate - doesnt work as all cases dead by 1y
#survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = visceral) %>% 
 # tbl_survfit(
  #  times = 365.25,
  #  label_header = "**1-year survival (95% CI)**"
 # )


nrow(visceral)

summary(visceral$time_between_diagnosis_date_death_date_days)

table(visceral$status, useNA = "ifany")

#errors with above attempts to problem solve - not working
fit <- survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = visceral)

# Extract components
fit_df <- data.frame(
  time = fit$time,
  surv = fit$surv,
  lower = fit$lower,
  upper = fit$upper
)

# Check the structure of the new data frame
str(fit)

report_time <- max(fit_df$time[fit_df$surv > 0])

tbl_survfit(
  fit,
  times = report_time,
  label_header = paste0("**Survival at ", report_time, " days (95% CI)**")
)




quantile(visceral$time_between_diagnosis_date_death_date_days, probs = c(0.25, 0.75), na.rm = TRUE)
range(visceral$time_between_diagnosis_date_death_date_days,na.rm = TRUE)
```

### Cutaneous

```{r}
survfit(Surv(cutaneous$time_between_diagnosis_date_death_date_days, cutaneous$status) ~ 1, data = cutaneous) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) +  add_confidence_interval() +
  add_risktable()

summary(cutaneous$time_between_diagnosis_date_death_date_days)
#median survival time is 5 days 

survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = cutaneous)
#probability of surviving to one year
summary(survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = cutaneous), times = 365.25)

#1 year survival rate
survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = cutaneous) %>% 
  tbl_survfit(
    times = 365.25,
    label_header = "**1-year survival (95% CI)**"
  )

quantile(cutaneous$time_between_diagnosis_date_death_date_days, probs = c(0.25, 0.75), na.rm = TRUE)
range(cutaneous$time_between_diagnosis_date_death_date_days,na.rm = TRUE)
```

### Other

```{r}
survfit(Surv(other$time_between_diagnosis_date_death_date_days, other$status) ~ 1, data = other) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) +  add_confidence_interval() +
  add_risktable()

summary(other$time_between_diagnosis_date_death_date_days)
#median survival time is 5 days 

survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = other)
#probability of surviving to one year
summary(survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = other), times = 365.25)

#1 year survival rate
survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = other) %>% 
  tbl_survfit(
    times = 365.25,
    label_header = "**1-year survival (95% CI)**"
  )
```

### Combined KM survival curve plot for all and all locations individual

```{r}
HSA_dogs4 <- HSA_dogs4 %>%
  mutate(cancer_group = recode(tracked_condition,
                               'Hemangiosarcoma - visceral'="Visceral",
                               'Hemangiosarcoma - splenic'= "Splenic",
                               'Hemangiosarcoma - cardiac'="Cardiac",
                               'Hemangiosarcoma - cutaneous'="Cutaneous",
                               'Hemangiosarcoma - other/not specified'="Other")
  )#


#set levels for cancer_group
HSA_dogs4 <- HSA_dogs4%>%
  mutate(cancer_group = factor(cancer_group, levels = c("Splenic","Cardiac"  ,"Visceral", "Cutaneous","Other")))

# Fit the survival model
fit <- survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ cancer_group, data = HSA_dogs4)

# Plot the Kaplan-Meier curves
ggsurvplot(fit, data = HSA_dogs4, palette = "Dark2", risk.table = TRUE)



library(RColorBrewer)

KM_plot_4 <- ggsurvplot(
  fit, 
  data = HSA_dogs4,
  palette = brewer.pal(n = 5, name = "Set1"),  # Use a scientific palette
  conf.int = TRUE,  # Add confidence intervals
  conf.int.alpha = 0.1,  # Adjust transparency (0 = fully transparent, 1 = fully opaque)
  risk.table = FALSE,  # Dont show risk table
  ggtheme = theme_minimal(),  # Use a clean theme for manuscripts
  legend.title = "Cancer Type",
  xlab = "Time (Days)",
  ylab = "Survival Probability"
)

KM_plot_4



ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls.svg", plot = KM_plot_4$plot, width = 15, height = 10, dpi = 300)

ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls.pdf", plot = KM_plot_4$plot, width = 15, height = 10, dpi = 300)


ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls.tif", plot = KM_plot_4$plot, width = 15, height = 10, dpi = 300)



#remove "other category from plot as fairly uninformative
HSA_dogs5 <- HSA_dogs4 %>%
  filter(cancer_group != "Other")
fit <- survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ cancer_group, data = HSA_dogs5)
library(RColorBrewer)

KM_plot_5 <- ggsurvplot(
  fit, 
  data = HSA_dogs5,
  palette = brewer.pal(n = 5, name = "Set1"),  # Use a scientific palette
  conf.int = TRUE,  # Add confidence intervals
  conf.int.alpha = 0.1,  # Adjust transparency (0 = fully transparent, 1 = fully opaque)
  risk.table = FALSE,  # Dont show risk table
  ggtheme = theme_minimal(),  # Use a clean theme for manuscripts
  legend.title = "Cancer Type",
  xlab = "Time (Days)",
  ylab = "Survival Probability"
)

KM_plot_5



ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_no_other.svg", plot = KM_plot_5$plot, width = 15, height = 10, dpi = 300)

ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_no_other.pdf", plot = KM_plot_5$plot, width = 15, height = 10, dpi = 300)


ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_no_other.tif", plot = KM_plot_5$plot, width = 15, height = 10, dpi = 300)

```

## Incidence per 100 dog years

Use profile dataset to calculate time between enrolled date and death_date OR 31st Dec 2024 if not dead yet

```{r}
source("C:/Users/ctaylor18/GitHub/GRLS_analyses/Code/functions/age_calc_funct.R")

#add death date 31/12/24 for all dogs with no death date recorded into death_date_2
profile <- profile %>%
  mutate(death_date_2 = if_else(death_date == "", "2024-12-31", death_date))
time_in_study <- age_calc(profile,"subject_id","enrolled_date","death_date_2")

#as numeric
time_in_study$time_between_enrolled_date_death_date_2_years <-as.numeric(time_in_study$time_between_enrolled_date_death_date_2_years)

# Sum total dog-years
total_dog_years <- sum(time_in_study$time_between_enrolled_date_death_date_2_years, na.rm = TRUE)

# Number of dogs
cases <- 490

# Incidence rate per 100 dog-years
incidence_rate <- (cases / total_dog_years) * 100

incidence_rate

#95%CI 
# Calculate 95% CI using Poisson distribution
alpha <- 0.05

lower_limit <- 0.5 * qchisq(alpha / 2, 2 * cases)   # lower bound for cases
upper_limit <- 0.5 * qchisq(1 - alpha / 2, 2 * (cases + 1)) # upper bound for cases

lower_rate <- (lower_limit / total_dog_years) * 100
upper_rate <- (upper_limit / total_dog_years) * 100

# Print results
cat("Incidence rate of first haemangiosarcoma diagnosis per 100 dog-years :", round(incidence_rate, 3), "\n")
cat("95% CI:", round(lower_rate, 3), "to", round(upper_rate, 3), "\n")



#incidence for multiple cases though:
# Number of dogs
cases <- 526

# Incidence rate per 100 dog-years
incidence_rate <- (cases / total_dog_years) * 100

incidence_rate

#95%CI 
# Calculate 95% CI using Poisson distribution
alpha <- 0.05

lower_limit <- 0.5 * qchisq(alpha / 2, 2 * cases)   # lower bound for cases
upper_limit <- 0.5 * qchisq(1 - alpha / 2, 2 * (cases + 1)) # upper bound for cases

lower_rate <- (lower_limit / total_dog_years) * 100
upper_rate <- (upper_limit / total_dog_years) * 100

# Print results
cat("Incidence rate of haemangiosarcoma per 100 dog-years:", round(incidence_rate, 3), "\n")
cat("95% CI:", round(lower_rate, 3), "to", round(upper_rate, 3), "\n")

```

## Confirmed diagnoses descriptives and MST and incidence

```{r}
cohort_info_for_tiers <- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Data/study_endpoints.csv")

cohort_info_for_tiers<- cohort_info_for_tiers %>%
  mutate(across(where(is.character), str_to_lower))%>%
  select(subject_id, tracked_condition,tier_of_confidence)%>%
   dplyr::filter(tier_of_confidence==1)%>%
  dplyr::filter(str_detect(tracked_condition, "hemangios"))

#create confirmed HSA cohort and then run analysis as above for that
HSA_dogs_confirmed <- cohort_info_for_tiers %>%
  left_join(HSA_dogs2, by="subject_id") %>%
  rename(tracked_condition =tracked_condition.x)%>%
  rename(tier_of_confidence = tier_of_confidence.x)
```

```{r}
study_times <- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_study_pop_with_final_date.csv")

#link to dog profile to calc age at enrollment and time enrolled in study]

study_times_HSA_dogs2_conf <- HSA_dogs_confirmed %>% left_join(study_times, by="subject_id")


#calc difference between enrollment date and final date

#all dogs
study_times_HSA_dogs2_conf <- study_times_HSA_dogs2_conf %>%
mutate(
  start_date = as.Date(enrolled_date),
  end_date = as.Date(final_date),
  days_between = as.numeric(end_date -start_date)
)

study_times_HSA_dogs2_conf <- study_times_HSA_dogs2_conf %>%
  mutate(
    years_between = days_between / 365.25
  )

all_dogs_conf <- study_times_HSA_dogs2_conf %>%
  summarise(
    median_years = median(years_between, na.rm = TRUE),
    min_years = min(years_between, na.rm = TRUE),
    max_years = max(years_between, na.rm = TRUE),
    IQR_years = IQR(years_between, na.rm = TRUE),
    q1_years = quantile(years_between, 0.25, na.rm = TRUE),
    q3_years = quantile(years_between, 0.75, na.rm = TRUE)
  )

#HSA dogs
HSA_dogs_summary_conf <- study_times_HSA_dogs2_conf %>%
  filter(had_HSA==1)%>%
  summarise(
    median_years = median(years_between, na.rm = TRUE),
    min_years = min(years_between, na.rm = TRUE),
    max_years = max(years_between, na.rm = TRUE),
    IQR_years = IQR(years_between, na.rm = TRUE),
    q1_years = quantile(years_between, 0.25, na.rm = TRUE),
    q3_years = quantile(years_between, 0.75, na.rm = TRUE)
  )


```

## Age

```{r}



HSA_dogs_summary_age_conf <- study_times_HSA_dogs2_conf %>%
  summarise(
    median_years = median(age_at_final_date, na.rm = TRUE),
    min_years = min(age_at_final_date, na.rm = TRUE),
    max_years = max(age_at_final_date, na.rm = TRUE),
    IQR_years = IQR(age_at_final_date, na.rm = TRUE),
    q1_years = quantile(age_at_final_date, 0.25, na.rm = TRUE),
    q3_years = quantile(age_at_final_date, 0.75, na.rm = TRUE)
  )
```

## Haemangiosarcoma locations

```{r}
#these are the broader locations -cardiac, splenic, visceral, cutaneous
locations_main2 <- HSA_dogs_confirmed %>%
  count(tracked_condition, name = "count") %>%
  mutate(percent = round(100 * count / sum(count), 1)) %>%
  arrange(desc(count))

#create table count of ALL locations from locations column NB. more than one location per dog
locations2 <- HSA_dogs_confirmed$location

# Step 1: Replace " and " with commas to standardize delimiters
locations_clean2 <- gsub(" and ", ", ", locations2, ignore.case = TRUE)

# Step 2: Remove unwanted strings (e.g., "suspect", "presum", "likely", directional terms)
# Use regex and ignore case for better coverage
locations_clean2 <- tolower(locations_clean2)  # convert early for consistent matching
locations_clean2 <- gsub("suspect[^,]*", "", locations_clean2)
locations_clean2 <- gsub("presum[^,]*", "", locations_clean2)
locations_clean2 <- gsub("likel[^,]*", "", locations_clean2)
locations_clean2 <- gsub("probable[^,]*", "", locations_clean2)
locations_clean2 <- gsub("\\bleft\\b", "", locations_clean2)
locations_clean2 <- gsub("\\bright\\b", "", locations_clean2)

# Optional: remove "mass on", "in the", etc. if those phrases occur
locations_clean2 <- gsub("mass on ", "", locations_clean2)
locations_clean2 <- gsub("in the ", "", locations_clean2)
locations_clean2 <- gsub("area of ", "", locations_clean2)
locations_clean2 <- gsub("\\s+", " ", locations_clean2)  # condense multiple spaces

# Step 3: Split into individual locations by comma
location_split2 <- strsplit(locations_clean2, ",")

# Step 4: Trim and flatten into a vector
all_locations2 <- trimws(unlist(location_split2))

# Step 5: Standardize common synonyms (e.g., "cardiac", "ventricle" → "heart")
all_locations2[all_locations2 %in% c("heart", "cardiac", "right atrial cavity", "ventricle", "right atrium","atrium","atrium/auricle","atrial","heart (atrium)","heart (ventricular wall)","heart (auricle)","subtle infiltration in section of atrium)","heart (mass","heart ( auricle)","ventricular free wall)","heart ( atrium)","heart ( auricle)", "near heart base","heart (atrium","heart (atrial tissue)" ,"heart (","heart base","heart; atrium","heart (interventricular septum just under the annulus of the posterior tricuspid valve leaflet)","myocardium/heart","myocardium","atrial cavity","atrio-ventricular groove", "atrium (","atrium of the heart","heart ( aricular)","heart; atrium","heart base","myocardium)","ventricle","heart ( auricle","ventricle)","auricle")] <- "heart"
all_locations2[all_locations2 %in% c("skin (lower lip", "skin", "skin (abdomen)", "scrotum","dermal","haired skin")] <- "skin"
all_locations2[all_locations2 %in% c("lungs", "lung", "thoracic cavity", "thorax","all lung lobes","lung.","lungs (","lungs (")] <- "lung"
all_locations2[all_locations2 %in% c("liver", "hepatic","liver (","liver: disseminated hemangiosarcoma")] <- "liver"
all_locations2[all_locations2 %in% c("spleen", "splenic","spleen (")] <- "spleen"
all_locations2[all_locations2 %in% c("abdominal cavity", "abdomen")] <- "abdomen"
all_locations2[all_locations2 %in% c("jejunum", "small intestine","intestinal serosa","intestines","small intestine (muscular wall)","small intestine (muscularis mucosa)")] <- "intestine"
all_locations2[all_locations2 %in% c("kidneys", "kidney","kidney (possible)","bilateral kidneys"  )] <- "kidney"
all_locations2[all_locations2 %in% c("retroperitoneal mass", "retroperitoneum ("  ,"retroperitoneal cavity","retroperitoneal","retroperitoneal space")] <- "retroperitoneum"
all_locations2[all_locations2 %in% c("cerebellum","cerebreum","cerebrum","cerebellum)","brain (cerebrum","brain (cerebrum)")] <- "brain"
all_locations2[all_locations2 %in% c("mesenteric adipose tissue","peripancreatic adipose","adipose tissue; cranial thigh","fibroadipose tissue:","fibroadipose tissue (","fat (mesentery?)","fat (tibial area)","adipose")] <- "adipose tissue"
all_locations2[all_locations2 %in% c("gastric serosa","stomach (muscularis mucosa)")] <- "stomach"

all_locations2[all_locations2 %in% c("to the urinary bladder caudal aspect of the abdomen","bladder","urinary bladder")] <- "bladder"
all_locations2[all_locations2 %in% c("tricept","skeletal muscle/adipose tissue ","epaxial muscle","muscle","skeletal muscle/adipose tissue","muscle ( flank)","skeletal muscle")] <- "muscle"
all_locations2[all_locations2 %in% c("adrenal")] <- "adrenal gland"
all_locations2[all_locations2 %in% c("rib"," bone (proximal humerus)","bone (proximal humerus)","carpus","15 vertebral body","l5 vertebral body","tibial)","bone (proximal humerus)")] <- "bone"
all_locations2[all_locations2 %in% c("muscle (diaphragm")] <- "diaphragm"
#NB skin techincally cutis
all_locations2[all_locations2 %in% c("subcutaneous","subcutaneous mass ( flank)","lateral mid thorax","subcutis")] <- "skin"
all_locations2[all_locations2 %in% c("adrenal glands")] <- "adrenal gland"
all_locations2[all_locations2 %in% c("mesentery/omentum","mesenetery")] <- "mesentery"
#combine all the weird misc ones into "other" and "unknown"
all_locations2[all_locations2 %in% c("abdomen)","abdominal fluid","abdominal mass","body wall","caudal abdominal mass","hindlimb amputation site","leg mass","mid abdomen -","mouth","mucosal lined tissue","side","shoulder mass", "shoulder/scapula","shoulder","front leg","thoaracic inlet","thoracic deep to scapula","chest","digit","abdomen")] <- "other"
all_locations2[all_locations2 %in% c("(","masses","possibly spleen or liver","previously found in spleen","unspecified sites: tisse
","spleen previously removed","unspecified sites: tisse")] <- "unknown"


# Step 6: Remove empty strings (resulting from deleted words)
all_locations2 <- all_locations2[all_locations2 != ""]



# Step 7: Create and print counts
location_counts2 <- sort(table(all_locations2), decreasing = TRUE)
print(location_counts2)

location_counts2 <- sort(table(all_locations2), decreasing = TRUE)


# Convert to data.frame
location_df2 <- as.data.frame(location_counts2)
colnames(location_df2) <- c("location", "count")

# Add percentage column
location_df2$percent <- round(100 * location_df2$count / sum(location_df2$count), 1)

# View result
print(location_df2)

# Combine count and percent into one column like "420 (80)"
location_df2$count_percent <- paste0(location_df2$count, " (", location_df2$percent, ")")

# Optional: keep just location and the combined column
location_summary2 <- location_df2[, c("location","count","count_percent", "count_percent")]

# View result
print(location_summary2)
#write to csv for manuscript
write.csv(location_summary2,"C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/GRLS_HSA_location_counts_confirmed_cases.csv")
```

Location barplot

```{r}
library(dplyr)
library(ggplot2)
library(forcats)

plot2<- location_df2 %>%
  ggplot(aes(x = fct_reorder(location, count), y = count)) +
  geom_col(fill = "grey30", colour = "black", width = 0.8) +
  geom_text(aes(label = count_percent),
            hjust = -0.15,
            size = 4) +
  coord_flip() +
  expand_limits(y = max(location_df2$count) * 1.15) +
  labs(
    title = "Distribution of Confirmed Haemangiosarcoma Locations",
    x = NULL,
    y = "Number of cases"
  ) +
  theme_classic(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

ggsave("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/GRLS_HSA_locations_barplot_confirmed.tiff", plot = plot2, width = 8, height = 6, dpi = 600)

```

## Haemangiosarcoma location MSTs

Survival time days calculation

```{r}
#funct to calc time between events
source("C:/Users/ctaylor18/GitHub/GRLS_analyses/Code/functions/age_calc_funct.R")

#convert char dates to date format
HSA_dogs_confirmed$diagnosis_date2 <- as.Date(HSA_dogs_confirmed$diagnosis_date)
HSA_dogs_confirmed$death_date2 <- as.Date(HSA_dogs_confirmed$death_date)


HSA_dogs3_conf <- age_calc(HSA_dogs_confirmed,"subject_id","diagnosis_date","death_date")

#across all these rows replace any negative values with 0
HSA_dogs3_conf <- HSA_dogs3_conf %>%
  mutate(across(c(time_between_diagnosis_date_death_date_days,time_between_diagnosis_date_death_date_weeks,time_between_diagnosis_date_death_date_years,time_between_diagnosis_date_death_date_months), ~ as.numeric(.))) %>%   # convert to numeric
  mutate(across(c(time_between_diagnosis_date_death_date_days,time_between_diagnosis_date_death_date_weeks,time_between_diagnosis_date_death_date_years,time_between_diagnosis_date_death_date_months), ~ ifelse(. < 0, 0, .))) # replace negatives with 0


#need to create a status variable for calc MST - status by 31/12/24. 0 = censored, 1=dead
#outcome where if do not have a death date = 0 and if do =1 
HSA_dogs3_conf <- HSA_dogs3_conf %>%
  mutate(outcome = if_else(time_between_diagnosis_date_death_date_days == "not recorded", 0, 1))



column_name <- "outcome"

HSA_dogs4_conf <- HSA_dogs3_conf %>%
  mutate(status = ifelse(is.na(.data[[column_name]]), 0, 1))

#ensure time between events are numeric
HSA_dogs4_conf <- HSA_dogs4_conf %>%
  mutate(across(c(time_between_diagnosis_date_death_date_days,time_between_diagnosis_date_death_date_weeks,time_between_diagnosis_date_death_date_years,time_between_diagnosis_date_death_date_months), as.numeric))

```

```{r}
splenic_conf<- HSA_dogs4_conf %>%
  filter(grepl("splen",tracked_condition))
cardiac_conf <- HSA_dogs4_conf %>%
  filter(grepl("card",tracked_condition))
visceral_conf<- HSA_dogs4_conf %>%
  filter(grepl("visc",tracked_condition))
cutaneous_conf<- HSA_dogs4 %>%
  filter(grepl("cut",tracked_condition))
other_conf<- HSA_dogs4_conf %>%
  filter(grepl("other",tracked_condition))

```

### All HSA

```{r}
survfit(Surv(HSA_dogs4_conf$time_between_diagnosis_date_death_date_days, HSA_dogs4_conf$status) ~ 1, data = HSA_dogs4_conf) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability of confirmed cases"
  ) +  add_confidence_interval() +
  add_risktable()

summary(HSA_dogs4_conf$time_between_diagnosis_date_death_date_days)

survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = HSA_dogs4_conf)
#probability of surviving to one year
summary(survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = HSA_dogs4_conf), times = 365.25)

#1 year survival rate
survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = HSA_dogs4_conf) %>% 
  tbl_survfit(
    times = 365.25,
    label_header = "**1-year survival (95% CI)**"
  )

quantile(HSA_dogs4_conf$time_between_diagnosis_date_death_date_days, probs = c(0.25, 0.75), na.rm = TRUE)

```

### Splenic

```{r}
survfit(Surv(splenic_conf$time_between_diagnosis_date_death_date_days, splenic_conf$status) ~ 1, data = splenic_conf) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) +  add_confidence_interval() +
  add_risktable()

summary(splenic_conf$time_between_diagnosis_date_death_date_days)
#median survival time is 5 days 

survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = splenic_conf)
#probability of surviving to one year
summary(survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = splenic_conf), times = 365.25)

#1 year survival rate
survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = splenic_conf) %>% 
  tbl_survfit(
    times = 365.25,
    label_header = "**1-year survival (95% CI)**"
  )

quantile(splenic_conf$time_between_diagnosis_date_death_date_days, probs = c(0.25, 0.75), na.rm = TRUE)
range(splenic_conf$time_between_diagnosis_date_death_date_days,na.rm = TRUE)
```

### Cardiac

```{r}
survfit(Surv(cardiac_conf$time_between_diagnosis_date_death_date_days, cardiac_conf$status) ~ 1, data = cardiac_conf) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) +  add_confidence_interval() +
  add_risktable()

summary(cardiac_conf$time_between_diagnosis_date_death_date_days)
#median survival time is 5 days 

survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = cardiac_conf)
#probability of surviving to one year
summary(survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = cardiac_conf), times = 365.25)

#1 year survival rate
survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = cardiac_conf) %>% 
  tbl_survfit(
    times = 365.25,
    label_header = "**1-year survival (95% CI)**"
  )

quantile(cardiac_conf$time_between_diagnosis_date_death_date_days, probs = c(0.25, 0.75), na.rm = TRUE)
range(cardiac_conf$time_between_diagnosis_date_death_date_days,na.rm = TRUE)
```

### Visceral

```{r}
survfit(Surv(visceral_conf$time_between_diagnosis_date_death_date_days, visceral_conf$status) ~ 1, data = visceral_conf) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) +  add_confidence_interval() +
  add_risktable()

summary(visceral_conf$time_between_diagnosis_date_death_date_days)

survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = visceral_conf)
#probability of surviving to one year
#NB needs to be set at 364 here as the single 365d case makes this not work as everyone has had the case by the 365d point
summary(survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = visceral_conf), times = 365.25)
 
#1 year survival rate - doesnt work as all cases dead by 1y
#survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = visceral_conf) %>% 
 # tbl_survfit(
 #   times = 365.25,
 #   label_header = "**1-year survival (95% CI)**"
 # )


nrow(visceral_conf)

summary(visceral_conf$time_between_diagnosis_date_death_date_days)

table(visceral_conf$status, useNA = "ifany")

#errors with above attempts to problem solve - not working
fit <- survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = visceral_conf)

# Extract components
fit_df <- data.frame(
  time = fit$time,
  surv = fit$surv,
  lower = fit$lower,
  upper = fit$upper
)

# Check the structure of the new data frame
str(fit_df)

report_time <- max(fit_df$time[fit_df$surv > 0])

tbl_survfit(
  fit,
  times = report_time,
  label_header = paste0("**Survival at ", report_time, " days (95% CI)**")
)




quantile(visceral_conf$time_between_diagnosis_date_death_date_days, probs = c(0.25, 0.75), na.rm = TRUE)
range(visceral_conf$time_between_diagnosis_date_death_date_days,na.rm = TRUE)
```

### Cutaneous

```{r}
survfit(Surv(cutaneous_conf$time_between_diagnosis_date_death_date_days, cutaneous_conf$status) ~ 1, data = cutaneous_conf) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) +  add_confidence_interval() +
  add_risktable()

summary(cutaneous_conf$time_between_diagnosis_date_death_date_days)
#median survival time is 5 days 

survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = cutaneous_conf)
#probability of surviving to one year
summary(survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = cutaneous_conf), times = 365.25)

#1 year survival rate
survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = cutaneous_conf) %>% 
  tbl_survfit(
    times = 365.25,
    label_header = "**1-year survival (95% CI)**"
  )

quantile(cutaneous_conf$time_between_diagnosis_date_death_date_days, probs = c(0.25, 0.75), na.rm = TRUE)
range(cutaneous_conf$time_between_diagnosis_date_death_date_days,na.rm = TRUE)
```

### Other

```{r}
survfit(Surv(other_conf$time_between_diagnosis_date_death_date_days, other_conf$status) ~ 1, data = other_conf) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) +  add_confidence_interval() +
  add_risktable()

summary(other_conf$time_between_diagnosis_date_death_date_days)
#median survival time is 5 days 

survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = other_conf)
#probability of surviving to one year
summary(survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = other_conf), times = 365.25)

#1 year survival rate
survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = other_conf) %>% 
  tbl_survfit(
    times = 365.25,
    label_header = "**1-year survival (95% CI)**"
  )
```

### Combined KM survival curve plot for all and all locations individual

```{r}
HSA_dogs4_conf <- HSA_dogs4_conf %>%
  mutate(cancer_group = recode(tracked_condition,
                               'hemangiosarcoma - visceral'="Visceral",
                               'hemangiosarcoma - splenic'= "Splenic",
                               'hemangiosarcoma - cardiac'="Cardiac",
                               'hemangiosarcoma - cutaneous'="Cutaneous",
                               'hemangiosarcoma - other/not specified'="Other")
  )#


#set levels for cancer_group
HSA_dogs4_conf <- HSA_dogs4_conf%>%
  mutate(cancer_group = factor(cancer_group, levels = c("Splenic","Cardiac"  ,"Visceral", "Cutaneous","Other")))

# Fit the survival model
fit <- survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ cancer_group, data = HSA_dogs4_conf)

# Plot the Kaplan-Meier curves
ggsurvplot(fit, data = HSA_dogs4_conf, palette = "Dark2", risk.table = TRUE)



library(RColorBrewer)

KM_plot_4 <- ggsurvplot(
  fit, 
  data = HSA_dogs4_conf,
  palette = brewer.pal(n = 5, name = "Set1"),  # Use a scientific palette
  conf.int = TRUE,  # Add confidence intervals
  conf.int.alpha = 0.1,  # Adjust transparency (0 = fully transparent, 1 = fully opaque)
  risk.table = FALSE,  # Dont show risk table
  ggtheme = theme_minimal(),  # Use a clean theme for manuscripts
  legend.title = "Cancer Type",
  xlab = "Time (Days)",
  ylab = "Survival Probability"
)

KM_plot_4



ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_confirmed.svg", plot = KM_plot_4$plot, width = 15, height = 10, dpi = 300)

ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_confirmed.pdf", plot = KM_plot_4$plot, width = 15, height = 10, dpi = 300)


ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_confirmed.tif", plot = KM_plot_4$plot, width = 15, height = 10, dpi = 300)



#remove "other category from plot as fairly uninformative
HSA_dogs5_conf <- HSA_dogs4_conf %>%
  filter(cancer_group != "Other")
fit <- survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ cancer_group, data = HSA_dogs5_conf)
library(RColorBrewer)

KM_plot_5 <- ggsurvplot(
  fit, 
  data = HSA_dogs5_conf,
  palette = brewer.pal(n = 5, name = "Set1"),  # Use a scientific palette
  conf.int = TRUE,  # Add confidence intervals
  conf.int.alpha = 0.1,  # Adjust transparency (0 = fully transparent, 1 = fully opaque)
  risk.table = FALSE,  # Dont show risk table
  ggtheme = theme_minimal(),  # Use a clean theme for manuscripts
  legend.title = "Cancer Type",
  xlab = "Time (Days)",
  ylab = "Survival Probability"
)

KM_plot_5



ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_no_other_conf.svg", plot = KM_plot_5$plot, width = 15, height = 10, dpi = 300)

ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_no_other_conf.pdf", plot = KM_plot_5$plot, width = 15, height = 10, dpi = 300)


ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_no_other_conf.tif", plot = KM_plot_5$plot, width = 15, height = 10, dpi = 300)

```

## Incidence per 100 dog years

Use profile dataset to calculate time between enrolled date and death_date OR 31st Dec 2024 if not dead yet

```{r}
source("C:/Users/ctaylor18/GitHub/GRLS_analyses/Code/functions/age_calc_funct.R")

#add death date 31/12/24 for all dogs with no death date recorded into death_date_2
profile <- profile %>%
  mutate(death_date_2 = if_else(death_date == "", "2024-12-31", death_date))
time_in_study <- age_calc(profile,"subject_id","enrolled_date","death_date_2")

#link profile data to HSA dog 4 conf


#as numeric
time_in_study$time_between_enrolled_date_death_date_2_years <-as.numeric(time_in_study$time_between_enrolled_date_death_date_2_years)

# Sum total dog-years
total_dog_years <- sum(time_in_study$time_between_enrolled_date_death_date_2_years, na.rm = TRUE)

# Number of dogs

cases <- length(unique(HSA_dogs4_conf$subject_id))
# Incidence rate per 100 dog-years
incidence_rate <- (cases / total_dog_years) * 100

incidence_rate

#95%CI 
# Calculate 95% CI using Poisson distribution
alpha <- 0.05

lower_limit <- 0.5 * qchisq(alpha / 2, 2 * cases)   # lower bound for cases
upper_limit <- 0.5 * qchisq(1 - alpha / 2, 2 * (cases + 1)) # upper bound for cases

lower_rate <- (lower_limit / total_dog_years) * 100
upper_rate <- (upper_limit / total_dog_years) * 100

# Print results
cat("Incidence rate of first haemangiosarcoma diagnosis (definitive) per 100 dog-years :", round(incidence_rate, 3), "\n")
cat("95% CI:", round(lower_rate, 3), "to", round(upper_rate, 3), "\n")



#incidence for multiple cases though:
# Number of dogs
cases <- length(HSA_dogs4_conf$subject_id)

# Incidence rate per 100 dog-years
incidence_rate <- (cases / total_dog_years) * 100

incidence_rate

#95%CI 
# Calculate 95% CI using Poisson distribution
alpha <- 0.05

lower_limit <- 0.5 * qchisq(alpha / 2, 2 * cases)   # lower bound for cases
upper_limit <- 0.5 * qchisq(1 - alpha / 2, 2 * (cases + 1)) # upper bound for cases

lower_rate <- (lower_limit / total_dog_years) * 100
upper_rate <- (upper_limit / total_dog_years) * 100

# Print results
cat("Incidence rate of haemangiosarcoma (definitive) per 100 dog-years:", round(incidence_rate, 3), "\n")
cat("95% CI:", round(lower_rate, 3), "to", round(upper_rate, 3), "\n")

```

### KM Survival curves of demographics

```{r}
#get df ready for KMs
cohort_df <- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_updated_analysis_df.csv")
#keep the age, sex, neut of only cases
cohort_df_cases <- cohort_df %>%
  dplyr::filter(had_HSA==1) %>%
  dplyr::select(c(subject_id,age_at_final_date,age_at_final_date_bio,sex,neuter,had_HSA))
#combine these cases to the HSA dogs 
cohort_df_cases_survival_times <- cohort_df_cases%>%
  left_join(HSA_dogs4,by="subject_id")

#confirmed cases only
cohort_df_cases_survival_times_confirmed <- cohort_df_cases_survival_times %>%
  dplyr::filter(tier_of_confidence==1)
```

#### Age

All cases:

```{r}
cohort_df_cases_survival_times <- cohort_df_cases_survival_times%>%
  mutate(age_at_final_date_bio = factor(age_at_final_date_bio, levels = c("<5y",">5-<8y",">8-<11y", ">11-<13y",">13y")))

# Fit the survival model
fit <- survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ age_at_final_date_bio, data = cohort_df_cases_survival_times)

# Plot the Kaplan-Meier curves
ggsurvplot(fit, data = cohort_df_cases_survival_times, palette = "Dark2", risk.table = TRUE)



library(RColorBrewer)

KM_plot_age <- ggsurvplot(
  fit, 
  data = cohort_df_cases_survival_times,
  palette = brewer.pal(n = 5, name = "Set1"),  # Use a scientific palette
  conf.int = TRUE,  # Add confidence intervals
  conf.int.alpha = 0.1,  # Adjust transparency (0 = fully transparent, 1 = fully opaque)
  risk.table = FALSE,  # Dont show risk table
  ggtheme = theme_minimal(),  # Use a clean theme for manuscripts
  legend.title = "Age at diagnosis groups",
  xlab = "Time (Days)",
  ylab = "Survival Probability"
)

KM_plot_age



ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_age_diagnosis.svg", plot = KM_plot_age$plot, width = 15, height = 10, dpi = 300)

ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_age_diagnosis.pdf", plot = KM_plot_age$plot, width = 15, height = 10, dpi = 300)


ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_age_diagnosis.tif", plot = KM_plot_age$plot, width = 15, height = 10, dpi = 300)

```

Tidier chatgpt sugestions

```{r}
library(dplyr)
library(survival)
library(survminer)
library(ggplot2)
library(grid)

survival_age_data <- cohort_df_cases_survival_times %>%
  mutate(
    age_at_final_date_bio = factor(
      age_at_final_date_bio,
      levels = c(
        "<5y",
        ">5-<8y",
        ">8-<11y",
        ">11-<13y",
        ">13y"
      ),
      labels = c(
        "<5 years",
        "5–<8 years",
        "8–<11 years",
        "11–<13 years",
        "≥13 years"
      )
    )
  ) %>%
  filter(
    !is.na(age_at_final_date_bio),
    !is.na(time_between_diagnosis_date_death_date_days),
    !is.na(status)
  ) %>%
  droplevels()

fit_age <- survfit(
  Surv(
    time_between_diagnosis_date_death_date_days,
    status
  ) ~ age_at_final_date_bio,
  data = survival_age_data
)

age_n <- table(
  survival_age_data$age_at_final_date_bio
)

age_legend_labels <- paste0(
  names(age_n),
  " (n=",
  as.integer(age_n),
  ")"
)
KM_plot_age <- ggsurvplot(
  fit_age,
  data = survival_age_data,

  conf.int = FALSE,

  # Censor marks
  censor = TRUE,
  censor.shape = 124,
  censor.size = 3,

  # Main manuscript time window
  xlim = c(0, 365),
  break.time.by = 90,

  # Risk table
  risk.table = TRUE,
  risk.table.height = 0.25,
  risk.table.title = "Number at risk",
  risk.table.y.text = TRUE,
  risk.table.y.text.col = FALSE,
  risk.table.col = "black",

  # Curves remain distinguishable in greyscale
  linetype = "strata",
  size = 0.9,

  palette = c(
    "#0072B2",
    "#E69F00",
    "#009E73",
    "#CC79A7",
    "#000000"
  ),

  xlab = "Days since HSA diagnosis",
  ylab = "Survival probability",

  legend.title = "Age at diagnosis",
  legend.labs = age_legend_labels,

  ggtheme = theme_classic(base_size = 12),
  tables.theme = theme_cleantable(base_size = 10)
)

KM_plot_age$plot <- KM_plot_age$plot +
  scale_y_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, by = 0.2),
    labels = scales::label_percent(accuracy = 1)
  ) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.key.width = unit(1.5, "cm"),
    legend.text = element_text(size = 9),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

KM_plot_age$table <- KM_plot_age$table +
  theme(
    axis.title.x = element_blank(),
    axis.text = element_text(size = 9)
  )

KM_plot_age
```

Confirmed cases only:

```{r}
cohort_df_cases_survival_times_confirmed <- cohort_df_cases_survival_times_confirmed%>%
  mutate(age_at_final_date_bio = factor(age_at_final_date_bio, levels = c("<5y",">5-<8y",">8-<11y", ">11-<13y",">13y")))
# Fit the survival model
fit <- survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ age_at_final_date_bio, data = cohort_df_cases_survival_times_confirmed)

# Plot the Kaplan-Meier curves
ggsurvplot(fit, data = cohort_df_cases_survival_times_confirmed, palette = "Dark2", risk.table = TRUE)



library(RColorBrewer)

KM_plot_age2 <- ggsurvplot(
  fit, 
  data = cohort_df_cases_survival_times_confirmed,
  palette = brewer.pal(n = 5, name = "Set1"),  # Use a scientific palette
  conf.int = TRUE,  # Add confidence intervals
  conf.int.alpha = 0.1,  # Adjust transparency (0 = fully transparent, 1 = fully opaque)
  risk.table = FALSE,  # Dont show risk table
  ggtheme = theme_minimal(),  # Use a clean theme for manuscripts
  legend.title = "Age at diagnosis groups (confirmed cases)",
  xlab = "Time (Days)",
  ylab = "Survival Probability"
)

KM_plot_age2



ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_age_diagnosis_conf_cases.svg", plot = KM_plot_age2$plot, width = 15, height = 10, dpi = 300)

ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_age_diagnosis_conf_cases.pdf", plot = KM_plot_age2$plot, width = 15, height = 10, dpi = 300)


ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_age_diagnosis_conf_cases.tif", plot = KM_plot_age2$plot, width = 15, height = 10, dpi = 300)


```

Sex

All cases:

```{r}

# Fit the survival model
fit <- survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ sex, data = cohort_df_cases_survival_times)

# Plot the Kaplan-Meier curves
ggsurvplot(fit, data = cohort_df_cases_survival_times, palette = "Dark2", risk.table = TRUE)



library(RColorBrewer)

KM_plot_sex <- ggsurvplot(
  fit, 
  data = cohort_df_cases_survival_times,
  palette = brewer.pal(n = 5, name = "Set1"),  # Use a scientific palette
  conf.int = TRUE,  # Add confidence intervals
  conf.int.alpha = 0.1,  # Adjust transparency (0 = fully transparent, 1 = fully opaque)
  risk.table = FALSE,  # Dont show risk table
  ggtheme = theme_minimal(),  # Use a clean theme for manuscripts
  legend.title = "Sex groups",
  xlab = "Time (Days)",
  ylab = "Survival Probability"
)

KM_plot_sex



ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_sex_diagnosis.svg", plot = KM_plot_sex$plot, width = 15, height = 10, dpi = 300)

ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_sex_diagnosis.pdf", plot = KM_plot_sex$plot, width = 15, height = 10, dpi = 300)


ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_sex_diagnosis.tif", plot = KM_plot_sex$plot, width = 15, height = 10, dpi = 300)

```

Confirmed cases:

```{r}

# Fit the survival model
fit <- survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ sex, data = cohort_df_cases_survival_times_confirmed)

# Plot the Kaplan-Meier curves
ggsurvplot(fit, data = cohort_df_cases_survival_times_confirmed, palette = "Dark2", risk.table = TRUE)



library(RColorBrewer)

KM_plot_sex2<- ggsurvplot(
  fit, 
  data = cohort_df_cases_survival_times_confirmed,
  palette = brewer.pal(n = 5, name = "Set1"),  # Use a scientific palette
  conf.int = TRUE,  # Add confidence intervals
  conf.int.alpha = 0.1,  # Adjust transparency (0 = fully transparent, 1 = fully opaque)
  risk.table = FALSE,  # Dont show risk table
  ggtheme = theme_minimal(),  # Use a clean theme for manuscripts
  legend.title = "Sex in confirmed cases",
  xlab = "Time (Days)",
  ylab = "Survival Probability"
)

KM_plot_sex2



ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_sex_diagnosis_conf_cases.svg", plot = KM_plot_sex2$plot, width = 15, height = 10, dpi = 300)

ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_sex_diagnosis_conf_cases.pdf", plot = KM_plot_sex2$plot, width = 15, height = 10, dpi = 300)


ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_sex_diagnosis_conf_cases.tif", plot = KM_plot_sex$plot, width = 15, height = 10, dpi = 300)

```

#### Neuter status

All cases:

```{r}

# Fit the survival model
fit <- survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ neuter, data = cohort_df_cases_survival_times)

# Plot the Kaplan-Meier curves
ggsurvplot(fit, data = cohort_df_cases_survival_times, palette = "Dark2", risk.table = TRUE)



library(RColorBrewer)

KM_plot_neut<- ggsurvplot(
  fit, 
  data = cohort_df_cases_survival_times,
  palette = brewer.pal(n = 5, name = "Set1"),  # Use a scientific palette
  conf.int = TRUE,  # Add confidence intervals
  conf.int.alpha = 0.1,  # Adjust transparency (0 = fully transparent, 1 = fully opaque)
  risk.table = FALSE,  # Dont show risk table
  ggtheme = theme_minimal(),  # Use a clean theme for manuscripts
  legend.title = "Neutering status groups",
  xlab = "Time (Days)",
  ylab = "Survival Probability"
)

KM_plot_neut



ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_neut_diagnosis.svg", plot = KM_plot_neut$plot, width = 15, height = 10, dpi = 300)

ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_neut_diagnosis.pdf", plot = KM_plot_neut$plot, width = 15, height = 10, dpi = 300)


ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_neut_diagnosis.tif", plot = KM_plot_neut$plot, width = 15, height = 10, dpi = 300)

```

Confirmed cases:

```{r}

# Fit the survival model
fit <- survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ neuter, data = cohort_df_cases_survival_times)

# Plot the Kaplan-Meier curves
ggsurvplot(fit, data = cohort_df_cases_survival_times, palette = "Dark2", risk.table = TRUE)



library(RColorBrewer)

KM_plot_neut2<- ggsurvplot(
  fit, 
  data = cohort_df_cases_survival_times,
  palette = brewer.pal(n = 5, name = "Set1"),  # Use a scientific palette
  conf.int = TRUE,  # Add confidence intervals
  conf.int.alpha = 0.1,  # Adjust transparency (0 = fully transparent, 1 = fully opaque)
  risk.table = FALSE,  # Dont show risk table
  ggtheme = theme_minimal(),  # Use a clean theme for manuscripts
  legend.title = "Age at diagnosis groups",
  xlab = "Time (Days)",
  ylab = "Survival Probability"
)

KM_plot_neut2



ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_neut_diagnosis_conf_cases.svg", plot = KM_plot_neut2$plot, width = 15, height = 10, dpi = 300)

ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_neut_diagnosis_conf_cases.pdf", plot = KM_plot_neut2$plot, width = 15, height = 10, dpi = 300)


ggsave(filename = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/km_hsa_grls_neut_diagnosis_conf_cases.tif", plot = KM_plot_neut2$plot, width = 15, height = 10, dpi = 300)

```

Create an image of the 3 demographics in a row with second col of confirmed

```{r}

```

### Plot of combined all cases and histo case locations of tumours reported

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)

# Combine all-case and histopath-confirmed location data
location_comparison_df <- bind_rows(
  location_df %>%
    transmute(
      location = as.character(location),
      count = as.integer(count),
      case_definition = "All HSA cases"
    ),

  location_df2 %>%
    transmute(
      location = as.character(location),
      count = as.integer(count),
      case_definition = "Histopathologically confirmed HSA"
    )
) %>%
  group_by(case_definition) %>%
  mutate(
    # Percentage of all recorded locations within each case definition
    percent = 100 * count / sum(count)
  ) %>%
  ungroup()


# Order locations according to frequency in the all-case analysis
location_order <- location_comparison_df %>%
  filter(case_definition == "All HSA cases") %>%
  arrange(percent) %>%
  pull(location)


# Add zero rows where a location is absent from one case definition
location_comparison_df <- location_comparison_df %>%
  complete(
    location,
    case_definition,
    fill = list(
      count = 0L,
      percent = 0
    )
  ) %>%
  mutate(
    case_definition = factor(
      case_definition,
      levels = c(
        "All HSA cases",
        "Histopathologically confirmed HSA"
      )
    ),

    location_label = str_to_sentence(location),

    location_label = factor(
      location_label,
      levels = str_to_sentence(location_order)
    ),

    # Labels displayed as count (percentage)
    bar_label = if_else(
      count == 0,
      "",
      sprintf("%d (%.1f%%)", count, percent)
    )
  )
```

```{r}
bar_dodge <- position_dodge(width = 0.78)

location_comparison_plot <- ggplot(
  location_comparison_df,
  aes(
    x = location_label,
    y = percent,
    fill = case_definition
  )
) +
  geom_col(
    position = bar_dodge,
    width = 0.70,
    colour = "black",
    linewidth = 0.25
  ) +
  geom_text(
    aes(label = bar_label),
    position = bar_dodge,
    hjust = -0.08,
    size = 3.0
  ) +
  coord_flip(clip = "off") +
  scale_fill_manual(
    values = c(
      "All HSA cases" = "#1b9e77",
      "Histopathologically confirmed HSA" = "#d95f02"
    ),
    name = NULL
  ) +
  scale_y_continuous(
    labels = function(x) paste0(x, "%"),
    expand = expansion(mult = c(0, 0.30))
  ) +
  labs(
    x = NULL,
    y = "Percentage of recorded tumour locations",
    caption = paste(
      "Dogs could contribute more than one tumour location.",
      "Labels show the number and percentage of recorded locations."
    )
  ) +
  theme_classic(base_size = 12) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.text = element_text(size = 11),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 12),
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    plot.caption = element_text(
      hjust = 0,
      size = 9,
      margin = margin(t = 8)
    ),
    plot.margin = margin(
      t = 5.5,
      r = 65,
      b = 5.5,
      l = 5.5
    )
  )

location_comparison_plot
```

```{r}
ggsave(
  "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/GRLS_HSA_locations_combined.tiff",
  plot = location_comparison_plot,
  width = 8.5,
  height = 10,
  units = "in",
  dpi = 600,
  compression = "lzw"
)

ggsave(
  "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/GRLS_HSA_locations_combined.pdf",
  plot = location_comparison_plot,
  width = 8.5,
  height = 10,
  units = "in"
)
```

Compare for main 4 locations whether sig difference in number of histo confirmed

```{r}
# ================================================================
# Compare the four main HSA locations between:
#   1. Histopathologically confirmed HSA cases
#   2. Non-histopathologically confirmed HSA cases
#
# Main locations:
#   spleen, heart, lung and liver
#
# Each dog contributes a maximum of once to each location.
#
# Pearson's chi-squared tests are used for each comparison.
# Odds ratios and Wald 95% confidence intervals are calculated
# from the corresponding 2 x 2 contingency tables.
#
# P-values are unadjusted as these are prespecified,
# exploratory/descriptive comparisons.
# ================================================================


library(dplyr)
library(tidyr)
library(stringr)
library(purrr)
library(tibble)


# ------------------------------------------------
# 1. Check required variables are present
# ------------------------------------------------

stopifnot(
  all(c("subject_id", "location") %in% names(HSA_dogs2)),
  "subject_id" %in% names(HSA_dogs_confirmed)
)


# ------------------------------------------------
# 2. Identify histopathologically confirmed dogs
# ------------------------------------------------

confirmed_ids <- HSA_dogs_confirmed %>%
  distinct(subject_id) %>%
  pull(subject_id)


# ------------------------------------------------
# 3. Create one row per HSA dog
#
# If a dog occurs more than once in HSA_dogs2, combine its
# location strings before splitting individual locations.
# ------------------------------------------------

hsa_case_status <- HSA_dogs2 %>%
  select(subject_id, location) %>%
  group_by(subject_id) %>%
  summarise(
    location_raw = paste(
      unique(location[!is.na(location) & location != ""]),
      collapse = ", "
    ),
    .groups = "drop"
  ) %>%
  mutate(
    histopath_confirmed = subject_id %in% confirmed_ids,

    confirmation_group = factor(
      if_else(
        histopath_confirmed,
        "Histopathologically confirmed",
        "Not histopathologically confirmed"
      ),
      levels = c(
        "Not histopathologically confirmed",
        "Histopathologically confirmed"
      )
    )
  )


# Check numbers in each confirmation group
hsa_case_status %>%
  count(confirmation_group)


# ------------------------------------------------
# 4. Clean and split location field
# ------------------------------------------------

location_segments <- hsa_case_status %>%
  select(subject_id, location_raw) %>%
  mutate(

    location_clean = str_to_lower(location_raw),

    # Standardise delimiters
    location_clean = str_replace_all(
      location_clean,
      regex(" and ", ignore_case = TRUE),
      ", "
    ),

    # Remove uncertain wording
    location_clean = str_remove_all(
      location_clean,
      "suspect[^,]*"
    ),

    location_clean = str_remove_all(
      location_clean,
      "presum[^,]*"
    ),

    location_clean = str_remove_all(
      location_clean,
      "likel[^,]*"
    ),

    location_clean = str_remove_all(
      location_clean,
      "probable[^,]*"
    ),

    # Remove directional terms
    location_clean = str_remove_all(
      location_clean,
      "\\bleft\\b|\\bright\\b"
    ),

    # Remove unnecessary wording
    location_clean = str_replace_all(
      location_clean,
      fixed("mass on "),
      ""
    ),

    location_clean = str_replace_all(
      location_clean,
      fixed("in the "),
      ""
    ),

    location_clean = str_replace_all(
      location_clean,
      fixed("area of "),
      ""
    ),

    location_clean = str_squish(location_clean)
  ) %>%

  # Split dogs with multiple recorded locations
  separate_rows(
    location_clean,
    sep = ","
  ) %>%

  mutate(
    location_clean = str_squish(location_clean),

    # Do not classify explicitly ambiguous spleen/liver wording
    location_clean = if_else(
      str_detect(
        location_clean,
        "possibly spleen or liver|spleen or liver"
      ),
      NA_character_,
      location_clean
    )
  ) %>%

  filter(
    !is.na(location_clean),
    location_clean != ""
  )


# ------------------------------------------------
# 5. Identify the four main tumour locations
#
# A dog may have more than one location.
# Each dog contributes only once to each location.
# ------------------------------------------------

dog_location_long <- location_segments %>%
  transmute(

    subject_id,

    spleen = str_detect(
      location_clean,
      "\\b(spleen|splenic)\\b"
    ),

    heart = str_detect(
      location_clean,
      paste0(
        "\\b(",
        "heart|cardiac|myocardium|myocardial|",
        "atrium|atrial|auricle|",
        "ventricle|ventricular|",
        "atrioventricular",
        ")\\b"
      )
    ),

    lung = str_detect(
      location_clean,
      "\\b(lung|lungs|thorax|thoracic)\\b"
    ),

    liver = str_detect(
      location_clean,
      "\\b(liver|hepatic)\\b"
    )
  ) %>%

  pivot_longer(
    cols = c(
      spleen,
      heart,
      lung,
      liver
    ),
    names_to = "location",
    values_to = "location_present"
  ) %>%

  filter(location_present) %>%

  # Prevent duplicate contribution of the same dog
  # to the same anatomical location
  distinct(
    subject_id,
    location
  )


# Check dog-level numbers with each location
dog_location_long %>%
  count(
    location,
    sort = TRUE
  )


# ------------------------------------------------
# 6. Define analysis population
#
# Restrict to dogs with some recorded tumour-location data.
#
# Dogs with a different recorded site (e.g. skin) remain in the
# denominator and count as not having spleen/heart/lung/liver
# involvement as appropriate.
# ------------------------------------------------

location_analysis_dogs <- hsa_case_status %>%
  filter(
    !is.na(location_raw),
    str_squish(location_raw) != ""
  )


# Check denominators
location_analysis_dogs %>%
  count(confirmation_group)


# ------------------------------------------------
# 7. Run Pearson chi-squared test for each of
#    the four major tumour locations
# ------------------------------------------------

main_locations <- c(
  "spleen",
  "heart",
  "lung",
  "liver"
)


main_location_tests <- map_dfr(
  main_locations,
  function(site) {

    # Dogs with this location
    dogs_with_site <- dog_location_long %>%
      filter(location == site) %>%
      distinct(subject_id) %>%
      pull(subject_id)


    # Add location present/absent to the whole analysis cohort
    site_data <- location_analysis_dogs %>%
      mutate(
        location_present =
          subject_id %in% dogs_with_site
      )


    # ------------------------------------------------
    # Construct 2 x 2 table
    #
    #                        Confirmed    Not confirmed
    #
    # Location present          a              b
    #
    # Location absent           c              d
    #
    # ------------------------------------------------

    a <- sum(
      site_data$histopath_confirmed &
        site_data$location_present
    )

    b <- sum(
      !site_data$histopath_confirmed &
        site_data$location_present
    )

    c <- sum(
      site_data$histopath_confirmed &
        !site_data$location_present
    )

    d <- sum(
      !site_data$histopath_confirmed &
        !site_data$location_present
    )


    location_table <- matrix(
      c(
        a, b,
        c, d
      ),
      nrow = 2,
      byrow = TRUE,
      dimnames = list(
        location_status = c(
          "Location present",
          "Location absent"
        ),
        confirmation_status = c(
          "Histopathologically confirmed",
          "Not histopathologically confirmed"
        )
      )
    )


    # ------------------------------------------------
    # Pearson chi-squared test
    #
    # correct = FALSE turns off Yates continuity correction
    # ------------------------------------------------

    chi_result <- chisq.test(
      location_table,
      correct = FALSE
    )


    # Store expected counts so we can check that
    # chi-squared assumptions are satisfied
    expected_counts <- chi_result$expected

    minimum_expected_count <- min(expected_counts)


    # ------------------------------------------------
    # Calculate ordinary 2 x 2 odds ratio
    #
    # OR = (a/b) / (c/d)
    #    = (a*d) / (b*c)
    #
    # OR > 1:
    # location proportion higher among confirmed cases
    #
    # OR < 1:
    # location proportion lower among confirmed cases
    # ------------------------------------------------

    OR <- (a * d) / (b * c)


    # ------------------------------------------------
    # Wald 95% CI for the odds ratio
    #
    # SE(log OR) =
    # sqrt(1/a + 1/b + 1/c + 1/d)
    # ------------------------------------------------

    SE_log_OR <- sqrt(
      (1 / a) +
      (1 / b) +
      (1 / c) +
      (1 / d)
    )


    CI_lower <- exp(
      log(OR) -
        1.96 * SE_log_OR
    )


    CI_upper <- exp(
      log(OR) +
        1.96 * SE_log_OR
    )


    # ------------------------------------------------
    # Store results
    # ------------------------------------------------

    tibble(

      location = site,

      confirmed_with_location = a,
      confirmed_without_location = c,
      confirmed_total = a + c,

      confirmed_percent =
        100 * a / (a + c),

      not_confirmed_with_location = b,
      not_confirmed_without_location = d,
      not_confirmed_total = b + d,

      not_confirmed_percent =
        100 * b / (b + d),

      # Confirmed minus non-confirmed percentage
      percentage_point_difference =
        confirmed_percent -
        not_confirmed_percent,

      OR_for_histopath_confirmation = OR,

      CI_lower = CI_lower,

      CI_upper = CI_upper,

      chi_square = unname(
        chi_result$statistic
      ),

      degrees_freedom = unname(
        chi_result$parameter
      ),

      chi_square_p_value =
        chi_result$p.value,

      # Useful check for chi-squared assumptions
      minimum_expected_count =
        minimum_expected_count
    )
  }
) %>%

  mutate(

    interpretation = case_when(

      chi_square_p_value < 0.05 &
        OR_for_histopath_confirmation > 1 ~
        "More frequently recorded among histopathologically confirmed cases",

      chi_square_p_value < 0.05 &
        OR_for_histopath_confirmation < 1 ~
        "Less frequently recorded among histopathologically confirmed cases",

      TRUE ~
        "No evidence of a difference"
    ),

    # Preserve preferred location order
    location = factor(
      location,
      levels = main_locations
    )
  ) %>%

  arrange(location)


# ------------------------------------------------
# 8. View complete numerical results
# ------------------------------------------------

main_location_tests


# ------------------------------------------------
# 9. Check chi-squared expected cell counts
#
# Ideally all should be >= 5
# ------------------------------------------------

main_location_tests %>%
  select(
    location,
    minimum_expected_count
  )


# Optional warning if any expected cell count is <5
if (
  any(
    main_location_tests$minimum_expected_count < 5
  )
) {

  warning(
    "At least one location has an expected cell count <5; consider Fisher's exact test for that comparison."
  )

}


# ------------------------------------------------
# 10. Function to format p-values
# ------------------------------------------------

format_p_value <- function(p) {

  case_when(

    is.na(p) ~
      NA_character_,

    p < 0.001 ~
      "<0.001",

    TRUE ~
      sprintf("%.3f", p)
  )
}


# ------------------------------------------------
# 11. Create manuscript-ready results table
# ------------------------------------------------

main_location_results_table <-
  main_location_tests %>%

  transmute(

    Location =
      str_to_sentence(
        as.character(location)
      ),

    `Histopathologically confirmed, n/N (%)` =
      sprintf(
        "%d/%d (%.1f%%)",
        confirmed_with_location,
        confirmed_total,
        confirmed_percent
      ),

    `Not histopathologically confirmed, n/N (%)` =
      sprintf(
        "%d/%d (%.1f%%)",
        not_confirmed_with_location,
        not_confirmed_total,
        not_confirmed_percent
      ),

    `Difference in percentage points` =
      sprintf(
        "%+.1f",
        percentage_point_difference
      ),

    `OR (95% CI)` =
      sprintf(
        "%.2f (%.2f–%.2f)",
        OR_for_histopath_confirmation,
        CI_lower,
        CI_upper
      ),

    `Chi-squared statistic` =
      sprintf(
        "%.2f",
        chi_square
      ),

    `Chi-squared p-value` =
      format_p_value(
        chi_square_p_value
      ),

    Interpretation =
      interpretation
  )


# View manuscript-ready table
main_location_results_table


# ------------------------------------------------
# 12. Save full numerical results
# ------------------------------------------------

write.csv(
  main_location_tests,
  "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_main_location_histopath_tests_chisq.csv",
  row.names = FALSE
)


# ------------------------------------------------
# 13. Save manuscript-ready formatted table
# ------------------------------------------------

write.csv(
  main_location_results_table,
  "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_main_location_histopath_results_table_chisq.csv",
  row.names = FALSE
)
```

## Survival time calculations for dogs with \>= 1 day of survival time ie. intention to treat proxy

```{r}
survival_time_1 <- HSA_dogs4 %>%
  dplyr::filter(time_between_diagnosis_date_death_date_days > 0)


survfit(Surv(survival_time_1$time_between_diagnosis_date_death_date_days, survival_time_1$status) ~ 1, data = survival_time_1) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) +  add_confidence_interval() +
  add_risktable()

summary(survival_time_1$time_between_diagnosis_date_death_date_days)

survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = survival_time_1)
#probability of surviving to one year
summary(survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = survival_time_1), times = 365.25)

#1 year survival rate
survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = survival_time_1) %>% 
  tbl_survfit(
    times = 365.25,
    label_header = "**1-year survival (95% CI)**"
  )

quantile(survival_time_1$time_between_diagnosis_date_death_date_days, probs = c(0.25, 0.75), na.rm = TRUE)
```

By locations:

```{r}
splenic<- survival_time_1 %>%
  filter(grepl("splen",tracked_condition))
cardiac <- survival_time_1 %>%
  filter(grepl("card",tracked_condition))
visceral<- survival_time_1 %>%
  filter(grepl("visc",tracked_condition))
cutaneous<- survival_time_1 %>%
  filter(grepl("cut",tracked_condition))
other<- survival_time_1 %>%
  filter(grepl("other",tracked_condition))

```


### Splenic

```{r}
survfit(Surv(splenic$time_between_diagnosis_date_death_date_days, splenic$status) ~ 1, data = splenic) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) +  add_confidence_interval() +
  add_risktable()

summary(splenic$time_between_diagnosis_date_death_date_days)
#median survival time is 5 days 

survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = splenic)
#probability of surviving to one year
summary(survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = splenic), times = 365.25)

#1 year survival rate
survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = splenic) %>% 
  tbl_survfit(
    times = 365.25,
    label_header = "**1-year survival (95% CI)**"
  )

quantile(splenic$time_between_diagnosis_date_death_date_days, probs = c(0.25, 0.75), na.rm = TRUE)
range(splenic$time_between_diagnosis_date_death_date_days,na.rm = TRUE)
```

### Cardiac

```{r}
survfit(Surv(cardiac$time_between_diagnosis_date_death_date_days, cardiac$status) ~ 1, data = cardiac) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) +  add_confidence_interval() +
  add_risktable()

summary(cardiac$time_between_diagnosis_date_death_date_days)
#median survival time is 5 days 

survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = cardiac)
#probability of surviving to one year
summary(survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = cardiac), times = 365.25)

#1 year survival rate
survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = cardiac) %>% 
  tbl_survfit(
    times = 365.25,
    label_header = "**1-year survival (95% CI)**"
  )

quantile(cardiac$time_between_diagnosis_date_death_date_days, probs = c(0.25, 0.75), na.rm = TRUE)
range(cardiac$time_between_diagnosis_date_death_date_days,na.rm = TRUE)
```

### Visceral

```{r}
survfit(Surv(visceral$time_between_diagnosis_date_death_date_days, visceral$status) ~ 1, data = visceral) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) +  add_confidence_interval() +
  add_risktable()

summary(visceral$time_between_diagnosis_date_death_date_days)

survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = visceral)
#probability of surviving to one year
#NB needs to be set at 364 here as the single 365d case makes this not work as everyone has had the case by the 365d point
summary(survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = visceral), times = 365.25)
 
#1 year survival rate - doesnt work as all cases dead by 1y
#survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = visceral) %>% 
 # tbl_survfit(
  #  times = 365.25,
  #  label_header = "**1-year survival (95% CI)**"
 # )


nrow(visceral)

summary(visceral$time_between_diagnosis_date_death_date_days)

table(visceral$status, useNA = "ifany")

#errors with above attempts to problem solve - not working
fit <- survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = visceral)

# Extract components
fit_df <- data.frame(
  time = fit$time,
  surv = fit$surv,
  lower = fit$lower,
  upper = fit$upper
)

# Check the structure of the new data frame
str(fit)

report_time <- max(fit_df$time[fit_df$surv > 0])

tbl_survfit(
  fit,
  times = report_time,
  label_header = paste0("**Survival at ", report_time, " days (95% CI)**")
)




quantile(visceral$time_between_diagnosis_date_death_date_days, probs = c(0.25, 0.75), na.rm = TRUE)
range(visceral$time_between_diagnosis_date_death_date_days,na.rm = TRUE)
```

### Cutaneous

```{r}
survfit(Surv(cutaneous$time_between_diagnosis_date_death_date_days, cutaneous$status) ~ 1, data = cutaneous) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) +  add_confidence_interval() +
  add_risktable()

summary(cutaneous$time_between_diagnosis_date_death_date_days)
#median survival time is 5 days 

survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = cutaneous)
#probability of surviving to one year
summary(survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = cutaneous), times = 365.25)

#1 year survival rate
survfit(Surv(time_between_diagnosis_date_death_date_days, status) ~ 1, data = cutaneous) %>% 
  tbl_survfit(
    times = 365.25,
    label_header = "**1-year survival (95% CI)**"
  )

quantile(cutaneous$time_between_diagnosis_date_death_date_days, probs = c(0.25, 0.75), na.rm = TRUE)
range(cutaneous$time_between_diagnosis_date_death_date_days,na.rm = TRUE)
```
