---
title: "GRLS medications usage exploration"
format: html
editor: visual
---

## GRLS medications usage exploration:

Want to explore non-use, intermediate and chronic use of medications and their association/or not with cancer development.

Need to :

-   use the active ingredient groups initially, then use the WHO ATC grouping?

-   calculate non-use, intermediate, chronic use of meds in the \>1y prior to diagnosis (NB. med usage is only done in annual vet questionaire not real-time so wont have exact date prescribed but just date of questionaire filling)

-   get in diagnosis data (currently use end points dataset) - could use study year

-   

Meds data

```{r}
meds <- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/GRLS_medications_initial_tidy.csv" )

#remove rows where there are just all NAs for drug info e.g first 2 rows of df

meds_2 <- meds %>%
  filter(medication_name != "NA" & medication_ingredients != "NA")
  #name, ingredient, dose coiumns all empty = remove
```

Endpoints:

-   NB might need to consider groupings e.g for cancer location e.g kidney cancer

-   \~1200 end point cancers out of 1700 end points overall

-   for each cancer, each case is a case of that cancer and the controls are those that didnt have that cancer?

```{r}
endpoints <-read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Data/study_endpoints.csv")

end_points <- endpoints %>% 
  mutate(across(where(is.character), tolower))
uniqueIDs_anydeath <- unique(end_points$subject_id)
# seperate out only neoplasia deaths with key words
neoplasia_at_death <- end_points %>% filter(grepl('tumor|oma|leuk|neopla',tracked_condition),ignore.case=TRUE)
uniqueIDs_neoplasia_at_death <- unique(neoplasia_at_death$subject_id)
```
