---
title: "GRLS HSA vs non-HSA logistic regression"
format: html
editor: visual
---

## HSA GRLS cohort

HSA cases vs non-case control logistic regression

```{r}
library(tidyverse)
library(tidyverse)
library(dplyr)
library(survival)
library(ggplot2)
library(tibble)
library(kableExtra)
library(epiDisplay)
library(lme4)
#devtools::install_github("zabore/ezfun")
ezfun::set_ccf_palette("contrast")
library(lubridate)
library(ggsurvfit)
library(gtsummary)
library(tidycmprsk)
library(gtable)
#devtools::install_github("zabore/condsurv")
library(condsurv)
library(survminer)
library(cardx)
library(finalfit)
library(crosstable)
library(kableExtra)
library(officer)
library(flextable)
cohort_abr <-read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_cohort_all_RFs_data_tidied_from_cox.csv")
```

Converting columns to factors for analyses

```{r}

#replace all missing with NA (so "" and NA)
cohort_abr <- cohort_abr %>%
  mutate(across(everything(), as.character)) %>%  # Ensure all columns are character
  mutate(across(everything(), ~na_if(.x, "n/a"))) %>%
  mutate(across(everything(), ~na_if(.x, "NA"))) %>%
  mutate(across(everything(), ~na_if(.x, "nr"))) %>%
  mutate(across(everything(), ~na_if(.x, "not recorded"))) %>%
  mutate(across(everything(), ~na_if(.x, ""))) %>%
  mutate(across(everything(), ~replace_na(.x, "not_recorded")))
cohort_abr$survival_time_months <- as.numeric(cohort_abr$survival_time_months)


cohort_abr$status_factor <- as.factor(as.character(cohort_abr$had_HSA))
cohort_abr$status <-as.numeric(cohort_abr$had_HSA) 
cohort_abr$survival_time_months <- as.numeric(cohort_abr$survival_time_months)


#change spayed while in heat to account for males and entire females
cohort_abr <- cohort_abr %>%
  mutate(spayed_while_in_heat = case_when(
    sex== "M" ~ "Male",
    sex_status == "Intact Female" ~ "Intact Female",
    spayed_while_in_heat == "0" ~ "0",
    spayed_while_in_heat == "1" ~"1"
  ))
#same as above account for unnetered and males 
cohort_abr$heats_before_neutering <- as.numeric(cohort_abr$heats_before_neutering)
heats <- cohort_abr %>%
# Convert to numeric
  filter(heats_before_neutering != 0) %>% 
# Remove rows where the column is 0
  filter(heats_before_neutering != "not_recorded")%>%
  dplyr::select(heats_before_neutering)# Keep only this column
heats <- as.numeric(heats$heats_before_neutering)

# Function to calculate quartiles or terciles dynamically
get_breaks <- function(x, probs) {
  breaks <- quantile(x, probs = probs, na.rm = TRUE)
  
  # If breaks are not unique, switch to terciles
  if (length(unique(breaks)) < length(probs)) {
    message("Quartiles not unique, switching to terciles.")
    breaks <- quantile(x, probs = c(0, 1/3, 2/3, 1), na.rm = TRUE)
  }
  
  return(breaks)
}

breaks <- get_breaks(heats, c(0, 0.25, 0.5, 0.75, 1))


# Assign labels dynamically
labels <- paste0("≥", round(head(breaks, -1), 2), " - <", round(tail(breaks, -1), 2))


cohort_abr <- cohort_abr %>%
  mutate(heats_before_neutering_terciles = case_when(
    sex== "M" ~ "Male",
    sex_status == "Intact Female" ~ "Intact Female",
    heats_before_neutering == "0" ~ "0",
    heats_before_neutering == "1" ~"1-<=2",
    heats_before_neutering == "2" ~"1-<=2",
    heats_before_neutering == "3" ~">2-<=4",
    heats_before_neutering == "4" ~">2-<=4",
    heats_before_neutering == "5" ~">4-12",
    heats_before_neutering == "6" ~">4-12",
    heats_before_neutering == "7" ~">4-12",
    heats_before_neutering == "8" ~">4-12",
    heats_before_neutering == "9" ~">4-12",
    heats_before_neutering == "10" ~">4-12",
    heats_before_neutering == "11" ~">4-12",
    heats_before_neutering == "12" ~">4-12"
  ))


#create combined sex neuter variable
cohort_abr <- cohort_abr %>%
  mutate(sex_neuter = paste(sex, neuter, sep = " "))



#ensure not_recorded is default not baseline
convert_and_relevel <- function(df) {
  df <- df %>%
    mutate(across(where(is.character), ~ factor(.))) %>%  # Convert all character columns to factors
    mutate(across(where(is.factor), ~ fct_relevel(., "not_recorded", after = Inf)))  # Move "not_recorded" to last place
  return(df)
}

# Apply to your dataset
cohort_abr <- convert_and_relevel(cohort_abr)


#swap not_recorded for NA and exclude from univar this way
cohort_no_not_rec <- cohort_abr %>%
  mutate(across(
    where(~ is.character(.) | is.factor(.)),
    ~ factor(na_if(as.character(.), "not_recorded"))
  ))



cohort_abr <- cohort_abr%>%
  mutate(heats_before_neutering_terciles = factor(heats_before_neutering_terciles, levels = c("0", "1-<=2", ">2-<=4", ">4-12","Intact Female","Male","not_recorded")))


#neuter reason levels
cohort_abr <- cohort_abr%>%
  mutate(neutered_reason = factor(neutered_reason, levels = c("Elective",  "Medical Reason", "Entire","Unknown")))


cohort_abr <- cohort_abr%>%
  mutate(mode_purina_BCS_5y_prior_endpoint = factor(mode_purina_BCS_5y_prior_endpoint, levels = c("5", "3", "4", "6","7","8","9","not_recorded")))

#change BCS to be over underweight
cohort_abr <- cohort_abr %>%
  mutate(mode_BCS_classify = case_when(
    mode_purina_BCS_5y_prior_endpoint == "1" ~"underweight",
    mode_purina_BCS_5y_prior_endpoint == "2" ~"underweight",
    mode_purina_BCS_5y_prior_endpoint == "3" ~"underweight",
    mode_purina_BCS_5y_prior_endpoint == "4" ~"ideal",
    mode_purina_BCS_5y_prior_endpoint == "5" ~"ideal",
    mode_purina_BCS_5y_prior_endpoint == "6" ~"overweight",
    mode_purina_BCS_5y_prior_endpoint == "7" ~"overweight",
    mode_purina_BCS_5y_prior_endpoint == "8" ~"overweight",
    mode_purina_BCS_5y_prior_endpoint == "9" ~"overweight",
  ))

#set age levels
cohort_abr <- cohort_abr%>%
  mutate(age_at_final_date_quartiles = factor(age_at_final_date_quartiles, levels = c("0.74-<8.47 years", ">=8.47-<10.68 years", ">=10.68-<11.64 years", ">=11.64-14.51 years")))

cohort_abr <- cohort_abr%>%
  mutate(age_at_final_date_bio = factor(age_at_final_date_bio, levels = c(">5-<8y","<5y",  ">8-<11y", ">11-<13y",">13y")))

cohort_abr <- cohort_abr%>%
  mutate(age_at_final_date_GB = factor(age_at_final_date_GB, levels = c(">7-<9y","<5y", ">5-<7y",  ">9-<11y",">11-<13y",">13y")))


#manually assign height, weight, BCS factor levles and age at neuter and age at diagnosis 
#height
cohort_abr <- cohort_abr%>%
  mutate(avg_height_5y_prior_endpoint_quartiles = factor(avg_height_5y_prior_endpoint_quartiles, levels = c("49.13-<56.07cm", ">=56.07-<58.42cm", ">=58.42-<60.32cm", ">=60.32cm-67.73cm","not_recorded")))
#weight
cohort_abr <- cohort_abr%>%
  mutate(avg_weight_5y_prior_endpoint_quartiles = factor(avg_weight_5y_prior_endpoint_quartiles, levels = c("9.75-<21.31kg", ">=21.31-<24.66kg", ">=24.66-<29.37kg", ">=29.37-41.52kg","not_recorded")))
#bcs
cohort_abr <- cohort_abr%>%
  mutate(avg_purina_BCS_5y_prior_endpoint_quartiles = factor(avg_purina_BCS_5y_prior_endpoint_quartiles, levels = c("3.3-<5", ">=5-<5.28", ">=5.28-<5.86", ">=5.86 - 9","not_recorded")))
#neut age

cohort_abr <- cohort_abr%>%
  mutate(age_at_neuter_years_quartiles = factor(age_at_neuter_years_quartiles, levels = c("0.15-<0.52y", ">=0.52-<0.83y", ">=0.83-2.44y", ">=2.44-9.59y","Entire","not_recorded")))
#diagnosis age
cohort_abr <- cohort_abr%>%
  mutate(age_at_final_date_quartiles = factor(age_at_final_date_quartiles, levels = c("0.74-<8.47 years", ">=8.47-<10.68 years", ">=10.68-<11.64 years", ">=11.64-14.51 years","not_recorded")))


```

## Univariable logstic regression

-   Counts

```{r}
dependent <- "had_HSA"

# All other variables as explanatory variables
explanatory = c(
  #demog
  "age_at_final_date_GB","age_at_final_date_bio", "age_at_final_date_quartiles","age_at_neuter_years_quartiles","sex_neuter",
   "sex", "neuter", "mismating_ever",
             "neutered_reason", "neutered_ever_bred", "spayed_while_in_heat", "spayed_while_pregnant", "prevent_heat_ever", 
             "no_pregnancy_ever", "intact_sired_litters_natural_ever", "neutered_females_bred_with_count_ever", 
             "neutered_sired_litters_natural_ever", "heats_before_neutering", "heats_before_neutering_terciles",
             "preg_pre_neuter",
 "num_breeding_before_neutering_terciles", "ever_bred_from", 
              "avg_weight_5y_prior_endpoint_quartiles", "avg_height_5y_prior_endpoint_quartiles", 
             "avg_purina_BCS_5y_prior_endpoint_quartiles", "mode_purina_BCS_5y_prior_endpoint", "mode_BCS_classify",
  #lifestyle/activity
  "main_lifestyle_category", 
 "main_lifestyle_category_mode_5yrs_prior", "USDA_30min_sy1_2", "activity_1h_min_sy1_2", 
             "activity_KC_2h_min_sy1_2", "avg_activity_duration_sy1_2_quartiles",
 "avg_activity_freq_sy1_2_terciles", 
             "avg_activity_freq_sy3_terciles", "avg_activity_intensity_sy3_quartiles", "avg_intensity_5y_prior_endpoint_quartiles", "avg_frequency_5y_prior_endpoint_quartiles",
  #home/environmental
  "MDI_quintile","owner_MDI_quintile","area_type_mode_5yrs_prior","house_type_mode_5yrs_prior", 
             "water_source_mode_5yrs_prior", "pipes_metal_any_mode_5yrs_prior", "pipes_plastic_any_mode_5yrs_prior", 
             "heating_fuel_primary_mode_5yrs_prior", "cooking_fuel_primary_mode_5yrs_prior", "heating_fuel_secondary_mode_5yrs_prior", 
             "cooking_fuel_secondary_mode_5yrs_prior", "region_mode_5y_prior","use_aerosol_study_years_early_life", "use_air_cleaner_study_years_early_life", "use_hepa_filter_study_years_early_life", 
             "use_moth_balls_study_years_early_life", "use_incense_or_candles_study_years_early_life", "smoke_exposure_study_years_early_life",  "use_aerosol_study_years_rest_of_life", 
             "use_air_cleaner_study_years_rest_of_life", "use_hepa_filter_study_years_rest_of_life", "use_moth_balls_study_years_rest_of_life", 
             "use_incense_or_candles_study_years_rest_of_life", "smoke_exposure_study_years_rest_of_life", "any_treated_weeds_study_years_rest_of_life", 
             "any_treated_insects_study_years_rest_of_life", "any_treated_fertilizer_study_years_rest_of_life","use_aerosol_study_years_whole_life", 
             "use_air_cleaner_study_years_whole_life", "use_hepa_filter_study_years_whole_life", "use_moth_balls_study_years_whole_life", 
             "use_incense_or_candles_study_years_whole_life", "smoke_exposure_study_years_whole_life", "any_treated_weeds_study_years_whole_life", 
             "any_treated_insects_study_years_whole_life", "any_treated_fertilizer_study_years_whole_life", "hours_of_smoke_early_life_total_dosage_terciles", "hours_of_smoke_rest_of_life_total_dosage_terciles", 
             "hours_of_smoke_whole_life_total_dosage_terciles", "in.the.garage_YN_study_years_early_life", "in.the.house_YN_study_years_early_life", 
             "outside_YN_study_years_early_life",  "in.the.house_YN_sum_early_life_terciles", 
             "in.the.garage_YN_study_years_rest_of_life", "in.the.house_YN_study_years_rest_of_life", "outside_YN_study_years_rest_of_life", 
              "in.the.house_YN_sum_rest_of_life_quartiles",  
             "in.the.garage_YN_study_years_whole_life", "in.the.house_YN_study_years_whole_life", "outside_YN_study_years_whole_life", 
             "in.the.house_YN_sum_whole_life_quartiles",  
             "majority_sleep_location_lifetime","sleep_location_mode_5yrs_prior", 
             "use_aerosol_exposed_5yrs_prior", "use_air_cleaner_exposed_5yrs_prior", "use_hepa_filter_exposed_5yrs_prior", 
             "use_moth_balls_exposed_5yrs_prior", "use_incense_or_candles_exposed_5yrs_prior", "smoke_exposure_exposed_5yrs_prior", 
             "any_treated_weeds_exposed_5yrs_prior", "any_treated_insects_exposed_5yrs_prior", "any_treated_fertilizer_exposed_5yrs_prior", 
             "hours_of_smoke_exposure_hours_5yrs_prior_terciles",
  #comorbidities
  "chronic_inflam_present_lifetime","number_comorbs_present_lifetime","chronic_inflam_present.x","chronic_inflam_present_5y","number_comorbs_present_5y","chronic_inflam_present_no_min_year","number_comorbs_present_no_min_year","intestinal_parasitism","tick_borne_parasitism","other_parasitism","all_parasitism","infectious","immune_mediated","cardiovascular","gastrointestinal","inflammatory_other","orthopaedic","chronic_inflammatory","lifelong_comorb_present.x","lifelong_orthopaedic","lifelong_immune_mediated","lifelong_cardiovascular","lifelong_gastrointestinal","lifelong_inflammatory_other","lifelong_chronic_inflammatory","number_lifelong_comorbs_present","malig_cancers_count","benign_cancers_count","all_cancers_count","all_cancers_YN","benign_cancers_YN","malig_cancers_YN","all_cancers_count_no_HSA","malig_cancers_count_no_HSA","malig_cancers_YN_no_HSA","all_cancers_YN_no_HSA",
  #medications
  "half_anti_microbials_lifetime", "half_dietary_supplements_lifetime", 
             "half_enteric_lifetime", "half_fluid_metabolites_lifetime", "half_hormones_and_related_lifetime", "half_immunologicals_lifetime", 
             "half_misc_lifetime", "half_neurological_lifetime", "half_anti_parasites_lifetime", "half_anti_septics_lifetime", 
             "half_cardio_respiratory_lifetime", "half_diuretics_lifetime", "half_anti_hist_lifetime", 
             "quartile_anti_inflam_lifetime", "quartile_anti_microbials_lifetime", "quartile_dietary_supplements_lifetime", 
 "quartile_hormones_and_related_lifetime", 
             "quartile_immunologicals_lifetime", "quartile_misc_lifetime", "quartile_neurological_lifetime", 
             "quartile_anti_parasites_lifetime", 
             "half_anti_inflam_5y_prior_endpoint", "half_anti_microbials_5y_prior_endpoint", "half_enteric_5y_prior_endpoint", 
             "half_immunologicals_5y_prior_endpoint", "half_misc_5y_prior_endpoint", "half_neurological_5y_prior_endpoint", 
             "half_anti_parasites_5y_prior_endpoint", "half_dietary_supplements_5y_prior_endpoint", 
             "half_anti_septics_5y_prior_endpoint", "half_hormones_and_related_5y_prior_endpoint", 
             "half_cardio_respiratory_5y_prior_endpoint", "half_fluid_metabolites_5y_prior_endpoint", 
             "half_diuretics_5y_prior_endpoint", "half_anti_hist_5y_prior_endpoint", 
             "quartile_anti_inflam_5y_prior_endpoint", "quartile_anti_microbials_5y_prior_endpoint",  "quartile_immunologicals_5y_prior_endpoint", 
             "quartile_misc_5y_prior_endpoint", "quartile_neurological_5y_prior_endpoint", 
             "quartile_anti_parasites_5y_prior_endpoint", "quartile_dietary_supplements_5y_prior_endpoint", 
 "quartile_hormones_and_related_5y_prior_endpoint", "NSAID_5y","NSAID_ever","steroid_5y","steroid_ever","fenbendazole_5y","fenbendazole_ever"
)

# Create a univariable logistic regression table
univar_table <- cohort_abr %>%summary_factorlist(dependent, explanatory, 
  p=TRUE,  add_dependent_label=TRUE,column = TRUE)
#column = TRUE does case and contorl counts for each variable category rather than rowise

univar_table2 <- cohort_abr %>%summary_factorlist(dependent, explanatory, 
  p=TRUE,  add_dependent_label=TRUE,na_include = TRUE,column=TRUE)
# View the results
print(univar_table)
write.csv(univar_table,"C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_cross_table_case_control_output.csv")

#save as worddoc
# Install and load the required packages
library(officer)
library(flextable)

# Create a Word document
doc <- read_docx()

# Convert finalfit output to a flextable
results_flextable <- flextable(univar_table)

# Add the flextable to the Word document
doc <- doc %>%
  body_add_flextable(results_flextable) %>%
  body_add_par("Finalfit Regression Results", style = "heading 1")

# Save the Word document
print(doc, target = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_cross_table_case_control_output.docx")
```

-   Univar log reg with ORs

```{r}
source("C:/Users/ctaylor18/GitHub/HSA-VC-enviro-analyses/functions/univar_log_reg_auto.R")

cohort_abr$had_HSA <- as.factor(cohort_abr$had_HSA)

#leaving out the continuous variables as missing data causing issues
#"max_tumour_cm_value" "age_at_diagnosis_cont"
cohort_abr2 <- cohort_abr %>%
  dplyr::select(#demog
  "age_at_final_date_GB","age_at_final_date_bio", "age_at_final_date_quartiles","age_at_neuter_years_quartiles","sex_neuter",
   "sex", "neuter", "mismating_ever",
             "neutered_reason", "neutered_ever_bred", "spayed_while_in_heat", "spayed_while_pregnant", "prevent_heat_ever", 
             "no_pregnancy_ever", "intact_sired_litters_natural_ever", "neutered_females_bred_with_count_ever", 
             "neutered_sired_litters_natural_ever", "heats_before_neutering", "heats_before_neutering_terciles",
             "preg_pre_neuter",
 "num_breeding_before_neutering_terciles", "ever_bred_from", 
              "avg_weight_5y_prior_endpoint_quartiles", "avg_height_5y_prior_endpoint_quartiles", 
             "avg_purina_BCS_5y_prior_endpoint_quartiles", "mode_purina_BCS_5y_prior_endpoint", "mode_BCS_classify",
  #lifestyle/activity
  "main_lifestyle_category", 
 "main_lifestyle_category_mode_5yrs_prior", "USDA_30min_sy1_2", "activity_1h_min_sy1_2", 
             "activity_KC_2h_min_sy1_2", "avg_activity_duration_sy1_2_quartiles",
 "avg_activity_freq_sy1_2_terciles", 
             "avg_activity_freq_sy3_terciles", "avg_activity_intensity_sy3_quartiles", "avg_intensity_5y_prior_endpoint_quartiles", "avg_frequency_5y_prior_endpoint_quartiles",
  #home/environmental
  "MDI_quintile","owner_MDI_quintile","area_type_mode_5yrs_prior","house_type_mode_5yrs_prior", 
             "water_source_mode_5yrs_prior", "pipes_metal_any_mode_5yrs_prior", "pipes_plastic_any_mode_5yrs_prior", 
             "heating_fuel_primary_mode_5yrs_prior", "cooking_fuel_primary_mode_5yrs_prior", "heating_fuel_secondary_mode_5yrs_prior", 
             "cooking_fuel_secondary_mode_5yrs_prior", "region_mode_5y_prior","use_aerosol_study_years_early_life", "use_air_cleaner_study_years_early_life", "use_hepa_filter_study_years_early_life", 
             "use_moth_balls_study_years_early_life", "use_incense_or_candles_study_years_early_life", "smoke_exposure_study_years_early_life",  "use_aerosol_study_years_rest_of_life", 
             "use_air_cleaner_study_years_rest_of_life", "use_hepa_filter_study_years_rest_of_life", "use_moth_balls_study_years_rest_of_life", 
             "use_incense_or_candles_study_years_rest_of_life", "smoke_exposure_study_years_rest_of_life", "any_treated_weeds_study_years_rest_of_life", 
             "any_treated_insects_study_years_rest_of_life", "any_treated_fertilizer_study_years_rest_of_life","use_aerosol_study_years_whole_life", 
             "use_air_cleaner_study_years_whole_life", "use_hepa_filter_study_years_whole_life", "use_moth_balls_study_years_whole_life", 
             "use_incense_or_candles_study_years_whole_life", "smoke_exposure_study_years_whole_life", "any_treated_weeds_study_years_whole_life", 
             "any_treated_insects_study_years_whole_life", "any_treated_fertilizer_study_years_whole_life", "hours_of_smoke_early_life_total_dosage_terciles", "hours_of_smoke_rest_of_life_total_dosage_terciles", 
             "hours_of_smoke_whole_life_total_dosage_terciles", "in.the.garage_YN_study_years_early_life", "in.the.house_YN_study_years_early_life", 
             "outside_YN_study_years_early_life",  "in.the.house_YN_sum_early_life_terciles", 
             "in.the.garage_YN_study_years_rest_of_life", "in.the.house_YN_study_years_rest_of_life", "outside_YN_study_years_rest_of_life", 
              "in.the.house_YN_sum_rest_of_life_quartiles",  
             "in.the.garage_YN_study_years_whole_life", "in.the.house_YN_study_years_whole_life", "outside_YN_study_years_whole_life", 
             "in.the.house_YN_sum_whole_life_quartiles",  
             "majority_sleep_location_lifetime","sleep_location_mode_5yrs_prior", 
             "use_aerosol_exposed_5yrs_prior", "use_air_cleaner_exposed_5yrs_prior", "use_hepa_filter_exposed_5yrs_prior", 
             "use_moth_balls_exposed_5yrs_prior", "use_incense_or_candles_exposed_5yrs_prior", "smoke_exposure_exposed_5yrs_prior", 
             "any_treated_weeds_exposed_5yrs_prior", "any_treated_insects_exposed_5yrs_prior", "any_treated_fertilizer_exposed_5yrs_prior", 
             "hours_of_smoke_exposure_hours_5yrs_prior_terciles",
  #comorbidities
  "chronic_inflam_present_lifetime","number_comorbs_present_lifetime","chronic_inflam_present.x","chronic_inflam_present_5y","number_comorbs_present_5y","chronic_inflam_present_no_min_year","number_comorbs_present_no_min_year","intestinal_parasitism","tick_borne_parasitism","other_parasitism","all_parasitism","infectious","immune_mediated","cardiovascular","gastrointestinal","inflammatory_other","orthopaedic","chronic_inflammatory","lifelong_comorb_present.x","lifelong_orthopaedic","lifelong_immune_mediated","lifelong_cardiovascular","lifelong_gastrointestinal","lifelong_inflammatory_other","lifelong_chronic_inflammatory","number_lifelong_comorbs_present","malig_cancers_count","benign_cancers_count","all_cancers_count","all_cancers_YN","benign_cancers_YN","malig_cancers_YN","all_cancers_count_no_HSA","malig_cancers_count_no_HSA","malig_cancers_YN_no_HSA","all_cancers_YN_no_HSA",
  #medications
  "half_anti_microbials_lifetime", "half_dietary_supplements_lifetime", 
             "half_enteric_lifetime", "half_fluid_metabolites_lifetime", "half_hormones_and_related_lifetime", "half_immunologicals_lifetime", 
             "half_misc_lifetime", "half_neurological_lifetime", "half_anti_parasites_lifetime", "half_anti_septics_lifetime", 
             "half_cardio_respiratory_lifetime", "half_diuretics_lifetime", "half_anti_hist_lifetime", 
             "quartile_anti_inflam_lifetime", "quartile_anti_microbials_lifetime", "quartile_dietary_supplements_lifetime", 
 "quartile_hormones_and_related_lifetime", 
             "quartile_immunologicals_lifetime", "quartile_misc_lifetime", "quartile_neurological_lifetime", 
             "quartile_anti_parasites_lifetime", 
             "half_anti_inflam_5y_prior_endpoint", "half_anti_microbials_5y_prior_endpoint", "half_enteric_5y_prior_endpoint", 
             "half_immunologicals_5y_prior_endpoint", "half_misc_5y_prior_endpoint", "half_neurological_5y_prior_endpoint", 
             "half_anti_parasites_5y_prior_endpoint", "half_dietary_supplements_5y_prior_endpoint", 
             "half_anti_septics_5y_prior_endpoint", "half_hormones_and_related_5y_prior_endpoint", 
             "half_cardio_respiratory_5y_prior_endpoint", "half_fluid_metabolites_5y_prior_endpoint", 
             "half_diuretics_5y_prior_endpoint", "half_anti_hist_5y_prior_endpoint", 
             "quartile_anti_inflam_5y_prior_endpoint", "quartile_anti_microbials_5y_prior_endpoint",  "quartile_immunologicals_5y_prior_endpoint", 
             "quartile_misc_5y_prior_endpoint", "quartile_neurological_5y_prior_endpoint", 
             "quartile_anti_parasites_5y_prior_endpoint", "quartile_dietary_supplements_5y_prior_endpoint", 
 "quartile_hormones_and_related_5y_prior_endpoint", "NSAID_5y","NSAID_ever","steroid_5y","steroid_ever","fenbendazole_5y","fenbendazole_ever","had_HSA")

cohort_abr2<- cohort_abr2%>%
  mutate_all(~factor(ifelse(is.na(.), "not_recorded", as.character(.))))



cohort_abr2 <- cohort_abr2%>%
  mutate(heats_before_neutering_terciles = factor(heats_before_neutering_terciles, levels = c("0", "1-<=2", ">2-<=4", ">4-12","Intact Female","Male","not_recorded")))


#neuter reason levels
cohort_abr2<- cohort_abr2%>%
  mutate(neutered_reason = factor(neutered_reason, levels = c("Elective",  "Medical Reason", "Entire","Unknown","not_recorded")))


cohort_abr2 <- cohort_abr2%>%
  mutate(mode_purina_BCS_5y_prior_endpoint = factor(mode_purina_BCS_5y_prior_endpoint, levels = c("5", "3", "4", "6","7","8","9","not_recorded")))

#change BCS to be over underweight
cohort_abr2 <- cohort_abr2 %>%
  mutate(mode_BCS_classify = case_when(
    mode_purina_BCS_5y_prior_endpoint == "1" ~"underweight",
    mode_purina_BCS_5y_prior_endpoint == "2" ~"underweight",
    mode_purina_BCS_5y_prior_endpoint == "3" ~"underweight",
    mode_purina_BCS_5y_prior_endpoint == "4" ~"ideal",
    mode_purina_BCS_5y_prior_endpoint == "5" ~"ideal",
    mode_purina_BCS_5y_prior_endpoint == "6" ~"overweight",
    mode_purina_BCS_5y_prior_endpoint == "7" ~"overweight",
    mode_purina_BCS_5y_prior_endpoint == "8" ~"overweight",
    mode_purina_BCS_5y_prior_endpoint == "9" ~"overweight",
    mode_purina_BCS_5y_prior_endpoint == "not_recorded" ~"not_recorded",
  ))

#set age levels
cohort_abr2 <- cohort_abr2%>%
  mutate(age_at_final_date_quartiles = factor(age_at_final_date_quartiles, levels = c("0.74-<8.47 years", ">=8.47-<10.68 years", ">=10.68-<11.64 years", ">=11.64-14.51 years")))

cohort_abr2 <- cohort_abr2%>%
  mutate(age_at_final_date_bio = factor(age_at_final_date_bio, levels = c(">5-<8y","<5y",  ">8-<11y", ">11-<13y",">13y")))

cohort_abr2 <- cohort_abr2%>%
  mutate(age_at_final_date_GB = factor(age_at_final_date_GB, levels = c(">7-<9y","<5y", ">5-<7y",  ">9-<11y",">11-<13y",">13y")))


#manually assign height, weight, BCS factor levles and age at neuter and age at diagnosis 
#height
cohort_abr2 <- cohort_abr2%>%
  mutate(avg_height_5y_prior_endpoint_quartiles = factor(avg_height_5y_prior_endpoint_quartiles, levels = c("49.13-<56.07cm", ">=56.07-<58.42cm", ">=58.42-<60.32cm", ">=60.32cm-67.73cm","not_recorded")))
#weight
cohort_abr2 <- cohort_abr2%>%
  mutate(avg_weight_5y_prior_endpoint_quartiles = factor(avg_weight_5y_prior_endpoint_quartiles, levels = c("9.75-<21.31kg", ">=21.31-<24.66kg", ">=24.66-<29.37kg", ">=29.37-41.52kg","not_recorded")))
#bcs
cohort_abr2 <- cohort_abr2%>%
  mutate(avg_purina_BCS_5y_prior_endpoint_quartiles = factor(avg_purina_BCS_5y_prior_endpoint_quartiles, levels = c("3.3-<5", ">=5-<5.28", ">=5.28-<5.86", ">=5.86 - 9","not_recorded")))
#neut age

cohort_abr2<- cohort_abr2 %>%
  mutate(age_at_neuter_years_quartiles = factor(age_at_neuter_years_quartiles, levels = c("0.15-<0.52y", ">=0.52-<0.83y", ">=0.83-2.44y", ">=2.44-9.59y","Entire","not_recorded")))
#diagnosis age
cohort_abr2 <- cohort_abr2%>%
  mutate(age_at_final_date_quartiles = factor(age_at_final_date_quartiles, levels = c("0.74-<8.47 years", ">=8.47-<10.68 years", ">=10.68-<11.64 years", ">=11.64-14.51 years","not_recorded")))




# Now run the function
univar_results <- univar_logistic_regression(cohort_abr2, "had_HSA", explanatory)

# If you want to write the output to CSV after running:
write.csv(univar_results, "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_log_reg_univar.csv", row.names = FALSE)



```

Tidy up csv of univar log reg into more readable format:
```{r}
#create tidy table from univar output
source("C:/Users/ctaylor18/GitHub/HSA-VC-enviro-analyses/functions/univar_log_reg_table.R")
#read in data
univar <- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_log_reg_univar.csv")

#function to tidy data and write to csv in a table ready to use for univar
table <- univar_logistic_regression_table(univar,"C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_log_reg_univar_tidied_table.csv")

#remove all the redunt variable name bits of table so just one row per variable 
blank_repeated_variable <- function(df, fill = ""){
  df$Variable <- ifelse(duplicated(df$Variable), fill, df$Variable)
  return(df)
}


table <- blank_repeated_variable(table, fill = " ") 

#crop table to keep relevant cols
table_2 <- table %>%
  dplyr::select(c(Variable,term,Case_summary,Control_summary,OR_CI,p.value,LR_T_pvalue,))
print(table_2)

print(knitr::kable(table_2, row.names=FALSE, 
    align=c("l", "l", "r", "r", "r")))

write.csv(table_2,"C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_log_reg_univar_tidied_table2.csv")

#save table_2 as a word doc file too
doc <- read_docx()

# Convert finalfit output to a flextable
results_flextable_2 <- flextable(table_2)

# Add the flextable to the Word document
doc <- doc %>%
  body_add_flextable(results_flextable_2) %>%
  body_add_par("Finalfit Regression Results", style = "heading 1")

# Save the Word document
print(doc, target = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_log_reg_univar_tidied_table2.docx")
```



Combine the final fit chi sq table with counts with the univar logistic regression table




Run univar log reg to only retain variables where they are sig not just due to NA/not recorded variable

```{r}

check_levels <- function(df) {
  sapply(df, function(x) length(unique(na.omit(x))) >= 2)
}

# Example usage
result <- check_levels(cohort_no_not_rec)

# Print columns that do not meet the condition
names(result[result == FALSE])




cohort_no_not_rec$had_HSA <- as.factor(cohort_no_not_rec$had_HSA)

#leaving out the continuous variables as missing data causing issues
#"max_tumour_cm_value" "age_at_diagnosis_cont"
cohort_no_not_rec<- cohort_no_not_rec %>%
  dplyr::select("MDI_quintile","owner_MDI_quintile","area_type_mode_5yrs_prior","house_type_mode_5yrs_prior", 
             "water_source_mode_5yrs_prior", "pipes_metal_any_mode_5yrs_prior", "pipes_plastic_any_mode_5yrs_prior", 
             "heating_fuel_primary_mode_5yrs_prior", "cooking_fuel_primary_mode_5yrs_prior", "heating_fuel_secondary_mode_5yrs_prior", 
             "cooking_fuel_secondary_mode_5yrs_prior", "region_mode_5y_prior"
            ,"chronic_inflam_present_lifetime","number_comorbs_present_lifetime","chronic_inflam_present.x","chronic_inflam_present_5y","number_comorbs_present_5y","chronic_inflam_present_no_min_year","number_comorbs_present_no_min_year","intestinal_parasitism","tick_borne_parasitism","other_parasitism","all_parasitism","infectious","immune_mediated","cardiovascular","gastrointestinal","inflammatory_other","orthopaedic","chronic_inflammatory","lifelong_comorb_present.x","lifelong_orthopaedic","lifelong_immune_mediated","lifelong_cardiovascular","lifelong_gastrointestinal","lifelong_inflammatory_other","lifelong_chronic_inflammatory","number_lifelong_comorbs_present",
            "USDA_30min_sy1_2", "activity_1h_min_sy1_2", 
             "activity_KC_2h_min_sy1_2", "avg_activity_duration_sy1_2_quartiles",#"avg_activity_intensity_sy1_2_terciles",
 "avg_activity_freq_sy1_2_terciles", 
             "avg_activity_freq_sy3_terciles", "avg_activity_intensity_sy3_quartiles", "avg_intensity_5y_prior_endpoint_quartiles", "avg_frequency_5y_prior_endpoint_quartiles", 
             "use_aerosol_study_years_early_life", "use_air_cleaner_study_years_early_life", "use_hepa_filter_study_years_early_life", 
             "use_moth_balls_study_years_early_life", "use_incense_or_candles_study_years_early_life", "smoke_exposure_study_years_early_life", 
       #      "any_treated_weeds_study_years_early_life", "any_treated_insects_study_years_early_life", #"any_treated_fertilizer_study_years_early_life", 
         #    "use_aerosol_sum_early_life", "use_air_cleaner_sum_early_life", "use_hepa_filter_sum_early_life", "use_moth_balls_sum_early_life", 
          #   "use_incense_or_candles_sum_early_life", 
 #"smoke_exposure_sum_early_life",
 #"any_treated_weeds_sum_early_life", 
    #         "any_treated_insects_sum_early_life", #"any_treated_fertilizer_sum_early_life", 
 "use_aerosol_study_years_rest_of_life", 
             "use_air_cleaner_study_years_rest_of_life", "use_hepa_filter_study_years_rest_of_life", "use_moth_balls_study_years_rest_of_life", 
             "use_incense_or_candles_study_years_rest_of_life", "smoke_exposure_study_years_rest_of_life", "any_treated_weeds_study_years_rest_of_life", 
             "any_treated_insects_study_years_rest_of_life", "any_treated_fertilizer_study_years_rest_of_life", #"use_aerosol_sum_rest_of_life", 
 #            "use_air_cleaner_sum_rest_of_life", "use_hepa_filter_sum_rest_of_life", "use_moth_balls_sum_rest_of_life", 
          #   "use_incense_or_candles_sum_rest_of_life", 
# "smoke_exposure_sum_rest_of_life", 
# "any_treated_weeds_sum_rest_of_life", 
            # "any_treated_insects_sum_rest_of_life", "any_treated_fertilizer_sum_rest_of_life", 
"use_aerosol_study_years_whole_life", 
             "use_air_cleaner_study_years_whole_life", "use_hepa_filter_study_years_whole_life", "use_moth_balls_study_years_whole_life", 
             "use_incense_or_candles_study_years_whole_life", "smoke_exposure_study_years_whole_life", "any_treated_weeds_study_years_whole_life", 
             "any_treated_insects_study_years_whole_life", "any_treated_fertilizer_study_years_whole_life",# "use_aerosol_sum_whole_life", 
           #  "use_air_cleaner_sum_whole_life", "use_hepa_filter_sum_whole_life", "use_moth_balls_sum_whole_life",# "use_incense_or_candles_sum_whole_life", 
            # "smoke_exposure_sum_whole_life", 
 #"any_treated_weeds_sum_whole_life", "any_treated_insects_sum_whole_life", 
            # "any_treated_fertilizer_sum_whole_life",
 "hours_of_smoke_early_life_total_dosage_terciles", "hours_of_smoke_rest_of_life_total_dosage_terciles", 
             "hours_of_smoke_whole_life_total_dosage_terciles", "in.the.garage_YN_study_years_early_life", "in.the.house_YN_study_years_early_life", 
             "outside_YN_study_years_early_life",  "in.the.house_YN_sum_early_life_terciles", 
             "in.the.garage_YN_study_years_rest_of_life", "in.the.house_YN_study_years_rest_of_life", "outside_YN_study_years_rest_of_life", 
              "in.the.house_YN_sum_rest_of_life_quartiles",  
             "in.the.garage_YN_study_years_whole_life", "in.the.house_YN_study_years_whole_life", "outside_YN_study_years_whole_life", 
             "in.the.house_YN_sum_whole_life_quartiles",  
             "majority_sleep_location_lifetime",  "main_lifestyle_category", 
 "main_lifestyle_category_mode_5yrs_prior", "sleep_location_mode_5yrs_prior", 
             "use_aerosol_exposed_5yrs_prior", "use_air_cleaner_exposed_5yrs_prior", "use_hepa_filter_exposed_5yrs_prior", 
             "use_moth_balls_exposed_5yrs_prior", "use_incense_or_candles_exposed_5yrs_prior", "smoke_exposure_exposed_5yrs_prior", 
             "any_treated_weeds_exposed_5yrs_prior", "any_treated_insects_exposed_5yrs_prior", "any_treated_fertilizer_exposed_5yrs_prior", 
             "hours_of_smoke_exposure_hours_5yrs_prior_terciles", "sex", "neuter", "mismating_ever",
             "neutered_reason", "neutered_ever_bred", "spayed_while_in_heat", "spayed_while_pregnant", "prevent_heat_ever", 
             "no_pregnancy_ever", "intact_sired_litters_natural_ever", "neutered_females_bred_with_count_ever", 
             "neutered_sired_litters_natural_ever", "heats_before_neutering", "heats_before_neutering_terciles",
             "preg_pre_neuter",# "mated_pre_neuter",
 "num_breeding_before_neutering_terciles", "ever_bred_from", 
              "avg_weight_5y_prior_endpoint_quartiles", "avg_height_5y_prior_endpoint_quartiles", 
             "avg_purina_BCS_5y_prior_endpoint_quartiles", "mode_purina_BCS_5y_prior_endpoint", "mode_BCS_classify", "half_anti_microbials_lifetime", "half_dietary_supplements_lifetime", 
             "half_enteric_lifetime", "half_fluid_metabolites_lifetime", "half_hormones_and_related_lifetime", "half_immunologicals_lifetime", 
             "half_misc_lifetime", "half_neurological_lifetime", "half_anti_parasites_lifetime", "half_anti_septics_lifetime", 
             "half_cardio_respiratory_lifetime", "half_diuretics_lifetime", "half_anti_hist_lifetime", 
             "quartile_anti_inflam_lifetime", "quartile_anti_microbials_lifetime", "quartile_dietary_supplements_lifetime", 
            #not enough data for quartiles "quartile_enteric_lifetime", "quartile_fluid_metabolites_lifetime", 
 "quartile_hormones_and_related_lifetime", 
             "quartile_immunologicals_lifetime", "quartile_misc_lifetime", "quartile_neurological_lifetime", 
             "quartile_anti_parasites_lifetime", 
 #not enough data for quartiles "quartile_anti_septics_lifetime", "quartile_cardio_respiratory_lifetime",   "quartile_diuretics_lifetime", "quartile_anti_hist_lifetime", 
             "half_anti_inflam_5y_prior_endpoint", "half_anti_microbials_5y_prior_endpoint", "half_enteric_5y_prior_endpoint", 
             "half_immunologicals_5y_prior_endpoint", "half_misc_5y_prior_endpoint", "half_neurological_5y_prior_endpoint", 
             "half_anti_parasites_5y_prior_endpoint", "half_dietary_supplements_5y_prior_endpoint", 
             "half_anti_septics_5y_prior_endpoint", "half_hormones_and_related_5y_prior_endpoint", 
             "half_cardio_respiratory_5y_prior_endpoint", "half_fluid_metabolites_5y_prior_endpoint", 
             "half_diuretics_5y_prior_endpoint", "half_anti_hist_5y_prior_endpoint", 
             "quartile_anti_inflam_5y_prior_endpoint", "quartile_anti_microbials_5y_prior_endpoint", 
            # "quartile_enteric_5y_prior_endpoint",
 "quartile_immunologicals_5y_prior_endpoint", 
             "quartile_misc_5y_prior_endpoint", "quartile_neurological_5y_prior_endpoint", 
             "quartile_anti_parasites_5y_prior_endpoint", "quartile_dietary_supplements_5y_prior_endpoint", 
       #      "quartile_anti_septics_5y_prior_endpoint", 
 "quartile_hormones_and_related_5y_prior_endpoint", 
          #   "quartile_cardio_respiratory_5y_prior_endpoint", #"quartile_fluid_metabolites_5y_prior_endpoint", 
         #    "quartile_diuretics_5y_prior_endpoint", "quartile_anti_hist_5y_prior_endpoint", 
             #"quartile_anti_viral_5y_prior_endpoint",
 "age_at_final_date_quartiles","age_at_neuter_years_quartiles",
"malig_cancers_count","benign_cancers_count","all_cancers_count","all_cancers_YN","benign_cancers_YN","malig_cancers_YN","all_cancers_count_no_HSA","malig_cancers_count_no_HSA","malig_cancers_YN_no_HSA","all_cancers_YN_no_HSA","NSAID_5y","NSAID_ever","steroid_5y","steroid_ever","fenbendazole_5y","fenbendazole_ever","had_HSA","age_at_final_date_bio","age_at_final_date_GB","sex_neuter")


univar_logistic_regression(cohort_no_not_rec, "had_HSA", "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_log_reg_univar_no_missing.csv")

univar_table_no_missing <- cohort_no_not_rec %>%summary_factorlist(dependent, explanatory, 
  add_dependent_label=TRUE,na_include = FALSE, p=TRUE,column=TRUE)
# View the results
print(univar_table_no_missing)

write.csv(univar_table_no_missing,"C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_log_reg_univar_no_missing.csv")

```

Potentially significant univariable variables (with NA included as "not_recorded" level)

```{r}
lrt_pvalues <-read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_log_reg_univar.csv")
multivar_variables <- lrt_pvalues %>%
  filter(LR_T_pvalue <= 0.2) %>%
  distinct(predictor)
cols_to_keep = multivar_variables$predictor
cols_to_keep <- append(cols_to_keep, "had_HSA")
cohort_abr_sig <- cohort_abr2[,cols_to_keep]

#missing data removed for p val calculation
pvalues_no_miss <-read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_log_reg_univar_no_missing.csv")
multivar_variables2 <- pvalues_no_miss %>%
  filter(p <= 0.2) %>%
  distinct(Dependent..had_HSA)
cols_to_keep2 = multivar_variables2$Dependent..had_HSA
cols_to_keep2 <- append(cols_to_keep2, "had_HSA")
cols_to_keep2 <- cols_to_keep2[cols_to_keep2 != ""]

#keeping the not recorded variables
cohort_abr_sig2 <- cohort_no_not_rec[,cols_to_keep2]



#keep only these ones for the original cohort with "not_recorded" level as only want to take these ones that are actually univar sig forward to multivar model build

cohort_abr_sig <- cohort_abr2[,cols_to_keep2]

```

## Multivariable logistic regression

Assessing collinearity of potentially sig univariable variables:

```{r}
source("C:/Users/ctaylor18/GitHub/HSA-VC-enviro-analyses/functions/corr_matrix_creation.R")
#use sig variables list generated above

create_corr_matrix(cohort_abr_sig,c(1:100), "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_logreg.csv", "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_logreg2.csv","C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_logreg_corrvalue.csv","C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_logreg.png")
```

Correlated variables

```{r}
corr_values <- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_logreg_corrvalue.csv")

#ditch rows where variable 1 and 2 values are same as in another row :
corr_values2 <- corr_values %>%
  dplyr::select(-c(X))

corr_values3 <- corr_values2 %>%
  mutate(Var1_sorted = pmin(Variable1, Variable2),  
         Var2_sorted = pmax(Variable1, Variable2)) %>%  
  distinct(Var1_sorted, Var2_sorted, Correlation, .keep_all = TRUE) %>%  
  dplyr::select(-c(Var1_sorted, Var2_sorted))  # Remove helper columns


print(corr_values3)
```

```{r}
source("C:/Users/ctaylor18/GitHub/HSA-VC-enviro-analyses/functions/corr_matrix_creation.R")
#use sig variables list generated above

create_corr_matrix(cohort_abr_sig2,c(1:100), "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_logreg_no_missing.csv", "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_logreg2_no_missing.csv","C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_logreg_corrvalue_no_missing.csv","C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_logreg_no_missing.png")





corr_values <- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_logreg_corrvalue_no_missing.csv")

#ditch rows where variable 1 and 2 values are same as in another row :
corr_values2 <- corr_values %>%
  dplyr::select(-c(X))

corr_values3 <- corr_values2 %>%
  mutate(Var1_sorted = pmin(Variable1, Variable2),  
         Var2_sorted = pmax(Variable1, Variable2)) %>%  
  distinct(Var1_sorted, Var2_sorted, Correlation, .keep_all = TRUE) %>%  
  dplyr::select(-c(Var1_sorted, Var2_sorted))  # Remove helper columns


print(corr_values3)
```

Correlated variables (98):

-   Comorbidities

    -   5y vs lifetime

    -   comorb inflam vs chronic inflam

    -   time between comorb and endpoint and diagnosis after comorb

    -   Malig vs all cancer counts and YN variables

-   Activity

    -   activity freq and intensity for SY3+

-   Enviro

    -   hours of smoke whole life and smoke exposure YN early life

    -   for all the exposure YN variables corr between early vs whole vs rest vs 5y prior

    -   fertilizer and weeds 5y prior

-   Repro

    -   sex and spayed in heat, no preg ever, heats before neutering,age at neuter quartiles (and amongst all these variables)

    -   

-   Medications

    -   half vs quartile and lifetime vs 5y for most are correlated

    -   fluid metabolites and diuretic and antihistamines

    -   5y and ever variables for fenbendazole, NSAID and steroids

#### Automated log reg model building - does nto work

```{r}
sig_variables <- unique(multivar_variables$predictor)

# Construct formula for full model
full_formula <- as.formula(paste("had_HSA ~", paste(sig_variables, collapse = " + ")))

# Build the null model
model_null <- glm(had_HSA ~ 1, data = cohort_abr_sig, family = binomial)

# Build the full model
model_full <- glm(full_formula, data = cohort_abr_sig, family = binomial)

# Display model summary
summary(model_full)
#check VIF of full model - doesnt run as aliased coeffs
#vif_values <- car::vif(model_full)

# Stepwise selection using both directions and LRT test
step(model_null,
     scope = list(upper = model_full),
     direction = "both",
     test = "LRT",
     data = df_for_multivar)
```

Automated model final:

```{r,eval=FALSE,error=FALSE}
model_auto <- glm(formula = had_HSA ~ age_at_final_date_quartiles + in.the.house_YN_study_years_rest_of_life + 
    avg_frequency_5y_prior_endpoint_quartiles + mode_purina_BCS_5y_prior_endpoint + 
    avg_height_5y_prior_endpoint_quartiles + cooking_fuel_secondary_mode_5yrs_prior + 
    in.the.house_YN_sum_whole_life_quartiles + quartile_anti_parasites_5y_prior_endpoint + 
    quartile_anti_parasites_lifetime + avg_weight_5y_prior_endpoint_quartiles + 
    no_pregnancy_ever + comorb_immune_mediated_chronic + number_comorbs_present_lifetime + 
    comorb_inflammatory_other + in.the.house_YN_sum_rest_of_life_quartiles + 
    mismating_ever + smoke_exposure_study_years_early_life + 
    main_lifestyle_category + outside_YN_study_years_early_life, 
    family = binomial, data = cohort_abr_sig)
summary(model_auto)
#car::vif(model_auto)

dependent ="had_HSA"
explanatory2 <- c(
  "age_at_final_date_quartiles", 
  "in.the.house_YN_study_years_rest_of_life", 
  "avg_frequency_5y_prior_endpoint_quartiles", 
  "mode_purina_BCS_5y_prior_endpoint", 
  "avg_height_5y_prior_endpoint_quartiles", 
  "cooking_fuel_secondary_mode_5yrs_prior", 
  "in.the.house_YN_sum_whole_life_quartiles", 
  "quartile_anti_parasites_5y_prior_endpoint", 
  "quartile_anti_parasites_lifetime", 
  "avg_weight_5y_prior_endpoint_quartiles", 
  "no_pregnancy_ever", 
  "comorb_immune_mediated_chronic", 
  "number_comorbs_present_lifetime", 
  "comorb_inflammatory_other", 
  "in.the.house_YN_sum_rest_of_life_quartiles", 
  "mismating_ever", 
  "smoke_exposure_study_years_early_life", 
  "main_lifestyle_category", 
  "outside_YN_study_years_early_life"
)

# only presenting factors not management factors - explanatory_2 <- c("haem_cs_present","mass_cs_present","no_cs_present","splenic_interest")
cohort_abr_sig %>%
  finalfit(dependent, explanatory2, metrics=TRUE) -> t2
knitr::kable(t2[[1]], row.names=FALSE, align=c("l", "l", "r", "r", "r", "r"))
knitr::kable(t2[[2]], row.names=FALSE, col.names="")
```

### Manual multivariable model building

from most significant to least (p\<0.2) :

```{r}
null_model <- glm(had_HSA ~ 1, data = cohort_abr_sig, family = binomial)

```

-   age

```{r}
#age quartiles
age_model <- glm(had_HSA ~ age_at_final_date_quartiles, data = cohort_abr_sig, family = binomial)

summary(age_model)
logistic.display(age_model)
lrtest(null_model,age_model)


#age bio
age_model2 <- glm(had_HSA ~ age_at_final_date_bio, data = cohort_abr_sig, family = binomial)

summary(age_model2)
logistic.display(age_model2)
lrtest(null_model,age_model2)

#age GB
age_model3 <- glm(had_HSA ~ age_at_final_date_GB, data = cohort_abr_sig, family = binomial)

summary(age_model3)
logistic.display(age_model3)
lrtest(null_model,age_model3)
```

-   in house rest of life YN, whole life quartiles

```{r}
model_2a <- glm(had_HSA ~ age_at_final_date_GB + in.the.house_YN_sum_rest_of_life_quartiles, data = cohort_abr_sig, family = binomial)

summary(model_2a)
logistic.display(model_2a)
lrtest(age_model,model_2a)


model_2b <- glm(had_HSA ~ age_at_final_date_GB + in.the.house_YN_sum_whole_life_quartiles, data = cohort_abr_sig, family = binomial)

summary(model_2b)
logistic.display(model_2b)
lrtest(age_model,model_2b)
```

-   all cancers count and YN

```{r}
model_4a<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA, data = cohort_abr_sig, family = binomial)

summary(model_4a)
logistic.display(model_4a)
lrtest(age_model,model_4a)
car::vif(model_4a)


model_4b<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_YN_no_HSA, data = cohort_abr_sig, family = binomial)

summary(model_4b)
logistic.display(model_4b)
lrtest(age_model,model_4b)
car::vif(model_4b)
```

-   freq and intensity 5y quartiles

```{r}
model_5a<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles , data = cohort_abr_sig, family = binomial)

#freq non sig
summary(model_5a)
logistic.display(model_5a)
lrtest(model_4a,model_5a)
car::vif(model_5a)


model_5b<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_intensity_5y_prior_endpoint_quartiles , data = cohort_abr_sig, family = binomial)
model_5c<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA  , data = cohort_abr_sig, family = binomial)
summary(model_5b)
lrtest(model_4a,model_5b)
summary(model_5c)
logistic.display(model_5b)

```

-   BCS whole value
    -   only not_recorded category sig = not retained

```{r}

model_6b<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_intensity_5y_prior_endpoint_quartiles + mode_purina_BCS_5y_prior_endpoint, data = cohort_abr_sig, family = binomial)
summary(model_6b)
logistic.display(model_6b)
car::vif(model_6b)
lrtest(model_5a,model_6b)

```

-   fertiliser YN whole and rest of life
    -   doesnt improve

```{r}
model_7<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life, data = cohort_abr_sig, family = binomial)
summary(model_7)
logistic.display(model_7)
car::vif(model_7)
lrtest(model_5a,model_7)
```

-   quartile antiparasites 5y prior, half antiparasites lifetime, quartile antiparasites lifetime

```{r}
model_8a<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint, data = cohort_abr_sig, family = binomial)

summary(model_8a)
logistic.display(model_8a)
car::vif(model_8a)
lrtest(model_7,model_8a)

```

-   main lifestyle 5y
    -   no levels significant exc not_recorded

```{r}
model_9<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint+ main_lifestyle_category_mode_5yrs_prior, data = cohort_abr_sig, family = binomial)
summary(model_9)
logistic.display(model_9)
car::vif(model_9)
lrtest(model_8a,model_9)
```

-   malig cancer YN, malig cancer count cant keep in two cancer variables so assess if YN or all cancer count is better

```{r}
model_10a<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint + malig_cancers_YN_no_HSA, data = cohort_abr_sig, family = binomial)
summary(model_10a)
logistic.display(model_10a)
car::vif(model_10a)
lrtest(model_8a,model_10a)



model_10b<- glm(had_HSA ~ age_at_final_date_GB  + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint + malig_cancers_YN_no_HSA, data = cohort_abr_sig, family = binomial)
summary(model_10b)
logistic.display(model_10b)

model_10c<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint , data = cohort_abr_sig, family = binomial)
summary(model_10c)
logistic.display(model_10c)


model_10d<- glm(had_HSA ~ age_at_final_date_GB  +malig_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint , data = cohort_abr_sig, family = binomial)
summary(model_10d)
logistic.display(model_10d)
car::vif(model_10a)
logistic.display(model_10b)
car::vif(model_10a)
```

-   fenbendazole ever

```{r}
model_11<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint + fenbendazole_ever, data = cohort_abr_sig, family = binomial)

summary(model_11)
logistic.display(model_11)
car::vif(model_11)
lrtest(model_8a,model_11)
```

-   NSAID ever
-   aliased with fenbendazole_ever

```{r}
model_12<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever, data = cohort_abr_sig, family = binomial)

summary(model_12)
summary(model_11)
logistic.display(model_12)
car::vif(model_12)
lrtest(model_8a,model_12)
```

-   cooking fuel secondary 5y prior
    -   only not recorded is sig = not retain

```{r}
model_13<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever + cooking_fuel_secondary_mode_5yrs_prior, data = cohort_abr_sig, family = binomial)

summary(model_13)
logistic.display(model_13)
car::vif(model_13)
lrtest(model_12,model_13)
```

-   heats_before_neutering_terciles

```{r}
model_13<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever + heats_before_neutering_terciles, data = cohort_abr_sig, family = binomial)

summary(model_13)
logistic.display(model_13)
car::vif(model_13)
lrtest(model_12,model_13)
```

-   heating fuel
-   only not recorded category sig =not retain variable

```{r}
model_14<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever + heats_before_neutering_terciles + heating_fuel_primary_mode_5yrs_prior, data = cohort_abr_sig, family = binomial)

summary(model_14)
logistic.display(model_14)
car::vif(model_14)
lrtest(model_13,model_14)
```

-   insect treatment rest of life

```{r}
model_15<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever + heats_before_neutering_terciles + any_treated_insects_study_years_rest_of_life, data = cohort_abr_sig, family = binomial)

summary(model_15)
logistic.display(model_15)
car::vif(model_15)
lrtest(model_13,model_15)
```

-   half antimicrobials lifetime, quartile antimicrobials lifetime
    -   non sig

```{r}
model_16a<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever + heats_before_neutering_terciles + any_treated_insects_study_years_rest_of_life + half_anti_microbials_lifetime, data = cohort_abr_sig, family = binomial)

summary(model_16a)
logistic.display(model_16a)
car::vif(model_16a)
lrtest(model_15,model_16a)



model_16b<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever + heats_before_neutering_terciles + any_treated_insects_study_years_rest_of_life + quartile_anti_microbials_lifetime, data = cohort_abr_sig, family = binomial)

summary(model_16b)
logistic.display(model_16b)
car::vif(model_16b)
lrtest(model_15,model_16b)
```

-   steroid ever
    -   NSAID ever preffered

```{r}
model_17<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever + heats_before_neutering_terciles + any_treated_insects_study_years_rest_of_life + steroid_ever, data = cohort_abr_sig, family = binomial)

summary(model_17)
logistic.display(model_17)
car::vif(model_17)
#steroid ever and nsaid ever aliased coeffs - assess which is better
#nsaid model
summary(model_12) 
model_17a<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  + heats_before_neutering_terciles + any_treated_insects_study_years_rest_of_life + steroid_ever, data = cohort_abr_sig, family = binomial)

summary(model_17a)
logistic.display(model_17a)
lrtest(model_15,model_17)
```

-   GI comorb 5y

```{r}
model_18<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever + heats_before_neutering_terciles + any_treated_insects_study_years_rest_of_life + gastrointestinal, data = cohort_abr_sig, family = binomial)

summary(model_18)
logistic.display(model_18)
car::vif(model_18)
lrtest(model_15,model_18)
```

-   age at neutering
    -   non sig

```{r}
model_19<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever + heats_before_neutering_terciles + any_treated_insects_study_years_rest_of_life + gastrointestinal + age_at_neuter_years_quartiles, data = cohort_abr_sig, family = binomial)

summary(model_19)
logistic.display(model_19)
car::vif(model_19)
lrtest(model_18,model_19)
```

-   quartile antiinflam lifetime
    -   non sig

```{r}
model_20<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever + heats_before_neutering_terciles + any_treated_insects_study_years_rest_of_life + gastrointestinal + quartile_anti_inflam_lifetime, data = cohort_abr_sig, family = binomial)

summary(model_20)
logistic.display(model_20)
car::vif(model_20)
lrtest(model_18,model_20)
```

-   weight 5y prior

```{r}
model_21<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever + heats_before_neutering_terciles + any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles, data = cohort_abr_sig, family = binomial)

summary(model_21)
logistic.display(model_21)
car::vif(model_21)
lrtest(model_18,model_21)
```

-   immunologicals half lifetime, quartiles
    -   non sig

```{r}
model_22a<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever + heats_before_neutering_terciles + any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + half_immunologicals_lifetime, data = cohort_abr_sig, family = binomial)

summary(model_22a)


model_22b<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever + heats_before_neutering_terciles + any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + quartile_immunologicals_lifetime, data = cohort_abr_sig, family = binomial)

summary(model_22b)
logistic.display(model_22a)
car::vif(model_22a)
lrtest(model_21,model_22a)
```

-   NSAIDs 5y prior
    -   ever is better

```{r}
#ever
model_23a<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever + heats_before_neutering_terciles + any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles , data = cohort_abr_sig, family = binomial)
summary(model_23a)
#5y
model_23b<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_5y + heats_before_neutering_terciles + any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles , data = cohort_abr_sig, family = binomial)
summary(model_23b)
```

-   half, quartiles anti-inflam 5y prior
    -   neither sig

```{r}
#half
model_24a<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever + heats_before_neutering_terciles + any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + half_anti_inflam_5y_prior_endpoint, data = cohort_abr_sig, family = binomial)
summary(model_24a)
car::vif(model_24a)
logistic.display(model_24a)
lrtest(model_21,model_24a)
#quartiles
model_24b<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_5y + heats_before_neutering_terciles + any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + quartile_anti_inflam_5y_prior_endpoint, data = cohort_abr_sig, family = binomial)
summary(model_24b)
```

-   sex
    -   cant keep heats before neutering and sex in together so select which is better - keep sex

```{r}
model_25<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex, data = cohort_abr_sig, family = binomial)
summary(model_25)
logistic.display(model_25)
model_25a<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + heats_before_neutering_terciles, data = cohort_abr_sig, family = binomial)
logistic.display(model_25a)
summary(model_25a)
car::vif(model_25)
logistic.display(model_25)
lrtest(model_21,model_25)
```

-   quartiles, half immunologicals 5y prior
    -   non sig

```{r}
model_26<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + quartile_immunologicals_5y_prior_endpoint, data = cohort_abr_sig, family = binomial)
summary(model_26)
logistic.display(model_26)
car::vif(model_26)
lrtest(model_25,model_26)
```

-   neuter

```{r}
model_27<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter, data = cohort_abr_sig, family = binomial)
summary(model_27)
logistic.display(model_27)
car::vif(model_27)
lrtest(model_25,model_27)
```

-   half enteric 5y prior, lifetime
    -   non sig

```{r}
model_28<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter + half_enteric_5y_prior_endpoint, data = cohort_abr_sig, family = binomial)
summary(model_28)
logistic.display(model_28)
car::vif(model_28)
lrtest(model_27,model_28)

model_28<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter + half_enteric_lifetime, data = cohort_abr_sig, family = binomial)
summary(model_28)
logistic.display(model_28)
car::vif(model_28)
lrtest(model_27,model_28)
```

-   height
    -   non sig

```{r}
model_29<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter + avg_height_5y_prior_endpoint_quartiles, data = cohort_abr_sig, family = binomial)
summary(model_29)
logistic.display(model_29)
car::vif(model_29)
lrtest(model_27,model_29)
```

-   region
    -   non sig

```{r}
model_30<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter + region_mode_5y_prior, data = cohort_abr_sig, family = binomial)
summary(model_30)
logistic.display(model_30)
car::vif(model_30)
lrtest(model_27,model_30)
```

-   number comorbs present lfietime - 2 levels have no variables , not include

```{r}
model_31<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter + number_comorbs_present_lifetime, data = cohort_abr_sig, family = binomial)
summary(model_31)
logistic.display(model_31)
car::vif(model_31)
lrtest(model_27,model_31)
```

-   outside YN early years and whole life
    -   non sig

```{r}
model_32<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter + outside_YN_study_years_early_life, data = cohort_abr_sig, family = binomial)
summary(model_32)
logistic.display(model_32)
car::vif(model_32)
lrtest(model_27,model_32)
```

-   fenbendazole 5y
    -   non sig

```{r}
model_33<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter + fenbendazole_5y, data = cohort_abr_sig, family = binomial)
summary(model_33)
logistic.display(model_33)
car::vif(model_33)
lrtest(model_27,model_33)
```

-   half antimicrobials 5y prior
    -   non sig

```{r}
model_34<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter + half_anti_microbials_5y_prior_endpoint, data = cohort_abr_sig, family = binomial)
summary(model_34)
logistic.display(model_34)
car::vif(model_34)
lrtest(model_27,model_34)
```

-   number comorbs 5y
    -   non sig

```{r}
model_35<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter + number_comorbs_present_5y, data = cohort_abr_sig, family = binomial)
summary(model_35)
logistic.display(model_35)
car::vif(model_35)
lrtest(model_27,model_35)
```

-   dietary half and quartiles lifetime
    -   non sig

```{r}
model_36<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter + quartile_dietary_supplements_lifetime, data = cohort_abr_sig, family = binomial)
summary(model_36)
logistic.display(model_36)
car::vif(model_36)
lrtest(model_27,model_36)
model_36<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter + half_dietary_supplements_lifetime, data = cohort_abr_sig, family = binomial)
summary(model_36)
logistic.display(model_36)
car::vif(model_36)
lrtest(model_27,model_36)
```

-   incense rest of life
    -   non sig

```{r}
model_37<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter + use_incense_or_candles_study_years_rest_of_life, data = cohort_abr_sig, family = binomial)
summary(model_37)
logistic.display(model_37)
car::vif(model_37)
lrtest(model_27,model_37)
```

-   chronic inflam 5y
    -   non sig

```{r}
model_38<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter + chronic_inflammatory, data = cohort_abr_sig, family = binomial)
summary(model_38)
logistic.display(model_38)
car::vif(model_38)
lrtest(model_27,model_38)
```

-   avg purina BCS 5y quartiles
    -   non sig

```{r}
model_39<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter + avg_purina_BCS_5y_prior_endpoint_quartiles, data = cohort_abr_sig, family = binomial)
summary(model_39)
logistic.display(model_39)
car::vif(model_39)
lrtest(model_27,model_39)
```

-   lifelong comorbs present YN, lifelong chronic inflam
    -   non sig

```{r}
model_40<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter + lifelong_chronic_inflammatory, data = cohort_abr_sig, family = binomial)
summary(model_40)
model_40<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter + lifelong_comorb_present.x, data = cohort_abr_sig, family = binomial)
summary(model_40)
logistic.display(model_40)
car::vif(model_40)
lrtest(model_27,model_40)
```

-   chronic inflam present YN
    -   non sig

```{r}
model_41<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter + chronic_inflam_present.x, data = cohort_abr_sig, family = binomial)
summary(model_41)
logistic.display(model_41)
car::vif(model_41)
lrtest(model_27,model_41)
```

-   smoke exposure YN early lfie
    -   non sig

```{r}
model_42<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter + smoke_exposure_study_years_early_life, data = cohort_abr_sig, family = binomial)
summary(model_42)
logistic.display(model_42)
car::vif(model_42)
lrtest(model_27,model_42)
```

-   all parasite comorbs 5y
    -   non sig

```{r}
model_43<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter + all_parasitism, data = cohort_abr_sig, family = binomial)
summary(model_43)
logistic.display(model_43)
car::vif(model_43)
lrtest(model_27,model_43)
```

-   dietary supplement 5y and lifetime
    -   non sig

```{r}
model_44<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter + half_dietary_supplements_5y_prior_endpoint, data = cohort_abr_sig, family = binomial)
summary(model_44)
logistic.display(model_44)
car::vif(model_44)
lrtest(model_27,model_44)
```

-   majority sleep location lifetime
    -   non sig

```{r}
model_45<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter + majority_sleep_location_lifetime, data = cohort_abr_sig, family = binomial)
summary(model_45)
logistic.display(model_45)
car::vif(model_45)
lrtest(model_27,model_45)
```

-   steroids 5y prior
    -   not sig

```{r}
model_46<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter + steroid_5y, data = cohort_abr_sig, family = binomial)
summary(model_46)
logistic.display(model_46)
car::vif(model_46)
lrtest(model_27,model_46)
```

-   quartiles antimicrobial 5y prior
    -   not sig

```{r}
model_47<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter + quartile_anti_microbials_5y_prior_endpoint, data = cohort_abr_sig, family = binomial)
summary(model_47)
logistic.display(model_47)
car::vif(model_47)
lrtest(model_27,model_47)
```

Final model

```{r}
model_f<- glm(had_HSA ~ age_at_final_date_GB  +all_cancers_count_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + any_treated_fertilizer_study_years_whole_life + quartile_anti_parasites_5y_prior_endpoint  +NSAID_ever +  any_treated_insects_study_years_rest_of_life + gastrointestinal + avg_weight_5y_prior_endpoint_quartiles + sex + neuter, data = cohort_abr_sig, family = binomial)
summary(model_f)
logistic.display(model_f)
car::vif(model_f)
```

```{r}
dependent <- "had_HSA"
explanatory <- c("age_at_final_date_GB","all_cancers_YN_no_HSA","avg_frequency_5y_prior_endpoint_quartiles","any_treated_fertilizer_study_years_whole_life","quartile_anti_parasites_5y_prior_endpoint","NSAID_ever","any_treated_insects_study_years_rest_of_life","gastrointestinal","avg_weight_5y_prior_endpoint_quartiles","sex","neuter")
final_model_table <-cohort_abr_sig %>% 
  finalfit(dependent, explanatory, metrics = TRUE,column=TRUE) %>%
  kable(format = "html", digits = 2, escape = FALSE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

final_model_table

final_model_table2<- cohort_abr_sig %>% 
  finalfit(dependent, explanatory, metrics = TRUE,column = TRUE) 
final_model_table2_df <- as.data.frame(final_model_table2)

doc <- read_docx() %>%
results_flextable <-flextable(final_model_table2_df)



# Create a Word document
doc <- read_docx()


# Add the flextable to the Word document
doc <- doc %>%
  body_add_flextable(results_flextable) %>%
  body_add_par("Finalfit Regression Results", style = "heading 1")
print(results_flextable, target = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/finalfit_logreg_table_all_variables.docx")

```

```{r}
cohort_abr_sig %>%
  finalfit(dependent, explanatory, method = "glm", family = "binomial") -> logreg_model

plot <- or_plot(.data = cohort_abr_sig,
                 dependent = "had_HSA", 
                 explanatory = c("age_at_final_date_GB",
                                 "all_cancers_YN_no_HSA",
                                 "avg_frequency_5y_prior_endpoint_quartiles",
                                 "any_treated_fertilizer_study_years_whole_life",
                                 "quartile_anti_parasites_5y_prior_endpoint",
                                 "NSAID_ever",
                                 "any_treated_insects_study_years_rest_of_life",
                                 "gastrointestinal",
                                 "avg_weight_5y_prior_endpoint_quartiles",
                                 "sex"))


plot2 <-cohort_abr_sig %>%
  or_plot(dependent, explanatory,column_space = c(-0.5, -0.1, 0.5),table_text_size = 3, plot_opts = list(xlim(0.1, 40)))
plot2
print(plot2)
```

### Model build with only 5y prior or no time point variables

```{r}
cols_to_keep_5y_or_no_time <- c("heating_fuel_primary_mode_5yrs_prior","cooking_fuel_secondary_mode_5yrs_prior",
                              "region_mode_5y_prior","number_comorbs_present_5y","all_parasitism","gastrointestinal","inflammatory_other","chronic_inflammatory","avg_intensity_5y_prior_endpoint_quartiles","avg_frequency_5y_prior_endpoint_quartiles","main_lifestyle_category_mode_5yrs_prior","use_hepa_filter_exposed_5yrs_prior","any_treated_weeds_exposed_5yrs_prior","any_treated_insects_exposed_5yrs_prior","any_treated_fertilizer_exposed_5yrs_prior","sex","neuter","spayed_while_in_heat","avg_weight_5y_prior_endpoint_quartiles","avg_height_5y_prior_endpoint_quartiles","avg_purina_BCS_5y_prior_endpoint_quartiles","mode_purina_BCS_5y_prior_endpoint","half_anti_inflam_5y_prior_endpoint","half_anti_microbials_5y_prior_endpoint","half_enteric_5y_prior_endpoint","half_immunologicals_5y_prior_endpoint", "half_misc_5y_prior_endpoint","half_anti_parasites_5y_prior_endpoint","half_dietary_supplements_5y_prior_endpoint","half_fluid_metabolites_5y_prior_endpoint","quartile_anti_inflam_5y_prior_endpoint","quartile_anti_microbials_5y_prior_endpoint","quartile_immunologicals_5y_prior_endpoint","quartile_misc_5y_prior_endpoint","quartile_anti_parasites_5y_prior_endpoint","quartile_dietary_supplements_5y_prior_endpoint","age_at_final_date_GB","age_at_neuter_years_quartiles","malig_cancers_count_no_HSA","all_cancers_count_no_HSA","all_cancers_count","malig_cancers_YN_no_HSA","all_cancers_YN_no_HSA","NSAID_5y","steroid_5y","fenbendazole_5y","age_at_final_date_quartiles","age_at_final_date_bio","sex_neuter","lifelong_comorb_present.x","had_HSA")


#set levels for purina BCS mode
cohort_abr_sig_5y <- cohort_abr_sig_5y%>%
  mutate(mode_purina_BCS_5y_prior_endpoint = factor(mode_purina_BCS_5y_prior_endpoint, levels = c("5", "3", "4", "6","7","8","9","not_recorded")))



cohort_abr_sig_5y <- cohort_abr_sig_5y%>%
  mutate(age_at_final_date_GB = factor(age_at_final_date_GB, levels = c(">7-<9y","<5y", ">5-<7y",  ">9-<11y",">11-<13y",">13y")))
cohort_abr_sig_5y <- cohort_abr2[,cols_to_keep_5y_or_no_time]

```

Null model

```{r}
null_model <- glm(had_HSA ~ 1, data = cohort_abr_sig_5y, family = binomial)

```

Age

```{r}
model_age1 <- glm(had_HSA ~ age_at_final_date_quartiles, data=cohort_abr_sig_5y,family=binomial)


summary(model_age1)
logistic.display(model_age1)
#car::vif(model_a)
lrtest(null_model,model_age1)

model_age2 <- glm(had_HSA ~ age_at_final_date_bio, data=cohort_abr_sig_5y,family=binomial)


summary(model_age2)
logistic.display(model_age2)
#car::vif(model_a)
lrtest(null_model,model_age2)


model_age3 <- glm(had_HSA ~ age_at_final_date_GB, data=cohort_abr_sig_5y,family=binomial)


summary(model_age3)
logistic.display(model_age3)
#car::vif(model_a)
lrtest(null_model,model_age3)
```

Cancer counts and YN

-   all cancer YN has lowest AIC but all are sig, elect to take forward only malig_cancers_YN variable as removes poss issues with benign masses

```{r}
model_cancer1 <- glm(had_HSA ~ age_at_final_date_GB + all_cancers_YN_no_HSA, data=cohort_abr_sig_5y,family=binomial)


summary(model_cancer1)
logistic.display(model_cancer1)
car::vif(model_cancer1)
lrtest(model_age3,model_cancer1)


model_cancer2<- glm(had_HSA ~ age_at_final_date_GB + all_cancers_count_no_HSA, data=cohort_abr_sig_5y,family=binomial)


summary(model_cancer2)
logistic.display(model_cancer2)
car::vif(model_cancer2)
lrtest(model_age3,model_cancer2)




model_cancer3<- glm(had_HSA ~ age_at_final_date_GB + malig_cancers_count_no_HSA, data=cohort_abr_sig_5y,family=binomial)


summary(model_cancer3)
logistic.display(model_cancer3)
car::vif(model_cancer3)
lrtest(model_age3,model_cancer3)




model_cancer4<- glm(had_HSA ~ age_at_final_date_GB + malig_cancers_YN_no_HSA, data=cohort_abr_sig_5y,family=binomial)


summary(model_cancer4)
logistic.display(model_cancer4)
car::vif(model_cancer4)
lrtest(model_age3,model_cancer4)
```

Frequency and intensity 5y prior

-   keep freq

```{r}
model_freq<- glm(had_HSA ~ age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles, data=cohort_abr_sig_5y,family=binomial)


summary(model_freq)
logistic.display(model_freq)
car::vif(model_freq)
lrtest(model_cancer1,model_freq)

model_int<- glm(had_HSA ~ age_at_final_date_GB + all_cancers_YN_no_HSA +  avg_intensity_5y_prior_endpoint_quartiles, data=cohort_abr_sig_5y,family=binomial)


summary(model_int)
logistic.display(model_int)
car::vif(model_int)
lrtest(model_cancer1,model_int)
```

Weeds 5y

```{r}
model_weed<- glm(had_HSA ~ age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles +any_treated_weeds_exposed_5yrs_prior, data=cohort_abr_sig_5y,family=binomial)


summary(model_weed)
logistic.display(model_weed)
car::vif(model_weed)
lrtest(model_weed,model_freq)
```

Fertiliser 5y

-   non sig

```{r}
model_fert<- glm(had_HSA ~ age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles +any_treated_weeds_exposed_5yrs_prior + any_treated_fertilizer_exposed_5yrs_prior, data=cohort_abr_sig_5y,family=binomial)


summary(model_fert)
logistic.display(model_fert)
car::vif(model_fert)
lrtest(model_weed,model_fert)
```

NSAID 5y

-   sig but weed no longer

```{r}
model_a<- glm(had_HSA ~ age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles +any_treated_weeds_exposed_5yrs_prior + NSAID_5y, data=cohort_abr_sig_5y,family=binomial)



summary(model_a)
logistic.display(model_a)
car::vif(model_a)
lrtest(model_weed,model_a)

model_a1<- glm(had_HSA ~ age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles + NSAID_5y, data=cohort_abr_sig_5y,family=binomial)
```

Half anti-inflam 5y/quartile anti-inflam

-   both non sig in model (and not collinear with VIF)

```{r}
model_b1 <- glm(had_HSA ~ age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y + half_anti_inflam_5y_prior_endpoint, data=cohort_abr_sig_5y,family=binomial)


summary(model_b1)
logistic.display(model_b1)
car::vif(model_b1)
lrtest(model_a1,model_b1)

model_b2 <- glm(had_HSA ~ age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y +  quartile_anti_inflam_5y_prior_endpoint, data=cohort_abr_sig_5y,family=binomial)


summary(model_b2)
logistic.display(model_b2)
car::vif(model_b2)
lrtest(model_a1,model_b2)
```

Half immunologicals 5y/quartiles

-   non sig

```{r}
model_c1 <- glm(had_HSA ~age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y  + half_immunologicals_5y_prior_endpoint, data=cohort_abr_sig_5y,family=binomial)


summary(model_c1)
logistic.display(model_c1)
car::vif(model_c1)
lrtest(model_a1,model_c1)

model_c2 <- glm(had_HSA ~ age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y + quartile_immunologicals_5y_prior_endpoint, data=cohort_abr_sig_5y,family=binomial)


summary(model_c2)
logistic.display(model_c2)
car::vif(model_c2)
lrtest(model_a1,model_c2)
```

Half antiparasites 5y

-   not sig

```{r}
model_d <- glm(had_HSA ~ age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y + half_anti_parasites_5y_prior_endpoint, data=cohort_abr_sig_5y,family=binomial)


summary(model_d)
logistic.display(model_d)
car::vif(model_d)
lrtest(model_a1,model_d)
```

sex-neuter, sex, neuter

-   sex-neuter lowestAIC

```{r}
#sex
model_e1 <- glm(had_HSA ~ age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y  + sex, data=cohort_abr_sig_5y,family=binomial)


summary(model_e1)
logistic.display(model_e1)
car::vif(model_e1)
lrtest(model_a1,model_e1)
#sex neut
model_e2 <- glm(had_HSA ~ age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y  + sex_neuter, data=cohort_abr_sig_5y,family=binomial)


summary(model_e2)
logistic.display(model_e2)
car::vif(model_e2)
lrtest(model_a1,model_e2)
#neut

model_e3 <- glm(had_HSA ~ age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y  + neuter, data=cohort_abr_sig_5y,family=binomial)


summary(model_e3)
logistic.display(model_e3)
car::vif(model_e3)
lrtest(model_a1,model_e3)
```

region mode 5y

-non sig

```{r}

model_f <- glm(had_HSA ~age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y  + sex_neuter + region_mode_5y_prior, data=cohort_abr_sig_5y,family=binomial)
summary(model_f)
logistic.display(model_f)
car::vif(model_f)
lrtest(model_e2,model_f)
```

age at neuter quartiles

-   cant have with age at final date variable

```{r}

model_g <- glm(had_HSA ~age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y  + sex_neuter + age_at_neuter_years_quartiles, data=cohort_abr_sig_5y,family=binomial)
summary(model_g)
logistic.display(model_g)
car::vif(model_g)
lrtest(model_e2,model_g)
```

half dietary supp 5y

-   not sig

```{r}
model_h <- glm(had_HSA ~age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y  + sex_neuter + half_dietary_supplements_5y_prior_endpoint, data=cohort_abr_sig_5y,family=binomial)
summary(model_h)
logistic.display(model_h)
car::vif(model_h)
lrtest(model_e2,model_h)
```

fenbendazole 5y

-   collinear with NSAID_5y

-   NSAID lower AIC so keep instead

```{r}
model_i <- glm(had_HSA ~age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y  + sex_neuter + fenbendazole_5y, data=cohort_abr_sig_5y,family=binomial)
summary(model_i)
logistic.display(model_i)
car::vif(model_i)
lrtest(model_e2,model_i)


model_i2 <- glm(had_HSA ~age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles   + sex_neuter + fenbendazole_5y, data=cohort_abr_sig_5y,family=binomial)
summary(model_i2)
logistic.display(model_i2)
car::vif(model_i2)
lrtest(model_e2,model_i2)
```

main lifestyle category 5y

-   only not recorded var csategory is sig= not retain

```{r}
model_j <- glm(had_HSA ~age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y  + sex_neuter + main_lifestyle_category_mode_5yrs_prior, data=cohort_abr_sig_5y,family=binomial)
summary(model_j)
logistic.display(model_j)
car::vif(model_j)
lrtest(model_e2,model_j)
```

half antimicrobials 5y

-   noyt sig

```{r}
model_k <- glm(had_HSA ~age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y  + sex_neuter + half_anti_microbials_5y_prior_endpoint, data=cohort_abr_sig_5y,family=binomial)
summary(model_k)
logistic.display(model_k)
car::vif(model_k)
lrtest(model_e2,model_k)
```

avg purina BCS 5y

-   not sig

```{r}
model_l <- glm(had_HSA ~age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y  + sex_neuter + avg_purina_BCS_5y_prior_endpoint_quartiles, data=cohort_abr_sig_5y,family=binomial)
summary(model_l)
logistic.display(model_l)
car::vif(model_l)
lrtest(model_e2,model_l)
```

lifelong comorb present

-   not sig

```{r}
model_m <- glm(had_HSA ~age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y  + sex_neuter + lifelong_comorb_present.x, data=cohort_abr_sig_5y,family=binomial)
summary(model_m)
logistic.display(model_m)
car::vif(model_m)
lrtest(model_e2,model_m)

```

chronic inflam 5y

-   not sig

```{r}
model_n <- glm(had_HSA ~age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y  + sex_neuter + chronic_inflammatory, data=cohort_abr_sig_5y,family=binomial)
summary(model_n)
logistic.display(model_n)
car::vif(model_n)
lrtest(model_e2,model_n)
```

inflammatory other

-   not sig

```{r}
model_o <- glm(had_HSA ~age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y  + sex_neuter + inflammatory_other, data=cohort_abr_sig_5y,family=binomial)
summary(model_o)
logistic.display(model_o)
car::vif(model_o)
lrtest(model_e2,model_o)
```

all parasitism

-   not sig

```{r}
model_p <- glm(had_HSA ~age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y  + sex_neuter + all_parasitism, data=cohort_abr_sig_5y,family=binomial)
summary(model_p)
logistic.display(model_p)
car::vif(model_p)
lrtest(model_e2,model_p)
```

steroid 5y

-   collinear with NSAID 5y, NSAID better fit

```{r}
model_q <- glm(had_HSA ~age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y  + sex_neuter + steroid_5y, data=cohort_abr_sig_5y,family=binomial)
summary(model_q)
logistic.display(model_q)
car::vif(model_q)
lrtest(model_e2,model_q)
```

number comorbs present 5y

-   not sig

```{r}
model_r <- glm(had_HSA ~age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y  + sex_neuter + number_comorbs_present_5y, data=cohort_abr_sig_5y,family=binomial)
summary(model_r)
logistic.display(model_r)
car::vif(model_r)
lrtest(model_e2,model_r)
```

hepa filter 5y

-not sig

```{r}
model_s <- glm(had_HSA ~age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y  + sex_neuter + use_hepa_filter_exposed_5yrs_prior, data=cohort_abr_sig_5y,family=binomial)
summary(model_s)
logistic.display(model_s)
car::vif(model_s)
lrtest(model_e2,model_s)
```

quartile antimicrobials 5y

-   not isg

```{r}
model_t <- glm(had_HSA ~age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y  + sex_neuter + use_hepa_filter_exposed_5yrs_prior, data=cohort_abr_sig_5y,family=binomial)
summary(model_t)
logistic.display(model_t)
car::vif(model_t)
lrtest(model_e2,model_t)
```

mode purina BCS 5y

-   sig although only for not recorded catregory

```{r}
model_u <- glm(had_HSA ~age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y  + sex_neuter + mode_purina_BCS_5y_prior_endpoint, data=cohort_abr_sig_5y,family=binomial)
summary(model_u)
logistic.display(model_u)
car::vif(model_u)
lrtest(model_e2,model_u)
```

height 5y

-   not sig

```{r}
model_v<- glm(had_HSA ~age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y  + sex_neuter + avg_height_5y_prior_endpoint_quartiles, data=cohort_abr_sig_5y,family=binomial)
summary(model_v)
logistic.display(model_v)
car::vif(model_v)
lrtest(model_e2,model_v)
```

heating fuel primary 5y

```{r}
model_w <- glm(had_HSA ~age_at_final_date_GB + all_cancers_YN_no_HSA + avg_frequency_5y_prior_endpoint_quartiles  + NSAID_5y  + sex_neuter + heating_fuel_primary_mode_5yrs_prior, data=cohort_abr_sig_5y,family=binomial)
summary(model_w)
logistic.display(model_w)
car::vif(model_w)
lrtest(model_e2,model_w)
```

Final model output with time restricted variables:
H&L p=0.07 Cstatistics 0.78


```{r}
dependent <- "had_HSA"
explanatory <- c("age_at_final_date_GB","sex_neuter","all_cancers_YN_no_HSA","avg_frequency_5y_prior_endpoint_quartiles","NSAID_5y","heating_fuel_primary_mode_5yrs_prior")
final_model_table <-cohort_abr_sig_5y %>% 
  finalfit(dependent, explanatory, metrics = TRUE,column=TRUE) %>%
  kable(format = "html", digits = 2, escape = FALSE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

final_model_table

final_model_table2<- cohort_abr_sig_5y %>% 
  finalfit(dependent, explanatory, metrics = FALSE) 
final_model_table2_df <- as.data.frame(final_model_table2)

results_flextable <-flextable(final_model_table2_df)



# Create a Word document
doc <- read_docx()


# Add the flextable to the Word document
doc <- doc %>%
  body_add_flextable(results_flextable) %>%
  body_add_par("Finalfit Regression Results", style = "heading 1")
print(results_flextable, target = "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/finalfit_logreg_table_5y_restricted_variables.docx")

```

```{r}
cohort_abr_sig_5y %>%
  finalfit(dependent, explanatory, method = "glm", family = "binomial") -> logreg_model

plot <- or_plot(.data = cohort_abr_sig,
                 dependent = "had_HSA", 
                 explanatory = c("age_at_final_date_GB","sex_neuter","all_cancers_YN_no_HSA","avg_frequency_5y_prior_endpoint_quartiles","NSAID_5y","heating_fuel_primary_mode_5yrs_prior"))


plot2 <-cohort_abr_sig_5y %>%
  or_plot(dependent, explanatory,column_space = c(-0.5, -0.1, 0.5),table_text_size = 3, plot_opts = list(xlim(0, 8)))
plot2
print(plot2)
```

