---
title: "GRLS log reg HSA revised analysis approach"
format: html
editor: visual
---

# Variables associated with HSA in GRLS cohort

Load in dataset and packages

```{r}
library(tidyverse)
library(tidyverse)
library(dplyr)
library(survival)
library(ggplot2)
library(tibble)
library(kableExtra)
library(epiDisplay)
library(lme4)
#devtools::install_github("zabore/ezfun")
ezfun::set_ccf_palette("contrast")
library(lubridate)
library(ggsurvfit)
library(gtsummary)
library(tidycmprsk)
library(gtable)
#devtools::install_github("zabore/condsurv")
library(condsurv)
library(survminer)
library(cardx)
library(finalfit)
library(crosstable)
library(kableExtra)
library(officer)
library(flextable)
library(writexl)
library(readr)
# data was preliminarily tidied in GRLS cox HSA time to diagnosis script before changing to log reg analyses
cohort_abr <-read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_cohort_all_RFs_data_tidied_from_cox.csv")
```

## Updated weight data

Incorporate in updated weight from Zoe's work (she has modified the erroneus lb kg issues also now all only adult weight so no \<18m weights in here)

```{r}
updated_weight <- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Data/weight_cleaned_grls.csv")

#existing weight columns are average and median 5y prior to endpoint and either split into quartiles or biological

#use the updated_weight df to create these same variables again + whole life weight

weight_df <- updated_weight %>%
  mutate(
    weight_kg = case_when(
      str_to_lower(weight_units) %in% c("lb", "lbs") ~ weight * 0.45359237,
      str_to_lower(weight_units) == "kg" ~ weight,
      TRUE ~ NA_real_
    )
  )

median_weight <- weight_df %>%
  filter(!is.na(weight_kg)) %>%
  group_by(subject_id) %>%
  arrange(year_in_study, .by_group = TRUE) %>%
  summarise(
    median_weight_lifetime_updated = median(weight_kg, na.rm = TRUE),

    median_weight_5y_prior_endpoint_updated = median(
      weight_kg[year_in_study %in% tail(sort(unique(year_in_study)), 5)],
      na.rm = TRUE
    ),

    .groups = "drop"
  )

#add into main df and see how different values are
cohort_abr_updated_weight2<- cohort_abr %>%
  left_join(median_weight, by="subject_id")


#check how diff
  
  
summarise(
  cohort_abr_updated_weight2,
  old_mean = mean(as.numeric(na_if(median_weight_5y_prior_endpoint_kg, "not_recorded")), na.rm = TRUE),
  old_median = median(as.numeric(na_if(median_weight_5y_prior_endpoint_kg, "not_recorded")), na.rm = TRUE),
  old_q1 = quantile(as.numeric(na_if(median_weight_5y_prior_endpoint_kg, "not_recorded")), 0.25, na.rm = TRUE),
  old_q3 = quantile(as.numeric(na_if(median_weight_5y_prior_endpoint_kg, "not_recorded")), 0.75, na.rm = TRUE),

  new_mean = mean(median_weight_5y_prior_endpoint_updated, na.rm = TRUE),
  new_median = median(median_weight_5y_prior_endpoint_updated, na.rm = TRUE),
  new_q1 = quantile(median_weight_5y_prior_endpoint_updated, 0.25, na.rm = TRUE),
  new_q3 = quantile(median_weight_5y_prior_endpoint_updated, 0.75, na.rm = TRUE)
)  


#remove old weight rows from df
cohort_abr_updated_weight2<- cohort_abr_updated_weight2 %>%
  dplyr::select(-c("median_weight_5y_prior_endpoint_kg","avg_weight_5y_prior_endpoint_kg","avg_weight_5y_prior_endpoint_kg_outliers_removed","median_weight_5y_prior_endpoint_any_records_kg","median_weight_5y_prior_endpoint_kg_outliers_removed","median_weight_5y_prior_endpoint_any_records_kg","median_weight_5y_prior_endpoint_any_records_kg_outliers_removed",
                  "avg_weight_5y_prior_endpoint_quartiles",
                  "median_weight_5y_prior_endpoint_kg_outliers_removed_quartiles","avg_weight_5y_prior_endpoint_kg_biological","median_weight_5y_prior_endpoint_any_records_kg_outliers_removed_biological",
                  "median_weight_5y_prior_endpoint_kg_outliers_removed_biological"
                  ))

#now recode the new weight data into quartiles
cuts <- quantile(
  cohort_abr_updated_weight2$median_weight_5y_prior_endpoint_updated,
  probs = c(0, 0.25, 0.5, 0.75, 1),
  na.rm = TRUE
)

labels <- c(
  paste0(round(cuts[1], 1), "–", round(cuts[2], 1), " kg"),
  paste0(round(cuts[2], 1), "–", round(cuts[3], 1), " kg"),
  paste0(round(cuts[3], 1), "–", round(cuts[4], 1), " kg"),
  paste0(round(cuts[4], 1), "–", round(cuts[5], 1), " kg")
)

cohort_abr_updated_weight2 <- cohort_abr_updated_weight2%>%
  mutate(
    weight_quartile = case_when(
      is.na(median_weight_5y_prior_endpoint_updated) ~ "not_recorded",
      TRUE ~ as.character(cut(
        median_weight_5y_prior_endpoint_updated,
        breaks = cuts,
        include.lowest = TRUE,
        labels = labels
      ))
    ),
    weight_quartile = factor(
      weight_quartile,
      levels = c(labels, "not_recorded")
    )
  )


#also creating a 'biological' weight variable
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    weight_biological = case_when(
      is.na(median_weight_5y_prior_endpoint_updated) ~ "not_recorded",
      median_weight_5y_prior_endpoint_updated <= 25 ~ "≤25 kg",
      median_weight_5y_prior_endpoint_updated <= 35 ~ ">25–≤35 kg",
      median_weight_5y_prior_endpoint_updated > 35 ~ ">35 kg"
    ),
    weight_biological = factor(
      weight_biological,
      levels = c(
        "≤25 kg",
        ">25–≤35 kg",
        ">35 kg",
        "not_recorded"
      )
    )
  )
```

Converting to factors for analysis:

```{r}

cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    across(
      everything(),
      as.character
    )
  ) %>%
  mutate(
    across(
      everything(),
      ~ na_if(.x, "n/a")
    )
  ) %>%
  mutate(
    across(
      everything(),
      ~ na_if(.x, "NA")
    )
  ) %>%
  mutate(
    across(
      everything(),
      ~ na_if(.x, "nr")
    )
  ) %>%
  mutate(
    across(
      everything(),
      ~ na_if(.x, "not recorded")
    )
  ) %>%
  mutate(
    across(
      everything(),
      ~ na_if(.x, "")
    )  ) %>%
  mutate(
    across(
      everything(),
      ~ replace_na(.x, "not_recorded")
    )
  )

# Factor version of outcome for descriptive tables
cohort_abr_updated_weight2$status_factor <- factor(
  cohort_abr_updated_weight2$had_HSA
)

# Numeric binary outcome for logistic regression
cohort_abr_updated_weight2$status <- as.numeric(
  as.character(cohort_abr_updated_weight2$had_HSA)
)

cohort_abr_updated_weight2$had_HSA <- as.numeric(
  as.character(cohort_abr_updated_weight2$had_HSA)
)

# Convert survival time back to numeric
cohort_abr_updated_weight2$survival_time_months <- as.numeric(
  as.character(cohort_abr_updated_weight2$survival_time_months)
)


#change spayed while in heat to account for males and entire females
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    spayed_while_in_heat = case_when(
      sex == "M" ~ "Male",
      sex_status == "Intact Female" ~ "Intact Female",
      spayed_while_in_heat == "0" ~ "0",
      spayed_while_in_heat == "1" ~ "1",
      TRUE ~ "not_recorded"
    )
  )


# Convert heats before neutering to numeric temporarily
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    heats_before_neutering_numeric = case_when(
      heats_before_neutering == "not_recorded" ~ NA_real_,
      TRUE ~ suppressWarnings(
        as.numeric(as.character(heats_before_neutering))
      )
    )
  )

# Create a temporary vector for inspecting non-zero recorded values
heats <- cohort_abr_updated_weight2 %>%
  filter(
    !is.na(heats_before_neutering_numeric),
    heats_before_neutering_numeric != 0
  ) %>%
  pull(heats_before_neutering_numeric)


# Function to calculate quartile or tercile breaks dynamically
get_breaks <- function(x, probs) {

  breaks <- quantile(
    x,
    probs = probs,
    na.rm = TRUE
  )

  if (length(unique(breaks)) < length(probs)) {

    message(
      "Quartile breaks are not unique; switching to terciles."
    )

    breaks <- quantile(
      x,
      probs = c(0, 1 / 3, 2 / 3, 1),
      na.rm = TRUE
    )
  }

  breaks
}


breaks <- get_breaks(
  heats,
  c(0, 0.25, 0.5, 0.75, 1)
)


labels <- paste0(
  "≥",
  round(head(breaks, -1), 2),
  " - <",
  round(tail(breaks, -1), 2)
)


# Create manually grouped heats-before-neutering variable
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    heats_before_neutering_tercile = case_when(
      sex == "M" ~ "Male",
      sex_status == "Intact Female" ~ "Intact Female",

      heats_before_neutering_numeric == 0 ~ "0",

      heats_before_neutering_numeric %in% c(
        1, 2
      ) ~ "1-<=2",

      heats_before_neutering_numeric %in% c(
        3, 4
      ) ~ ">2-<=4",

      heats_before_neutering_numeric >= 5 &
        heats_before_neutering_numeric <= 12 ~ ">4-12",

      TRUE ~ "not_recorded"
    )
  )



# Create combined sex-neuter variable
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    sex_neuter = paste(
      sex,
      neuter,
      sep = " "
    )
  )


#change no_int_recorded to not_recorded
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  dplyr::mutate(
    mode_intensity_5y_prior_endpoint =
      forcats::fct_recode(
        mode_intensity_5y_prior_endpoint,
        not_recorded = "no_int_recorded"
      )
  )

#------------------------------------------------------------
# CONVERT CHARACTER VARIABLES TO FACTORS
#------------------------------------------------------------

# Convert all remaining character columns to factors.
# Move "not_recorded" to the final level so it is not used
# as the default reference category.

convert_and_relevel <- function(df) {
df %>%
    mutate(
      across(
        where(is.character),
        factor
      )
    ) %>%
    mutate(
      across(
        where(is.factor),
        ~ {
          if ("not_recorded" %in% levels(.x)) {
            forcats::fct_relevel(
              .x,
              "not_recorded",
              after = Inf
            )
          } else {
            .x
          }
        }
      )
    )
}


cohort_abr_updated_weight2 <- convert_and_relevel(
  cohort_abr_updated_weight2
)
#------------------------------------------------------------
# RESTORE CONTINUOUS VARIABLES TO NUMERIC
#------------------------------------------------------------

# These variables must be numeric so that logistic regression fits
# one coefficient per one-unit increase rather than one coefficient
# for every observed value.

continuous_variables <- c(
  "mean_activity_level_overview",

  "whole_activity_mean",
  "early_activity_mean",
  "rest_activity_mean",
  "prior5y_activity_mean",

  "avg_activity_duration_sy1_2",
  "avg_activity_intensity_sy1_2",
  "avg_activity_freq_sy1_2",

  "avg_activity_freq_sy3",
  "avg_activity_intensity_sy3",

  "avg_intensity_5y_prior_endpoint",
  "avg_frequency_5y_prior_endpoint",

  "avg_height_5y_prior_endpoint",

  "median_weight_lifetime_updated",
  "median_weight_5y_prior_endpoint_updated"
)

# Check that all named variables exist
missing_continuous_variables <- setdiff(
  continuous_variables,
  names(cohort_abr_updated_weight2)
)

if (length(missing_continuous_variables) > 0) {

  warning(
    paste(
      "These continuous variables are missing from the dataframe:",
      paste(
        missing_continuous_variables,
        collapse = ", "
      )
    )
  )
}


# Retain only continuous variables that actually exist
continuous_variables_present <- intersect(
  continuous_variables,
  names(cohort_abr_updated_weight2)
)

# Convert factor -> character -> numeric
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    across(
      all_of(continuous_variables_present),
      ~ suppressWarnings(
        as.numeric(as.character(.x))
      )
    )
  )


#------------------------------------------------------------
# CHECK NUMERIC CONVERSION
#------------------------------------------------------------
continuous_variable_classes <- tibble(
  variable = continuous_variables_present,
  variable_class = vapply(
    cohort_abr_updated_weight2[
      continuous_variables_present
    ],
    function(x) {
      paste(
        class(x),
        collapse = "/"
      )
    },
    character(1)
  ),
  number_missing = vapply(
    cohort_abr_updated_weight2[
      continuous_variables_present
    ],
    function(x) {
      sum(is.na(x))
    },
    integer(1)
  ),
  number_unique = vapply(
    cohort_abr_updated_weight2[
      continuous_variables_present
    ],
    function(x) {
      dplyr::n_distinct(
        x,
        na.rm = TRUE
      )
    },
    integer(1)
  )
)
continuous_variable_classes


# Stop the script if any continuous variable is still not numeric
non_numeric_continuous_variables <- continuous_variable_classes %>%
  filter(
    variable_class != "numeric"
  )

if (nrow(non_numeric_continuous_variables) > 0) {

  stop(
    paste(
      "The following intended continuous variables are not numeric:",
      paste(
        non_numeric_continuous_variables$variable,
        collapse = ", "
      )
    )
  )
}


#------------------------------------------------------------
# SET MANUAL FACTOR LEVELS
#------------------------------------------------------------
# Heats before neutering
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    heats_before_neutering_tercile = factor(
      heats_before_neutering_tercile,
      levels = c(
        "0",
        "1-<=2",
        ">2-<=4",
        ">4-12",
        "Intact Female",
        "Male",
        "not_recorded"
      )
    )
  )


# Neutering reason
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    neutered_reason = factor(
      neutered_reason,
      levels = c(
        "Elective",
        "Medical Reason",
        "Entire",
        "Unknown",
        "not_recorded"
      )
    )
  )
# Mode BCS
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    mode_purina_BCS_5y_prior_endpoint = factor(
      mode_purina_BCS_5y_prior_endpoint,
      levels = c(
        "5",
        "3",
        "4",
        "6",
        "7",
        "8",
        "9",
        "1",
        "2",
        "not_recorded"
      )
    )
  )
# Reclassify mode BCS
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    mode_BCS_classify = case_when(
      as.character(
        mode_purina_BCS_5y_prior_endpoint
      ) %in% c(
        "1", "2", "3"
      ) ~ "underweight",

      as.character(
        mode_purina_BCS_5y_prior_endpoint
      ) %in% c(
        "4", "5"
      ) ~ "ideal",

      as.character(
        mode_purina_BCS_5y_prior_endpoint
      ) %in% c(
        "6", "7", "8", "9"
      ) ~ "overweight",

      TRUE ~ "not_recorded"
    ),
    mode_BCS_classify = factor(
      mode_BCS_classify,
      levels = c(
        "ideal",
        "underweight",
        "overweight",
        "not_recorded"
      )
    )
  )
# Age quartiles
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    age_at_final_date_quartiles = factor(
      age_at_final_date_quartiles,
      levels = c(
        "0.74-<8.47 years",
        ">=8.47-<10.68 years",
        ">=10.68-<11.64 years",
        ">=11.64-14.51 years",
        "not_recorded"
      )
    )
  )


# Biological age groups
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    age_at_final_date_bio = factor(
      age_at_final_date_bio,
      levels = c(
        ">5-<8y",
        "<5y",
        ">8-<11y",
        ">11-<13y",
        ">13y",
        "not_recorded"
      )
    )
  )


# Golden Retriever Lifetime Study age groups
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    age_at_final_date_GB = factor(
      age_at_final_date_GB,
      levels = c(
        ">7-<9y",
        "<5y",
        ">5-<7y",
        ">9-<11y",
        ">11-<13y",
        ">13y",
        "not_recorded"
      )
    )
  )


# Height quartiles
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    avg_height_5y_prior_endpoint_quartiles = factor(
      avg_height_5y_prior_endpoint_quartiles,
      levels = c(
        "49.80-<56.30cm",
        ">=56.30-<58.29cm",
        ">=58.29-<60.32cm",
        ">=60.32cm-67.73cm",
        "not_recorded"
      )
    )
  )
# Average BCS quartiles
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    avg_purina_BCS_5y_prior_endpoint_quartiles = factor(
      avg_purina_BCS_5y_prior_endpoint_quartiles,
      levels = c(
        "3.3-<5",
        ">=5-<5.28",
        ">=5.28-<5.86",
        ">=5.86 - 9",
        "not_recorded"
      )
    )
  )


# Age at neutering quartiles
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    age_at_neuter_years_quartiles = factor(
      age_at_neuter_years_quartiles,
      levels = c(
        "0.15-<0.52y",
        ">=0.52-<0.83y",
        ">=0.83-2.44y",
        ">=2.44-9.59y",
        "Entire",
        "not_recorded"
      )
    )
  )
# Number bred before neutering
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    num_breeding_before_neutering_terciles = factor(
      num_breeding_before_neutering_terciles,
      levels = c(
        ">0-<3",
        ">=3-<9",
        ">=9-142",
        "not_bred",
        "not_recorded"
      )
    )
  )


# NSAID exposure
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    NSAID_5y_quantiles = factor(
      NSAID_5y_quantiles,
      levels = c(
        "no_NSAID_admin",
        "Q1(1)",
        "Q2(1-2)",
        "Q3(2-3)",
        "Q4(3-11)",
        "no_med_use_recorded",
        "not_recorded"
      )
    )
  )
# Endoparasiticide exposure
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    endo_5y_quantiles = factor(
      endo_5y_quantiles,
      levels = c(
        "no_endo_admin",
        "Q1(1-2)",
        "Q2(2-3)",
        "Q3(3-4)",
        "Q4(4-17)",
        "no_med_use_recorded",
        "not_recorded"
      )
    )
  )


# Ectoparasiticide exposure
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    ecto_5y_quantiles = factor(
      ecto_5y_quantiles,
      levels = c(
        "no_ecto_admin",
        "Q1(1)",
        "Q2(1-2)",
        "Q3(2-4)",
        "Q4(4-17)",
        "no_med_use_recorded",
        "not_recorded"
      )
    )
  )


# Biological weight categories
cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  mutate(
    weight_biological = factor(
      weight_biological,
      levels = c(
        "≤25 kg",
        ">25–≤35 kg",
        ">35 kg",
        "not_recorded"
      )
    )
  )


# Weight quartiles
if ("weight_quartile" %in% names(cohort_abr_updated_weight2)) {

  current_weight_quartile_levels <- levels(
    factor(
      cohort_abr_updated_weight2$weight_quartile
    )
  )

  recorded_weight_quartile_levels <- setdiff(
    current_weight_quartile_levels,
    "not_recorded"
  )

  cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
    mutate(
      weight_quartile = factor(
        weight_quartile,
        levels = c(
          recorded_weight_quartile_levels,
          "not_recorded"
        )
      )
    )
}


#------------------------------------------------------------
# MOVE MEDICATION MISSING CATEGORY TO FINAL LEVEL
#------------------------------------------------------------

# Prefer named variables rather than fixed column positions because column
# positions may change when variables are added or removed.

medication_variables <- unique(
  unlist(
    list(
      grep(
        "^half_",
        names(cohort_abr_updated_weight2),
        value = TRUE
      ),
      grep(
        "^quartile_",
        names(cohort_abr_updated_weight2),
        value = TRUE
      ),
      c(
        "NSAID_5y_quantiles",
        "endo_5y_quantiles",
        "ecto_5y_quantiles"
      )
    )
  )
)


medication_variables <- intersect(
  medication_variables,
  names(cohort_abr_updated_weight2)
)


for (variable in medication_variables) {

  if (is.factor(cohort_abr_updated_weight2[[variable]])) {

    current_levels <- levels(
      cohort_abr_updated_weight2[[variable]]
    )

    if (
      "no_medication_records_for_this_time_period" %in%
        current_levels
    ) {

      cohort_abr_updated_weight2[[variable]] <- factor(
        cohort_abr_updated_weight2[[variable]],
        levels = c(
          setdiff(
            current_levels,
            "no_medication_records_for_this_time_period"
          ),
          "no_medication_records_for_this_time_period"
        )
      )
    }
  }
}


#------------------------------------------------------------
# FINAL VALIDATION OF CONTINUOUS VARIABLES
#------------------------------------------------------------

# Check again after all factor-level changes to ensure no continuous
# variables have accidentally been converted back to factors.

final_continuous_check <- tibble(
  variable = continuous_variables_present,
  variable_class = vapply(
    cohort_abr_updated_weight2[
      continuous_variables_present
    ],
    function(x) {
      paste(
        class(x),
        collapse = "/"
      )
    },
    character(1)
  ),
  number_unique = vapply(
    cohort_abr_updated_weight2[
      continuous_variables_present
    ],
    function(x) {
      dplyr::n_distinct(
        x,
        na.rm = TRUE
      )
    },
    integer(1)
  ),
  number_missing = vapply(
    cohort_abr_updated_weight2[
      continuous_variables_present
    ],
    function(x) {
      sum(is.na(x))
    },
    integer(1)
  )
)

final_continuous_check


if (
  any(
    final_continuous_check$variable_class != "numeric"
  )
) {

  stop(
    paste(
      "Continuous-variable conversion failed for:",
      paste(
        final_continuous_check$variable[
          final_continuous_check$variable_class !=
            "numeric"
        ],
        collapse = ", "
      )
    )
  )
}


# Specific check for the activity variable that previously caused problems
class(
  cohort_abr_updated_weight2$avg_activity_freq_sy3
)

is.numeric(
  cohort_abr_updated_weight2$avg_activity_freq_sy3
)

is.factor(
  cohort_abr_updated_weight2$avg_activity_freq_sy3
)


# Expected output:
#
# [1] "numeric"
# [1] TRUE
# [1] FALSE


#------------------------------------------------------------
# OPTIONAL DATAFRAME WITH "not_recorded" REPLACED BY NA
#------------------------------------------------------------

# Use this version if you want not_recorded categories excluded
# automatically by complete-case regression models.

cohort_no_not_rec <- cohort_abr_updated_weight2 %>%
 mutate(
   across(
    where(
       ~ is.character(.x) || is.factor(.x)
      ),
     ~ factor(
        na_if(
         as.character(.x),
          "not_recorded"
      )
     )
    )
  )


# Recreate BCS classification in the no-not-recorded dataframe
cohort_no_not_rec <- cohort_no_not_rec %>%
  mutate(
    mode_BCS_classify = case_when(
      as.character(
        mode_purina_BCS_5y_prior_endpoint
      ) %in% c(
        "1", "2", "3"
      ) ~ "underweight",

      as.character(
        mode_purina_BCS_5y_prior_endpoint
      ) %in% c(
        "4", "5"
      ) ~ "ideal",

      as.character(
        mode_purina_BCS_5y_prior_endpoint
      ) %in% c(
        "6", "7", "8", "9"
      ) ~ "overweight",

      TRUE ~ NA_character_
    ),

    mode_BCS_classify = factor(
      mode_BCS_classify,
      levels = c(
        "ideal",
        "underweight",
        "overweight"
      )
    )
  )


#------------------------------------------------------------
# WRITE COLUMN NAMES
#------------------------------------------------------------

colnames_df <- tibble(
  variable = names(cohort_no_not_rec)
)



colnames_df <- colnames(cohort_no_not_rec)
write.csv(colnames_df,"C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_col_names.csv")
```

Harmonise the missing data recording name

```{r}
medication_missing_labels <- c(
  "no_med_use_recorded",
  "drug_not_administered"
)

new_medication_missing_label <-
  "No medication records for this time period"

cohort_abr_updated_weight2 <- cohort_abr_updated_weight2 %>%
  dplyr::mutate(
    dplyr::across(
      where(~ is.character(.x) || is.factor(.x)),
      ~ {
        was_factor <- is.factor(.x)
        was_ordered <- is.ordered(.x)

        x_character <- as.character(.x)

        x_character[
          x_character %in% medication_missing_labels
        ] <- new_medication_missing_label

        if (was_factor) {
          updated_levels <- levels(.x)

          updated_levels[
            updated_levels %in% medication_missing_labels
          ] <- new_medication_missing_label

          factor(
            x_character,
            levels = unique(updated_levels),
            ordered = was_ordered
          )
        } else {
          x_character
        }
      }
    )
  )
```

Save dataset for work on confirmed HSA dogs only modelling

```{r}
write.csv(cohort_abr_updated_weight2,"C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_updated_analysis_df.csv")
```

## Regression analyses

**Fixed/forced covariates ie. in ALL models**

```{r}
exposome_data <- cohort_abr_updated_weight2 %>%
  mutate(
    had_HSA = as.integer(as.character(had_HSA))
  )

#forced covariates ie. the model will keep these in always and the exposures will be added in and out of different models
forced_covariates <- c(
  "age_at_final_date_bio",
  "sex",
  "neuter"
  #or possibly sex_status which is sex and neuter status combined
)
```

### **Exposome groups**

#### Built environment and socioeconomic

#### Household chemical exposure

#### Lifestyle/sleep location

#### Dog activity

#### Body condition/weight

#### Reproductive exposures

#### Medication types

##### NSAIDs(anti-inflams)

##### Antimicrobials

##### Dietary supplements

##### Enteric medications

##### Hormones

##### Immunologicals

##### Neurologicals

##### Anti-parasites

##### Anti-septics

##### Cardiorespiratory

##### Diuretics

##### Antihistamines

##### Antivirals

#### Comorbidities

##### Cardiovascular

##### Orthopaedic

##### Immune mediated

##### Chronic inflammatory

##### Parasitism

```{r}
#exposome families
exposure_families <- list(

  #------------------------------------------------------------
  # BUILT ENVIRONMENT AND SOCIOECONOMIC
  #------------------------------------------------------------

  built_environment_socioeconomic = c(
    "MDI_quintile",
    "owner_MDI_quintile",
    "area_type_mode_5yrs_prior",
    "house_type_mode_5yrs_prior",
    "water_source_mode_5yrs_prior",
    "pipes_metal_any_mode_5yrs_prior",
    "pipes_plastic_any_mode_5yrs_prior",
    "heating_fuel_primary_mode_5yrs_prior",
    "cooking_fuel_primary_mode_5yrs_prior",
    "heating_fuel_secondary_mode_5yrs_prior",
    "cooking_fuel_secondary_mode_5yrs_prior",
    "region_mode_5y_prior"
  ),

  #------------------------------------------------------------
  # HOUSEHOLD CHEMICAL EXPOSURE
  #------------------------------------------------------------

  household_chemical_exposure = c(
    "use_aerosol_study_years_early_life",
    "use_aerosol_study_years_rest_of_life",
    "use_aerosol_study_years_whole_life",
    "use_aerosol_exposed_5yrs_prior",

    "use_air_cleaner_study_years_early_life",
    "use_air_cleaner_study_years_rest_of_life",
    "use_air_cleaner_study_years_whole_life",
    "use_air_cleaner_exposed_5yrs_prior",

    "use_hepa_filter_study_years_early_life",
    "use_hepa_filter_study_years_rest_of_life",
    "use_hepa_filter_study_years_whole_life",
    "use_hepa_filter_exposed_5yrs_prior",

    "use_moth_balls_study_years_early_life",
    "use_moth_balls_study_years_rest_of_life",
    "use_moth_balls_study_years_whole_life",
    "use_moth_balls_exposed_5yrs_prior",

    "use_incense_or_candles_study_years_early_life",
    "use_incense_or_candles_study_years_rest_of_life",
    "use_incense_or_candles_study_years_whole_life",
    "use_incense_or_candles_exposed_5yrs_prior",

    "smoke_exposure_study_years_early_life",
    "smoke_exposure_study_years_rest_of_life",
    "smoke_exposure_study_years_whole_life",
    "smoke_exposure_exposed_5yrs_prior",

    "hours_of_smoke_early_life_total_dosage_terciles",
    "hours_of_smoke_rest_of_life_total_dosage_terciles",
    "hours_of_smoke_whole_life_total_dosage_terciles",
    "hours_of_smoke_exposure_hours_5yrs_prior_terciles",

    "any_treated_weeds_study_years_early_life",
    "any_treated_weeds_study_years_rest_of_life",
    "any_treated_weeds_study_years_whole_life",
    "any_treated_weeds_exposed_5yrs_prior",

    "any_treated_insects_study_years_early_life",
    "any_treated_insects_study_years_rest_of_life",
    "any_treated_insects_study_years_whole_life",
    "any_treated_insects_exposed_5yrs_prior",

    "any_treated_fertilizer_study_years_early_life",
    "any_treated_fertilizer_study_years_rest_of_life",
    "any_treated_fertilizer_study_years_whole_life",
    "any_treated_fertilizer_exposed_5yrs_prior"
  ),

  #------------------------------------------------------------
  # LIFESTYLE / SLEEP LOCATION
  #------------------------------------------------------------

  lifestyle_sleep_location = c(
    "majority_sleep_location_lifetime",
    "main_lifestyle_category",
    "main_lifestyle_category_mode_5yrs_prior",
    "sleep_location_mode_5yrs_prior",

    "in.the.garage_YN_study_years_early_life",
    "in.the.garage_YN_study_years_rest_of_life",
    "in.the.garage_YN_study_years_whole_life",

    "in.the.house_YN_study_years_early_life",
    "in.the.house_YN_study_years_rest_of_life",
    "in.the.house_YN_study_years_whole_life",

    "outside_YN_study_years_early_life",
    "outside_YN_study_years_rest_of_life",
    "outside_YN_study_years_whole_life"

   # "in.the.house_YN_sum_early_life_quartile",
    #"in.the.house_YN_sum_rest_of_life_quartile",
    #"in.the.house_YN_sum_whole_life_quartile"
  ),

  #------------------------------------------------------------
  # DOG ACTIVITY
  #------------------------------------------------------------

  dog_activity = c(
### removing all these avg,mean,median,min and max for intensity and frequency as they were a weird attempt to make an ordinal score for activity data up to 100 but do not make sense to interpret. keeping mode across these as allows to capture the free text of owner survey
  # Overall activity
 # "mean_activity_level_overview",
  "mode_activity_level_overview",

  # Time spent exercising
  "USDA_30min",
  "activity_1h_min",
  "activity_KC_2h_min",
#  "avg_activity_duration_sy1_2",
 # "avg_activity_duration_sy1_2_quartile",

  # Intensity
  #"avg_activity_intensity_sy1_2",
  #"avg_activity_intensity_sy3",
  #"avg_intensity_5y_prior_endpoint",
  #"avg_activity_intensity_sy3_quartile",
  #"avg_intensity_5y_prior_endpoint_quartile",
  "mode_intensity_5y_prior_endpoint",

  # Frequency
  #"avg_activity_freq_sy1_2",
  #"avg_activity_freq_sy3",
  #"avg_frequency_5y_prior_endpoint",
  #"avg_activity_freq_sy1_2_quartile",
  #"avg_activity_freq_sy3_quartile",
  #"avg_frequency_5y_prior_endpoint_quartile",
  "mode_frequency_5y_prior_endpoint",

  # Whole-life summaries
  "whole_activity_mode",
  #"whole_activity_mean",
  #"whole_activity_median",
  #"whole_activity_min",
  #"whole_activity_max",

  # Early-life summaries
  "early_activity_mode",
  #"early_activity_mean",
  #"early_activity_median",
  #"early_activity_min",
  #"early_activity_max",

  # Rest-of-life summaries
  "rest_activity_mode",
  #"rest_activity_mean",
  #"rest_activity_median",
  #"rest_activity_min",
  #"rest_activity_max",

  # Five-year-prior summaries
  "prior5y_activity_mode"
 # "prior5y_activity_mean",
  #"prior5y_activity_median",
  #"prior5y_activity_min",
  #"prior5y_activity_max"

  ),

  #------------------------------------------------------------
  # BODY CONDITION / WEIGHT
  #------------------------------------------------------------

  body_condition_weight = c(
    "weight_biological",
    "weight_quartile",
    "avg_height_5y_prior_endpoint_quartiles",
   # "avg_purina_BCS_5y_prior_endpoint_quartiles", <- doesnt really make sense to do avg when its a whole number value 
    "mode_BCS_classify",
   "mode_purina_BCS_5y_prior_endpoint"
  ),

  #------------------------------------------------------------
  # REPRODUCTIVE EXPOSURES
  #------------------------------------------------------------

  reproductive_exposures = c(
    "age_at_neuter_years_quartiles",
    "early_neuter_groups",
    "mismating_ever",
    "neutered_reason",
    "neutered_ever_bred",
    "spayed_while_in_heat",
    "spayed_while_pregnant",
    "prevent_heat_ever",
    "no_pregnancy_ever",
    "heats_before_neutering_terciles",
    "preg_pre_neuter",
    "mated_pre_neuter",
    "num_breeding_before_neutering_terciles",
    "ever_bred_from"
  ),

  #------------------------------------------------------------
  # INDIVIDUAL MEDICATION FAMILIES
  #------------------------------------------------------------

  nsaids_anti_inflammatories = c(
    "NSAID_ever",
    "NSAID_5y",
    "NSAID_5y_quantiles",
    "half_anti_inflam_lifetime",
    "quartile_anti_inflam_lifetime",
    "half_anti_inflam_5y_prior_endpoint",
    "quartile_anti_inflam_5y_prior_endpoint"
  ),

  antimicrobials = c(
    "half_anti_microbials_lifetime",
    "quartile_anti_microbials_lifetime",
    "half_anti_microbials_5y_prior_endpoint",
    "quartile_anti_microbials_5y_prior_endpoint"
  ),

  dietary_supplements = c(
    "half_dietary_supplements_lifetime",
    "quartile_dietary_supplements_lifetime",
    "half_dietary_supplements_5y_prior_endpoint",
    "quartile_dietary_supplements_5y_prior_endpoint"
  ),

  enteric_medications = c(
    "half_enteric_lifetime",
    "quartile_enteric_lifetime",
    "half_enteric_5y_prior_endpoint",
    "quartile_enteric_5y_prior_endpoint"
  ),

  hormones = c(
    "half_hormones_and_related_lifetime",
    "quartile_hormones_and_related_lifetime",
    "half_hormones_and_related_5y_prior_endpoint",
    "quartile_hormones_and_related_5y_prior_endpoint"
  ),
  steroids=c(
    "steroid_ever",
    "steroid_5y"
  ),

  immunologicals = c(
    "half_immunologicals_lifetime",
    "quartile_immunologicals_lifetime",
    "half_immunologicals_5y_prior_endpoint",
    "quartile_immunologicals_5y_prior_endpoint"
  ),

  neurologicals = c(
    "half_neurological_lifetime",
    "quartile_neurological_lifetime",
    "half_neurological_5y_prior_endpoint",
    "quartile_neurological_5y_prior_endpoint"
  ),

  anti_parasites = c(
    "half_anti_parasites_lifetime",
    "quartile_anti_parasites_lifetime",
    "half_anti_parasites_5y_prior_endpoint",
    "quartile_anti_parasites_5y_prior_endpoint",
    "ecto_5y_quantiles",
    "endo_5y_quantiles",
    "fenbendazole_ever",
    "fenbendazole_5y"
  ),

  anti_septics = c(
    "half_anti_septics_lifetime",
    "quartile_anti_septics_lifetime",
    "half_anti_septics_5y_prior_endpoint",
    "quartile_anti_septics_5y_prior_endpoint"
  ),

  cardiorespiratory_medications = c(
    "half_cardio_respiratory_lifetime",
    "quartile_cardio_respiratory_lifetime",
    "half_cardio_respiratory_5y_prior_endpoint",
    "quartile_cardio_respiratory_5y_prior_endpoint"
  ),

  diuretics = c(
    "half_diuretics_lifetime",
    "quartile_diuretics_lifetime",
    "half_diuretics_5y_prior_endpoint",
    "quartile_diuretics_5y_prior_endpoint"
  ),

  antihistamines = c(
    "half_anti_hist_lifetime",
    "quartile_anti_hist_lifetime",
    "half_anti_hist_5y_prior_endpoint",
    "quartile_anti_hist_5y_prior_endpoint"
  ),

  antivirals = c(
    "half_anti_viral_lifetime",
    "quartile_anti_viral_lifetime",
    "half_anti_viral_5y_prior_endpoint",
    "quartile_anti_viral_5y_prior_endpoint"
  ),

  #------------------------------------------------------------
  # COMORBIDITIES: OVERALL
  #------------------------------------------------------------

  comorbidities = c(
    "chronic_inflam_present_lifetime",
    "number_comorbs_present_lifetime",
    "chronic_inflam_present_5y",
    "number_comorbs_present_5y",
    "chronic_inflam_present_no_min_year",
    "number_comorbs_present_no_min_year",
    "lifelong_comorb_present.x",
    "number_lifelong_comorbs_present"
  ),

  #------------------------------------------------------------
  # INDIVIDUAL COMORBIDITY FAMILIES
  #------------------------------------------------------------

  cardiovascular_comorbidity = c(
    "cardiovascular",
    "lifelong_cardiovascular",
    "comorb_cardiovascular_all"
  ),

  orthopaedic_comorbidity = c(
    "orthopaedic",
    "lifelong_orthopaedic",
    "comorb_orthopaedic_all"
  ),

  immune_mediated_comorbidity = c(
    "immune_mediated",
    "lifelong_immune_mediated",
    "comorb_immune_med_all"
  ),

  chronic_inflammatory_comorbidity = c(
    "chronic_inflammatory",
    "lifelong_chronic_inflammatory",
    "lifelong_inflammatory_other",
    "comorb_chronic_inflam_all"
  ),

  #------------------------------------------------------------
  # PARASITISM
  #------------------------------------------------------------

  parasitism = c(
    "intestinal_parasitism",
    "tick_borne_parasitism",
    "other_parasitism",
    "all_parasitism"
  )
)
```

```{r}
exposure_families_requested <- exposure_families

missing_family_variables <- imap_dfr(
  exposure_families_requested,
  function(variables, family) {

    missing <- setdiff(
      variables,
      names(exposome_data)
    )

    tibble(
      exposure_family = family,
      missing_variable = missing
    )
  }
)

missing_family_variables
```

```{r}
analysis_df <- cohort_abr_updated_weight2
outcome_var <- "had_HSA"
requested_exposure_families <- exposure_families
missing_exposure_variables <- imap_dfr(
  requested_exposure_families,
  function(variables, family) {

    missing_variables <- setdiff(
      variables,
      names(analysis_df)
    )

    tibble(
      exposure_family = family,
      variable = missing_variables
    )
  }
)
exposure_families <- map(
  requested_exposure_families,
  ~ intersect(.x, names(analysis_df))
)
exposure_families <- exposure_families[
  lengths(exposure_families) > 0
]
tibble(
  exposure_family = names(exposure_families),
  number_variables = lengths(exposure_families)
)
```

```{r}
screen_exposure <- function(
    data,
    exposure,
    outcome = "had_HSA"
) {

  x <- data[[exposure]]
  y <- data[[outcome]]

  complete <- !is.na(x) & !is.na(y)

  x_complete <- x[complete]
  y_complete <- y[complete]

  tibble(
    exposure = exposure,
    variable_class = paste(class(x), collapse = "/"),
    n_total = nrow(data),
    n_observed = sum(!is.na(x)),
    n_missing = sum(is.na(x)),
    percent_missing = round(
      100 * mean(is.na(x)),
      1
    ),
    n_complete_outcome_exposure = sum(complete),
    cases_complete = sum(y_complete == 1),
    controls_complete = sum(y_complete == 0),
    n_unique_observed = n_distinct(x_complete),
    suitable_for_model = (
      sum(complete) > 0 &&
      n_distinct(x_complete) >= 2 &&
      n_distinct(y_complete) == 2
    )
  )
}
```

```{r}
exposure_screening <- imap_dfr(
  exposure_families,
  function(variables, family) {

    map_dfr(
      variables,
      ~ screen_exposure(
        data = analysis_df,
        exposure = .x,
        outcome = outcome_var
      )
    ) %>%
      mutate(
        exposure_family = family,
        .before = 1
      )
  }
)

exposure_screening %>%
  filter(!suitable_for_model)
```

```{r}
usable_exposure_families <- split(
  exposure_screening$exposure[
    exposure_screening$suitable_for_model
  ],
  exposure_screening$exposure_family[
    exposure_screening$suitable_for_model
  ]
)
```

```{r}
#ensure the numeric activity columns are not being turned back into factors again!
# Identify every column containing "mean" or "avg" in its name
mean_avg_variables <- names(analysis_df)[
  stringr::str_detect(
    names(analysis_df),
    stringr::regex("mean|avg", ignore_case = TRUE)
  ) &
  !stringr::str_detect(
    names(analysis_df),
    stringr::regex(
      "quart|quantile|tercile|categor|class",
      ignore_case = TRUE
    )
  )
]

# Review which columns have been selected
mean_avg_variables


# Convert selected columns from factor/character to numeric
#
# "not_recorded" and other non-numeric values become NA.
# factor -> character -> numeric prevents factor level numbers
# being used accidentally.

analysis_df <- analysis_df %>%
  dplyr::mutate(
    dplyr::across(
      dplyr::all_of(mean_avg_variables),
      ~ suppressWarnings(
        as.numeric(
          dplyr::na_if(
            as.character(.x),
            "not_recorded"
          )
        )
      )
    )
  )


#------------------------------------------------------------
# CHECK THAT CONVERSION WORKED
#------------------------------------------------------------

mean_avg_variable_check <- tibble::tibble(
  variable = mean_avg_variables,

  variable_class = vapply(
    analysis_df[mean_avg_variables],
    function(x) {
      paste(
        class(x),
        collapse = "/"
      )
    },
    character(1)
  ),

  number_observed = vapply(
    analysis_df[mean_avg_variables],
    function(x) {
      sum(!is.na(x))
    },
    integer(1)
  ),

  number_missing = vapply(
    analysis_df[mean_avg_variables],
    function(x) {
      sum(is.na(x))
    },
    integer(1)
  ),

  number_unique = vapply(
    analysis_df[mean_avg_variables],
    function(x) {
      dplyr::n_distinct(
        x,
        na.rm = TRUE
      )
    },
    integer(1)
  )
)

mean_avg_variable_check


# Specifically confirm the problem variable
class(
  analysis_df$avg_activity_freq_sy3
)

is.numeric(
  analysis_df$avg_activity_freq_sy3
)

is.factor(
  analysis_df$avg_activity_freq_sy3
)

```

\

```{r}
make_exposure_counts <- function(
    data,
    exposure,
    outcome = "had_HSA"
) {

  data %>%
    transmute(
      exposure_level = as.character(.data[[exposure]]),
      outcome_value = .data[[outcome]]
    ) %>%
    mutate(
      exposure_level = replace_na(
        exposure_level,
        "missing"
      ),
      outcome_value = as.character(outcome_value)
    ) %>%
    count(
      exposure_level,
      outcome_value,
      name = "n",
      .drop = FALSE
    ) %>%
    pivot_wider(
      names_from = outcome_value,
      values_from = n,
      values_fill = 0,
      names_prefix = "outcome_"
    ) %>%
    mutate(
      exposure = exposure,
      .before = 1
    )
}
```

```{r}
exposure_level_counts <- imap_dfr(
  usable_exposure_families,
  function(variables, family) {

    map_dfr(
      variables,
      ~ make_exposure_counts(
        data = analysis_df,
        exposure = .x,
        outcome = outcome_var
      )
    ) %>%
      mutate(
        exposure_family = family,
        .before = 1
      )
  }
)
```

```{r}
make_exposure_counts <- function(
    data,
    exposure,
    outcome = "had_HSA"
) {

  # Treat genuinely continuous variables as one exposure,
  # rather than producing one count row per numeric value
  if (is.numeric(data[[exposure]])) {

    data %>%
      dplyr::summarise(
        exposure = exposure,
        exposure_level = "continuous",
        outcome_0 = sum(
          .data[[outcome]] == 0 &
            !is.na(.data[[exposure]]),
          na.rm = TRUE
        ),
        outcome_1 = sum(
          .data[[outcome]] == 1 &
            !is.na(.data[[exposure]]),
          na.rm = TRUE
        ),
        missing = sum(
          is.na(.data[[exposure]])
        )
      )

  } else {

    # Categorical variables retain one row per exposure level
    data %>%
      dplyr::transmute(
        exposure_level = as.character(
          .data[[exposure]]
        ),
        outcome_value = .data[[outcome]]
      ) %>%
      dplyr::mutate(
        exposure_level = tidyr::replace_na(
          exposure_level,
          "missing"
        ),
        outcome_value = as.character(
          outcome_value
        )
      ) %>%
      dplyr::count(
        exposure_level,
        outcome_value,
        name = "n",
        .drop = FALSE
      ) %>%
      tidyr::pivot_wider(
        names_from = outcome_value,
        values_from = n,
        values_fill = 0,
        names_prefix = "outcome_"
      ) %>%
      dplyr::mutate(
        exposure = exposure,
        missing = NA_integer_,
        .before = 1
      )
  }
}


#------------------------------------------------------------
# CREATE EXPOSURE LEVEL COUNTS
#------------------------------------------------------------

exposure_level_counts <- purrr::imap_dfr(
  usable_exposure_families,
  function(variables, family) {

    purrr::map_dfr(
      variables,
      ~ make_exposure_counts(
        data = analysis_df,
        exposure = .x,
        outcome = outcome_var
      )
    ) %>%
      dplyr::mutate(
        exposure_family = family,
        .before = 1
      )
  }
)


exposure_level_counts
```

```{r}
exposure_level_counts

```

```{r}
write_csv(
  exposure_level_counts,
  "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/exposome_exposure_level_counts.csv"
)
```

### Univariable models

```{r}
fit_crude_exposure_model <- function(
    data,
    exposure,
    outcome = "had_HSA"
) {

  required_variables <- c(
    outcome,
    exposure
  )

  missing_variables <- setdiff(
    required_variables,
    names(data)
  )

  if (length(missing_variables) > 0) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = NA_character_,
        n = NA_integer_,
        cases = NA_integer_,
        controls = NA_integer_,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = NA_real_,
        overall_LRT_df = NA_real_,
        overall_LRT_p = NA_real_,
        AIC = NA_real_,
        converged = FALSE,
        warning = paste(
          "Missing variable(s):",
          paste(
            missing_variables,
            collapse = ", "
          )
        )
      )
    )
  }

  model_data <- data %>%
    dplyr::select(
      dplyr::all_of(
        required_variables
      )
    ) %>%
    tidyr::drop_na()

  n_model <- nrow(model_data)

  n_cases <- sum(
    model_data[[outcome]] == 1,
    na.rm = TRUE
  )

  n_controls <- sum(
    model_data[[outcome]] == 0,
    na.rm = TRUE
  )

  outcome_variation <- dplyr::n_distinct(
    model_data[[outcome]]
  )

  exposure_variation <- dplyr::n_distinct(
    model_data[[exposure]]
  )

  if (
    n_model == 0 ||
      outcome_variation < 2 ||
      exposure_variation < 2
  ) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = NA_character_,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = NA_real_,
        overall_LRT_df = NA_real_,
        overall_LRT_p = NA_real_,
        AIC = NA_real_,
        converged = FALSE,
        warning = "Insufficient variation in outcome or exposure"
      )
    )
  }

  full_formula <- stats::reformulate(
    termlabels = exposure,
    response = outcome
  )

  null_formula <- stats::as.formula(
    paste0(outcome, " ~ 1")
  )

  warning_messages <- character()

  full_model <- tryCatch(

    withCallingHandlers(

      stats::glm(
        formula = full_formula,
        data = model_data,
        family = stats::binomial()
      ),

      warning = function(w) {

        warning_messages <<- c(
          warning_messages,
          conditionMessage(w)
        )

        invokeRestart("muffleWarning")
      }
    ),

    error = function(e) {
      e
    }
  )

  if (inherits(full_model, "error")) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = NA_character_,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = NA_real_,
        overall_LRT_df = NA_real_,
        overall_LRT_p = NA_real_,
        AIC = NA_real_,
        converged = FALSE,
        warning = conditionMessage(full_model)
      )
    )
  }

  null_model <- tryCatch(

    stats::glm(
      formula = null_formula,
      data = model_data,
      family = stats::binomial()
    ),

    error = function(e) {
      e
    }
  )

  if (inherits(null_model, "error")) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = NA_character_,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = NA_real_,
        overall_LRT_df = NA_real_,
        overall_LRT_p = NA_real_,
        AIC = stats::AIC(full_model),
        converged = full_model$converged,
        warning = conditionMessage(null_model)
      )
    )
  }

  LRT_table <- tryCatch(

    stats::anova(
      null_model,
      full_model,
      test = "LRT"
    ),

    error = function(e) {
      NULL
    }
  )

  if (is.null(LRT_table)) {

    overall_LRT_chisq <- NA_real_
    overall_LRT_df <- NA_real_
    overall_LRT_p <- NA_real_

  } else {

    overall_LRT_chisq <- LRT_table$Deviance[2]
    overall_LRT_df <- LRT_table$Df[2]
    overall_LRT_p <- LRT_table$`Pr(>Chi)`[2]
  }

  exposure_object <- model_data[[exposure]]

  if (
    is.factor(exposure_object) ||
      is.character(exposure_object)
  ) {

    factor_exposure <- factor(exposure_object)

    reference_level <- levels(
      factor_exposure
    )[1]

  } else {

    reference_level <- "Per one-unit increase"
  }

  tidy_results <- tryCatch(

    broom::tidy(
      full_model,
      conf.int = TRUE,
      exponentiate = FALSE
    ),

    error = function(e) {
      NULL
    }
  )

  if (is.null(tidy_results)) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = reference_level,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = overall_LRT_chisq,
        overall_LRT_df = overall_LRT_df,
        overall_LRT_p = overall_LRT_p,
        AIC = stats::AIC(full_model),
        converged = full_model$converged,
        warning = "Could not tidy model coefficients"
      )
    )
  }

  tidy_results <- tidy_results %>%
    dplyr::filter(
      term != "(Intercept)"
    )

  if (nrow(tidy_results) == 0) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = reference_level,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = overall_LRT_chisq,
        overall_LRT_df = overall_LRT_df,
        overall_LRT_p = overall_LRT_p,
        AIC = stats::AIC(full_model),
        converged = full_model$converged,
        warning = "No exposure coefficient returned"
      )
    )
  }

  final_warning <- if (
    length(warning_messages) == 0
  ) {

    NA_character_

  } else {

    paste(
      unique(warning_messages),
      collapse = "; "
    )
  }

  tidy_results %>%
    dplyr::transmute(
      exposure = exposure,
      term = term,
      reference_level = reference_level,
      n = n_model,
      cases = n_cases,
      controls = n_controls,
      OR = exp(estimate),
      CI_lower = exp(conf.low),
      CI_upper = exp(conf.high),
      wald_z = statistic,
      wald_p = p.value,
      overall_LRT_chisq = overall_LRT_chisq,
      overall_LRT_df = overall_LRT_df,
      overall_LRT_p = overall_LRT_p,
      AIC = stats::AIC(full_model),
      converged = full_model$converged,
      warning = final_warning
    )
}


```

```{r}
crude_results <- imap_dfr(
  usable_exposure_families,
  function(variables, family) {

    map_dfr(
      variables,
      ~ fit_crude_exposure_model(
        data = analysis_df,
        exposure = .x,
        outcome = outcome_var
      )
    ) %>%
      mutate(
        exposure_family = family,
        model_type = "Crude",
        .before = 1
      )
  }
)
```

```{r}
test_exposure <- usable_exposure_families[[1]][1]

test_exposure
```

```{r}
test_result <- fit_crude_exposure_model(
  data = analysis_df,
  exposure = test_exposure,
  outcome = outcome_var
)

test_result
```

```{r}
crude_results <- purrr::imap_dfr(
  usable_exposure_families,
  function(variables, family) {

    family_results <- purrr::map_dfr(
      variables,
      function(variable_name) {

        fit_crude_exposure_model(
          data = analysis_df,
          exposure = variable_name,
          outcome = outcome_var
        )
      }
    )

    family_results %>%
      dplyr::mutate(
        exposure_family = family,
        model_type = "Crude",
        .before = 1
      )
  }
)
```

```{r}
crude_results %>%
  dplyr::distinct(
    exposure_family,
    exposure
  ) %>%
  nrow()
```

Checking model problems

```{r}
crude_model_problems <- crude_results %>%
  dplyr::filter(
    !converged |
      !is.na(warning)
  ) %>%
  dplyr::distinct(
    exposure_family,
    exposure,
    n,
    cases,
    controls,
    converged,
    warning
  ) %>%
  dplyr::arrange(
    exposure_family,
    exposure
  )
```

One row per exposure variable table

```{r}
crude_overall_tests <- crude_results %>%
  dplyr::distinct(
    exposure_family,
    exposure,
    n,
    cases,
    controls,
    overall_LRT_chisq,
    overall_LRT_df,
    overall_LRT_p,
    AIC,
    converged,
    warning
  )
crude_overall_tests <- crude_overall_tests %>%
  dplyr::group_by(
    exposure_family
  ) %>%
  dplyr::mutate(
    LRT_p_FDR_within_family = stats::p.adjust(
      overall_LRT_p,
      method = "BH"
    )
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    LRT_p_FDR_global = stats::p.adjust(
      overall_LRT_p,
      method = "BH"
    )
  ) %>%
  dplyr::arrange(
    exposure_family,
    overall_LRT_p
  )
```

```{r}
crude_results <- crude_results %>%
  dplyr::left_join(
    crude_overall_tests %>%
      dplyr::select(
        exposure_family,
        exposure,
        LRT_p_FDR_within_family,
        LRT_p_FDR_global
      ),
    by = c(
      "exposure_family",
      "exposure"
    )
  )
```

```{r}
format_p_value <- function(p) {

  dplyr::case_when(
    is.na(p) ~ NA_character_,
    p < 0.001 ~ "<0.001",
    TRUE ~ sprintf("%.3f", p)
  )
}
```

```{r}
crude_results_formatted <- crude_results %>%
  dplyr::mutate(

    OR_95CI = dplyr::case_when(
      is.na(OR) ~ NA_character_,
      TRUE ~ sprintf(
        "%.2f (%.2f–%.2f)",
        OR,
        CI_lower,
        CI_upper
      )
    ),

    Wald_p_display = format_p_value(
      wald_p
    ),

    LRT_p_display = format_p_value(
      overall_LRT_p
    ),

    family_FDR_display = format_p_value(
      LRT_p_FDR_within_family
    ),

    global_FDR_display = format_p_value(
      LRT_p_FDR_global
    )
  ) %>%
  dplyr::select(
    exposure_family,
    exposure,
    term,
    reference_level,
    n,
    cases,
    controls,
    OR_95CI,
    OR,
    CI_lower,
    CI_upper,
    wald_z,
    wald_p,
    Wald_p_display,
    overall_LRT_chisq,
    overall_LRT_df,
    overall_LRT_p,
    LRT_p_display,
    LRT_p_FDR_within_family,
    family_FDR_display,
    LRT_p_FDR_global,
    global_FDR_display,
    AIC,
    converged,
    warning
  ) %>%
  dplyr::arrange(
    exposure_family,
    overall_LRT_p,
    exposure,
    term
  )


#write to csv
write.csv(crude_results_formatted,"C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_log_reg_family_grouped.csv")


```

Table per family

Need to produce prettier tables

-   p\<0.001, values to 2 dp

```{r}
format_p <- function(x, digits = 3) {
  case_when(
    is.na(x) ~ NA_character_,
    x < 0.001 ~ "<0.001",
    TRUE ~ formatC(x, format = "f", digits = digits)
  )
}

crude_results_pretty <- crude_results %>%
  mutate(
    # OR and confidence limits to 3 decimal places
    OR_display = sprintf("%.3f", OR),
    CI_lower_display = sprintf("%.3f", CI_lower),
    CI_upper_display = sprintf("%.3f", CI_upper),

    # Combined OR (95% CI)
    OR_95CI = sprintf(
      "%.3f (%.3f–%.3f)",
      OR,
      CI_lower,
      CI_upper
    ),

    # Other numerical results
    wald_z_display = sprintf("%.3f", wald_z),
    overall_LRT_chisq_display = sprintf("%.3f", overall_LRT_chisq),
    AIC_display = sprintf("%.3f", AIC),

    # P-values
    Wald_p_display = format_p(wald_p),
    LRT_p_display = format_p(overall_LRT_p),
    family_FDR_display = format_p(LRT_p_FDR_within_family),
    global_FDR_display = format_p(LRT_p_FDR_global)
  )


crude_results_pretty 
```

Format further differences:

-   reference level in table not as column and also family grouping

```{r}
library(dplyr)
library(stringr)
library(purrr)
library(tibble)


#===============================================================================
# Formatting functions
#===============================================================================

format_p <- function(x, digits = 3) {
  dplyr::case_when(
    is.na(x)  ~ "",
    x < 0.001 ~ "<0.001",
    TRUE ~ formatC(
      x,
      format = "f",
      digits = digits
    )
  )
}


format_count_percent <- function(count, denominator, digits = 1) {
  dplyr::case_when(
    is.na(count) |
      is.na(denominator) |
      denominator == 0 ~ "",

    TRUE ~ sprintf(
      paste0("%d (%.", digits, "f%%)"),
      as.integer(count),
      100 * count / denominator
    )
  )
}


# Apply identical cleaning to:
# 1. coefficient categories
# 2. count-table categories
# 3. reference categories

pretty_category <- function(x) {
  x <- as.character(x)

  dplyr::case_when(
    is.na(x) ~ NA_character_,

    x == "" |
      x == "continuous" ~
      "Continuous",

    x == "no_med_use_recorded" ~
      "No medication use recorded",

    x == "no_medication_records_for_this_time_period" ~
      "No medication records for this time period",

    TRUE ~ x
  )
}



crude_model_display <- crude_results %>%
  dplyr::mutate(
    exposure = as.character(exposure),
    term = as.character(term),

    # For continuous variables, term usually equals exposure.
    # For factors, remove the exposure name from the start of term.
    Category_raw = dplyr::case_when(
      term == exposure ~ "continuous",

      stringr::str_starts(term, exposure) ~
        stringr::str_sub(
          term,
          start = stringr::str_length(exposure) + 1L
        ),

      TRUE ~ term
    ),

    Category = pretty_category(Category_raw),

    `Crude OR (95% CI)` = dplyr::case_when(
      is.na(OR) |
        is.na(CI_lower) |
        is.na(CI_upper) ~ "",

      TRUE ~ sprintf(
        "%.3f (%.3f–%.3f)",
        OR,
        CI_lower,
        CI_upper
      )
    ),

    `Wald p-value` = format_p(wald_p)
  ) %>%
  dplyr::select(
    exposure_family,
    exposure,
    Category,
    `Crude OR (95% CI)`,
    `Wald p-value`
  )



reference_lookup <- crude_results %>%
  dplyr::filter(
    !is.na(reference_level),
    reference_level != ""
  ) %>%
  dplyr::transmute(
    exposure_family,
    exposure,
    ReferenceCategory = pretty_category(reference_level)
  ) %>%
  dplyr::distinct()




reference_lookup %>%
  dplyr::count(
    exposure_family,
    exposure
  ) %>%
  dplyr::filter(n > 1)


crude_overall_p <- crude_results %>%
  dplyr::group_by(
    exposure_family,
    exposure
  ) %>%
  dplyr::summarise(
    `LRT p-value` = format_p(
      dplyr::first(
        overall_LRT_p[!is.na(overall_LRT_p)],
        default = NA_real_
      )
    ),

    `Family FDR p-value` = format_p(
      dplyr::first(
        LRT_p_FDR_within_family[
          !is.na(LRT_p_FDR_within_family)
        ],
        default = NA_real_
      )
    ),

    `Global FDR p-value` = format_p(
      dplyr::first(
        LRT_p_FDR_global[
          !is.na(LRT_p_FDR_global)
        ],
        default = NA_real_
      )
    ),

    .groups = "drop"
  )


crude_counts_display <- exposure_level_counts %>%
  dplyr::group_by(
    exposure_family,
    exposure
  ) %>%
  dplyr::mutate(
    category_original_order = dplyr::row_number(),

    total_cases = sum(
      outcome_1,
      na.rm = TRUE
    ),

    total_non_cases = sum(
      outcome_0,
      na.rm = TRUE
    ),

    `Cases (%)` = format_count_percent(
      outcome_1,
      total_cases
    ),

    `Non-cases (%)` = format_count_percent(
      outcome_0,
      total_non_cases
    ),

    Category = pretty_category(exposure_level)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::select(
    exposure_family,
    exposure,
    Category,
    category_original_order,
    `Cases (%)`,
    `Non-cases (%)`
  )


family_order_lookup <- tibble::tibble(
  exposure_family = names(exposure_families),
  family_order = seq_along(exposure_families)
)


exposure_order_lookup <- purrr::imap_dfr(
  exposure_families,
  function(variables, family_name) {
    tibble::tibble(
      exposure_family = family_name,
      exposure = variables,
      exposure_order = seq_along(variables)
    )
  }
) %>%
  dplyr::distinct(
    exposure_family,
    exposure,
    .keep_all = TRUE
  )


crude_results_table <- crude_counts_display %>%

  # Add the reference category to every row for each exposure
  dplyr::left_join(
    reference_lookup,
    by = c(
      "exposure_family",
      "exposure"
    )
  ) %>%

  # Add OR and Wald p-value to the relevant non-reference rows
  dplyr::left_join(
    crude_model_display,
    by = c(
      "exposure_family",
      "exposure",
      "Category"
    )
  ) %>%

  # Add the overall LRT and FDR p-values
  dplyr::left_join(
    crude_overall_p,
    by = c(
      "exposure_family",
      "exposure"
    )
  ) %>%

  # Add the desired family and exposure order
  dplyr::left_join(
    family_order_lookup,
    by = "exposure_family"
  ) %>%

  dplyr::left_join(
    exposure_order_lookup,
    by = c(
      "exposure_family",
      "exposure"
    )
  ) %>%

  dplyr::mutate(
    is_reference =
      !is.na(ReferenceCategory) &
      Category == ReferenceCategory,

    # Put the reference category first
    reference_order = dplyr::if_else(
      is_reference,
      0L,
      1L
    ),

    # Add the reference OR manually
    `Crude OR (95% CI)` = dplyr::case_when(
      is_reference ~ "1.000 (reference)",
      TRUE ~ dplyr::coalesce(
        `Crude OR (95% CI)`,
        ""
      )
    ),

    # Reference categories have no coefficient-level Wald p-value
    `Wald p-value` = dplyr::case_when(
      is_reference ~ "",
      TRUE ~ dplyr::coalesce(
        `Wald p-value`,
        ""
      )
    ),

    # Put unrecognised families/exposures at the end
    family_order = dplyr::coalesce(
      family_order,
      length(exposure_families) + 1L
    ),

    exposure_order = dplyr::coalesce(
      exposure_order,
      999999L
    )
  ) %>%

  dplyr::arrange(
    family_order,
    exposure_order,
    reference_order,
    category_original_order
  )



crude_results_table <- crude_results_table %>%
  dplyr::group_by(
    exposure_family,
    exposure
  ) %>%
  dplyr::mutate(
    `LRT p-value` = dplyr::if_else(
      dplyr::row_number() == 1L,
      `LRT p-value`,
      ""
    ),

    `Family FDR p-value` = dplyr::if_else(
      dplyr::row_number() == 1L,
      `Family FDR p-value`,
      ""
    ),

    `Global FDR p-value` = dplyr::if_else(
      dplyr::row_number() == 1L,
      `Global FDR p-value`,
      ""
    )
  ) %>%
  dplyr::ungroup()




#tidier ready for manuscript table
crude_results_table2 <- crude_results_table %>%
  dplyr::transmute(
    Group = exposure_family,
    Variable = exposure,
    Category,
    `Cases (%)`,
    `Non-cases (%)`,
    `Crude OR (95% CI)`,
    `Wald p-value`,
    `LRT p-value`,
    `Family FDR p-value`,
    `Global FDR p-value`
  )



crude_results_table3 <- crude_results_table2 %>%
  dplyr::mutate(
    Group_original = Group,
    Variable_original = Variable
  ) %>%

  # Display each variable name only once
  dplyr::group_by(
    Group_original,
    Variable_original
  ) %>%
  dplyr::mutate(
    Variable = dplyr::if_else(
      dplyr::row_number() == 1L,
      Variable,
      ""
    )
  ) %>%
  dplyr::ungroup() %>%

  # Display each exposure family only once
  dplyr::group_by(Group_original) %>%
  dplyr::mutate(
    Group = dplyr::if_else(
      dplyr::row_number() == 1L,
      Group,
      ""
    )
  ) %>%
  dplyr::ungroup() %>%

  dplyr::select(
    -Group_original,
    -Variable_original
  )
```

Need to ensure correct reference level (need to refactorise at start before this analysis)

```{r}
output_folder <- "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output"

ft <- flextable(crude_results_table3) %>%
  theme_booktabs() %>%
  bold(part = "header") %>%
  fontsize(size = 8, part = "all") %>%
  align(align = "left", part = "all") %>%
  valign(valign = "top", part = "all") %>%
  autofit() %>%
  set_table_properties(
    layout = "autofit",
    width = 1
  )

doc <- read_docx() %>%
  body_add_par(
    "Crude logistic regression results",
    style = "heading 1"
  ) %>%
  body_add_flextable(ft) %>%
  body_end_section_landscape()



writexl::write_xlsx(
  list("Crude results" = crude_results_table3),
  file.path(output_folder, "crude_results_table.xlsx")
)

readr::write_excel_csv(
  crude_results_table3,
  file.path(output_folder, "crude_results_table.csv"),
  na = ""
)

print(
  doc,
  target = file.path(output_folder, "crude_results_table.docx")
)
```

#### Univariable output interpretation:

Rules:

-   if significant not just for the "not_recorded" value

-   p\<0.2 for LRT and family FDR \<0.1 to consider and then consider against variables within exposure family if multiple significant

-   Looking for consistent association across a variable e.g. in quartiles is it increasing as dose incrases etc.

Built environment/socioeconomic

-   primary heating fuel source 5y prior

-   region mode 5y prior

-   aerosol use rest of life

-   incense/candles rest of life

-   smoke early life

-   weed treatment rest of life and whole life and 5y prior

-   insect treatment rest and whole life and 5y prior

-   fertiliser treatment rest and whole and 5y prior

Lifestyle sleep location

-   main lifestyle category ever and 5y prior

-   outside sleep location early life and whole life

Dog activity

-   mode activity level overview (whole life dataset) - NB mean is also sig but not normall distriution etc and its the numericising of a categorical information thing so ignore

-   mode intensity 5y prior endpoint

-   mode frequency 5y prior

-   whole life and rest of life mode intensity

BCS/weight

-   none are significant if we exclude the not_recorded variable

Repro exposures

-   age at neuter quartiles

-   early neuter groups

-   rest of repro variables, the only sig ones are due to putting "Entire" as a reason in a neuter variable or similar

NSAID

-   ever administered and 5y prior and 5y prior quantiles and lifetime quartiles

Antimicrobials

-   none

Dietary

-   half dietary supplements lifetime and 5y prior endpoint

Hormones

-   none

Steroid

-   none, only due to missing category

Immunologicals

-   quartiles lifetime and 5y prior

Neurologicals

-   none

Anti-parasites

-   half and quarter antiparasites lifetime and 5y prior

-   ectoparasite 5y prior quantile and endo 5y prior quantile

-   fenbendazole ever and 5y prior

Antiseptics

-   none

Cardiorespiratory

-   none

Diuretics

-   none

Anti-histamines

-   none

Anti-virals

-   none

Comorbidities

-   chronic inflam comor present

Creating several further univariable models with only complete cases retained to assess if significance only driven by missing data for the following:

(also running the age at neuter etc to exclude entire dogs)

```{r}
candidate_exposure_families <- list(

  #---------------------------------------------------------------------------
  # BUILT ENVIRONMENT / SOCIOECONOMIC
  #---------------------------------------------------------------------------

  built_environment_socioeconomic = c(
    "heating_fuel_primary_mode_5yrs_prior",
    "region_mode_5y_prior"
  ),

  #---------------------------------------------------------------------------
  # HOUSEHOLD CHEMICAL EXPOSURE
  #---------------------------------------------------------------------------

  household_chemical_exposure = c(
    "use_aerosol_study_years_rest_of_life",
    "use_incense_or_candles_study_years_rest_of_life",
    "smoke_exposure_study_years_early_life",

    "any_treated_weeds_study_years_rest_of_life",
    "any_treated_weeds_study_years_whole_life",
    "any_treated_weeds_exposed_5yrs_prior",

    "any_treated_insects_study_years_rest_of_life",
    "any_treated_insects_study_years_whole_life",
    "any_treated_insects_exposed_5yrs_prior",

    "any_treated_fertilizer_study_years_rest_of_life",
    "any_treated_fertilizer_study_years_whole_life",
    "any_treated_fertilizer_exposed_5yrs_prior"

    # Optional liberal candidate:
    # "use_hepa_filter_study_years_rest_of_life"
  ),

  #---------------------------------------------------------------------------
  # LIFESTYLE / SLEEP LOCATION
  #---------------------------------------------------------------------------

  lifestyle_sleep_location = c(
    "main_lifestyle_category",
    "main_lifestyle_category_mode_5yrs_prior",
    "outside_YN_study_years_early_life",
    "outside_YN_study_years_whole_life"
  ),

  #---------------------------------------------------------------------------
  # DOG ACTIVITY
  #---------------------------------------------------------------------------

  dog_activity = c(
    "mode_activity_level_overview",
    "mode_intensity_5y_prior_endpoint",
    "mode_frequency_5y_prior_endpoint",
    "whole_activity_mode",
    "rest_activity_mode"
  ),

  #---------------------------------------------------------------------------
  # REPRODUCTIVE EXPOSURES
  #---------------------------------------------------------------------------

  reproductive_exposures = c(
    "age_at_neuter_years_quartiles",
    "early_neuter_groups"
  ),

  #---------------------------------------------------------------------------
  # NSAIDS / ANTI-INFLAMMATORIES
  #---------------------------------------------------------------------------

  nsaids_anti_inflammatories = c(
    "NSAID_ever",
    "NSAID_5y",
    "NSAID_5y_quantiles",
    "quartile_anti_inflam_lifetime"
  ),

  #---------------------------------------------------------------------------
  # DIETARY SUPPLEMENTS
  #---------------------------------------------------------------------------

  dietary_supplements = c(
    "half_dietary_supplements_lifetime",
    "half_dietary_supplements_5y_prior_endpoint"
  ),

  #---------------------------------------------------------------------------
  # IMMUNOLOGICALS
  #---------------------------------------------------------------------------

  immunologicals = c(
    "quartile_immunologicals_lifetime",
    "quartile_immunologicals_5y_prior_endpoint"
  ),

  #---------------------------------------------------------------------------
  # ANTI-PARASITES
  #---------------------------------------------------------------------------

  anti_parasites = c(
    "half_anti_parasites_lifetime",
    "quartile_anti_parasites_lifetime",
    "half_anti_parasites_5y_prior_endpoint",
    "quartile_anti_parasites_5y_prior_endpoint",
    "ecto_5y_quantiles",
    "endo_5y_quantiles",
    "fenbendazole_ever",
    "fenbendazole_5y"
  ),

  #---------------------------------------------------------------------------
  # CHRONIC INFLAMMATORY COMORBIDITY
  #---------------------------------------------------------------------------

  chronic_inflammatory_comorbidity = c(
    "comorb_chronic_inflam_all"
  )
)
candidate_variables <- unique(
  unlist(
    candidate_exposure_families,
    use.names = FALSE
  )
)

missing_candidate_variables <- setdiff(
  candidate_variables,
  names(analysis_df)
)

missing_candidate_variables
```

Complete cases by converting missing record sback to NA

```{r}
missing_category_labels <- c(
  "not_recorded",
  "not recorded",
  "no_med_use_recorded",
  "no_medication_records_for_this_time_period",
  "No medication use recorded",
  "No medication records for this time period",
  "no_int_recorded",
  "no_activity_records_for_this_time_period",
  "drug_not_administered",
 "no_activity_records_for_this_time_period"

)


remove_missing_categories <- function(
    x,
    extra_missing = character(0)
) {

  missing_values <- c(
    missing_category_labels,
    extra_missing
  )

  if (is.factor(x)) {

    original_levels <- levels(x)
    x_character <- as.character(x)

    x_character[
      x_character %in% missing_values
    ] <- NA_character_

    factor(
      x_character,
      levels = setdiff(
        original_levels,
        missing_values
      )
    )

  } else if (is.character(x)) {

    x[
      x %in% missing_values
    ] <- NA_character_

    factor(x)

  } else {

    x
  }
}
```

```{r}
analysis_cc <- analysis_df %>%
  mutate(
    had_HSA = as.integer(
      as.character(had_HSA)
    )
  ) %>%
  mutate(
    across(
      all_of(candidate_variables),
      remove_missing_categories
    )
  )

#ditching entire category in neuter age
analysis_cc <- analysis_cc %>%
  mutate(
    age_at_neuter_years_quartiles =
      remove_missing_categories(
        age_at_neuter_years_quartiles,
        extra_missing = "Entire"
      )
  )
```

Running complete case models for these variables:

```{r}
fit_crude_complete_case <- function(
    data,
    exposure,
    exposure_family,
    outcome = "had_HSA"
) {

  # Select only the outcome and this particular exposure
  model_data <- data[
    ,
    c(outcome, exposure),
    drop = FALSE
  ]

  # Exposure-specific complete cases
  model_data <- model_data[
    complete.cases(model_data),
    ,
    drop = FALSE
  ]

  # Remove factor levels left empty after excluding missing values
  model_data <- droplevels(model_data)
  
  # Use treatment contrasts so categorical exposures give one OR per level
if (is.factor(model_data[[exposure]])) {
  model_data[[exposure]] <- factor(
    as.character(model_data[[exposure]]),
    levels = levels(model_data[[exposure]]),
    ordered = FALSE
  )
}

  # Ensure character exposures are factors
  if (is.character(model_data[[exposure]])) {
    model_data[[exposure]] <- factor(
      model_data[[exposure]]
    )
  }

  n_model <- nrow(model_data)

  cases <- sum(
    model_data[[outcome]] == 1,
    na.rm = TRUE
  )

  controls <- sum(
    model_data[[outcome]] == 0,
    na.rm = TRUE
  )

  # Check that model is estimable
  if (
    n_model == 0 ||
    dplyr::n_distinct(model_data[[outcome]]) < 2 ||
    dplyr::n_distinct(model_data[[exposure]]) < 2
  ) {

    return(
      tibble::tibble(
        exposure_family = exposure_family,
        exposure = exposure,
        term = NA_character_,
        n = n_model,
        cases = cases,
        controls = controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = NA_real_,
        overall_LRT_df = NA_real_,
        overall_LRT_p = NA_real_,
        converged = FALSE,
        warning = "Insufficient outcome or exposure variation"
      )
    )
  }

  full_formula <- reformulate(
    exposure,
    response = outcome
  )

  full_model <- tryCatch(
    glm(
      full_formula,
      data = model_data,
      family = binomial()
    ),
    error = function(e) e
  )

  if (inherits(full_model, "error")) {

    return(
      tibble::tibble(
        exposure_family = exposure_family,
        exposure = exposure,
        term = NA_character_,
        n = n_model,
        cases = cases,
        controls = controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = NA_real_,
        overall_LRT_df = NA_real_,
        overall_LRT_p = NA_real_,
        converged = FALSE,
        warning = conditionMessage(full_model)
      )
    )
  }

  null_model <- glm(
    as.formula(
      paste(outcome, "~ 1")
    ),
    data = model_data,
    family = binomial()
  )

  lrt <- anova(
    null_model,
    full_model,
    test = "LRT"
  )

  coefficient_table <- summary(
    full_model
  )$coefficients

  # Remove the intercept
  coefficient_table <- coefficient_table[
    rownames(coefficient_table) != "(Intercept)",
    ,
    drop = FALSE
  ]

  estimates <- coefficient_table[, "Estimate"]
  standard_errors <- coefficient_table[, "Std. Error"]

  tibble::tibble(
    exposure_family = exposure_family,
    exposure = exposure,
    term = rownames(coefficient_table),

    n = n_model,
    cases = cases,
    controls = controls,

    OR = exp(estimates),

    CI_lower = exp(
      estimates - 1.96 * standard_errors
    ),

    CI_upper = exp(
      estimates + 1.96 * standard_errors
    ),

    wald_z = coefficient_table[, "z value"],
    wald_p = coefficient_table[, "Pr(>|z|)"],

    overall_LRT_chisq = lrt$Deviance[2],
    overall_LRT_df = lrt$Df[2],
    overall_LRT_p = lrt$`Pr(>Chi)`[2],

    converged = full_model$converged,

    warning = ifelse(
      full_model$converged,
      NA_character_,
      "Model did not converge"
    )
  )
}
candidate_lookup <- purrr::imap_dfr(
  candidate_exposure_families,
  function(variables, family_name) {

    tibble::tibble(
      exposure_family = family_name,
      exposure = variables
    )
  }
)


crude_cc_results <- purrr::map2_dfr(
  candidate_lookup$exposure,
  candidate_lookup$exposure_family,

  function(exposure_name, family_name) {

    fit_crude_complete_case(
      data = analysis_cc,
      exposure = exposure_name,
      exposure_family = family_name,
      outcome = "had_HSA"
    )
  }
)

crude_cc_overall_p <- crude_cc_results %>%
  distinct(
    exposure_family,
    exposure,
    overall_LRT_p
  ) %>%
  group_by(exposure_family) %>%
  mutate(
    LRT_p_FDR_within_family = p.adjust(
      overall_LRT_p,
      method = "BH"
    )
  ) %>%
  ungroup() %>%
  mutate(
    LRT_p_FDR_global = p.adjust(
      overall_LRT_p,
      method = "BH"
    )
  )


crude_cc_results <- crude_cc_results %>%
  select(
    -any_of(
      c(
        "LRT_p_FDR_within_family",
        "LRT_p_FDR_global"
      )
    )
  ) %>%
  left_join(
    crude_cc_overall_p,
    by = c(
      "exposure_family",
      "exposure",
      "overall_LRT_p"
    )
  )

format_p <- function(x, digits = 3) {
  dplyr::case_when(
    is.na(x) ~ "",
    x < 0.001 ~ "<0.001",
    TRUE ~ formatC(
      x,
      format = "f",
      digits = digits
    )
  )
}

format_count_percent <- function(count, denominator) {
  dplyr::case_when(
    is.na(count) |
      is.na(denominator) |
      denominator == 0 ~ "",

    TRUE ~ sprintf(
      "%d (%.1f%%)",
      as.integer(count),
      100 * count / denominator
    )
  )
}

first_non_missing <- function(x) {
  x <- x[!is.na(x)]

  if (length(x) == 0) {
    NA
  } else {
    x[[1]]
  }
}

make_cc_level_counts <- function(
    data,
    exposure,
    exposure_family,
    outcome = "had_HSA"
) {

  model_data <- data[
    ,
    c(outcome, exposure),
    drop = FALSE
  ]

  model_data <- model_data[
    complete.cases(model_data),
    ,
    drop = FALSE
  ]

  model_data <- droplevels(model_data)

  x <- model_data[[exposure]]
  y <- model_data[[outcome]]

  # Convert characters, logical variables and ordered factors
  # to ordinary unordered factors
  if (is.character(x) || is.logical(x)) {
    x <- factor(x)
  }

  if (is.ordered(x)) {
    x <- factor(
      as.character(x),
      levels = levels(x),
      ordered = FALSE
    )
  }

  total_cases <- sum(y == 1, na.rm = TRUE)
  total_non_cases <- sum(y == 0, na.rm = TRUE)

  # Categorical factor exposure
  if (is.factor(x)) {

    exposure_levels <- levels(x)

    tab <- table(
      factor(x, levels = exposure_levels),
      factor(y, levels = c(0, 1))
    )

    return(
      tibble::tibble(
        exposure_family = exposure_family,
        exposure = exposure,
        Category = exposure_levels,
        category_order = seq_along(exposure_levels),
        model_type = "factor",
        reference_level = exposure_levels[[1]],
        comparison_level = NA_character_,
        outcome_0 = as.integer(tab[, "0"]),
        outcome_1 = as.integer(tab[, "1"]),
        total_cases = total_cases,
        total_non_cases = total_non_cases
      )
    )
  }

  distinct_values <- sort(unique(x))

  # Numeric binary exposure such as 0/1
  if (
    is.numeric(x) &&
      length(distinct_values) == 2
  ) {

    x_factor <- factor(
      x,
      levels = distinct_values
    )

    tab <- table(
      x_factor,
      factor(y, levels = c(0, 1))
    )

    return(
      tibble::tibble(
        exposure_family = exposure_family,
        exposure = exposure,
        Category = as.character(distinct_values),
        category_order = seq_along(distinct_values),
        model_type = "binary_numeric",
        reference_level =
          as.character(distinct_values[[1]]),
        comparison_level =
          as.character(distinct_values[[2]]),
        outcome_0 = as.integer(tab[, "0"]),
        outcome_1 = as.integer(tab[, "1"]),
        total_cases = total_cases,
        total_non_cases = total_non_cases
      )
    )
  }

  # Continuous numeric exposure
  tibble::tibble(
    exposure_family = exposure_family,
    exposure = exposure,
    Category = "Continuous",
    category_order = 1L,
    model_type = "continuous",
    reference_level = NA_character_,
    comparison_level = NA_character_,
    outcome_0 = total_non_cases,
    outcome_1 = total_cases,
    total_cases = total_cases,
    total_non_cases = total_non_cases
  )
}


cc_level_counts <- purrr::map2_dfr(
  candidate_lookup$exposure,
  candidate_lookup$exposure_family,

  function(exposure_name, family_name) {
    make_cc_level_counts(
      data = analysis_cc,
      exposure = exposure_name,
      exposure_family = family_name,
      outcome = "had_HSA"
    )
  }
) %>%
  dplyr::mutate(
    `Cases (%)` = format_count_percent(
      outcome_1,
      total_cases
    ),

    `Non-cases (%)` = format_count_percent(
      outcome_0,
      total_non_cases
    )
  )


cc_exposure_metadata <- cc_level_counts %>%
  dplyr::distinct(
    exposure_family,
    exposure,
    model_type,
    reference_level,
    comparison_level
  )

crude_cc_model_display <- crude_cc_results %>%
  dplyr::left_join(
    cc_exposure_metadata,
    by = c(
      "exposure_family",
      "exposure"
    )
  ) %>%
  dplyr::mutate(
    Category = dplyr::case_when(
      model_type == "continuous" ~
        "Continuous",

      model_type == "binary_numeric" ~
        comparison_level,

      model_type == "factor" &
        !is.na(term) &
        startsWith(term, exposure) ~
        substring(
          term,
          nchar(exposure) + 1L
        ),

      TRUE ~ term
    ),

    `Crude OR (95% CI)` = dplyr::case_when(
      is.na(OR) |
        is.na(CI_lower) |
        is.na(CI_upper) ~ "",

      TRUE ~ sprintf(
        "%.3f (%.3f–%.3f)",
        OR,
        CI_lower,
        CI_upper
      )
    ),

    `Wald p-value` = format_p(wald_p)
  ) %>%
  dplyr::select(
    exposure_family,
    exposure,
    Category,
    `Crude OR (95% CI)`,
    `Wald p-value`
  )

crude_cc_overall_display <- crude_cc_results %>%
  dplyr::group_by(
    exposure_family,
    exposure
  ) %>%
  dplyr::summarise(
    `Model n` = first_non_missing(n),

    `LRT p-value` = format_p(
      first_non_missing(overall_LRT_p)
    ),

    `Family FDR p-value` = format_p(
      first_non_missing(
        LRT_p_FDR_within_family
      )
    ),

    `Global FDR p-value` = format_p(
      first_non_missing(
        LRT_p_FDR_global
      )
    ),

    .groups = "drop"
  )
```

```{r}
crude_cc_results_table <- cc_level_counts %>%
  dplyr::left_join(
    crude_cc_model_display,
    by = c(
      "exposure_family",
      "exposure",
      "Category"
    )
  ) %>%
  dplyr::left_join(
    crude_cc_overall_display,
    by = c(
      "exposure_family",
      "exposure"
    )
  ) %>%
  dplyr::mutate(
    is_reference =
      !is.na(reference_level) &
      Category == reference_level,

    `Crude OR (95% CI)` = dplyr::case_when(
      is_reference ~ "1.000 (reference)",

      TRUE ~ dplyr::coalesce(
        `Crude OR (95% CI)`,
        ""
      )
    ),

    `Wald p-value` = dplyr::case_when(
      is_reference ~ "",

      TRUE ~ dplyr::coalesce(
        `Wald p-value`,
        ""
      )
    )
  )


family_order_lookup <- tibble::tibble(
  exposure_family =
    names(candidate_exposure_families),

  family_order =
    seq_along(candidate_exposure_families)
)

exposure_order_lookup <- purrr::imap_dfr(
  candidate_exposure_families,

  function(variables, family_name) {
    tibble::tibble(
      exposure_family = family_name,
      exposure = variables,
      exposure_order = seq_along(variables)
    )
  }
)

crude_cc_results_table <- crude_cc_results_table %>%
  dplyr::left_join(
    family_order_lookup,
    by = "exposure_family"
  ) %>%
  dplyr::left_join(
    exposure_order_lookup,
    by = c(
      "exposure_family",
      "exposure"
    )
  ) %>%
  dplyr::arrange(
    family_order,
    exposure_order,
    category_order
  )


crude_cc_results_table <- crude_cc_results_table %>%
  dplyr::group_by(
    exposure_family,
    exposure
  ) %>%
  dplyr::mutate(
    `Model n` = dplyr::if_else(
      dplyr::row_number() == 1L,
      as.character(`Model n`),
      ""
    ),

    `LRT p-value` = dplyr::if_else(
      dplyr::row_number() == 1L,
      `LRT p-value`,
      ""
    ),

    `Family FDR p-value` = dplyr::if_else(
      dplyr::row_number() == 1L,
      `Family FDR p-value`,
      ""
    ),

    `Global FDR p-value` = dplyr::if_else(
      dplyr::row_number() == 1L,
      `Global FDR p-value`,
      ""
    )
  ) %>%
  dplyr::ungroup()


crude_cc_results_table <- crude_cc_results_table %>%
  dplyr::transmute(
    Group = exposure_family,
    Variable = exposure,
    Category,
    `Model n`,
    `Cases (%)`,
    `Non-cases (%)`,
    `Crude OR (95% CI)`,
    `Wald p-value`,
    `LRT p-value`,
    `Family FDR p-value`,
    `Global FDR p-value`
  )


crude_cc_results_table
```

Save output

```{r}
output_folder <- "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output"

readr::write_excel_csv(
  crude_cc_results_table,
  file.path(
    output_folder,
    "crude_complete_case_results_table.csv"
  ),
  na = ""
)



ft_cc <- flextable(
  crude_cc_results_table
) %>%
  theme_booktabs() %>%
  bold(part = "header") %>%
  fontsize(size = 7, part = "all") %>%
  valign(valign = "top", part = "all") %>%
  align(
    j = c("Group", "Variable", "Category"),
    align = "left",
    part = "all"
  ) %>%
  align(
    j = c(
      "Model n",
      "Cases (%)",
      "Non-cases (%)",
      "Crude OR (95% CI)",
      "Wald p-value",
      "LRT p-value",
      "Family FDR p-value",
      "Global FDR p-value"
    ),
    align = "center",
    part = "all"
  ) %>%
  autofit() %>%
  set_table_properties(
    layout = "autofit",
    width = 1,
    opts_word = list(
      split = TRUE
    )
  )

doc_cc <- read_docx() %>%
  body_add_par(
    "Complete-case crude logistic regression results",
    style = "heading 1"
  ) %>%
  body_add_flextable(ft_cc) %>%
  body_end_section_landscape()


print(
  doc_cc,
  target = file.path(
    output_folder,
    "crude_complete_case_results_table.docx"
  )
)
```

From this to take forward:

-   heating fuel primary mode 5y prior

-   aerosol, incense, smoke early life

-   weeds and insects and fertiliser rest and whole life

-   main lifestyle category mode

-   outside YN early or whole

-   mode activity level overview

-   mode activity intensity 5y prior

-   mode freq 5y prior

-   whole activity mode

-   antiparasites quartiles lifetime and 5y prior

-   comorb chronic inflam all

## Adjusted single exposure models

Running models with forced covariates in for the following variables:

```{r}
adjusted_candidate_exposure_families <- list(

  built_environment_socioeconomic = c(
    "heating_fuel_primary_mode_5yrs_prior"
  ),

  household_chemical_exposure = c(
    "use_aerosol_study_years_rest_of_life",
    "use_hepa_filter_study_years_rest_of_life",
    "use_incense_or_candles_study_years_rest_of_life",
    "smoke_exposure_study_years_early_life",

    "any_treated_weeds_study_years_rest_of_life",
    "any_treated_weeds_study_years_whole_life",
    "any_treated_weeds_exposed_5yrs_prior",

    "any_treated_insects_study_years_rest_of_life",
    "any_treated_insects_study_years_whole_life",
    "any_treated_insects_exposed_5yrs_prior",

    "any_treated_fertilizer_study_years_rest_of_life",
    "any_treated_fertilizer_study_years_whole_life",
    "any_treated_fertilizer_exposed_5yrs_prior"
  ),

  lifestyle_sleep_location = c(
    "main_lifestyle_category",
    "main_lifestyle_category_mode_5yrs_prior",
    "outside_YN_study_years_early_life",
    "outside_YN_study_years_whole_life"
  ),

  dog_activity = c(
    "mode_activity_level_overview",
    "mode_intensity_5y_prior_endpoint",
    "mode_frequency_5y_prior_endpoint",
    "whole_activity_mode",
    "rest_activity_mode"
  ),

  NSAIDs_anti_inflammatories = c(
    "NSAID_ever",
    "NSAID_5y",
    "NSAID_5y_quantiles",
    "quartile_anti_inflam_lifetime"
  ),

  dietary_supplements = c(
    "half_dietary_supplements_lifetime",
    "half_dietary_supplements_5y_prior_endpoint"
  ),

  immunologicals = c(
    "quartile_immunologicals_lifetime",
    "quartile_immunologicals_5y_prior_endpoint"
  ),

  anti_parasites = c(
    "half_anti_parasites_lifetime",
    "quartile_anti_parasites_lifetime",
    "half_anti_parasites_5y_prior_endpoint",
    "quartile_anti_parasites_5y_prior_endpoint",
    "ecto_5y_quantiles",
    "endo_5y_quantiles",
    "fenbendazole_ever",
    "fenbendazole_5y"
  ),

  chronic_inflammatory_comorbidity = c(
    "comorb_chronic_inflam_all"
  )
)
```

```{r}
forced_covariates <- c(
  "age_at_final_date_bio",
  "sex",
  "neuter"
)

adjusted_candidate_lookup <- purrr::imap_dfr(
  adjusted_candidate_exposure_families,
  function(variables, family_name) {
    tibble::tibble(
      exposure_family = family_name,
      exposure = variables
    )
  }
)

required_names <- c(
  "had_HSA",
  forced_covariates,
  adjusted_candidate_lookup$exposure
)

missing_names <- setdiff(
  unique(required_names),
  names(analysis_df)
)

missing_names
```

```{r}
extract_vif_diagnostics <- function(model, exposure) {

  if (!requireNamespace("car", quietly = TRUE)) {
    return(
      list(
        exposure_GVIF = NA_real_,
        exposure_GVIF_df = NA_real_,
        exposure_GVIF_adjusted = NA_real_,
        max_GVIF_adjusted = NA_real_
      )
    )
  }

  vif_object <- tryCatch(
    car::vif(model),
    error = function(e) NULL
  )

  if (is.null(vif_object)) {
    return(
      list(
        exposure_GVIF = NA_real_,
        exposure_GVIF_df = NA_real_,
        exposure_GVIF_adjusted = NA_real_,
        max_GVIF_adjusted = NA_real_
      )
    )
  }

  if (is.matrix(vif_object)) {

    terms <- rownames(vif_object)
    raw_gvif <- vif_object[, "GVIF"]
    gvif_df <- vif_object[, "Df"]

    adjusted_column <- grep(
      "1/(2",
      colnames(vif_object),
      fixed = TRUE
    )

    if (length(adjusted_column) > 0) {
      adjusted_gvif <- vif_object[, adjusted_column[[1]]]
    } else {
      adjusted_gvif <- raw_gvif^(1 / (2 * gvif_df))
    }

  } else {

    terms <- names(vif_object)
    raw_gvif <- as.numeric(vif_object)
    gvif_df <- rep(1, length(raw_gvif))

    # Comparable scale to GVIF^(1/(2*df))
    adjusted_gvif <- sqrt(raw_gvif)
  }

  exposure_position <- match(
    exposure,
    terms
  )

  maximum_adjusted <- if (
    length(adjusted_gvif) == 0 ||
      all(is.na(adjusted_gvif))
  ) {
    NA_real_
  } else {
    max(adjusted_gvif, na.rm = TRUE)
  }

  list(
    exposure_GVIF = if (is.na(exposure_position)) {
      NA_real_
    } else {
      unname(raw_gvif[exposure_position])
    },

    exposure_GVIF_df = if (is.na(exposure_position)) {
      NA_real_
    } else {
      unname(gvif_df[exposure_position])
    },

    exposure_GVIF_adjusted = if (is.na(exposure_position)) {
      NA_real_
    } else {
      unname(adjusted_gvif[exposure_position])
    },

    max_GVIF_adjusted = maximum_adjusted
  )
}

empty_adjusted_result <- function(
    exposure,
    exposure_family,
    n = 0L,
    cases = 0L,
    controls = 0L,
    warning_message = NA_character_,
    rank_deficient = NA
) {

  tibble::tibble(
    exposure_family = exposure_family,
    exposure = exposure,
    term = NA_character_,
    reference_level = NA_character_,

    n = n,
    cases = cases,
    controls = controls,

    adjusted_OR = NA_real_,
    adjusted_CI_lower = NA_real_,
    adjusted_CI_upper = NA_real_,
    adjusted_wald_z = NA_real_,
    adjusted_wald_p = NA_real_,

    adjusted_LRT_chisq = NA_real_,
    adjusted_LRT_df = NA_real_,
    adjusted_LRT_p = NA_real_,

    exposure_GVIF = NA_real_,
    exposure_GVIF_df = NA_real_,
    exposure_GVIF_adjusted = NA_real_,
    max_GVIF_adjusted = NA_real_,
    rank_deficient = rank_deficient,

    AIC = NA_real_,
    converged = FALSE,
    warning = warning_message
  )
}
```

```{r}
fit_adjusted_single_exposure <- function(
    data,
    exposure,
    exposure_family,
    outcome = "had_HSA",
    forced_covariates
) {

  required_variables <- unique(
    c(
      outcome,
      exposure,
      forced_covariates
    )
  )

  model_data <- data[
    ,
    required_variables,
    drop = FALSE
  ]

  # Removes only genuine NA values.
  # Character levels such as "not_recorded" are retained.
  model_data <- model_data[
    stats::complete.cases(model_data),
    ,
    drop = FALSE
  ]

  model_data <- droplevels(model_data)

  # Ensure outcome is numeric 0/1
  if (
    is.character(model_data[[outcome]]) ||
      is.factor(model_data[[outcome]])
  ) {

    model_data[[outcome]] <- suppressWarnings(
      as.integer(
        as.character(model_data[[outcome]])
      )
    )
  }

  if (is.logical(model_data[[outcome]])) {
    model_data[[outcome]] <- as.integer(
      model_data[[outcome]]
    )
  }

  if (
    any(
      !model_data[[outcome]] %in% c(0, 1),
      na.rm = TRUE
    )
  ) {

    return(
      empty_adjusted_result(
        exposure = exposure,
        exposure_family = exposure_family,
        n = nrow(model_data),
        warning_message =
          "Outcome is not coded solely as 0 and 1"
      )
    )
  }

  predictor_variables <- c(
    forced_covariates,
    exposure
  )

  # Convert categorical predictors to ordinary unordered factors
  for (variable in predictor_variables) {

    if (is.ordered(model_data[[variable]])) {

      old_levels <- levels(
        model_data[[variable]]
      )

      model_data[[variable]] <- factor(
        as.character(model_data[[variable]]),
        levels = old_levels,
        ordered = FALSE
      )

    } else if (
      is.character(model_data[[variable]]) ||
        is.logical(model_data[[variable]])
    ) {

      model_data[[variable]] <- factor(
        model_data[[variable]]
      )
    }
  }

  n_model <- nrow(model_data)

  cases <- sum(
    model_data[[outcome]] == 1,
    na.rm = TRUE
  )

  controls <- sum(
    model_data[[outcome]] == 0,
    na.rm = TRUE
  )

  if (
    n_model == 0 ||
      dplyr::n_distinct(model_data[[outcome]]) < 2
  ) {

    return(
      empty_adjusted_result(
        exposure = exposure,
        exposure_family = exposure_family,
        n = n_model,
        cases = cases,
        controls = controls,
        warning_message =
          "Insufficient outcome variation"
      )
    )
  }

  invariant_predictors <- predictor_variables[
    vapply(
      predictor_variables,
      function(variable) {
        dplyr::n_distinct(
          model_data[[variable]]
        ) < 2
      },
      logical(1)
    )
  ]

  if (length(invariant_predictors) > 0) {

    return(
      empty_adjusted_result(
        exposure = exposure,
        exposure_family = exposure_family,
        n = n_model,
        cases = cases,
        controls = controls,
        warning_message = paste(
          "No variation in:",
          paste(
            invariant_predictors,
            collapse = ", "
          )
        )
      )
    )
  }

  reduced_formula <- reformulate(
    termlabels = forced_covariates,
    response = outcome
  )

  full_formula <- reformulate(
    termlabels = c(
      forced_covariates,
      exposure
    ),
    response = outcome
  )

  # Check exact redundancy before fitting
  design_matrix <- tryCatch(
    model.matrix(
      full_formula,
      data = model_data
    ),
    error = function(e) e
  )

  if (inherits(design_matrix, "error")) {

    return(
      empty_adjusted_result(
        exposure = exposure,
        exposure_family = exposure_family,
        n = n_model,
        cases = cases,
        controls = controls,
        warning_message =
          conditionMessage(design_matrix)
      )
    )
  }

  rank_deficient <- (
    qr(design_matrix)$rank <
      ncol(design_matrix)
  )

  if (rank_deficient) {

    return(
      empty_adjusted_result(
        exposure = exposure,
        exposure_family = exposure_family,
        n = n_model,
        cases = cases,
        controls = controls,
        warning_message =
          "Rank-deficient design matrix: exact predictor redundancy",
        rank_deficient = TRUE
      )
    )
  }

  captured_warnings <- character()

  fit_model <- function(formula) {

    tryCatch(
      withCallingHandlers(
        glm(
          formula,
          data = model_data,
          family = binomial()
        ),
        warning = function(w) {

          captured_warnings <<- c(
            captured_warnings,
            conditionMessage(w)
          )

          invokeRestart("muffleWarning")
        }
      ),
      error = function(e) e
    )
  }

  reduced_model <- fit_model(
    reduced_formula
  )

  full_model <- fit_model(
    full_formula
  )

  if (
    inherits(reduced_model, "error") ||
      inherits(full_model, "error")
  ) {

    error_messages <- c(
      if (inherits(reduced_model, "error")) {
        conditionMessage(reduced_model)
      },

      if (inherits(full_model, "error")) {
        conditionMessage(full_model)
      }
    )

    return(
      empty_adjusted_result(
        exposure = exposure,
        exposure_family = exposure_family,
        n = n_model,
        cases = cases,
        controls = controls,
        warning_message = paste(
          unique(error_messages),
          collapse = "; "
        ),
        rank_deficient = FALSE
      )
    )
  }

  # Overall exposure test after adjustment
  lrt <- anova(
    reduced_model,
    full_model,
    test = "LRT"
  )

  coefficient_table <- summary(
    full_model
  )$coefficients

  coefficient_names <- rownames(
    coefficient_table
  )

  exposure_rows <- (
    coefficient_names == exposure |
      startsWith(
        coefficient_names,
        exposure
      )
  )

  coefficient_table <- coefficient_table[
    exposure_rows,
    ,
    drop = FALSE
  ]

  if (nrow(coefficient_table) == 0) {

    return(
      empty_adjusted_result(
        exposure = exposure,
        exposure_family = exposure_family,
        n = n_model,
        cases = cases,
        controls = controls,
        warning_message =
          "Could not identify exposure coefficient rows",
        rank_deficient = FALSE
      )
    )
  }

  estimates <- coefficient_table[
    ,
    "Estimate"
  ]

  standard_errors <- coefficient_table[
    ,
    "Std. Error"
  ]

  reference_level <- if (
    is.factor(model_data[[exposure]])
  ) {
    levels(model_data[[exposure]])[[1]]
  } else {
    NA_character_
  }

  vif_diagnostics <- extract_vif_diagnostics(
    full_model,
    exposure
  )

  warning_text <- paste(
    unique(captured_warnings),
    collapse = "; "
  )

  if (warning_text == "") {
    warning_text <- NA_character_
  }

  if (!full_model$converged) {

    warning_text <- paste(
      na.omit(
        c(
          warning_text,
          "Full model did not converge"
        )
      ),
      collapse = "; "
    )
  }

  tibble::tibble(
    exposure_family = exposure_family,
    exposure = exposure,
    term = rownames(coefficient_table),
    reference_level = reference_level,

    n = n_model,
    cases = cases,
    controls = controls,

    adjusted_OR = exp(estimates),

    adjusted_CI_lower = exp(
      estimates - 1.96 * standard_errors
    ),

    adjusted_CI_upper = exp(
      estimates + 1.96 * standard_errors
    ),

    adjusted_wald_z =
      coefficient_table[, "z value"],

    adjusted_wald_p =
      coefficient_table[, "Pr(>|z|)"],

    adjusted_LRT_chisq =
      lrt$Deviance[2],

    adjusted_LRT_df =
      lrt$Df[2],

    adjusted_LRT_p =
      lrt$`Pr(>Chi)`[2],

    exposure_GVIF =
      vif_diagnostics$exposure_GVIF,

    exposure_GVIF_df =
      vif_diagnostics$exposure_GVIF_df,

    exposure_GVIF_adjusted =
      vif_diagnostics$exposure_GVIF_adjusted,

    max_GVIF_adjusted =
      vif_diagnostics$max_GVIF_adjusted,

    rank_deficient = FALSE,

    AIC = stats::AIC(
      full_model
    ),

    converged =
      full_model$converged,

    warning = warning_text
  )
}
```

Run model build functions

```{r}
adjusted_single_results <- purrr::map2_dfr(
  adjusted_candidate_lookup$exposure,
  adjusted_candidate_lookup$exposure_family,

  function(exposure_name, family_name) {

    fit_adjusted_single_exposure(
      data = analysis_df,
      exposure = exposure_name,
      exposure_family = family_name,
      outcome = "had_HSA",
      forced_covariates = forced_covariates
    )
  }
)
```

```{r}
adjusted_overall_p <- adjusted_single_results %>%
  dplyr::distinct(
    exposure_family,
    exposure,
    adjusted_LRT_p
  ) %>%
  dplyr::group_by(
    exposure_family
  ) %>%
  dplyr::mutate(
    adjusted_LRT_FDR_within_family =
      p.adjust(
        adjusted_LRT_p,
        method = "BH"
      )
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    adjusted_LRT_FDR_global =
      p.adjust(
        adjusted_LRT_p,
        method = "BH"
      )
  )

adjusted_single_results <- adjusted_single_results %>%
  dplyr::left_join(
    adjusted_overall_p,
    by = c(
      "exposure_family",
      "exposure",
      "adjusted_LRT_p"
    )
  )
```

```{r}
format_p <- function(x, digits = 3) {
  dplyr::case_when(
    is.na(x) ~ "",
    x < 0.001 ~ "<0.001",
    TRUE ~ formatC(
      x,
      format = "f",
      digits = digits
    )
  )
}

adjusted_exposure_summary <- adjusted_single_results %>%
  dplyr::distinct(
    exposure_family,
    exposure,
    n,
    cases,
    controls,
    adjusted_LRT_chisq,
    adjusted_LRT_df,
    adjusted_LRT_p,
    adjusted_LRT_FDR_within_family,
    adjusted_LRT_FDR_global,
    exposure_GVIF_adjusted,
    max_GVIF_adjusted,
    rank_deficient,
    AIC,
    converged,
    warning
  ) %>%
  dplyr::mutate(
    `Adjusted LRT p` =
      format_p(adjusted_LRT_p),

    `Family FDR p` =
      format_p(
        adjusted_LRT_FDR_within_family
      ),

    `Global FDR p` =
      format_p(
        adjusted_LRT_FDR_global
      )
  ) %>%
  dplyr::arrange(
    adjusted_LRT_p
  )

View(adjusted_exposure_summary)
```

```{r}
adjusted_exposure_summary %>%
  dplyr::filter(
    rank_deficient |
      !converged |
      !is.na(warning)
  )
```

```{r}
adjusted_exposure_summary %>%
  dplyr::filter(
    !is.na(exposure_GVIF_adjusted),
    exposure_GVIF_adjusted > sqrt(5)
  ) %>%
  dplyr::arrange(
    dplyr::desc(
      exposure_GVIF_adjusted
    )
  )
```

```{r}
format_p <- function(x, digits = 3) {
  dplyr::case_when(
    is.na(x) ~ "",
    x < 0.001 ~ "<0.001",
    TRUE ~ formatC(
      x,
      format = "f",
      digits = digits
    )
  )
}

format_count_percent <- function(count, denominator) {
  dplyr::case_when(
    is.na(count) |
      is.na(denominator) |
      denominator == 0 ~ "",

    TRUE ~ sprintf(
      "%d (%.1f%%)",
      as.integer(count),
      100 * count / denominator
    )
  )
}

first_non_missing <- function(x) {
  x <- x[!is.na(x)]

  if (length(x) == 0) {
    NA
  } else {
    x[[1]]
  }
}

make_adjusted_level_counts <- function(
    data,
    exposure,
    exposure_family,
    outcome = "had_HSA",
    forced_covariates
) {

  required_variables <- unique(
    c(
      outcome,
      exposure,
      forced_covariates
    )
  )

  model_data <- data[
    ,
    required_variables,
    drop = FALSE
  ]

  # Same complete-case population as the adjusted model
  model_data <- model_data[
    complete.cases(model_data),
    ,
    drop = FALSE
  ]

  model_data <- droplevels(model_data)

  # Outcome as 0/1
  if (
    is.factor(model_data[[outcome]]) ||
      is.character(model_data[[outcome]])
  ) {
    model_data[[outcome]] <- suppressWarnings(
      as.integer(
        as.character(model_data[[outcome]])
      )
    )
  }

  if (is.logical(model_data[[outcome]])) {
    model_data[[outcome]] <-
      as.integer(model_data[[outcome]])
  }

  x <- model_data[[exposure]]
  y <- model_data[[outcome]]

  # Match transformations in the modelling function
  if (is.ordered(x)) {
    x <- factor(
      as.character(x),
      levels = levels(x),
      ordered = FALSE
    )
  } else if (
    is.character(x) ||
      is.logical(x)
  ) {
    x <- factor(x)
  }

  total_cases <- sum(
    y == 1,
    na.rm = TRUE
  )

  total_non_cases <- sum(
    y == 0,
    na.rm = TRUE
  )

  # Factor exposure
  if (is.factor(x)) {

    exposure_levels <- levels(x)

    tab <- table(
      factor(
        x,
        levels = exposure_levels
      ),
      factor(
        y,
        levels = c(0, 1)
      )
    )

    return(
      tibble::tibble(
        exposure_family = exposure_family,
        exposure = exposure,
        Category = exposure_levels,
        category_order = seq_along(exposure_levels),
        model_type = "factor",
        reference_level = exposure_levels[[1]],
        comparison_level = NA_character_,

        outcome_0 = as.integer(
          tab[, "0"]
        ),

        outcome_1 = as.integer(
          tab[, "1"]
        ),

        total_cases = total_cases,
        total_non_cases = total_non_cases,
        model_n = nrow(model_data)
      )
    )
  }

  distinct_values <- sort(
    unique(x)
  )

  # Numeric binary exposure
  if (
    is.numeric(x) &&
      length(distinct_values) == 2
  ) {

    x_factor <- factor(
      x,
      levels = distinct_values
    )

    tab <- table(
      x_factor,
      factor(
        y,
        levels = c(0, 1)
      )
    )

    return(
      tibble::tibble(
        exposure_family = exposure_family,
        exposure = exposure,
        Category = as.character(
          distinct_values
        ),
        category_order =
          seq_along(distinct_values),
        model_type = "binary_numeric",

        reference_level =
          as.character(distinct_values[[1]]),

        comparison_level =
          as.character(distinct_values[[2]]),

        outcome_0 = as.integer(
          tab[, "0"]
        ),

        outcome_1 = as.integer(
          tab[, "1"]
        ),

        total_cases = total_cases,
        total_non_cases = total_non_cases,
        model_n = nrow(model_data)
      )
    )
  }

  # Continuous exposure
  tibble::tibble(
    exposure_family = exposure_family,
    exposure = exposure,
    Category = "Continuous",
    category_order = 1L,
    model_type = "continuous",
    reference_level = NA_character_,
    comparison_level = NA_character_,
    outcome_0 = total_non_cases,
    outcome_1 = total_cases,
    total_cases = total_cases,
    total_non_cases = total_non_cases,
    model_n = nrow(model_data)
  )
}

adjusted_level_counts <- purrr::map2_dfr(
  adjusted_candidate_lookup$exposure,
  adjusted_candidate_lookup$exposure_family,

  function(exposure_name, family_name) {

    make_adjusted_level_counts(
      data = analysis_df,
      exposure = exposure_name,
      exposure_family = family_name,
      outcome = "had_HSA",
      forced_covariates = forced_covariates
    )
  }
) %>%
  mutate(
    `Cases (%)` = format_count_percent(
      outcome_1,
      total_cases
    ),

    `Non-cases (%)` =
      format_count_percent(
        outcome_0,
        total_non_cases
      )
  )

adjusted_exposure_metadata <-
  adjusted_level_counts %>%
  distinct(
    exposure_family,
    exposure,
    model_type,
    reference_level,
    comparison_level
  )

adjusted_model_display <-
  adjusted_single_results %>%

  left_join(
    adjusted_exposure_metadata,
    by = c(
      "exposure_family",
      "exposure"
    ),
    suffix = c("", "_counts")
  ) %>%

  mutate(
    Category = case_when(

      model_type == "continuous" ~
        "Continuous",

      model_type == "binary_numeric" ~
        comparison_level,

      model_type == "factor" &
        !is.na(term) &
        startsWith(term, exposure) ~
        substring(
          term,
          nchar(exposure) + 1L
        ),

      TRUE ~ term
    ),

    `Adjusted OR (95% CI)` =
      case_when(
        is.na(adjusted_OR) |
          is.na(adjusted_CI_lower) |
          is.na(adjusted_CI_upper) ~ "",

        TRUE ~ sprintf(
          "%.3f (%.3f–%.3f)",
          adjusted_OR,
          adjusted_CI_lower,
          adjusted_CI_upper
        )
      ),

    `Adjusted Wald p-value` =
      format_p(
        adjusted_wald_p
      )
  ) %>%

  select(
    exposure_family,
    exposure,
    Category,
    `Adjusted OR (95% CI)`,
    `Adjusted Wald p-value`
  )

adjusted_overall_display <-
  adjusted_single_results %>%

  group_by(
    exposure_family,
    exposure
  ) %>%

  summarise(
    `Adjusted LRT p-value` =
      format_p(
        first_non_missing(
          adjusted_LRT_p
        )
      ),

    `Family FDR p-value` =
      format_p(
        first_non_missing(
          adjusted_LRT_FDR_within_family
        )
      ),

    `Global FDR p-value` =
      format_p(
        first_non_missing(
          adjusted_LRT_FDR_global
        )
      ),

    `Exposure adjusted GVIF` =
      round(
        first_non_missing(
          exposure_GVIF_adjusted
        ),
        3
      ),

    `Maximum adjusted GVIF` =
      round(
        first_non_missing(
          max_GVIF_adjusted
        ),
        3
      ),

    converged = all(
      converged,
      na.rm = TRUE
    ),

    warning = paste(
      unique(
        warning[
          !is.na(warning) &
            warning != ""
        ]
      ),
      collapse = "; "
    ),

    .groups = "drop"
  )
```

```{r}
adjusted_results_table <-
  adjusted_level_counts %>%

  left_join(
    adjusted_model_display,
    by = c(
      "exposure_family",
      "exposure",
      "Category"
    )
  ) %>%

  left_join(
    adjusted_overall_display,
    by = c(
      "exposure_family",
      "exposure"
    )
  ) %>%

  mutate(
    is_reference =
      !is.na(reference_level) &
      Category == reference_level,

    `Adjusted OR (95% CI)` =
      case_when(
        is_reference ~
          "1.000 (reference)",

        TRUE ~ coalesce(
          `Adjusted OR (95% CI)`,
          ""
        )
      ),

    `Adjusted Wald p-value` =
      case_when(
        is_reference ~ "",

        TRUE ~ coalesce(
          `Adjusted Wald p-value`,
          ""
        )
      )
  )
family_order_lookup <- tibble::tibble(
  exposure_family =
    names(
      adjusted_candidate_exposure_families
    ),

  family_order =
    seq_along(
      adjusted_candidate_exposure_families
    )
)

exposure_order_lookup <- purrr::imap_dfr(
  adjusted_candidate_exposure_families,

  function(variables, family_name) {
    tibble::tibble(
      exposure_family = family_name,
      exposure = variables,
      exposure_order =
        seq_along(variables)
    )
  }
)
adjusted_results_table <-
  adjusted_results_table %>%

  left_join(
    family_order_lookup,
    by = "exposure_family"
  ) %>%

  left_join(
    exposure_order_lookup,
    by = c(
      "exposure_family",
      "exposure"
    )
  ) %>%

  arrange(
    family_order,
    exposure_order,
    category_order
  )

```

```{r}
adjusted_results_table <-
  adjusted_results_table %>%

  group_by(
    exposure_family,
    exposure
  ) %>%

  mutate(
    `Model n` = if_else(
      row_number() == 1L,
      as.character(model_n),
      ""
    ),

    `Adjusted LRT p-value` =
      if_else(
        row_number() == 1L,
        `Adjusted LRT p-value`,
        ""
      ),

    `Family FDR p-value` =
      if_else(
        row_number() == 1L,
        `Family FDR p-value`,
        ""
      ),

    `Global FDR p-value` =
      if_else(
        row_number() == 1L,
        `Global FDR p-value`,
        ""
      ),

    `Exposure adjusted GVIF` =
      if_else(
        row_number() == 1L,
        as.character(
          `Exposure adjusted GVIF`
        ),
        ""
      ),

    `Maximum adjusted GVIF` =
      if_else(
        row_number() == 1L,
        as.character(
          `Maximum adjusted GVIF`
        ),
        ""
      ),

    Converged = if_else(
      row_number() == 1L,
      as.character(converged),
      ""
    ),

    Warning = if_else(
      row_number() == 1L,
      coalesce(warning, ""),
      ""
    )
  ) %>%

  ungroup()
adjusted_results_table <-
  adjusted_results_table %>%

  transmute(
    Group = exposure_family,
    Variable = exposure,
    Category,
    `Model n`,
    `Cases (%)`,
    `Non-cases (%)`,
    `Adjusted OR (95% CI)`,
    `Adjusted Wald p-value`,
    `Adjusted LRT p-value`,
    `Family FDR p-value`,
    `Global FDR p-value`,
    `Exposure adjusted GVIF`,
    `Maximum adjusted GVIF`,
    Converged,
    Warning
  )
adjusted_results_table <-
  adjusted_results_table %>%

  mutate(
    Group_original = Group,
    Variable_original = Variable
  ) %>%

  group_by(
    Group_original,
    Variable_original
  ) %>%

  mutate(
    Variable = if_else(
      row_number() == 1L,
      Variable,
      ""
    )
  ) %>%

  ungroup() %>%

  group_by(Group_original) %>%

  mutate(
    Group = if_else(
      row_number() == 1L,
      Group,
      ""
    )
  ) %>%

  ungroup() %>%

  select(
    -Group_original,
    -Variable_original
  )
```

Save output

```{r}
output_folder <- "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output"


dir.create(
  output_folder,
  recursive = TRUE,
  showWarnings = FALSE
)

readr::write_excel_csv(
  adjusted_results_table,
  file.path(
    output_folder,
    "adjusted_single_exposure_detailed_results.csv"
  ),
  na = ""
)
#need to format differently shorter col names etc for word

adjusted_results_word <- adjusted_results_table %>%
  dplyr::select(
    Group,
    Variable,
    Category,
    `Model n`,
    `Cases (%)`,
    `Non-cases (%)`,
    `Adjusted OR (95% CI)`,
    `Adjusted Wald p-value`,
    `Adjusted LRT p-value`,
    `Family FDR p-value`,
    `Global FDR p-value`
  )

ft_adj <- flextable::flextable(
  adjusted_results_word
) %>%

  flextable::set_header_labels(
    Group = "Group",
    Variable = "Variable",
    Category = "Category",
    `Model n` = "n",
    `Cases (%)` = "Cases (%)",
    `Non-cases (%)` = "Non-cases (%)",
    `Adjusted OR (95% CI)` = "aOR (95% CI)",
    `Adjusted Wald p-value` = "Wald p",
    `Adjusted LRT p-value` = "LRT p",
    `Family FDR p-value` = "Family FDR",
    `Global FDR p-value` = "Global FDR"
  ) %>%

  flextable::theme_booktabs() %>%

  flextable::bold(
    part = "header"
  ) %>%

  flextable::fontsize(
    size = 7.5,
    part = "all"
  ) %>%

  flextable::font(
    fontname = "Arial",
    part = "all"
  ) %>%

  flextable::valign(
    valign = "top",
    part = "all"
  ) %>%

  flextable::align(
    j = c(
      "Group",
      "Variable",
      "Category"
    ),
    align = "left",
    part = "all"
  ) %>%

  flextable::align(
    j = c(
      "Model n",
      "Cases (%)",
      "Non-cases (%)",
      "Adjusted OR (95% CI)",
      "Adjusted Wald p-value",
      "Adjusted LRT p-value",
      "Family FDR p-value",
      "Global FDR p-value"
    ),
    align = "center",
    part = "all"
  ) %>%

  # Reduce wasted space inside cells
  flextable::padding(
    padding.top = 1,
    padding.bottom = 1,
    padding.left = 2,
    padding.right = 2,
    part = "all"
  ) %>%

  # Fixed widths in inches
  flextable::width(
    j = "Group",
    width = 1.10
  ) %>%

  flextable::width(
    j = "Variable",
    width = 1.70
  ) %>%

  flextable::width(
    j = "Category",
    width = 1.35
  ) %>%

  flextable::width(
    j = "Model n",
    width = 0.45
  ) %>%

  flextable::width(
    j = "Cases (%)",
    width = 0.75
  ) %>%

  flextable::width(
    j = "Non-cases (%)",
    width = 0.85
  ) %>%

  flextable::width(
    j = "Adjusted OR (95% CI)",
    width = 1.30
  ) %>%

  flextable::width(
    j = c(
      "Adjusted Wald p-value",
      "Adjusted LRT p-value",
      "Family FDR p-value",
      "Global FDR p-value"
    ),
    width = 0.65
  ) %>%

  flextable::set_table_properties(
    layout = "fixed",
    width = 1,
    opts_word = list(
      split = TRUE
    )
  )

doc_adj <- officer::read_docx() %>%

  officer::body_add_par(
    "Core-adjusted single-exposure logistic regression results",
    style = "heading 1"
  ) %>%

  flextable::body_add_flextable(
    ft_adj
  ) %>%

  officer::body_end_section_landscape()

print(
  doc_adj,
  target = file.path(
    output_folder,
    "adjusted_single_exposure_detailed_results.docx"
  )
)
```

From these adjusted models retain for multivariable:

-   primary heating fuel

-   weed treatment rest or whole life (exact same case control counts)

-   fertiliser rest of life or whole life (exact same cxase control counts)

-   main lifestyle category mode 5y prior

-   mode activity intensity for 5y prior

-   mode activity frequency 5y prior

-   NSAID ever Y N and NSAID 5y prior YN

-   NSAID quantiles

-   quartile parasite lifetime and 5y prior endpoint

-   ecto 5y prior endo 5y prior

-   fenbendazole ever

Assess between dog activity variables for inclusion:

```{r}
activity_compare_data <- analysis_df %>%
  dplyr::select(
    had_HSA,
    age_at_final_date_bio,
    sex,
    neuter,
    mode_intensity_5y_prior_endpoint,
    mode_frequency_5y_prior_endpoint,
    rest_activity_mode
  ) %>%
  dplyr::filter(
    complete.cases(.)
  ) %>%
  droplevels()

#intensity
mode_intensity_5y_prior_endpoint <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    mode_intensity_5y_prior_endpoint,
  data = activity_compare_data,
  family = binomial()
)
#frequency
mode_frequency_5y_prior_endpoint <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    mode_intensity_5y_prior_endpoint,
  data = activity_compare_data,
  family = binomial()
)
#rest of life activity
rest_activity_mode <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    rest_activity_mode,
  data = activity_compare_data,
  family = binomial()
)

#compare
AIC(
  mode_intensity_5y_prior_endpoint,
  mode_frequency_5y_prior_endpoint,
  rest_activity_mode
)

BIC(
   mode_intensity_5y_prior_endpoint,
  mode_frequency_5y_prior_endpoint,
  rest_activity_mode
)
```

Not much difference exc intensity and freq 5y prior sl lower AIC - frequency is more interpret-able so take forward

Assess between NSAID variables for inclusion:

```{r}
nsaid_compare_data <- analysis_df %>%
  dplyr::select(
    had_HSA,
    age_at_final_date_bio,
    sex,
    neuter,
    NSAID_5y,
    NSAID_5y_quantiles,
    NSAID_ever
  ) %>%
  dplyr::filter(
    complete.cases(.)
  ) %>%
  droplevels()

#5y YN
nsaid_binary_model <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    NSAID_5y,
  data = nsaid_compare_data,
  family = binomial()
)
#ever YN
nsaid_binary_model_ever <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    NSAID_ever,
  data = nsaid_compare_data,
  family = binomial()
)
#5y categorical 
nsaid_category_model <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    NSAID_5y_quantiles,
  data = nsaid_compare_data,
  family = binomial()
)

#compare
AIC(
  nsaid_binary_model,
  nsaid_binary_model_ever,
  nsaid_category_model
)

BIC(
  nsaid_binary_model,
  nsaid_binary_model_ever,
  nsaid_category_model
)
```

Both binaries are equivocal and categorical worse from a BIC perspective = use a binary

Assess between parasite tx for inclusion:

```{r}
parasite_compare_data <- analysis_df %>%
  dplyr::select(
    had_HSA,
    age_at_final_date_bio,
    sex,
    neuter,
    quartile_anti_parasites_lifetime,
    half_anti_parasites_lifetime,
    quartile_anti_parasites_5y_prior_endpoint,
    ecto_5y_quantiles,
    endo_5y_quantiles,
    fenbendazole_ever
  ) %>%
  dplyr::filter(
    complete.cases(.)
  ) %>%
  droplevels()
# Lifetime anti-parasite use: halves
half_anti_parasites_lifetime_model <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    half_anti_parasites_lifetime,
  data = parasite_compare_data,
  family = binomial()
)

# Lifetime anti-parasite use: quartiles
quartile_anti_parasites_lifetime_model <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    quartile_anti_parasites_lifetime,
  data = parasite_compare_data,
  family = binomial()
)


# Five-year anti-parasite use: quartiles
quartile_anti_parasites_5y_model <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    quartile_anti_parasites_5y_prior_endpoint,
  data = parasite_compare_data,
  family = binomial()
)

# Five-year ectoparasiticide use
ecto_5y_model <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    ecto_5y_quantiles,
  data = parasite_compare_data,
  family = binomial()
)

# Five-year endoparasiticide use
endo_5y_model <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    endo_5y_quantiles,
  data = parasite_compare_data,
  family = binomial()
)

# Fenbendazole ever
fenbendazole_ever_model <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    fenbendazole_ever,
  data = parasite_compare_data,
  family = binomial()
)



AIC(
  half_anti_parasites_lifetime_model,
  quartile_anti_parasites_lifetime_model,
  quartile_anti_parasites_5y_model,
  ecto_5y_model,
  endo_5y_model,
  fenbendazole_ever_model
)

BIC(
  half_anti_parasites_lifetime_model,
  quartile_anti_parasites_lifetime_model,
  quartile_anti_parasites_5y_model,
  ecto_5y_model,
  endo_5y_model,
  fenbendazole_ever_model
)
```

## Multivariable modelling

Variables to consider:

-   primary heating fuel

-   weed treatment whole life

-   fertiliser whole life

-   main lifestyle category mode 5y prior

-   mode activity frequency 5y prior

-   NSAID ever Y N

-   half anti parasites lifetime

```{r}
candidate_exposures <- c(
  "heating_fuel_primary_mode_5yrs_prior",
  "any_treated_weeds_study_years_whole_life",
  "any_treated_fertilizer_study_years_whole_life",
  "main_lifestyle_category_mode_5yrs_prior",
  "mode_frequency_5y_prior_endpoint",
  "NSAID_ever",
  "half_anti_parasites_lifetime"
)
```

```{r}
forced_covariates <- c(
  "age_at_final_date_bio",
  "sex",
  "neuter"
)

multivar_variables <- c(
  forced_covariates,
  candidate_exposures
)

multivar_data <- analysis_df %>%
  dplyr::select(
    had_HSA,
    dplyr::all_of(multivar_variables)
  ) %>%
  dplyr::filter(
    complete.cases(.)
  ) %>%
  droplevels()


for (variable in multivar_variables) {

  if (is.ordered(multivar_data[[variable]])) {

    original_levels <- levels(
      multivar_data[[variable]]
    )

    multivar_data[[variable]] <- factor(
      as.character(multivar_data[[variable]]),
      levels = original_levels,
      ordered = FALSE
    )

  } else if (
    is.character(multivar_data[[variable]]) ||
      is.logical(multivar_data[[variable]])
  ) {

    multivar_data[[variable]] <- factor(
      multivar_data[[variable]]
    )
  }
}
```

```{r}
full_trial_model <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    heating_fuel_primary_mode_5yrs_prior +
    any_treated_weeds_study_years_whole_life +
    any_treated_fertilizer_study_years_whole_life +
    main_lifestyle_category_mode_5yrs_prior +
    mode_frequency_5y_prior_endpoint +
    NSAID_ever +
    half_anti_parasites_lifetime,
  data = multivar_data,
  family = binomial()
)

summary(full_trial_model)
logistic.display(full_trial_model)
```

```{r}
drop1(
  full_trial_model,
  test = "LRT"
)
```

```{r}
car::vif(
  full_trial_model
)
```

From this output:

-   no VIF issue but several variables become NS

-   retain: fixed covariates of age, sex, neuter and frequencey of exercise and assess how it performs vs fuller model

```{r}
reduced_trial_model <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    mode_frequency_5y_prior_endpoint ,
  data = multivar_data,
  family = binomial()
)

summary(reduced_trial_model)
logistic.display(reduced_trial_model)
```

```{r}
anova(
  reduced_trial_model,
  full_trial_model,
  test = "LRT"
)
AIC(
  reduced_trial_model,
  full_trial_model
)

BIC(
  reduced_trial_model,
  full_trial_model
)
```

```{r}
AIC(
  reduced_trial_model,
  full_trial_model
)

BIC(
  reduced_trial_model,
  full_trial_model
)
```

According to LRT - full model significantly better and likewise for AIC although BIC reduced model performs better

Issue with variable seperation as not recorded and never each have 0 cases or non cases respectively

trial on complete cases df

```{r}
analysis_df <- analysis_df %>%
  dplyr::mutate(
    mode_frequency_5y_clean = dplyr::na_if(
      as.character(mode_frequency_5y_prior_endpoint),
      "not_recorded"
    ),

    mode_frequency_5y_clean = factor(
      mode_frequency_5y_clean,
      levels = c(
        "daily",
        "multi_daily",
        "min_weekly",
        "less_than_weekly",
        "never"
      )
    )
  )


candidate_exposures <- c(
  "heating_fuel_primary_mode_5yrs_prior",
  "any_treated_weeds_study_years_whole_life",
  "any_treated_fertilizer_study_years_whole_life",
  "main_lifestyle_category_mode_5yrs_prior",
  "mode_frequency_5y_clean",
  "NSAID_ever",
  "half_anti_parasites_lifetime"
)

model_building_data <- analysis_df %>%
  dplyr::select(
    had_HSA,
    age_at_final_date_bio,
    sex,
    neuter,
    dplyr::all_of(candidate_exposures)
  ) %>%
  dplyr::filter(
    complete.cases(.)
  ) %>%
  droplevels()


core_model <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter,
  data = model_building_data,
  family = binomial()
)


```

```{r}
core_activity_model <- update(
  core_model,
  . ~ . + mode_frequency_5y_clean
)

anova(
  core_model,
  core_activity_model,
  test = "LRT"
)

AIC(
  core_model,
  core_activity_model
)

BIC(
  core_model,
  core_activity_model
)
```

Do forced covar model vs full and then all variables these stepwise ones

```{r}
core_model <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter,
  data = model_building_data,
  family = binomial()
)
```

```{r}
core_heating_model <- update(
  core_model,
  . ~ . + heating_fuel_primary_mode_5yrs_prior
)

core_weeds_model <- update(
  core_model,
  . ~ . + any_treated_weeds_study_years_whole_life
)

core_fertiliser_model <- update(
  core_model,
  . ~ . + any_treated_fertilizer_study_years_whole_life
)

core_lifestyle_model <- update(
  core_model,
  . ~ . + main_lifestyle_category_mode_5yrs_prior
)

core_activity_model <- update(
  core_model,
  . ~ . + mode_frequency_5y_clean
)

core_NSAID_model <- update(
  core_model,
  . ~ . + NSAID_ever
)

core_parasite_model <- update(
  core_model,
  . ~ . + half_anti_parasites_lifetime
)
```

```{r}
anova(core_model, core_heating_model, test = "LRT")
anova(core_model, core_weeds_model, test = "LRT")
anova(core_model, core_fertiliser_model, test = "LRT")
anova(core_model, core_lifestyle_model, test = "LRT")
anova(core_model, core_activity_model, test = "LRT")
anova(core_model, core_NSAID_model, test = "LRT")
anova(core_model, core_parasite_model, test = "LRT")
```

```{r}
AIC(
  core_model,
  core_heating_model,
  core_weeds_model,
  core_fertiliser_model,
  core_lifestyle_model,
  core_activity_model,
  core_NSAID_model,
  core_parasite_model
)

BIC(
  core_model,
  core_heating_model,
  core_weeds_model,
  core_fertiliser_model,
  core_lifestyle_model,
  core_activity_model,
  core_NSAID_model,
  core_parasite_model
)
```

From LRT:

Adding in parasite, nsaid, weed, heating fuel and lifestyle each improve vs core model

From AIC:

All better than just core

From BIC:

only fertiliser better

Build onwards from fertiliser

```{r}
fertiliser_model<- glm(
had_HSA ~ age_at_final_date_bio +
  sex +
  neuter +
  any_treated_fertilizer_study_years_whole_life, data=model_building_data,family=binomial())

logistic.display(fertiliser_model)
```

Trial other variables alongside fertiliser

```{r}
fertiliser_parasite_model <- update(
  fertiliser_model,
  . ~ . + half_anti_parasites_lifetime
)

fertiliser_NSAID_model <- update(
  fertiliser_model,
  . ~ . + NSAID_ever
)

fertiliser_weeds_model <- update(
  fertiliser_model,
  . ~ . + any_treated_weeds_study_years_whole_life
)

fertiliser_heating_model <- update(
  fertiliser_model,
  . ~ . + heating_fuel_primary_mode_5yrs_prior
)

fertiliser_lifestyle_model <- update(
  fertiliser_model,
  . ~ . + main_lifestyle_category_mode_5yrs_prior
)
```

```{r}
anova(
  fertiliser_model,
  fertiliser_parasite_model,
  test = "LRT"
)

anova(
  fertiliser_model,
  fertiliser_NSAID_model,
  test = "LRT"
)

anova(
  fertiliser_model,
  fertiliser_weeds_model,
  test = "LRT"
)

anova(
  fertiliser_model,
  fertiliser_heating_model,
  test = "LRT"
)

anova(
  fertiliser_model,
  fertiliser_lifestyle_model,
  test = "LRT"
)
```

```{r}
AIC(
  fertiliser_model,
  fertiliser_parasite_model,
  fertiliser_NSAID_model,
  fertiliser_weeds_model,
  fertiliser_heating_model,
  fertiliser_lifestyle_model
)

BIC(
  fertiliser_model,
  fertiliser_parasite_model,
  fertiliser_NSAID_model,
  fertiliser_weeds_model,
  fertiliser_heating_model,
  fertiliser_lifestyle_model
)
```

From this:

-   remove weeds and assess the other variables in again stepwise (using lowest AIC of heating value then others in)

```{r}
model_stage_2 <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    any_treated_fertilizer_study_years_whole_life +
    heating_fuel_primary_mode_5yrs_prior,
  data = model_building_data,
  family = binomial()
)
```

```{r}
anova(
  fertiliser_model,
  model_stage_2,
  test = "LRT"
)

AIC(
  fertiliser_model,
  model_stage_2
)

BIC(
  fertiliser_model,
  model_stage_2
)
```

```{r}
#remaining variables trial in
stage2_parasite_model <- update(
  model_stage_2,
  . ~ . + half_anti_parasites_lifetime
)

stage2_NSAID_model <- update(
  model_stage_2,
  . ~ . + NSAID_ever
)

stage2_lifestyle_model <- update(
  model_stage_2,
  . ~ . + main_lifestyle_category_mode_5yrs_prior
)
```

```{r}
anova(
  model_stage_2,
  stage2_parasite_model,
  test = "LRT"
)

anova(
  model_stage_2,
  stage2_NSAID_model,
  test = "LRT"
)

anova(
  model_stage_2,
  stage2_lifestyle_model,
  test = "LRT"
)
AIC(
  model_stage_2,
  stage2_parasite_model,
  stage2_NSAID_model,
  stage2_lifestyle_model
)

BIC(
  model_stage_2,
  stage2_parasite_model,
  stage2_NSAID_model,
  stage2_lifestyle_model
)
```

All variables: worse on BIC, same AIC values, NSAID sig on LRT

```{r}
drop1(
  model_stage_2,
  test = "LRT"
)

logistic.display(
  model_stage_2
)

table(
  model_building_data$heating_fuel_primary_mode_5yrs_prior,
  model_building_data$had_HSA
)
```

Confirm if it's just the missing data driving fuel inclusion by removing "not_recorded" and assessing performance without it:

```{r}
heating_recorded_data <- model_building_data %>%
  dplyr::filter(
    heating_fuel_primary_mode_5yrs_prior != "not_recorded"
  ) %>%
  droplevels()


fertiliser_model_heating_cc <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    any_treated_fertilizer_study_years_whole_life,
  data = heating_recorded_data,
  family = binomial()
)

fertiliser_heating_model_cc <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    any_treated_fertilizer_study_years_whole_life +
    heating_fuel_primary_mode_5yrs_prior,
  data = heating_recorded_data,
  family = binomial()
)
```

```{r}
anova(
  fertiliser_model_heating_cc,
  fertiliser_heating_model_cc,
  test = "LRT"
)

AIC(
  fertiliser_model_heating_cc,
  fertiliser_heating_model_cc
)

BIC(
  fertiliser_model_heating_cc,
  fertiliser_heating_model_cc
)

drop1(
  fertiliser_heating_model_cc,
  test = "LRT"
)
```

:

```{r}
fertiliser_model <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    any_treated_fertilizer_study_years_whole_life,
  data = model_building_data,
  family = binomial()
)

fertiliser_NSAID_model <- update(
  fertiliser_model,
  . ~ . + NSAID_ever
)

fertiliser_parasite_model <- update(
  fertiliser_model,
  . ~ . + half_anti_parasites_lifetime
)

fertiliser_lifestyle_model <- update(
  fertiliser_model,
  . ~ . + main_lifestyle_category_mode_5yrs_prior
)
```

```{r}
anova(
  fertiliser_model,
  fertiliser_NSAID_model,
  test = "LRT"
)

anova(
  fertiliser_model,
  fertiliser_parasite_model,
  test = "LRT"
)

anova(
  fertiliser_model,
  fertiliser_lifestyle_model,
  test = "LRT"
)
```

```{r}
AIC(
  fertiliser_model,
  fertiliser_NSAID_model,
  fertiliser_parasite_model,
  fertiliser_lifestyle_model
)

BIC(
  fertiliser_model,
  fertiliser_NSAID_model,
  fertiliser_parasite_model,
  fertiliser_lifestyle_model
)
```

All improve on LRT and AIC but not BIC

```{r}
expanded_candidate_model <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    any_treated_fertilizer_study_years_whole_life +
    NSAID_ever +
    half_anti_parasites_lifetime +
    main_lifestyle_category_mode_5yrs_prior,
  data = model_building_data,
  family = binomial()
)

drop1(
  expanded_candidate_model,
  test = "LRT"
)
```

```{r}
anova(
  fertiliser_model,
  expanded_candidate_model,
  test = "LRT"
)

AIC(
  fertiliser_model,
  expanded_candidate_model
)

BIC(
  fertiliser_model,
  expanded_candidate_model
)
```

Only lifestyle sig in this final model of lifestyle, NSAID, parasite, fertiliser

Although lifestyle is only sig due to not recorded category = ignore

Asssess these alongside lifestyle

```{r}
lifestyle_model <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    main_lifestyle_category_mode_5yrs_prior,
  data = model_building_data,
  family = binomial()
)

anova(
  lifestyle_model,
  expanded_candidate_model,
  test = "LRT"
)

AIC(
  lifestyle_model,
  expanded_candidate_model
)

BIC(
  lifestyle_model,
  expanded_candidate_model
)
```

```{r}
logistic.display(lifestyle_model)
```

Once the missing category is removed heating is NS, so continue building without

```{r}
#with not recorded removed 
analysis_df <- analysis_df %>%
  dplyr::mutate(
    lifestyle_5y_clean = dplyr::na_if(
      as.character(main_lifestyle_category_mode_5yrs_prior),
      "not_recorded"
    ),

    lifestyle_5y_clean = factor(
      lifestyle_5y_clean
    ),

    NSAID_ever_clean = dplyr::na_if(
      as.character(NSAID_ever),
      "No medication records for this time period"
    ),

    NSAID_ever_clean = factor(
      NSAID_ever_clean,
      levels = c("0", "1")
    ),

    anti_parasites_lifetime_clean = dplyr::na_if(
      as.character(half_anti_parasites_lifetime),
      "No medication records for this time period"
    ),

    anti_parasites_lifetime_clean = factor(
      anti_parasites_lifetime_clean,
      levels = c(
        "lower_half",
        "upper_half"
      )
    )
  )

model_building_data_clean <- analysis_df %>%
  dplyr::select(
    had_HSA,
    age_at_final_date_bio,
    sex,
    neuter,
    any_treated_fertilizer_study_years_whole_life,
    lifestyle_5y_clean,
    NSAID_ever_clean,
    anti_parasites_lifetime_clean
  ) %>%
  dplyr::filter(
    complete.cases(.)
  ) %>%
  droplevels()

core_model_clean <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter,
  data = model_building_data_clean,
  family = binomial()
)

fertiliser_model_clean <- update(
  core_model_clean,
  . ~ . +
    any_treated_fertilizer_study_years_whole_life
)

lifestyle_model_clean <- update(
  core_model_clean,
  . ~ . + lifestyle_5y_clean
)

NSAID_model_clean <- update(
  core_model_clean,
  . ~ . + NSAID_ever_clean
)

parasite_model_clean <- update(
  core_model_clean,
  . ~ . + anti_parasites_lifetime_clean
)
```

```{r}
anova(
  core_model_clean,
  lifestyle_model_clean,
  test = "LRT"
)

AIC(
  core_model_clean,
  fertiliser_model_clean,
  lifestyle_model_clean,
  NSAID_model_clean,
  parasite_model_clean
)

BIC(
  core_model_clean,
  fertiliser_model_clean,
  lifestyle_model_clean,
  NSAID_model_clean,
  parasite_model_clean
)
```

Next final steps to explore NSIAD and parasite

```{r}
model_no_lifestyle<- glm(had_HSA ~
  age_at_final_date_bio +
  sex +
  neuter +
  any_treated_fertilizer_study_years_whole_life +
  NSAID_ever_clean +
  anti_parasites_lifetime_clean, data=analysis_df, family = binomial()
)

drop1(
  model_no_lifestyle,
  test = "LRT"
)

# Exact observations used in the larger model
comparison_data <- model.frame(
  model_no_lifestyle
)

nrow(comparison_data)
nobs(model_no_lifestyle)

fertiliser_model_same_data <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    any_treated_fertilizer_study_years_whole_life,
  data = comparison_data,
  family = binomial()
)
```

```{r}
anova(
  fertiliser_model_same_data,
  model_no_lifestyle,
  test = "LRT"
)

AIC(
  fertiliser_model_same_data,
  model_no_lifestyle
)

BIC(
  fertiliser_model_same_data,
  model_no_lifestyle
)
```

```{r}
fertiliser_NSAID_same_data <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    any_treated_fertilizer_study_years_whole_life +
    NSAID_ever_clean,
  data = comparison_data,
  family = binomial()
)

fertiliser_parasite_same_data <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    any_treated_fertilizer_study_years_whole_life +
    anti_parasites_lifetime_clean,
  data = comparison_data,
  family = binomial()
)
```

```{r}
anova(
  fertiliser_model_same_data,
  fertiliser_NSAID_same_data,
  test = "LRT"
)

anova(
  fertiliser_model_same_data,
  fertiliser_parasite_same_data,
  test = "LRT"
)

AIC(
  fertiliser_model_same_data,
  fertiliser_NSAID_same_data,
  fertiliser_parasite_same_data,
  model_no_lifestyle
)

BIC(
  fertiliser_model_same_data,
  fertiliser_NSAID_same_data,
  fertiliser_parasite_same_data,
  model_no_lifestyle
)
```

```{r}
#Final model from these
fertiliser_NSAID_same_data <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    any_treated_fertilizer_study_years_whole_life +
    NSAID_ever_clean,
  data = comparison_data,
  family = binomial()
)
logistic.display(fertiliser_NSAID_same_data)
```

Building the final model wtih missing data included

```{r}
final_model <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    any_treated_fertilizer_study_years_whole_life +
    NSAID_ever,
  data = cohort_abr_updated_weight2,
  family = binomial()
)
logistic.display(final_model)
```

```{r}

library(dplyr)
library(forcats)
library(finalfit)

finalfit_data <- cohort_abr_updated_weight2 %>%
  mutate(
    # Outcome: second level is the event modelled
    had_HSA = factor(
      as.character(had_HSA),
      levels = c("0", "1"),
      labels = c("No HSA", "HSA")
    ),

    # Set intended reference levels
    age_at_final_date_bio = forcats::fct_relevel(
      factor(age_at_final_date_bio),
      ">5-<8y"
    ),

    sex = forcats::fct_relevel(
      factor(sex),
      "F"
    ),

    neuter = forcats::fct_relevel(
      factor(neuter),
      "Entire"
    ),

    any_treated_fertilizer_study_years_whole_life =
      forcats::fct_relevel(
        factor(
          any_treated_fertilizer_study_years_whole_life
        ),
        "Not within whole life"
      ),

    # Clean and label NSAID categories
    NSAID_ever = case_when(
      as.character(NSAID_ever) == "0" ~
        "No",

      as.character(NSAID_ever) == "1" ~
        "Yes",

      as.character(NSAID_ever) %in% c(
        "drug_not_administered",
        "no_med_use_recorded",
        "No medication records for this time period"
      ) ~
        "No medication records for this time period",

      TRUE ~ as.character(NSAID_ever)
    ),

    NSAID_ever = factor(
      NSAID_ever,
      levels = c(
        "No",
        "Yes",
        "No medication records for this time period"
      )
    )
  )

```

```{r}
final_variables <- c(
  "had_HSA",
  "age_at_final_date_bio",
  "sex",
  "neuter",
  "any_treated_fertilizer_study_years_whole_life",
  "NSAID_ever"
)

finalfit_data <- finalfit_data %>%
  select(
    all_of(final_variables)
  ) %>%
  filter(
    complete.cases(.)
  ) %>%
  droplevels()
```

```{r}

```

```{r}
dependent <- "had_HSA"

explanatory <- c(
  "age_at_final_date_bio",
  "sex",
  "neuter",
  "any_treated_fertilizer_study_years_whole_life",
  "NSAID_ever"
)



final_model_table <-finalfit_data %>% 
  finalfit(dependent, explanatory, metrics = TRUE,column=TRUE) %>%
  kable(format = "html", digits = 2, escape = FALSE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

final_model_table
```

```{r}
final_model <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    any_treated_fertilizer_study_years_whole_life +
    NSAID_ever,
  data = cohort_abr_updated_weight2,
  family = binomial()
)

final_model_data <- model.frame(
  final_model
)

final_model_without_NSAID <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    any_treated_fertilizer_study_years_whole_life,
  data = final_model_data,
  family = binomial()
)


anova(
  final_model_without_NSAID,
  final_model,
  test = "LRT"
)

drop1(
  final_model,
  test = "LRT"
)

AIC(
  final_model_without_NSAID,
  final_model
)

BIC(
  final_model_without_NSAID,
  final_model
)

data.frame(
  criterion = c("AIC", "BIC"),

  without_NSAID = c(
    AIC(final_model_without_NSAID),
    BIC(final_model_without_NSAID)
  ),

  with_NSAID = c(
    AIC(final_model),
    BIC(final_model)
  )
) %>%
  dplyr::mutate(
    difference_with_minus_without =
      with_NSAID - without_NSAID
  )
```

VChecking if sex-neuter worth putting in instead of seperate sex and neuter

```{r}
with(
  model.frame(final_model),
  table(
    sex,
    neuter,
    had_HSA,
    useNA = "ifany"
  )
)
```

```{r}
final_model_sex_neuter_interaction <- update(
  final_model,
  . ~ . + sex:neuter
)


anova(
  final_model,
  final_model_sex_neuter_interaction,
  test = "LRT"
)

AIC(
  final_model,
  final_model_sex_neuter_interaction
)

BIC(
  final_model,
  final_model_sex_neuter_interaction
)

drop1(
  final_model_sex_neuter_interaction,
  test = "LRT"
)
```

```{r}
final_model_data <- model.frame(final_model) %>%
  dplyr::mutate(
    sex_neuter = interaction(
      sex,
      neuter,
      sep = "_",
      drop = TRUE
    )
  )

table(
  final_model_data$sex_neuter,
  final_model_data$had_HSA
)


final_model_data <- final_model_data %>%
  dplyr::mutate(
    sex_neuter = forcats::fct_relevel(
      sex_neuter,
      "F_Entire"
    )
  )


final_model_sex_neuter_combined <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex_neuter +
    any_treated_fertilizer_study_years_whole_life +
    NSAID_ever,
  data = final_model_data,
  family = binomial()
)

logistic.display(final_model_sex_neuter_combined)
AIC(
  final_model_sex_neuter_interaction,
  final_model_sex_neuter_combined
)

logLik(final_model_sex_neuter_interaction)
logLik(final_model_sex_neuter_combined)



#trial without NSAID
final_model_sex_neuter_combined_noNSAID <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex_neuter +
    any_treated_fertilizer_study_years_whole_life, 
  data = final_model_data,
  family = binomial()
)

AIC(final_model_sex_neuter_combined_noNSAID,final_model_sex_neuter_combined)
BIC(final_model_sex_neuter_combined_noNSAID,final_model_sex_neuter_combined)

anova(
  final_model_sex_neuter_combined_noNSAID,
  final_model_sex_neuter_combined,
  test = "LRT"
)

```

```{r}
final_model_data <- final_model_data %>%
  mutate(
    had_HSA = factor(
      had_HSA,
      levels = c(0, 1),
      labels = c("No HSA", "HSA")
    )
  )

dependent <- "had_HSA"

explanatory <- c(
  "age_at_final_date_bio",
  "sex_neuter",
  "any_treated_fertilizer_study_years_whole_life",
  "NSAID_ever"
)

final_model_table2 <- final_model_data %>%
  finalfit::finalfit(
    dependent = dependent,
    explanatory = explanatory,
    metrics = TRUE,
    column = TRUE
  ) %>%
  knitr::kable(
    format = "html",
    digits = 2,
    escape = FALSE
  ) %>%
  kableExtra::kable_styling(
    full_width = FALSE,
    bootstrap_options = c("striped", "hover")
  )

final_model_table2
```
