---
title: "GRLS HSA vs non-HSA logistic regression"
format: html
editor: visual
---

## HSA GRLS cohort

HSA cases vs non-case control logistic regression

```{r}
library(tidyverse)
library(tidyverse)
library(dplyr)
library(survival)
library(ggplot2)
library(tibble)
library(kableExtra)
library(epiDisplay)
library(lme4)
#devtools::install_github("zabore/ezfun")
ezfun::set_ccf_palette("contrast")
library(lubridate)
library(ggsurvfit)
library(gtsummary)
library(tidycmprsk)
library(gtable)
#devtools::install_github("zabore/condsurv")
library(condsurv)
library(survminer)
library(cardx)
library(finalfit)
library(crosstable)
library(kableExtra)
cohort_abr <-read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_cohort_all_RFs_data_tidied_from_cox.csv")
```

Converting columns to factors for analyses

```{r}

#replace all missing with NA (so "" and NA)
cohort_abr <- cohort_abr %>%
  mutate(across(everything(), as.character)) %>%  # Ensure all columns are character
  mutate(across(everything(), ~na_if(.x, "n/a"))) %>%
  mutate(across(everything(), ~na_if(.x, "NA"))) %>%
  mutate(across(everything(), ~na_if(.x, "nr"))) %>%
  mutate(across(everything(), ~na_if(.x, "not recorded"))) %>%
  mutate(across(everything(), ~na_if(.x, ""))) %>%
  mutate(across(everything(), ~replace_na(.x, "not_recorded")))
cohort_abr$survival_time_months <- as.numeric(cohort_abr$survival_time_months)


cohort_abr$status_factor <- as.factor(as.character(cohort_abr$had_HSA))
cohort_abr$status <-as.numeric(cohort_abr$had_HSA) 
cohort_abr$survival_time_months <- as.numeric(cohort_abr$survival_time_months)


#manually assign height, weight, BCS factor levles and age at neuter and age at diagnosis 
#height
cohort_abr <- cohort_abr%>%
  mutate(avg_height_5y_prior_endpoint_quartiles = factor(avg_height_5y_prior_endpoint_quartiles, levels = c("49.13-<56.07cm", ">=56.07-<58.42cm", ">=58.42-<60.32cm", ">=60.32cm-67.73cm","not_recorded")))
#weight
cohort_abr <- cohort_abr%>%
  mutate(avg_weight_5y_prior_endpoint_quartiles = factor(avg_weight_5y_prior_endpoint_quartiles, levels = c("9.75-<21.31kg", ">=21.31-<24.66kg", ">=24.66-<29.37kg", ">=29.37-41.52kg","not_recorded")))
#bcs
cohort_abr <- cohort_abr%>%
  mutate(avg_purina_BCS_5y_prior_endpoint_quartiles = factor(avg_purina_BCS_5y_prior_endpoint_quartiles, levels = c("3.3-<5", ">=5-<5.28", ">=5.28-<5.86", ">=5.86 - 9","not_recorded")))
#neut age

cohort_abr <- cohort_abr%>%
  mutate(age_at_neuter_years_quartiles = factor(age_at_neuter_years_quartiles, levels = c("0.15-<0.52y", ">=0.52-<0.83y", ">=0.83-2.44y", ">=2.44-9.59y","Entire","not_recorded")))
#diagnosis age
cohort_abr <- cohort_abr%>%
  mutate(age_at_final_date_quartiles = factor(age_at_final_date_quartiles, levels = c("0.74-<8.47 years", ">=8.47-<10.68 years", ">=10.68-<11.64 years", ">=11.64-14.51 years","not_recorded")))

#change spayed while in heat to account for males and entire females
cohort_abr <- cohort_abr %>%
  mutate(spayed_while_in_heat = case_when(
    sex== "M" ~ "Male",
    sex_status == "Intact Female" ~ "Intact Female",
    spayed_while_in_heat == "0" ~ "0",
    spayed_while_in_heat == "1" ~"1"
  ))
#same as above account for unnetered and males 
cohort_abr$heats_before_neutering <- as.numeric(cohort_abr$heats_before_neutering)
heats <- cohort_abr %>%
# Convert to numeric
  filter(heats_before_neutering != 0) %>% 
# Remove rows where the column is 0
  filter(heats_before_neutering != "not_recorded")%>%
  dplyr::select(heats_before_neutering)# Keep only this column
heats <- as.numeric(heats$heats_before_neutering)

# Function to calculate quartiles or terciles dynamically
get_breaks <- function(x, probs) {
  breaks <- quantile(x, probs = probs, na.rm = TRUE)
  
  # If breaks are not unique, switch to terciles
  if (length(unique(breaks)) < length(probs)) {
    message("Quartiles not unique, switching to terciles.")
    breaks <- quantile(x, probs = c(0, 1/3, 2/3, 1), na.rm = TRUE)
  }
  
  return(breaks)
}

breaks <- get_breaks(heats, c(0, 0.25, 0.5, 0.75, 1))


# Assign labels dynamically
labels <- paste0("≥", round(head(breaks, -1), 2), " - <", round(tail(breaks, -1), 2))


cohort_abr <- cohort_abr %>%
  mutate(heats_before_neutering_terciles = case_when(
    sex== "M" ~ "Male",
    sex_status == "Intact Female" ~ "Intact Female",
    heats_before_neutering == "0" ~ "0",
    heats_before_neutering == "1" ~"1-<=2",
    heats_before_neutering == "2" ~"1-<=2",
    heats_before_neutering == "3" ~">2-<=4",
    heats_before_neutering == "4" ~">2-<=4",
    heats_before_neutering == "5" ~">4-12",
    heats_before_neutering == "6" ~">4-12",
    heats_before_neutering == "7" ~">4-12",
    heats_before_neutering == "8" ~">4-12",
    heats_before_neutering == "9" ~">4-12",
    heats_before_neutering == "10" ~">4-12",
    heats_before_neutering == "11" ~">4-12",
    heats_before_neutering == "12" ~">4-12"
  ))

cohort_abr <- cohort_abr%>%
  mutate(heats_before_neutering_terciles = factor(heats_before_neutering_terciles, levels = c("0", "1-<=2", ">2-<=4", ">4-12","Intact Female","Male","not_recorded")))


#neuter reason levels
cohort_abr <- cohort_abr%>%
  mutate(neutered_reason = factor(neutered_reason, levels = c("Elective",  "Medical Reason", "Entire","Unknown")))


cohort_abr <- cohort_abr%>%
  mutate(mode_purina_BCS_5y_prior_endpoint = factor(mode_purina_BCS_5y_prior_endpoint, levels = c("5", "3", "4", "6","7","8","9","not_recorded")))

#change BCS to be over underweight
cohort_abr <- cohort_abr %>%
  mutate(mode_BCS_classify = case_when(
    mode_purina_BCS_5y_prior_endpoint == "1" ~"underweight",
    mode_purina_BCS_5y_prior_endpoint == "2" ~"underweight",
    mode_purina_BCS_5y_prior_endpoint == "3" ~"underweight",
    mode_purina_BCS_5y_prior_endpoint == "4" ~"ideal",
    mode_purina_BCS_5y_prior_endpoint == "5" ~"ideal",
    mode_purina_BCS_5y_prior_endpoint == "6" ~"overweight",
    mode_purina_BCS_5y_prior_endpoint == "7" ~"overweight",
    mode_purina_BCS_5y_prior_endpoint == "8" ~"overweight",
    mode_purina_BCS_5y_prior_endpoint == "9" ~"overweight",
  ))

#ensure not_recorded is default not baseline
convert_and_relevel <- function(df) {
  df <- df %>%
    mutate(across(where(is.character), ~ factor(.))) %>%  # Convert all character columns to factors
    mutate(across(where(is.factor), ~ fct_relevel(., "not_recorded", after = Inf)))  # Move "not_recorded" to last place
  return(df)
}

# Apply to your dataset
cohort_abr <- convert_and_relevel(cohort_abr)
```

## Univariable logstic regression

-   Counts

```{r}
dependent <- "had_HSA"

# All other variables as explanatory variables
explanatory = c("MDI_quintile","owner_MDI_quintile","area_type_mode_5yrs_prior","house_type_mode_5yrs_prior", 
             "water_source_mode_5yrs_prior", "pipes_metal_any_mode_5yrs_prior", "pipes_plastic_any_mode_5yrs_prior", 
             "heating_fuel_primary_mode_5yrs_prior", "cooking_fuel_primary_mode_5yrs_prior", "heating_fuel_secondary_mode_5yrs_prior", 
             "cooking_fuel_secondary_mode_5yrs_prior", "region_mode_5y_prior",  "number_comorbs_present_lifetime", 
             "study_year_first_chronic_inflam", "chronic_inflam_present_5y", "number_comorbs_present_5y", 
             "chronic_inflam_present_no_min_year", "number_comorbs_present_no_min_year",  "diagnosis_after_comorb",  "time_between_first_comorb_and_endpoint",  
              "num_lifetime_chronic_inflam", "comorb_intestinal_parasitism_chronic","comorb_tick_borne_parasitism_chronic","comorb_other_parasitism_chronic","comorb_other_parasitism_chronic","comorb_orthopaedic_chronic","comorb_immune_mediated_chronic","comorb_cardiovascular_chronic","comorb_gastrointestinal_chronic","comorb_inflammatory_other_chronic","comorb_orthopaedic","comorb_immune_mediated","comorb_cardiovascular","comorb_gastrointestinal","comorb_inflammatory_other",
 "USDA_30min_sy1_2", "activity_1h_min_sy1_2", 
             "activity_KC_2h_min_sy1_2", "avg_activity_duration_sy1_2_quartiles",#"avg_activity_intensity_sy1_2_terciles",
 "avg_activity_freq_sy1_2_terciles", 
             "avg_activity_freq_sy3_terciles", "avg_activity_intensity_sy3_quartiles", "avg_intensity_5y_prior_endpoint_quartiles", "avg_frequency_5y_prior_endpoint_quartiles", 
             "use_aerosol_study_years_early_life", "use_air_cleaner_study_years_early_life", "use_hepa_filter_study_years_early_life", 
             "use_moth_balls_study_years_early_life", "use_incense_or_candles_study_years_early_life", "smoke_exposure_study_years_early_life", 
       #      "any_treated_weeds_study_years_early_life", "any_treated_insects_study_years_early_life", #"any_treated_fertilizer_study_years_early_life", 
         #    "use_aerosol_sum_early_life", "use_air_cleaner_sum_early_life", "use_hepa_filter_sum_early_life", "use_moth_balls_sum_early_life", 
          #   "use_incense_or_candles_sum_early_life", 
 #"smoke_exposure_sum_early_life",
 #"any_treated_weeds_sum_early_life", 
    #         "any_treated_insects_sum_early_life", #"any_treated_fertilizer_sum_early_life", 
 "use_aerosol_study_years_rest_of_life", 
             "use_air_cleaner_study_years_rest_of_life", "use_hepa_filter_study_years_rest_of_life", "use_moth_balls_study_years_rest_of_life", 
             "use_incense_or_candles_study_years_rest_of_life", "smoke_exposure_study_years_rest_of_life", "any_treated_weeds_study_years_rest_of_life", 
             "any_treated_insects_study_years_rest_of_life", "any_treated_fertilizer_study_years_rest_of_life", #"use_aerosol_sum_rest_of_life", 
 #            "use_air_cleaner_sum_rest_of_life", "use_hepa_filter_sum_rest_of_life", "use_moth_balls_sum_rest_of_life", 
          #   "use_incense_or_candles_sum_rest_of_life", 
# "smoke_exposure_sum_rest_of_life", 
# "any_treated_weeds_sum_rest_of_life", 
            # "any_treated_insects_sum_rest_of_life", "any_treated_fertilizer_sum_rest_of_life", 
"use_aerosol_study_years_whole_life", 
             "use_air_cleaner_study_years_whole_life", "use_hepa_filter_study_years_whole_life", "use_moth_balls_study_years_whole_life", 
             "use_incense_or_candles_study_years_whole_life", "smoke_exposure_study_years_whole_life", "any_treated_weeds_study_years_whole_life", 
             "any_treated_insects_study_years_whole_life", "any_treated_fertilizer_study_years_whole_life",# "use_aerosol_sum_whole_life", 
           #  "use_air_cleaner_sum_whole_life", "use_hepa_filter_sum_whole_life", "use_moth_balls_sum_whole_life",# "use_incense_or_candles_sum_whole_life", 
            # "smoke_exposure_sum_whole_life", 
 #"any_treated_weeds_sum_whole_life", "any_treated_insects_sum_whole_life", 
            # "any_treated_fertilizer_sum_whole_life",
 "hours_of_smoke_early_life_total_dosage_terciles", "hours_of_smoke_rest_of_life_total_dosage_terciles", 
             "hours_of_smoke_whole_life_total_dosage_terciles", "in.the.garage_YN_study_years_early_life", "in.the.house_YN_study_years_early_life", 
             "outside_YN_study_years_early_life",  "in.the.house_YN_sum_early_life_terciles", 
             "in.the.garage_YN_study_years_rest_of_life", "in.the.house_YN_study_years_rest_of_life", "outside_YN_study_years_rest_of_life", 
              "in.the.house_YN_sum_rest_of_life_quartiles",  
             "in.the.garage_YN_study_years_whole_life", "in.the.house_YN_study_years_whole_life", "outside_YN_study_years_whole_life", 
             "in.the.house_YN_sum_whole_life_quartiles",  
             "majority_sleep_location_lifetime",  "main_lifestyle_category", 
 "main_lifestyle_category_mode_5yrs_prior", "sleep_location_mode_5yrs_prior", 
             "use_aerosol_exposed_5yrs_prior", "use_air_cleaner_exposed_5yrs_prior", "use_hepa_filter_exposed_5yrs_prior", 
             "use_moth_balls_exposed_5yrs_prior", "use_incense_or_candles_exposed_5yrs_prior", "smoke_exposure_exposed_5yrs_prior", 
             "any_treated_weeds_exposed_5yrs_prior", "any_treated_insects_exposed_5yrs_prior", "any_treated_fertilizer_exposed_5yrs_prior", 
             "hours_of_smoke_exposure_hours_5yrs_prior_terciles", "sex", "neuter", "mismating_ever",
             "neutered_reason", "neutered_ever_bred", "spayed_while_in_heat", "spayed_while_pregnant", "prevent_heat_ever", 
             "no_pregnancy_ever", "intact_sired_litters_natural_ever", "neutered_females_bred_with_count_ever", 
             "neutered_sired_litters_natural_ever", "heats_before_neutering", "heats_before_neutering_terciles",
             "preg_pre_neuter",# "mated_pre_neuter",
 "num_breeding_before_neutering_quartiles", "ever_bred_from", 
              "avg_weight_5y_prior_endpoint_quartiles", "avg_height_5y_prior_endpoint_quartiles", 
             "avg_purina_BCS_5y_prior_endpoint_quartiles", "mode_purina_BCS_5y_prior_endpoint", "mode_BCS_classify","half_anti_microbials_lifetime", "half_dietary_supplements_lifetime", 
             "half_enteric_lifetime", "half_fluid_metabolites_lifetime", "half_hormones_and_related_lifetime", "half_immunologicals_lifetime", 
             "half_misc_lifetime", "half_neurological_lifetime", "half_anti_parasites_lifetime", "half_anti_septics_lifetime", 
             "half_cardio_respiratory_lifetime", "half_diuretics_lifetime", "half_anti_hist_lifetime", 
             "quartile_anti_inflam_lifetime", "quartile_anti_microbials_lifetime", "quartile_dietary_supplements_lifetime", 
            #not enough data for quartiles "quartile_enteric_lifetime", "quartile_fluid_metabolites_lifetime", 
 "quartile_hormones_and_related_lifetime", 
             "quartile_immunologicals_lifetime", "quartile_misc_lifetime", "quartile_neurological_lifetime", 
             "quartile_anti_parasites_lifetime", 
 #not enough data for quartiles "quartile_anti_septics_lifetime", "quartile_cardio_respiratory_lifetime",   "quartile_diuretics_lifetime", "quartile_anti_hist_lifetime", 
             "half_anti_inflam_5y_prior_endpoint", "half_anti_microbials_5y_prior_endpoint", "half_enteric_5y_prior_endpoint", 
             "half_immunologicals_5y_prior_endpoint", "half_misc_5y_prior_endpoint", "half_neurological_5y_prior_endpoint", 
             "half_anti_parasites_5y_prior_endpoint", "half_dietary_supplements_5y_prior_endpoint", 
             "half_anti_septics_5y_prior_endpoint", "half_hormones_and_related_5y_prior_endpoint", 
             "half_cardio_respiratory_5y_prior_endpoint", "half_fluid_metabolites_5y_prior_endpoint", 
             "half_diuretics_5y_prior_endpoint", "half_anti_hist_5y_prior_endpoint", 
             "quartile_anti_inflam_5y_prior_endpoint", "quartile_anti_microbials_5y_prior_endpoint", 
            # "quartile_enteric_5y_prior_endpoint",
 "quartile_immunologicals_5y_prior_endpoint", 
             "quartile_misc_5y_prior_endpoint", "quartile_neurological_5y_prior_endpoint", 
             "quartile_anti_parasites_5y_prior_endpoint", "quartile_dietary_supplements_5y_prior_endpoint", 
       #      "quartile_anti_septics_5y_prior_endpoint", 
 "quartile_hormones_and_related_5y_prior_endpoint", 
          #   "quartile_cardio_respiratory_5y_prior_endpoint", #"quartile_fluid_metabolites_5y_prior_endpoint", 
         #    "quartile_diuretics_5y_prior_endpoint", "quartile_anti_hist_5y_prior_endpoint", 
             #"quartile_anti_viral_5y_prior_endpoint",
 "age_at_final_date_quartiles","age_at_neuter_years_quartiles")

# Create a univariable logistic regression table
univar_table <- cohort_abr %>%summary_factorlist(dependent, explanatory, 
  p=TRUE,  add_dependent_label=TRUE)


univar_table2 <- cohort_abr %>%summary_factorlist(dependent, explanatory, 
  p=TRUE,  add_dependent_label=TRUE,na_include = TRUE)
# View the results
print(univar_table)
```

-   Univar log reg with ORs

```{r}
source("C:/Users/ctaylor18/GitHub/HSA-VC-enviro-analyses/functions/univar_log_reg_auto.R")

cohort_abr$had_HSA <- as.factor(cohort_abr$had_HSA)

#leaving out the continuous variables as missing data causing issues
#"max_tumour_cm_value" "age_at_diagnosis_cont"
cohort_abr2 <- cohort_abr %>%
  dplyr::select("MDI_quintile","owner_MDI_quintile","area_type_mode_5yrs_prior","house_type_mode_5yrs_prior", 
             "water_source_mode_5yrs_prior", "pipes_metal_any_mode_5yrs_prior", "pipes_plastic_any_mode_5yrs_prior", 
             "heating_fuel_primary_mode_5yrs_prior", "cooking_fuel_primary_mode_5yrs_prior", "heating_fuel_secondary_mode_5yrs_prior", 
             "cooking_fuel_secondary_mode_5yrs_prior", "region_mode_5y_prior",  "number_comorbs_present_lifetime", 
             "study_year_first_chronic_inflam", "chronic_inflam_present_5y", "number_comorbs_present_5y", 
             "chronic_inflam_present_no_min_year", "number_comorbs_present_no_min_year",  "diagnosis_after_comorb",  "time_between_first_comorb_and_endpoint",  
              "num_lifetime_chronic_inflam", "comorb_intestinal_parasitism_chronic","comorb_tick_borne_parasitism_chronic","comorb_other_parasitism_chronic","comorb_other_parasitism_chronic","comorb_orthopaedic_chronic","comorb_immune_mediated_chronic","comorb_cardiovascular_chronic","comorb_gastrointestinal_chronic","comorb_inflammatory_other_chronic","comorb_orthopaedic","comorb_immune_mediated","comorb_cardiovascular","comorb_gastrointestinal","comorb_inflammatory_other",
 "USDA_30min_sy1_2", "activity_1h_min_sy1_2", 
             "activity_KC_2h_min_sy1_2", "avg_activity_duration_sy1_2_quartiles",#"avg_activity_intensity_sy1_2_terciles",
 "avg_activity_freq_sy1_2_terciles", 
             "avg_activity_freq_sy3_terciles", "avg_activity_intensity_sy3_quartiles", "avg_intensity_5y_prior_endpoint_quartiles", "avg_frequency_5y_prior_endpoint_quartiles", 
             "use_aerosol_study_years_early_life", "use_air_cleaner_study_years_early_life", "use_hepa_filter_study_years_early_life", 
             "use_moth_balls_study_years_early_life", "use_incense_or_candles_study_years_early_life", "smoke_exposure_study_years_early_life", 
       #      "any_treated_weeds_study_years_early_life", "any_treated_insects_study_years_early_life", #"any_treated_fertilizer_study_years_early_life", 
         #    "use_aerosol_sum_early_life", "use_air_cleaner_sum_early_life", "use_hepa_filter_sum_early_life", "use_moth_balls_sum_early_life", 
          #   "use_incense_or_candles_sum_early_life", 
 #"smoke_exposure_sum_early_life",
 #"any_treated_weeds_sum_early_life", 
    #         "any_treated_insects_sum_early_life", #"any_treated_fertilizer_sum_early_life", 
 "use_aerosol_study_years_rest_of_life", 
             "use_air_cleaner_study_years_rest_of_life", "use_hepa_filter_study_years_rest_of_life", "use_moth_balls_study_years_rest_of_life", 
             "use_incense_or_candles_study_years_rest_of_life", "smoke_exposure_study_years_rest_of_life", "any_treated_weeds_study_years_rest_of_life", 
             "any_treated_insects_study_years_rest_of_life", "any_treated_fertilizer_study_years_rest_of_life", #"use_aerosol_sum_rest_of_life", 
 #            "use_air_cleaner_sum_rest_of_life", "use_hepa_filter_sum_rest_of_life", "use_moth_balls_sum_rest_of_life", 
          #   "use_incense_or_candles_sum_rest_of_life", 
# "smoke_exposure_sum_rest_of_life", 
# "any_treated_weeds_sum_rest_of_life", 
            # "any_treated_insects_sum_rest_of_life", "any_treated_fertilizer_sum_rest_of_life", 
"use_aerosol_study_years_whole_life", 
             "use_air_cleaner_study_years_whole_life", "use_hepa_filter_study_years_whole_life", "use_moth_balls_study_years_whole_life", 
             "use_incense_or_candles_study_years_whole_life", "smoke_exposure_study_years_whole_life", "any_treated_weeds_study_years_whole_life", 
             "any_treated_insects_study_years_whole_life", "any_treated_fertilizer_study_years_whole_life",# "use_aerosol_sum_whole_life", 
           #  "use_air_cleaner_sum_whole_life", "use_hepa_filter_sum_whole_life", "use_moth_balls_sum_whole_life",# "use_incense_or_candles_sum_whole_life", 
            # "smoke_exposure_sum_whole_life", 
 #"any_treated_weeds_sum_whole_life", "any_treated_insects_sum_whole_life", 
            # "any_treated_fertilizer_sum_whole_life",
 "hours_of_smoke_early_life_total_dosage_terciles", "hours_of_smoke_rest_of_life_total_dosage_terciles", 
             "hours_of_smoke_whole_life_total_dosage_terciles", "in.the.garage_YN_study_years_early_life", "in.the.house_YN_study_years_early_life", 
             "outside_YN_study_years_early_life",  "in.the.house_YN_sum_early_life_terciles", 
             "in.the.garage_YN_study_years_rest_of_life", "in.the.house_YN_study_years_rest_of_life", "outside_YN_study_years_rest_of_life", 
              "in.the.house_YN_sum_rest_of_life_quartiles",  
             "in.the.garage_YN_study_years_whole_life", "in.the.house_YN_study_years_whole_life", "outside_YN_study_years_whole_life", 
             "in.the.house_YN_sum_whole_life_quartiles",  
             "majority_sleep_location_lifetime",  "main_lifestyle_category", 
 "main_lifestyle_category_mode_5yrs_prior", "sleep_location_mode_5yrs_prior", 
             "use_aerosol_exposed_5yrs_prior", "use_air_cleaner_exposed_5yrs_prior", "use_hepa_filter_exposed_5yrs_prior", 
             "use_moth_balls_exposed_5yrs_prior", "use_incense_or_candles_exposed_5yrs_prior", "smoke_exposure_exposed_5yrs_prior", 
             "any_treated_weeds_exposed_5yrs_prior", "any_treated_insects_exposed_5yrs_prior", "any_treated_fertilizer_exposed_5yrs_prior", 
             "hours_of_smoke_exposure_hours_5yrs_prior_terciles", "sex", "neuter", "mismating_ever",
             "neutered_reason", "neutered_ever_bred", "spayed_while_in_heat", "spayed_while_pregnant", "prevent_heat_ever", 
             "no_pregnancy_ever", "intact_sired_litters_natural_ever", "neutered_females_bred_with_count_ever", 
             "neutered_sired_litters_natural_ever", "heats_before_neutering", "heats_before_neutering_terciles",
             "preg_pre_neuter",# "mated_pre_neuter",
 "num_breeding_before_neutering_quartiles", "ever_bred_from", 
              "avg_weight_5y_prior_endpoint_quartiles", "avg_height_5y_prior_endpoint_quartiles", 
             "avg_purina_BCS_5y_prior_endpoint_quartiles", "mode_purina_BCS_5y_prior_endpoint", "mode_BCS_classify", "half_anti_microbials_lifetime", "half_dietary_supplements_lifetime", 
             "half_enteric_lifetime", "half_fluid_metabolites_lifetime", "half_hormones_and_related_lifetime", "half_immunologicals_lifetime", 
             "half_misc_lifetime", "half_neurological_lifetime", "half_anti_parasites_lifetime", "half_anti_septics_lifetime", 
             "half_cardio_respiratory_lifetime", "half_diuretics_lifetime", "half_anti_hist_lifetime", 
             "quartile_anti_inflam_lifetime", "quartile_anti_microbials_lifetime", "quartile_dietary_supplements_lifetime", 
            #not enough data for quartiles "quartile_enteric_lifetime", "quartile_fluid_metabolites_lifetime", 
 "quartile_hormones_and_related_lifetime", 
             "quartile_immunologicals_lifetime", "quartile_misc_lifetime", "quartile_neurological_lifetime", 
             "quartile_anti_parasites_lifetime", 
 #not enough data for quartiles "quartile_anti_septics_lifetime", "quartile_cardio_respiratory_lifetime",   "quartile_diuretics_lifetime", "quartile_anti_hist_lifetime", 
             "half_anti_inflam_5y_prior_endpoint", "half_anti_microbials_5y_prior_endpoint", "half_enteric_5y_prior_endpoint", 
             "half_immunologicals_5y_prior_endpoint", "half_misc_5y_prior_endpoint", "half_neurological_5y_prior_endpoint", 
             "half_anti_parasites_5y_prior_endpoint", "half_dietary_supplements_5y_prior_endpoint", 
             "half_anti_septics_5y_prior_endpoint", "half_hormones_and_related_5y_prior_endpoint", 
             "half_cardio_respiratory_5y_prior_endpoint", "half_fluid_metabolites_5y_prior_endpoint", 
             "half_diuretics_5y_prior_endpoint", "half_anti_hist_5y_prior_endpoint", 
             "quartile_anti_inflam_5y_prior_endpoint", "quartile_anti_microbials_5y_prior_endpoint", 
            # "quartile_enteric_5y_prior_endpoint",
 "quartile_immunologicals_5y_prior_endpoint", 
             "quartile_misc_5y_prior_endpoint", "quartile_neurological_5y_prior_endpoint", 
             "quartile_anti_parasites_5y_prior_endpoint", "quartile_dietary_supplements_5y_prior_endpoint", 
       #      "quartile_anti_septics_5y_prior_endpoint", 
 "quartile_hormones_and_related_5y_prior_endpoint", 
          #   "quartile_cardio_respiratory_5y_prior_endpoint", #"quartile_fluid_metabolites_5y_prior_endpoint", 
         #    "quartile_diuretics_5y_prior_endpoint", "quartile_anti_hist_5y_prior_endpoint", 
             #"quartile_anti_viral_5y_prior_endpoint",
 "age_at_final_date_quartiles","age_at_neuter_years_quartiles", "had_HSA")

cohort_abr2<- cohort_abr2%>%
  mutate_all(~factor(ifelse(is.na(.), "not_recorded", as.character(.))))
univar_logistic_regression(cohort_abr2, "had_HSA", "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_log_reg_univar.csv")
```

Potentially significant univariable variables:

```{r}
lrt_pvalues <-read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_log_reg_univar.csv")
multivar_variables <- lrt_pvalues %>%
  filter(LR_T_pvalue <= 0.2) %>%
  distinct(predictor)
cols_to_keep = multivar_variables$predictor
cols_to_keep <- append(cols_to_keep, "had_HSA")
cohort_abr_sig <- cohort_abr2[,cols_to_keep]


```

## Multivariable logistic regression

Assessing collinearity of potentially sig univariable variables:

```{r}
source("C:/Users/ctaylor18/GitHub/HSA-VC-enviro-analyses/functions/corr_matrix_creation.R")
#use sig variables list generated above

create_corr_matrix(cohort_abr_sig,c(1:110), "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_logreg.csv", "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_logreg2.csv","C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_logreg_corrvalue.csv","C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_logreg.png")
```

Correlated variables

```{r}
corr_values <- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/hsa_corr_matrix_quartiles_logreg_corrvalue.csv")

#ditch rows where variable 1 and 2 values are same as in another row :
corr_values2 <- corr_values %>%
  dplyr::select(-c(X))

corr_values3 <- corr_values2 %>%
  mutate(Var1_sorted = pmin(Variable1, Variable2),  
         Var2_sorted = pmax(Variable1, Variable2)) %>%  
  distinct(Var1_sorted, Var2_sorted, Correlation, .keep_all = TRUE) %>%  
  dplyr::select(-c(Var1_sorted, Var2_sorted))  # Remove helper columns


print(corr_values3)
```

Correlated variables (99):

-   Comorbidities

    -   5y vs lifetime

    -   comorb inflam vs chronic inflam

    -   time between comorb and endpoint and diagnosis after comorb

-   Activity

    -   activity freq and intensity for SY3+

-   Enviro

    -   hours of smoke whole life and smoke exposure YN early life

    -   for all the exposure YN variables corr between early vs whole vs rest vs 5y prior

    -   fertilizer and weeds 5y prior

-   Repro

    -   sex and spayed in heat, no preg ever, heats before neutering,age at neuter quartiles (and amongst all these variables)

    -   

-   Medications

    -   half vs quartile and lifetime vs 5y for most are correlated

    -   fluid metabolites and diuretic and antihistamines

#### Automated log reg model building

```{r}
sig_variables <- unique(multivar_variables$predictor)

# Construct formula for full model
full_formula <- as.formula(paste("had_HSA ~", paste(sig_variables, collapse = " + ")))

# Build the null model
model_null <- glm(had_HSA ~ 1, data = cohort_abr_sig, family = binomial)

# Build the full model
model_full <- glm(full_formula, data = cohort_abr_sig, family = binomial)

# Display model summary
summary(model_full)
#check VIF of full model - doesnt run as aliased coeffs
#vif_values <- car::vif(model_full)

# Stepwise selection using both directions and LRT test
step(model_null,
     scope = list(upper = model_full),
     direction = "both",
     test = "LRT",
     data = df_for_multivar)
```

Automated model final:

```{r}
model_auto <- glm(formula = had_HSA ~ age_at_final_date_quartiles + in.the.house_YN_study_years_rest_of_life + 
    avg_frequency_5y_prior_endpoint_quartiles + mode_purina_BCS_5y_prior_endpoint + 
    avg_height_5y_prior_endpoint_quartiles + cooking_fuel_secondary_mode_5yrs_prior + 
    in.the.house_YN_sum_whole_life_quartiles + quartile_anti_parasites_5y_prior_endpoint + 
    quartile_anti_parasites_lifetime + avg_weight_5y_prior_endpoint_quartiles + 
    no_pregnancy_ever + comorb_immune_mediated_chronic + number_comorbs_present_lifetime + 
    comorb_inflammatory_other + in.the.house_YN_sum_rest_of_life_quartiles + 
    mismating_ever + smoke_exposure_study_years_early_life + 
    main_lifestyle_category + outside_YN_study_years_early_life, 
    family = binomial, data = cohort_abr_sig)
summary(model_auto)
#car::vif(model_auto)

dependent ="had_HSA"
explanatory2 <- c(
  "age_at_final_date_quartiles", 
  "in.the.house_YN_study_years_rest_of_life", 
  "avg_frequency_5y_prior_endpoint_quartiles", 
  "mode_purina_BCS_5y_prior_endpoint", 
  "avg_height_5y_prior_endpoint_quartiles", 
  "cooking_fuel_secondary_mode_5yrs_prior", 
  "in.the.house_YN_sum_whole_life_quartiles", 
  "quartile_anti_parasites_5y_prior_endpoint", 
  "quartile_anti_parasites_lifetime", 
  "avg_weight_5y_prior_endpoint_quartiles", 
  "no_pregnancy_ever", 
  "comorb_immune_mediated_chronic", 
  "number_comorbs_present_lifetime", 
  "comorb_inflammatory_other", 
  "in.the.house_YN_sum_rest_of_life_quartiles", 
  "mismating_ever", 
  "smoke_exposure_study_years_early_life", 
  "main_lifestyle_category", 
  "outside_YN_study_years_early_life"
)

# only presenting factors not management factors - explanatory_2 <- c("haem_cs_present","mass_cs_present","no_cs_present","splenic_interest")
cohort_abr_sig %>%
  finalfit(dependent, explanatory2, metrics=TRUE) -> t2
knitr::kable(t2[[1]], row.names=FALSE, align=c("l", "l", "r", "r", "r", "r"))
knitr::kable(t2[[2]], row.names=FALSE, col.names="")
```

### Manual multivariable model building

from most significant to least (p\<0.2) :

-   age

```{r}

```

-   in house
