---
title: "GRLS HSA updated analysis confirmed cases"
format: html
editor: visual
---

# Variables associated with HSA in GRLS cohort - confirmed cases only

```{r}
library(tidyverse)
library(tidyverse)
library(dplyr)
library(survival)
library(ggplot2)
library(tibble)
library(kableExtra)
library(epiDisplay)
library(lme4)
#devtools::install_github("zabore/ezfun")
ezfun::set_ccf_palette("contrast")
library(lubridate)
library(ggsurvfit)
library(gtsummary)
library(tidycmprsk)
library(gtable)
#devtools::install_github("zabore/condsurv")
library(condsurv)
library(survminer)
library(cardx)
library(finalfit)
library(crosstable)
library(kableExtra)
library(officer)
library(flextable)
library(writexl)
library(readr)
#updated df from midway through the updated log reg of cohort so now should only need to remove the tier 2 and 3 dogs prior to rerun
cohort_df <- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_updated_analysis_df.csv")
#classification of dogs into tiers
cohort_info_for_tiers <- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Data/study_endpoints.csv")

cohort_info_for_tiers1<- cohort_info_for_tiers %>%
  mutate(across(where(is.character), str_to_lower))%>%
  dplyr::select(subject_id, tracked_condition,tier_of_confidence)%>%
   dplyr::filter(tier_of_confidence==1)%>%
  dplyr::filter(str_detect(tracked_condition, "hemangios"))



```

```{r}
#now remove any of the cohort had HSA = 1 where they are not in the tier_of_confidence==1 list
cohort_df_tier_1_cases <-cohort_df %>%
  filter(
    had_HSA == 0 |
      subject_id %in% cohort_info_for_tiers1$subject_id
  )


#check all removed are correctly removed
removed_cases <- cohort_df %>%
  filter(
    had_HSA == 1,
    !subject_id %in% cohort_info_for_tiers1$subject_id
  )
```

```{r}
#------------------------------------------------------------
# CONVERT CHARACTER VARIABLES TO FACTORS
#------------------------------------------------------------

# Convert all remaining character columns to factors.
# Move "not_recorded" to the final level so it is not used
# as the default reference category.

convert_and_relevel <- function(df) {
df %>%
    mutate(
      across(
        where(is.character),
        factor
      )
    ) %>%
    mutate(
      across(
        where(is.factor),
        ~ {
          if ("not_recorded" %in% levels(.x)) {
            forcats::fct_relevel(
              .x,
              "not_recorded",
              after = Inf
            )
          } else {
            .x
          }
        }
      )
    )
}


cohort_df_tier_1_cases <- convert_and_relevel(
  cohort_df_tier_1_cases
)


convert_and_relevel2 <- function(df) {
df %>%
    mutate(
      across(
        where(is.character),
        factor
      )
    ) %>%
    mutate(
      across(
        where(is.factor),
        ~ {
          if ("no_medication_records_for_this_time_period" %in% levels(.x)) {
            forcats::fct_relevel(
              .x,
              "no_medication_records_for_this_time_period",
              after = Inf
            )
          } else {
            .x
          }
        }
      )
    )
}


cohort_df_tier_1_cases <- convert_and_relevel2(
  cohort_df_tier_1_cases
)


#perform these re-levels for the whole df also as build final model same
cohort_df_all <-convert_and_relevel(cohort_df)
cohort_df_all <-convert_and_relevel2(cohort_df)

```

Running univariable models with tier 1 cases only:

```{r}
exposome_data <- cohort_df_tier_1_cases%>%
  mutate(
    had_HSA = as.integer(as.character(had_HSA))
  )

#forced covariates ie. the model will keep these in always and the exposures will be added in and out of different models
forced_covariates <- c(
  "age_at_final_date_bio",
  "sex",
  "neuter"
  #or possibly sex_status which is sex and neuter status combined
)
#exposome families
exposure_families <- list(

  #------------------------------------------------------------
  # BUILT ENVIRONMENT AND SOCIOECONOMIC
  #------------------------------------------------------------

  built_environment_socioeconomic = c(
    "MDI_quintile",
    "owner_MDI_quintile",
    "area_type_mode_5yrs_prior",
    "house_type_mode_5yrs_prior",
    "water_source_mode_5yrs_prior",
    "pipes_metal_any_mode_5yrs_prior",
    "pipes_plastic_any_mode_5yrs_prior",
    "heating_fuel_primary_mode_5yrs_prior",
    "cooking_fuel_primary_mode_5yrs_prior",
    "heating_fuel_secondary_mode_5yrs_prior",
    "cooking_fuel_secondary_mode_5yrs_prior",
    "region_mode_5y_prior"
  ),

  #------------------------------------------------------------
  # HOUSEHOLD CHEMICAL EXPOSURE
  #------------------------------------------------------------

  household_chemical_exposure = c(
    "use_aerosol_study_years_early_life",
    "use_aerosol_study_years_rest_of_life",
    "use_aerosol_study_years_whole_life",
    "use_aerosol_exposed_5yrs_prior",

    "use_air_cleaner_study_years_early_life",
    "use_air_cleaner_study_years_rest_of_life",
    "use_air_cleaner_study_years_whole_life",
    "use_air_cleaner_exposed_5yrs_prior",

    "use_hepa_filter_study_years_early_life",
    "use_hepa_filter_study_years_rest_of_life",
    "use_hepa_filter_study_years_whole_life",
    "use_hepa_filter_exposed_5yrs_prior",

    "use_moth_balls_study_years_early_life",
    "use_moth_balls_study_years_rest_of_life",
    "use_moth_balls_study_years_whole_life",
    "use_moth_balls_exposed_5yrs_prior",

    "use_incense_or_candles_study_years_early_life",
    "use_incense_or_candles_study_years_rest_of_life",
    "use_incense_or_candles_study_years_whole_life",
    "use_incense_or_candles_exposed_5yrs_prior",

    "smoke_exposure_study_years_early_life",
    "smoke_exposure_study_years_rest_of_life",
    "smoke_exposure_study_years_whole_life",
    "smoke_exposure_exposed_5yrs_prior",

    "hours_of_smoke_early_life_total_dosage_terciles",
    "hours_of_smoke_rest_of_life_total_dosage_terciles",
    "hours_of_smoke_whole_life_total_dosage_terciles",
    "hours_of_smoke_exposure_hours_5yrs_prior_terciles",

    "any_treated_weeds_study_years_early_life",
    "any_treated_weeds_study_years_rest_of_life",
    "any_treated_weeds_study_years_whole_life",
    "any_treated_weeds_exposed_5yrs_prior",

    "any_treated_insects_study_years_early_life",
    "any_treated_insects_study_years_rest_of_life",
    "any_treated_insects_study_years_whole_life",
    "any_treated_insects_exposed_5yrs_prior",

    "any_treated_fertilizer_study_years_early_life",
    "any_treated_fertilizer_study_years_rest_of_life",
    "any_treated_fertilizer_study_years_whole_life",
    "any_treated_fertilizer_exposed_5yrs_prior"
  ),

  #------------------------------------------------------------
  # LIFESTYLE / SLEEP LOCATION
  #------------------------------------------------------------

  lifestyle_sleep_location = c(
    "majority_sleep_location_lifetime",
    "main_lifestyle_category",
    "main_lifestyle_category_mode_5yrs_prior",
    "sleep_location_mode_5yrs_prior",

    "in.the.garage_YN_study_years_early_life",
    "in.the.garage_YN_study_years_rest_of_life",
    "in.the.garage_YN_study_years_whole_life",

    "in.the.house_YN_study_years_early_life",
    "in.the.house_YN_study_years_rest_of_life",
    "in.the.house_YN_study_years_whole_life",

    "outside_YN_study_years_early_life",
    "outside_YN_study_years_rest_of_life",
    "outside_YN_study_years_whole_life"

   # "in.the.house_YN_sum_early_life_quartile",
    #"in.the.house_YN_sum_rest_of_life_quartile",
    #"in.the.house_YN_sum_whole_life_quartile"
  ),

  #------------------------------------------------------------
  # DOG ACTIVITY
  #------------------------------------------------------------

  dog_activity = c(
### removing all these avg,mean,median,min and max for intensity and frequency as they were a weird attempt to make an ordinal score for activity data up to 100 but do not make sense to interpret. keeping mode across these as allows to capture the free text of owner survey
  # Overall activity
 # "mean_activity_level_overview",
  "mode_activity_level_overview",

  # Time spent exercising
  "USDA_30min",
  "activity_1h_min",
  "activity_KC_2h_min",
#  "avg_activity_duration_sy1_2",
 # "avg_activity_duration_sy1_2_quartile",

  # Intensity
  #"avg_activity_intensity_sy1_2",
  #"avg_activity_intensity_sy3",
  #"avg_intensity_5y_prior_endpoint",
  #"avg_activity_intensity_sy3_quartile",
  #"avg_intensity_5y_prior_endpoint_quartile",
  "mode_intensity_5y_prior_endpoint",

  # Frequency
  #"avg_activity_freq_sy1_2",
  #"avg_activity_freq_sy3",
  #"avg_frequency_5y_prior_endpoint",
  #"avg_activity_freq_sy1_2_quartile",
  #"avg_activity_freq_sy3_quartile",
  #"avg_frequency_5y_prior_endpoint_quartile",
  "mode_frequency_5y_prior_endpoint",

  # Whole-life summaries
  "whole_activity_mode",
  #"whole_activity_mean",
  #"whole_activity_median",
  #"whole_activity_min",
  #"whole_activity_max",

  # Early-life summaries
  "early_activity_mode",
  #"early_activity_mean",
  #"early_activity_median",
  #"early_activity_min",
  #"early_activity_max",

  # Rest-of-life summaries
  "rest_activity_mode",
  #"rest_activity_mean",
  #"rest_activity_median",
  #"rest_activity_min",
  #"rest_activity_max",

  # Five-year-prior summaries
  "prior5y_activity_mode"
 # "prior5y_activity_mean",
  #"prior5y_activity_median",
  #"prior5y_activity_min",
  #"prior5y_activity_max"

  ),

  #------------------------------------------------------------
  # BODY CONDITION / WEIGHT
  #------------------------------------------------------------

  body_condition_weight = c(
    "weight_biological",
    "weight_quartile",
    "avg_height_5y_prior_endpoint_quartiles",
   # "avg_purina_BCS_5y_prior_endpoint_quartiles", <- doesnt really make sense to do avg when its a whole number value 
    "mode_BCS_classify",
   "mode_purina_BCS_5y_prior_endpoint"
  ),

  #------------------------------------------------------------
  # REPRODUCTIVE EXPOSURES
  #------------------------------------------------------------

  reproductive_exposures = c(
    "age_at_neuter_years_quartiles",
    "early_neuter_groups",
    "mismating_ever",
    "neutered_reason",
    "neutered_ever_bred",
    "spayed_while_in_heat",
    "spayed_while_pregnant",
    "prevent_heat_ever",
    "no_pregnancy_ever",
    "heats_before_neutering_terciles",
    "preg_pre_neuter",
    "mated_pre_neuter",
    "num_breeding_before_neutering_terciles",
    "ever_bred_from"
  ),

  #------------------------------------------------------------
  # INDIVIDUAL MEDICATION FAMILIES
  #------------------------------------------------------------

  nsaids_anti_inflammatories = c(
    "NSAID_ever",
    "NSAID_5y",
    "NSAID_5y_quantiles",
    "half_anti_inflam_lifetime",
    "quartile_anti_inflam_lifetime",
    "half_anti_inflam_5y_prior_endpoint",
    "quartile_anti_inflam_5y_prior_endpoint"
  ),

  antimicrobials = c(
    "half_anti_microbials_lifetime",
    "quartile_anti_microbials_lifetime",
    "half_anti_microbials_5y_prior_endpoint",
    "quartile_anti_microbials_5y_prior_endpoint"
  ),

  dietary_supplements = c(
    "half_dietary_supplements_lifetime",
    "quartile_dietary_supplements_lifetime",
    "half_dietary_supplements_5y_prior_endpoint",
    "quartile_dietary_supplements_5y_prior_endpoint"
  ),

  enteric_medications = c(
    "half_enteric_lifetime",
    "quartile_enteric_lifetime",
    "half_enteric_5y_prior_endpoint",
    "quartile_enteric_5y_prior_endpoint"
  ),

  hormones = c(
    "half_hormones_and_related_lifetime",
    "quartile_hormones_and_related_lifetime",
    "half_hormones_and_related_5y_prior_endpoint",
    "quartile_hormones_and_related_5y_prior_endpoint"
  ),
  steroids=c(
    "steroid_ever",
    "steroid_5y"
  ),

  immunologicals = c(
    "half_immunologicals_lifetime",
    "quartile_immunologicals_lifetime",
    "half_immunologicals_5y_prior_endpoint",
    "quartile_immunologicals_5y_prior_endpoint"
  ),

  neurologicals = c(
    "half_neurological_lifetime",
    "quartile_neurological_lifetime",
    "half_neurological_5y_prior_endpoint",
    "quartile_neurological_5y_prior_endpoint"
  ),

  anti_parasites = c(
    "half_anti_parasites_lifetime",
    "quartile_anti_parasites_lifetime",
    "half_anti_parasites_5y_prior_endpoint",
    "quartile_anti_parasites_5y_prior_endpoint",
    "ecto_5y_quantiles",
    "endo_5y_quantiles",
    "fenbendazole_ever",
    "fenbendazole_5y"
  ),

  anti_septics = c(
    "half_anti_septics_lifetime",
    "quartile_anti_septics_lifetime",
    "half_anti_septics_5y_prior_endpoint",
    "quartile_anti_septics_5y_prior_endpoint"
  ),

  cardiorespiratory_medications = c(
    "half_cardio_respiratory_lifetime",
    "quartile_cardio_respiratory_lifetime",
    "half_cardio_respiratory_5y_prior_endpoint",
    "quartile_cardio_respiratory_5y_prior_endpoint"
  ),

  diuretics = c(
    "half_diuretics_lifetime",
    "quartile_diuretics_lifetime",
    "half_diuretics_5y_prior_endpoint",
    "quartile_diuretics_5y_prior_endpoint"
  ),

  antihistamines = c(
    "half_anti_hist_lifetime",
    "quartile_anti_hist_lifetime",
    "half_anti_hist_5y_prior_endpoint",
    "quartile_anti_hist_5y_prior_endpoint"
  ),

  antivirals = c(
    "half_anti_viral_lifetime",
    "quartile_anti_viral_lifetime",
    "half_anti_viral_5y_prior_endpoint",
    "quartile_anti_viral_5y_prior_endpoint"
  ),

  #------------------------------------------------------------
  # COMORBIDITIES: OVERALL
  #------------------------------------------------------------

  comorbidities = c(
    "chronic_inflam_present_lifetime",
    "number_comorbs_present_lifetime",
    "chronic_inflam_present_5y",
    "number_comorbs_present_5y",
    "chronic_inflam_present_no_min_year",
    "number_comorbs_present_no_min_year",
    "lifelong_comorb_present.x",
    "number_lifelong_comorbs_present"
  ),

  #------------------------------------------------------------
  # INDIVIDUAL COMORBIDITY FAMILIES
  #------------------------------------------------------------

  cardiovascular_comorbidity = c(
    "cardiovascular",
    "lifelong_cardiovascular",
    "comorb_cardiovascular_all"
  ),

  orthopaedic_comorbidity = c(
    "orthopaedic",
    "lifelong_orthopaedic",
    "comorb_orthopaedic_all"
  ),

  immune_mediated_comorbidity = c(
    "immune_mediated",
    "lifelong_immune_mediated",
    "comorb_immune_med_all"
  ),

  chronic_inflammatory_comorbidity = c(
    "chronic_inflammatory",
    "lifelong_chronic_inflammatory",
    "lifelong_inflammatory_other",
    "comorb_chronic_inflam_all"
  ),

  #------------------------------------------------------------
  # PARASITISM
  #------------------------------------------------------------

  parasitism = c(
    "intestinal_parasitism",
    "tick_borne_parasitism",
    "other_parasitism",
    "all_parasitism"
  )
)
```

```{r}
exposure_families_requested <- exposure_families

missing_family_variables <- imap_dfr(
  exposure_families_requested,
  function(variables, family) {

    missing <- setdiff(
      variables,
      names(exposome_data)
    )

    tibble(
      exposure_family = family,
      missing_variable = missing
    )
  }
)

missing_family_variables
```

```{r}
analysis_df <-cohort_df_tier_1_cases
outcome_var <- "had_HSA"
requested_exposure_families <- exposure_families
missing_exposure_variables <- imap_dfr(
  requested_exposure_families,
  function(variables, family) {

    missing_variables <- setdiff(
      variables,
      names(analysis_df)
    )

    tibble(
      exposure_family = family,
      variable = missing_variables
    )
  }
)
exposure_families <- map(
  requested_exposure_families,
  ~ intersect(.x, names(analysis_df))
)
exposure_families <- exposure_families[
  lengths(exposure_families) > 0
]
tibble(
  exposure_family = names(exposure_families),
  number_variables = lengths(exposure_families)
)
```

```{r}
screen_exposure <- function(
    data,
    exposure,
    outcome = "had_HSA"
) {

  x <- data[[exposure]]
  y <- data[[outcome]]

  complete <- !is.na(x) & !is.na(y)

  x_complete <- x[complete]
  y_complete <- y[complete]

  tibble(
    exposure = exposure,
    variable_class = paste(class(x), collapse = "/"),
    n_total = nrow(data),
    n_observed = sum(!is.na(x)),
    n_missing = sum(is.na(x)),
    percent_missing = round(
      100 * mean(is.na(x)),
      1
    ),
    n_complete_outcome_exposure = sum(complete),
    cases_complete = sum(y_complete == 1),
    controls_complete = sum(y_complete == 0),
    n_unique_observed = n_distinct(x_complete),
    suitable_for_model = (
      sum(complete) > 0 &&
      n_distinct(x_complete) >= 2 &&
      n_distinct(y_complete) == 2
    )
  )
}
```

```{r}
exposure_screening <- imap_dfr(
  exposure_families,
  function(variables, family) {

    map_dfr(
      variables,
      ~ screen_exposure(
        data = analysis_df,
        exposure = .x,
        outcome = outcome_var
      )
    ) %>%
      mutate(
        exposure_family = family,
        .before = 1
      )
  }
)

exposure_screening %>%
  filter(!suitable_for_model)
```

```{r}
usable_exposure_families <- split(
  exposure_screening$exposure[
    exposure_screening$suitable_for_model
  ],
  exposure_screening$exposure_family[
    exposure_screening$suitable_for_model
  ]
)
```

```{r}
#ensure the numeric activity columns are not being turned back into factors again!
# Identify every column containing "mean" or "avg" in its name
mean_avg_variables <- names(analysis_df)[
  stringr::str_detect(
    names(analysis_df),
    stringr::regex("mean|avg", ignore_case = TRUE)
  ) &
  !stringr::str_detect(
    names(analysis_df),
    stringr::regex(
      "quart|quantile|tercile|categor|class",
      ignore_case = TRUE
    )
  )
]

# Review which columns have been selected
mean_avg_variables


# Convert selected columns from factor/character to numeric
#
# "not_recorded" and other non-numeric values become NA.
# factor -> character -> numeric prevents factor level numbers
# being used accidentally.

analysis_df <- analysis_df %>%
  dplyr::mutate(
    dplyr::across(
      dplyr::all_of(mean_avg_variables),
      ~ suppressWarnings(
        as.numeric(
          dplyr::na_if(
            as.character(.x),
            "not_recorded"
          )
        )
      )
    )
  )

cohort_df_all <-cohort_df_all %>%
  dplyr::mutate(
    dplyr::across(
      dplyr::all_of(mean_avg_variables),
      ~ suppressWarnings(
        as.numeric(
          dplyr::na_if(
            as.character(.x),
            "not_recorded"
          )
        )
      )
    )
  )

#------------------------------------------------------------
# CHECK THAT CONVERSION WORKED
#------------------------------------------------------------

mean_avg_variable_check <- tibble::tibble(
  variable = mean_avg_variables,

  variable_class = vapply(
    analysis_df[mean_avg_variables],
    function(x) {
      paste(
        class(x),
        collapse = "/"
      )
    },
    character(1)
  ),

  number_observed = vapply(
    analysis_df[mean_avg_variables],
    function(x) {
      sum(!is.na(x))
    },
    integer(1)
  ),

  number_missing = vapply(
    analysis_df[mean_avg_variables],
    function(x) {
      sum(is.na(x))
    },
    integer(1)
  ),

  number_unique = vapply(
    analysis_df[mean_avg_variables],
    function(x) {
      dplyr::n_distinct(
        x,
        na.rm = TRUE
      )
    },
    integer(1)
  )
)

mean_avg_variable_check


# Specifically confirm the problem variable
class(
  analysis_df$avg_activity_freq_sy3
)

is.numeric(
  analysis_df$avg_activity_freq_sy3
)

is.factor(
  analysis_df$avg_activity_freq_sy3
)

analysis_df <- analysis_df %>%
  mutate(
    age_at_final_date_bio = factor(
      age_at_final_date_bio,
      levels = c(
        ">5-<8y",
        "<5y",
        ">8-<11y",
        ">11-<13y",
        ">13y",
        "not_recorded"
      )
    )
  )
#




#



#
```

```{r}
make_exposure_counts <- function(
    data,
    exposure,
    outcome = "had_HSA"
) {

  data %>%
    transmute(
      exposure_level = as.character(.data[[exposure]]),
      outcome_value = .data[[outcome]]
    ) %>%
    mutate(
      exposure_level = replace_na(
        exposure_level,
        "missing"
      ),
      outcome_value = as.character(outcome_value)
    ) %>%
    count(
      exposure_level,
      outcome_value,
      name = "n",
      .drop = FALSE
    ) %>%
    pivot_wider(
      names_from = outcome_value,
      values_from = n,
      values_fill = 0,
      names_prefix = "outcome_"
    ) %>%
    mutate(
      exposure = exposure,
      .before = 1
    )
}
```

```{r}
exposure_level_counts <- imap_dfr(
  usable_exposure_families,
  function(variables, family) {

    map_dfr(
      variables,
      ~ make_exposure_counts(
        data = analysis_df,
        exposure = .x,
        outcome = outcome_var
      )
    ) %>%
      mutate(
        exposure_family = family,
        .before = 1
      )
  }
)
```

```{r}
make_exposure_counts <- function(
    data,
    exposure,
    outcome = "had_HSA"
) {

  # Treat genuinely continuous variables as one exposure,
  # rather than producing one count row per numeric value
  if (is.numeric(data[[exposure]])) {

    data %>%
      dplyr::summarise(
        exposure = exposure,
        exposure_level = "continuous",
        outcome_0 = sum(
          .data[[outcome]] == 0 &
            !is.na(.data[[exposure]]),
          na.rm = TRUE
        ),
        outcome_1 = sum(
          .data[[outcome]] == 1 &
            !is.na(.data[[exposure]]),
          na.rm = TRUE
        ),
        missing = sum(
          is.na(.data[[exposure]])
        )
      )

  } else {

    # Categorical variables retain one row per exposure level
    data %>%
      dplyr::transmute(
        exposure_level = as.character(
          .data[[exposure]]
        ),
        outcome_value = .data[[outcome]]
      ) %>%
      dplyr::mutate(
        exposure_level = tidyr::replace_na(
          exposure_level,
          "missing"
        ),
        outcome_value = as.character(
          outcome_value
        )
      ) %>%
      dplyr::count(
        exposure_level,
        outcome_value,
        name = "n",
        .drop = FALSE
      ) %>%
      tidyr::pivot_wider(
        names_from = outcome_value,
        values_from = n,
        values_fill = 0,
        names_prefix = "outcome_"
      ) %>%
      dplyr::mutate(
        exposure = exposure,
        missing = NA_integer_,
        .before = 1
      )
  }
}


#------------------------------------------------------------
# CREATE EXPOSURE LEVEL COUNTS
#------------------------------------------------------------

exposure_level_counts <- purrr::imap_dfr(
  usable_exposure_families,
  function(variables, family) {

    purrr::map_dfr(
      variables,
      ~ make_exposure_counts(
        data = analysis_df,
        exposure = .x,
        outcome = outcome_var
      )
    ) %>%
      dplyr::mutate(
        exposure_family = family,
        .before = 1
      )
  }
)


exposure_level_counts
```

```{r}
write_csv(
  exposure_level_counts,
  "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/exposome_exposure_level_counts_confirmed_cases.csv"
)
```

Univariable

```{r}
fit_crude_exposure_model <- function(
    data,
    exposure,
    outcome = "had_HSA"
) {

  required_variables <- c(
    outcome,
    exposure
  )

  missing_variables <- setdiff(
    required_variables,
    names(data)
  )

  if (length(missing_variables) > 0) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = NA_character_,
        n = NA_integer_,
        cases = NA_integer_,
        controls = NA_integer_,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = NA_real_,
        overall_LRT_df = NA_real_,
        overall_LRT_p = NA_real_,
        AIC = NA_real_,
        converged = FALSE,
        warning = paste(
          "Missing variable(s):",
          paste(
            missing_variables,
            collapse = ", "
          )
        )
      )
    )
  }

  model_data <- data %>%
    dplyr::select(
      dplyr::all_of(
        required_variables
      )
    ) %>%
    tidyr::drop_na()

  n_model <- nrow(model_data)

  n_cases <- sum(
    model_data[[outcome]] == 1,
    na.rm = TRUE
  )

  n_controls <- sum(
    model_data[[outcome]] == 0,
    na.rm = TRUE
  )

  outcome_variation <- dplyr::n_distinct(
    model_data[[outcome]]
  )

  exposure_variation <- dplyr::n_distinct(
    model_data[[exposure]]
  )

  if (
    n_model == 0 ||
      outcome_variation < 2 ||
      exposure_variation < 2
  ) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = NA_character_,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = NA_real_,
        overall_LRT_df = NA_real_,
        overall_LRT_p = NA_real_,
        AIC = NA_real_,
        converged = FALSE,
        warning = "Insufficient variation in outcome or exposure"
      )
    )
  }

  full_formula <- stats::reformulate(
    termlabels = exposure,
    response = outcome
  )

  null_formula <- stats::as.formula(
    paste0(outcome, " ~ 1")
  )

  warning_messages <- character()

  full_model <- tryCatch(

    withCallingHandlers(

      stats::glm(
        formula = full_formula,
        data = model_data,
        family = stats::binomial()
      ),

      warning = function(w) {

        warning_messages <<- c(
          warning_messages,
          conditionMessage(w)
        )

        invokeRestart("muffleWarning")
      }
    ),

    error = function(e) {
      e
    }
  )

  if (inherits(full_model, "error")) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = NA_character_,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = NA_real_,
        overall_LRT_df = NA_real_,
        overall_LRT_p = NA_real_,
        AIC = NA_real_,
        converged = FALSE,
        warning = conditionMessage(full_model)
      )
    )
  }

  null_model <- tryCatch(

    stats::glm(
      formula = null_formula,
      data = model_data,
      family = stats::binomial()
    ),

    error = function(e) {
      e
    }
  )

  if (inherits(null_model, "error")) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = NA_character_,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = NA_real_,
        overall_LRT_df = NA_real_,
        overall_LRT_p = NA_real_,
        AIC = stats::AIC(full_model),
        converged = full_model$converged,
        warning = conditionMessage(null_model)
      )
    )
  }

  LRT_table <- tryCatch(

    stats::anova(
      null_model,
      full_model,
      test = "LRT"
    ),

    error = function(e) {
      NULL
    }
  )

  if (is.null(LRT_table)) {

    overall_LRT_chisq <- NA_real_
    overall_LRT_df <- NA_real_
    overall_LRT_p <- NA_real_

  } else {

    overall_LRT_chisq <- LRT_table$Deviance[2]
    overall_LRT_df <- LRT_table$Df[2]
    overall_LRT_p <- LRT_table$`Pr(>Chi)`[2]
  }

  exposure_object <- model_data[[exposure]]

  if (
    is.factor(exposure_object) ||
      is.character(exposure_object)
  ) {

    factor_exposure <- factor(exposure_object)

    reference_level <- levels(
      factor_exposure
    )[1]

  } else {

    reference_level <- "Per one-unit increase"
  }

  tidy_results <- tryCatch(

    broom::tidy(
      full_model,
      conf.int = TRUE,
      exponentiate = FALSE
    ),

    error = function(e) {
      NULL
    }
  )

  if (is.null(tidy_results)) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = reference_level,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = overall_LRT_chisq,
        overall_LRT_df = overall_LRT_df,
        overall_LRT_p = overall_LRT_p,
        AIC = stats::AIC(full_model),
        converged = full_model$converged,
        warning = "Could not tidy model coefficients"
      )
    )
  }

  tidy_results <- tidy_results %>%
    dplyr::filter(
      term != "(Intercept)"
    )

  if (nrow(tidy_results) == 0) {

    return(
      tibble::tibble(
        exposure = exposure,
        term = NA_character_,
        reference_level = reference_level,
        n = n_model,
        cases = n_cases,
        controls = n_controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = overall_LRT_chisq,
        overall_LRT_df = overall_LRT_df,
        overall_LRT_p = overall_LRT_p,
        AIC = stats::AIC(full_model),
        converged = full_model$converged,
        warning = "No exposure coefficient returned"
      )
    )
  }

  final_warning <- if (
    length(warning_messages) == 0
  ) {

    NA_character_

  } else {

    paste(
      unique(warning_messages),
      collapse = "; "
    )
  }

  tidy_results %>%
    dplyr::transmute(
      exposure = exposure,
      term = term,
      reference_level = reference_level,
      n = n_model,
      cases = n_cases,
      controls = n_controls,
      OR = exp(estimate),
      CI_lower = exp(conf.low),
      CI_upper = exp(conf.high),
      wald_z = statistic,
      wald_p = p.value,
      overall_LRT_chisq = overall_LRT_chisq,
      overall_LRT_df = overall_LRT_df,
      overall_LRT_p = overall_LRT_p,
      AIC = stats::AIC(full_model),
      converged = full_model$converged,
      warning = final_warning
    )
}


```

```{r}
crude_results <- imap_dfr(
  usable_exposure_families,
  function(variables, family) {

    map_dfr(
      variables,
      ~ fit_crude_exposure_model(
        data = analysis_df,
        exposure = .x,
        outcome = outcome_var
      )
    ) %>%
      mutate(
        exposure_family = family,
        model_type = "Crude",
        .before = 1
      )
  }
)
```

```{r}
crude_results <- purrr::imap_dfr(
  usable_exposure_families,
  function(variables, family) {

    family_results <- purrr::map_dfr(
      variables,
      function(variable_name) {

        fit_crude_exposure_model(
          data = analysis_df,
          exposure = variable_name,
          outcome = outcome_var
        )
      }
    )

    family_results %>%
      dplyr::mutate(
        exposure_family = family,
        model_type = "Crude",
        .before = 1
      )
  }
)
```

```{r}
crude_model_problems <- crude_results %>%
  dplyr::filter(
    !converged |
      !is.na(warning)
  ) %>%
  dplyr::distinct(
    exposure_family,
    exposure,
    n,
    cases,
    controls,
    converged,
    warning
  ) %>%
  dplyr::arrange(
    exposure_family,
    exposure
  )
```

```{r}
crude_overall_tests <- crude_results %>%
  dplyr::distinct(
    exposure_family,
    exposure,
    n,
    cases,
    controls,
    overall_LRT_chisq,
    overall_LRT_df,
    overall_LRT_p,
    AIC,
    converged,
    warning
  )
crude_overall_tests <- crude_overall_tests %>%
  dplyr::group_by(
    exposure_family
  ) %>%
  dplyr::mutate(
    LRT_p_FDR_within_family = stats::p.adjust(
      overall_LRT_p,
      method = "BH"
    )
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    LRT_p_FDR_global = stats::p.adjust(
      overall_LRT_p,
      method = "BH"
    )
  ) %>%
  dplyr::arrange(
    exposure_family,
    overall_LRT_p
  )

crude_results <- crude_results %>%
  dplyr::left_join(
    crude_overall_tests %>%
      dplyr::select(
        exposure_family,
        exposure,
        LRT_p_FDR_within_family,
        LRT_p_FDR_global
      ),
    by = c(
      "exposure_family",
      "exposure"
    )
  )

format_p_value <- function(p) {

  dplyr::case_when(
    is.na(p) ~ NA_character_,
    p < 0.001 ~ "<0.001",
    TRUE ~ sprintf("%.3f", p)
  )
}

crude_results_formatted <- crude_results %>%
  dplyr::mutate(

    OR_95CI = dplyr::case_when(
      is.na(OR) ~ NA_character_,
      TRUE ~ sprintf(
        "%.2f (%.2f–%.2f)",
        OR,
        CI_lower,
        CI_upper
      )
    ),

    Wald_p_display = format_p_value(
      wald_p
    ),

    LRT_p_display = format_p_value(
      overall_LRT_p
    ),

    family_FDR_display = format_p_value(
      LRT_p_FDR_within_family
    ),

    global_FDR_display = format_p_value(
      LRT_p_FDR_global
    )
  ) %>%
  dplyr::select(
    exposure_family,
    exposure,
    term,
    reference_level,
    n,
    cases,
    controls,
    OR_95CI,
    OR,
    CI_lower,
    CI_upper,
    wald_z,
    wald_p,
    Wald_p_display,
    overall_LRT_chisq,
    overall_LRT_df,
    overall_LRT_p,
    LRT_p_display,
    LRT_p_FDR_within_family,
    family_FDR_display,
    LRT_p_FDR_global,
    global_FDR_display,
    AIC,
    converged,
    warning
  ) %>%
  dplyr::arrange(
    exposure_family,
    overall_LRT_p,
    exposure,
    term
  )


#write to csv
write.csv(crude_results_formatted,"C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_GRLS_log_reg_family_grouped_confirmed_cases.csv")


```

Formatted output

```{r}
format_p <- function(x, digits = 3) {
  case_when(
    is.na(x) ~ NA_character_,
    x < 0.001 ~ "<0.001",
    TRUE ~ formatC(x, format = "f", digits = digits)
  )
}

crude_results_pretty <- crude_results %>%
  mutate(
    # OR and confidence limits to 3 decimal places
    OR_display = sprintf("%.3f", OR),
    CI_lower_display = sprintf("%.3f", CI_lower),
    CI_upper_display = sprintf("%.3f", CI_upper),

    # Combined OR (95% CI)
    OR_95CI = sprintf(
      "%.3f (%.3f–%.3f)",
      OR,
      CI_lower,
      CI_upper
    ),

    # Other numerical results
    wald_z_display = sprintf("%.3f", wald_z),
    overall_LRT_chisq_display = sprintf("%.3f", overall_LRT_chisq),
    AIC_display = sprintf("%.3f", AIC),

    # P-values
    Wald_p_display = format_p(wald_p),
    LRT_p_display = format_p(overall_LRT_p),
    family_FDR_display = format_p(LRT_p_FDR_within_family),
    global_FDR_display = format_p(LRT_p_FDR_global)
  )


crude_results_pretty 
```

```{r}
#===============================================================================
# Formatting functions
#===============================================================================

format_p <- function(x, digits = 3) {
  dplyr::case_when(
    is.na(x)  ~ "",
    x < 0.001 ~ "<0.001",
    TRUE ~ formatC(
      x,
      format = "f",
      digits = digits
    )
  )
}


format_count_percent <- function(count, denominator, digits = 1) {
  dplyr::case_when(
    is.na(count) |
      is.na(denominator) |
      denominator == 0 ~ "",

    TRUE ~ sprintf(
      paste0("%d (%.", digits, "f%%)"),
      as.integer(count),
      100 * count / denominator
    )
  )
}


# Apply identical cleaning to:
# 1. coefficient categories
# 2. count-table categories
# 3. reference categories

pretty_category <- function(x) {
  x <- as.character(x)

  dplyr::case_when(
    is.na(x) ~ NA_character_,

    x == "" |
      x == "continuous" ~
      "Continuous",

    x == "no_med_use_recorded" ~
      "No medication use recorded",

    x == "no_medication_records_for_this_time_period" ~
      "No medication records for this time period",

    TRUE ~ x
  )
}



crude_model_display <- crude_results %>%
  dplyr::mutate(
    exposure = as.character(exposure),
    term = as.character(term),

    # For continuous variables, term usually equals exposure.
    # For factors, remove the exposure name from the start of term.
    Category_raw = dplyr::case_when(
      term == exposure ~ "continuous",

      stringr::str_starts(term, exposure) ~
        stringr::str_sub(
          term,
          start = stringr::str_length(exposure) + 1L
        ),

      TRUE ~ term
    ),

    Category = pretty_category(Category_raw),

    `Crude OR (95% CI)` = dplyr::case_when(
      is.na(OR) |
        is.na(CI_lower) |
        is.na(CI_upper) ~ "",

      TRUE ~ sprintf(
        "%.3f (%.3f–%.3f)",
        OR,
        CI_lower,
        CI_upper
      )
    ),

    `Wald p-value` = format_p(wald_p)
  ) %>%
  dplyr::select(
    exposure_family,
    exposure,
    Category,
    `Crude OR (95% CI)`,
    `Wald p-value`
  )



reference_lookup <- crude_results %>%
  dplyr::filter(
    !is.na(reference_level),
    reference_level != ""
  ) %>%
  dplyr::transmute(
    exposure_family,
    exposure,
    ReferenceCategory = pretty_category(reference_level)
  ) %>%
  dplyr::distinct()




reference_lookup %>%
  dplyr::count(
    exposure_family,
    exposure
  ) %>%
  dplyr::filter(n > 1)


crude_overall_p <- crude_results %>%
  dplyr::group_by(
    exposure_family,
    exposure
  ) %>%
  dplyr::summarise(
    `LRT p-value` = format_p(
      dplyr::first(
        overall_LRT_p[!is.na(overall_LRT_p)],
        default = NA_real_
      )
    ),

    `Family FDR p-value` = format_p(
      dplyr::first(
        LRT_p_FDR_within_family[
          !is.na(LRT_p_FDR_within_family)
        ],
        default = NA_real_
      )
    ),

    `Global FDR p-value` = format_p(
      dplyr::first(
        LRT_p_FDR_global[
          !is.na(LRT_p_FDR_global)
        ],
        default = NA_real_
      )
    ),

    .groups = "drop"
  )


crude_counts_display <- exposure_level_counts %>%
  dplyr::group_by(
    exposure_family,
    exposure
  ) %>%
  dplyr::mutate(
    category_original_order = dplyr::row_number(),

    total_cases = sum(
      outcome_1,
      na.rm = TRUE
    ),

    total_non_cases = sum(
      outcome_0,
      na.rm = TRUE
    ),

    `Cases (%)` = format_count_percent(
      outcome_1,
      total_cases
    ),

    `Non-cases (%)` = format_count_percent(
      outcome_0,
      total_non_cases
    ),

    Category = pretty_category(exposure_level)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::select(
    exposure_family,
    exposure,
    Category,
    category_original_order,
    `Cases (%)`,
    `Non-cases (%)`
  )


family_order_lookup <- tibble::tibble(
  exposure_family = names(exposure_families),
  family_order = seq_along(exposure_families)
)


exposure_order_lookup <- purrr::imap_dfr(
  exposure_families,
  function(variables, family_name) {
    tibble::tibble(
      exposure_family = family_name,
      exposure = variables,
      exposure_order = seq_along(variables)
    )
  }
) %>%
  dplyr::distinct(
    exposure_family,
    exposure,
    .keep_all = TRUE
  )


crude_results_table <- crude_counts_display %>%

  # Add the reference category to every row for each exposure
  dplyr::left_join(
    reference_lookup,
    by = c(
      "exposure_family",
      "exposure"
    )
  ) %>%

  # Add OR and Wald p-value to the relevant non-reference rows
  dplyr::left_join(
    crude_model_display,
    by = c(
      "exposure_family",
      "exposure",
      "Category"
    )
  ) %>%

  # Add the overall LRT and FDR p-values
  dplyr::left_join(
    crude_overall_p,
    by = c(
      "exposure_family",
      "exposure"
    )
  ) %>%

  # Add the desired family and exposure order
  dplyr::left_join(
    family_order_lookup,
    by = "exposure_family"
  ) %>%

  dplyr::left_join(
    exposure_order_lookup,
    by = c(
      "exposure_family",
      "exposure"
    )
  ) %>%

  dplyr::mutate(
    is_reference =
      !is.na(ReferenceCategory) &
      Category == ReferenceCategory,

    # Put the reference category first
    reference_order = dplyr::if_else(
      is_reference,
      0L,
      1L
    ),

    # Add the reference OR manually
    `Crude OR (95% CI)` = dplyr::case_when(
      is_reference ~ "1.000 (reference)",
      TRUE ~ dplyr::coalesce(
        `Crude OR (95% CI)`,
        ""
      )
    ),

    # Reference categories have no coefficient-level Wald p-value
    `Wald p-value` = dplyr::case_when(
      is_reference ~ "",
      TRUE ~ dplyr::coalesce(
        `Wald p-value`,
        ""
      )
    ),

    # Put unrecognised families/exposures at the end
    family_order = dplyr::coalesce(
      family_order,
      length(exposure_families) + 1L
    ),

    exposure_order = dplyr::coalesce(
      exposure_order,
      999999L
    )
  ) %>%

  dplyr::arrange(
    family_order,
    exposure_order,
    reference_order,
    category_original_order
  )



crude_results_table <- crude_results_table %>%
  dplyr::group_by(
    exposure_family,
    exposure
  ) %>%
  dplyr::mutate(
    `LRT p-value` = dplyr::if_else(
      dplyr::row_number() == 1L,
      `LRT p-value`,
      ""
    ),

    `Family FDR p-value` = dplyr::if_else(
      dplyr::row_number() == 1L,
      `Family FDR p-value`,
      ""
    ),

    `Global FDR p-value` = dplyr::if_else(
      dplyr::row_number() == 1L,
      `Global FDR p-value`,
      ""
    )
  ) %>%
  dplyr::ungroup()




#tidier ready for manuscript table
crude_results_table2 <- crude_results_table %>%
  dplyr::transmute(
    Group = exposure_family,
    Variable = exposure,
    Category,
    `Cases (%)`,
    `Non-cases (%)`,
    `Crude OR (95% CI)`,
    `Wald p-value`,
    `LRT p-value`,
    `Family FDR p-value`,
    `Global FDR p-value`
  )



crude_results_table3 <- crude_results_table2 %>%
  dplyr::mutate(
    Group_original = Group,
    Variable_original = Variable
  ) %>%

  # Display each variable name only once
  dplyr::group_by(
    Group_original,
    Variable_original
  ) %>%
  dplyr::mutate(
    Variable = dplyr::if_else(
      dplyr::row_number() == 1L,
      Variable,
      ""
    )
  ) %>%
  dplyr::ungroup() %>%

  # Display each exposure family only once
  dplyr::group_by(Group_original) %>%
  dplyr::mutate(
    Group = dplyr::if_else(
      dplyr::row_number() == 1L,
      Group,
      ""
    )
  ) %>%
  dplyr::ungroup() %>%

  dplyr::select(
    -Group_original,
    -Variable_original
  )
```

```{r}
output_folder <- "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output"

ft <- flextable(crude_results_table3) %>%
  theme_booktabs() %>%
  bold(part = "header") %>%
  fontsize(size = 8, part = "all") %>%
  align(align = "left", part = "all") %>%
  valign(valign = "top", part = "all") %>%
  autofit() %>%
  set_table_properties(
    layout = "autofit",
    width = 1
  )

doc <- read_docx() %>%
  body_add_par(
    "Crude logistic regression results",
    style = "heading 1"
  ) %>%
  body_add_flextable(ft) %>%
  body_end_section_landscape()



writexl::write_xlsx(
  list("Crude results" = crude_results_table3),
  file.path(output_folder, "crude_results_table_confirmed_cases.xlsx")
)

readr::write_excel_csv(
  crude_results_table3,
  file.path(output_folder, "crude_results_table_confirmed_cases.csv"),
  na = ""
)

print(
  doc,
  target = file.path(output_folder, "crude_results_table_confirmed_cases.docx")
)
```

#### Univariable output interpretation:

Rules:

-   if significant not just for the "not_recorded" value

-   p\<0.2 for LRT and family FDR \<0.1 to consider and then consider against variables within exposure family if multiple significant

-   Looking for consistent association across a variable e.g. in quartiles is it increasing as dose incrases etc.

**Built enviro/socioeconomic**

-   heating fuel primary 5y prior

**Household chemical exposure**

-   aerosol rest of life

-   hepa filter rest of life and 5y prior

-   candles rest of life

-   weeds rest/whole/5y prior (same thing)

-   treated insects rest of life/whole life

-   fertiliser rest whole 5y prior

-   main lifestyle 5y prior

Lifestyle sleep location

-   garage early years

-   house rest of life

Activity

-   mode activity level overview

-   mode intensity 5y prior

-   mode frequency 5y prior

-   mode whole life activity

-   rest activity mode

-   prior 5y mode activity

Repro exposures

-   early neuter groups

Medication

-   NSAID ever/5y and 5y quantiles

-   parasites half lifetime and 5y prior, quartile lifetime

-   ecto and endo 5y quantiles

Univar models with fixed covariates included

```{r}
candidate_exposure_families <- list(

  #---------------------------------------------------------------------------
  # BUILT ENVIRONMENT / SOCIOECONOMIC
  #---------------------------------------------------------------------------

  built_environment_socioeconomic = c(
    "heating_fuel_primary_mode_5yrs_prior"
  ),

  #---------------------------------------------------------------------------
  # HOUSEHOLD CHEMICAL EXPOSURE
  #---------------------------------------------------------------------------

  household_chemical_exposure = c(
    "use_aerosol_study_years_rest_of_life",
    "use_incense_or_candles_study_years_rest_of_life",
    "use_hepa_filter_study_years_rest_of_life",
    "use_hepa_filter_exposed_5yrs_prior",

    "any_treated_weeds_study_years_rest_of_life",
    "any_treated_weeds_study_years_whole_life",
    "any_treated_weeds_exposed_5yrs_prior",

    "any_treated_insects_study_years_rest_of_life",
    "any_treated_insects_study_years_whole_life",
    "any_treated_fertilizer_study_years_whole_life",
    "any_treated_fertilizer_exposed_5yrs_prior"
  ),

  #---------------------------------------------------------------------------
  # LIFESTYLE / SLEEP LOCATION
  #---------------------------------------------------------------------------

  lifestyle_sleep_location = c(
    "main_lifestyle_category_mode_5yrs_prior"
  ),

  #---------------------------------------------------------------------------
  # DOG ACTIVITY
  #---------------------------------------------------------------------------

  dog_activity = c(
    "mode_activity_level_overview",
    "mode_intensity_5y_prior_endpoint",
    "mode_frequency_5y_prior_endpoint",
    "whole_activity_mode",
    "rest_activity_mode",
    "prior5y_activity_mode"
  ),


  #---------------------------------------------------------------------------
  # NSAIDS / ANTI-INFLAMMATORIES
  #---------------------------------------------------------------------------

  nsaids_anti_inflammatories = c(
    "NSAID_ever",
    "NSAID_5y",
    "NSAID_5y_quantiles"
  ),


  anti_parasites = c(
    "half_anti_parasites_lifetime",
    "quartile_anti_parasites_lifetime",
    "half_anti_parasites_5y_prior_endpoint",
    "quartile_anti_parasites_5y_prior_endpoint",
    "ecto_5y_quantiles",
    "endo_5y_quantiles"

  )


)
candidate_variables <- unique(
  unlist(
    candidate_exposure_families,
    use.names = FALSE
  )
)

missing_candidate_variables <- setdiff(
  candidate_variables,
  names(analysis_df)
)

missing_candidate_variables
```

Running complete cases to ensure not just sig due to missing data

```{r}
missing_category_labels <- c(
  "not_recorded",
  "not recorded",
  "no_med_use_recorded",
  "no_medication_records_for_this_time_period",
  "No medication use recorded",
  "No medication records for this time period",
  "no_int_recorded",
  "no_activity_records_for_this_time_period",
  "drug_not_administered",
 "no_activity_records_for_this_time_period"

)


remove_missing_categories <- function(
    x,
    extra_missing = character(0)
) {

  missing_values <- c(
    missing_category_labels,
    extra_missing
  )

  if (is.factor(x)) {

    original_levels <- levels(x)
    x_character <- as.character(x)

    x_character[
      x_character %in% missing_values
    ] <- NA_character_

    factor(
      x_character,
      levels = setdiff(
        original_levels,
        missing_values
      )
    )

  } else if (is.character(x)) {

    x[
      x %in% missing_values
    ] <- NA_character_

    factor(x)

  } else {

    x
  }
}
```

```{r}
analysis_cc <- analysis_df %>%
  mutate(
    had_HSA = as.integer(
      as.character(had_HSA)
    )
  ) %>%
  mutate(
    across(
      all_of(candidate_variables),
      remove_missing_categories
    )
  )

#ditching entire category in neuter age
analysis_cc <- analysis_cc %>%
  mutate(
    age_at_neuter_years_quartiles =
      remove_missing_categories(
        age_at_neuter_years_quartiles,
        extra_missing = "Entire"
      )
  )
```

```{r}
fit_crude_complete_case <- function(
    data,
    exposure,
    exposure_family,
    outcome = "had_HSA"
) {

  # Select only the outcome and this particular exposure
  model_data <- data[
    ,
    c(outcome, exposure),
    drop = FALSE
  ]

  # Exposure-specific complete cases
  model_data <- model_data[
    complete.cases(model_data),
    ,
    drop = FALSE
  ]

  # Remove factor levels left empty after excluding missing values
  model_data <- droplevels(model_data)
  
  # Use treatment contrasts so categorical exposures give one OR per level
if (is.factor(model_data[[exposure]])) {
  model_data[[exposure]] <- factor(
    as.character(model_data[[exposure]]),
    levels = levels(model_data[[exposure]]),
    ordered = FALSE
  )
}

  # Ensure character exposures are factors
  if (is.character(model_data[[exposure]])) {
    model_data[[exposure]] <- factor(
      model_data[[exposure]]
    )
  }

  n_model <- nrow(model_data)

  cases <- sum(
    model_data[[outcome]] == 1,
    na.rm = TRUE
  )

  controls <- sum(
    model_data[[outcome]] == 0,
    na.rm = TRUE
  )

  # Check that model is estimable
  if (
    n_model == 0 ||
    dplyr::n_distinct(model_data[[outcome]]) < 2 ||
    dplyr::n_distinct(model_data[[exposure]]) < 2
  ) {

    return(
      tibble::tibble(
        exposure_family = exposure_family,
        exposure = exposure,
        term = NA_character_,
        n = n_model,
        cases = cases,
        controls = controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = NA_real_,
        overall_LRT_df = NA_real_,
        overall_LRT_p = NA_real_,
        converged = FALSE,
        warning = "Insufficient outcome or exposure variation"
      )
    )
  }

  full_formula <- reformulate(
    exposure,
    response = outcome
  )

  full_model <- tryCatch(
    glm(
      full_formula,
      data = model_data,
      family = binomial()
    ),
    error = function(e) e
  )

  if (inherits(full_model, "error")) {

    return(
      tibble::tibble(
        exposure_family = exposure_family,
        exposure = exposure,
        term = NA_character_,
        n = n_model,
        cases = cases,
        controls = controls,
        OR = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        wald_z = NA_real_,
        wald_p = NA_real_,
        overall_LRT_chisq = NA_real_,
        overall_LRT_df = NA_real_,
        overall_LRT_p = NA_real_,
        converged = FALSE,
        warning = conditionMessage(full_model)
      )
    )
  }

  null_model <- glm(
    as.formula(
      paste(outcome, "~ 1")
    ),
    data = model_data,
    family = binomial()
  )

  lrt <- anova(
    null_model,
    full_model,
    test = "LRT"
  )

  coefficient_table <- summary(
    full_model
  )$coefficients

  # Remove the intercept
  coefficient_table <- coefficient_table[
    rownames(coefficient_table) != "(Intercept)",
    ,
    drop = FALSE
  ]

  estimates <- coefficient_table[, "Estimate"]
  standard_errors <- coefficient_table[, "Std. Error"]

  tibble::tibble(
    exposure_family = exposure_family,
    exposure = exposure,
    term = rownames(coefficient_table),

    n = n_model,
    cases = cases,
    controls = controls,

    OR = exp(estimates),

    CI_lower = exp(
      estimates - 1.96 * standard_errors
    ),

    CI_upper = exp(
      estimates + 1.96 * standard_errors
    ),

    wald_z = coefficient_table[, "z value"],
    wald_p = coefficient_table[, "Pr(>|z|)"],

    overall_LRT_chisq = lrt$Deviance[2],
    overall_LRT_df = lrt$Df[2],
    overall_LRT_p = lrt$`Pr(>Chi)`[2],

    converged = full_model$converged,

    warning = ifelse(
      full_model$converged,
      NA_character_,
      "Model did not converge"
    )
  )
}
candidate_lookup <- purrr::imap_dfr(
  candidate_exposure_families,
  function(variables, family_name) {

    tibble::tibble(
      exposure_family = family_name,
      exposure = variables
    )
  }
)


crude_cc_results <- purrr::map2_dfr(
  candidate_lookup$exposure,
  candidate_lookup$exposure_family,

  function(exposure_name, family_name) {

    fit_crude_complete_case(
      data = analysis_cc,
      exposure = exposure_name,
      exposure_family = family_name,
      outcome = "had_HSA"
    )
  }
)

crude_cc_overall_p <- crude_cc_results %>%
  distinct(
    exposure_family,
    exposure,
    overall_LRT_p
  ) %>%
  group_by(exposure_family) %>%
  mutate(
    LRT_p_FDR_within_family = p.adjust(
      overall_LRT_p,
      method = "BH"
    )
  ) %>%
  ungroup() %>%
  mutate(
    LRT_p_FDR_global = p.adjust(
      overall_LRT_p,
      method = "BH"
    )
  )


crude_cc_results <- crude_cc_results %>%
  select(
    -any_of(
      c(
        "LRT_p_FDR_within_family",
        "LRT_p_FDR_global"
      )
    )
  ) %>%
  left_join(
    crude_cc_overall_p,
    by = c(
      "exposure_family",
      "exposure",
      "overall_LRT_p"
    )
  )

format_p <- function(x, digits = 3) {
  dplyr::case_when(
    is.na(x) ~ "",
    x < 0.001 ~ "<0.001",
    TRUE ~ formatC(
      x,
      format = "f",
      digits = digits
    )
  )
}

format_count_percent <- function(count, denominator) {
  dplyr::case_when(
    is.na(count) |
      is.na(denominator) |
      denominator == 0 ~ "",

    TRUE ~ sprintf(
      "%d (%.1f%%)",
      as.integer(count),
      100 * count / denominator
    )
  )
}

first_non_missing <- function(x) {
  x <- x[!is.na(x)]

  if (length(x) == 0) {
    NA
  } else {
    x[[1]]
  }
}

make_cc_level_counts <- function(
    data,
    exposure,
    exposure_family,
    outcome = "had_HSA"
) {

  model_data <- data[
    ,
    c(outcome, exposure),
    drop = FALSE
  ]

  model_data <- model_data[
    complete.cases(model_data),
    ,
    drop = FALSE
  ]

  model_data <- droplevels(model_data)

  x <- model_data[[exposure]]
  y <- model_data[[outcome]]

  # Convert characters, logical variables and ordered factors
  # to ordinary unordered factors
  if (is.character(x) || is.logical(x)) {
    x <- factor(x)
  }

  if (is.ordered(x)) {
    x <- factor(
      as.character(x),
      levels = levels(x),
      ordered = FALSE
    )
  }

  total_cases <- sum(y == 1, na.rm = TRUE)
  total_non_cases <- sum(y == 0, na.rm = TRUE)

  # Categorical factor exposure
  if (is.factor(x)) {

    exposure_levels <- levels(x)

    tab <- table(
      factor(x, levels = exposure_levels),
      factor(y, levels = c(0, 1))
    )

    return(
      tibble::tibble(
        exposure_family = exposure_family,
        exposure = exposure,
        Category = exposure_levels,
        category_order = seq_along(exposure_levels),
        model_type = "factor",
        reference_level = exposure_levels[[1]],
        comparison_level = NA_character_,
        outcome_0 = as.integer(tab[, "0"]),
        outcome_1 = as.integer(tab[, "1"]),
        total_cases = total_cases,
        total_non_cases = total_non_cases
      )
    )
  }

  distinct_values <- sort(unique(x))

  # Numeric binary exposure such as 0/1
  if (
    is.numeric(x) &&
      length(distinct_values) == 2
  ) {

    x_factor <- factor(
      x,
      levels = distinct_values
    )

    tab <- table(
      x_factor,
      factor(y, levels = c(0, 1))
    )

    return(
      tibble::tibble(
        exposure_family = exposure_family,
        exposure = exposure,
        Category = as.character(distinct_values),
        category_order = seq_along(distinct_values),
        model_type = "binary_numeric",
        reference_level =
          as.character(distinct_values[[1]]),
        comparison_level =
          as.character(distinct_values[[2]]),
        outcome_0 = as.integer(tab[, "0"]),
        outcome_1 = as.integer(tab[, "1"]),
        total_cases = total_cases,
        total_non_cases = total_non_cases
      )
    )
  }

  # Continuous numeric exposure
  tibble::tibble(
    exposure_family = exposure_family,
    exposure = exposure,
    Category = "Continuous",
    category_order = 1L,
    model_type = "continuous",
    reference_level = NA_character_,
    comparison_level = NA_character_,
    outcome_0 = total_non_cases,
    outcome_1 = total_cases,
    total_cases = total_cases,
    total_non_cases = total_non_cases
  )
}


cc_level_counts <- purrr::map2_dfr(
  candidate_lookup$exposure,
  candidate_lookup$exposure_family,

  function(exposure_name, family_name) {
    make_cc_level_counts(
      data = analysis_cc,
      exposure = exposure_name,
      exposure_family = family_name,
      outcome = "had_HSA"
    )
  }
) %>%
  dplyr::mutate(
    `Cases (%)` = format_count_percent(
      outcome_1,
      total_cases
    ),

    `Non-cases (%)` = format_count_percent(
      outcome_0,
      total_non_cases
    )
  )


cc_exposure_metadata <- cc_level_counts %>%
  dplyr::distinct(
    exposure_family,
    exposure,
    model_type,
    reference_level,
    comparison_level
  )

crude_cc_model_display <- crude_cc_results %>%
  dplyr::left_join(
    cc_exposure_metadata,
    by = c(
      "exposure_family",
      "exposure"
    )
  ) %>%
  dplyr::mutate(
    Category = dplyr::case_when(
      model_type == "continuous" ~
        "Continuous",

      model_type == "binary_numeric" ~
        comparison_level,

      model_type == "factor" &
        !is.na(term) &
        startsWith(term, exposure) ~
        substring(
          term,
          nchar(exposure) + 1L
        ),

      TRUE ~ term
    ),

    `Crude OR (95% CI)` = dplyr::case_when(
      is.na(OR) |
        is.na(CI_lower) |
        is.na(CI_upper) ~ "",

      TRUE ~ sprintf(
        "%.3f (%.3f–%.3f)",
        OR,
        CI_lower,
        CI_upper
      )
    ),

    `Wald p-value` = format_p(wald_p)
  ) %>%
  dplyr::select(
    exposure_family,
    exposure,
    Category,
    `Crude OR (95% CI)`,
    `Wald p-value`
  )

crude_cc_overall_display <- crude_cc_results %>%
  dplyr::group_by(
    exposure_family,
    exposure
  ) %>%
  dplyr::summarise(
    `Model n` = first_non_missing(n),

    `LRT p-value` = format_p(
      first_non_missing(overall_LRT_p)
    ),

    `Family FDR p-value` = format_p(
      first_non_missing(
        LRT_p_FDR_within_family
      )
    ),

    `Global FDR p-value` = format_p(
      first_non_missing(
        LRT_p_FDR_global
      )
    ),

    .groups = "drop"
  )
```

```{r}
crude_cc_results_table <- cc_level_counts %>%
  dplyr::left_join(
    crude_cc_model_display,
    by = c(
      "exposure_family",
      "exposure",
      "Category"
    )
  ) %>%
  dplyr::left_join(
    crude_cc_overall_display,
    by = c(
      "exposure_family",
      "exposure"
    )
  ) %>%
  dplyr::mutate(
    is_reference =
      !is.na(reference_level) &
      Category == reference_level,

    `Crude OR (95% CI)` = dplyr::case_when(
      is_reference ~ "1.000 (reference)",

      TRUE ~ dplyr::coalesce(
        `Crude OR (95% CI)`,
        ""
      )
    ),

    `Wald p-value` = dplyr::case_when(
      is_reference ~ "",

      TRUE ~ dplyr::coalesce(
        `Wald p-value`,
        ""
      )
    )
  )


family_order_lookup <- tibble::tibble(
  exposure_family =
    names(candidate_exposure_families),

  family_order =
    seq_along(candidate_exposure_families)
)

exposure_order_lookup <- purrr::imap_dfr(
  candidate_exposure_families,

  function(variables, family_name) {
    tibble::tibble(
      exposure_family = family_name,
      exposure = variables,
      exposure_order = seq_along(variables)
    )
  }
)

crude_cc_results_table <- crude_cc_results_table %>%
  dplyr::left_join(
    family_order_lookup,
    by = "exposure_family"
  ) %>%
  dplyr::left_join(
    exposure_order_lookup,
    by = c(
      "exposure_family",
      "exposure"
    )
  ) %>%
  dplyr::arrange(
    family_order,
    exposure_order,
    category_order
  )


crude_cc_results_table <- crude_cc_results_table %>%
  dplyr::group_by(
    exposure_family,
    exposure
  ) %>%
  dplyr::mutate(
    `Model n` = dplyr::if_else(
      dplyr::row_number() == 1L,
      as.character(`Model n`),
      ""
    ),

    `LRT p-value` = dplyr::if_else(
      dplyr::row_number() == 1L,
      `LRT p-value`,
      ""
    ),

    `Family FDR p-value` = dplyr::if_else(
      dplyr::row_number() == 1L,
      `Family FDR p-value`,
      ""
    ),

    `Global FDR p-value` = dplyr::if_else(
      dplyr::row_number() == 1L,
      `Global FDR p-value`,
      ""
    )
  ) %>%
  dplyr::ungroup()


crude_cc_results_table <- crude_cc_results_table %>%
  dplyr::transmute(
    Group = exposure_family,
    Variable = exposure,
    Category,
    `Model n`,
    `Cases (%)`,
    `Non-cases (%)`,
    `Crude OR (95% CI)`,
    `Wald p-value`,
    `LRT p-value`,
    `Family FDR p-value`,
    `Global FDR p-value`
  )


crude_cc_results_table
```

```{r}
output_folder <- "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output"

readr::write_excel_csv(
  crude_cc_results_table,
  file.path(
    output_folder,
    "crude_complete_case_results_table_confirmed_cases.csv"
  ),
  na = ""
)



ft_cc <- flextable(
  crude_cc_results_table
) %>%
  theme_booktabs() %>%
  bold(part = "header") %>%
  fontsize(size = 7, part = "all") %>%
  valign(valign = "top", part = "all") %>%
  align(
    j = c("Group", "Variable", "Category"),
    align = "left",
    part = "all"
  ) %>%
  align(
    j = c(
      "Model n",
      "Cases (%)",
      "Non-cases (%)",
      "Crude OR (95% CI)",
      "Wald p-value",
      "LRT p-value",
      "Family FDR p-value",
      "Global FDR p-value"
    ),
    align = "center",
    part = "all"
  ) %>%
  autofit() %>%
  set_table_properties(
    layout = "autofit",
    width = 1,
    opts_word = list(
      split = TRUE
    )
  )

doc_cc <- read_docx() %>%
  body_add_par(
    "Complete-case crude logistic regression results",
    style = "heading 1"
  ) %>%
  body_add_flextable(ft_cc) %>%
  body_end_section_landscape()


print(
  doc_cc,
  target = file.path(
    output_folder,
    "crude_complete_case_results_table_confirmed_cases.docx"
  )
)
```

From this take all of the candidate exposure list forward into adjusted models:

```{r}
forced_covariates <- c(
  "age_at_final_date_bio",
  "sex",
  "neuter"
)

adjusted_candidate_lookup <- purrr::imap_dfr(
  candidate_exposure_families,
  function(variables, family_name) {
    tibble::tibble(
      exposure_family = family_name,
      exposure = variables
    )
  }
)

required_names <- c(
  "had_HSA",
  forced_covariates,
  candidate_lookup$exposure
)

missing_names <- setdiff(
  unique(required_names),
  names(analysis_df)
)

missing_names
```

```{r}
extract_vif_diagnostics <- function(model, exposure) {

  if (!requireNamespace("car", quietly = TRUE)) {
    return(
      list(
        exposure_GVIF = NA_real_,
        exposure_GVIF_df = NA_real_,
        exposure_GVIF_adjusted = NA_real_,
        max_GVIF_adjusted = NA_real_
      )
    )
  }

  vif_object <- tryCatch(
    car::vif(model),
    error = function(e) NULL
  )

  if (is.null(vif_object)) {
    return(
      list(
        exposure_GVIF = NA_real_,
        exposure_GVIF_df = NA_real_,
        exposure_GVIF_adjusted = NA_real_,
        max_GVIF_adjusted = NA_real_
      )
    )
  }

  if (is.matrix(vif_object)) {

    terms <- rownames(vif_object)
    raw_gvif <- vif_object[, "GVIF"]
    gvif_df <- vif_object[, "Df"]

    adjusted_column <- grep(
      "1/(2",
      colnames(vif_object),
      fixed = TRUE
    )

    if (length(adjusted_column) > 0) {
      adjusted_gvif <- vif_object[, adjusted_column[[1]]]
    } else {
      adjusted_gvif <- raw_gvif^(1 / (2 * gvif_df))
    }

  } else {

    terms <- names(vif_object)
    raw_gvif <- as.numeric(vif_object)
    gvif_df <- rep(1, length(raw_gvif))

    # Comparable scale to GVIF^(1/(2*df))
    adjusted_gvif <- sqrt(raw_gvif)
  }

  exposure_position <- match(
    exposure,
    terms
  )

  maximum_adjusted <- if (
    length(adjusted_gvif) == 0 ||
      all(is.na(adjusted_gvif))
  ) {
    NA_real_
  } else {
    max(adjusted_gvif, na.rm = TRUE)
  }

  list(
    exposure_GVIF = if (is.na(exposure_position)) {
      NA_real_
    } else {
      unname(raw_gvif[exposure_position])
    },

    exposure_GVIF_df = if (is.na(exposure_position)) {
      NA_real_
    } else {
      unname(gvif_df[exposure_position])
    },

    exposure_GVIF_adjusted = if (is.na(exposure_position)) {
      NA_real_
    } else {
      unname(adjusted_gvif[exposure_position])
    },

    max_GVIF_adjusted = maximum_adjusted
  )
}

empty_adjusted_result <- function(
    exposure,
    exposure_family,
    n = 0L,
    cases = 0L,
    controls = 0L,
    warning_message = NA_character_,
    rank_deficient = NA
) {

  tibble::tibble(
    exposure_family = exposure_family,
    exposure = exposure,
    term = NA_character_,
    reference_level = NA_character_,

    n = n,
    cases = cases,
    controls = controls,

    adjusted_OR = NA_real_,
    adjusted_CI_lower = NA_real_,
    adjusted_CI_upper = NA_real_,
    adjusted_wald_z = NA_real_,
    adjusted_wald_p = NA_real_,

    adjusted_LRT_chisq = NA_real_,
    adjusted_LRT_df = NA_real_,
    adjusted_LRT_p = NA_real_,

    exposure_GVIF = NA_real_,
    exposure_GVIF_df = NA_real_,
    exposure_GVIF_adjusted = NA_real_,
    max_GVIF_adjusted = NA_real_,
    rank_deficient = rank_deficient,

    AIC = NA_real_,
    converged = FALSE,
    warning = warning_message
  )
}
```

```{r}
adjusted_single_results <- purrr::map2_dfr(
  adjusted_candidate_lookup$exposure,
  adjusted_candidate_lookup$exposure_family,

  function(exposure_name, family_name) {

    fit_adjusted_single_exposure(
      data = analysis_df,
      exposure = exposure_name,
      exposure_family = family_name,
      outcome = "had_HSA",
      forced_covariates = forced_covariates
    )
  }
)
```

```{r}
adjusted_overall_p <- adjusted_single_results %>%
  dplyr::distinct(
    exposure_family,
    exposure,
    adjusted_LRT_p
  ) %>%
  dplyr::group_by(
    exposure_family
  ) %>%
  dplyr::mutate(
    adjusted_LRT_FDR_within_family =
      p.adjust(
        adjusted_LRT_p,
        method = "BH"
      )
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    adjusted_LRT_FDR_global =
      p.adjust(
        adjusted_LRT_p,
        method = "BH"
      )
  )

adjusted_single_results <- adjusted_single_results %>%
  dplyr::left_join(
    adjusted_overall_p,
    by = c(
      "exposure_family",
      "exposure",
      "adjusted_LRT_p"
    )
  )
```

```{r}
format_p <- function(x, digits = 3) {
  dplyr::case_when(
    is.na(x) ~ "",
    x < 0.001 ~ "<0.001",
    TRUE ~ formatC(
      x,
      format = "f",
      digits = digits
    )
  )
}

adjusted_exposure_summary <- adjusted_single_results %>%
  dplyr::distinct(
    exposure_family,
    exposure,
    n,
    cases,
    controls,
    adjusted_LRT_chisq,
    adjusted_LRT_df,
    adjusted_LRT_p,
    adjusted_LRT_FDR_within_family,
    adjusted_LRT_FDR_global,
    exposure_GVIF_adjusted,
    max_GVIF_adjusted,
    rank_deficient,
    AIC,
    converged,
    warning
  ) %>%
  dplyr::mutate(
    `Adjusted LRT p` =
      format_p(adjusted_LRT_p),

    `Family FDR p` =
      format_p(
        adjusted_LRT_FDR_within_family
      ),

    `Global FDR p` =
      format_p(
        adjusted_LRT_FDR_global
      )
  ) %>%
  dplyr::arrange(
    adjusted_LRT_p
  )

View(adjusted_exposure_summary)

adjusted_exposure_summary %>%
  dplyr::filter(
    rank_deficient |
      !converged |
      !is.na(warning)
  )
```

```{r}
adjusted_exposure_summary %>%
  dplyr::filter(
    !is.na(exposure_GVIF_adjusted),
    exposure_GVIF_adjusted > sqrt(5)
  ) %>%
  dplyr::arrange(
    dplyr::desc(
      exposure_GVIF_adjusted
    )
  )
```

```{r}

```

```{r}

```

```{r}
adjusted_results_table <-
  adjusted_results_table %>%

  group_by(
    exposure_family,
    exposure
  ) %>%

  mutate(
    `Model n` = if_else(
      row_number() == 1L,
      as.character(model_n),
      ""
    ),

    `Adjusted LRT p-value` =
      if_else(
        row_number() == 1L,
        `Adjusted LRT p-value`,
        ""
      ),

    `Family FDR p-value` =
      if_else(
        row_number() == 1L,
        `Family FDR p-value`,
        ""
      ),

    `Global FDR p-value` =
      if_else(
        row_number() == 1L,
        `Global FDR p-value`,
        ""
      ),

    `Exposure adjusted GVIF` =
      if_else(
        row_number() == 1L,
        as.character(
          `Exposure adjusted GVIF`
        ),
        ""
      ),

    `Maximum adjusted GVIF` =
      if_else(
        row_number() == 1L,
        as.character(
          `Maximum adjusted GVIF`
        ),
        ""
      ),

    Converged = if_else(
      row_number() == 1L,
      as.character(converged),
      ""
    ),

    Warning = if_else(
      row_number() == 1L,
      coalesce(warning, ""),
      ""
    )
  ) %>%

  ungroup()
adjusted_results_table <-
  adjusted_results_table %>%

  transmute(
    Group = exposure_family,
    Variable = exposure,
    Category,
    `Model n`,
    `Cases (%)`,
    `Non-cases (%)`,
    `Adjusted OR (95% CI)`,
    `Adjusted Wald p-value`,
    `Adjusted LRT p-value`,
    `Family FDR p-value`,
    `Global FDR p-value`,
    `Exposure adjusted GVIF`,
    `Maximum adjusted GVIF`,
    Converged,
    Warning
  )
adjusted_results_table <-
  adjusted_results_table %>%

  mutate(
    Group_original = Group,
    Variable_original = Variable
  ) %>%

  group_by(
    Group_original,
    Variable_original
  ) %>%

  mutate(
    Variable = if_else(
      row_number() == 1L,
      Variable,
      ""
    )
  ) %>%

  ungroup() %>%

  group_by(Group_original) %>%

  mutate(
    Group = if_else(
      row_number() == 1L,
      Group,
      ""
    )
  ) %>%

  ungroup() %>%

  select(
    -Group_original,
    -Variable_original
  )
```

```{r}
output_folder <- "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output"


dir.create(
  output_folder,
  recursive = TRUE,
  showWarnings = FALSE
)

readr::write_excel_csv(
  adjusted_results_table,
  file.path(
    output_folder,
    "adjusted_single_exposure_detailed_results_confirmed_cases.csv"
  ),
  na = ""
)
#need to format differently shorter col names etc for word

adjusted_results_word <- adjusted_results_table %>%
  dplyr::select(
    Group,
    Variable,
    Category,
    `Model n`,
    `Cases (%)`,
    `Non-cases (%)`,
    `Adjusted OR (95% CI)`,
    `Adjusted Wald p-value`,
    `Adjusted LRT p-value`,
    `Family FDR p-value`,
    `Global FDR p-value`
  )

ft_adj <- flextable::flextable(
  adjusted_results_word
) %>%

  flextable::set_header_labels(
    Group = "Group",
    Variable = "Variable",
    Category = "Category",
    `Model n` = "n",
    `Cases (%)` = "Cases (%)",
    `Non-cases (%)` = "Non-cases (%)",
    `Adjusted OR (95% CI)` = "aOR (95% CI)",
    `Adjusted Wald p-value` = "Wald p",
    `Adjusted LRT p-value` = "LRT p",
    `Family FDR p-value` = "Family FDR",
    `Global FDR p-value` = "Global FDR"
  ) %>%

  flextable::theme_booktabs() %>%

  flextable::bold(
    part = "header"
  ) %>%

  flextable::fontsize(
    size = 7.5,
    part = "all"
  ) %>%

  flextable::font(
    fontname = "Arial",
    part = "all"
  ) %>%

  flextable::valign(
    valign = "top",
    part = "all"
  ) %>%

  flextable::align(
    j = c(
      "Group",
      "Variable",
      "Category"
    ),
    align = "left",
    part = "all"
  ) %>%

  flextable::align(
    j = c(
      "Model n",
      "Cases (%)",
      "Non-cases (%)",
      "Adjusted OR (95% CI)",
      "Adjusted Wald p-value",
      "Adjusted LRT p-value",
      "Family FDR p-value",
      "Global FDR p-value"
    ),
    align = "center",
    part = "all"
  ) %>%

  # Reduce wasted space inside cells
  flextable::padding(
    padding.top = 1,
    padding.bottom = 1,
    padding.left = 2,
    padding.right = 2,
    part = "all"
  ) %>%

  # Fixed widths in inches
  flextable::width(
    j = "Group",
    width = 1.10
  ) %>%

  flextable::width(
    j = "Variable",
    width = 1.70
  ) %>%

  flextable::width(
    j = "Category",
    width = 1.35
  ) %>%

  flextable::width(
    j = "Model n",
    width = 0.45
  ) %>%

  flextable::width(
    j = "Cases (%)",
    width = 0.75
  ) %>%

  flextable::width(
    j = "Non-cases (%)",
    width = 0.85
  ) %>%

  flextable::width(
    j = "Adjusted OR (95% CI)",
    width = 1.30
  ) %>%

  flextable::width(
    j = c(
      "Adjusted Wald p-value",
      "Adjusted LRT p-value",
      "Family FDR p-value",
      "Global FDR p-value"
    ),
    width = 0.65
  ) %>%

  flextable::set_table_properties(
    layout = "fixed",
    width = 1,
    opts_word = list(
      split = TRUE
    )
  )

doc_adj <- officer::read_docx() %>%

  officer::body_add_par(
    "Core-adjusted single-exposure logistic regression results",
    style = "heading 1"
  ) %>%

  flextable::body_add_flextable(
    ft_adj
  ) %>%

  officer::body_end_section_landscape()

print(
  doc_adj,
  target = file.path(
    output_folder,
    "adjusted_single_exposure_detailed_results_confirmed_cases.docx"
  )
)
```

```{r}

```

From these adjusted models retain for multivariable:

-   Model build:

    -   prior 5y activity mode

    -   NSAID ever

    -   parasite lifetime half

    -   endo 5y quantile

```{r}
candidate_exposures <- c(
  "prior5y_activity_mode",
  "NSAID_ever",
  "half_anti_parasites_lifetime",
  "endo_5y_quantiles"
)
```

```{r}
forced_covariates <- c(
  "age_at_final_date_bio",
  "sex",
  "neuter"
)

multivar_variables <- c(
  forced_covariates,
  candidate_exposures
)

multivar_data <- analysis_df %>%
  dplyr::select(
    had_HSA,
    dplyr::all_of(multivar_variables)
  ) %>%
  dplyr::filter(
    complete.cases(.)
  ) %>%
  droplevels()


for (variable in multivar_variables) {

  if (is.ordered(multivar_data[[variable]])) {

    original_levels <- levels(
      multivar_data[[variable]]
    )

    multivar_data[[variable]] <- factor(
      as.character(multivar_data[[variable]]),
      levels = original_levels,
      ordered = FALSE
    )

  } else if (
    is.character(multivar_data[[variable]]) ||
      is.logical(multivar_data[[variable]])
  ) {

    multivar_data[[variable]] <- factor(
      multivar_data[[variable]]
    )
  }
}
```

```{r}
full_trial_model <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    prior5y_activity_mode +
    NSAID_ever +
    half_anti_parasites_lifetime,
  data = multivar_data,
  family = binomial()
)

summary(full_trial_model)
logistic.display(full_trial_model)
```

```{r}
drop1(
  full_trial_model,
  test = "LRT"
)
```

Remove parasites:

```{r}
full_trial_model2 <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    prior5y_activity_mode +
    NSAID_ever ,
  data = multivar_data,
  family = binomial()
)

summary(full_trial_model2)
logistic.display(full_trial_model2)
```

```{r}
drop1(
  full_trial_model2,
  test = "LRT"
)
```

Age activity likely confounders, exploring:

```{r}
model_without_age <- glm(
  had_HSA ~
    sex +
    neuter +
    prior5y_activity_mode +
    NSAID_ever,
  data = multivar_data,
  family = binomial()
)

model_with_age <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    prior5y_activity_mode +
    NSAID_ever,
  data = multivar_data,
  family = binomial()
)
```

```{r}
activity_terms <- grep(
  "^prior5y_activity_mode",
  names(coef(model_with_age)),
  value = TRUE
)

activity_confounding <- data.frame(
  term = activity_terms,

  OR_without_age = exp(
    coef(model_without_age)[activity_terms]
  ),

  OR_with_age = exp(
    coef(model_with_age)[activity_terms]
  ),

  percent_change_log_OR = 100 * abs(
    (
      coef(model_with_age)[activity_terms] -
        coef(model_without_age)[activity_terms]
    ) /
      coef(model_without_age)[activity_terms]
  )
)

activity_confounding
```

```{r}
terms_to_check <- c(
  grep(
    "^neuter",
    names(coef(model_with_age)),
    value = TRUE
  ),
  grep(
    "^NSAID_ever",
    names(coef(model_with_age)),
    value = TRUE
  )
)

other_age_confounding <- data.frame(
  term = terms_to_check,

  OR_without_age = exp(
    coef(model_without_age)[terms_to_check]
  ),

  OR_with_age = exp(
    coef(model_with_age)[terms_to_check]
  ),

  percent_change_log_OR = 100 * abs(
    (
      coef(model_with_age)[terms_to_check] -
        coef(model_without_age)[terms_to_check]
    ) /
      coef(model_without_age)[terms_to_check]
  )
)

other_age_confounding
```

```{r}
age_activity_table <- with(
  multivar_data,
  table(
    age_at_final_date_bio,
    prior5y_activity_mode
  )
)

age_activity_table

prop.table(
  age_activity_table,
  margin = 1
)
```

NSAID and activity

```{r}
model_without_activity <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    NSAID_ever,
  data = multivar_data,
  family = binomial()
)

model_with_activity <- glm(
  had_HSA ~
    age_at_final_date_bio +
    sex +
    neuter +
    prior5y_activity_mode +
    NSAID_ever,
  data = multivar_data,
  family = binomial()
)
```

```{r}
NSAID_terms <- grep(
  "^NSAID_ever",
  names(coef(model_with_activity)),
  value = TRUE
)

NSAID_activity_confounding <- data.frame(
  term = NSAID_terms,

  OR_without_activity = exp(
    coef(model_without_activity)[NSAID_terms]
  ),

  OR_with_activity = exp(
    coef(model_with_activity)[NSAID_terms]
  ),

  percent_change_log_OR = 100 * abs(
    (
      coef(model_with_activity)[NSAID_terms] -
        coef(model_without_activity)[NSAID_terms]
    ) /
      coef(model_without_activity)[NSAID_terms]
  )
)

NSAID_activity_confounding
```

Check activity definitely worth having in model : (yes)

```{r}
lrtest(model_with_activity,model_without_activity)
AIC(model_with_activity)
AIC(model_without_activity)
BIC(model_with_activity)
BIC(model_without_activity)
```

Check for interaction sex:neuter - significant

```{r}
final_model_sex_neuter_interaction <- update(
  model_with_activity,
  . ~ . + sex:neuter
)


anova(
  model_with_activity,
  final_model_sex_neuter_interaction,
  test = "LRT"
)

AIC(
  model_with_activity,
  final_model_sex_neuter_interaction
)

BIC(
  model_with_activity,
  final_model_sex_neuter_interaction
)

drop1(
  final_model_sex_neuter_interaction,
  test = "LRT"
)
```

Final model

```{r}


final_model_data <- model.frame(full_trial_model2) %>%
  dplyr::mutate(
    sex_neuter = interaction(
      sex,
      neuter,
      sep = " ",
      drop = TRUE
    )
  )

table(
  final_model_data$sex_neuter,
  final_model_data$had_HSA
)


final_model_data <- final_model_data %>%
  dplyr::mutate(
    sex_neuter = forcats::fct_relevel(
      sex_neuter,
      "F_Entire"
    )
  )

final_model_data <- final_model_data %>%
  mutate(
    had_HSA = factor(
      had_HSA,
      levels = c(0, 1),
      labels = c("No HSA", "HSA")
    )
  )

dependent <- "had_HSA"

explanatory <- c(
  "age_at_final_date_bio",
  "sex_neuter",
  "prior5y_activity_mode",
  "NSAID_ever"
)

final_model_table2 <- final_model_data %>%
  finalfit::finalfit(
    dependent = dependent,
    explanatory = explanatory,
    metrics = TRUE,
    column = TRUE
  ) %>%
  knitr::kable(
    format = "html",
    digits = 2,
    escape = FALSE
  ) %>%
  kableExtra::kable_styling(
    full_width = FALSE,
    bootstrap_options = c("striped", "hover")
  )

final_model_table2
```

Attempt to create forest plot of both the final all cases model and the final histopath cases model:

```{r}
library(pROC)
#recode 0 and 1 in NSAID ever to "no NSAID prescription recorded" 0 "NSAID prescription recorded" 1
final_model_data <- final_model_data %>%
  mutate(
    NSAID_ever = forcats::fct_recode(
      factor(NSAID_ever),
      "No NSAID prescription recorded" = "0",
      "NSAID prescription recorded" = "1"
    )
  )

cohort_df_all <- cohort_df_all %>%
  mutate(
    NSAID_ever = forcats::fct_recode(
      factor(NSAID_ever),
      "No NSAID prescription recorded" = "0",
      "NSAID prescription recorded" = "1"
    )
  )
model_histopath <- glm(had_HSA ~
    age_at_final_date_bio +
    sex_neuter +
    NSAID_ever       +
    prior5y_activity_mode ,
    data=final_model_data,family=binomial)

#
model_all_cases <- glm(had_HSA ~
    age_at_final_date_bio +
    sex_neuter +
    NSAID_ever       +
    any_treated_fertilizer_study_years_whole_life,
    data=cohort_df_all,family=binomial)

roc_histo <- pROC::roc(final_model_data$had_HSA, fitted(model_histopath))
roc_histo
roc_all <- pROC::roc(cohort_df_all$had_HSA, fitted(model_all_cases))
roc_all
```

```{r}
#extract forest plot info
extract_forest_model <- function(
    model,
    model_label
) {

  model_terms <- attr(
    terms(model),
    "term.labels"
  )

  model_matrix <- model.matrix(model)

  coefficient_lookup <- tibble::tibble(
    term = colnames(model_matrix),
    term_number = attr(
      model_matrix,
      "assign"
    )
  ) %>%
    dplyr::filter(
      term != "(Intercept)"
    ) %>%
    dplyr::mutate(
      predictor =
        model_terms[term_number]
    ) %>%
    dplyr::select(
      term,
      predictor
    )


  # All factor levels, including reference levels
  factor_level_lookup <- purrr::imap_dfr(
    model$xlevels,

    function(
        factor_levels,
        predictor_name
    ) {

      tibble::tibble(
        predictor = predictor_name,
        category = factor_levels,
        category_order =
          seq_along(factor_levels),
        is_reference =
          category_order == 1L
      )
    }
  )


  # Estimated non-reference coefficients
  coefficient_results <- broom::tidy(
    model
  ) %>%

    dplyr::filter(
      term != "(Intercept)"
    ) %>%

    dplyr::left_join(
      coefficient_lookup,
      by = "term"
    ) %>%

    dplyr::mutate(
      category = dplyr::case_when(

        predictor %in%
          names(model$xlevels) ~

          substring(
            term,
            nchar(predictor) + 1L
          ),

        TRUE ~
          "Per unit increase"
      )
    ) %>%

    dplyr::left_join(
      factor_level_lookup,
      by = c(
        "predictor",
        "category"
      )
    ) %>%

    dplyr::mutate(
      category_order =
        dplyr::coalesce(
          category_order,
          1L
        ),

      is_reference = FALSE,

      adjusted_OR = exp(estimate),

      CI_lower = exp(
        estimate -
          1.96 * std.error
      ),

      CI_upper = exp(
        estimate +
          1.96 * std.error
      ),

      p_value = p.value
    ) %>%

    dplyr::select(
      predictor,
      term,
      category,
      category_order,
      is_reference,
      adjusted_OR,
      CI_lower,
      CI_upper,
      p_value
    )


  # Synthetic rows for the factor reference categories
  reference_results <- factor_level_lookup %>%

    dplyr::filter(
      is_reference
    ) %>%

    dplyr::mutate(
      term = paste0(
        predictor,
        "__reference"
      ),

      adjusted_OR = 1,
      CI_lower = 1,
      CI_upper = 1,
      p_value = NA_real_
    ) %>%

    dplyr::select(
      predictor,
      term,
      category,
      category_order,
      is_reference,
      adjusted_OR,
      CI_lower,
      CI_upper,
      p_value
    )


  predictor_order_lookup <- tibble::tibble(
    predictor = model_terms,
    predictor_order =
      seq_along(model_terms)
  )


  dplyr::bind_rows(
    reference_results,
    coefficient_results
  ) %>%

    dplyr::left_join(
      predictor_order_lookup,
      by = "predictor"
    ) %>%

    dplyr::mutate(
      model = model_label,

      OR_text = dplyr::if_else(
        is_reference,
        "1.00 (1.00–1.00)",
        sprintf(
          "%.2f (%.2f–%.2f)",
          adjusted_OR,
          CI_lower,
          CI_upper
        )
      )
    ) %>%

    dplyr::arrange(
      predictor_order,
      category_order
    ) %>%

    dplyr::select(
      model,
      predictor,
      predictor_order,
      term,
      category,
      category_order,
      is_reference,
      adjusted_OR,
      CI_lower,
      CI_upper,
      p_value,
      OR_text
    )
}
```

```{r}
forest_all_cases <- extract_forest_model(
  model = model_all_cases,
  model_label = "All HSA cases"
)

forest_histopath <- extract_forest_model(
  model = model_histopath,
  model_label =
    "Histopath-confirmed HSA"
)
```

```{r}
all_case_predictors <- unique(
  forest_all_cases$predictor
)

histopath_predictors <- unique(
  forest_histopath$predictor
)

shared_predictors <- intersect(
  all_case_predictors,
  histopath_predictors
)

all_case_only_predictors <- setdiff(
  all_case_predictors,
  histopath_predictors
)

histopath_only_predictors <- setdiff(
  histopath_predictors,
  all_case_predictors
)

shared_predictors
all_case_only_predictors
histopath_only_predictors
```

```{r}
forest_comparison <- bind_rows(
  forest_all_cases,
  forest_histopath
) %>%
  mutate(
    section = case_when(

      predictor %in%
        shared_predictors ~
        "Shared predictors",

      predictor %in%
        all_case_only_predictors ~
        "All-case model only",

      predictor %in%
        histopath_only_predictors ~
        paste(
          "Histopath-confirmed",
          "model only"
        ),

      TRUE ~
        "Other"
    ),

    section = factor(
      section,
      levels = c(
        "Shared predictors",
        "All-case model only",
        paste(
          "Histopath-confirmed",
          "model only"
        )
      )
    ),

    row_id = paste(
      predictor,
      term,
      sep = "__"
    ),

    OR_text = sprintf(
      "%.2f (%.2f–%.2f)",
      adjusted_OR,
      CI_lower,
      CI_upper
    )
  )

predictor_labels <- c(
  age_at_final_date_bio =
    "Age",

  sex_neuter =
    "Sex and neuter status",


  NSAID_ever =
    "NSAID exposure during whole life",

  any_treated_fertilizer_study_years_whole_life =
    "Fertiliser treatment during whole life",

  prior5y_activity_mode =
    "Most frequently reported activity level during previous 5 years"
)

forest_comparison <- forest_comparison %>%
  mutate(
    predictor_label = recode(
      predictor,
      !!!predictor_labels,
      .default = predictor
    ),

    category_label = str_remove(
      term,
      fixed(predictor)
    ),

    category_label = str_replace_all(
      category_label,
      "_",
      " "
    ),

    category_label = str_squish(
      category_label
    ),

    display_label = case_when(
      category_label == "" ~
        predictor_label,

      TRUE ~ paste0(
        predictor_label,
        ": ",
        category_label
      )
    )
  )

forest_comparison %>%
  distinct(
    predictor,
    term,
    display_label
  )
```

```{r}
shared_term_check <- forest_comparison %>%
  filter(
    predictor %in%
      shared_predictors
  ) %>%
  count(
    predictor,
    term,
    name = "number_of_models"
  )

shared_term_check %>%
  filter(
    number_of_models != 2
  )
```

```{r}
all_case_row_order <- forest_all_cases %>%
  mutate(
    row_id = paste(
      predictor,
      term,
      sep = "__"
    )
  ) %>%
  pull(row_id)

histopath_row_order <- forest_histopath %>%
  mutate(
    row_id = paste(
      predictor,
      term,
      sep = "__"
    )
  ) %>%
  pull(row_id)

row_order <- unique(
  c(
    all_case_row_order,
    histopath_row_order
  )
)

forest_comparison <- forest_comparison %>%
  mutate(
    row_id = factor(
      row_id,
      levels = rev(row_order)
    )
  )
```

```{r}
row_labels <- forest_comparison %>%
  distinct(
    row_id,
    display_label
  ) %>%
  deframe()
```

```{r}
forest_comparison <- forest_comparison %>%
  mutate(
    predictor_label = recode(
      predictor,
      !!!predictor_labels,
      .default = predictor
    ),

    # Use the actual model factor level, not the coefficient term
    category_label = as.character(category),

    category_label = str_replace_all(
      category_label,
      "_",
      " "
    ),

    category_label = str_squish(
      category_label
    ),

    display_label = paste0(
      predictor_label,
      ": ",
      category_label,
      if_else(
        is_reference,
        " (reference)",
        ""
      )
    ),

    # Shared row identifier must also use the actual category
    row_id = paste(
      predictor,
      category,
      sep = "__"
    )
  )
```

Deal with infinity

```{r}
forest_comparison <- forest_comparison %>%
  mutate(
    adjusted_OR = if_else(
      term == "prior5y_activity_modenot_recorded",
      NA_real_,
      adjusted_OR
    ),
    CI_lower = if_else(
      term == "prior5y_activity_modenot_recorded",
      NA_real_,
      CI_lower
    ),
    CI_upper = if_else(
      term == "prior5y_activity_modenot_recorded",
      NA_real_,
      CI_upper
    ),
    OR_text = if_else(
      term == "prior5y_activity_modenot_recorded",
      "Not estimated",
      OR_text
    )
  )

forest_comparison <- forest_comparison %>%
  filter(term != "prior5y_activity_modenot_recorded")
```

```{r}

forest_comparison <- forest_comparison %>%
  mutate(
    unstable_estimate =
      model == "Histopath-confirmed HSA" &
      predictor == "prior5y_activity_mode" &
      category == "not_recorded",

    adjusted_OR = if_else(
      unstable_estimate,
      NA_real_,
      adjusted_OR
    ),

    CI_lower = if_else(
      unstable_estimate,
      NA_real_,
      CI_lower
    ),

    CI_upper = if_else(
      unstable_estimate,
      NA_real_,
      CI_upper
    ),

    OR_text = if_else(
      unstable_estimate,
      "Not estimable",
      OR_text
    )
  )

forest_comparison <- forest_comparison %>%
  mutate(
    text_x = case_when(
      is_reference & model == "All HSA cases" ~ 1.12,
      is_reference & model == "Histopath-confirmed HSA" ~ 1.45,

      model == "All HSA cases" ~ CI_upper * 1.08,
      model == "Histopath-confirmed HSA" ~ CI_upper * 1.30,

      TRUE ~ CI_upper * 1.08
    )
  )
#==============================================================================
# EXPLICIT PREDICTOR AND CATEGORY ORDER FOR THE PLOT
#==============================================================================

forest_comparison <- forest_comparison %>%
  mutate(
    predictor_order_plot = case_when(
      predictor == "age_at_final_date_bio" ~ 1L,
      predictor == "sex_neuter" ~ 2L,
      predictor == "NSAID_ever" ~ 3L,

      predictor ==
        "any_treated_fertilizer_study_years_whole_life" ~ 4L,

      predictor == "prior5y_activity_mode" ~ 5L,

      TRUE ~ as.integer(predictor_order)
    ),

    category_order_plot = case_when(
      predictor == "age_at_final_date_bio" ~
        match(
          as.character(category),
          c(
            ">5-<8y",
            "<5y",
            ">8-<11y",
            ">11-<13y",
            ">13y"
          )
        ),

      predictor == "sex_neuter" ~
        match(
          as.character(category),
          c(
            "F Entire",
            "F Neutered",
            "M Entire",
            "M Neutered"
          )
        ),

      predictor == "NSAID_ever" ~
        match(
          as.character(category),
          c(
            "No NSAID prescription recorded",
            "NSAID prescription recorded",
            "No medication records for this time period"
          )
        ),

      predictor ==
        "any_treated_fertilizer_study_years_whole_life" ~
        match(
          as.character(category),
          c(
            "Not within whole life",
            "Within whole life"
          )
        ),

      predictor == "prior5y_activity_mode" ~
        match(
          as.character(category),
          c(
            "Little",
            "Moderate",
            "Very active",
            "not_recorded"
          )
        ),

      TRUE ~ as.integer(category_order)
    ),

    # Fall back to the model-derived order if a category label
    # does not match one of the labels above
    category_order_plot = coalesce(
      category_order_plot,
      as.integer(category_order)
    )
  )

display_order <- forest_comparison %>%
  arrange(
    section,
    predictor_order,
    category_order
  ) %>%
  distinct(
    row_id,
    .keep_all = TRUE
  )

label_df <- display_order %>%
  arrange(
    section,
    predictor_order,
    category_order
  ) %>%
  group_by(section, predictor) %>%
  mutate(
    display_label_short = case_when(
      row_number() == 1 ~ paste0(
        predictor_label, ": ", category_label,
        if_else(is_reference, " (reference)", "")
      ),
      TRUE ~ paste0(
        "   ", category_label,
        if_else(is_reference, " (reference)", "")
      )
    )
  ) %>%
  ungroup()

row_labels <- stats::setNames(
  label_df$display_label_short,
  as.character(label_df$row_id)
)

forest_comparison <- forest_comparison %>%
  mutate(
    row_id = factor(
      as.character(row_id),
      levels = rev(
        as.character(display_order$row_id)
      )
    )
  )

row_labels <- stats::setNames(
  display_order$display_label,
  as.character(display_order$row_id)
)
```

```         
```

rejig so dont need such long bit of info before each variable level, plot sizing nonsense

```{r}
# Ensure the panel order is fixed
forest_comparison <- forest_comparison %>%
  mutate(
    section = factor(
      section,
      levels = c(
        "Shared predictors",
        "All-case model only",
        "Histopath-confirmed model only"
      )
    )
  )


# One row per category for the y-axis
display_order <- forest_comparison %>%
  arrange(
    section,
    predictor_order_plot,
    category_order_plot
  ) %>%
  distinct(
    section,
    row_id,
    .keep_all = TRUE
  ) %>%
  group_by(
    section,
    predictor
  ) %>%
  mutate(
    axis_label = if_else(
      row_number() == 1L,

      # Variable name only on the reference/first row
      paste0(
        predictor_label,
        ": ",
        category_label,
        if_else(
          is_reference,
          " (reference)",
          ""
        )
      ),

      # Subsequent rows show only the category
      paste0(
        "   ",
        category_label
      )
    )
  ) %>%
  ungroup()


# Factor order: reference first, then remaining levels
forest_comparison <- forest_comparison %>%
  mutate(
    row_id = factor(
      as.character(row_id),
      levels = rev(
        unique(
          as.character(
            display_order$row_id
          )
        )
      )
    )
  )


row_labels <- stats::setNames(
  display_order$axis_label,
  as.character(display_order$row_id)
)

#==============================================================================
# VALUES TO BE PLOTTED
#==============================================================================


x_lower <- 0.1
x_upper <- 50

forest_plot_values <- forest_comparison %>%
  filter(
    !is_reference,
    !is.na(adjusted_OR),
    !is.na(CI_lower),
    !is.na(CI_upper),
    is.finite(adjusted_OR),
    is.finite(CI_lower),
    is.finite(CI_upper)
  ) %>%
  mutate(
    OR_text = sprintf(
      "%.2f (%.2f–%.2f)",
      adjusted_OR,
      CI_lower,
      CI_upper
    ),

    # Initial desired label positions
    desired_right_x = CI_upper * 1.07,
    desired_left_x  = CI_lower / 1.07,

    # Keep anchors inside the plotting box.
    # When a label would exceed an edge, reverse its alignment.
    text_x = case_when(
      model == "All HSA cases" &
        desired_right_x <= 32 ~
        desired_right_x,

      model == "All HSA cases" ~
        47,

      model == "Histopath-confirmed HSA" &
        desired_left_x >= 0.18 ~
        desired_left_x,

      model == "Histopath-confirmed HSA" ~
        0.11
    ),

    text_hjust = case_when(
      # Normal right-hand label
      model == "All HSA cases" &
        desired_right_x <= 32 ~ 0,

      # Near right boundary: right-align inside box
      model == "All HSA cases" ~ 1,

      # Normal left-hand label
      model == "Histopath-confirmed HSA" &
        desired_left_x >= 0.18 ~ 1,

      # Near left boundary: left-align inside box
      model == "Histopath-confirmed HSA" ~ 0
    )
  )

forest_axis_rows <- display_order %>%
  transmute(
    section,
    row_id = factor(
      as.character(row_id),
      levels = levels(forest_comparison$row_id)
    )
  )
```

```{r}
# small vertical offsets for the two models
nudge_all   <- position_nudge(y = -0.16)
nudge_histo <- position_nudge(y =  0.16)

comparison_forest_plot <- ggplot() +

  # keep all rows visible, including references
  geom_blank(
    data = forest_axis_rows,
    aes(
      x = 1,
      y = row_id
    )
  ) +

  geom_vline(
    xintercept = 1,
    linetype = "dashed",
    linewidth = 0.5,
    colour = "grey45"
  ) +

  #========================================================
  # ALL HSA CASES
  #========================================================

  geom_errorbar(
    data = forest_plot_values %>%
      filter(model == "All HSA cases"),
    aes(
      y = row_id,
      xmin = CI_lower,
      xmax = CI_upper,
      colour = model
    ),
    orientation = "y",
    position = nudge_all,
    width = 0.00,
    linewidth = 0.65
  ) +

  geom_point(
    data = forest_plot_values %>%
      filter(model == "All HSA cases"),
    aes(
      x = adjusted_OR,
      y = row_id,
      colour = model,
      shape = model
    ),
    position = nudge_all,
    size = 2.8
  ) +

  geom_label(
  data = forest_plot_values %>%
    filter(model == "All HSA cases"),
  aes(
    x = text_x,
    y = row_id,
    label = OR_text,
    colour = model,
    hjust = text_hjust
  ),
  position = nudge_all,
  size = 2.4,
  show.legend = FALSE,
  fill = "white",
  label.size = NA,
  label.padding = unit(0.08, "lines")
) +

  #========================================================
  # HISTOPATH-CONFIRMED HSA
  #========================================================

  geom_errorbar(
    data = forest_plot_values %>%
      filter(model == "Histopath-confirmed HSA"),
    aes(
      y = row_id,
      xmin = CI_lower,
      xmax = CI_upper,
      colour = model
    ),
    orientation = "y",
    position = nudge_histo,
    width = 0.00,
    linewidth = 0.65
  ) +

  geom_point(
    data = forest_plot_values %>%
      filter(model == "Histopath-confirmed HSA"),
    aes(
      x = adjusted_OR,
      y = row_id,
      colour = model,
      shape = model
    ),
    position = nudge_histo,
    size = 2.8
  ) +


geom_label(
  data = forest_plot_values %>%
    filter(model == "Histopath-confirmed HSA"),
  aes(
    x = text_x,
    y = row_id,
    label = OR_text,
    colour = model,
    hjust = text_hjust
  ),
  position = nudge_histo,
  size = 2.4,
  show.legend = FALSE,
  fill = "white",
  label.size = NA,
  label.padding = unit(0.08, "lines")
)+

  facet_grid(
    section ~ .,
    scales = "free_y",
    space = "free_y",
    switch = "y"
  ) +

  scale_y_discrete(
    labels = row_labels
  ) +

  scale_x_log10(
    breaks = c(0.1, 0.25, 0.5, 1, 2, 5, 10, 20, 50),
    labels = c("0.1", "0.25", "0.5", "1", "2", "5", "10", "20", "50"),
    expand = expansion(mult = c(0.015, 0.015))
  ) +

  coord_cartesian(
    xlim = c(0.1, 50),
    clip = "on"
  ) +

  scale_colour_manual(
    values = c(
      "All HSA cases" = "#1b9e77",
      "Histopath-confirmed HSA" = "#d95f02"
    )
  ) +

  scale_shape_manual(
    values = c(
      "All HSA cases" = 16,
      "Histopath-confirmed HSA" = 17
    )
  ) +

  labs(
    title = "Adjusted associations in alternative HSA models",
    subtitle = "Shared predictors are displayed for both models; model-specific predictors are displayed once",
    x = "Adjusted odds ratio (log scale, 95% CI)",
    y = NULL,
    colour = "Final model",
    shape = "Final model"
  ) +

  theme_bw(base_size = 11) +
  theme(
    legend.position = "top",
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    strip.placement = "outside",
    strip.background = element_blank(),
    strip.text.y.left = element_text(
      angle = 0,
      face = "bold",
      hjust = 1
    ),
    axis.text.y = element_text(size = 9),
    plot.title = element_text(face = "bold"),
    plot.margin = margin(
      t = 10,
      r = 10,
      b = 10,
      l = 10
    )
  )

comparison_forest_plot
```

```{r}
output_folder <-
  "C:/Users/ctaylor18/GitHub/GRLS_analyses/Output"

dir.create(
  output_folder,
  recursive = TRUE,
  showWarnings = FALSE
)

ggsave(
  filename = file.path(
    output_folder,
    "forest_plot_all_vs_histopath_models.png"
  ),
  plot = comparison_forest_plot,
  width = 12,
  height = 8,
  units = "in",
  dpi = 300,
  bg = "white"
)
```
