---
title: "GRLS activity classifier"
format: html
editor: visual
---

## GRLS activity classifier

Packages:

```{r}
library(tidyverse)
library(ggplot2)
library(lubridate)
```

Determine whether dogs are active or sedentary by incorporating the following activity datasets:

```{r}
#activity data from study y 1 and 2
activity_sy1_2 <- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Data/activity_details_through_SY2.csv")
#check # dogs in it
act_1_2 <- unique(activity_sy1_2$subject_id)
#activity data from study y 3 + 
activity_sy3_beyond <- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Data/activity_details_SY3_beyond.csv")
act_3_plus<- unique(activity_sy3_beyond$subject_id)
#activity overview dataset
activity_overview<- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Data/activity_overview.csv")
act_over<-unique(activity_overview$subject_id)
```

-   Create an early life active vs sedentary variable (using datasets from first 2y study?/first 2y life)

-   Create rest of life active vs sedentary variable

-   Create an overall lifetime active vs sedentary variable

    -   Rather than time to cancer event just calculate until endpoint in study

Additionally use the dog details data to calculate values for first 2y of life:

```{r}
dog_enrollment<- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Data/data version Feb 2024/dog_details.csv")
```

## Early life classifier

### Visualise data

Create plots of activity data to assess what has most variance and what thresholds might make sense to use:

-   what values exist in each column

```{r}
freq_table <- activity_sy1_2 %>%
  dplyr:: select(c(-"subject_id","year_in_study","record_date")) %>%
  dplyr::select(where(is.character)) %>%  # Select character columns
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  count(column, value) %>%
  arrange(desc(n))
```

Based on this, there are in general \> several hundred values in each row for different categories except for the 'walk_reason_other_specify' which seems like quite messy free text:

```{r}
head(table(activity_sy1_2$walk_reason_other_specify))
```

Plots of:

-   Aerobic datasets involve dogs that are allowed time off leash - 50% of dogs have off-leash time so these Qs are only valid to half dataset

-   aerobic duration

```{r}
aero_duration <- activity_sy1_2 %>%
  dplyr::select(aerobic_duration) %>%
  # empty rows to recoded as not_record
   mutate(aerobic_duration= if_else(aerobic_duration == "", "not recorded", aerobic_duration))
  
# Create a long format dataframe for plotting
data_aero <- aero_duration %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value")

# Plot bar plots for each column
ggplot(data_aero, aes(x = value)) +
  geom_bar() +
  facet_wrap(~ variable, scales = "free", ncol = 5)+
  labs(title = "Aerobic duration ", x = "Category", y = "Count") 
```

-   aerobic frequency

```{r}
aero_freq <- activity_sy1_2 %>%
  dplyr::select(aerobic_frequency) %>%
  # empty rows to recoded as not_record
   mutate(aerobic_frequency= if_else(aerobic_frequency == "", "not recorded", aerobic_frequency))
  
# Create a long format dataframe for plotting
data_aero_f <- aero_freq %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value")

# Plot bar plots for each column
ggplot(data_aero_f, aes(x = value)) +
  geom_bar() +
  facet_wrap(~ variable, scales = "free", ncol = 5)+
  labs(title = "Aerobic freq ", x = "Category", y = "Count") 
```

-   aerobic pace

```{r}
aero_pace <- activity_sy1_2 %>%
  dplyr::select(aerobic_pace) %>%
  # empty rows to recoded as not_record
   mutate(aerobic_pace= if_else(aerobic_pace == "", "not recorded", aerobic_pace))
  
# Create a long format dataframe for plotting
data_aero_p <- aero_pace %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value")

# Plot bar plots for each column
ggplot(data_aero_p, aes(x = value)) +
  geom_bar() +
  facet_wrap(~ variable, scales = "free", ncol = 5)+
  labs(title = "Aerobic pace", x = "Category", y = "Count") 
```

-   fetch games frequency

```{r}
fetch_freq <- activity_sy1_2 %>%
  dplyr::select(fetch_games_frequency) %>%
  # empty rows to recoded as not_record
   mutate(fetch_games_frequency= if_else(fetch_games_frequency == "", "not recorded", fetch_games_frequency))
  
# Create a long format dataframe for plotting
data_fetch_f <- fetch_freq %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value")

# Plot bar plots for each column
ggplot(data_fetch_f, aes(x = value)) +
  geom_bar() +
  facet_wrap(~ variable, scales = "free", ncol = 5)+
  labs(title = "fetch freq ", x = "Category", y = "Count") 
```

-   walk without leash

```{r}
leash <- activity_sy1_2 %>%
  mutate(walk_without_leash=as.character(walk_without_leash))%>%
  dplyr::select(walk_without_leash) %>%
  # empty rows to recoded as not_record
   mutate(walk_without_leash= if_else(walk_without_leash == "", "not recorded", walk_without_leash))
  
# Create a long format dataframe for plotting
leash <- leash %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value")

# Plot bar plots for each column
ggplot(leash, aes(x = value)) +
  geom_bar() +
  facet_wrap(~ variable, scales = "free", ncol = 5)+
  labs(title = "walk without leash", x = "Category", y = "Count") 
```

-   walk duration

```{r}
walk_dur <- activity_sy1_2 %>%
  dplyr::select(walk_duration) %>%
  # empty rows to recoded as not_record
   mutate(walk_duration= if_else(walk_duration == "", "not recorded",walk_duration))
  
# Create a long format dataframe for plotting
walk_dur <- walk_dur %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value")

# Plot bar plots for each column
ggplot(walk_dur, aes(x = value)) +
  geom_bar() +
  facet_wrap(~ variable, scales = "free", ncol = 5)+
  labs(title = "walk duration ", x = "Category", y = "Count") 
```

-   walk pace

```{r}
walk_pace <- activity_sy1_2 %>%
  dplyr::select(walk_pace) %>%
  # empty rows to recoded as not_record
   mutate(walk_pace= if_else(walk_pace == "", "not recorded",walk_pace))
  
# Create a long format dataframe for plotting
walk_pace <- walk_pace %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value")

# Plot bar plots for each column
ggplot(walk_pace, aes(x = value)) +
  geom_bar() +
  facet_wrap(~ variable, scales = "free", ncol = 5)+
  labs(title = "walk pace ", x = "Category", y = "Count") 
```

-   walk frequency

```{r}
walk_freq <- activity_sy1_2 %>%
  dplyr::select(walk_frequency) %>%
  # empty rows to recoded as not_record
   mutate(walk_frequency= if_else(walk_frequency == "", "not recorded",walk_frequency))
  
# Create a long format dataframe for plotting
walk_freq <- walk_freq %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value")

# Plot bar plots for each column
ggplot(walk_freq, aes(x = value)) +
  geom_bar() +
  facet_wrap(~ variable, scales = "free", ncol = 5)+
  labs(title = "walk freq", x = "Category", y = "Count") 
```

-   walk reason bathroom

```{r}
bathroom <- activity_sy1_2 %>%
  mutate(walk_reason_bathroom=as.character(walk_reason_bathroom))%>%
  dplyr::select(walk_reason_bathroom) %>%
  # empty rows to recoded as not_record
   mutate(walk_reason_bathroom= if_else(walk_reason_bathroom == "", "not recorded", walk_reason_bathroom))
  
# Create a long format dataframe for plotting
bathroom<- bathroom%>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value")

# Plot bar plots for each column
ggplot(bathroom, aes(x = value)) +
  geom_bar() +
  facet_wrap(~ variable, scales = "free", ncol = 5)+
  labs(title = "walk for bathroom", x = "Category", y = "Count") 
```

-   walk general enjoyment

```{r}
general_enjoyment <- activity_sy1_2 %>%
  mutate(walk_reason_general_enjoyment=as.character(walk_reason_general_enjoyment))%>%
  dplyr::select(walk_reason_general_enjoyment) %>%
  # empty rows to recoded as not_record
   mutate(walk_reason_general_enjoyment= if_else(walk_reason_general_enjoyment == "", "not recorded", walk_reason_general_enjoyment))
  
# Create a long format dataframe for plotting
general_enjoyment<- general_enjoyment%>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value")

# Plot bar plots for each column
ggplot(general_enjoyment, aes(x = value)) +
  geom_bar() +
  facet_wrap(~ variable, scales = "free", ncol = 5)+
  labs(title = "walk for general_enjoyment", x = "Category", y = "Count") 
```

-   walk reason training

```{r}
training <- activity_sy1_2 %>%
  mutate(walk_reason_training=as.character(walk_reason_training))%>%
  dplyr::select(walk_reason_training) %>%
  # empty rows to recoded as not_record
   mutate(walk_reason_training= if_else(walk_reason_training == "", "not recorded", walk_reason_training))
  
# Create a long format dataframe for plotting
training<- training%>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value")

# Plot bar plots for each column
ggplot(training, aes(x = value)) +
  geom_bar() +
  facet_wrap(~ variable, scales = "free", ncol = 5)+
  labs(title = "walk for training", x = "Category", y = "Count") 
```

Assess if worthwhile creating combined walk + aerobic column for duration +/- intensity to not lose data on dogs that are off leash:

```{r}
#of dogs that are walked off leash how many of these dont have activity recorded in walk_duration
dogs_off_leash <- activity_sy1_2 %>%
  filter(walk_without_leash ==1)

table(dogs_off_leash$walk_duration)
table(dogs_off_leash$aerobic_duration)
table(dogs_off_leash$aerobic_pace)

#create a new column of total activity to include both lead walk and off lead activity
activity_sy1_2 <- activity_sy1_2 %>%
  mutate(total_activity_duration = paste(walk_duration,aerobic_duration,  sep = "_")) %>%
  mutate(total_activity_freq= paste(walk_frequency,aerobic_frequency,sep="_")) %>%
  mutate(total_activity_intensity =paste(walk_pace,aerobic_pace,sep="_"))


#duration recoding
  
less_than_30min <- c("_10-30 minutes", "10-30 minutes_","Less than 10 minutes_","Less than 10 minutes_Less than 10 minutes","_Less than 10 minutes","10-30 minutes_Less than 10 minutes")
#cardiac CS
between_31_60min <-c("_30-60 minutes","10-30 minutes_30-60 minutes","30-60 minutes_","Less than 10 minutes_10-30 minutes","10-30 minutes_10-30 minutes","30-60 minutes_Less than 10 minutes","Less than 10 minutes_30-60 minutes","30-60 minutes_10-30 minutes")
#respiratory CS
between_61_90min<- c("Greater than 60 minutes_","_Greater than 60 minute","10-30 minutes_Greater than 60 minutes","Greater than 60 minutes_10-30 minutes","_Greater than 60 minutes","Greater than 60 minutes_Less than 10 minutes","Greater than 60 minutes_Less than 10 minutes","Less than 10 minutes_Greater than 60 minutes")
# GI CS
between_91_120min <- c("Greater than 60 minutes_30-60 minutes","30-60 minutes_Greater than 60 minutes","30-60 minutes_30-60 minutes")
# urinary
more_than_120min<- c("Greater than 60 minutes_Greater than 60 minutes")

all_groups <- c(less_than_30min,between_31_60min,between_61_90min,
                between_91_120min,more_than_120min)

# Find any cateogories that are not in any group
duration <- setdiff(activity_sy1_2$total_activity_duration, all_groups)

# Display unclassified clinical signs
print(duration)

#frequency recoding
never <- c("Never_","Never_Less than once/month") 
less_than_weekly <- c("Less than once/month_","Less than once/month_Less than once/month", "Less than once/month_Less than once/week","Less than once/week_Less than once/month","Never_Less than once/week","Less than once/week_",
                      "Less than once/week_Less than once/week")
min_weekly<-c("Once or twice/week_Once or twice/week","Once or twice/week_","Never_Once or twice/week","Once or twice/week_Less than once/month",
              "Once or twice/week_Less than once/week","Less than once/week_Once or twice/week","Less than once/month_Once or twice/week")
      
daily<-c("Once/day_","Less than once/week_Once/day","Never_Once/day", "Once or twice/week_Once/day","Once/day_Once or twice/week","Once/day_Less than once/week",
         "Once/day_Less than once/month","Once/day_Once or twice/week",
         "Once/day_Less than once/week", "Once/day_Less than once/month",
         "Once or twice/week_Once/day","Less than once/month_Once/day"
         )
multi_daily <-c("Once/day_Once/day","Once/day_More than once a day","More than once a day_","Once or twice/week_More than once a day","Less than once/month_More than once a day","More than once a day_Once/day","Never_More than once a day","More than once a day_Less than once/month","More than once a day_Once or twice/week","Less than once/week_More than once a day","More than once a day_More than once a day",
                "More than once a day_Less than once/week")

all_groups_freq <- c(never,less_than_weekly,min_weekly,daily,multi_daily)
freq <- setdiff(activity_sy1_2$total_activity_freq, all_groups_freq)

# Display unclassified clinical signs
print(freq)
#intensity recoding
low_only <-c("_Bursts of exercise less than 30% of the time","Slow_","Slow_Bursts of exercise less than 30% of the time"
             )
moderate <-c("Average_","_Engages in moderate exercise (greater than 30% of the time but less than 60% of the time)","Brisk_","Brisk_Bursts of exercise less than 30% of the time","Average_Engages in moderate exercise (greater than 30% of the time but less than 60% of the time)","Jog_")
high <-c(" _Engages in strenuous exercise most of the time (greater than 60%)","Run_","Average_Engages in strenuous exercise most of the time (greater than 60%)","Brisk_Engages in strenuous exercise most of the time (greater than 60%)","_Engages in strenuous exercise most of the time (greater than 60%)","Jog_Engages in strenuous exercise most of the time (greater than 60%)","Run_Engages in strenuous exercise most of the time (greater than 60%)")

all_groups_int <- c(low_only,moderate,high)
int <- setdiff(activity_sy1_2$total_activity_intensity, all_groups_int)
print(int)

```

To take forward:

-   walk duration (although how does this differ from aerobic?)

-   walk pace

-   walk frequency

-   fetch freq - this could represent a high intensity ?

Ditch:

-   aerobic duration

-   aerobic freq

-   aerobic pace

    -   high proportion of not recorded across all aerobic variables (\~50%), as they need to have answered

-   walk without leash (almost equal split of Y N)

-   walk reason bathroom

-   walk general enjoyment

-   walk reason training

-   

### Giving dogs a score for the classifier

-   Convert characters eg. duration/freq to numeric

-   Convert characters into:

    -   USDA blanket \>30 min activity/dog/day recommendation - above or below this

    -   UKKC Golden retriever \>2h/day recommendation

-   Create an annual score in a classifier and an average count of activity

-   Average over the 2y of this dataset

Keeping relevant columns:

```{r}
#keeping only relevant columns
activity_sy1_2_use <- activity_sy1_2 %>%
  dplyr::select(c(subject_id,year_in_study,record_date,walk_duration,walk_pace,walk_frequency,fetch_games_frequency))


#create column for USDA activity <30min or >30 min
activity_sy1_2 <- activity_sy1_2 %>%
  mutate(USDA_30min_walk = case_when (
    walk_duration %in% c("10-30 minutes", "Less than 10 minutes") ~ "0",  walk_duration %in% c("30-60 minutes", "Greater than 60 minutes") ~ "1", 
                                 TRUE ~ "not_recorded"))

#create column for UKKC recommendaiton <2h but cannot do this one so >1h
activity_sy1_2 <- activity_sy1_2 %>%
  mutate(walk_1h_min = case_when (
    walk_duration %in% c("10-30 minutes", "Less than 10 minutes","30-60 minutes") ~ "0",  walk_duration %in% c( "Greater than 60 minutes") ~ "1", 
                                 TRUE ~ "not_recorded"))




```
