---
title: "HSA cohort data check and descriptive statistics"
format: html
editor: visual
---

# HSA cohort data check and descriptive stats

```{r}
library(tidyverse)
HSA <- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_cohort_all_RFs_data.csv")

HSA <- HSA %>%
  dplyr::select(-c(X,X,year_in_study_diagnosis_or_final_record_year.x,year_in_study_diagnosis_or_final_record_year.y,year_in_study_diagnosis_or_final_record_year.y.y))

#rename had HSA col
HSA <- HSA %>%
  rename(had_HSA = had_HSA.x.x)
```

Datachecks:

```{r}
table(HSA$had_HSA)
```

```{r}
df_types <- HSA %>%
  summarise(across(everything(), ~ class(.x)[1])) %>%
  pivot_longer(cols = everything(), names_to = "Column", values_to = "Data_Type")

print(df_types)

```

## Summary stats for columns:

### Household:

#### House location

```{r}
cross_table <- table(HSA$area_type_mode_5yrs_prior, HSA$had_HSA)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)


```

#### MDI

```{r}
cross_table <- table(HSA$had_HSA,HSA$MDI_quintile)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

#### House type

```{r}
cross_table <- table(HSA$had_HSA,HSA$house_type_mode_5yrs_prior)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

#### Water source

```{r}
cross_table <- table(HSA$had_HSA,HSA$water_source_mode_5yrs_prior)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

Primary heat source

```{r}
cross_table <- table(HSA$had_HSA,HSA$heating_fuel_primary_mode_5yrs_prior)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

#### Primary cooking fuel

```{r}
cross_table <- table(HSA$had_HSA,HSA$cooking_fuel_primary_mode_5yrs_prior)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

#### Metal pipes

```{r}
cross_table <- table(HSA$had_HSA,HSA$pipes_metal_any_mode_5yrs_prior)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

#### Plastic pipes

```{r}
cross_table <- table(HSA$had_HSA,HSA$pipes_plastic_any_mode_5yrs_prior)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

#### Region

```{r}
cross_table <- table(HSA$had_HSA,HSA$region_name_mode_5yrs_prior)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

#### State

```{r}
cross_table <- table(HSA$had_HSA,HSA$state_mode_5yrs_prior)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

### Exposures:

#### Aerosol

```{r}
cross_table <- table(HSA$had_HSA,HSA$use_aerosol_exposed_5yrs_prior)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

Air cleaner

```{r}
cross_table <- table(HSA$had_HSA,HSA$use_air_cleaner_exposed_5yrs_prior)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

#### Hepa filter

```{r}
cross_table <- table(HSA$had_HSA,HSA$use_hepa_filter_exposed_5yrs_prior)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

#### Moth balls

```{r}
cross_table <- table(HSA$had_HSA,HSA$use_moth_balls_exposed_5yrs_prior)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

#### Incense/candles

```{r}
cross_table <- table(HSA$had_HSA,HSA$use_incense_or_candles_exposed_5yrs_prior)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

#### Smoke exposure

```{r}
cross_table <- table(HSA$had_HSA,HSA$smoke_exposure_exposed_5yrs_prior)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

#### Smoke hours

-   quartiles split this

```{r}
HSA %>%
    summarize(
    mean_smoke = mean(hours_of_smoke_exposure_hours_5yrs_prior, na.rm = TRUE),
    median_smoke = median(hours_of_smoke_exposure_hours_5yrs_prior, na.rm = TRUE),
    sd_smoke = sd(hours_of_smoke_exposure_hours_5yrs_prior, na.rm = TRUE),
    min_smoke = min(hours_of_smoke_exposure_hours_5yrs_prior, na.rm = TRUE),
    max_smoke= max(hours_of_smoke_exposure_hours_5yrs_prior, na.rm = TRUE),
    n = n(),
    n_missing = sum(is.na(hours_of_smoke_exposure_hours_5yrs_prior))
  )
```

#### Insecticide around house

```{r}
cross_table <- table(HSA$had_HSA,HSA$any_treated_insects_exposed_5yrs_prior)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

#### Weed killer around house

```{r}
cross_table <- table(HSA$had_HSA,HSA$any_treated_weeds_exposed_5yrs_prior)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

#### Fertiliser around house

```{r}
cross_table <- table(HSA$had_HSA,HSA$any_treated_fertilizer_exposed_5yrs_prior)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

### Comorbidities:

#### Did dogs have a chronic comorbidity within 5y?

NB. many dogs did not have a comorbidity, this may need removing?

```{r}
cross_table <- table(HSA$had_HSA,HSA$chronic_inflam_present_5y)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

#### Number of comorbdities recorded in lifetime:

```{r}

```

#### Specific chronic comorbidities

```{r}
table(HSA$all_comorbidities_ever)
```

#### Group chronic comorbs thematically

```{r}
comorbs_unique <- unique(HSA$comorbidities_lifetime)
comorbs_unique2 <- unique(HSA$comorbidities_no_min_year)
intestinal_parasitism <- c("giardia","hookworms", "coccidia","tapeworms","whipworms","roundworms","eimeria","isospora")
tick_borne_parasitism <-c("lyme_disease","anaplasma","rocky_mountain_spotted_fever","ticks","ehrlichia")
other_parasitism <-c("fleas","fungal_infection","heartworm","ear_mites")
infectious <- c("leptospirosis","parvovirus")
orthopaedic <-c("osteoarthritis")
immune_mediated <- c("uveitis","autoimmune_thyroiditis","keratoconjunctivitis_sicca","megaesophagus","diabetes_mellitus","cushings_disease","hypothyroidism","granuloma")
cardiovascular <- c("cardiomyopathy","hypertension")
gastrointestinal <-c("chronic_colitis","pancreatitis","malabsorptive_disorder")

inflammatory_other <- c("tracheobronchitis", "hypercalcaemia","pigmentary_uveitis")

#create new columns based on these lists
# Function to check presence of any condition in a category
check_condition <- function(condition_list, condition_column) {
  sapply(condition_column, function(x) any(str_detect(x, paste(condition_list, collapse = "|"))))
}

# Create new columns
HSA <- HSA %>%
  mutate(
    comorb_intestinal_parasitism_chronic = check_condition(intestinal_parasitism, comorbidities_5y),
    comorb_tick_borne_parasitism_chronic = check_condition(tick_borne_parasitism, comorbidities_5y),
    comorb_other_parasitism_chronic = check_condition(other_parasitism, comorbidities_5y),
    comorb_infectious_chronic = check_condition(infectious, comorbidities_5y),
    comorb_orthopaedic_chronic = check_condition(orthopaedic,comorbidities_5y),
    comorb_immune_mediated_chronic = check_condition(immune_mediated,comorbidities_5y),
    comorb_cardiovascular_chronic = check_condition(cardiovascular, comorbidities_5y),
    comorb_gastrointestinal_chronic = check_condition(gastrointestinal, comorbidities_5y),
    comorb_inflammatory_other_chronic = check_condition(inflammatory_other, comorbidities_5y)
  )

```

Count for these new groupings:

Parasitism

```{r}
cross_table <- table(HSA$had_HSA,HSA$comorb_intestinal_parasitism_chronic)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

```{r}
cross_table <- table(HSA$had_HSA,HSA$comorb_tick_borne_parasitism_chronic)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

```{r}
cross_table <- table(HSA$had_HSA,HSA$comorb_other_parasitism_chronic)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

```{r}
cross_table <- table(HSA$had_HSA,HSA$comorb_infectious_chronic)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

```{r}
cross_table <- table(HSA$had_HSA,HSA$comorb_orthopaedic_chronic)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

```{r}
cross_table <- table(HSA$had_HSA,HSA$comorb_immune_mediated_chronic)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

```{r}
cross_table <- table(HSA$had_HSA,HSA$comorb_cardiovascular_chronic)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

```{r}
cross_table <- table(HSA$had_HSA,HSA$comorb_gastrointestinal_chronic)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

```{r}
cross_table <- table(HSA$had_HSA,HSA$comorb_inflammatory_other_chronic)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

#### Examining conditions that are considered likely chronic inflammatory once diagnosed without needing to be there for \>1y in records (but only within final 5y)

```{r}
HSA <- HSA %>%
  mutate(
    comorb_orthopaedic = check_condition(orthopaedic,comorbidities_no_min_year),
    comorb_immune_mediated = check_condition(immune_mediated,comorbidities_no_min_year),
    comorb_cardiovascular = check_condition(cardiovascular, comorbidities_no_min_year),
    comorb_gastrointestinal = check_condition(gastrointestinal, comorbidities_no_min_year),
    comorb_inflammatory_other = check_condition(inflammatory_other, comorbidities_no_min_year)
  )

```

```{r}
cross_table <- table(HSA$had_HSA,HSA$comorb_orthopaedic)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

```{r}
cross_table <- table(HSA$had_HSA,HSA$comorb_immune_mediated)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

```{r}
cross_table <- table(HSA$had_HSA,HSA$comorb_cardiovascular)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

```{r}
cross_table <- table(HSA$had_HSA,HSA$comorb_gastrointestinal_chronic)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

```{r}
cross_table <- table(HSA$had_HSA,HSA$comorb_inflammatory_other_chronic)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

#### Study year of first chronic inflammatory condition recorded

```{r}

HSA %>%
    summarize(
    mean_freq = mean(study_year_first_chronic_inflam, na.rm = TRUE),
    median_freq = median(study_year_first_chronic_inflam, na.rm = TRUE),
    sd_freq = sd(study_year_first_chronic_inflam, na.rm = TRUE),
    min_freq = min(study_year_first_chronic_inflam, na.rm = TRUE),
    max_freq= max(study_year_first_chronic_inflam, na.rm = TRUE),
    n = n(),
    n_missing = sum(is.na(study_year_first_chronic_inflam))
  )
```

#### Time between first chronic inflammatory condition and endpoint record

```{r}


HSA %>%
    summarize(
    mean_freq = mean(time_between_first_comorb_and_endpoint, na.rm = TRUE),
    median_freq = median(time_between_first_comorb_and_endpoint, na.rm = TRUE),
    sd_freq = sd(time_between_first_comorb_and_endpoint, na.rm = TRUE),
    min_freq = min(time_between_first_comorb_and_endpoint, na.rm = TRUE),
    max_freq= max(time_between_first_comorb_and_endpoint, na.rm = TRUE),
    n = n(),
    n_missing = sum(is.na(time_between_first_comorb_and_endpoint))
  )
```

### Activity and lifestyle:

#### Activity frequency

```{r}
HSA %>%
    summarize(
    mean_freq = mean(avg_frequency_5y_prior_endpoint, na.rm = TRUE),
    median_freq = median(avg_frequency_5y_prior_endpoint, na.rm = TRUE),
    sd_freq = sd(avg_frequency_5y_prior_endpoint, na.rm = TRUE),
    min_freq = min(avg_frequency_5y_prior_endpoint, na.rm = TRUE),
    max_freq= max(avg_frequency_5y_prior_endpoint, na.rm = TRUE),
    n = n(),
    n_missing = sum(is.na(avg_frequency_5y_prior_endpoint))
  )
```

#### Activity intensity

```{r}
HSA %>%
    summarize(
    mean_intensity = mean(avg_intensity_5y_prior_endpoint, na.rm = TRUE),
    median_intensity = median(avg_intensity_5y_prior_endpoint, na.rm = TRUE),
    sd_intensity = sd(avg_intensity_5y_prior_endpoint, na.rm = TRUE),
    min_intensity = min(avg_intensity_5y_prior_endpoint, na.rm = TRUE),
    max_intensity= max(avg_intensity_5y_prior_endpoint, na.rm = TRUE),
    n = n(),
    n_missing = sum(is.na(avg_intensity_5y_prior_endpoint))
  )
```

#### Lifestyle

```{r}
cross_table <- table(HSA$had_HSA,HSA$main_lifestyle_category_mode_5yrs_prior)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

#### Weight (NB. in LB - may wish to convert kg)

```{r}
HSA %>%
    summarize(
    mean_weight = mean(avg_weight_5y_prior_endpoint, na.rm = TRUE),
    median_weight = median(avg_weight_5y_prior_endpoint, na.rm = TRUE),
    sd_weight = sd(avg_weight_5y_prior_endpoint, na.rm = TRUE),
    min_weight = min(avg_weight_5y_prior_endpoint, na.rm = TRUE),
    max_weight= max(avg_weight_5y_prior_endpoint, na.rm = TRUE),
    n = n(),
    n_missing = sum(is.na(avg_weight_5y_prior_endpoint))
  )
```

Converting to kg and removing outliers

```{r}
HSA <- HSA %>%
  mutate(avg_weight_5y_prior_endpoint_kg = avg_weight_5y_prior_endpoint * 0.453)


#how does data look now
HSA %>%
    summarize(
    mean_weight = mean(avg_weight_5y_prior_endpoint_kg, na.rm = TRUE),
    median_weight = median(avg_weight_5y_prior_endpoint_kg, na.rm = TRUE),
    sd_weight = sd(avg_weight_5y_prior_endpoint_kg, na.rm = TRUE),
    min_weight = min(avg_weight_5y_prior_endpoint_kg, na.rm = TRUE),
    max_weight= max(avg_weight_5y_prior_endpoint_kg, na.rm = TRUE),
    n = n(),
    n_missing = sum(is.na(avg_weight_5y_prior_endpoint_kg))
  )
#remove outliers >1.5X above IQR

Q1 <- quantile(HSA$avg_weight_5y_prior_endpoint_kg, 0.25, na.rm = TRUE)
Q3 <- quantile(HSA$avg_weight_5y_prior_endpoint_kg, 0.75, na.rm = TRUE)
IQR <- Q3 - Q1

HSA <- HSA %>%
  mutate(avg_weight_5y_prior_endpoint_kg_outliers_removed= ifelse(avg_weight_5y_prior_endpoint_kg < (Q1 - 1.5 * IQR) | avg_weight_5y_prior_endpoint_kg > (Q3 + 1.5 * IQR), NA, avg_weight_5y_prior_endpoint_kg))
```

```{r}
#outlier removed dataset

#how does data look now
HSA %>%
    summarize(
    mean_weight = mean(avg_weight_5y_prior_endpoint_kg_outliers_removed, na.rm = TRUE),
    median_weight = median(avg_weight_5y_prior_endpoint_kg_outliers_removed, na.rm = TRUE),
    sd_weight = sd(avg_weight_5y_prior_endpoint_kg_outliers_removed, na.rm = TRUE),
    min_weight = min(avg_weight_5y_prior_endpoint_kg_outliers_removed, na.rm = TRUE),
    max_weight= max(avg_weight_5y_prior_endpoint_kg_outliers_removed, na.rm = TRUE),
    n = n(),
    n_missing = sum(is.na(avg_weight_5y_prior_endpoint_kg_outliers_removed))
  )
```

#### Height (NB. in inches- may wish to convert to cm)

```{r}
HSA %>%
    summarize(
    mean_height = mean(avg_height_5y_prior_endpoint, na.rm = TRUE),
    median_height = median(avg_height_5y_prior_endpoint, na.rm = TRUE),
    sd_height = sd(avg_height_5y_prior_endpoint, na.rm = TRUE),
    min_height = min(avg_height_5y_prior_endpoint, na.rm = TRUE),
    max_height= max(avg_height_5y_prior_endpoint, na.rm = TRUE),
    n = n(),
    n_missing = sum(is.na(avg_height_5y_prior_endpoint))
  )
```

Convert to cm and remove outliers

```{r}
HSA <- HSA %>%
  mutate(avg_height_5y_prior_endpoint_cm = avg_height_5y_prior_endpoint * 2.54)


#how does data look now
HSA %>%
    summarize(
    mean_height = mean(avg_height_5y_prior_endpoint_cm, na.rm = TRUE),
    median_height = median(avg_height_5y_prior_endpoint_cm, na.rm = TRUE),
    sd_height = sd(avg_height_5y_prior_endpoint_cm, na.rm = TRUE),
    min_height = min(avg_height_5y_prior_endpoint_cm, na.rm = TRUE),
    max_height= max(avg_height_5y_prior_endpoint_cm, na.rm = TRUE),
    n = n(),
    n_missing = sum(is.na(avg_height_5y_prior_endpoint_cm))
  )
#remove outliers >1.5X above IQR

Q1 <- quantile(HSA$avg_height_5y_prior_endpoint_cm, 0.25, na.rm = TRUE)
Q3 <- quantile(HSA$avg_height_5y_prior_endpoint_cm, 0.75, na.rm = TRUE)
IQR <- Q3 - Q1

HSA <- HSA %>%
  mutate(avg_height_5y_prior_endpoint_cm_outliers_removed= ifelse(avg_height_5y_prior_endpoint_cm < (Q1 - 1.5 * IQR) | avg_height_5y_prior_endpoint_cm > (Q3 + 1.5 * IQR), NA, avg_height_5y_prior_endpoint_cm))

```

#### BCS

```{r}
HSA %>%
    summarize(
    mean_bcs = mean(avg_purina_BCS_5y_prior_endpoint, na.rm = TRUE),
    median_bcs = median(avg_purina_BCS_5y_prior_endpoint, na.rm = TRUE),
    sd_bcs= sd(avg_purina_BCS_5y_prior_endpoint, na.rm = TRUE),
    min_bcs = min(avg_purina_BCS_5y_prior_endpoint, na.rm = TRUE),
    max_bcs= max(avg_purina_BCS_5y_prior_endpoint, na.rm = TRUE),
    n = n(),
    n_missing = sum(is.na(avg_purina_BCS_5y_prior_endpoint))
  )


HSA %>%
    summarize(
    mean_bcs = mean(mode_purina_BCS_5y_prior_endpoint, na.rm = TRUE),
    median_bcs = median(mode_purina_BCS_5y_prior_endpoint, na.rm = TRUE),
    sd_bcs= sd(mode_purina_BCS_5y_prior_endpoint, na.rm = TRUE),
    min_bcs = min(mode_purina_BCS_5y_prior_endpoint, na.rm = TRUE),
    max_bcs= max(mode_purina_BCS_5y_prior_endpoint, na.rm = TRUE),
    n = n(),
    n_missing = sum(is.na(mode_purina_BCS_5y_prior_endpoint))
  )
```

### Reproduction

#### Neuter status

```{r}
cross_table <- table(HSA$had_HSA,HSA$neuter)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

#### Sex neuter status by final record

```{r}
cross_table <- table(HSA$had_HSA,HSA$sex_status)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

#### Sex status

```{r}
cross_table <- table(HSA$had_HSA,HSA$sex)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

#### Neutering reason

```{r}
cross_table <- table(HSA$had_HSA,HSA$neutered_reason)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

#### Age at neutering

NB. need to change the missing ones that are entire dogs to "not neutered"

```{r}
HSA %>%
    summarize(
    mean_age_at_neuter_years = mean(age_at_neuter_years, na.rm = TRUE),
    median_age_at_neuter_years= median(age_at_neuter_years, na.rm = TRUE),
    sd_age_at_neuter_years = sd(age_at_neuter_years, na.rm = TRUE),
    min_age_at_neuter_years = min(age_at_neuter_years, na.rm = TRUE),
    max_age_at_neuter_years= max(age_at_neuter_years, na.rm = TRUE),
    n = n(),
    n_missing = sum(is.na(age_at_neuter_years))
  )
```

#### Ever bred from neutered:

NB. need to combine this with entire dog records for "ever bred from variable" which will include neutered and entire dogs

```{r}
cross_table <- table(HSA$had_HSA,HSA$neutered_ever_bred)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

#### Dogs ever bred from (neutered and entire dogs, M and F dogs)

```{r}
cross_table <- table(HSA$had_HSA,HSA$ever_bred_from)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

### Medications:

#### 5y prior to endpoint:

##### Anti-inflammatories

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_anti_inflam_5y_prior_endpoint)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Anti-microbials

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_anti_microbials_5y_prior_endpoint)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Dietary supplements

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_dietary_supplements_5y_prior_endpoint)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Enteric medications

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_enteric_5y_prior_endpoint)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Fluid metabolites

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_fluid_metabolites_5y_prior_endpoint)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Hormones

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_hormones_and_related_5y_prior_endpoint)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Immunologicals

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_immunologicals_5y_prior_endpoint)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Neurological

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_neurological_5y_prior_endpoint)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Anti-parasite

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_anti_parasites_5y_prior_endpoint)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Anti-septic

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_anti_septics_5y_prior_endpoint)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Cardio-respiratory

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_cardio_respiratory_5y_prior_endpoint)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Diuretics

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_diuretics_5y_prior_endpoint)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Anti-histamines

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_anti_hist_5y_prior_endpoint)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Anti-virals

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_anti_viral_5y_prior_endpoint)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

#### Lifetime:

##### Anti-inflammatories

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_anti_inflam_lifetime)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Anti-microbials

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_anti_microbials_lifetime)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Dietary supplements

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_dietary_supplements_lifetime)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Enteric medications

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_enteric_lifetime)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

\*convert to had/didnt have

```{r}

```

##### Fluid metabolites

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_fluid_metabolites_lifetime)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Hormones

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_hormones_and_related_lifetime)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Immunologicals

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_immunologicals_lifetime)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Neurological

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_neurological_lifetime)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Anti-parasite

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_anti_parasites_lifetime)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Anti-septic

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_anti_septics_lifetime)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Cardio-respiratory

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_cardio_respiratory_lifetime)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Diuretics

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_diuretics_lifetime)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Anti-histamines

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_anti_hist_lifetime)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

##### Anti-virals

```{r}
cross_table <- table(HSA$had_HSA,HSA$quartile_anti_viral_lifetime)
cross_table_percent <- prop.table(cross_table, margin = 1) * 100
print(cross_table)
print(cross_table_percent)
```

```{r}
write.csv(HSA,"C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/HSA_cohort_all_RFs_data_tidied2.csv")
```
