---
title: "Lymphoma dog comorbs and meds cross checking"
format: html
editor: visual
---

## Data and packages

## Comorbidities data

```{r}
lymphoma_dogs <-read.csv("C:/Users/ctaylor18/OneDrive - Royal Veterinary College/post doc/GRLS/GRLS lymphoma RP2 projects/clinical_records_coding_compiled_CT_tidy_and_original_columns.csv")
comorbs_all <-read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/all_cohort_all_comorbs_ever.csv")
comorbs_all <- comorbs_all %>%
  dplyr::select(-X)
#rename so same patientID col name
lymphoma_dogs <- lymphoma_dogs %>%
  rename(subject_id = PatientID)
```

Filter the comorbs all for only lymphoma cases:

```{r}
#keep lymphoma IDs
lymphoma_ids <- unique(lymphoma_dogs$subject_id)
#keep relevant lymphoma records
lymphoma_dogs_abr <- lymphoma_dogs %>%
  dplyr::select(c(subject_id,Co.morbidities_YN,Comorbidity.1,Comorbidity.2,Pre.existing.medications_YN,Pre.existing.medication.1,Pre.existing.medication.2,Pre.existing.medication.3))


lymphoma_comorbs<- comorbs_all%>%
  dplyr::filter(subject_id %in% lymphoma_ids)
#now add in the RP2 read clinical record bits
lymphoma_comorbs_with_rp2 <- lymphoma_comorbs %>%
  left_join(lymphoma_dogs_abr,by="subject_id")
```

Write this to csv to examine

```{r}
write.csv(lymphoma_comorbs_with_rp2,"C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/GRLS_lymphoma_plus_records_comorbs3.csv")
```

Do counts for manual crosschecking now

```{r}
records_crosschecked <- read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/GRLS_lymphoma_plus_records_comorbs3_annotated.csv")

records_crosschecked_abr <- records_crosschecked %>%
  distinct(subject_id, .keep_all = TRUE)
```

```{r}
table(records_crosschecked_abr$mismatch_YN)

table(records_crosschecked_abr$mismatch_OK_YN)

table(records_crosschecked_abr$reason_for_mismatch_OK)
```

\~ 70% agreement between records

### Medications data

```{r}
meds_endpoints_5y<-read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/GRLS_endpoints_5y_prior_medications.csv")

#keep lymphoma only records
lymphoma_meds<- meds_endpoints_5y%>%
  dplyr::filter(subject_id %in% lymphoma_ids)

#keep relevant columns
lymphoma_meds_abr <- lymphoma_meds %>%
  dplyr::select(c(subject_id,year_in_study.x,record_date,medication_name,medication_name_specify,medication_ingredients,medication_ingredients_100,administration_method,duration,active_ingredient_groups,active_ingredient_groups_100))


#now add in the RP2 read clinical record bits
lymphoma_meds_with_rp2 <- lymphoma_meds_abr%>%
  left_join(lymphoma_dogs_abr,by="subject_id")


write.csv(lymphoma_meds_with_rp2,"C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/GRLS_lymphoma_plus_records_medications.csv")
```

Cross-check counting

```{r}
lymphoma_meds_annotated<-read.csv("C:/Users/ctaylor18/GitHub/GRLS_analyses/Output/GRLS_lymphoma_plus_records_medications_annotated.csv")

lymphoma_meds_annotated_unique <- lymphoma_meds_annotated %>%
  distinct(subject_id, .keep_all = TRUE)



```

```{r}
table(lymphoma_meds_annotated_unique$agreement_YN)
```

\~17% full agreement, remaining 50% had no meds in clinical records / no clinical records ... so of those with clinical records \~50% matched fully

```{r}
table(lymphoma_meds_annotated_unique$mismatch_OK_YN)
```

A further \~50% of those with clinical records that didnt fully mismatch had an "OK" reason to mismatch

```{r}
table(lymphoma_meds_annotated_unique$reason_for_mismatch_OK
      )
```

e.g 6 of them had 2/3 of the medications matching, \~10 only had a few early SY records so reasonable to mismatch

So, \~70% of dogs with available meds info and clinical records reading info agreement
