---
title: "MCT dataset variable creation"
format: html
editor: visual
---

## MCT case-control variable creation

```{r}
library(tidyverse)
library(lubridate)
source(here::here("R", "cohort_preparation.R"))
dog_pop <- read.csv(here::here("Output/MCT_cohort_250531.csv"))
```

Make sure population has the endpoint in date format (but also only to the year level to match to questinoaire etc. data which only annual)

```{r}
# Remove the excess record-date column and create endpoint values at year and
# month levels. Annual owner questionnaires do not justify day-level matching.
dog_pop <- prepare_cancer_cohort(dog_pop)
```

Datasets to join up to:

-   clinic and deprivation indices

-   Environment

-   lifestyle

-   reproduction

-   comorbidities

-   physical exam

-   medications

Need to create from these (where applicable):

-   early life (SY0-2)

-   rest of life (SY3 +)

-   whole life (SY0-final SY of records)

-   5y prior to endpoint

Had seperate files for all of these datasets for the HSA dataset creation but will add in here for other user of file ease

## Clinic and deprivation indices

created in this file for all the clinics based on 2019 values so can use these again as MDI not updated yet- GRLS clinic and deprivation indices linkup.qmd

```{r}
deprivation <- read.csv(here::here("Output/GRLS_dogs_MDI_2019.csv"))

# Join clinic and owner deprivation onto the cancer cohort by subject ID.
dog_pop_deprivation <- add_deprivation_features(dog_pop, deprivation)
```

## Environment exposures

NB. most of the enviro variables (or possibly all) do not have any missing data ie. they are all encoded 0 or 1 for the questionaires - so there is no 3rd level of "no_records/not_recorded" as with many other variables in the dataset

indoors exposures =

'use_aerosol','use_air_cleaner','use_hepa_filter,'use_moth_balls','use_incense_or_candles','smoke_exposure'

outside house exposures=

'any_treated_weeds','any_treated_insects','any_treated_fertilizer'

```{r}
# Environmental-condition and sleep-location questionnaires are separate source
# tables because sleep location can contain a different number of rows per dog.
source(here::here("Code", "GRLS_functions.R"))
source(here::here("R", "environment.R"))

environment_exposures <- read.csv(
  here::here("Data/environment_conditions.csv"),
  na.strings = c("", "NA")
)
sleep_environment <- read.csv(here::here("Data/environment_sleep_location.csv"),na.strings=c("","NA"))
sleep_environment <- sleep_environment %>% mutate(across(where(is.character), tolower))

```

### Tidy to binary exposure windows

Indoor exposures are aerosol use, air cleaners, HEPA filters, moth balls,
incense/candles, and household smoke. Outdoor exposures are treatment of weeds,
insects, and lawns with fertiliser.

For every exposure, the shared function creates indicators and counts for:

- early life (study years 0–2);
- rest of life (study years 3–10);
- whole observed life (study years 0–10); and
- the five years preceding the cancer endpoint.

The sleep-location table remains separate because dogs may have multiple sleep
locations within a questionnaire year.

```{r}
environment_features <- build_environment_binary_features(
  environment_exposures,
  dog_pop_deprivation
)

# Retain the established object names used by the smoke and sleep sections below.
indoors_outdoors <- environment_features$longitudinal %>%
  dplyr::select(-dplyr::all_of(setdiff(names(dog_pop_deprivation), "subject_id")))
dog_pop_deprivation_enviro__early_rest_whole4 <-
  environment_features$exposure_windows
dog_pop_deprivation_enviro5y <- environment_features$within_five_years
```

#### Hours of smoke exposure - early, rest, whole, 5y prior

Rows recorded as "see detail rows" do not contain a usable dosage and are
excluded. The shared helper converts the remaining values to numeric hours and
calculates total exposure in early life, the rest of life, the whole observed
life, and before diagnosis. It also retains the binary five-year exposure
features created above.

```{r}
dog_pop_deprivation_enviro_smoking_all <- build_smoke_dosage_features(
  dog_pop_deprivation_enviro__early_rest_whole4,
  dog_pop_deprivation_enviro5y
)
```

### Sleep location

Lots of dogs have multiple `specific_sleep_location` values in the same record
year. Duplicate rows are removed, but distinct locations are retained. The
three recorded locations are garage, house, and outside; the shared helper
creates early-life, rest-of-life, whole-life, and five-year indicators for each.

The main lifetime sleep location is the uniquely most frequent location. A tie
deliberately stops the analysis so that a subject-matter tie rule can be chosen
rather than silently selecting one location.

```{r}
dog_pop_deprivation_enviro_smoking_sleep_all <-
  build_sleep_location_features(
    sleep_environment,
    dog_pop_deprivation_enviro_smoking_all
  )
```

## Lifestyle

- As with the environmental and sleep data, every dog has lifestyle records, so
  this domain does not use a `not_recorded` or `records_not_available` category.
- The established broad groups are competitive, working, breed/showing,
  miscellaneous, companion/pet, and other. Category precedence remains the same
  as in the original notebook when a recorded role contains more than one term.
- A dog's main lifestyle is its most frequently recorded raw role. For the 403
  dogs with equally frequent roles, all tied roles are retained together and the
  broad category is `multiple`; this replaces the previous order-dependent
  selection of one role.

```{r}
source(here::here("R", "lifestyle.R"))

lifestyle <- read.csv(
  here::here("Data", "lifestyle.csv"),
  na.strings = c("", "NA")
)

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all <-
  add_lifestyle_features(
    dog_pop_deprivation_enviro_smoking_sleep_all,
    lifestyle
  )
```

## Household exposures

Household features come from `Data/house_details.csv`; residential location
features come from `Data/location_history.csv`. Text is normalized to lower case,
and missing or blank source values are represented as `no_records_available`.

Variables retained from this domain are:

- primary and secondary heating and cooking fuel;
- metal and plastic pipe indicators;
- water source, house type, area type, and mean recorded house age; and
- modal region, state, and owner ZIP.

Free-text fuel descriptions are mapped to the established electric, natural gas,
oil, propane, wood, and renewable groups. The curated spelling variants and
mixed-source descriptions are retained in `R/household.R`.

Pipe indicators combine the structured primary/secondary-house questions with
the curated free-text descriptions. If either structured material question has
no record, the derived material value remains `no_records_available`.

For water, municipal supply means a public water system managed by a local
authority or utility and treated to applicable US drinking-water standards. The
approved broad groups are municipal, groundwater, spring/surface, bottled, and
rainwater/cistern. These grouped values—not the original ungrouped answers—are
used for every derived water-source feature.

Early-life (study years 0–2), rest-of-life (3–10), and whole-life (0–10)
features use the most frequent recorded value. Five-year features use the most
frequent value in the five study years before the cohort endpoint. When values
tie, the output is explicitly `multiple` rather than selecting whichever row
appears first.

Dog `094-000461` has no
`year_in_study_diagnosis_or_final_record_year` in either frozen cancer cohort.
The historical notebook replaced this missing value with 1. The shared workflow
now stops so that the cohort endpoint definition must be corrected explicitly
before an analysis dataset is accepted.

```{r}
source(here::here("R", "household.R"))

house <- read.csv(here::here("Data", "house_details.csv"))
location <- read.csv(here::here("Data", "location_history.csv"))

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location2 <-
  build_household_features(
    house,
    location,
    dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all
  )
```

## Comorbidities

Load in all the comorbidity files:

```{r}
source(here::here("R/comorbidities.R"))

# Join domain extracts by the unique dog/year/to-date key. The extracts contain
# the same keys in different row orders, so row-wise cbind is not valid.
comorbs_all <- read_condition_domains(here::here("Data"))
conditions_neoplasia <- read.csv(
  here::here("Data", "conditions_neoplasia.csv")
)
```

Comorbs to find:

```{r}
#specific comorb groups 
intestinal_parasitism <- c("giardia", "hookworms", "coccidia", "tapeworms", "whipworms", "roundworms", "eimeria", "isospora","cryptosporidium","eimeria","tapeworms")
tick_borne_parasitism <-c("lyme_disease", "anaplasma", "rocky_mountain_spotted_fever", "ticks", "ehrlichia", "lyme disease", "lyme diseases, oa?", "lyme/anaplasma", "lyme+", "anaplasma and lyme +", "positive anaplasma", "erlichia +", "erlichia","valley_fever","babesia","anaplasma","ehrlichia","lyme_disease")
other_parasitism <-c("fleas", "fungal_infection", "heartworm", "ear_mites","other_ectoparasite","heartworm_infection")
all_parasitism<-c("giardia", "hookworms", "coccidia", "tapeworms", "whipworms", "roundworms", "eimeria", "isospora","lyme_disease", "anaplasma", "rocky_mountain_spotted_fever", "ticks", "ehrlichia", "lyme disease", "lyme diseases, oa?", "lyme/anaplasma", "lyme+", "anaplasma and lyme +", "positive anaplasma", "erlichia +", "erlichia","fleas", "fungal_infection", "heartworm", "ear_mites","heartworm_infection","other_ectoparasite")
infectious <- c("leptospirosis","parvovirus","bacterial cystitis/recurrent uti", "chronic uti's/ bacterial cystitis", "cystitis", "uti", "frequent utis", "previous utis", "chronic urinary incontinence","history of valley fever", "mma- coccidioides")
orthopaedic <-c("osteoarthritis", "l ccl partial tear", "oa of stifle", "hip dysplasia", "right elbow dysplasia grade ii", "ccl rupture", "acl tear", "tendonitis/ arthritis", "arthritis- cervical neck spine", "lh partial ccl tear, lh iliopsoas injury", "chronic lameness- hx of car accident", "lameness","hypothyroidism, chronic lameness, obesity","overweight, arthritis","dvd, larpar", "seizures, oa","tail fracture", "broken leg, histiocytoma","osteochondritis_dissecans","intervertebral_disc_disease","lameness","musculoskeletal_pain","osteomyeltiis","cruciate_ligament_rupture","hip_dysplasia","elbow_dysplasia","crepitus_decreased_range_of_motion","bone_fractures","limber_tail_syndrome","panosteitis")
immune_mediated <- c("uveitis", "autoimmune_thyroiditis", "keratoconjunctivitis_sicca", "megaesophagus", "diabetes_mellitus", "cushings_disease", "hypothyroidism", "granuloma", "immune complex formation", "suspected cushings/ thyroid carcinoma","dvd, larpar","immune_related_musculoskeletal_disorder","rheumatoid_arthritis","diabetes_insipidus","hypothyroidism","pancreatic_insuficiency","addisons_disease","autoimmune_thyroiditis","cushings_disease","hyperadrenalestrogenism","hypercalcemia")
cardiovascular <- c("cardiomyopathy", "hypertension", "heart base tumor", "heart base mass", "mild degenerative valve disease", "degenerative mitral and tricuspid valve disease", "mitral valve disease", "pulmonary thrombosis", "portal vein thrombus","arrhythmia","atrial enlargement","cardiomegaly","cardiomyopathy","congestive_heart_failure","hypertension","murmur","lower_airway_disease","myocardial_dysfunction","pleural_effusion","pulmonic_stenosis","pneumothorax","pulmonic_stenosis","subaortic_stenosis","tachycardia","valvular_disease")
gastrointestinal <-c("chronic_colitis", "pancreatitis", "malabsorptive_disorder", "chronic pancreatitis", "cystic endometrial hyperplasia", "intermittent mixed bowel diarrhea", "intermittent diarrhea - bacterial", "gastric ulcers","mesenteric volvulus","fb","esophagitis","diarrhea","dietary_indiscretion","food_sensitivity","gastrointestinal_foreign_body","gastorintestinal_ulcer","intussusception","malabsorptive_disorder","megaesophagus")

inflammatory_other <- c("tracheobronchitis", "hypercalcaemia", "pigmentary_uveitis", "ear infections", "chronic pyoderma", "chronic pruritis", "chronic rhinitis/ nasal discharge", "allergy, obesity", "pruritis", "severe allergies", "atopy", "allergies, chronic otitis", "chronic hot spots, skin infections", "many pyodera/dermatitis events", "chronic otitis, allergic dermatitis", "dermatitis", "food allergy", "skin allergies", "chronic otitis, overweight", "atopic dermatitis, allergies", "seasonal allergies/pyoderma", "chronic otitis, chronic allergies", "dermatitis lip fold","anal sac infection","otitis_externa","otitis_media","aural_haematoma","aural_foreign_body","pharyngitis","rhinitis","hypercalcaemia","keratoconjunctivitis_sicca","hypercalcemia","influenza","parvovirus")


chronic_inflammatory <-c("congestive_heart_failure",
  "cardiomyopathy",
  "hypertension",
  "heartworm_infection",
  "lower_airway_disease",
  "dental_tarter_calculus",
  "addisons_disease",
  "autoimmune_thyroiditis",
  "cushings_disease",
  "diabetes_insipidus",
  "diabetes_mellitus",
  "hyperadrenalestrogenism",
  "hypercalcemia",
  "hypothyroidism",
  "pancreatic_insufficiency",
  "keratoconjunctivitis_sicca",
  "pigmentary_uveitis",
  "uveitis",
  "chronic_colitis",
  "malabsorptive_disorder",
  "megaesophagus",
  "pancreatitis",
  "anaplasma",
  "babesia",
  "canine_herpesvirus",
  "coccidia",
  "cryptosporidium",
  "ear_mites",
  "ehrlichia",
  "eimeria",
  "fleas",
  "fungal_infection",
  "giardia",
  "granuloma",
  "hookworms",
  "influenza",
  "isospora",
  "leptospirosis",
  "lyme_disease",
  "other_ectoparasite",
  "parvovirus",
  "rocky_mountain_spotted_fever",
  "roundworms",
  "tapeworms",
  "ticks",
  "tracheobronchitis",
  "unspecified_gastrointestinal_parasite",
  "valley_fever",
  "whipworms",
  "immune_related_musculoskeletal_disorder",
  "osteoarthritis",
  "rheumatoid_arthritis",
  "laryngeal_paralysis",
  "meningitis",
  "meningoencephalitis",
  "myasthenia_gravis",
  "erythema_multiforme",
  "flea_allergy_dermatitis",
  "food_allergy_dermatitis",
  "r_o_atopy")
```

#### Lifetime and final-five-year condition groups

The shared helper calculates each condition group by its column names. Lifetime
flags use all available annual records. The final-five-year flags use each dog's
five most recent study years, and the six groups below also retain diagnoses
that are treated as permanent once recorded.

```{r}
comorbidity_groups <- list(
  intestinal_parasitism = intestinal_parasitism,
  tick_borne_parasitism = tick_borne_parasitism,
  other_parasitism = other_parasitism,
  all_parasitism = all_parasitism,
  infectious = infectious,
  orthopaedic = orthopaedic,
  immune_mediated = immune_mediated,
  cardiovascular = cardiovascular,
  gastrointestinal = gastrointestinal,
  inflammatory_other = inflammatory_other,
  chronic_inflammatory = chronic_inflammatory
)

comorbs_single_record <- c(
  "atrial_enlargement",
  "cardiomegaly",
  "cardiomyopathy",
  "congestive_heart_failure",
  "hypertension",
  "myocardial_dysfunction",
  "valvular_disease",
  "addisons_disease",
  "autoimmune_thyroiditis",
  "cushings_disease",
  "diabetes_insipidus",
  "diabetes_mellitus",
  "hyperadrenalestrogenism",
  "hypercalcemia",
  "hypothyroidism",
  "pancreatic_insufficiency",
  "cataracts",
  "glaucoma",
  "keratoconjunctivitis_sicca",
  "progressive_retinal_atrophy_or_degeneration",
  "pigmentary_uveitis",
  "lymphangiectasia",
  "malabsorptive_disorder",
  "megaesophagus",
  "motility_disorder",
  "elbow_dysplasia",
  "hip_dysplasia",
  "osteoarthritis",
  "rheumatoid_arthritis",
  "spondylosis"
)

GRLS_5yprior_lifetime <- build_condition_group_features(
  comorbs_all,
  comorbidity_groups,
  comorbs_single_record
)
```

#### Other neoplasia variables

The all-cancer count includes every named neoplasia field, including leukemia
and melanoma, except the cancer defining this analysis
(`mast_cell_tumor`). Each reported cancer is counted once per annual record.
The existing malignant and benign subgroup definition is retained separately.

```{r}
malig_cancers <- c(
  "hemangiosarcoma",
  "histiocytic_sarcoma",
  "melanoma",
  "multiple_myeloma",
  "osteosarcoma",
  "pancreatic_tumor",
  "plasma_cell_tumor",
  "plasmacytoma",
  "soft_tissue_sarcoma",
  "squamous_cell_carcinoma"
)

benign_cancers <- c(
  "adrenal_tumor",
  "basal_cell_tumor",
  "bile_duct_tumor",
  "bladder_tumor",
  "brain_spinal_cord_tumor",
  "breast_or_mammary_tumor",
  "eye_tumor",
  "hair_matrix_tumor",
  "heart_tumor",
  "kidney_tumor",
  "liver_tumor",
  "lung_tumor",
  "nasal_tumor",
  "perianal_adenoma",
  "pituitary_tumor",
  "prostate_tumor",
  "splenic_tumor",
  "stomach_intestinal_tumor",
  "testicular_tumor",
  "thymoma",
  "thyroid_tumor"
)

cancer_summary_5y_lifetime <- build_neoplasia_features(
  conditions_neoplasia,
  target_cancer = "mast_cell_tumor",
  malignant_cancers = malig_cancers,
  benign_cancers = benign_cancers
)
```

Combine all comorbs data back into the main df

```{r}
#combine cancer and other comorbs
GRLS_comorbs_cancer_5yprior_lifetime <- GRLS_5yprior_lifetime %>%
  left_join(cancer_summary_5y_lifetime,by="subject_id")

#combine back to main df
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location2 %>% left_join(GRLS_comorbs_cancer_5yprior_lifetime,by="subject_id")
```

## Medications

Using the df created in an earlier script of analyses which tidies up free text and the many drug categories etc. in this qmd GRLS medications.qmd

```{r}
medications <- read.csv(here::here("Output/GRLS_medications_initial_tidy.csv"))
source(here::here("R", "medications.R"))
```

### Lifetime medication calculations

```{r}
medication_features_lifetime <- summarise_medication_quartiles(
  medications,
  suffix = "lifetime",
  missing_value = "medication_not_prescribed"
)

meds_per_dog_lifetime <- medication_features_lifetime$counts
medication_summary_lifetime <- medication_features_lifetime$summary
meds_per_dog_quart_lifetime <- medication_features_lifetime$categorised
meds_wide_quart_lifetime <- medication_features_lifetime$wide

write.csv(medication_summary_lifetime,here::here("Output/MCT_study_medication_quartiles_summaries_lifetime.csv"))
meds_wide_quart_lifetime
```

### 5y prior to endpoint medication calculations

NB. there is 2889 dogs in this 5y - dogs with meds records in 5y prior to death, ensure when joining back into main 3044 dogs that these \~150 dogs have "no_records_available" coded on them

```{r}
medications_5y <- select_last_study_years(medications, 5)

medication_features_5y <- summarise_medication_quartiles(
  medications_5y,
  suffix = "5y",
  missing_value = "medication_not_used"
)

meds_per_dog_5y <- medication_features_5y$counts
medication_summary_5y <- medication_features_5y$summary
meds_per_dog_quart_5y <- medication_features_5y$categorised
meds_wide_quart_5y <- medication_features_5y$wide

write.csv(medication_summary_5y,here::here("Output/MCT_study_medication_quartiles_summaries_5y.csv"))
meds_wide_quart_5y
```

### Examine for specific potential RF medications:

-   NSAIDs exposure

    -   ever Y N and 5y prior Y N
    -   counts

-   steroids exposure

    -   ever Y N and 5y prior Y N
    -   counts

-   Fenbendazole exposure

    -   ever Y N and 5y prior Y N
    -   counts

```{r}
#free text drug names to look for
NSAID_names<-c("carprofen","meloxicam","firocoxib","deracoxib","flurbiprofen","piroxicam","ketotifen","ketoprofen","rimadyl","metacam","meloxidyl","previcox","vetprofen","ketorolac opthalmic ointment","toradol","acular","voltaren
","diclofenac","flurbiprofen eye drops
","carprovet","carprieve","truprofen","ofloxacin ophthalmic ointment + nsaid
")
steroid_names<-c("dexamethasone, neomycin sulfate, thiabendazole
","prednisolone acetate
","betamethasone, clotrimazole, gentamicin sulfate
","betamethasone, gentamicin sulfate
","dexamethasone","dexamethasone, miconazole nitrate
","dexamethasone, miconazole nitrate
","methylprednisolone","prednisone","betamethasone, florfenicol, terbinafine
","clotrimazole, gentamicin sulfate, mometasone furoate
","dexamethasone, methylprednisolone
","mometasone furoate, orbifloxacin, posaconazole
","dexamethasone, neomycin sulfate, polymyxin b
","bacitracin zinc, dexamethasone, neomycin sulfate, polymyxin b
","miconazole nitrate, polymyxin b, prednisolone acetate
","hydrocortisone, gentamicin sulfate, miconazole nitrate
","gentamicin sulfate, hydrocortisone, miconazole nitrate
","burow's solution, hydrocortisone
","acetic acid, boric acid, hydrocortisone, ketoconazole
","hydrocortisone","bacitracin zinc, hydrocortisone, neomycin sulfate, polymyxin b
","dexamethasone, enrofloxacin, miconazole nitrate
","clotrimazole, dexamethasone, enrofloxacin
","hydrocortisone, ketoconazole
","dexamethasone, ketoconazole, phytosphingosine
","betamethasone","acetic acid, boric acid, ceramide complex, hydrocortisone, ketoconazole
","hydrocortisone, troleandomycin
","dexamethasone, enrofloxacin, ketoconazole, tris-edta
","bacitracin zinc, neomycin sulfate, polymyxin b, unspecified steroid
","clotrimazole, dexamethasone
","dexamethasone, ketoconazole, tris-edta
","dexamethasone, ketoconazole
","betamethasone, clotrimazole, gentamicin sulfate, mometasone furoate
","dexamethasone, diphenhydramine
","beclomethasone, clotrimazole, neomycin sulfate
","dexamethasone, enrofloxacin, ketoconazole
","dexamethasone, dimethyl sulfoxide, enrofloxacin, lidocaine
","dexamethasone, enrofloxacin
","betamethasone, gentamicin sulfate, miconazole nitrate
","chloramphenicol, dexamethasone
","hydrocortisone, pramoxine
","dexamethasone, penicillin
","betamethasone, chloramphenicol, ketoconazole
","dexamethasone, tobramycin
","enrofloxacin, dexamethasone, tris-edta
","dexamethasone, miconazole nitrate, neomycin sulfate, nystatin, thiostrepton, triamcinolone acetonide
","betamethasone, burow's solution, gentamicin sulfate, hydrocortisone, miconazole nitrate
","ketoconazole, hydrocortisone
","bacitracin, hydrocortisone
","dexamethasone, miconazole nitrate, tris-edta
","chlorhexidine gluconate, dexamethasone
","dexamethasone, enrofloxacin, tris-edta
","dexamethasone, enrofloxacin, silver sulfadiazine
","dexamethasone, miconazole nitrate, neomycin sulfate, polymyxin b, prednisolone acetate, thiabendazole
","betamethasone, enrofloxacin
","dexamethasone, ear cleaner (epi-otic advanced), enrofloxacin
","betamethasone, tobramycin
","dexamethasone ophthalmic ointment
","miconazole + dexamethasone ear solution
","amikacin + dexamethasone ear solution
","depo-medrol
","prednisolone acetate ophthalmic suspension
","neo-poly-dex ophthalmic ointment
","neo-poly-bac ophthalmic ointment + dexamethasone
","betamethasone + gentamicin ear solution
","conofite + dexamethasone ear solution
","neo-poly-bac ophthalmic ointment + hydrocortisone
","baytril + conofite + dexamethasone ear solution
","betamethasone + gentamicin spray
","baytil + clotrimazole + dexamethasone ear solution
","baytril + dexamethasone + miconazole ear solution
","baytril + conzol + dexamethasone ear solution
","trizulta + keto + baytril + dexamethasone flush
","clotrimazole + dexamethasone ear solution
","burow's solution + hydrocortisone
","betamethasone + clotrimazole + gentamicin ear solution
","trizulta + keto + dexamethasone flush
","cortalone
","baytril + dexamethasone + ketoconazole ear solution
","betamethasone + gentamicin sulfate + miconazole nitrate ear solution
","dexamethasone otic
","baytril + trizulta + dexamethasone ear solution
","neomycin sulfate + nystatin + triamcinolone acetonide
","chloramphenicol + ketoconazole + triamcinolone acetonide ear solution
","triamcinolone","trizedta + dexamethasone + miconazole flush
","chlorhexiderm + dexamethasone spray
","trizedta + baytril + dexamethasone flush
","neomycin sulfate + tetracaine + triamcinolone acetonide
","clotrimazole + gentamicin + mometasone ear solution
","baytril + dexamethasone + silvadene ear solution
","itraconazole + ketoconazole + triamcinolone acetonide ear solution
","baytril + triamcinolone acetonide
","troleandomycin + hydrocortisone
","neo-poly-bac ophthalmic ointment + steroid
","t8 keto flush + dexamethasone
","dexamethasone + miconazole
","dermacool + hydrocortisone
","trizchlor 4 + hydrocortisone
","dexamethasone + diphenhydramine
","tear stimulation forte + betamethasone + ofloxacin
","baytril + dexamethasone + dimethyl sulfoxide + lidocaine
","dexamethasone + baytril
","ketoconazole + mupirocin + triamcinolone acetonide
","penicillin + dexamethasone
","tobramycin + betamethasone ophthalmic ointment
","lotramin + dexamethasone
","bacitracin + hydrocortisone
")
fenbendazole_names<-c("fenbendazole","fenbendazole, maropitant citrate, metronidazole, omeprazole, sucralfate
","cerenia + fenbendazole + metronidazole + omeprazole + sucralfate
","panacur","safe-guard")
```

Lifetime and final five years

```{r}
specific_medication_features <- summarise_specific_medication_flags(
  medications,
  medications_5y,
  list(
    NSAID = NSAID_names,
    steroid = steroid_names,
    fenbendazole = fenbendazole_names
  )
)

meds_5y_lifetime <- meds_wide_quart_5y |>
  dplyr::full_join(
    meds_wide_quart_lifetime,
    by = "subject_id",
    relationship = "one-to-one"
  )

medication_features <- meds_5y_lifetime |>
  dplyr::full_join(
    specific_medication_features,
    by = "subject_id",
    relationship = "one-to-one"
  )

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds <-
  join_medication_features(
    dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs,
    medication_features
  )
```

#### Examine also an NSAID number of times variable

```{r}
NSAID_meds <- c(
  "carprofen",
  "meloxicam",
  "firocoxib",
  "deracoxib",
  "flurbiprofen",
  "ketoprofen"
)

NSAID_5y_features <- summarise_medication_count_feature(
  medications_5y,
  NSAID_meds,
  "NSAID",
  function(count) {
    dplyr::case_when(
      count == 0L ~ "no_NSAID_admin",
      count == 1L ~ "Q1(1)",
      count == 2L ~ "Q2(1-2)",
      count == 3L ~ "Q3(2-3)",
      count > 3L ~ "Q4(3-8)"
    )
  },
  amount_column = "NSAID_5y_quantiles"
)

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID <-
  join_medication_features(
    dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds,
    NSAID_5y_features
  )
```

#### Examine also a steroid number of times variable

```{r}
steroids <- c(
  "prednisone",
  "predisolone acetate",
  "trimeprazine tartrate, prednisone",
  "triamcinolone acetonide",
  "methylpredisolone",
  "hydrocortisone, gentamicin sulfate, miconazole nitrate",
  "burow's solution, hydrocortisone",
  "dexamethasone, miconazole nitrate",
  "dimethyl sulfoxide, fluocinolone acetonide",
  "hydrocortisone",
  "dexamethasone, enrofloxacin, ketoconazole, tris-edta",
  "beclomethasone, clotrimazole, neomycin sulfate",
  "dexamethasone, ketoconazole, tris-edta",
  "acetic acid, boric acid, ceramide complex, hydrocortisone, ketoconazole",
  "betamethasone",
  "mometasone furoate",
  "dexamethasone",
  "betamethasone",
  "hydrocortisone",
  "prednisolone",
  "methylprednisolone",
  "triamcinolone acetonide",
  "difluprednate",
  "fluticasone",
  "dexamethasone, miconazole nitrate",
  "amikacin sulfate, dexamethasone",
  "dexamethasone, methylprednisolone",
  "dexamethasone, triamcinolone acetonide",
  "hydrocortisone, ketoconazole",
  "dexamethasone, ketoconazole, phytosphingosine",
  "dexamethasone, diphenhydramine",
  "betamethasone, clotrimazole, gentamicin sulfate, mometasone furoate",
  "dimethyl sulfoxide, fluocinolone acetonide, gentamicin sulfate, miconazole nitrate",
  "dimethyl sulfoxide, flunixin, fluocinolone acetonide",
  "hydrocortisone, troleandomycin",
  "Dexamethasone",
  "Betamethasone",
  "dexamethasone",
  "betamethasone",
  "hydrocortisone",
  "menthol, hydrocortisone",
  "dexamethasone, ketoconazole",
  "dexamethasone, enrofloxacin",
  "dexamethasone, dimethyl sulfoxide, enrofloxacin, lidocaine",
  "dexamethasone, enrofloxacin, ketoconazole",
  "dexamethasone, penicillin",
  "dexamethasone, tobramycin",
  "betamethasone, ofloxacin",
  "chloramphenicol, dexamethasone",
  "betamethasone, chloramphenicol, ketoconazole",
  "enrofloxacin, dexamethasone, tris-edta",
  "dexamethasone, miconazole nitrate, neomycin sulfate, nystatin, thiostrepton, triamcinolone acetonide",
  "betamethasone, burow's solution, gentamicin sulfate, hydrocortisone, miconazole nitrate",
  "ketoconazole, hydrocortisone",
  "bacitracin, hydrocortisone",
  "dexamethasone, miconazole nitrate, tris-edta",
  "chlorhexidine gluconate, dexamethasone",
  "dexamethasone, enrofloxacin, tris-edta",
  "dexamethasone, enrofloxacin, silver sulfadiazine",
  "dexamethasone, miconazole nitrate, neomycin sulfate, polymyxin b, prednisolone acetate, thiabendazole",
  "betamethasone, enrofloxacin",
  "dexamethasone, ear cleaner (epi-otic advanced), enrofloxacin",
  "betamethasone, tobramycin"
)

```

Use steroids list to do counting now

```{r}
steroid_5y_features <- summarise_medication_count_feature(
  medications_5y,
  steroids,
  "steroid",
  function(count) {
    dplyr::case_when(
      count == 0L ~ "no_steroid_admin",
      count == 1L ~ "Q1(1)",
      count == 2L ~ "Q2(1-2)",
      count == 3L ~ "Q3(2-3)",
      count > 3L ~ "Q4(3-17)"
    )
  },
  amount_column = "steroid_5y_quantiles"
)

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid <-
  join_medication_features(
    dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID,
    steroid_5y_features
  )
```

For MCT, also retain five-year administration counts for systemic
prednisolone/prednisone, lokivetmab, and oclacitinib. There was insufficient
variation to form meaningful quartiles, so positive values remain counts.

```{r}
systemic_preds <- c(
  "prednisone",
  "predisolone acetate",
  "trimeprazine tartrate, prednisone",
  "prednisolone"
)
lokivetmab <- c("lokivetmab", "cytopoint")
oclacitinib <- c("oclacitinib", "apoquel")

count_or_no_administration <- function(label) {
  function(count) {
    ifelse(count == 0L, paste0("no_", label, "_admin"), as.character(count))
  }
}

systemic_pred_features <- summarise_medication_count_feature(
  medications_5y,
  systemic_preds,
  "systemic_preds",
  count_or_no_administration("systemic_preds")
)
lokivetmab_features <- summarise_medication_count_feature(
  medications_5y,
  lokivetmab,
  "lokivetmab",
  count_or_no_administration("lokivetmab")
)
oclacitinib_features <- summarise_medication_count_feature(
  medications_5y,
  oclacitinib,
  "oclacitinib",
  count_or_no_administration("oclacitinib")
)

mct_specific_count_features <- systemic_pred_features |>
  dplyr::full_join(
    lokivetmab_features,
    by = "subject_id",
    relationship = "one-to-one"
  ) |>
  dplyr::full_join(
    oclacitinib_features,
    by = "subject_id",
    relationship = "one-to-one"
  )

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib <-
  join_medication_features(
    dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid,
    mct_specific_count_features
  )
```

Interim write to csv

```{r}
write.csv(dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib,here::here("Output/GRLS_MCT_variables_dataset_interim.csv"))
```

## Reproduction variables

Reproductive features use the annual female and male reproductive-history
tables plus neuter status and dates from `Data/dog_profile.csv`. The heat-detail
table is not used here because `heat_date` has substantially less coverage than
the annual `heat_last_year` indicator.

The first-heat age is therefore the dog's age at the first questionnaire that
reported a heat in the preceding year; it is an approximate reporting age, not
an exact heat date. `heat_reporting_years` counts distinct study years with a
positive `heat_last_year` response and must not be interpreted as the literal
number of heats.

Pregnancy and mating are defined as ever recorded, regardless of neuter status,
because some dogs are never neutered. Pregnancy is not applicable to male dogs,
and mating is not applicable to female dogs. Male mating uses either a positive
annual intact-breeding response or the neutered-ever-bred response.

Neuter age, time between neutering and the cohort endpoint, and age at endpoint
are derived from recorded dates. The workflow stops rather than replacing
negative intervals with zero. In the current data, dogs `094-000441` and
`094-023124` have neuter dates before birth. Ten dogs have neuter dates after
the MCT endpoint; four of these also affect lymphoma: `094-006352`,
`094-000677`, `094-026390`, and `094-027529`. These source/cohort dates must
be resolved before accepting new analysis datasets.

```{r}
source(here::here("R", "reproduction.R"))

repro_hx_female <- read.csv(
  here::here("Data", "reproductive_history_female.csv")
)
repro_hx_male <- read.csv(
  here::here("Data", "reproductive_history_male.csv")
)
profile <- read.csv(here::here("Data", "dog_profile.csv"))

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3 <-
  build_reproduction_features(
    dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib,
    profile,
    repro_hx_female,
    repro_hx_male
  )
```

## Weight and height data 5y prior to endpoint

Physical examination records from the endpoint study year and the five preceding
study years are used. The primary summaries require at least five examination
rows, preserving the established threshold; the `any_records` medians remain
available when fewer records exist. Weight is converted from pounds to kg using
0.453 and height from inches to cm using 2.54. Converted summaries are left
missing for dogs younger than 18 months at the endpoint (six dogs in the legacy
notes). Where BCS values tie for the mode, the maximum tied value is used.

```{r}
source(here::here("R", "weight.R"))

dog_weight <- read.csv(here::here("Data/exam_physical.csv"))
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_unique <-
  build_weight_features(
    dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3,
    dog_weight
  )
```

```{r}
write.csv(dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_unique,here::here("Output/GRLS_MCT_variables_dataset_interim.csv"))
```

## Activity

This feature set follows the original classification work in
`Code/GRLS activity classifcation method_modified.qmd`. The questionnaire
changed after study year 2, so the two periods retain their established scoring
systems:

- study years 0--2 combine reported walk and aerobic duration, frequency, and
  intensity;
- study year 3 onward scores activity frequency and pace, because duration is
  not collected in the later format;
- the overview questionnaire supplies the whole-period activity assessment.

Every activity record after the dog's endpoint is ignored before any summary is
calculated. Full-date source tables are compared with `end_date`; the
month-resolution overview table is compared with the endpoint month. The
five-year summaries use records from the endpoint study year and the five
preceding study years.

The SY3+ and overview fields are ordinary means of their scored records. This
corrects the legacy calculation that divided those means by record count a
second time. `avg_frequency_5y_prior_endpoint` is calculated from the combined
early and later records in the five-year window; it is not replaced by the
all-SY3 mean.

For ordinal modes (the early recommendation categories and, in the lymphoma
schema, overview activity score), the maximum value is used where modal values
tie. This is the documented ordinal tie rule; nominal modes elsewhere in the
pipeline continue to use `multiple`.

```{r}
source(here::here("R", "activity.R"))

activity_sy1_2 <- read.csv(
  here::here("Data/activity_details_through_SY2.csv")
)
activity_sy3_beyond <- read.csv(
  here::here("Data/activity_details_SY3_beyond.csv")
)
activity_overview <- read.csv(here::here("Data/activity_overview.csv"))

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_unique <-
  dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_unique |>
  dplyr::select(-c(year_in_study.x, year_in_study.y, record_date.y))

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity <-
  build_activity_features(
    dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_unique,
    activity_sy1_2,
    activity_sy3_beyond,
    activity_overview,
    include_overview_distribution = FALSE
  )
```

## Tidying up columns for final df

Dates and endpoint fields retain genuine `NA` values (for example, a dog that
was never withdrawn). Missing values in analysis feature columns are written as
`records_not_available`. The feature selection is name-based so it cannot
silently shift when an upstream column is added or removed.

CSV row names are disabled; spreadsheet-style `X` index columns are not part of
the analysis dataset.

```{r}
source(here::here("R", "final_dataset.R"))

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity_tidied <-
  finalise_cancer_dataset(
    dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity,
    case_column = "had_MCT"
  )

write.csv(
  dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity_tidied,
  here::here("Output/GRLS_MCT_variables_dataset.csv"),
  row.names = FALSE
)
```
