---
title: "MCT dataset variable creation"
format: html
editor: visual
---

## MCT case-control variable creation

```{r}
library(tidyverse)
library(lubridate)
source(here::here("R", "cohort_preparation.R"))
dog_pop <- read.csv(here::here("Output/MCT_cohort_250531.csv"))
```

Make sure population has the endpoint in date format (but also only to the year level to match to questinoaire etc. data which only annual)

```{r}
# Remove the excess record-date column and create endpoint values at year and
# month levels. Annual owner questionnaires do not justify day-level matching.
dog_pop <- prepare_cancer_cohort(dog_pop)
```

Datasets to join up to:

-   clinic and deprivation indices

-   Environment

-   lifestyle

-   reproduction

-   comorbidities

-   physical exam

-   medications

Need to create from these (where applicable):

-   early life (SY0-2)

-   rest of life (SY3 +)

-   whole life (SY0-final SY of records)

-   5y prior to endpoint

Had seperate files for all of these datasets for the HSA dataset creation but will add in here for other user of file ease

## Clinic and deprivation indices

created in this file for all the clinics based on 2019 values so can use these again as MDI not updated yet- GRLS clinic and deprivation indices linkup.qmd

```{r}
deprivation <- read.csv(here::here("Output/GRLS_dogs_MDI_2019.csv"))

# Join clinic and owner deprivation onto the cancer cohort by subject ID.
dog_pop_deprivation <- add_deprivation_features(dog_pop, deprivation)
```

## Environment exposures

NB. most of the enviro variables (or possibly all) do not have any missing data ie. they are all encoded 0 or 1 for the questionaires - so there is no 3rd level of "no_records/not_recorded" as with many other variables in the dataset

indoors exposures =

'use_aerosol','use_air_cleaner','use_hepa_filter,'use_moth_balls','use_incense_or_candles','smoke_exposure'

outside house exposures=

'any_treated_weeds','any_treated_insects','any_treated_fertilizer'

```{r}
# Environmental-condition and sleep-location questionnaires are separate source
# tables because sleep location can contain a different number of rows per dog.
source(here::here("Code", "GRLS_functions.R"))
source(here::here("R", "environment.R"))

environment_exposures <- read.csv(
  here::here("Data/environment_conditions.csv"),
  na.strings = c("", "NA")
)
sleep_environment <- read.csv(here::here("Data/environment_sleep_location.csv"),na.strings=c("","NA"))
sleep_environment <- sleep_environment %>% mutate(across(where(is.character), tolower))

```

### Tidy to binary exposure windows

Indoor exposures are aerosol use, air cleaners, HEPA filters, moth balls,
incense/candles, and household smoke. Outdoor exposures are treatment of weeds,
insects, and lawns with fertiliser.

For every exposure, the shared function creates indicators and counts for:

- early life (study years 0–2);
- rest of life (study years 3–10);
- whole observed life (study years 0–10); and
- the five years preceding the cancer endpoint.

The sleep-location table remains separate because dogs may have multiple sleep
locations within a questionnaire year.

```{r}
environment_features <- build_environment_binary_features(
  environment_exposures,
  dog_pop_deprivation
)

# Retain the established object names used by the smoke and sleep sections below.
indoors_outdoors <- environment_features$longitudinal %>%
  dplyr::select(-dplyr::all_of(setdiff(names(dog_pop_deprivation), "subject_id")))
dog_pop_deprivation_enviro__early_rest_whole4 <-
  environment_features$exposure_windows
dog_pop_deprivation_enviro5y <- environment_features$within_five_years
```

#### Hours of smoke exposure - early, rest, whole, 5y prior

Rows recorded as "see detail rows" do not contain a usable dosage and are
excluded. The shared helper converts the remaining values to numeric hours and
calculates total exposure in early life, the rest of life, the whole observed
life, and before diagnosis. It also retains the binary five-year exposure
features created above.

```{r}
dog_pop_deprivation_enviro_smoking_all <- build_smoke_dosage_features(
  dog_pop_deprivation_enviro__early_rest_whole4,
  dog_pop_deprivation_enviro5y
)
```

### Sleep location

Lots of dogs have multiple `specific_sleep_location` values in the same record
year. Duplicate rows are removed, but distinct locations are retained. The
three recorded locations are garage, house, and outside; the shared helper
creates early-life, rest-of-life, whole-life, and five-year indicators for each.

The main lifetime sleep location is the uniquely most frequent location. A tie
deliberately stops the analysis so that a subject-matter tie rule can be chosen
rather than silently selecting one location.

```{r}
dog_pop_deprivation_enviro_smoking_sleep_all <-
  build_sleep_location_features(
    sleep_environment,
    dog_pop_deprivation_enviro_smoking_all
  )
```

## Lifestyle

- As with the environmental and sleep data, every dog has lifestyle records, so
  this domain does not use a `not_recorded` or `records_not_available` category.
- The established broad groups are competitive, working, breed/showing,
  miscellaneous, companion/pet, and other. Category precedence remains the same
  as in the original notebook when a recorded role contains more than one term.
- A dog's main lifestyle is its most frequently recorded raw role. For the 403
  dogs with equally frequent roles, all tied roles are retained together and the
  broad category is `multiple`; this replaces the previous order-dependent
  selection of one role.

```{r}
source(here::here("R", "lifestyle.R"))

lifestyle <- read.csv(
  here::here("Data", "lifestyle.csv"),
  na.strings = c("", "NA")
)

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all <-
  add_lifestyle_features(
    dog_pop_deprivation_enviro_smoking_sleep_all,
    lifestyle
  )
```

## Household exposures

Household features come from `Data/house_details.csv`; residential location
features come from `Data/location_history.csv`. Text is normalized to lower case,
and missing or blank source values are represented as `no_records_available`.

Variables retained from this domain are:

- primary and secondary heating and cooking fuel;
- metal and plastic pipe indicators;
- water source, house type, area type, and mean recorded house age; and
- modal region, state, and owner ZIP.

Free-text fuel descriptions are mapped to the established electric, natural gas,
oil, propane, wood, and renewable groups. The curated spelling variants and
mixed-source descriptions are retained in `R/household.R`.

Pipe indicators combine the structured primary/secondary-house questions with
the curated free-text descriptions. If either structured material question has
no record, the derived material value remains `no_records_available`.

For water, municipal supply means a public water system managed by a local
authority or utility and treated to applicable US drinking-water standards. The
approved broad groups are municipal, groundwater, spring/surface, bottled, and
rainwater/cistern. These grouped values—not the original ungrouped answers—are
used for every derived water-source feature.

Early-life (study years 0–2), rest-of-life (3–10), and whole-life (0–10)
features use the most frequent recorded value. Five-year features use the most
frequent value in the five study years before the cohort endpoint. When values
tie, the output is explicitly `multiple` rather than selecting whichever row
appears first.

Dog `094-000461` has no
`year_in_study_diagnosis_or_final_record_year` in either frozen cancer cohort.
The historical notebook replaced this missing value with 1. The shared workflow
now stops so that the cohort endpoint definition must be corrected explicitly
before an analysis dataset is accepted.

```{r}
source(here::here("R", "household.R"))

house <- read.csv(here::here("Data", "house_details.csv"))
location <- read.csv(here::here("Data", "location_history.csv"))

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location2 <-
  build_household_features(
    house,
    location,
    dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all
  )
```

## Comorbidities

Load in all the comorbidity files:

```{r}
source(here::here("R/comorbidities.R"))

# Join domain extracts by the unique dog/year/to-date key. The extracts contain
# the same keys in different row orders, so row-wise cbind is not valid.
comorbs_all <- read_condition_domains(here::here("Data"))
comorbs_unique <- comorbs_all
#for 5y variables
comorbs_unique2 <- comorbs_all
#for other calcs
comorbs_unique3 <- comorbs_all
```

Comorbs to find:

```{r}
#potential chronic inflam conditions
comorbs_to_find <-c("congestive_heart_failure",
  "cardiomyopathy",
  "hypertension",
  "heartworm_infection",
  "lower_airway_disease",
  "dental_tarter_calculus",
  "addisons_disease",
  "autoimmune_thyroiditis",
  "cushings_disease",
  "diabetes_insipidus",
  "diabetes_mellitus",
  "hyperadrenalestrogenism",
  "hypercalcemia",
  "hypothyroidism",
  "pancreatic_insufficiency",
  "keratoconjunctivitis_sicca",
  "pigmentary_uveitis",
  "uveitis",
  "chronic_colitis",
  "malabsorptive_disorder",
  "megaesophagus",
  "pancreatitis",
  "anaplasma",
  "babesia",
  "canine_herpesvirus",
  "coccidia",
  "cryptosporidium",
  "ear_mites",
  "ehrlichia",
  "eimeria",
  "fleas",
  "fungal_infection",
  "giardia",
  "granuloma",
  "hookworms",
  "influenza",
  "isospora",
  "leptospirosis",
  "lyme_disease",
  "other_ectoparasite",
  "parvovirus",
  "rocky_mountain_spotted_fever",
  "roundworms",
  "tapeworms",
  "ticks",
  "tracheobronchitis",
  "unspecified_gastrointestinal_parasite",
  "valley_fever",
  "whipworms",
  "immune_related_musculoskeletal_disorder",
  "osteoarthritis",
  "rheumatoid_arthritis",
  "laryngeal_paralysis",
  "meningitis",
  "meningoencephalitis",
  "myasthenia_gravis",
  "erythema_multiforme",
  "flea_allergy_dermatitis",
  "food_allergy_dermatitis",
  "r_o_atopy")
#specific comorb groups 
intestinal_parasitism <- c("giardia", "hookworms", "coccidia", "tapeworms", "whipworms", "roundworms", "eimeria", "isospora","cryptosporidium","eimeria","tapeworms")
tick_borne_parasitism <-c("lyme_disease", "anaplasma", "rocky_mountain_spotted_fever", "ticks", "ehrlichia", "lyme disease", "lyme diseases, oa?", "lyme/anaplasma", "lyme+", "anaplasma and lyme +", "positive anaplasma", "erlichia +", "erlichia","valley_fever","babesia","anaplasma","ehrlichia","lyme_disease")
other_parasitism <-c("fleas", "fungal_infection", "heartworm", "ear_mites","other_ectoparasite","heartworm_infection")
all_parasitism<-c("giardia", "hookworms", "coccidia", "tapeworms", "whipworms", "roundworms", "eimeria", "isospora","lyme_disease", "anaplasma", "rocky_mountain_spotted_fever", "ticks", "ehrlichia", "lyme disease", "lyme diseases, oa?", "lyme/anaplasma", "lyme+", "anaplasma and lyme +", "positive anaplasma", "erlichia +", "erlichia","fleas", "fungal_infection", "heartworm", "ear_mites","heartworm_infection","other_ectoparasite")
infectious <- c("leptospirosis","parvovirus","bacterial cystitis/recurrent uti", "chronic uti's/ bacterial cystitis", "cystitis", "uti", "frequent utis", "previous utis", "chronic urinary incontinence","history of valley fever", "mma- coccidioides")
orthopaedic <-c("osteoarthritis", "l ccl partial tear", "oa of stifle", "hip dysplasia", "right elbow dysplasia grade ii", "ccl rupture", "acl tear", "tendonitis/ arthritis", "arthritis- cervical neck spine", "lh partial ccl tear, lh iliopsoas injury", "chronic lameness- hx of car accident", "lameness","hypothyroidism, chronic lameness, obesity","overweight, arthritis","dvd, larpar", "seizures, oa","tail fracture", "broken leg, histiocytoma","osteochondritis_dissecans","intervertebral_disc_disease","lameness","musculoskeletal_pain","osteomyeltiis","cruciate_ligament_rupture","hip_dysplasia","elbow_dysplasia","crepitus_decreased_range_of_motion","bone_fractures","limber_tail_syndrome","panosteitis")
immune_mediated <- c("uveitis", "autoimmune_thyroiditis", "keratoconjunctivitis_sicca", "megaesophagus", "diabetes_mellitus", "cushings_disease", "hypothyroidism", "granuloma", "immune complex formation", "suspected cushings/ thyroid carcinoma","dvd, larpar","immune_related_musculoskeletal_disorder","rheumatoid_arthritis","diabetes_insipidus","hypothyroidism","pancreatic_insuficiency","addisons_disease","autoimmune_thyroiditis","cushings_disease","hyperadrenalestrogenism","hypercalcemia")
cardiovascular <- c("cardiomyopathy", "hypertension", "heart base tumor", "heart base mass", "mild degenerative valve disease", "degenerative mitral and tricuspid valve disease", "mitral valve disease", "pulmonary thrombosis", "portal vein thrombus","arrhythmia","atrial enlargement","cardiomegaly","cardiomyopathy","congestive_heart_failure","hypertension","murmur","lower_airway_disease","myocardial_dysfunction","pleural_effusion","pulmonic_stenosis","pneumothorax","pulmonic_stenosis","subaortic_stenosis","tachycardia","valvular_disease")
gastrointestinal <-c("chronic_colitis", "pancreatitis", "malabsorptive_disorder", "chronic pancreatitis", "cystic endometrial hyperplasia", "intermittent mixed bowel diarrhea", "intermittent diarrhea - bacterial", "gastric ulcers","mesenteric volvulus","fb","esophagitis","diarrhea","dietary_indiscretion","food_sensitivity","gastrointestinal_foreign_body","gastorintestinal_ulcer","intussusception","malabsorptive_disorder","megaesophagus")

inflammatory_other <- c("tracheobronchitis", "hypercalcaemia", "pigmentary_uveitis", "ear infections", "chronic pyoderma", "chronic pruritis", "chronic rhinitis/ nasal discharge", "allergy, obesity", "pruritis", "severe allergies", "atopy", "allergies, chronic otitis", "chronic hot spots, skin infections", "many pyodera/dermatitis events", "chronic otitis, allergic dermatitis", "dermatitis", "food allergy", "skin allergies", "chronic otitis, overweight", "atopic dermatitis, allergies", "seasonal allergies/pyoderma", "chronic otitis, chronic allergies", "dermatitis lip fold","anal sac infection","otitis_externa","otitis_media","aural_haematoma","aural_foreign_body","pharyngitis","rhinitis","hypercalcaemia","keratoconjunctivitis_sicca","hypercalcemia","influenza","parvovirus")


chronic_inflammatory <-c("congestive_heart_failure",
  "cardiomyopathy",
  "hypertension",
  "heartworm_infection",
  "lower_airway_disease",
  "dental_tarter_calculus",
  "addisons_disease",
  "autoimmune_thyroiditis",
  "cushings_disease",
  "diabetes_insipidus",
  "diabetes_mellitus",
  "hyperadrenalestrogenism",
  "hypercalcemia",
  "hypothyroidism",
  "pancreatic_insufficiency",
  "keratoconjunctivitis_sicca",
  "pigmentary_uveitis",
  "uveitis",
  "chronic_colitis",
  "malabsorptive_disorder",
  "megaesophagus",
  "pancreatitis",
  "anaplasma",
  "babesia",
  "canine_herpesvirus",
  "coccidia",
  "cryptosporidium",
  "ear_mites",
  "ehrlichia",
  "eimeria",
  "fleas",
  "fungal_infection",
  "giardia",
  "granuloma",
  "hookworms",
  "influenza",
  "isospora",
  "leptospirosis",
  "lyme_disease",
  "other_ectoparasite",
  "parvovirus",
  "rocky_mountain_spotted_fever",
  "roundworms",
  "tapeworms",
  "ticks",
  "tracheobronchitis",
  "unspecified_gastrointestinal_parasite",
  "valley_fever",
  "whipworms",
  "immune_related_musculoskeletal_disorder",
  "osteoarthritis",
  "rheumatoid_arthritis",
  "laryngeal_paralysis",
  "meningitis",
  "meningoencephalitis",
  "myasthenia_gravis",
  "erythema_multiforme",
  "flea_allergy_dermatitis",
  "food_allergy_dermatitis",
  "r_o_atopy")
```

#### Chronic inflam conditions ever in all years of study

```{r}
source(here::here("Code/GRLS_functions.R"))

conditions_in_df_and_list <- intersect(colnames(comorbs_unique),comorbs_to_find)

comorbs_unique_2 <- comorbs_unique %>%
  #get rid of the to_date row for each dog
  filter(to_date !=1) %>%
  #get rid of the NA rows in the dataset - then add in later at the 3044 level to be "no_records_available" across the comorbs
  dplyr::select(-c(relationship_category,year_in_study,record_date,to_date))
###not using currently as want to keep the NA to tidy at end into "no_records_available"
#comorbs_unique_2<- comorbs_unique_2%>%
# replacing NA values with 0 across comorbidity columns
 # mutate_at(vars(3:196), ~replace_na(., 0))


#remove the rows in the comorbs_unique_2 that have got NAs across whole dataset for each dog. then when creating cohort variable at end can recode any of these dogs with no records as "no_records_available" (very few missing records across WHOLE study period, generally just occ year)

cols_to_check <- c(7:21,164:196)
comorbs_unique_2 <- comorbs_unique_2[complete.cases(comorbs_unique_2[, cols_to_check]), ]
#checking all dogs have some records to see if need to code any with "no_records_available" at end - all dogs have comorbs records in their lifetime
comorbs_unique <- unique(comorbs_unique_2$subject_id)

df_result2 <- comorbs_unique_2 %>%
  group_by(subject_id) %>%
  mutate(
    #is there any chronic inflam condition present in at least 1 row for subject id
    chronic_inflam_present_lifetime = if_else(
      rowSums(across(all_of(conditions_in_df_and_list))) > 0 & 
        rowSums(across(all_of(conditions_in_df_and_list), ~ sum(. == 1)) > 1), 
      1, 
      0
    ),
    #how many chronic inflam conditions
    number_comorbs_present_lifetime = if_else(
      chronic_inflam_present_lifetime == 1,
      rowSums(across(all_of(conditions_in_df_and_list))),
      0
    ),
    ,
    #what were the comorbidities
    comorbidities_lifetime = if_else(
      chronic_inflam_present_lifetime == 1,
      apply(across(all_of(conditions_in_df_and_list)), 1, function(x) paste(names(which(x == 1)), collapse = "_")),
      "")
  ) %>%
  ungroup()


```

#### Comorbs in final 5y prior to endpoint

```{r}
#keeping record_date for linking up to cancer diagnosis but remove to_date column
comorbs_unique_3 <- comorbs_unique2 %>%
#  mutate_at(vars(5:200), ~replace_na(., 0)) %>%
  filter(to_date !=1) %>%
  dplyr::select(-c(relationship_category,to_date)) %>%
  rename(year_in_study_comorb = year_in_study)
cols_to_check <- c(5:23,26:164,165:198)
comorbs_unique_3 <- comorbs_unique_3[complete.cases(comorbs_unique_3[, cols_to_check]), ]

##checking all dogs have some records to see if need to code any with "no_records_available" at end - all dogs have comorbs records for at least some of last 5y = dont need to do "record_unavailable"
comorbs_unique_5y <- unique(comorbs_unique_3$subject_id)

#run all functions above onto this dataset
df_result3 <- comorbs_unique_3 %>%
  group_by(subject_id) %>%
  mutate(
    #is there any chronic inflam condition present in at least 1 row for subject id
    chronic_inflam_present_5y = if_else(
      rowSums(across(all_of(conditions_in_df_and_list))) > 0 & 
        rowSums(across(all_of(conditions_in_df_and_list), ~ sum(. == 1)) > 1), 
      1, 
      0
    ),
    #how many chronic inflam conditions
    number_comorbs_present_5y = if_else(
      chronic_inflam_present_5y == 1,
      rowSums(across(all_of(conditions_in_df_and_list))),
      0
    ),
    ,
    #what were the comorbidities
    comorbidities_5y = if_else(
      chronic_inflam_present_5y == 1,
      apply(across(all_of(conditions_in_df_and_list)), 1, function(x) paste(names(which(x == 1)), collapse = "_")),
      "")
  ) %>%
  ungroup()


#keep only top 5 rows per patient 
df_result_restricted <- df_result3 %>%
  group_by(subject_id) %>%
  filter(year_in_study_comorb %in% sort(unique(year_in_study_comorb), decreasing = TRUE)[1:5]) %>%
  ungroup()
  


# Function to detect comorbidities for each subject_id based on index range
detect_comorbidity <- function(df, subject_id_col, index_range, comorbidity_groups) {
  # Extract actual column names within the index range
  valid_columns <- colnames(df)[index_range]

  df %>%
    group_by(!!sym(subject_id_col)) %>%  # Group by subject_id
    summarise(across(all_of(valid_columns), ~ any(. == 1), .names = "temp_{.col}"), .groups = "drop") %>%  # Check for 1s
    mutate(
      !!!map(comorbidity_groups, function(cols) {
        # Identify valid columns that exist in the dataset
        matched_cols <- intersect(paste0("temp_", cols), colnames(.))
        
        # Assign "Y" if any value in matched columns is 1, else "N"
        if (length(matched_cols) > 0) {
          ifelse(rowSums(dplyr::select(., all_of(matched_cols))) > 0, "Y", "N")
        } else {
          "N"  # If no matching columns exist, default to "N"
        }
      })
    ) %>%
    dplyr::select(-starts_with("temp_"))  # Remove temporary columns
}

# Define comorbidity groups (list of column names matching each category)
comorbidity_groups <- list(
  intestinal_parasitism = intestinal_parasitism,
  tick_borne_parasitism = tick_borne_parasitism,
  other_parasitism = other_parasitism,
  all_parasitism = all_parasitism,
  infectious = infectious,
  orthopaedic = orthopaedic,
  immune_mediated = immune_mediated,
  cardiovascular = cardiovascular,
  gastrointestinal = gastrointestinal,
  inflammatory_other = inflammatory_other,
  chronic_inflammatory = chronic_inflammatory
)

# Apply function to data
GRLS_data <- df_result_restricted %>%
  left_join(detect_comorbidity(df_result_restricted, "subject_id", 5:198, comorbidity_groups),
            by = "subject_id")


#keep only a single row for each patient
# Define the index range of comorbidity columns
comorbidity_index <- c(199:212)  # Keep subject_id out of this list

# Create a second dataframe with one row per subject_id, summarizing only the selected columns - not using this as need to ensure rows with a 1 are preferentially kept
#GRLS_summary <- GRLS_data %>%
#  group_by(subject_id) %>%
#   arrange(desc(number_comorbs_present_5y)) %>%
#  summarise(
#    subject_id = first(subject_id),  # Keep subject_id as is
#    across(all_of(colnames(GRLS_data)[comorbidity_index]), ~ ifelse(any(. == "Y"), "Y", "N")),
 #   .groups = "drop"
 # )


GRLS_summary <- GRLS_data %>%
  group_by(subject_id) %>%
  arrange(desc(number_comorbs_present_5y)) %>%  # prioritise higher values
  slice(1) %>%  # keep only the top row per subject
  ungroup()


### recode any NA as "no_record_available"
GRLS_summary <- GRLS_summary%>%
  mutate(across(where(~ is.character(.) | is.factor(.)| is.numeric(.)),
                ~ replace(., is.na(.) | . == "", "no_records_available")))
```

There are also a separate subset of conditions that if ever reported as diagnosed then will have for life (so do not need \>1 y study records with it available):

```{r}
comorbs_single_record <- c(
  "atrial_enlargement",
  "cardiomegaly",
  "cardiomyopathy",
  "congestive_heart_failure",
  "hypertension",
  "myocardial_dysfunction",
  "valvular_disease",
  "addisons_disease",
  "autoimmune_thyroiditis",
  "cushings_disease",
  "diabetes_insipidus",
  "diabetes_mellitus",
  "hyperadrenalestrogenism",
  "hypercalcemia",
  "hypothyroidism",
  "pancreatic_insufficiency",
  "cataracts",
  "glaucoma",
  "keratoconjunctivitis_sicca",
  "progressive_retinal_atrophy_or_degeneration",
  "pigmentary_uveitis",
  "lymphangiectasia",
  "malabsorptive_disorder",
  "megaesophagus",
  "motility_disorder",
  "elbow_dysplasia",
  "hip_dysplasia",
  "osteoarthritis",
  "rheumatoid_arthritis",
  "spondylosis"
)

#identify in records

comorbs_ever <- comorbs_unique3 %>%
  dplyr::select(-c(relationship_category,year_in_study,record_date,to_date))

#comorbs_ever<- comorbs_ever%>%
# replacing NA values with 0 across comorbidity columns
 # mutate_at(vars(3:196), ~replace_na(., 0))


### restrict columns so that those with all NAs are removed and then can recode back as "no_records_available" later
cols_to_check <- c(3:21,24:162,164:196)
comorbs_ever<- comorbs_ever[complete.cases(comorbs_ever[, cols_to_check]), ]
#all dogs have at least some records here
unique_comorbs_ever <- unique(comorbs_ever$subject_id)
#remove all the columns that are not in comorbs_single_record list or subject_id
comorbs_ever <- comorbs_ever %>%
  dplyr::select(subject_id, all_of(comorbs_single_record))

#
lifetime_comorbs_ever <- comorbs_ever %>%
  group_by(subject_id) %>%
  mutate(
    ## ANY lifetime comorb inflam present
    lifelong_comorb_present = if_else(rowSums(across(all_of(comorbs_single_record))) > 0, 1, 0),
    
    ### NUMBER OF chronic inflam conditions present
    number_lifelong_comorbs_present = rowSums(across(all_of(comorbs_single_record))),
    
    # what were the comorbidities
    lifelong_comorbidities = if_else(
      lifelong_comorb_present == 1,
      apply(across(all_of(comorbs_single_record)), 1, function(x) paste(names(which(x == 1)), collapse = "_")),
      NA_character_
    )
  ) %>%
  ungroup()

#keep count of comorbs and if present and then only a unique row for each patient
lifetime_comorbs_ever_abr <- lifetime_comorbs_ever %>%
    dplyr::select(c(subject_id,lifelong_comorb_present,number_lifelong_comorbs_present))%>%
  group_by(subject_id) %>%
  arrange(desc(number_lifelong_comorbs_present)) %>%  # prioritise higher values
  slice(1) %>%  # keep only the top row per subject
  ungroup()
```

Grouping the lifelong_comorbidities column into type of condition e.g endocrine, ocular etc based on earlier lists

```{r}

valid_comorbidity_groups <- lapply(comorbidity_groups, function(cols) {
  intersect(cols, colnames(lifetime_comorbs_ever))  # Keep only columns that exist in df
})

# Remove empty comorbidity groups
valid_comorbidity_groups <- valid_comorbidity_groups[lengths(valid_comorbidity_groups) > 0]





# Compute 'lifetime' flags for each comorbidity category
df_lifetime <- lifetime_comorbs_ever %>%
  group_by(subject_id) %>%
  summarise(across(everything(), \(x) max(x, na.rm = TRUE))) %>%  # Collapse per patient (1 if present, 0 otherwise)
  mutate(!!!setNames(
    lapply(valid_comorbidity_groups, function(cols) {
      if (length(cols) > 0) {
        as.integer(rowSums(.[cols], na.rm = TRUE) > 0)  # Correctly sum across selected columns
      } else {
        rep(0, n())  # Handle cases where no valid columns exist
      }
    }),
    paste0("lifelong_", names(valid_comorbidity_groups))  # Create new column names
  )) %>%
  select(subject_id, starts_with("lifelong_"))  # Keep only new columns

# View output
print(df_lifetime)
```

Combine the lifetime_comorbs_ever_abr and df_lifetime so that we have counts and groupings of these comorbs:

```{r}
lifelong_data <- df_lifetime %>%
  left_join(lifetime_comorbs_ever_abr,by="subject_id")

lifelong_data <- lifelong_data %>%
  dplyr::select(-c(lifelong_comorb_present.y))


GRLS_summary_lifelong<- lifelong_data %>%
  left_join(GRLS_summary,by="subject_id")

#add in _5y to the GRLS summary columns that are from 5y prior to final record
colnames(GRLS_summary_lifelong)[211:221] <- paste0(colnames(GRLS_summary_lifelong)[211:221], "_5y")

```

Lifetime comorbs now classify into the parasitism/infectious/cardio/ etc

```{r}

# Apply function to data
GRLS_data_lifetime <- df_result3 %>%
  left_join(detect_comorbidity(df_result_restricted, "subject_id", 5:198, comorbidity_groups),
            by = "subject_id")

#add in _5y to the GRLS summary columns that are from 5y prior to final record
#colnames(GRLS_data_lifetime)[202:212] <- paste0(colnames(GRLS_data_lifetime)[202:212], "_5y")

#keep only a single row for each patient
# Define the index range of comorbidity columns
comorbidity_index <- c(199:212)  # Keep subject_id out of this list

# Create a second dataframe with one row per subject_id, summarizing only the selected columns
GRLS_summary_lifetime <- GRLS_data_lifetime %>%
  group_by(subject_id) %>%
  summarise(
    subject_id = first(subject_id),  # Keep subject_id as is
    across(all_of(colnames(GRLS_data_lifetime)[comorbidity_index]), ~ ifelse(any(. == "Y"), "Y", "N")),
    .groups = "drop"
  )
#add lifetime to end of columns at 5:15 indices
names(GRLS_summary_lifetime)[5:15] <- paste0(names(GRLS_summary_lifetime)[5:15], "_lifetime")

```

Combine lifelong and 5y values into one as consider these comorbs to be permanent once diagnosed:

```{r}
#mutate the Y N in 5y cols to be 0 and 1 for counting in next step



GRLS_summary_lifelong <- GRLS_summary_lifelong %>%
  mutate(across(c(4:9, 211:221), ~ case_when(
    . == "N" ~ 0,
    . == "Y" ~ 1,
    TRUE ~ as.numeric(.)
  )))

#combining counts in 5y columns and then remove the lifelong ones 
GRLS_summary_lifelong <- GRLS_summary_lifelong %>%
  mutate(
    orthopaedic_5y2 = pmax(lifelong_orthopaedic, orthopaedic_5y, na.rm = TRUE),
    cardiovascular_5y2=pmax(lifelong_cardiovascular,cardiovascular_5y,na.rm=TRUE),
    immune_mediated_5y2=pmax(lifelong_immune_mediated,immune_mediated_5y,na.rm=TRUE),
    gastrointestinal_5y2=pmax(lifelong_gastrointestinal,gastrointestinal_5y,na.rm=TRUE),
    inflammatory_other_5y2=pmax(lifelong_inflammatory_other,inflammatory_other_5y,na.rm=TRUE),
    chronic_inflammatory_5y2=pmax(lifelong_chronic_inflammatory,chronic_inflammatory_5y,na.rm=TRUE))

#remove the earlier 5y columns and just keep the ones where lifelong + 5y ar ecombined
GRLS_summary_lifelong_2 <- GRLS_summary_lifelong %>%
  dplyr::select(c(subject_id,222:227))
```

Combine all of the 5y prior and ever comorb datasets

```{r}
GRLS_5yprior_lifetime <- GRLS_summary_lifelong_2 %>%
  left_join(GRLS_summary_lifetime,by="subject_id")

GRLS_5yprior_lifetime <- GRLS_5yprior_lifetime %>%
  dplyr::select(-c(8:10))

GRLS_5yprior_lifetime <- GRLS_5yprior_lifetime %>%
  mutate(across(c(8:18), ~ case_when(
    . == "N" ~ 0,
    . == "Y" ~ 1,
    TRUE ~ as.numeric(.)
  )))
```

#### Add neoplasia variables in comorb dataset - lifetime

NB. remove lymphoma from lists as will skew for all the cases to have a 1 and none of the controls by definition

```{r}
#ggrouping these
all_cancers <- c("adrenal_tumor","basal_cell_tumor","bile_duct_tumor","bladder_tumor","brain_spinal_cord_tumor","breast_or_mammary_tumor","eye_tumor","hair_matrix_tumor","heart_tumor","hemangiosarcoma",
                 "histiocytic_sarcoma","kidney_tumor","liver_tumor","lung_tumor","mast_cell_tumor","melanmoma","multiple_myeloma","nasal_tumor","osteosarcoma","pancreatic_tumor","perianal_adenoma","pituitary_tumor","plasma_cell_tumor","plasmacytoma","prostate_tumor","soft_tissue_sarcoma","splenic_tumor","squamous_cell_carcinoma","stomach_intestinal_tumor","testicular_tumor","thymoma","thyroid_tumor")
malig_cancers <- c("hemangiosarcoma",
  "histiocytic_sarcoma",  "melanoma", "multiple_myeloma", "osteosarcoma", "pancreatic_tumor", "plasma_cell_tumor", "plasmacytoma", "soft_tissue_sarcoma", "squamous_cell_carcinoma")
benign_cancers <- c("adrenal_tumor", "basal_cell_tumor", "bile_duct_tumor", "bladder_tumor", "brain_spinal_cord_tumor", "breast_or_mammary_tumor", "eye_tumor", "hair_matrix_tumor", "heart_tumor", "kidney_tumor", "liver_tumor", "lung_tumor", "mast_cell_tumor", "nasal_tumor", "perianal_adenoma", "pituitary_tumor", "prostate_tumor", "splenic_tumor", "stomach_intestinal_tumor", "testicular_tumor", "thymoma", "thyroid_tumor")

```

Create YN and counts for all cancers, benign and malig

```{r}
comorbs_neoplasia_2 <- conditions_neoplasia %>%
  filter(to_date != 1)%>%
  dplyr::select(-other_specify)

#for next step of 5y
comorbs_neoplasia_3 <- conditions_neoplasia %>%
  filter(to_date != 1)%>%
  dplyr::select(-other_specify)
#remove rows where comorbs NA
cols_to_check <- c(6:40)
comorbs_neoplasia_2 <- comorbs_neoplasia_2[complete.cases(comorbs_neoplasia_2[, cols_to_check]), ]


#check if any dogs have no neoplasia records - all do have some records at least for lifetime calcs
unique_cancer <- unique(comorbs_neoplasia_2$subject_id)
comorbs_neoplasia_2$all_cancers_count_lifetime <- 0
comorbs_neoplasia_2$malig_cancers_count_lifetime <- 0
comorbs_neoplasia_2$benign_cancers_count_lifetime <- 0


# Define the cancer columns indices (from index 2 to the last column)
cancer_columns <- 7:40

# Loop through the cancer columns by index
for (col_index in cancer_columns) {
  col_name <- colnames(comorbs_neoplasia_2)[col_index]  # Get column name
  
  # Count for all cancers
  comorbs_neoplasia_2$all_cancers_count_lifetime <- comorbs_neoplasia_2$all_cancers_count_lifetime + (comorbs_neoplasia_2[[col_name]] == 1)
  
  # Count for all cancers
  comorbs_neoplasia_2$all_cancers_count_lifetime <- comorbs_neoplasia_2$all_cancers_count_lifetime + (comorbs_neoplasia_2[[col_name]] == 1)
  
  # Count for malignant cancers
  if (col_name %in% malig_cancers) {
    comorbs_neoplasia_2$malig_cancers_count_lifetime <- comorbs_neoplasia_2$malig_cancers_count_lifetime + (comorbs_neoplasia_2[[col_name]] == 1)
  }
  
  # Count for benign cancers
  if (col_name %in% benign_cancers) {
    comorbs_neoplasia_2$benign_cancers_count_lifetime <- comorbs_neoplasia_2$benign_cancers_count_lifetime + (comorbs_neoplasia_2[[col_name]] == 1)
  }
}

#create Y N columns for each in addition to counts
comorbs_neoplasia_2 <- comorbs_neoplasia_2 %>%
  mutate(
    all_cancers_YN_lifetime = ifelse(all_cancers_count_lifetime >= 1, 1, 0),
    malig_cancers_YN_lifetime = ifelse(malig_cancers_count_lifetime >= 1, 1, 0),
    benign_cancers_YN_lifetime = ifelse(benign_cancers_count_lifetime >= 1, 1, 0),

  )


#code to get to unique rows and keep the max count for each dog
#get the max count value for each of the counts for each dog 
cancer_count_summary <- comorbs_neoplasia_2 %>%
  group_by(subject_id) %>%
  summarise(
    malig_cancers_count_lifetime = if_else(all(is.na(malig_cancers_count_lifetime)), 0, max(malig_cancers_count_lifetime, na.rm = TRUE)),
    benign_cancers_count_lifetime = if_else(all(is.na(benign_cancers_count_lifetime)), 0, max(benign_cancers_count_lifetime, na.rm = TRUE)),
    all_cancers_count_lifetime = if_else(all(is.na(all_cancers_count_lifetime)), 0, max(all_cancers_count_lifetime, na.rm = TRUE)),
    malig_cancers_count_lifetime = if_else(all(is.na(malig_cancers_count_lifetime)), 0, max(malig_cancers_count_lifetime, na.rm = TRUE)),
    all_cancers_count_lifetime = if_else(all(is.na(all_cancers_count_lifetime)), 0, max(all_cancers_count_lifetime, na.rm = TRUE)),

    malig_cancers_YN_lifetime = if_else(all(is.na(malig_cancers_YN_lifetime)), 0, max(malig_cancers_YN_lifetime, na.rm = TRUE)),
    benign_cancers_YN_lifetime = if_else(all(is.na(benign_cancers_YN_lifetime)), 0, max(benign_cancers_YN_lifetime, na.rm = TRUE)),
    all_cancers_YN_lifetime = if_else(all(is.na(all_cancers_YN_lifetime)), 0, max(all_cancers_YN_lifetime, na.rm = TRUE)),
    malig_cancers_YN_lifetime = if_else(all(is.na(malig_cancers_YN_lifetime)), 0, max(malig_cancers_YN_lifetime, na.rm = TRUE)),
    all_cancers_YN_lifetime = if_else(all(is.na(all_cancers_YN_lifetime)), 0, max(all_cancers_YN_lifetime, na.rm = TRUE)),
    
    .groups = "drop"
  )

```

#### Neoplasia YN and counts - 5y prior to endpoint

```{r}
#keep only the max 5 rows for that dog in year_in_study then same calcs as above
comorbs_neoplasia_3 <- comorbs_neoplasia_3 %>%
  group_by(subject_id) %>%
  #keeping only the 5 most recent records
  filter(year_in_study %in% sort(unique(year_in_study), decreasing = TRUE)[1:5]) %>%
  ungroup()


#remove rows where comorbs NA
cols_to_check <- c(6:40)
comorbs_neoplasia_3 <- comorbs_neoplasia_3[complete.cases(comorbs_neoplasia_3[, cols_to_check]), ]


#check if any dogs have no neoplasia records - only 2165 have records fo final 5y
unique_cancer2 <- unique(comorbs_neoplasia_3$subject_id)
  

comorbs_neoplasia_3$all_cancers_count_5y <- 0
comorbs_neoplasia_3$malig_cancers_count_5y <- 0
comorbs_neoplasia_3$benign_cancers_count_5y <- 0


# Define the cancer columns indices (from index 2 to the last column)
cancer_columns <- 6:41

# Loop through the cancer columns by index
for (col_index in cancer_columns) {
  col_name <- colnames(comorbs_neoplasia_3)[col_index]  # Get column name
  
  # Count for all cancers
  comorbs_neoplasia_3$all_cancers_count_5y <- comorbs_neoplasia_3$all_cancers_count_5y + (comorbs_neoplasia_3[[col_name]] == 1)
  
  # Count for all cancers
  comorbs_neoplasia_3$all_cancers_count_5y <- comorbs_neoplasia_3$all_cancers_count_5y + (comorbs_neoplasia_3[[col_name]] == 1)
  
  # Count for malignant cancers
  if (col_name %in% malig_cancers) {
    comorbs_neoplasia_3$malig_cancers_count_5y <- comorbs_neoplasia_3$malig_cancers_count_5y + (comorbs_neoplasia_3[[col_name]] == 1)
  }
  
  # Count for benign cancers
  if (col_name %in% benign_cancers) {
    comorbs_neoplasia_3$benign_cancers_count_5y <- comorbs_neoplasia_3$benign_cancers_count_5y + (comorbs_neoplasia_3[[col_name]] == 1)
  }
}

#create Y N columns for each in addition to counts
comorbs_neoplasia_3 <- comorbs_neoplasia_3 %>%
  mutate(
    all_cancers_YN_5y = ifelse(all_cancers_count_5y >= 1, 1, 0),
    malig_cancers_YN_5y = ifelse(malig_cancers_count_5y >= 1, 1, 0),
    benign_cancers_YN_5y = ifelse(benign_cancers_count_5y >= 1, 1, 0),

  )


#code to get to unique rows and keep the max count for each dog
#get the max count value for each of the counts for each dog 
cancer_count_summary_2 <- comorbs_neoplasia_3 %>%
  group_by(subject_id) %>%
  summarise(
    malig_cancers_count_5y = if_else(all(is.na(malig_cancers_count_5y)), 0, max(malig_cancers_count_5y, na.rm = TRUE)),
    benign_cancers_count_5y = if_else(all(is.na(benign_cancers_count_5y)), 0, max(benign_cancers_count_5y, na.rm = TRUE)),
    all_cancers_count_5y = if_else(all(is.na(all_cancers_count_5y)), 0, max(all_cancers_count_5y, na.rm = TRUE)),
    malig_cancers_count_5y = if_else(all(is.na(malig_cancers_count_5y)), 0, max(malig_cancers_count_5y, na.rm = TRUE)),
    all_cancers_count_5y = if_else(all(is.na(all_cancers_count_5y)), 0, max(all_cancers_count_5y, na.rm = TRUE)),

    malig_cancers_YN_5y = if_else(all(is.na(malig_cancers_YN_5y)), 0, max(malig_cancers_YN_5y, na.rm = TRUE)),
    benign_cancers_YN_5y = if_else(all(is.na(benign_cancers_YN_5y)), 0, max(benign_cancers_YN_5y, na.rm = TRUE)),
    all_cancers_YN_5y = if_else(all(is.na(all_cancers_YN_5y)), 0, max(all_cancers_YN_5y, na.rm = TRUE)),
    malig_cancers_YN_5y = if_else(all(is.na(malig_cancers_YN_5y)), 0, max(malig_cancers_YN_5y, na.rm = TRUE)),
    all_cancers_YN_5y = if_else(all(is.na(all_cancers_YN_5y)), 0, max(all_cancers_YN_5y, na.rm = TRUE)),
    
    .groups = "drop"
  )

#combine the 2 cancer summaries
cancer_summary_5y_lifetime <- cancer_count_summary %>%
  left_join(cancer_count_summary_2,by="subject_id")

#replace all NAs with "no_records_available" for the 5y prior ones which dont have any records for neoplasia in final 5y
cancer_summary_5y_lifetime <- cancer_summary_5y_lifetime %>%
  mutate(across(2:13, ~ case_when(
    is.na(.) ~ "no_records_available",
    TRUE ~ as.character(.)
  )))
```

Combine all comorbs data back into the main df

```{r}
#combine cancer and other comorbs
GRLS_comorbs_cancer_5yprior_lifetime <- GRLS_5yprior_lifetime %>%
  left_join(cancer_summary_5y_lifetime,by="subject_id")

#combine back to main df
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location2 %>% left_join(GRLS_comorbs_cancer_5yprior_lifetime,by="subject_id")
```

## Medications

Using the df created in an earlier script of analyses which tidies up free text and the many drug categories etc. in this qmd GRLS medications.qmd

```{r}
medications <- read.csv(here::here("Output/GRLS_medications_initial_tidy.csv"))
source(here::here("R", "medications.R"))
```

### Lifetime medication calculations

```{r}
medication_features_lifetime <- summarise_medication_quartiles(
  medications,
  suffix = "lifetime",
  missing_value = "medication_not_prescribed"
)

meds_per_dog_lifetime <- medication_features_lifetime$counts
medication_summary_lifetime <- medication_features_lifetime$summary
meds_per_dog_quart_lifetime <- medication_features_lifetime$categorised
meds_wide_quart_lifetime <- medication_features_lifetime$wide

write.csv(medication_summary_lifetime,here::here("Output/MCT_study_medication_quartiles_summaries_lifetime.csv"))
meds_wide_quart_lifetime
```

### 5y prior to endpoint medication calculations

NB. there is 2889 dogs in this 5y - dogs with meds records in 5y prior to death, ensure when joining back into main 3044 dogs that these \~150 dogs have "no_records_available" coded on them

```{r}
medications_5y <- select_last_study_years(medications, 5)

medication_features_5y <- summarise_medication_quartiles(
  medications_5y,
  suffix = "5y",
  missing_value = "medication_not_used"
)

meds_per_dog_5y <- medication_features_5y$counts
medication_summary_5y <- medication_features_5y$summary
meds_per_dog_quart_5y <- medication_features_5y$categorised
meds_wide_quart_5y <- medication_features_5y$wide

write.csv(medication_summary_5y,here::here("Output/MCT_study_medication_quartiles_summaries_5y.csv"))
meds_wide_quart_5y
```

### Examine for specific potential RF medications:

-   NSAIDs exposure

    -   ever Y N and 5y prior Y N
    -   counts

-   steroids exposure

    -   ever Y N and 5y prior Y N
    -   counts

-   Fenbendazole exposure

    -   ever Y N and 5y prior Y N
    -   counts

```{r}
#free text drug names to look for
NSAID_names<-c("carprofen","meloxicam","firocoxib","deracoxib","flurbiprofen","piroxicam","ketotifen","ketoprofen","rimadyl","metacam","meloxidyl","previcox","vetprofen","ketorolac opthalmic ointment","toradol","acular","voltaren
","diclofenac","flurbiprofen eye drops
","carprovet","carprieve","truprofen","ofloxacin ophthalmic ointment + nsaid
")
steroid_names<-c("dexamethasone, neomycin sulfate, thiabendazole
","prednisolone acetate
","betamethasone, clotrimazole, gentamicin sulfate
","betamethasone, gentamicin sulfate
","dexamethasone","dexamethasone, miconazole nitrate
","dexamethasone, miconazole nitrate
","methylprednisolone","prednisone","betamethasone, florfenicol, terbinafine
","clotrimazole, gentamicin sulfate, mometasone furoate
","dexamethasone, methylprednisolone
","mometasone furoate, orbifloxacin, posaconazole
","dexamethasone, neomycin sulfate, polymyxin b
","bacitracin zinc, dexamethasone, neomycin sulfate, polymyxin b
","miconazole nitrate, polymyxin b, prednisolone acetate
","hydrocortisone, gentamicin sulfate, miconazole nitrate
","gentamicin sulfate, hydrocortisone, miconazole nitrate
","burow's solution, hydrocortisone
","acetic acid, boric acid, hydrocortisone, ketoconazole
","hydrocortisone","bacitracin zinc, hydrocortisone, neomycin sulfate, polymyxin b
","dexamethasone, enrofloxacin, miconazole nitrate
","clotrimazole, dexamethasone, enrofloxacin
","hydrocortisone, ketoconazole
","dexamethasone, ketoconazole, phytosphingosine
","betamethasone","acetic acid, boric acid, ceramide complex, hydrocortisone, ketoconazole
","hydrocortisone, troleandomycin
","dexamethasone, enrofloxacin, ketoconazole, tris-edta
","bacitracin zinc, neomycin sulfate, polymyxin b, unspecified steroid
","clotrimazole, dexamethasone
","dexamethasone, ketoconazole, tris-edta
","dexamethasone, ketoconazole
","betamethasone, clotrimazole, gentamicin sulfate, mometasone furoate
","dexamethasone, diphenhydramine
","beclomethasone, clotrimazole, neomycin sulfate
","dexamethasone, enrofloxacin, ketoconazole
","dexamethasone, dimethyl sulfoxide, enrofloxacin, lidocaine
","dexamethasone, enrofloxacin
","betamethasone, gentamicin sulfate, miconazole nitrate
","chloramphenicol, dexamethasone
","hydrocortisone, pramoxine
","dexamethasone, penicillin
","betamethasone, chloramphenicol, ketoconazole
","dexamethasone, tobramycin
","enrofloxacin, dexamethasone, tris-edta
","dexamethasone, miconazole nitrate, neomycin sulfate, nystatin, thiostrepton, triamcinolone acetonide
","betamethasone, burow's solution, gentamicin sulfate, hydrocortisone, miconazole nitrate
","ketoconazole, hydrocortisone
","bacitracin, hydrocortisone
","dexamethasone, miconazole nitrate, tris-edta
","chlorhexidine gluconate, dexamethasone
","dexamethasone, enrofloxacin, tris-edta
","dexamethasone, enrofloxacin, silver sulfadiazine
","dexamethasone, miconazole nitrate, neomycin sulfate, polymyxin b, prednisolone acetate, thiabendazole
","betamethasone, enrofloxacin
","dexamethasone, ear cleaner (epi-otic advanced), enrofloxacin
","betamethasone, tobramycin
","dexamethasone ophthalmic ointment
","miconazole + dexamethasone ear solution
","amikacin + dexamethasone ear solution
","depo-medrol
","prednisolone acetate ophthalmic suspension
","neo-poly-dex ophthalmic ointment
","neo-poly-bac ophthalmic ointment + dexamethasone
","betamethasone + gentamicin ear solution
","conofite + dexamethasone ear solution
","neo-poly-bac ophthalmic ointment + hydrocortisone
","baytril + conofite + dexamethasone ear solution
","betamethasone + gentamicin spray
","baytil + clotrimazole + dexamethasone ear solution
","baytril + dexamethasone + miconazole ear solution
","baytril + conzol + dexamethasone ear solution
","trizulta + keto + baytril + dexamethasone flush
","clotrimazole + dexamethasone ear solution
","burow's solution + hydrocortisone
","betamethasone + clotrimazole + gentamicin ear solution
","trizulta + keto + dexamethasone flush
","cortalone
","baytril + dexamethasone + ketoconazole ear solution
","betamethasone + gentamicin sulfate + miconazole nitrate ear solution
","dexamethasone otic
","baytril + trizulta + dexamethasone ear solution
","neomycin sulfate + nystatin + triamcinolone acetonide
","chloramphenicol + ketoconazole + triamcinolone acetonide ear solution
","triamcinolone","trizedta + dexamethasone + miconazole flush
","chlorhexiderm + dexamethasone spray
","trizedta + baytril + dexamethasone flush
","neomycin sulfate + tetracaine + triamcinolone acetonide
","clotrimazole + gentamicin + mometasone ear solution
","baytril + dexamethasone + silvadene ear solution
","itraconazole + ketoconazole + triamcinolone acetonide ear solution
","baytril + triamcinolone acetonide
","troleandomycin + hydrocortisone
","neo-poly-bac ophthalmic ointment + steroid
","t8 keto flush + dexamethasone
","dexamethasone + miconazole
","dermacool + hydrocortisone
","trizchlor 4 + hydrocortisone
","dexamethasone + diphenhydramine
","tear stimulation forte + betamethasone + ofloxacin
","baytril + dexamethasone + dimethyl sulfoxide + lidocaine
","dexamethasone + baytril
","ketoconazole + mupirocin + triamcinolone acetonide
","penicillin + dexamethasone
","tobramycin + betamethasone ophthalmic ointment
","lotramin + dexamethasone
","bacitracin + hydrocortisone
")
fenbendazole_names<-c("fenbendazole","fenbendazole, maropitant citrate, metronidazole, omeprazole, sucralfate
","cerenia + fenbendazole + metronidazole + omeprazole + sucralfate
","panacur","safe-guard")
```

Lifetime

```{r}
#NSAID
NSAID_ever<- medications %>%
  mutate(NSAID_ever = apply(select(., medication_name, medication_name_specify, medication_ingredients), 
                            1, 
                            function(row) {
                              # Check each column in the row for a match with NSAID_names
                              any(tolower(str_trim(row)) %in% tolower(str_trim(NSAID_names)))
                            })) %>%
  mutate(NSAID_ever= ifelse(NSAID_ever, 1, 0)) %>%
  arrange(subject_id, desc(NSAID_ever)) %>%  # Ensure NSAID_5y = 1 rows come first
  distinct(subject_id, .keep_all = TRUE) %>%  # Keep first row for each subject_id
  select(subject_id, NSAID_ever)

#steroid
steroid_ever<- medications %>%
  mutate(steroid_ever = apply(select(., medication_name, medication_name_specify, medication_ingredients), 
                            1, 
                            function(row) {
                              # Check each column in the row for a match with NSAID_names
                              any(tolower(str_trim(row)) %in% tolower(str_trim(steroid_names)))
                            })) %>%
  mutate(steroid_ever = ifelse(steroid_ever, 1, 0)) %>%
  arrange(subject_id, desc(steroid_ever)) %>%  # Ensure NSAID_ever = 1 rows come first
  distinct(subject_id, .keep_all = TRUE) %>%  # Keep first row for each subject_id
  select(subject_id, steroid_ever)
#fenbendazole
fenbendazole_ever <- medications %>%
  mutate(fenbendazole_ever = apply(select(., medication_name, medication_name_specify, medication_ingredients), 
                            1, 
                            function(row) {
                              # Check each column in the row for a match with NSAID_names
                              any(tolower(str_trim(row)) %in% tolower(str_trim(fenbendazole_names)))
                            })) %>%
  mutate(fenbendazole_ever = ifelse(fenbendazole_ever, 1, 0)) %>%
  arrange(subject_id, desc(fenbendazole_ever)) %>%  # Ensure NSAID_ever = 1 rows come first
  distinct(subject_id, .keep_all = TRUE) %>%  # Keep first row for each subject_id
  select(subject_id, fenbendazole_ever)
```

5y prior

```{r}

#NSAID
NSAID_5y <-medications_5y %>%
  ungroup()%>%
  mutate(NSAID_5y = apply(select(., medication_name, medication_name_specify, medication_ingredients), 
                            1, 
                            function(row) {
                              # Check each column in the row for a match with NSAID_names
                              any(tolower(str_trim(row)) %in% tolower(str_trim(NSAID_names)))
                            })) %>%
  mutate(NSAID_5y = ifelse(NSAID_5y, 1, 0)) %>%
  arrange(subject_id, desc(NSAID_5y)) %>%  # Ensure NSAID_5y = 1 rows come first
  distinct(subject_id, .keep_all = TRUE) %>%  # Keep first row for each subject_id
  select(subject_id, NSAID_5y)

#steroid
steroid_5y <- medications_5y %>%
    ungroup()%>%
  mutate(steroid_5y = apply(select(., medication_name, medication_name_specify, medication_ingredients), 
                            1, 
                            function(row) {
                              # Check each column in the row for a match with NSAID_names
                              any(tolower(str_trim(row)) %in% tolower(str_trim(steroid_names)))
                            })) %>%
  mutate(steroid_5y = ifelse(steroid_5y, 1, 0)) %>%
  arrange(subject_id, desc(steroid_5y)) %>%  # Ensure NSAID_5y = 1 rows come first
  distinct(subject_id, .keep_all = TRUE) %>%  # Keep first row for each subject_id
  select(subject_id, steroid_5y)
#fenbendazole
fenbendazole_5y <-medications_5y %>%
    ungroup()%>%
  mutate(fenbendazole_5y = apply(select(., medication_name, medication_name_specify, medication_ingredients), 
                            1, 
                            function(row) {
                              # Check each column in the row for a match with NSAID_names
                              any(tolower(str_trim(row)) %in% tolower(str_trim(fenbendazole_names)))
                            })) %>%
  mutate(fenbendazole_5y = ifelse(fenbendazole_5y, 1, 0)) %>%
  arrange(subject_id, desc(fenbendazole_5y)) %>%  # Ensure NSAID_5y = 1 rows come first
  distinct(subject_id, .keep_all = TRUE) %>%  # Keep first row for each subject_id
  select(subject_id, fenbendazole_5y)
```

Combine these dfs

```{r}
#all med counts of interest
meds_5y_lifetime <- meds_wide_quart_5y %>% left_join (meds_wide_quart_lifetime,by="subject_id")

#specific drugs of interest
combined_df <- reduce(list(NSAID_ever,NSAID_5y,steroid_ever,steroid_5y,fenbendazole_ever,fenbendazole_5y), full_join, by = "subject_id")

#combine both of these
meds_5y_lifetime_NSAID_steroid_fenbendazole <- meds_5y_lifetime %>%
  left_join(combined_df,by="subject_id")


#add onto the comorbs and all other variables df
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs %>% left_join(meds_5y_lifetime_NSAID_steroid_fenbendazole,by="subject_id")


#now add onto these for the ~150 dogs that dont have medicine records that they "no_medical_records_available"
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds <-
  dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds %>%
  mutate(across(187:250, ~ case_when(
    is.na(.) ~ "no_medication_records_for_this_time_period",
    TRUE ~ as.character(.)
  )))
```

#### Examine also an NSAID number of times variable

```{r}
# use HSA medications df to get list of NSAIDs in df
table(medications_5y $medication_ingredients_100)
unique(medications_5y $medication_ingredients_100)
unique(medications_5y $medication_ingredients)
NSAID_meds <- c("carprofen","meloxicam","firocoxib","deracoxib","flurbiprofen","ketoprofen")

medications_5y  <- medications_5y  %>%
  mutate(NSAID_use = if_else(
    str_detect(tolower(medication_ingredients), str_c(NSAID_meds, collapse = "|")),
    "NSAID_yes",
    "NSAID_no"
  ))

#a count of these for each dog exc if NSAID no then just NSAID no
NSAID_5y_meds_per_dog <- medications_5y  %>%
  group_by(subject_id, NSAID_use) %>%
  summarise(count = n(), .groups = "drop")
  #remove rows where if subject has ANY NSAID_yes value then remove the NSAID_no row/s for that subject id
  # Step 1: Get subject_ids that have NSAID_yes
NSAID_yes_subjects <- NSAID_5y_meds_per_dog %>%
  filter(NSAID_use == "NSAID_yes") %>%
  pull(subject_id)

# Step 2: Keep NSAID_yes rows or NSAID_no only if no NSAID_yes exists for that dog
NSAID_5y_meds_per_dog_clean <- NSAID_5y_meds_per_dog %>%
  filter(NSAID_use == "NSAID_yes" | !(subject_id %in% NSAID_yes_subjects))

#calculate NSAID yes quartiles
NSAID_yes <- NSAID_5y_meds_per_dog_clean %>%
  filter(NSAID_use=="NSAID_yes")

NSAID_summary <- NSAID_yes %>%
  summarise(
    min=min(count),
    Q1 = quantile(count, 0.25, na.rm = TRUE),
    median_count_per_dog = quantile(count, 0.50, na.rm = TRUE),  # Use quantile for clarity
    Q3 = quantile(count, 0.75, na.rm = TRUE),
    max=max(count),
    .groups = "drop"
  )

NSAID_5y_meds_per_dog_clean_quartiles <- NSAID_5y_meds_per_dog_clean %>%
  mutate(NSAID_quantile = case_when(
    NSAID_use == "NSAID_no" ~ "no_NSAID_admin",
    NSAID_use == "NSAID_yes" & count <= 1 ~ "Q1(1)",
    NSAID_use == "NSAID_yes" & count > 1 & count <= 2 ~ "Q2(1-2)",
    NSAID_use == "NSAID_yes" & count > 2 & count <= 3 ~ "Q3(2-3)",
    NSAID_use == "NSAID_yes" & count > 3 ~ "Q4(3-8)"
  ))
#then filt down to unique single row for each dog of its count or NSAID no, then join back to the 3044 dogs and if no count or no NSAID no then "no_NSAID_recorded_in_time_period"
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds %>% left_join(NSAID_5y_meds_per_dog_clean_quartiles, by="subject_id")

#now replace NAs in NSAID_quantile column with "no_med_records_this_time_period"
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID %>%
    mutate(NSAID_quantile = replace_na(NSAID_quantile, "no_med_use_recorded"))

#ditch count column 
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID%>%
  dplyr::select(c(-"count"))
#rename NSAID_quantile to make clear its a 5y column
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID %>%
  rename(NSAID_5y_quantiles = NSAID_quantile)


dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID <-
  dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID %>%
  mutate(across(251:252, ~ case_when(
    is.na(.) ~ "no_medication_records_for_this_time_period",
    TRUE ~ as.character(.)
  )))
```

#### Examine also a steroid number of times variable

```{r}
steroids <- c(
  "prednisone",
  "predisolone acetate",
  "trimeprazine tartrate, prednisone",
  "triamcinolone acetonide",
  "methylpredisolone",
  "hydrocortisone, gentamicin sulfate, miconazole nitrate",
  "burow's solution, hydrocortisone",
  "dexamethasone, miconazole nitrate",
  "dimethyl sulfoxide, fluocinolone acetonide",
  "hydrocortisone",
  "dexamethasone, enrofloxacin, ketoconazole, tris-edta",
  "beclomethasone, clotrimazole, neomycin sulfate",
  "dexamethasone, ketoconazole, tris-edta",
  "acetic acid, boric acid, ceramide complex, hydrocortisone, ketoconazole",
  "betamethasone",
  "mometasone furoate",
  "dexamethasone",
  "betamethasone",
  "hydrocortisone",
  "prednisolone",
  "methylprednisolone",
  "triamcinolone acetonide",
  "difluprednate",
  "fluticasone",
  "dexamethasone, miconazole nitrate",
  "amikacin sulfate, dexamethasone",
  "dexamethasone, methylprednisolone",
  "dexamethasone, triamcinolone acetonide",
  "hydrocortisone, ketoconazole",
  "dexamethasone, ketoconazole, phytosphingosine",
  "dexamethasone, diphenhydramine",
  "betamethasone, clotrimazole, gentamicin sulfate, mometasone furoate",
  "dimethyl sulfoxide, fluocinolone acetonide, gentamicin sulfate, miconazole nitrate",
  "dimethyl sulfoxide, flunixin, fluocinolone acetonide",
  "hydrocortisone, troleandomycin",
  "Dexamethasone",
  "Betamethasone",
  "dexamethasone",
  "betamethasone",
  "hydrocortisone",
  "menthol, hydrocortisone",
  "dexamethasone, ketoconazole",
  "dexamethasone, enrofloxacin",
  "dexamethasone, dimethyl sulfoxide, enrofloxacin, lidocaine",
  "dexamethasone, enrofloxacin, ketoconazole",
  "dexamethasone, penicillin",
  "dexamethasone, tobramycin",
  "betamethasone, ofloxacin",
  "chloramphenicol, dexamethasone",
  "betamethasone, chloramphenicol, ketoconazole",
  "enrofloxacin, dexamethasone, tris-edta",
  "dexamethasone, miconazole nitrate, neomycin sulfate, nystatin, thiostrepton, triamcinolone acetonide",
  "betamethasone, burow's solution, gentamicin sulfate, hydrocortisone, miconazole nitrate",
  "ketoconazole, hydrocortisone",
  "bacitracin, hydrocortisone",
  "dexamethasone, miconazole nitrate, tris-edta",
  "chlorhexidine gluconate, dexamethasone",
  "dexamethasone, enrofloxacin, tris-edta",
  "dexamethasone, enrofloxacin, silver sulfadiazine",
  "dexamethasone, miconazole nitrate, neomycin sulfate, polymyxin b, prednisolone acetate, thiabendazole",
  "betamethasone, enrofloxacin",
  "dexamethasone, ear cleaner (epi-otic advanced), enrofloxacin",
  "betamethasone, tobramycin"
)

```

Use steroids list to do counting now

```{r}
# use HSA medications df to get list of NSAIDs in df
table(medications_5y $medication_ingredients_100)
unique(medications_5y $medication_ingredients_100)
unique(medications_5y $medication_ingredients)


medications_5y  <- medications_5y  %>%
  mutate(steroid_use = if_else(
    str_detect(tolower(medication_ingredients), str_c(steroids, collapse = "|")),
    "steroid_yes",
    "steroid_no"
  ))

#a count of these for each dog exc if steroid no then just steroid no
steroid_5y_meds_per_dog <- medications_5y  %>%
  group_by(subject_id, steroid_use) %>%
  summarise(count = n(), .groups = "drop")
  #remove rows where if subject has ANY steroid_yes value then remove the steroid_no row/s for that subject id
  # Step 1: Get subject_ids that have steroid_yes
steroid_yes_subjects <- steroid_5y_meds_per_dog %>%
  filter(steroid_use == "steroid_yes") %>%
  pull(subject_id)

# Step 2: Keep steroid_yes rows or steroid_no only if no steroid_yes exists for that dog
steroid_5y_meds_per_dog_clean <- steroid_5y_meds_per_dog %>%
  filter(steroid_use == "steroid_yes" | !(subject_id %in% steroid_yes_subjects))

#calculate steroid yes quartiles
steroid_yes <- steroid_5y_meds_per_dog_clean %>%
  filter(steroid_use=="steroid_yes")

steroid_summary <- steroid_yes %>%
  summarise(
    min=min(count),
    Q1 = quantile(count, 0.25, na.rm = TRUE),
    median_count_per_dog = quantile(count, 0.50, na.rm = TRUE),  # Use quantile for clarity
    Q3 = quantile(count, 0.75, na.rm = TRUE),
    max=max(count),
    .groups = "drop"
  )

steroid_5y_meds_per_dog_clean_quartiles <- steroid_5y_meds_per_dog_clean %>%
  mutate(steroid_quantile = case_when(
    steroid_use == "steroid_no" ~ "no_steroid_admin",
    steroid_use == "steroid_yes" & count <= 1 ~ "Q1(1)",
    steroid_use == "steroid_yes" & count > 1 & count <= 2 ~ "Q2(1-2)",
    steroid_use == "steroid_yes" & count > 2 & count <= 3 ~ "Q3(2-3)",
    steroid_use == "steroid_yes" & count > 3 ~ "Q4(3-17)"
  ))
#then filt down to unique single row for each dog of its count or steroid no, then join back to the 3044 dogs and if no count or no steroid no then "no_steroid_recorded_in_time_period"
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID %>% left_join(steroid_5y_meds_per_dog_clean_quartiles, by="subject_id")

#now replace NAs in steroid_quantile column with "no_med_records_this_time_period"
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid %>%
    mutate(steroid_quantile = replace_na(steroid_quantile, "no_med_use_recorded"))

#ditch count column 
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid<- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid%>%
  dplyr::select(c(-"count"))
#rename steroid_quantile to make clear its a 5y column
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid %>%
  rename(steroid_5y_quantiles = steroid_quantile)



dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid <-
  dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid %>%
  mutate(across(253:254, ~ case_when(
    is.na(.) ~ "no_medication_records_for_this_time_period",
    TRUE ~ as.character(.)
  )))
```

For MCT want to do:

#### Systemic preds

\- insufficient variation for quartiles so going to just leave counts as is and let ZS decide how to categorise eg. binary or something else

```{r}
systemic_preds<- c("prednisone",
  "predisolone acetate",
  "trimeprazine tartrate, prednisone",
   "prednisolone"
)
```

```{r}
# use HSA medications df to get list of NSAIDs in df
table(medications_5y $medication_ingredients_100)
unique(medications_5y $medication_ingredients_100)
unique(medications_5y $medication_ingredients)


medications_5y  <- medications_5y  %>%
  mutate(systemic_preds_use = if_else(
    str_detect(tolower(medication_ingredients), str_c(systemic_preds, collapse = "|")),
    "systemic_preds_yes",
    "systemic_preds_no"
  ))

#a count of these for each dog exc if systemic_preds no then just systemic_preds no
systemic_preds_5y_meds_per_dog <- medications_5y  %>%
  group_by(subject_id, systemic_preds_use) %>%
  summarise(count = n(), .groups = "drop")
  #remove rows where if subject has ANY systemic_preds_yes value then remove the systemic_preds_no row/s for that subject id
  # Step 1: Get subject_ids that have systemic_preds_yes
systemic_preds_yes_subjects <- systemic_preds_5y_meds_per_dog %>%
  filter(systemic_preds_use == "systemic_preds_yes") %>%
  pull(subject_id)

# Step 2: Keep systemic_preds_yes rows or systemic_preds_no only if no systemic_preds_yes exists for that dog
systemic_preds_5y_meds_per_dog_clean <- systemic_preds_5y_meds_per_dog %>%
  filter(systemic_preds_use == "systemic_preds_yes" | !(subject_id %in% systemic_preds_yes_subjects))

#calculate systemic_preds yes quartiles
systemic_preds_yes <- systemic_preds_5y_meds_per_dog_clean %>%
  filter(systemic_preds_use=="systemic_preds_yes")

systemic_preds_summary <- systemic_preds_yes %>%
  summarise(
    min=min(count),
    Q1 = quantile(count, 0.25, na.rm = TRUE),
    median_count_per_dog = quantile(count, 0.50, na.rm = TRUE),  # Use quantile for clarity
    Q3 = quantile(count, 0.75, na.rm = TRUE),
    max=max(count),
    .groups = "drop"
  )
#NB is not actually quartiles split
#NB also the count column is counting rows for each dog, so the no_systemic_preds_admins will also have a count for each medical row they DO NOT HAVE PREDS for 
systemic_preds_5y_meds_per_dog_clean_quartiles <- systemic_preds_5y_meds_per_dog_clean %>%
  mutate(systemic_preds_amount = case_when(
    systemic_preds_use == "systemic_preds_no" ~ "no_systemic_preds_admin",
    systemic_preds_use == "systemic_preds_yes" ~ as.character(count)
  ))
#then filt down to unique single row for each dog of its count or systemic_preds no, then join back to the 3044 dogs and if no count or no systemic_preds no then "no_systemic_preds_recorded_in_time_period"
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid %>% left_join(systemic_preds_5y_meds_per_dog_clean_quartiles, by="subject_id")

#now replace NAs in steroid_quantile column with "no_med_records_this_time_period"
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred %>%
    mutate(systemic_pred_amount = replace_na(systemic_preds_amount, "no_med_use_recorded"))

#ditch count column 
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred<- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred%>%
  dplyr::select(c(-"count"))
#rename steroid_quantile to make clear its a 5y column
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred %>%
  rename(systemic_preds_5y_amount = systemic_preds_amount)

###redo indices asnew column

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred <-
  dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred %>%
  mutate(across(256:258, ~ case_when(
    is.na(.) ~ "no_medication_records_for_this_time_period",
    TRUE ~ as.character(.)
  )))
```

Lokivetmab

```{r}
#
lokivetmab <-c("lokivetmab","cytopoint")
```

```{r}
medications_5y  <- medications_5y  %>%
  mutate(lokivetmab_use = if_else(
    str_detect(tolower(medication_ingredients), str_c(lokivetmab, collapse = "|")),
    "lokivetmab_yes",
    "lokivetmab_no"
  ))

#a count of these for each dog exc if lokivetmab no then just lokivetmab no
lokivetmab_5y_meds_per_dog <- medications_5y  %>%
  group_by(subject_id, lokivetmab_use) %>%
  summarise(count = n(), .groups = "drop")
  #remove rows where if subject has ANY lokivetmab_yes value then remove the lokivetmab_no row/s for that subject id
  # Step 1: Get subject_ids that have lokivetmab_yes
lokivetmab_yes_subjects <- lokivetmab_5y_meds_per_dog %>%
  filter(lokivetmab_use == "lokivetmab_yes") %>%
  pull(subject_id)

# Step 2: Keep lokivetmab_yes rows or lokivetmab_no only if no lokivetmab_yes exists for that dog
lokivetmab_5y_meds_per_dog_clean <- lokivetmab_5y_meds_per_dog %>%
  filter(lokivetmab_use == "lokivetmab_yes" | !(subject_id %in% lokivetmab_yes_subjects))

#calculate lokivetmab yes quartiles
lokivetmab_yes <- lokivetmab_5y_meds_per_dog_clean %>%
  filter(lokivetmab_use=="lokivetmab_yes")

lokivetmab_summary <- lokivetmab_yes %>%
  summarise(
    min=min(count),
    Q1 = quantile(count, 0.25, na.rm = TRUE),
    median_count_per_dog = quantile(count, 0.50, na.rm = TRUE),  # Use quantile for clarity
    Q3 = quantile(count, 0.75, na.rm = TRUE),
    max=max(count),
    .groups = "drop"
  )
#NB is not actually quartiles split
#NB also the count column is counting rows for each dog, so the no_lokivetmab_admins will also have a count for each medical row they DO NOT HAVE lokivetmabs for 
lokivetmab_5y_meds_per_dog_clean_quartiles <- lokivetmab_5y_meds_per_dog_clean %>%
  mutate(lokivetmab_amount = case_when(
    lokivetmab_use == "lokivetmab_no" ~ "no_lokivetmab_admin",
    lokivetmab_use == "lokivetmab_yes" ~ as.character(count)
  ))
#then filt down to unique single row for each dog of its count or lokivetmab no, then join back to the 3044 dogs and if no count or no lokivetmab no then "no_lokivetmab_recorded_in_time_period"
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred%>% left_join(lokivetmab_5y_meds_per_dog_clean_quartiles, by="subject_id")

#now replace NAs in steroid_quantile column with "no_med_records_this_time_period"
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab %>%
    mutate(lokivetmab_amount = replace_na(lokivetmab_amount, "no_med_use_recorded"))

#ditch count column 
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab<- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab%>%
  dplyr::select(c(-"count"))
#rename steroid_quantile to make clear its a 5y column
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab %>%
  rename(lokivetmab_5y_amount = lokivetmab_amount)

###redo indices asnew column

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab <-
  dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab %>%
  mutate(across(259:260, ~ case_when(
    is.na(.) ~ "no_medication_records_for_this_time_period",
    TRUE ~ as.character(.)
  )))
```

oclacitinib

```{r}
oclacitinib <-c("oclacitinib","apoquel")
```

```{r}
medications_5y  <- medications_5y  %>%
  mutate(oclacitinib_use = if_else(
    str_detect(tolower(medication_ingredients), str_c(oclacitinib, collapse = "|")),
    "oclacitinib_yes",
    "oclacitinib_no"
  ))

#a count of these for each dog exc if oclacitinib no then just oclacitinib no
oclacitinib_5y_meds_per_dog <- medications_5y  %>%
  group_by(subject_id, oclacitinib_use) %>%
  summarise(count = n(), .groups = "drop")
  #remove rows where if subject has ANY oclacitinib_yes value then remove the oclacitinib_no row/s for that subject id
  # Step 1: Get subject_ids that have oclacitinib_yes
oclacitinib_yes_subjects <- oclacitinib_5y_meds_per_dog %>%
  filter(oclacitinib_use == "oclacitinib_yes") %>%
  pull(subject_id)

# Step 2: Keep oclacitinib_yes rows or oclacitinib_no only if no oclacitinib_yes exists for that dog
oclacitinib_5y_meds_per_dog_clean <- oclacitinib_5y_meds_per_dog %>%
  filter(oclacitinib_use == "oclacitinib_yes" | !(subject_id %in% oclacitinib_yes_subjects))

#calculate oclacitinib yes quartiles
oclacitinib_yes <- oclacitinib_5y_meds_per_dog_clean %>%
  filter(oclacitinib_use=="oclacitinib_yes")

oclacitinib_summary <- oclacitinib_yes %>%
  summarise(
    min=min(count),
    Q1 = quantile(count, 0.25, na.rm = TRUE),
    median_count_per_dog = quantile(count, 0.50, na.rm = TRUE),  # Use quantile for clarity
    Q3 = quantile(count, 0.75, na.rm = TRUE),
    max=max(count),
    .groups = "drop"
  )
#NB is not actually quartiles split
#NB also the count column is counting rows for each dog, so the no_oclacitinib_admins will also have a count for each medical row they DO NOT HAVE oclacitinibs for 
oclacitinib_5y_meds_per_dog_clean_quartiles <- oclacitinib_5y_meds_per_dog_clean %>%
  mutate(oclacitinib_amount = case_when(
    oclacitinib_use == "oclacitinib_no" ~ "no_oclacitinib_admin",
    oclacitinib_use == "oclacitinib_yes" ~ as.character(count)
  ))
#then filt down to unique single row for each dog of its count or oclacitinib no, then join back to the 3044 dogs and if no count or no oclacitinib no then "no_oclacitinib_recorded_in_time_period"
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab%>% left_join(oclacitinib_5y_meds_per_dog_clean_quartiles, by="subject_id")

#now replace NAs in steroid_quantile column with "no_med_records_this_time_period"
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib %>%
    mutate(oclacitinib_amount = replace_na(oclacitinib_amount, "no_med_use_recorded"))

#ditch count column 
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib<- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib%>%
  dplyr::select(c(-"count"))
#rename steroid_quantile to make clear its a 5y column
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib %>%
  rename(oclacitinib_5y_amount = oclacitinib_amount)

###redo indices asnew column

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib <-
  dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib %>%
  mutate(across(260:262, ~ case_when(
    is.na(.) ~ "no_medication_records_for_this_time_period",
    TRUE ~ as.character(.)
  )))
```

Interim write to csv

```{r}
write.csv(dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib,here::here("Output/GRLS_MCT_variables_dataset_interim.csv"))
```

## Reproduction variables

Load in repro datasets

```{r}
# female information
repro_hx_female <- read.csv(here::here("Data/reproductive_history_female.csv"))
# not using df as hasnt been updated in past year or so of GRLS datasets so dont think being maintained spay_details <- read.csv(here::here("Data/details_spay_before_heat.csv"))
#join spay details to repro hx female sheet

repro_heats_female <- read.csv(here::here("Data/reproductive_history_heat.csv"))
#male 
repro_hx_male <- read.csv(here::here("Data/reproductive_history_male.csv"))

repro_hx_hormones <-read.csv(here::here("Data/reproductive_history_hormones.csv"))
#all empty cells encoded as NA
repro_hx_female <- repro_hx_female %>%
  mutate_if(~any(. == ""), na_if, "")

repro_hx_male <- repro_hx_male %>%
  mutate_if(~any(. == ""), na_if, "")

repro_heats_female <- repro_heats_female %>%
  mutate_if(~any(. == ""), na_if, "")

repro_hx_hormones <- repro_hx_hormones %>%
  mutate_if(~any(. == ""), na_if, "")

#what are the columns available in df
colnames(repro_hx_female)
colnames(repro_hx_male)
colnames(repro_heats_female)
colnames(repro_hx_hormones)

#dog profile also contains neuter date and neuter status
profile <- read.csv(here::here("Data/dog_profile.csv"))
```

First adding basic neuter info to the dataset = neuter status and age at neuter

```{r}
profile <- profile %>%
  dplyr::select(c(subject_id,sex_status,spay_neuter_date,birth_date))


profile <- profile %>% 
  mutate(sex=recode(
    sex_status,"Spayed Female"="F", "Intact Female" = "F", "Neutered Male" = "M",
    "Intact Male" = "M")
  )
#neuter status
profile <- profile%>%
  mutate(neuter=recode(
    sex_status, "Spayed Female"="Neutered", "Neutered Male"="Neutered","Intact Female"= "Entire", "Intact Male" = "Entire")
  )
#calculate age at neuter
profile$birth_date2 <- as.Date(profile$birth_date)
profile$spay_neuter_date2 <- as.Date(profile$spay_neuter_date)


profile <- profile %>%
mutate(
    age_at_neuter_years = case_when(
      is.na(spay_neuter_date2) ~ "not_neutered",
      #age at neuter if negative value replace with age = 0

      time_length(interval(birth_date2, spay_neuter_date2), "years") < 0 ~ "0",
      TRUE ~ as.character(round(time_length(interval(birth_date2, spay_neuter_date2), "years"), 1))
    )
  )
profile_abr <- profile %>%
  dplyr::select(subject_id,neuter, sex, sex_status,birth_date2, age_at_neuter_years,spay_neuter_date2)



## add back onto df
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib_neutering <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib %>% left_join(profile_abr, by="subject_id")
```

Info on number of seasons etc

```{r}
repro_status <-bind_rows(repro_hx_female,repro_hx_male)

#add dog profile
demog_neut <- repro_status %>%
  left_join(profile_abr, by="subject_id")
demog_neut <- demog_neut%>% 
  group_by(subject_id)%>%
  fill(c(spay_neuter_date2,neutered_reason),.direction="downup")


demog_neut <- demog_neut %>% 
  mutate_if(is.character, ~ ifelse(is.na(.), "not_recorded", .))

demog_neut$record_date2 <- as.Date(demog_neut$record_date)
#all dogs have neuter records
unique_neut <- unique(demog_neut$subject_id)

```

### Age at first heat variable

```{r}
source(here::here("Code/GRLS_functions.R"))
neut_age_seasons2<- first_exposure_point(demog_neut,heat_last_year,record_date2)
neut_age_seasons3  <- first_exposure_age2(neut_age_seasons2,"heat_last_year_first_exposure_date", "birth_date2")

neut_age_seasons3 <- neut_age_seasons3 %>%
  rename(age_first_heat_years=heat_last_year_first_exposure_date_age_years)
  

#recode so that if M in sex column recode the NAs for these as = "male_dog" in age_first_heat_years
#recode so that remaining NAs = "not_recorded"

neut_age_seasons3 <- neut_age_seasons3 %>%
  mutate(
    age_first_heat_years = case_when(
      is.na(age_first_heat_years) & sex == "M" ~ "male_dog",
      is.na(age_first_heat_years) ~ "not_recorded",
      TRUE ~ as.character(age_first_heat_years)
    ),
    heat_last_year_first_exposure_date_age = case_when(
      is.na(age_first_heat_years) & sex == "M" ~ "male_dog",
      is.na(heat_last_year_first_exposure_date_age) ~ "not_recorded",
      TRUE ~ as.character(heat_last_year_first_exposure_date_age)
    )
  )
```

### Num heats before neutering

```{r}
neut_age_seasons4 <- num_exposures(neut_age_seasons3,"heat_last_year","record_date2","spay_neuter_date2")

#replace 0s coded in the "heat_last_year_num_exposures" with "male_dogs" and "not_recorded" dogs if dogs are M or have no heat records
neut_age_seasons4 <- neut_age_seasons4 %>%
  mutate(
    heat_last_year_num_exposures = case_when(
      heat_last_year_num_exposures == 0 & sex == "M" ~ "male_dog",
      TRUE ~ as.character(heat_last_year_num_exposures)
    )
    )

```

### Ever pregnant variable

```{r}
#pregnancy before neuter Y N
neut_age_seasons5<- neut_age_seasons4 %>%
  group_by(subject_id) %>%
  mutate(preg_pre_neuter=ifelse(any(!is.na(age_at_neuter_years) & no_pregnancy_last_year==1,na.rm=TRUE),1,0))    

neut_age_seasons5 <- neut_age_seasons5 %>%
  mutate(
    no_pregnancy_last_year = case_when(
      no_pregnancy_last_year == 0 & sex == "M" ~ "male_dog",
      TRUE ~ as.character(no_pregnancy_last_year)
    )
    )
```

### Males ever bred from variable

```{r}

neut_age_M <- neut_age_seasons5 %>%
  mutate(
    mated_pre_neuter = if_else(sex == "M" & neuter == "Neutered" & 
                                  any(!is.na(age_at_neuter_years) & intact_naturally_bred_last_year == 1),
                                1,
                                0)
  )

```

### Tidying up of repro data so all values are valid for sex/neuter status e.g males dont have number of pregnancy 0 but just number of pregnancy - not applciable or similar

```{r}
neut_data_single_row <- neut_age_M %>%
  distinct(subject_id,.keep_all = TRUE)


#check male variables not include females - there are none of these

#check neut variables not including entire dogs
neut_data_single_row <- neut_data_single_row %>%
  mutate(
    preg_pre_neuter = case_when(
      preg_pre_neuter == 0 & neuter == "Entire" ~ "Entire",
      TRUE ~ as.character(preg_pre_neuter)
    )
    )

neut_data_single_row <- neut_data_single_row %>%
  mutate(
    mated_pre_neuter = case_when(
      mated_pre_neuter == 0 & neuter == "Entire" ~ "Entire",
      TRUE ~ as.character(mated_pre_neuter)
    )
    )
#keep relevant cols
neut_data_single_row <- neut_data_single_row %>%
  dplyr::select(1,37:42,47:50)


```

### Time between neuter date and cancer diagnosis/endpoint date variable

perform this when only single row per dog rather than all the repro rows as in the prev dfs

```{r}
#add neut data to the main df then calculate 
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib2 <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib %>% left_join(neut_data_single_row,by="subject_id")
#delete dupe columns
#dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib2 <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib2 %>%
#  dplyr::select(-c(neuter.x,sex.x,birth_date2.x,age_at_neuter_years.x,spay_neuter_date2.x))
#remove the .y suffix from the columns now
#dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib2 <-
 # dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib2 %>%
 # rename(sex= sex.y,
  #       neuter = neuter.y,
   #      birth_date2 = birth_date2.y,
     #    age_at_neuter_years = age_at_neuter_years.y,
     #    spay_neuter_date2 = spay_neuter_date2.y)
#calculate time between spay_neuter_date2 and end date
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3 <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib2 %>%
mutate(
    time_between_neuter_endpoint_diagnosis_years = case_when(
      is.na(spay_neuter_date2) ~ "not_neutered",
      #age at neuter if negative value replace with age = 0

      time_length(interval(spay_neuter_date2,end_date), "years") < 0 ~ "0",
      TRUE ~ as.character(round(time_length(interval(spay_neuter_date2, end_date), "years"), 1))
    )
  )

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3 <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3 %>%
mutate(
    time_between_neuter_endpoint_diagnosis_months = case_when(
      is.na(spay_neuter_date2) ~ "not_neutered",
      #age at neuter if negative value replace with age = 0

      time_length(interval(spay_neuter_date2,end_date), "months") < 0 ~ "0",
      TRUE ~ as.character(round(time_length(interval(spay_neuter_date2, end_date), "months"), 1))
    )
  )
```

### Age at endpoint diagnosis

```{r}
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3 <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3 %>%
  mutate(
    birth_date2 = as.Date(birth_date2, format = "%Y-%m-%d"),
    end_date2= as.Date(end_date, format = "%Y-%m-%d")
  ) %>%
  mutate(age_at_final_date = as.numeric(interval(birth_date2, end_date2) / years(1)))


```

## Weight and height data 5y prior to endpoint

```{r}
###add in weight and BCS data for the dogs
dog_weight <- read.csv(here::here("Data/exam_physical.csv"))
dog_weight <- dog_weight %>%
  dplyr::select(c(subject_id,year_in_study,height,weight,purina_body_condition_score))
#remerge back to big df 
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight <-  dog_weight%>%
  left_join(dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3,by="subject_id")
###for record years 5y prior to endpoint record year do an average of weight
get_mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))] # Returns the most frequent value
}

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight2 <- 
  dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight %>%
  mutate(year_in_study_diagnosis_or_final_record_year = as.numeric(year_in_study_diagnosis_or_final_record_year)) %>%
  group_by(subject_id) %>%
  filter(year_in_study >= (year_in_study_diagnosis_or_final_record_year - 5)) %>%
  mutate(
    count_rows = n(),
    avg_weight_5y_prior_endpoint = if (count_rows[1] >= 5) mean(weight, na.rm = TRUE) else NA_real_,
    avg_height_5y_prior_endpoint = if (count_rows[1] >= 5) mean(height, na.rm = TRUE) else NA_real_,
    median_weight_5y_prior_endpoint = if (count_rows[1] >= 5) median(weight, na.rm = TRUE) else NA_real_,   
    median_weight_5y_prior_endpoint_any_records = median(weight, na.rm = TRUE),
    median_height_5y_prior_endpoint = if (count_rows[1] >= 5) median(height, na.rm = TRUE) else NA_real_,
    median_height_5y_prior_endpoint_any_records = median(height, na.rm = TRUE),
    avg_purina_BCS_5y_prior_endpoint = if (count_rows[1] >= 5) mean(purina_body_condition_score, na.rm = TRUE) else NA_real_,
    mode_purina_BCS_5y_prior_endpoint = get_mode(purina_body_condition_score)
  ) %>%
  ungroup()


```

tidy these to remove outliers and convert to kg and cm (but not going to do quartiles/biological splits etc. - leave for student to decide how to split)

```{r}
#avg weight to kg
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight3<- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight2 %>%
  mutate(avg_weight_5y_prior_endpoint_kg = avg_weight_5y_prior_endpoint * 0.453)

#### remove weight values for dogs that have their age at final date as <18m (ie. non adult weights) = 6 dogs
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight3 <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight3 %>%
  mutate(avg_weight_5y_prior_endpoint_kg = if_else(age_at_final_date < 1.5, NA_real_, avg_weight_5y_prior_endpoint_kg))

#median weight
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight3<- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight3 %>%
  mutate(median_weight_5y_prior_endpoint_kg = median_weight_5y_prior_endpoint * 0.453)

#### remove weight values for dogs that have their age at final date as <18m (ie. non adult weights) = 6 dogs
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight3 <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight3 %>%
  mutate(median_weight_5y_prior_endpoint_kg = if_else(age_at_final_date < 1.5, NA_real_, median_weight_5y_prior_endpoint_kg))

# Convert both height summaries without overwriting the first derived column.
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4 <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight3 %>%
  mutate(
    avg_height_5y_prior_endpoint_cm = avg_height_5y_prior_endpoint * 2.54,
    median_height_5y_prior_endpoint_cm = median_height_5y_prior_endpoint * 2.54,
    avg_height_5y_prior_endpoint_cm = if_else(
      age_at_final_date < 1.5,
      NA_real_,
      avg_height_5y_prior_endpoint_cm
    ),
    median_height_5y_prior_endpoint_cm = if_else(
      age_at_final_date < 1.5,
      NA_real_,
      median_height_5y_prior_endpoint_cm
    )
  )

weight_feature_columns <- c(
  "avg_weight_5y_prior_endpoint",
  "avg_height_5y_prior_endpoint",
  "median_weight_5y_prior_endpoint",
  "median_weight_5y_prior_endpoint_any_records",
  "median_height_5y_prior_endpoint",
  "median_height_5y_prior_endpoint_any_records",
  "avg_purina_BCS_5y_prior_endpoint",
  "mode_purina_BCS_5y_prior_endpoint",
  "avg_weight_5y_prior_endpoint_kg",
  "median_weight_5y_prior_endpoint_kg",
  "avg_height_5y_prior_endpoint_cm",
  "median_height_5y_prior_endpoint_cm"
)

weight_features <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4 %>%
  dplyr::select(subject_id, all_of(weight_feature_columns)) %>%
  distinct()

if (anyDuplicated(weight_features$subject_id)) {
  stop("Weight feature values are inconsistent within at least one subject_id.")
}
if (anyDuplicated(dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3$subject_id)) {
  stop("The pre-weight MCT dataset is not one row per subject_id.")
}

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_unique <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3 %>%
  left_join(weight_features, by = "subject_id", relationship = "one-to-one")
```

```{r}
write.csv(dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_unique,here::here("Output/GRLS_MCT_variables_dataset_interim.csv"))
```

## Activity

taken from this script-GRLS activity classifcation method_modified.qmd

Loading in datasets:

```{r}
#activity data from study y 1 and 2
activity_sy1_2 <- read.csv(here::here("Data/activity_details_through_SY2.csv"))
#check # dogs in it
act_1_2 <- unique(activity_sy1_2$subject_id)
#activity data from study y 3 + 
activity_sy3_beyond <- read.csv(here::here("Data/activity_details_SY3_beyond.csv"))
act_3_plus<- unique(activity_sy3_beyond$subject_id)
#activity overview dataset
activity_overview<- read.csv(here::here("Data/activity_overview.csv"))
act_over<-unique(activity_overview$subject_id)
```

### Early years dataset variable creation

```{r}

#create a new column of total activity to include both lead walk and off lead activity
activity_sy1_2 <- activity_sy1_2 %>%
  mutate(total_activity_duration = paste(walk_duration,aerobic_duration,  sep = "_")) %>%
  mutate(total_activity_freq= paste(walk_frequency,aerobic_frequency,sep="_")) %>%
  mutate(total_activity_intensity =paste(walk_pace,aerobic_pace,sep="_"))

#check all dogs have sy1_2 recorded - they do
unique_actsy1_2 <- unique(activity_sy1_2$subject_id)
#duration recoding
  
less_than_30min <- c("_10-30 minutes", "10-30 minutes_","Less than 10 minutes_","Less than 10 minutes_Less than 10 minutes","_Less than 10 minutes","10-30 minutes_Less than 10 minutes","_")
#
between_31_60min <-c("_30-60 minutes","10-30 minutes_30-60 minutes","30-60 minutes_","Less than 10 minutes_10-30 minutes","10-30 minutes_10-30 minutes","30-60 minutes_Less than 10 minutes","Less than 10 minutes_30-60 minutes","30-60 minutes_10-30 minutes")
#
between_61_90min<- c("Greater than 60 minutes_","_Greater than 60 minute","10-30 minutes_Greater than 60 minutes","Greater than 60 minutes_10-30 minutes","_Greater than 60 minutes","Greater than 60 minutes_Less than 10 minutes","Greater than 60 minutes_Less than 10 minutes","Less than 10 minutes_Greater than 60 minutes")
#
between_91_120min <- c("Greater than 60 minutes_30-60 minutes","30-60 minutes_Greater than 60 minutes","30-60 minutes_30-60 minutes")
#
more_than_120min<- c("Greater than 60 minutes_Greater than 60 minutes")

all_groups <- c(less_than_30min,between_31_60min,between_61_90min,
                between_91_120min,more_than_120min)

# Find any cateogories that are not in any group
duration <- setdiff(activity_sy1_2$total_activity_duration, all_groups)

# Display unclassified clinical signs
print(duration)


#recode in df
activity_sy1_2 <- activity_sy1_2  %>%
  mutate(
    total_activity_duration_regroup = case_when(
      total_activity_duration %in% less_than_30min ~ "less_than_30min",
      total_activity_duration %in% between_31_60min ~ "between_31_60min",
      total_activity_duration %in% between_61_90min ~ "between_61_90min",
      total_activity_duration %in% between_91_120min ~ "between_91_120min",
      total_activity_duration %in% more_than_120min ~ "more_than_120min",
      TRUE ~ "no_duration_recorded" # For values not in any group
    )
  )



#frequency recoding
never <- c("Never_","Never_Less than once/month") 
less_than_weekly <- c("Less than once/month_","Less than once/month_Less than once/month", "Less than once/month_Less than once/week","Less than once/week_Less than once/month","Never_Less than once/week","Less than once/week_",
                      "Less than once/week_Less than once/week")
min_weekly<-c("Once or twice/week_Once or twice/week","Once or twice/week_","Never_Once or twice/week","Once or twice/week_Less than once/month",
              "Once or twice/week_Less than once/week","Less than once/week_Once or twice/week","Less than once/month_Once or twice/week")
      
daily<-c("Once/day_","Less than once/week_Once/day","Never_Once/day", "Once or twice/week_Once/day","Once/day_Once or twice/week","Once/day_Less than once/week",
         "Once/day_Less than once/month","Once/day_Once or twice/week",
         "Once/day_Less than once/week", "Once/day_Less than once/month",
         "Once or twice/week_Once/day","Less than once/month_Once/day"
         )
multi_daily <-c("Once/day_Once/day","Once/day_More than once a day","More than once a day_","Once or twice/week_More than once a day","Less than once/month_More than once a day","More than once a day_Once/day","Never_More than once a day","More than once a day_Less than once/month","More than once a day_Once or twice/week","Less than once/week_More than once a day","More than once a day_More than once a day",
                "More than once a day_Less than once/week")

all_groups_freq <- c(never,less_than_weekly,min_weekly,daily,multi_daily)
freq <- setdiff(activity_sy1_2$total_activity_freq, all_groups_freq)

# Display unclassified freq
print(freq)

#recode in df
activity_sy1_2 <- activity_sy1_2  %>%
  mutate(
    total_activity_freq_regroup = case_when(
      total_activity_freq %in% never ~ "never",
      total_activity_freq %in% less_than_weekly ~ "less_than_weekly",
      total_activity_freq %in% min_weekly ~ "min_weekly",
      total_activity_freq %in% daily ~ "daily",
      total_activity_freq %in% multi_daily ~ "multi_daily",
      TRUE ~ "no_freq_recorded" # For values not in any group
    )
  )
#create recoded freq column

#intensity recoding
low_only <-c("_Bursts of exercise less than 30% of the time","Slow_","Slow_Bursts of exercise less than 30% of the time"
             )
moderate <-c("Average_","_Engages in moderate exercise (greater than 30% of the time but less than 60% of the time)","Brisk_","Brisk_Bursts of exercise less than 30% of the time","Average_Engages in moderate exercise (greater than 30% of the time but less than 60% of the time)","Jog_","Average_Bursts of exercise less than 30% of the time","Jog_Bursts of exercise less than 30% of the time","Brisk_Engages in moderate exercise (greater than 30% of the time but less than 60% of the time)","Slow_Engages in moderate exercise (greater than 30% of the time but less than 60% of the time)")
high <-c(" _Engages in strenuous exercise most of the time (greater than 60%)","Run_","Average_Engages in strenuous exercise most of the time (greater than 60%)","Brisk_Engages in strenuous exercise most of the time (greater than 60%)","_Engages in strenuous exercise most of the time (greater than 60%)","Jog_Engages in strenuous exercise most of the time (greater than 60%)","Run_Engages in strenuous exercise most of the time (greater than 60%)","Run_Engages in moderate exercise (greater than 30% of the time but less than 60% of the time)", "Slow_Engages in strenuous exercise most of the time (greater than 60%)",
         "Jog_Engages in moderate exercise (greater than 30% of the time but less than 60% of the time)","Run_Bursts of exercise less than 30% of the time")

all_groups_int <- c(low_only,moderate,high)
int <- setdiff(activity_sy1_2$total_activity_intensity, all_groups_int)
print(int)


#create recoded intensity column
activity_sy1_2 <- activity_sy1_2  %>%
  mutate(
    total_activity_intensity_regroup = case_when(
      total_activity_intensity %in% low_only ~ "low_int",
      total_activity_intensity %in% moderate ~ "mod_int",
      total_activity_intensity %in% high ~ "high_int",
      TRUE ~ "no_int_recorded" # For values not in any group
    )
  )

activity_sy1_2 <- activity_sy1_2  %>%
  mutate(
    total_activity_duration_regroup_num = case_when(
      total_activity_duration %in% less_than_30min ~ "20",
      total_activity_duration %in% between_31_60min ~ "40",
      total_activity_duration %in% between_61_90min ~ "60",
      total_activity_duration %in% between_91_120min ~ "80",
      total_activity_duration %in% more_than_120min ~ "100",
      TRUE ~ "0" # For values not in any group
    )
  )

activity_sy1_2 <- activity_sy1_2  %>%
  mutate(
    total_activity_freq_regroup_num = case_when(
      total_activity_freq %in% never ~ "20",
      total_activity_freq %in% less_than_weekly ~ "40",
      total_activity_freq %in% min_weekly ~ "60",
      total_activity_freq %in% daily ~ "80",
      total_activity_freq %in% multi_daily ~ "100",
      TRUE ~ "0" # For values not in any group
    )
  )




#create recoded intensity column
activity_sy1_2 <- activity_sy1_2  %>%
  mutate(
    total_activity_intensity_regroup_num = case_when(
      total_activity_intensity %in% low_only ~ "33.3",
      total_activity_intensity %in% moderate ~ "66.7",
      total_activity_intensity %in% high ~ "100",
      TRUE ~ "0" # For values not in any group
    )
  )



#create column for USDA activity <30min or >30 min
activity_sy1_2 <- activity_sy1_2 %>%
  mutate(USDA_30min = case_when (
    total_activity_duration_regroup %in% c("less_than_30min") ~ "below_USDA",  total_activity_duration_regroup %in% c("between_31_60min", "between_61_90min","between_91_120min","more_than_120min") ~ "more_than_USDA", 
                                 TRUE ~ NA ))

#create column for UKKC recommendaiton <2h but cannot do this one so >1h
activity_sy1_2 <- activity_sy1_2 %>%
  mutate(activity_1h_min = case_when (
total_activity_duration_regroup %in% c("less_than_30min","between_31_60min") ~ "below_1h",  total_activity_duration_regroup %in% c("between_61_90min","between_91_120min","more_than_120min") ~ "more_than_1h", 
                                 TRUE ~ NA ))



activity_sy1_2 <- activity_sy1_2 %>%
  mutate(activity_KC_2h_min = case_when (
total_activity_duration_regroup %in% c("less_than_30min","between_31_60min","between_61_90min","between_91_120min") ~ "below_KC",  total_activity_duration_regroup %in% c("more_than_120min") ~ "more_than_KC", 
                                 TRUE ~ NA ))

#convert to numeric for score systems
activity_sy1_2 <- activity_sy1_2 %>%
  mutate(across(24:26, as.numeric))

totals_avg <- activity_sy1_2 %>%
  group_by(subject_id) %>%
  summarise(
    avg_activity_duration = (mean(sum(total_activity_duration_regroup_num, na.rm = TRUE)) /
                              (n() * 100)) * 100,
    avg_activity_intensity = (mean(sum(total_activity_intensity_regroup_num, na.rm = TRUE)) /
                              (n() * 100)) * 100,
    avg_activity_freq = (mean(sum(total_activity_freq_regroup_num, na.rm = TRUE)) /
                              (n() * 100)) * 100

  )

# Join the average values back to the original dataset
activity_sy1_2 <- activity_sy1_2 %>%
  left_join(totals_avg, by = "subject_id")



```

### Create rest of life variables (SY3+)

Recode pace and frequency (there is no duration in SY3+)

```{r}

#check unique IDs in SY3+ to check all dogs have records - they do not, approx 400 dogs missing records so need to account for this later on when adding back to 3044 dog df
unique_sy3 <- unique(activity_sy3_beyond$subject_id)
activity_sy3_beyond <- activity_sy3_beyond %>%
  mutate(pace_num = case_when(
    pace == "Slow walk" ~ 20,
    pace == "Average walk" ~ 40,
    pace == "Brisk walk" ~ 60,
    pace == "Jog" ~ 80,
    pace == "Run" ~ 100,
    pace == "" ~ 0  # empty cells replace with 0 for calculation purposes
  ))


activity_sy3_beyond <- activity_sy3_beyond %>%
  mutate(frequency_num = case_when(
    frequency == "Rarely" ~ 20,
    frequency == "Monthly" ~ 40,
    frequency == "Weekly" ~ 60,
    frequency == "Daily" ~ 80,
    frequency == "More than daily" ~ 100,
    frequency == "" ~ 0  # empty cells replace with 0 for calculation purposes
  ))

totals_y3 <- activity_sy3_beyond %>%
  group_by(subject_id) %>%
  summarise(
   # avg_activity_duration = (mean(duration_num, na.rm = TRUE)/
    #                          (n() * 100)) * 100,
    avg_activity_intensity = (mean(pace_num, na.rm = TRUE)/
                              (n() * 100)) * 100,
    avg_activity_freq=(mean(frequency_num,na.rm=TRUE)/
                              (n() * 100)) * 100# Replace with your column
  )




# Join the totals back to the original dataset
activity_sy3_beyond <- activity_sy3_beyond %>%
  left_join(totals_y3, by = "subject_id")
```

### Overview dataset for whole study period

```{r}
activity_overview <- activity_overview %>%
  mutate(activity_level_num = case_when(
    activity_level == "None" ~ 25,
    activity_level == "Little" ~ 50,
    activity_level == "Moderate" ~ 75,
    activity_level == "Very active" ~ 100,
    activity_level == "" ~ 0)
    )



totals <- activity_overview %>%
  group_by(subject_id) %>%
  summarise(mean_activity_level=(mean(activity_level_num, na.rm=TRUE)/(n() * 100)) * 100)
                              

# Join the totals back to the original dataset
activity_overview <- activity_overview %>%
  left_join(totals, by = "subject_id")
```

Combining all 3 datasets together for whole life of dog:

```{r}
activity_sy1_2_abr_3 <- activity_sy1_2 %>%
  dplyr::select(c(subject_id,year_in_study, USDA_30min,activity_1h_min,activity_KC_2h_min,avg_activity_duration,avg_activity_intensity,avg_activity_freq))


activity_sy3_beyond_abr_3 <- activity_sy3_beyond %>%
  dplyr::select(c(subject_id,year_in_study,
                  #frequency,pace,duration,frequency_num,pace_num,duration_num,<- not keeping as multi rows same year
                  avg_activity_freq,avg_activity_intensity))

activity_overview_abr_3 <- activity_overview %>%
  dplyr::select(c(subject_id,year_in_study,mean_activity_level))


activity_sy1_2_abr_3 <- activity_sy1_2_abr_3 %>%
  distinct(subject_id, .keep_all = TRUE)  %>%
 # mutate(study_year = "y1_2")%>%
  rename_with(~ paste0(., "_sy1_2"), .cols = 3:8)%>%
  dplyr::select(-c(year_in_study))
#sy 3+ 
activity_sy3_beyond_abr_3 <- activity_sy3_beyond_abr_3 %>%
  distinct(subject_id,  .keep_all = TRUE)  %>%
  #  mutate(study_year = "y3_on")%>%
  rename_with(~ paste0(., "_sy3"), .cols = 3:4)%>%
  dplyr::select(-c(year_in_study))
  
#overview
activity_overview_abr_3 <- activity_overview_abr_3 %>%
  distinct(subject_id, .keep_all = TRUE)  %>%
 # mutate(study_year="all_years") %>%
  rename_with(~ paste0(., "_overview"), .cols = 3:3)%>%
  dplyr::select(-c(year_in_study))


activity_all_3 <- activity_overview_abr_3 %>%
  left_join(activity_sy1_2_abr_3, by= "subject_id")
activity_all_3_3 <- activity_all_3 %>%
  left_join(activity_sy3_beyond_abr_3,by="subject_id")

#NAs recode as "no_records_available" for the SY3
activity_all_3_3 <- activity_all_3_3 %>%
  mutate(across(9:10, ~case_when(
      is.na(.) ~ "no_records_for_this_time_period",
    TRUE ~ as.character(.)  
  )))

```

Add in all these variables to main df

```{r}
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_unique <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_unique %>%
  dplyr::select(-c(year_in_study.x,year_in_study.y,record_date.y))
  
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_unique %>%
 left_join(activity_all_3_3, by="subject_id")
```

### 5y prior to endpoint calculations

Add the 3044 dogs to the multi rows of activity dataset:

```{r}
#keep SY0-2 and SY3 beyond same cols only
activity_sy1_2_same <- activity_sy1_2 %>%
  dplyr::select(c(subject_id,year_in_study,total_activity_intensity_regroup_num,total_activity_freq_regroup_num))
#rename cols to match SY3
activity_sy1_2_same <- activity_sy1_2_same %>%
  rename(pace_num = total_activity_intensity_regroup_num, frequency_num = total_activity_freq_regroup_num)
activity_sy3_same<- activity_sy3_beyond %>%
    dplyr::select(c(subject_id,year_in_study,pace_num,frequency_num))

activity_all_y_all_rows <- rbind(activity_sy1_2_same,activity_sy3_same)

#join to cohort for 5y backwards calculations
#abridge cohort to only keep relevant columns then join back to ~200 col df after
cohort_abr <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity %>%
  dplyr::select(c(subject_id,end_date2,had_MCT,year_in_study_diagnosis_or_final_record_year))
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_all_rows <- activity_all_y_all_rows %>%
  left_join(cohort_abr,by="subject_id")


#many to many match from step above resolved by only keeping unique row for each patient
#only singhle unique row per patient]
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_all_rows_2<- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_all_rows[!duplicated(dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_all_rows), ]

#make sure year is in numeric
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_all_rows_2$year_in_study_diagnosis_or_final_record_year  <- as.numeric(dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_all_rows_2$year_in_study_diagnosis_or_final_record_year )

cohort_5y_prior <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_all_rows_2 %>%
  group_by(subject_id) %>%
  filter(year_in_study >= (year_in_study_diagnosis_or_final_record_year - 5)) %>%
  summarise(
    avg_intensity_5y_prior_endpoint = mean(pace_num, na.rm = TRUE),
    avg_frequency_5y_prior_endpoint = mean(frequency_num, na.rm = TRUE),
    .groups = "drop"
  )

#combine all the activity data back again
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity %>% 
  left_join(cohort_5y_prior, by="subject_id")

#for any dogs with NAs recode "reocrds_not_available"
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity %>%
  mutate(across(287:295, ~case_when(
      is.na(.) ~ "no_records_for_this_time_period",
    TRUE ~ as.character(.)  
  )))



```

### Reclassify values like in HSA log reg (like moderate activity from numeric score etc

```{r}
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity%>%
  mutate(avg_activity_duration_sy1_2 = as.numeric(as.character(avg_activity_duration_sy1_2))) %>%
  mutate(avg_activity_duration_sy1_2_splits = case_when(
    avg_activity_duration_sy1_2 <= 20 ~ "less_than_30min",
    avg_activity_duration_sy1_2 >20 & avg_activity_duration_sy1_2 <= 40 ~ "between_30_60_min",
    avg_activity_duration_sy1_2 >40 & avg_activity_duration_sy1_2<= 60 ~ "between_60_90_min",
    avg_activity_duration_sy1_2 >60 & avg_activity_duration_sy1_2 <= 80 ~ "between_90_120_min",
    avg_activity_duration_sy1_2 >80 & avg_activity_duration_sy1_2 <= 100 ~ "more_than_120_min"
  ))

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity %>%
  mutate(avg_activity_freq_sy1_2 = as.numeric(as.character(avg_activity_freq_sy1_2))) %>%
  mutate(avg_activity_freq_sy1_2_splits = case_when(
    avg_activity_freq_sy1_2 <= 20 ~ "never",
    avg_activity_freq_sy1_2 >20 & avg_activity_freq_sy1_2<= 40 ~ "less_than_weekly",
    avg_activity_freq_sy1_2 >40 & avg_activity_freq_sy1_2<= 60 ~ "min_weekly",
    avg_activity_freq_sy1_2 >60 & avg_activity_freq_sy1_2 <= 80 ~ "daily",
    avg_activity_freq_sy1_2 >80 & avg_activity_freq_sy1_2<= 100 ~ "multi_daily"
  ))


dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity <-dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity%>%
  mutate(avg_activity_intensity_sy1_2 = as.numeric(as.character(avg_activity_intensity_sy1_2))) %>%
  mutate(avg_activity_intensity_sy1_2_splits = case_when(
    avg_activity_intensity_sy1_2 <= 33.3 ~ "low",
    avg_activity_intensity_sy1_2 >33.3 & avg_activity_intensity_sy1_2<= 66.7 ~ "moderate",
    avg_activity_intensity_sy1_2 >66.7 & avg_activity_intensity_sy1_2<= 100 ~ "high",

  ))

#SY3 + intensity and freq


dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity %>%
  mutate(avg_activity_freq_sy3 = as.numeric(as.character(avg_activity_freq_sy3))) %>%
  mutate(avg_activity_freq_sy3_splits = case_when(
    avg_activity_freq_sy3 <= 20 ~ "rarely",
    avg_activity_freq_sy3 >20 & avg_activity_freq_sy3<= 40 ~ "monthly",
    avg_activity_freq_sy3 >40 & avg_activity_freq_sy3<= 60 ~ "weekly",
    avg_activity_freq_sy3 >60 & avg_activity_freq_sy3 <= 80 ~ "daily",
    avg_activity_freq_sy3 >80 & avg_activity_freq_sy3<= 100 ~ "multi_daily",
    is.na(avg_activity_freq_sy3) ~ "no_records_for_this_time_period" 
  ))


dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity %>%
  mutate(avg_activity_intensity_sy3 = as.numeric(as.character(avg_activity_intensity_sy3))) %>%
  mutate(avg_activity_intensity_sy3_splits = case_when(
    avg_activity_intensity_sy3 <= 20 ~ "slow_walks",
    avg_activity_intensity_sy3 >20 & avg_activity_intensity_sy3<= 40 ~ "average_walks",
    avg_activity_intensity_sy3 >40 & avg_activity_intensity_sy3<= 60 ~ "brisk_walks",
    avg_activity_intensity_sy3 >60 & avg_activity_intensity_sy3 <= 80 ~ "jogs",
    avg_activity_intensity_sy3 >80 & avg_activity_intensity_sy3<= 100 ~ "runs",
    is.na(avg_activity_intensity_sy3) ~ "no_records_for_this_time_period" 
  ))



#5y prior endpoint  freq


dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity %>%
  mutate(avg_frequency_5y_prior_endpoint = as.numeric(as.character(avg_activity_freq_sy3))) %>%
  mutate(avg_frequency_5y_prior_endpoint_splits = case_when(
    avg_frequency_5y_prior_endpoint <= 20 ~ "rarely",
    avg_frequency_5y_prior_endpoint >20 & avg_frequency_5y_prior_endpoint<= 40 ~ "monthly",
    avg_frequency_5y_prior_endpoint >40 & avg_frequency_5y_prior_endpoint<= 60 ~ "weekly",
    avg_frequency_5y_prior_endpoint >60 & avg_frequency_5y_prior_endpoint <= 80 ~ "daily",
    avg_frequency_5y_prior_endpoint >80 &avg_frequency_5y_prior_endpoint<= 100 ~ "multi_daily",
    is.na(avg_frequency_5y_prior_endpoint) ~"no_records_for_this_time_period"
  ))
#5y prior endpoint intensity
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity %>%
  mutate(avg_intensity_5y_prior_endpoint = as.numeric(as.character(avg_intensity_5y_prior_endpoint))) %>%
  mutate(avg_intensity_5y_prior_endpoint_splits = case_when(
    avg_intensity_5y_prior_endpoint <= 20 ~ "slow_walks",
    avg_intensity_5y_prior_endpoint >20 & avg_intensity_5y_prior_endpoint<= 40 ~ "average_walks",
    avg_intensity_5y_prior_endpoint >40 & avg_intensity_5y_prior_endpoint<= 60 ~ "brisk_walks",
    avg_intensity_5y_prior_endpoint >60 & avg_intensity_5y_prior_endpoint <= 80 ~ "jogs",
    avg_intensity_5y_prior_endpoint >80 & avg_intensity_5y_prior_endpoint<= 100 ~ "runs",
    is.na(avg_intensity_5y_prior_endpoint) ~"no_records_for_this_time_period"
  ))
#recode into larger groupings e.g grouping together rarely + monthly for frequency
#freq
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity <-dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity %>%
  mutate(avg_activity_freq_sy3_splits_grouped = case_when(
    avg_activity_freq_sy3_splits == "monthly" ~ "rarely",
    avg_activity_freq_sy3_splits == "multi_daily"~"daily_or_more" ,
    avg_activity_freq_sy3_splits == "daily" ~"daily_or_more" ,
    avg_activity_freq_sy3_splits == "weekly" ~"weekly" ,
    avg_activity_freq_sy3_splits == "rarely" ~ "rarely",
    avg_activity_freq_sy3_splits == "rarely" ~ "rarely",
  ))

#intensity
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity <-dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity %>%
  mutate(avg_activity_intensity_sy3_splits_grouped = case_when(
    avg_activity_intensity_sy3_splits == "jogs" ~ "high_intensity",
    avg_activity_intensity_sy3_splits == "runs" ~"high_intensity" ,
    avg_activity_intensity_sy3_splits == "brisk_walks" ~"moderate_intensity",
    avg_activity_intensity_sy3_splits =="average_walks" ~ "low_intensity",
     avg_activity_intensity_sy3_splits =="slow_walks" ~ "low_intensity"
  ))
#5y prior freq
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity %>%
  mutate(avg_frequency_5y_prior_endpoint_splits_grouped = case_when(
    avg_frequency_5y_prior_endpoint_splits == "monthly" ~ "rarely",
    avg_frequency_5y_prior_endpoint_splits == "multi_daily"~"daily_or_more" ,
    avg_frequency_5y_prior_endpoint_splits == "daily" ~"daily_or_more" ,
    avg_frequency_5y_prior_endpoint_splits == "weekly" ~"weekly" ,
    avg_frequency_5y_prior_endpoint_splits == "rarely" ~"rarely" ,
    avg_frequency_5y_prior_endpoint_splits == "no_records_for_this_time_period" ~"no_records_for_this_time_period" 
  ))

#5y prior intensity
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity<-dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity %>%
  mutate(avg_intensity_5y_prior_endpoint_splits_grouped = case_when(
    avg_intensity_5y_prior_endpoint_splits == "jogs" ~ "high_intensity",
    avg_intensity_5y_prior_endpoint_splits == "runs" ~"high_intensity" ,
    avg_intensity_5y_prior_endpoint_splits == "brisk_walks" ~"moderate_intensity",
    avg_intensity_5y_prior_endpoint_splits =="average_walks" ~ "low_intensity",
    avg_intensity_5y_prior_endpoint_splits =="slow_walks" ~ "low_intensity",
    avg_intensity_5y_prior_endpoint_splits =="no_records_for_this_time_period" ~ "no_records_for_this_time_period"
  ))
```

## Tidying up columns for final df

-   need to ensure columns (where indicated) have the correct NA e.g dogs with no medical records have "no_medical_records_for_time_point" rather than NA/not_recorded etc

```{r}
colnames(dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity)

#remove unnecessary columns:
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity2 <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity %>%
  dplyr::select(-c(record_date.x,X))

sum(is.na(dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity2))

#check which columns have most NAs and decide how best to replace
na_summary <- colSums(is.na(dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity2))

na_summary_df <- data.frame(
  column = names(na_summary),
  na_count = na_summary
) %>%
  arrange(desc(na_count))

na_summary_df
#majority of NAs are fine e.g in date withdrawn column

#ensure the rest are replaced by NAs, doing sequentially to ignore date columns which keep messing up the is.na conversion
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity3 <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity2 %>%
  mutate(across(2:21, ~ replace(., is.na(.), "records_not_available")))
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity3 <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity3 %>%
  mutate(across(36:260, ~ replace(., is.na(.), "records_not_available")))
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity3 <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity3 %>%
  mutate(across(273:304, ~ replace(., is.na(.), "records_not_available")))

#check NAs now
na_summary2 <- colSums(is.na(dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity3))

na_summary_df2 <- data.frame(
  column = names(na_summary2),
  na_count = na_summary2
) %>%
  arrange(desc(na_count))

na_summary_df2


#manually fix avg_weight_5y_prior_endpoint and MDI_quintile
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity3 <- 
  dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity3 %>%
  mutate(avg_weight_5y_prior_endpoint = replace(avg_weight_5y_prior_endpoint, 
                                                 is.na(avg_weight_5y_prior_endpoint), 
                                                 "records_not_available"))


dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity_tidied <- 
  dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity3 %>%
  mutate(MDI_quintile = replace(MDI_quintile, 
                                                 is.na(MDI_quintile), 
                                                 "records_not_available"))


```

```{r}
write.csv(dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity_tidied,here::here("Output/GRLS_MCT_variables_dataset.csv"))
```
