---
title: "MCT dataset variable creation"
format: html
editor: visual
---

## MCT case-control variable creation

```{r}
library(tidyverse)
library(lubridate)
source(here::here("R", "cohort_preparation.R"))
dog_pop <- read.csv(here::here("Output/MCT_cohort_250531.csv"))
```

Make sure population has the endpoint in date format (but also only to the year level to match to questinoaire etc. data which only annual)

```{r}
# Remove the excess record-date column and create endpoint values at year and
# month levels. Annual owner questionnaires do not justify day-level matching.
dog_pop <- prepare_cancer_cohort(dog_pop)
```

Datasets to join up to:

-   clinic and deprivation indices

-   Environment

-   lifestyle

-   reproduction

-   comorbidities

-   physical exam

-   medications

Need to create from these (where applicable):

-   early life (SY0-2)

-   rest of life (SY3 +)

-   whole life (SY0-final SY of records)

-   5y prior to endpoint

Had seperate files for all of these datasets for the HSA dataset creation but will add in here for other user of file ease

## Clinic and deprivation indices

created in this file for all the clinics based on 2019 values so can use these again as MDI not updated yet- GRLS clinic and deprivation indices linkup.qmd

```{r}
deprivation <- read.csv(here::here("Output/GRLS_dogs_MDI_2019.csv"))

# Join clinic and owner deprivation onto the cancer cohort by subject ID.
dog_pop_deprivation <- add_deprivation_features(dog_pop, deprivation)
```

## Environment exposures

NB. most of the enviro variables (or possibly all) do not have any missing data ie. they are all encoded 0 or 1 for the questionaires - so there is no 3rd level of "no_records/not_recorded" as with many other variables in the dataset

indoors exposures =

'use_aerosol','use_air_cleaner','use_hepa_filter,'use_moth_balls','use_incense_or_candles','smoke_exposure'

outside house exposures=

'any_treated_weeds','any_treated_insects','any_treated_fertilizer'

```{r}
# Environmental-condition and sleep-location questionnaires are separate source
# tables because sleep location can contain a different number of rows per dog.
source(here::here("Code", "GRLS_functions.R"))
source(here::here("R", "environment.R"))

environment_exposures <- read.csv(
  here::here("Data/environment_conditions.csv"),
  na.strings = c("", "NA")
)
sleep_environment <- read.csv(here::here("Data/environment_sleep_location.csv"),na.strings=c("","NA"))
sleep_environment <- sleep_environment %>% mutate(across(where(is.character), tolower))

```

### Tidy to binary exposure windows

Indoor exposures are aerosol use, air cleaners, HEPA filters, moth balls,
incense/candles, and household smoke. Outdoor exposures are treatment of weeds,
insects, and lawns with fertiliser.

For every exposure, the shared function creates indicators and counts for:

- early life (study years 0–2);
- rest of life (study years 3–10);
- whole observed life (study years 0–10); and
- the five years preceding the cancer endpoint.

The sleep-location table remains separate because dogs may have multiple sleep
locations within a questionnaire year.

```{r}
environment_features <- build_environment_binary_features(
  environment_exposures,
  dog_pop_deprivation
)

# Retain the established object names used by the smoke and sleep sections below.
indoors_outdoors <- environment_features$longitudinal %>%
  dplyr::select(-dplyr::all_of(setdiff(names(dog_pop_deprivation), "subject_id")))
dog_pop_deprivation_enviro__early_rest_whole4 <-
  environment_features$exposure_windows
dog_pop_deprivation_enviro5y <- environment_features$within_five_years
```

#### Hours of smoke exposure - early, rest, whole, 5y prior

Rows recorded as "see detail rows" do not contain a usable dosage and are
excluded. The shared helper converts the remaining values to numeric hours and
calculates total exposure in early life, the rest of life, the whole observed
life, and before diagnosis. It also retains the binary five-year exposure
features created above.

```{r}
dog_pop_deprivation_enviro_smoking_all <- build_smoke_dosage_features(
  dog_pop_deprivation_enviro__early_rest_whole4,
  dog_pop_deprivation_enviro5y
)
```

### Sleep location

Lots of dogs have multiple `specific_sleep_location` values in the same record
year. Duplicate rows are removed, but distinct locations are retained. The
three recorded locations are garage, house, and outside; the shared helper
creates early-life, rest-of-life, whole-life, and five-year indicators for each.

The main lifetime sleep location is the uniquely most frequent location. A tie
deliberately stops the analysis so that a subject-matter tie rule can be chosen
rather than silently selecting one location.

```{r}
dog_pop_deprivation_enviro_smoking_sleep_all <-
  build_sleep_location_features(
    sleep_environment,
    dog_pop_deprivation_enviro_smoking_all
  )
```

## Lifestyle

- As with the environmental and sleep data, every dog has lifestyle records, so
  this domain does not use a `not_recorded` or `records_not_available` category.
- The established broad groups are competitive, working, breed/showing,
  miscellaneous, companion/pet, and other. Category precedence remains the same
  as in the original notebook when a recorded role contains more than one term.
- A dog's main lifestyle is its most frequently recorded raw role. For the 403
  dogs with equally frequent roles, all tied roles are retained together and the
  broad category is `multiple`; this replaces the previous order-dependent
  selection of one role.

```{r}
source(here::here("R", "lifestyle.R"))

lifestyle <- read.csv(
  here::here("Data", "lifestyle.csv"),
  na.strings = c("", "NA")
)

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all <-
  add_lifestyle_features(
    dog_pop_deprivation_enviro_smoking_sleep_all,
    lifestyle
  )
```

## Household exposures

Household features come from `Data/house_details.csv`; residential location
features come from `Data/location_history.csv`. Text is normalized to lower case,
and missing or blank source values are represented as `no_records_available`.

Variables retained from this domain are:

- primary and secondary heating and cooking fuel;
- metal and plastic pipe indicators;
- water source, house type, area type, and mean recorded house age; and
- modal region, state, and owner ZIP.

Free-text fuel descriptions are mapped to the established electric, natural gas,
oil, propane, wood, and renewable groups. The curated spelling variants and
mixed-source descriptions are retained in `R/household.R`.

Pipe indicators combine the structured primary/secondary-house questions with
the curated free-text descriptions. If either structured material question has
no record, the derived material value remains `no_records_available`.

For water, municipal supply means a public water system managed by a local
authority or utility and treated to applicable US drinking-water standards. The
approved broad groups are municipal, groundwater, spring/surface, bottled, and
rainwater/cistern. These grouped values—not the original ungrouped answers—are
used for every derived water-source feature.

Early-life (study years 0–2), rest-of-life (3–10), and whole-life (0–10)
features use the most frequent recorded value. Five-year features use the most
frequent value in the five study years before the cohort endpoint. When values
tie, the output is explicitly `multiple` rather than selecting whichever row
appears first.

Dog `094-000461` has no
`year_in_study_diagnosis_or_final_record_year` in either frozen cancer cohort.
The historical notebook replaced this missing value with 1. The shared workflow
now stops so that the cohort endpoint definition must be corrected explicitly
before an analysis dataset is accepted.

```{r}
source(here::here("R", "household.R"))

house <- read.csv(here::here("Data", "house_details.csv"))
location <- read.csv(here::here("Data", "location_history.csv"))

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location2 <-
  build_household_features(
    house,
    location,
    dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all
  )
```

## Comorbidities

Load in all the comorbidity files:

```{r}
source(here::here("R/comorbidities.R"))

# Join domain extracts by the unique dog/year/to-date key. The extracts contain
# the same keys in different row orders, so row-wise cbind is not valid.
comorbs_all <- read_condition_domains(here::here("Data"))
comorbs_unique <- comorbs_all
#for 5y variables
comorbs_unique2 <- comorbs_all
#for other calcs
comorbs_unique3 <- comorbs_all
```

Comorbs to find:

```{r}
#potential chronic inflam conditions
comorbs_to_find <-c("congestive_heart_failure",
  "cardiomyopathy",
  "hypertension",
  "heartworm_infection",
  "lower_airway_disease",
  "dental_tarter_calculus",
  "addisons_disease",
  "autoimmune_thyroiditis",
  "cushings_disease",
  "diabetes_insipidus",
  "diabetes_mellitus",
  "hyperadrenalestrogenism",
  "hypercalcemia",
  "hypothyroidism",
  "pancreatic_insufficiency",
  "keratoconjunctivitis_sicca",
  "pigmentary_uveitis",
  "uveitis",
  "chronic_colitis",
  "malabsorptive_disorder",
  "megaesophagus",
  "pancreatitis",
  "anaplasma",
  "babesia",
  "canine_herpesvirus",
  "coccidia",
  "cryptosporidium",
  "ear_mites",
  "ehrlichia",
  "eimeria",
  "fleas",
  "fungal_infection",
  "giardia",
  "granuloma",
  "hookworms",
  "influenza",
  "isospora",
  "leptospirosis",
  "lyme_disease",
  "other_ectoparasite",
  "parvovirus",
  "rocky_mountain_spotted_fever",
  "roundworms",
  "tapeworms",
  "ticks",
  "tracheobronchitis",
  "unspecified_gastrointestinal_parasite",
  "valley_fever",
  "whipworms",
  "immune_related_musculoskeletal_disorder",
  "osteoarthritis",
  "rheumatoid_arthritis",
  "laryngeal_paralysis",
  "meningitis",
  "meningoencephalitis",
  "myasthenia_gravis",
  "erythema_multiforme",
  "flea_allergy_dermatitis",
  "food_allergy_dermatitis",
  "r_o_atopy")
#specific comorb groups 
intestinal_parasitism <- c("giardia", "hookworms", "coccidia", "tapeworms", "whipworms", "roundworms", "eimeria", "isospora","cryptosporidium","eimeria","tapeworms")
tick_borne_parasitism <-c("lyme_disease", "anaplasma", "rocky_mountain_spotted_fever", "ticks", "ehrlichia", "lyme disease", "lyme diseases, oa?", "lyme/anaplasma", "lyme+", "anaplasma and lyme +", "positive anaplasma", "erlichia +", "erlichia","valley_fever","babesia","anaplasma","ehrlichia","lyme_disease")
other_parasitism <-c("fleas", "fungal_infection", "heartworm", "ear_mites","other_ectoparasite","heartworm_infection")
all_parasitism<-c("giardia", "hookworms", "coccidia", "tapeworms", "whipworms", "roundworms", "eimeria", "isospora","lyme_disease", "anaplasma", "rocky_mountain_spotted_fever", "ticks", "ehrlichia", "lyme disease", "lyme diseases, oa?", "lyme/anaplasma", "lyme+", "anaplasma and lyme +", "positive anaplasma", "erlichia +", "erlichia","fleas", "fungal_infection", "heartworm", "ear_mites","heartworm_infection","other_ectoparasite")
infectious <- c("leptospirosis","parvovirus","bacterial cystitis/recurrent uti", "chronic uti's/ bacterial cystitis", "cystitis", "uti", "frequent utis", "previous utis", "chronic urinary incontinence","history of valley fever", "mma- coccidioides")
orthopaedic <-c("osteoarthritis", "l ccl partial tear", "oa of stifle", "hip dysplasia", "right elbow dysplasia grade ii", "ccl rupture", "acl tear", "tendonitis/ arthritis", "arthritis- cervical neck spine", "lh partial ccl tear, lh iliopsoas injury", "chronic lameness- hx of car accident", "lameness","hypothyroidism, chronic lameness, obesity","overweight, arthritis","dvd, larpar", "seizures, oa","tail fracture", "broken leg, histiocytoma","osteochondritis_dissecans","intervertebral_disc_disease","lameness","musculoskeletal_pain","osteomyeltiis","cruciate_ligament_rupture","hip_dysplasia","elbow_dysplasia","crepitus_decreased_range_of_motion","bone_fractures","limber_tail_syndrome","panosteitis")
immune_mediated <- c("uveitis", "autoimmune_thyroiditis", "keratoconjunctivitis_sicca", "megaesophagus", "diabetes_mellitus", "cushings_disease", "hypothyroidism", "granuloma", "immune complex formation", "suspected cushings/ thyroid carcinoma","dvd, larpar","immune_related_musculoskeletal_disorder","rheumatoid_arthritis","diabetes_insipidus","hypothyroidism","pancreatic_insuficiency","addisons_disease","autoimmune_thyroiditis","cushings_disease","hyperadrenalestrogenism","hypercalcemia")
cardiovascular <- c("cardiomyopathy", "hypertension", "heart base tumor", "heart base mass", "mild degenerative valve disease", "degenerative mitral and tricuspid valve disease", "mitral valve disease", "pulmonary thrombosis", "portal vein thrombus","arrhythmia","atrial enlargement","cardiomegaly","cardiomyopathy","congestive_heart_failure","hypertension","murmur","lower_airway_disease","myocardial_dysfunction","pleural_effusion","pulmonic_stenosis","pneumothorax","pulmonic_stenosis","subaortic_stenosis","tachycardia","valvular_disease")
gastrointestinal <-c("chronic_colitis", "pancreatitis", "malabsorptive_disorder", "chronic pancreatitis", "cystic endometrial hyperplasia", "intermittent mixed bowel diarrhea", "intermittent diarrhea - bacterial", "gastric ulcers","mesenteric volvulus","fb","esophagitis","diarrhea","dietary_indiscretion","food_sensitivity","gastrointestinal_foreign_body","gastorintestinal_ulcer","intussusception","malabsorptive_disorder","megaesophagus")

inflammatory_other <- c("tracheobronchitis", "hypercalcaemia", "pigmentary_uveitis", "ear infections", "chronic pyoderma", "chronic pruritis", "chronic rhinitis/ nasal discharge", "allergy, obesity", "pruritis", "severe allergies", "atopy", "allergies, chronic otitis", "chronic hot spots, skin infections", "many pyodera/dermatitis events", "chronic otitis, allergic dermatitis", "dermatitis", "food allergy", "skin allergies", "chronic otitis, overweight", "atopic dermatitis, allergies", "seasonal allergies/pyoderma", "chronic otitis, chronic allergies", "dermatitis lip fold","anal sac infection","otitis_externa","otitis_media","aural_haematoma","aural_foreign_body","pharyngitis","rhinitis","hypercalcaemia","keratoconjunctivitis_sicca","hypercalcemia","influenza","parvovirus")


chronic_inflammatory <-c("congestive_heart_failure",
  "cardiomyopathy",
  "hypertension",
  "heartworm_infection",
  "lower_airway_disease",
  "dental_tarter_calculus",
  "addisons_disease",
  "autoimmune_thyroiditis",
  "cushings_disease",
  "diabetes_insipidus",
  "diabetes_mellitus",
  "hyperadrenalestrogenism",
  "hypercalcemia",
  "hypothyroidism",
  "pancreatic_insufficiency",
  "keratoconjunctivitis_sicca",
  "pigmentary_uveitis",
  "uveitis",
  "chronic_colitis",
  "malabsorptive_disorder",
  "megaesophagus",
  "pancreatitis",
  "anaplasma",
  "babesia",
  "canine_herpesvirus",
  "coccidia",
  "cryptosporidium",
  "ear_mites",
  "ehrlichia",
  "eimeria",
  "fleas",
  "fungal_infection",
  "giardia",
  "granuloma",
  "hookworms",
  "influenza",
  "isospora",
  "leptospirosis",
  "lyme_disease",
  "other_ectoparasite",
  "parvovirus",
  "rocky_mountain_spotted_fever",
  "roundworms",
  "tapeworms",
  "ticks",
  "tracheobronchitis",
  "unspecified_gastrointestinal_parasite",
  "valley_fever",
  "whipworms",
  "immune_related_musculoskeletal_disorder",
  "osteoarthritis",
  "rheumatoid_arthritis",
  "laryngeal_paralysis",
  "meningitis",
  "meningoencephalitis",
  "myasthenia_gravis",
  "erythema_multiforme",
  "flea_allergy_dermatitis",
  "food_allergy_dermatitis",
  "r_o_atopy")
```

#### Chronic inflam conditions ever in all years of study

```{r}
source(here::here("Code/GRLS_functions.R"))

conditions_in_df_and_list <- intersect(colnames(comorbs_unique),comorbs_to_find)

comorbs_unique_2 <- comorbs_unique %>%
  #get rid of the to_date row for each dog
  filter(to_date !=1) %>%
  #get rid of the NA rows in the dataset - then add in later at the 3044 level to be "no_records_available" across the comorbs
  dplyr::select(-c(relationship_category,year_in_study,record_date,to_date))
###not using currently as want to keep the NA to tidy at end into "no_records_available"
#comorbs_unique_2<- comorbs_unique_2%>%
# replacing NA values with 0 across comorbidity columns
 # mutate_at(vars(3:196), ~replace_na(., 0))


#remove the rows in the comorbs_unique_2 that have got NAs across whole dataset for each dog. then when creating cohort variable at end can recode any of these dogs with no records as "no_records_available" (very few missing records across WHOLE study period, generally just occ year)

cols_to_check <- c(7:21,164:196)
comorbs_unique_2 <- comorbs_unique_2[complete.cases(comorbs_unique_2[, cols_to_check]), ]
#checking all dogs have some records to see if need to code any with "no_records_available" at end - all dogs have comorbs records in their lifetime
comorbs_unique <- unique(comorbs_unique_2$subject_id)

df_result2 <- comorbs_unique_2 %>%
  group_by(subject_id) %>%
  mutate(
    #is there any chronic inflam condition present in at least 1 row for subject id
    chronic_inflam_present_lifetime = if_else(
      rowSums(across(all_of(conditions_in_df_and_list))) > 0 & 
        rowSums(across(all_of(conditions_in_df_and_list), ~ sum(. == 1)) > 1), 
      1, 
      0
    ),
    #how many chronic inflam conditions
    number_comorbs_present_lifetime = if_else(
      chronic_inflam_present_lifetime == 1,
      rowSums(across(all_of(conditions_in_df_and_list))),
      0
    ),
    ,
    #what were the comorbidities
    comorbidities_lifetime = if_else(
      chronic_inflam_present_lifetime == 1,
      apply(across(all_of(conditions_in_df_and_list)), 1, function(x) paste(names(which(x == 1)), collapse = "_")),
      "")
  ) %>%
  ungroup()


```

#### Comorbs in final 5y prior to endpoint

```{r}
#keeping record_date for linking up to cancer diagnosis but remove to_date column
comorbs_unique_3 <- comorbs_unique2 %>%
#  mutate_at(vars(5:200), ~replace_na(., 0)) %>%
  filter(to_date !=1) %>%
  dplyr::select(-c(relationship_category,to_date)) %>%
  rename(year_in_study_comorb = year_in_study)
cols_to_check <- c(5:23,26:164,165:198)
comorbs_unique_3 <- comorbs_unique_3[complete.cases(comorbs_unique_3[, cols_to_check]), ]

##checking all dogs have some records to see if need to code any with "no_records_available" at end - all dogs have comorbs records for at least some of last 5y = dont need to do "record_unavailable"
comorbs_unique_5y <- unique(comorbs_unique_3$subject_id)

#run all functions above onto this dataset
df_result3 <- comorbs_unique_3 %>%
  group_by(subject_id) %>%
  mutate(
    #is there any chronic inflam condition present in at least 1 row for subject id
    chronic_inflam_present_5y = if_else(
      rowSums(across(all_of(conditions_in_df_and_list))) > 0 & 
        rowSums(across(all_of(conditions_in_df_and_list), ~ sum(. == 1)) > 1), 
      1, 
      0
    ),
    #how many chronic inflam conditions
    number_comorbs_present_5y = if_else(
      chronic_inflam_present_5y == 1,
      rowSums(across(all_of(conditions_in_df_and_list))),
      0
    ),
    ,
    #what were the comorbidities
    comorbidities_5y = if_else(
      chronic_inflam_present_5y == 1,
      apply(across(all_of(conditions_in_df_and_list)), 1, function(x) paste(names(which(x == 1)), collapse = "_")),
      "")
  ) %>%
  ungroup()


#keep only top 5 rows per patient 
df_result_restricted <- df_result3 %>%
  group_by(subject_id) %>%
  filter(year_in_study_comorb %in% sort(unique(year_in_study_comorb), decreasing = TRUE)[1:5]) %>%
  ungroup()
  


# Function to detect comorbidities for each subject_id based on index range
detect_comorbidity <- function(df, subject_id_col, index_range, comorbidity_groups) {
  # Extract actual column names within the index range
  valid_columns <- colnames(df)[index_range]

  df %>%
    group_by(!!sym(subject_id_col)) %>%  # Group by subject_id
    summarise(across(all_of(valid_columns), ~ any(. == 1), .names = "temp_{.col}"), .groups = "drop") %>%  # Check for 1s
    mutate(
      !!!map(comorbidity_groups, function(cols) {
        # Identify valid columns that exist in the dataset
        matched_cols <- intersect(paste0("temp_", cols), colnames(.))
        
        # Assign "Y" if any value in matched columns is 1, else "N"
        if (length(matched_cols) > 0) {
          ifelse(rowSums(dplyr::select(., all_of(matched_cols))) > 0, "Y", "N")
        } else {
          "N"  # If no matching columns exist, default to "N"
        }
      })
    ) %>%
    dplyr::select(-starts_with("temp_"))  # Remove temporary columns
}

# Define comorbidity groups (list of column names matching each category)
comorbidity_groups <- list(
  intestinal_parasitism = intestinal_parasitism,
  tick_borne_parasitism = tick_borne_parasitism,
  other_parasitism = other_parasitism,
  all_parasitism = all_parasitism,
  infectious = infectious,
  orthopaedic = orthopaedic,
  immune_mediated = immune_mediated,
  cardiovascular = cardiovascular,
  gastrointestinal = gastrointestinal,
  inflammatory_other = inflammatory_other,
  chronic_inflammatory = chronic_inflammatory
)

# Apply function to data
GRLS_data <- df_result_restricted %>%
  left_join(detect_comorbidity(df_result_restricted, "subject_id", 5:198, comorbidity_groups),
            by = "subject_id")


#keep only a single row for each patient
# Define the index range of comorbidity columns
comorbidity_index <- c(199:212)  # Keep subject_id out of this list

# Create a second dataframe with one row per subject_id, summarizing only the selected columns - not using this as need to ensure rows with a 1 are preferentially kept
#GRLS_summary <- GRLS_data %>%
#  group_by(subject_id) %>%
#   arrange(desc(number_comorbs_present_5y)) %>%
#  summarise(
#    subject_id = first(subject_id),  # Keep subject_id as is
#    across(all_of(colnames(GRLS_data)[comorbidity_index]), ~ ifelse(any(. == "Y"), "Y", "N")),
 #   .groups = "drop"
 # )


GRLS_summary <- GRLS_data %>%
  group_by(subject_id) %>%
  arrange(desc(number_comorbs_present_5y)) %>%  # prioritise higher values
  slice(1) %>%  # keep only the top row per subject
  ungroup()


### recode any NA as "no_record_available"
GRLS_summary <- GRLS_summary%>%
  mutate(across(where(~ is.character(.) | is.factor(.)| is.numeric(.)),
                ~ replace(., is.na(.) | . == "", "no_records_available")))
```

There are also a separate subset of conditions that if ever reported as diagnosed then will have for life (so do not need \>1 y study records with it available):

```{r}
comorbs_single_record <- c(
  "atrial_enlargement",
  "cardiomegaly",
  "cardiomyopathy",
  "congestive_heart_failure",
  "hypertension",
  "myocardial_dysfunction",
  "valvular_disease",
  "addisons_disease",
  "autoimmune_thyroiditis",
  "cushings_disease",
  "diabetes_insipidus",
  "diabetes_mellitus",
  "hyperadrenalestrogenism",
  "hypercalcemia",
  "hypothyroidism",
  "pancreatic_insufficiency",
  "cataracts",
  "glaucoma",
  "keratoconjunctivitis_sicca",
  "progressive_retinal_atrophy_or_degeneration",
  "pigmentary_uveitis",
  "lymphangiectasia",
  "malabsorptive_disorder",
  "megaesophagus",
  "motility_disorder",
  "elbow_dysplasia",
  "hip_dysplasia",
  "osteoarthritis",
  "rheumatoid_arthritis",
  "spondylosis"
)

#identify in records

comorbs_ever <- comorbs_unique3 %>%
  dplyr::select(-c(relationship_category,year_in_study,record_date,to_date))

#comorbs_ever<- comorbs_ever%>%
# replacing NA values with 0 across comorbidity columns
 # mutate_at(vars(3:196), ~replace_na(., 0))


### restrict columns so that those with all NAs are removed and then can recode back as "no_records_available" later
cols_to_check <- c(3:21,24:162,164:196)
comorbs_ever<- comorbs_ever[complete.cases(comorbs_ever[, cols_to_check]), ]
#all dogs have at least some records here
unique_comorbs_ever <- unique(comorbs_ever$subject_id)
#remove all the columns that are not in comorbs_single_record list or subject_id
comorbs_ever <- comorbs_ever %>%
  dplyr::select(subject_id, all_of(comorbs_single_record))

#
lifetime_comorbs_ever <- comorbs_ever %>%
  group_by(subject_id) %>%
  mutate(
    ## ANY lifetime comorb inflam present
    lifelong_comorb_present = if_else(rowSums(across(all_of(comorbs_single_record))) > 0, 1, 0),
    
    ### NUMBER OF chronic inflam conditions present
    number_lifelong_comorbs_present = rowSums(across(all_of(comorbs_single_record))),
    
    # what were the comorbidities
    lifelong_comorbidities = if_else(
      lifelong_comorb_present == 1,
      apply(across(all_of(comorbs_single_record)), 1, function(x) paste(names(which(x == 1)), collapse = "_")),
      NA_character_
    )
  ) %>%
  ungroup()

#keep count of comorbs and if present and then only a unique row for each patient
lifetime_comorbs_ever_abr <- lifetime_comorbs_ever %>%
    dplyr::select(c(subject_id,lifelong_comorb_present,number_lifelong_comorbs_present))%>%
  group_by(subject_id) %>%
  arrange(desc(number_lifelong_comorbs_present)) %>%  # prioritise higher values
  slice(1) %>%  # keep only the top row per subject
  ungroup()
```

Grouping the lifelong_comorbidities column into type of condition e.g endocrine, ocular etc based on earlier lists

```{r}

valid_comorbidity_groups <- lapply(comorbidity_groups, function(cols) {
  intersect(cols, colnames(lifetime_comorbs_ever))  # Keep only columns that exist in df
})

# Remove empty comorbidity groups
valid_comorbidity_groups <- valid_comorbidity_groups[lengths(valid_comorbidity_groups) > 0]





# Compute 'lifetime' flags for each comorbidity category
df_lifetime <- lifetime_comorbs_ever %>%
  group_by(subject_id) %>%
  summarise(across(everything(), \(x) max(x, na.rm = TRUE))) %>%  # Collapse per patient (1 if present, 0 otherwise)
  mutate(!!!setNames(
    lapply(valid_comorbidity_groups, function(cols) {
      if (length(cols) > 0) {
        as.integer(rowSums(.[cols], na.rm = TRUE) > 0)  # Correctly sum across selected columns
      } else {
        rep(0, n())  # Handle cases where no valid columns exist
      }
    }),
    paste0("lifelong_", names(valid_comorbidity_groups))  # Create new column names
  )) %>%
  select(subject_id, starts_with("lifelong_"))  # Keep only new columns

# View output
print(df_lifetime)
```

Combine the lifetime_comorbs_ever_abr and df_lifetime so that we have counts and groupings of these comorbs:

```{r}
lifelong_data <- df_lifetime %>%
  left_join(lifetime_comorbs_ever_abr,by="subject_id")

lifelong_data <- lifelong_data %>%
  dplyr::select(-c(lifelong_comorb_present.y))


GRLS_summary_lifelong<- lifelong_data %>%
  left_join(GRLS_summary,by="subject_id")

#add in _5y to the GRLS summary columns that are from 5y prior to final record
colnames(GRLS_summary_lifelong)[211:221] <- paste0(colnames(GRLS_summary_lifelong)[211:221], "_5y")

```

Lifetime comorbs now classify into the parasitism/infectious/cardio/ etc

```{r}

# Apply function to data
GRLS_data_lifetime <- df_result3 %>%
  left_join(detect_comorbidity(df_result_restricted, "subject_id", 5:198, comorbidity_groups),
            by = "subject_id")

#add in _5y to the GRLS summary columns that are from 5y prior to final record
#colnames(GRLS_data_lifetime)[202:212] <- paste0(colnames(GRLS_data_lifetime)[202:212], "_5y")

#keep only a single row for each patient
# Define the index range of comorbidity columns
comorbidity_index <- c(199:212)  # Keep subject_id out of this list

# Create a second dataframe with one row per subject_id, summarizing only the selected columns
GRLS_summary_lifetime <- GRLS_data_lifetime %>%
  group_by(subject_id) %>%
  summarise(
    subject_id = first(subject_id),  # Keep subject_id as is
    across(all_of(colnames(GRLS_data_lifetime)[comorbidity_index]), ~ ifelse(any(. == "Y"), "Y", "N")),
    .groups = "drop"
  )
#add lifetime to end of columns at 5:15 indices
names(GRLS_summary_lifetime)[5:15] <- paste0(names(GRLS_summary_lifetime)[5:15], "_lifetime")

```

Combine lifelong and 5y values into one as consider these comorbs to be permanent once diagnosed:

```{r}
#mutate the Y N in 5y cols to be 0 and 1 for counting in next step



GRLS_summary_lifelong <- GRLS_summary_lifelong %>%
  mutate(across(c(4:9, 211:221), ~ case_when(
    . == "N" ~ 0,
    . == "Y" ~ 1,
    TRUE ~ as.numeric(.)
  )))

#combining counts in 5y columns and then remove the lifelong ones 
GRLS_summary_lifelong <- GRLS_summary_lifelong %>%
  mutate(
    orthopaedic_5y2 = pmax(lifelong_orthopaedic, orthopaedic_5y, na.rm = TRUE),
    cardiovascular_5y2=pmax(lifelong_cardiovascular,cardiovascular_5y,na.rm=TRUE),
    immune_mediated_5y2=pmax(lifelong_immune_mediated,immune_mediated_5y,na.rm=TRUE),
    gastrointestinal_5y2=pmax(lifelong_gastrointestinal,gastrointestinal_5y,na.rm=TRUE),
    inflammatory_other_5y2=pmax(lifelong_inflammatory_other,inflammatory_other_5y,na.rm=TRUE),
    chronic_inflammatory_5y2=pmax(lifelong_chronic_inflammatory,chronic_inflammatory_5y,na.rm=TRUE))

#remove the earlier 5y columns and just keep the ones where lifelong + 5y ar ecombined
GRLS_summary_lifelong_2 <- GRLS_summary_lifelong %>%
  dplyr::select(c(subject_id,222:227))
```

Combine all of the 5y prior and ever comorb datasets

```{r}
GRLS_5yprior_lifetime <- GRLS_summary_lifelong_2 %>%
  left_join(GRLS_summary_lifetime,by="subject_id")

GRLS_5yprior_lifetime <- GRLS_5yprior_lifetime %>%
  dplyr::select(-c(8:10))

GRLS_5yprior_lifetime <- GRLS_5yprior_lifetime %>%
  mutate(across(c(8:18), ~ case_when(
    . == "N" ~ 0,
    . == "Y" ~ 1,
    TRUE ~ as.numeric(.)
  )))
```

#### Add neoplasia variables in comorb dataset - lifetime

NB. remove lymphoma from lists as will skew for all the cases to have a 1 and none of the controls by definition

```{r}
#ggrouping these
all_cancers <- c("adrenal_tumor","basal_cell_tumor","bile_duct_tumor","bladder_tumor","brain_spinal_cord_tumor","breast_or_mammary_tumor","eye_tumor","hair_matrix_tumor","heart_tumor","hemangiosarcoma",
                 "histiocytic_sarcoma","kidney_tumor","liver_tumor","lung_tumor","mast_cell_tumor","melanmoma","multiple_myeloma","nasal_tumor","osteosarcoma","pancreatic_tumor","perianal_adenoma","pituitary_tumor","plasma_cell_tumor","plasmacytoma","prostate_tumor","soft_tissue_sarcoma","splenic_tumor","squamous_cell_carcinoma","stomach_intestinal_tumor","testicular_tumor","thymoma","thyroid_tumor")
malig_cancers <- c("hemangiosarcoma",
  "histiocytic_sarcoma",  "melanoma", "multiple_myeloma", "osteosarcoma", "pancreatic_tumor", "plasma_cell_tumor", "plasmacytoma", "soft_tissue_sarcoma", "squamous_cell_carcinoma")
benign_cancers <- c("adrenal_tumor", "basal_cell_tumor", "bile_duct_tumor", "bladder_tumor", "brain_spinal_cord_tumor", "breast_or_mammary_tumor", "eye_tumor", "hair_matrix_tumor", "heart_tumor", "kidney_tumor", "liver_tumor", "lung_tumor", "mast_cell_tumor", "nasal_tumor", "perianal_adenoma", "pituitary_tumor", "prostate_tumor", "splenic_tumor", "stomach_intestinal_tumor", "testicular_tumor", "thymoma", "thyroid_tumor")

```

Create YN and counts for all cancers, benign and malig

```{r}
comorbs_neoplasia_2 <- conditions_neoplasia %>%
  filter(to_date != 1)%>%
  dplyr::select(-other_specify)

#for next step of 5y
comorbs_neoplasia_3 <- conditions_neoplasia %>%
  filter(to_date != 1)%>%
  dplyr::select(-other_specify)
#remove rows where comorbs NA
cols_to_check <- c(6:40)
comorbs_neoplasia_2 <- comorbs_neoplasia_2[complete.cases(comorbs_neoplasia_2[, cols_to_check]), ]


#check if any dogs have no neoplasia records - all do have some records at least for lifetime calcs
unique_cancer <- unique(comorbs_neoplasia_2$subject_id)
comorbs_neoplasia_2$all_cancers_count_lifetime <- 0
comorbs_neoplasia_2$malig_cancers_count_lifetime <- 0
comorbs_neoplasia_2$benign_cancers_count_lifetime <- 0


# Define the cancer columns indices (from index 2 to the last column)
cancer_columns <- 7:40

# Loop through the cancer columns by index
for (col_index in cancer_columns) {
  col_name <- colnames(comorbs_neoplasia_2)[col_index]  # Get column name
  
  # Count for all cancers
  comorbs_neoplasia_2$all_cancers_count_lifetime <- comorbs_neoplasia_2$all_cancers_count_lifetime + (comorbs_neoplasia_2[[col_name]] == 1)
  
  # Count for all cancers
  comorbs_neoplasia_2$all_cancers_count_lifetime <- comorbs_neoplasia_2$all_cancers_count_lifetime + (comorbs_neoplasia_2[[col_name]] == 1)
  
  # Count for malignant cancers
  if (col_name %in% malig_cancers) {
    comorbs_neoplasia_2$malig_cancers_count_lifetime <- comorbs_neoplasia_2$malig_cancers_count_lifetime + (comorbs_neoplasia_2[[col_name]] == 1)
  }
  
  # Count for benign cancers
  if (col_name %in% benign_cancers) {
    comorbs_neoplasia_2$benign_cancers_count_lifetime <- comorbs_neoplasia_2$benign_cancers_count_lifetime + (comorbs_neoplasia_2[[col_name]] == 1)
  }
}

#create Y N columns for each in addition to counts
comorbs_neoplasia_2 <- comorbs_neoplasia_2 %>%
  mutate(
    all_cancers_YN_lifetime = ifelse(all_cancers_count_lifetime >= 1, 1, 0),
    malig_cancers_YN_lifetime = ifelse(malig_cancers_count_lifetime >= 1, 1, 0),
    benign_cancers_YN_lifetime = ifelse(benign_cancers_count_lifetime >= 1, 1, 0),

  )


#code to get to unique rows and keep the max count for each dog
#get the max count value for each of the counts for each dog 
cancer_count_summary <- comorbs_neoplasia_2 %>%
  group_by(subject_id) %>%
  summarise(
    malig_cancers_count_lifetime = if_else(all(is.na(malig_cancers_count_lifetime)), 0, max(malig_cancers_count_lifetime, na.rm = TRUE)),
    benign_cancers_count_lifetime = if_else(all(is.na(benign_cancers_count_lifetime)), 0, max(benign_cancers_count_lifetime, na.rm = TRUE)),
    all_cancers_count_lifetime = if_else(all(is.na(all_cancers_count_lifetime)), 0, max(all_cancers_count_lifetime, na.rm = TRUE)),
    malig_cancers_count_lifetime = if_else(all(is.na(malig_cancers_count_lifetime)), 0, max(malig_cancers_count_lifetime, na.rm = TRUE)),
    all_cancers_count_lifetime = if_else(all(is.na(all_cancers_count_lifetime)), 0, max(all_cancers_count_lifetime, na.rm = TRUE)),

    malig_cancers_YN_lifetime = if_else(all(is.na(malig_cancers_YN_lifetime)), 0, max(malig_cancers_YN_lifetime, na.rm = TRUE)),
    benign_cancers_YN_lifetime = if_else(all(is.na(benign_cancers_YN_lifetime)), 0, max(benign_cancers_YN_lifetime, na.rm = TRUE)),
    all_cancers_YN_lifetime = if_else(all(is.na(all_cancers_YN_lifetime)), 0, max(all_cancers_YN_lifetime, na.rm = TRUE)),
    malig_cancers_YN_lifetime = if_else(all(is.na(malig_cancers_YN_lifetime)), 0, max(malig_cancers_YN_lifetime, na.rm = TRUE)),
    all_cancers_YN_lifetime = if_else(all(is.na(all_cancers_YN_lifetime)), 0, max(all_cancers_YN_lifetime, na.rm = TRUE)),
    
    .groups = "drop"
  )

```

#### Neoplasia YN and counts - 5y prior to endpoint

```{r}
#keep only the max 5 rows for that dog in year_in_study then same calcs as above
comorbs_neoplasia_3 <- comorbs_neoplasia_3 %>%
  group_by(subject_id) %>%
  #keeping only the 5 most recent records
  filter(year_in_study %in% sort(unique(year_in_study), decreasing = TRUE)[1:5]) %>%
  ungroup()


#remove rows where comorbs NA
cols_to_check <- c(6:40)
comorbs_neoplasia_3 <- comorbs_neoplasia_3[complete.cases(comorbs_neoplasia_3[, cols_to_check]), ]


#check if any dogs have no neoplasia records - only 2165 have records fo final 5y
unique_cancer2 <- unique(comorbs_neoplasia_3$subject_id)
  

comorbs_neoplasia_3$all_cancers_count_5y <- 0
comorbs_neoplasia_3$malig_cancers_count_5y <- 0
comorbs_neoplasia_3$benign_cancers_count_5y <- 0


# Define the cancer columns indices (from index 2 to the last column)
cancer_columns <- 6:41

# Loop through the cancer columns by index
for (col_index in cancer_columns) {
  col_name <- colnames(comorbs_neoplasia_3)[col_index]  # Get column name
  
  # Count for all cancers
  comorbs_neoplasia_3$all_cancers_count_5y <- comorbs_neoplasia_3$all_cancers_count_5y + (comorbs_neoplasia_3[[col_name]] == 1)
  
  # Count for all cancers
  comorbs_neoplasia_3$all_cancers_count_5y <- comorbs_neoplasia_3$all_cancers_count_5y + (comorbs_neoplasia_3[[col_name]] == 1)
  
  # Count for malignant cancers
  if (col_name %in% malig_cancers) {
    comorbs_neoplasia_3$malig_cancers_count_5y <- comorbs_neoplasia_3$malig_cancers_count_5y + (comorbs_neoplasia_3[[col_name]] == 1)
  }
  
  # Count for benign cancers
  if (col_name %in% benign_cancers) {
    comorbs_neoplasia_3$benign_cancers_count_5y <- comorbs_neoplasia_3$benign_cancers_count_5y + (comorbs_neoplasia_3[[col_name]] == 1)
  }
}

#create Y N columns for each in addition to counts
comorbs_neoplasia_3 <- comorbs_neoplasia_3 %>%
  mutate(
    all_cancers_YN_5y = ifelse(all_cancers_count_5y >= 1, 1, 0),
    malig_cancers_YN_5y = ifelse(malig_cancers_count_5y >= 1, 1, 0),
    benign_cancers_YN_5y = ifelse(benign_cancers_count_5y >= 1, 1, 0),

  )


#code to get to unique rows and keep the max count for each dog
#get the max count value for each of the counts for each dog 
cancer_count_summary_2 <- comorbs_neoplasia_3 %>%
  group_by(subject_id) %>%
  summarise(
    malig_cancers_count_5y = if_else(all(is.na(malig_cancers_count_5y)), 0, max(malig_cancers_count_5y, na.rm = TRUE)),
    benign_cancers_count_5y = if_else(all(is.na(benign_cancers_count_5y)), 0, max(benign_cancers_count_5y, na.rm = TRUE)),
    all_cancers_count_5y = if_else(all(is.na(all_cancers_count_5y)), 0, max(all_cancers_count_5y, na.rm = TRUE)),
    malig_cancers_count_5y = if_else(all(is.na(malig_cancers_count_5y)), 0, max(malig_cancers_count_5y, na.rm = TRUE)),
    all_cancers_count_5y = if_else(all(is.na(all_cancers_count_5y)), 0, max(all_cancers_count_5y, na.rm = TRUE)),

    malig_cancers_YN_5y = if_else(all(is.na(malig_cancers_YN_5y)), 0, max(malig_cancers_YN_5y, na.rm = TRUE)),
    benign_cancers_YN_5y = if_else(all(is.na(benign_cancers_YN_5y)), 0, max(benign_cancers_YN_5y, na.rm = TRUE)),
    all_cancers_YN_5y = if_else(all(is.na(all_cancers_YN_5y)), 0, max(all_cancers_YN_5y, na.rm = TRUE)),
    malig_cancers_YN_5y = if_else(all(is.na(malig_cancers_YN_5y)), 0, max(malig_cancers_YN_5y, na.rm = TRUE)),
    all_cancers_YN_5y = if_else(all(is.na(all_cancers_YN_5y)), 0, max(all_cancers_YN_5y, na.rm = TRUE)),
    
    .groups = "drop"
  )

#combine the 2 cancer summaries
cancer_summary_5y_lifetime <- cancer_count_summary %>%
  left_join(cancer_count_summary_2,by="subject_id")

#replace all NAs with "no_records_available" for the 5y prior ones which dont have any records for neoplasia in final 5y
cancer_summary_5y_lifetime <- cancer_summary_5y_lifetime %>%
  mutate(across(2:13, ~ case_when(
    is.na(.) ~ "no_records_available",
    TRUE ~ as.character(.)
  )))
```

Combine all comorbs data back into the main df

```{r}
#combine cancer and other comorbs
GRLS_comorbs_cancer_5yprior_lifetime <- GRLS_5yprior_lifetime %>%
  left_join(cancer_summary_5y_lifetime,by="subject_id")

#combine back to main df
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs <- dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location2 %>% left_join(GRLS_comorbs_cancer_5yprior_lifetime,by="subject_id")
```

## Medications

Using the df created in an earlier script of analyses which tidies up free text and the many drug categories etc. in this qmd GRLS medications.qmd

```{r}
medications <- read.csv(here::here("Output/GRLS_medications_initial_tidy.csv"))
source(here::here("R", "medications.R"))
```

### Lifetime medication calculations

```{r}
medication_features_lifetime <- summarise_medication_quartiles(
  medications,
  suffix = "lifetime",
  missing_value = "medication_not_prescribed"
)

meds_per_dog_lifetime <- medication_features_lifetime$counts
medication_summary_lifetime <- medication_features_lifetime$summary
meds_per_dog_quart_lifetime <- medication_features_lifetime$categorised
meds_wide_quart_lifetime <- medication_features_lifetime$wide

write.csv(medication_summary_lifetime,here::here("Output/MCT_study_medication_quartiles_summaries_lifetime.csv"))
meds_wide_quart_lifetime
```

### 5y prior to endpoint medication calculations

NB. there is 2889 dogs in this 5y - dogs with meds records in 5y prior to death, ensure when joining back into main 3044 dogs that these \~150 dogs have "no_records_available" coded on them

```{r}
medications_5y <- select_last_study_years(medications, 5)

medication_features_5y <- summarise_medication_quartiles(
  medications_5y,
  suffix = "5y",
  missing_value = "medication_not_used"
)

meds_per_dog_5y <- medication_features_5y$counts
medication_summary_5y <- medication_features_5y$summary
meds_per_dog_quart_5y <- medication_features_5y$categorised
meds_wide_quart_5y <- medication_features_5y$wide

write.csv(medication_summary_5y,here::here("Output/MCT_study_medication_quartiles_summaries_5y.csv"))
meds_wide_quart_5y
```

### Examine for specific potential RF medications:

-   NSAIDs exposure

    -   ever Y N and 5y prior Y N
    -   counts

-   steroids exposure

    -   ever Y N and 5y prior Y N
    -   counts

-   Fenbendazole exposure

    -   ever Y N and 5y prior Y N
    -   counts

```{r}
#free text drug names to look for
NSAID_names<-c("carprofen","meloxicam","firocoxib","deracoxib","flurbiprofen","piroxicam","ketotifen","ketoprofen","rimadyl","metacam","meloxidyl","previcox","vetprofen","ketorolac opthalmic ointment","toradol","acular","voltaren
","diclofenac","flurbiprofen eye drops
","carprovet","carprieve","truprofen","ofloxacin ophthalmic ointment + nsaid
")
steroid_names<-c("dexamethasone, neomycin sulfate, thiabendazole
","prednisolone acetate
","betamethasone, clotrimazole, gentamicin sulfate
","betamethasone, gentamicin sulfate
","dexamethasone","dexamethasone, miconazole nitrate
","dexamethasone, miconazole nitrate
","methylprednisolone","prednisone","betamethasone, florfenicol, terbinafine
","clotrimazole, gentamicin sulfate, mometasone furoate
","dexamethasone, methylprednisolone
","mometasone furoate, orbifloxacin, posaconazole
","dexamethasone, neomycin sulfate, polymyxin b
","bacitracin zinc, dexamethasone, neomycin sulfate, polymyxin b
","miconazole nitrate, polymyxin b, prednisolone acetate
","hydrocortisone, gentamicin sulfate, miconazole nitrate
","gentamicin sulfate, hydrocortisone, miconazole nitrate
","burow's solution, hydrocortisone
","acetic acid, boric acid, hydrocortisone, ketoconazole
","hydrocortisone","bacitracin zinc, hydrocortisone, neomycin sulfate, polymyxin b
","dexamethasone, enrofloxacin, miconazole nitrate
","clotrimazole, dexamethasone, enrofloxacin
","hydrocortisone, ketoconazole
","dexamethasone, ketoconazole, phytosphingosine
","betamethasone","acetic acid, boric acid, ceramide complex, hydrocortisone, ketoconazole
","hydrocortisone, troleandomycin
","dexamethasone, enrofloxacin, ketoconazole, tris-edta
","bacitracin zinc, neomycin sulfate, polymyxin b, unspecified steroid
","clotrimazole, dexamethasone
","dexamethasone, ketoconazole, tris-edta
","dexamethasone, ketoconazole
","betamethasone, clotrimazole, gentamicin sulfate, mometasone furoate
","dexamethasone, diphenhydramine
","beclomethasone, clotrimazole, neomycin sulfate
","dexamethasone, enrofloxacin, ketoconazole
","dexamethasone, dimethyl sulfoxide, enrofloxacin, lidocaine
","dexamethasone, enrofloxacin
","betamethasone, gentamicin sulfate, miconazole nitrate
","chloramphenicol, dexamethasone
","hydrocortisone, pramoxine
","dexamethasone, penicillin
","betamethasone, chloramphenicol, ketoconazole
","dexamethasone, tobramycin
","enrofloxacin, dexamethasone, tris-edta
","dexamethasone, miconazole nitrate, neomycin sulfate, nystatin, thiostrepton, triamcinolone acetonide
","betamethasone, burow's solution, gentamicin sulfate, hydrocortisone, miconazole nitrate
","ketoconazole, hydrocortisone
","bacitracin, hydrocortisone
","dexamethasone, miconazole nitrate, tris-edta
","chlorhexidine gluconate, dexamethasone
","dexamethasone, enrofloxacin, tris-edta
","dexamethasone, enrofloxacin, silver sulfadiazine
","dexamethasone, miconazole nitrate, neomycin sulfate, polymyxin b, prednisolone acetate, thiabendazole
","betamethasone, enrofloxacin
","dexamethasone, ear cleaner (epi-otic advanced), enrofloxacin
","betamethasone, tobramycin
","dexamethasone ophthalmic ointment
","miconazole + dexamethasone ear solution
","amikacin + dexamethasone ear solution
","depo-medrol
","prednisolone acetate ophthalmic suspension
","neo-poly-dex ophthalmic ointment
","neo-poly-bac ophthalmic ointment + dexamethasone
","betamethasone + gentamicin ear solution
","conofite + dexamethasone ear solution
","neo-poly-bac ophthalmic ointment + hydrocortisone
","baytril + conofite + dexamethasone ear solution
","betamethasone + gentamicin spray
","baytil + clotrimazole + dexamethasone ear solution
","baytril + dexamethasone + miconazole ear solution
","baytril + conzol + dexamethasone ear solution
","trizulta + keto + baytril + dexamethasone flush
","clotrimazole + dexamethasone ear solution
","burow's solution + hydrocortisone
","betamethasone + clotrimazole + gentamicin ear solution
","trizulta + keto + dexamethasone flush
","cortalone
","baytril + dexamethasone + ketoconazole ear solution
","betamethasone + gentamicin sulfate + miconazole nitrate ear solution
","dexamethasone otic
","baytril + trizulta + dexamethasone ear solution
","neomycin sulfate + nystatin + triamcinolone acetonide
","chloramphenicol + ketoconazole + triamcinolone acetonide ear solution
","triamcinolone","trizedta + dexamethasone + miconazole flush
","chlorhexiderm + dexamethasone spray
","trizedta + baytril + dexamethasone flush
","neomycin sulfate + tetracaine + triamcinolone acetonide
","clotrimazole + gentamicin + mometasone ear solution
","baytril + dexamethasone + silvadene ear solution
","itraconazole + ketoconazole + triamcinolone acetonide ear solution
","baytril + triamcinolone acetonide
","troleandomycin + hydrocortisone
","neo-poly-bac ophthalmic ointment + steroid
","t8 keto flush + dexamethasone
","dexamethasone + miconazole
","dermacool + hydrocortisone
","trizchlor 4 + hydrocortisone
","dexamethasone + diphenhydramine
","tear stimulation forte + betamethasone + ofloxacin
","baytril + dexamethasone + dimethyl sulfoxide + lidocaine
","dexamethasone + baytril
","ketoconazole + mupirocin + triamcinolone acetonide
","penicillin + dexamethasone
","tobramycin + betamethasone ophthalmic ointment
","lotramin + dexamethasone
","bacitracin + hydrocortisone
")
fenbendazole_names<-c("fenbendazole","fenbendazole, maropitant citrate, metronidazole, omeprazole, sucralfate
","cerenia + fenbendazole + metronidazole + omeprazole + sucralfate
","panacur","safe-guard")
```

Lifetime and final five years

```{r}
specific_medication_features <- summarise_specific_medication_flags(
  medications,
  medications_5y,
  list(
    NSAID = NSAID_names,
    steroid = steroid_names,
    fenbendazole = fenbendazole_names
  )
)

meds_5y_lifetime <- meds_wide_quart_5y |>
  dplyr::full_join(
    meds_wide_quart_lifetime,
    by = "subject_id",
    relationship = "one-to-one"
  )

medication_features <- meds_5y_lifetime |>
  dplyr::full_join(
    specific_medication_features,
    by = "subject_id",
    relationship = "one-to-one"
  )

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds <-
  join_medication_features(
    dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs,
    medication_features
  )
```

#### Examine also an NSAID number of times variable

```{r}
NSAID_meds <- c(
  "carprofen",
  "meloxicam",
  "firocoxib",
  "deracoxib",
  "flurbiprofen",
  "ketoprofen"
)

NSAID_5y_features <- summarise_medication_count_feature(
  medications_5y,
  NSAID_meds,
  "NSAID",
  function(count) {
    dplyr::case_when(
      count == 0L ~ "no_NSAID_admin",
      count == 1L ~ "Q1(1)",
      count == 2L ~ "Q2(1-2)",
      count == 3L ~ "Q3(2-3)",
      count > 3L ~ "Q4(3-8)"
    )
  },
  amount_column = "NSAID_5y_quantiles"
)

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID <-
  join_medication_features(
    dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds,
    NSAID_5y_features
  )
```

#### Examine also a steroid number of times variable

```{r}
steroids <- c(
  "prednisone",
  "predisolone acetate",
  "trimeprazine tartrate, prednisone",
  "triamcinolone acetonide",
  "methylpredisolone",
  "hydrocortisone, gentamicin sulfate, miconazole nitrate",
  "burow's solution, hydrocortisone",
  "dexamethasone, miconazole nitrate",
  "dimethyl sulfoxide, fluocinolone acetonide",
  "hydrocortisone",
  "dexamethasone, enrofloxacin, ketoconazole, tris-edta",
  "beclomethasone, clotrimazole, neomycin sulfate",
  "dexamethasone, ketoconazole, tris-edta",
  "acetic acid, boric acid, ceramide complex, hydrocortisone, ketoconazole",
  "betamethasone",
  "mometasone furoate",
  "dexamethasone",
  "betamethasone",
  "hydrocortisone",
  "prednisolone",
  "methylprednisolone",
  "triamcinolone acetonide",
  "difluprednate",
  "fluticasone",
  "dexamethasone, miconazole nitrate",
  "amikacin sulfate, dexamethasone",
  "dexamethasone, methylprednisolone",
  "dexamethasone, triamcinolone acetonide",
  "hydrocortisone, ketoconazole",
  "dexamethasone, ketoconazole, phytosphingosine",
  "dexamethasone, diphenhydramine",
  "betamethasone, clotrimazole, gentamicin sulfate, mometasone furoate",
  "dimethyl sulfoxide, fluocinolone acetonide, gentamicin sulfate, miconazole nitrate",
  "dimethyl sulfoxide, flunixin, fluocinolone acetonide",
  "hydrocortisone, troleandomycin",
  "Dexamethasone",
  "Betamethasone",
  "dexamethasone",
  "betamethasone",
  "hydrocortisone",
  "menthol, hydrocortisone",
  "dexamethasone, ketoconazole",
  "dexamethasone, enrofloxacin",
  "dexamethasone, dimethyl sulfoxide, enrofloxacin, lidocaine",
  "dexamethasone, enrofloxacin, ketoconazole",
  "dexamethasone, penicillin",
  "dexamethasone, tobramycin",
  "betamethasone, ofloxacin",
  "chloramphenicol, dexamethasone",
  "betamethasone, chloramphenicol, ketoconazole",
  "enrofloxacin, dexamethasone, tris-edta",
  "dexamethasone, miconazole nitrate, neomycin sulfate, nystatin, thiostrepton, triamcinolone acetonide",
  "betamethasone, burow's solution, gentamicin sulfate, hydrocortisone, miconazole nitrate",
  "ketoconazole, hydrocortisone",
  "bacitracin, hydrocortisone",
  "dexamethasone, miconazole nitrate, tris-edta",
  "chlorhexidine gluconate, dexamethasone",
  "dexamethasone, enrofloxacin, tris-edta",
  "dexamethasone, enrofloxacin, silver sulfadiazine",
  "dexamethasone, miconazole nitrate, neomycin sulfate, polymyxin b, prednisolone acetate, thiabendazole",
  "betamethasone, enrofloxacin",
  "dexamethasone, ear cleaner (epi-otic advanced), enrofloxacin",
  "betamethasone, tobramycin"
)

```

Use steroids list to do counting now

```{r}
steroid_5y_features <- summarise_medication_count_feature(
  medications_5y,
  steroids,
  "steroid",
  function(count) {
    dplyr::case_when(
      count == 0L ~ "no_steroid_admin",
      count == 1L ~ "Q1(1)",
      count == 2L ~ "Q2(1-2)",
      count == 3L ~ "Q3(2-3)",
      count > 3L ~ "Q4(3-17)"
    )
  },
  amount_column = "steroid_5y_quantiles"
)

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid <-
  join_medication_features(
    dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID,
    steroid_5y_features
  )
```

For MCT, also retain five-year administration counts for systemic
prednisolone/prednisone, lokivetmab, and oclacitinib. There was insufficient
variation to form meaningful quartiles, so positive values remain counts.

```{r}
systemic_preds <- c(
  "prednisone",
  "predisolone acetate",
  "trimeprazine tartrate, prednisone",
  "prednisolone"
)
lokivetmab <- c("lokivetmab", "cytopoint")
oclacitinib <- c("oclacitinib", "apoquel")

count_or_no_administration <- function(label) {
  function(count) {
    ifelse(count == 0L, paste0("no_", label, "_admin"), as.character(count))
  }
}

systemic_pred_features <- summarise_medication_count_feature(
  medications_5y,
  systemic_preds,
  "systemic_preds",
  count_or_no_administration("systemic_preds")
)
lokivetmab_features <- summarise_medication_count_feature(
  medications_5y,
  lokivetmab,
  "lokivetmab",
  count_or_no_administration("lokivetmab")
)
oclacitinib_features <- summarise_medication_count_feature(
  medications_5y,
  oclacitinib,
  "oclacitinib",
  count_or_no_administration("oclacitinib")
)

mct_specific_count_features <- systemic_pred_features |>
  dplyr::full_join(
    lokivetmab_features,
    by = "subject_id",
    relationship = "one-to-one"
  ) |>
  dplyr::full_join(
    oclacitinib_features,
    by = "subject_id",
    relationship = "one-to-one"
  )

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib <-
  join_medication_features(
    dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid,
    mct_specific_count_features
  )
```

Interim write to csv

```{r}
write.csv(dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib,here::here("Output/GRLS_MCT_variables_dataset_interim.csv"))
```

## Reproduction variables

Reproductive features use the annual female and male reproductive-history
tables plus neuter status and dates from `Data/dog_profile.csv`. The heat-detail
table is not used here because `heat_date` has substantially less coverage than
the annual `heat_last_year` indicator.

The first-heat age is therefore the dog's age at the first questionnaire that
reported a heat in the preceding year; it is an approximate reporting age, not
an exact heat date. `heat_reporting_years` counts distinct study years with a
positive `heat_last_year` response and must not be interpreted as the literal
number of heats.

Pregnancy and mating are defined as ever recorded, regardless of neuter status,
because some dogs are never neutered. Pregnancy is not applicable to male dogs,
and mating is not applicable to female dogs. Male mating uses either a positive
annual intact-breeding response or the neutered-ever-bred response.

Neuter age, time between neutering and the cohort endpoint, and age at endpoint
are derived from recorded dates. The workflow stops rather than replacing
negative intervals with zero. In the current data, dogs `094-000441` and
`094-023124` have neuter dates before birth. Ten dogs have neuter dates after
the MCT endpoint; four of these also affect lymphoma: `094-006352`,
`094-000677`, `094-026390`, and `094-027529`. These source/cohort dates must
be resolved before accepting new analysis datasets.

```{r}
source(here::here("R", "reproduction.R"))

repro_hx_female <- read.csv(
  here::here("Data", "reproductive_history_female.csv")
)
repro_hx_male <- read.csv(
  here::here("Data", "reproductive_history_male.csv")
)
profile <- read.csv(here::here("Data", "dog_profile.csv"))

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3 <-
  build_reproduction_features(
    dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib,
    profile,
    repro_hx_female,
    repro_hx_male
  )
```

## Weight and height data 5y prior to endpoint

Physical examination records from the endpoint study year and the five preceding
study years are used. The primary summaries require at least five examination
rows, preserving the established threshold; the `any_records` medians remain
available when fewer records exist. Weight is converted from pounds to kg using
0.453 and height from inches to cm using 2.54. Converted summaries are left
missing for dogs younger than 18 months at the endpoint (six dogs in the legacy
notes). Where BCS values tie for the mode, the maximum tied value is used.

```{r}
source(here::here("R", "weight.R"))

dog_weight <- read.csv(here::here("Data/exam_physical.csv"))
dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_unique <-
  build_weight_features(
    dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3,
    dog_weight
  )
```

```{r}
write.csv(dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_unique,here::here("Output/GRLS_MCT_variables_dataset_interim.csv"))
```

## Activity

This feature set follows the original classification work in
`Code/GRLS activity classifcation method_modified.qmd`. The questionnaire
changed after study year 2, so the two periods retain their established scoring
systems:

- study years 0--2 combine reported walk and aerobic duration, frequency, and
  intensity;
- study year 3 onward scores activity frequency and pace, because duration is
  not collected in the later format;
- the overview questionnaire supplies the whole-period activity assessment.

Every activity record after the dog's endpoint is ignored before any summary is
calculated. Full-date source tables are compared with `end_date`; the
month-resolution overview table is compared with the endpoint month. The
five-year summaries use records from the endpoint study year and the five
preceding study years.

The SY3+ and overview fields are ordinary means of their scored records. This
corrects the legacy calculation that divided those means by record count a
second time. `avg_frequency_5y_prior_endpoint` is calculated from the combined
early and later records in the five-year window; it is not replaced by the
all-SY3 mean.

For ordinal modes (the early recommendation categories and, in the lymphoma
schema, overview activity score), the maximum value is used where modal values
tie. This is the documented ordinal tie rule; nominal modes elsewhere in the
pipeline continue to use `multiple`.

```{r}
source(here::here("R", "activity.R"))

activity_sy1_2 <- read.csv(
  here::here("Data/activity_details_through_SY2.csv")
)
activity_sy3_beyond <- read.csv(
  here::here("Data/activity_details_SY3_beyond.csv")
)
activity_overview <- read.csv(here::here("Data/activity_overview.csv"))

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_unique <-
  dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_unique |>
  dplyr::select(-c(year_in_study.x, year_in_study.y, record_date.y))

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity <-
  build_activity_features(
    dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_unique,
    activity_sy1_2,
    activity_sy3_beyond,
    activity_overview,
    include_overview_distribution = FALSE
  )
```

## Tidying up columns for final df

Dates and endpoint fields retain genuine `NA` values (for example, a dog that
was never withdrawn). Missing values in analysis feature columns are written as
`records_not_available`. The feature selection is name-based so it cannot
silently shift when an upstream column is added or removed.

CSV row names are disabled; spreadsheet-style `X` index columns are not part of
the analysis dataset.

```{r}
source(here::here("R", "final_dataset.R"))

dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity_tidied <-
  finalise_cancer_dataset(
    dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity,
    case_column = "had_MCT"
  )

write.csv(
  dog_pop_deprivation_enviro_smoking_sleep_lifestyle_all_house_location_comorbs_meds_NSAID_steroid_systemic_pred_lokivetmab_oclacitinib3_weight4_activity_endpoint_activity_tidied,
  here::here("Output/GRLS_MCT_variables_dataset.csv"),
  row.names = FALSE
)
```
